### 2013-10-02 paper/paper.tex
Added methods paper

%\documentclass[manuscript]{aastex} %one-column, double-spaced document
\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

\usepackage{hyperref}
%\usepackage{graphicx}
%\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{bm}

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{f_{\rm M}}
\newcommand{\fD}{f_{\rm D}}
\newcommand{\vD}{\vec{D}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}

%\slugcomment{}
%\shorttitle{}
%\shortauthors{}

\begin{document}

\title{A Bayesian method for inference of fundamental stellar parameters}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} and Sean M. Andrews\altaffilmark{1}
%Author2\altaffilmark{2},
}

\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street MS 10, Cambridge, MA 02138}
%\altaffiltext{2}{Institution 2}
\email{iczekala@cfa.harvard.edu}

%\begin{abstract}
%\end{abstract}
%\keywords{}

%\begin{figure}[htb]
%\begin{center}
%\includegraphics{file}
%\caption{}
%\label{fig:}
%\end{center}
%\end{figure}

\section{The model}

To summarize our model, we use Bayes rule to design a posterior probability function
\begin{equation}
  \overbrace{p(\vt | \vD)}^{\rm posterior} \propto \underbrace{p(\vD | \vt)}_{\rm likelihood} \; \overbrace{p(\vt)}^{\rm prior}
  \label{eqn:posterior}
\end{equation}
where the parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$. Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), v \sin i, v_z, A_v \}$ while the data vector $\vD = \fD(\lambda)$ represents the a high resolution spectrum. In our immediate case of our companion paper, this is a 51-order echelle spectrum from TRES at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$) and spanning the full optical range.

Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit. Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis.\footnote{We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).  We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid.} 
\begin{equation}
 p(\vD | \vt) = {\cal L} \propto \exp \left(-\frac{\chi^2}{2} \right)
\end{equation}

\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - k(\lambda | \vN) \fD(\lambda) }{\sigma(\lambda)} \right ]^2
\end{equation}
In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment. To start, we will assume flat priors on $\vstar$ and $\vN$ and then relax this assumption later. Then the logarithm of our posterior probability function is then
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - k(\lambda | \vN) \fD(\lambda) }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob}
\end{equation}
To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating between grid points of a high-resolution PHOENIX model stellar spectra ($R = 500,000$), rotationally broadened, instrumentally broadened, Doppler shifted, extinction corrected, and downsampled to the exact pixels of the TRES spectrum. $\sigma(\lambda)$ is the RMS scatter of the continuum (measured in a line-free region) inversely scaled by the blaze function in each order, to account for Poisson photon counting errors. For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead. The data prefactor $k(\lambda | \vN)$ is a systematic error term that aims to account for errors in the blaze-correction or flux-calibration. In our case, $\vN = \{c_0, c_1, \ldots, c_N\}$ are a set of 4 Chebyshev polynomial coefficients for each order. This amounts to a ``re-fluxing'' following the techniques of \citet{elh+06}. Additionally, we can add priors to limit the degree of re-fluxing (for example 20\%) to be consistent with a systematic noise floor for flux-calibration determined independently through calibration tests on TRES using standard stars.

\subsection{Nuisance parameters}
In order to understand the effects of the $k(\lambda | \vN)$ term, we now consider its effects on just one order. The first four Chebyshev polynomials are 
\begin{align*}
  T_0(x) &= 1\\
  T_1(x) &= x\\
  T_2(x) &= 2 x^2 - 1\\
  T_3(x) &= 4 x^3 - 3x\\
\end{align*}
In our implementation, we map the full range of a specific order $\lambda \in [\lambda_{\rm min}, \lambda_{\rm max}]$ to $x \in [-1, 1]$ (or we can map pixel number). Then, for a given set of Chebyshev coefficients, $\vN = \vc = \{ c_0, c_1, c_2, \ldots, c_N \}$, we have

\begin{align}
  k(\lambda | \vN) &= c_0 T_0(\lambda) + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \\
  k(\lambda | \vN) &= \sum^N_{i = 0} c_i T_i(\lambda)
\end{align}

using this, we can rewrite Equation~\ref{eqn:lnprob} as 
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ] \fD(\lambda) }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob2}
\end{equation}
To simplify the following discussion, we consider just one wavelength $\lambda_i$ and then generalize this to a sum over $\lambda$ later. If we expand out the square in Equation~\ref{eqn:lnprob2}, then for a given $\lambda_i$ we have
\begin{equation}
 \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fM^2 - 2 \fM \fD \sum_n c_n T_n + \fD^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
 \label{eqn:lnprob_lambda}
 \end{equation}
 where $\sigma$, $\fM$, $\fD$, and $T_{n,m}$ are all evaluated at $\lambda_i$

To simplify this math, we can rewrite the Chebyshev coefficients and polynomials as column vectors $\vc$ and $\vec{T}(\lambda)$, respectively
\begin{equation}
  \vc = 
  \begin{bmatrix}
    c_0\\
    c_1\\
    \vdots\\
    c_N
  \end{bmatrix}
  \hspace{3cm}
\vec{T}(\lambda) = 
\begin{bmatrix}
T_0(\lambda)\\
T_1(\lambda)\\
\vdots\\
T_N(\lambda)\\
\end{bmatrix}
\end{equation}

then 
\begin{equation}
  k(\lambda | \vN) = \sum^N_{i = 0} c_i T_i = \vec{T}^\trans \cdot \vc
\end{equation}
If we let 
\begin{equation}
  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
  \begin{bmatrix}
T_0 T_0 & T_0 T_1 &  \hdots & T_0 T_N \\
T_1 T_0 & T_1 T_1 &  \hdots & T_1 T_N \\
\vdots  & \vdots  &  \ddots & \vdots \\
T_N T_0 & T_N T_1 &  \hdots & T_N T_N \\
  \end{bmatrix}
\end{equation}
then we have
\begin{equation}
  \sum_n \sum_m c_n T_n c_m T_m = \vc^\trans \cdot {\bm W} \cdot \vc
\end{equation}
and we can rewrite Equation~\ref{eqn:lnprob_lambda} as 
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fM^2 - 2 \fM \fD\, \vec{T}^\trans \vc \; + \fD^2 \, \vc^\trans {\bm W}  \vc \bigr ]
\end{equation}

Because matrix multiplication is associative, we can sum ${\bm W(\lambda)}$ and $\vec{T}^\trans(\lambda)$ across all $\lambda$, and define
\begin{align}
  {\bm A} &= \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
  \vec{B}^\trans(\vstar) &= \sum_\lambda \frac{\fM(\lambda | \vstar) \fD(\lambda)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
  g(\vstar) &= -\frac{1}{2} \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)}\\
\end{align}
Now we can rewrite Equation~\ref{eqn:lnprob2} as 
\begin{equation}
  \ln \bigl (p(\vstar, \vN | \vD) \bigr) \propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar)
  \label{eqn:lnprob_matrix}
\end{equation}
where remember $\vN = \{c_0, c_1, \ldots, c_N \}$. Now we have
\begin{equation}
  p(\vt | \vD) = p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )
  \label{eqn:posterior_vec}
\end{equation}
Because this is a multi-dimensional Gaussian in the Chebyshev coefficients and the parameters we are most interested in are $\vstar$, we can analytically marginalize out the nuisance parameters
\begin{equation}
  p(\vstar | \vD) = \int p(\vstar, \vN | \vD) \dd \vN
\end{equation}
The analytic multi-dimensional Gaussian integral with linear term \citep{sgd+09} yields 
\begin{align}
  p(\vstar | \vD) &= \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right ) \dd \vc\\
  p(\vstar | \vD) &= \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
\end{align}
This result means that for any given set of parameters $\vstar$, we only need to calculate $\vec{B}(\vstar)$ and $g(\vstar)$; ${\bm A}(\vD)$ can be computed once and stored for the remainder of the MCMC run.

\subsection{Including Gaussian priors on nuisance parameters}
In order to constrain the degree of ``re-fluxing'' to be consistent with an independent determination of the systematic error floor in flux-calibration (for example, 20\%), we can add Gaussian priors that limit the coefficient of each Chebyshev polynomial to within a percentage of no correction. For example, we might wish to limit the multiplicative linear term $c_1$ to within an amplitude of $\pm 0.2$.

This corresponds to a prior on the nuisance parameters of
\begin{equation}
  p(\vN) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
  \label{eqn:nuisance_prior} 
\end{equation}
where 
\begin{equation}
  {\bm D} = 
  \begin{bmatrix}
    \sigma_0^{-2} & 0 & \hdots & 0 \\
    0 & \sigma_1^{-2} & \hdots & 0 \\
    \vdots & \vdots & \ddots & 0 \\
    0 & 0 & \hdots & \sigma_N^{-2} \\
  \end{bmatrix}
\end{equation}
$\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient. We can rewrite Equation~\ref{eqn:posterior} with priors as
\begin{align}
  p(\vstar, \vN | \vD) &\propto p( \vD | \vstar, \vN) p(\vN) p(\vstar)\\
  p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar)
  \label{eqn:posterior_prior}
\end{align}
we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form 
\begin{align}
  -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) &= \frac{1}{2}\left ( -\vc^\trans {\bm D} \vc + \vc^\trans {\bm D} \vec{\mu} + \vec{\mu}^\trans {\bm D} \vc - \vec{\mu}^\trans {\bm D} \vec{\mu} \right )\\
  &= -\frac{1}{2} \vc^\trans {\bm D} \vc + ({\bm D} \vec{\mu})^\trans \vc - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}
\end{align}
then we can rewrite
\begin{align}
  {\bm A}^\prime({\bm D}) &= \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) + {\bm D}\\
  \vec{B}^{\prime\trans}(\vstar | \vec{\mu}, {\bm D}) &= \sum_\lambda \frac{\fM(\lambda | \vstar) \fD(\lambda)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda) + ({\bm D} \vec{\mu})^\trans \\
  g^\prime(\vstar | \vec{\mu}, {\bm D}) &= -\frac{1}{2} \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)}- \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}\\
\end{align}
The result is the posterior probability function for our stellar parameters $\vt$ that has already been marginalized over the nuisance parameters with Gaussian priors.
\begin{equation}
  \boxed{
  p(\vstar | \vD) = \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
}
\end{equation}

\acknowledgments
IC would like to graciously acknowledge Gregory Green, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy.

%\appendix
%
%\section{}
%
%\begin{deluxetable}{ll}
%\tablecaption{\label{table:} Title}
%\tablehead{\colhead{Col1} & \colhead{Col2}}
%\startdata
%\enddata
%\tablecomments{}
%\end{deluxetable}

\bibliography{disks}
\bibliographystyle{hapj}
\end{document}

### 2013-10-02 cheb.py
Added methods paper

@@ -3,14 +3,16 @@ from numpy.polynomial import Chebyshev as Ch
 import numpy as np
 
 def test_chebyshev():
-    coef = np.array([0.,0.,1.,0.01,0.01])
-    coef2 = np.array([0.,0.,1,-1,0])
-    myCh = Ch(coef)
-    myCh2 = Ch(coef2)
+    '''Domain controls the x-range, while window controls the y-range.'''
+    coef = np.array([0.,1.])
+    #coef2 = np.array([0.,0.,1,-1,0])
+    myCh = Ch(coef,window=[-10,10])
+    #Domain c
+    #myCh2 = Ch(coef2)
     #xs = np.linspace(0,3.)
     x0 = np.linspace(-1,1)
-    #plt.plot(x0, myCh(x0))
-    plt.plot(x0, myCh2(x0))
+    plt.plot(x0, myCh(x0))
+    #plt.plot(x0, myCh2(x0))
     #plt.plot(xs, myCh2(xs))
 
     plt.show()

### 2013-10-03 paper/paper.tex
finally understand numpy fft implementation

@@
 This result means that for any given set of parameters $\vstar$, we only need to calculate $\vec{B}(\vstar)$ and $g(\vstar)$; ${\bm A}(\vD)$ can be computed once and stored for the remainder of the MCMC run.
+We have also reduced the dimensionality of our posterior space from $150+$ parameters ($\sim 6$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 6$, which will enable more rapid convergence of the Markov chain.
 
@@
 \end{equation}
+If we increase the permissible degree of re-fluxing by increasing $\sigma_i$, this will have the effect of smearing out or widening the posterior on a given $\vstar$.
 
@@
 IC would like to graciously acknowledge Gregory Green, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy.
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{art+13}.
 

### 2013-10-03 test_ft.py

@@ -1,0 +1,18 @@
+import numpy as np
+from numpy.fft import fft,ifft,fftshift,ifftshift,fftfreq
+import matplotlib.pyplot as plt
+
+__author__ = 'ian'
+
+'''
+Questions to address:
+
+When should you do ifftshift, fftshift
+Different between odd numbered and even arrays?
+When should you use fftfreqs?
+When should you use fft vs ifft?
+When should you use windowing?
+How do you do padding for sinc interpolation?
+
+'''
+

### 2013-10-03 fft_interpolate.py
finally understand numpy fft implementation

@@ -1,14 +1,20 @@
 import numpy as np
-import sys
-from scipy.interpolate import interp1d,UnivariateSpline,griddata
-from scipy.integrate import trapz
+from scipy.interpolate import interp1d,InterpolatedUnivariateSpline,griddata
 import matplotlib.pyplot as plt
 import model as m
 from scipy.special import hyp0f1,struve,j1
-#import PHOENIX_tools as pt
-import gc
+import PHOENIX_tools as pt
 
 c_kms = 2.99792458e5 #km s^-1
+
+f_full = pt.load_flux_full(5900,3.5,True)
+w_full = pt.w_full
+#Truncate to Dave's order
+#ind = (w_full > 5122) & (w_full < 5218)
+ind = (w_full > 3000) & (w_full < 12000.6)
+w_full = w_full[ind]
+f_full = f_full[ind]
+
 
 def calc_lam_grid(v=1.,start=3700.,end=10000):
     '''Returns a grid evenly spaced in velocity'''
@@ -24,55 +30,13 @@ def calc_lam_grid(v=1.,start=3700.,end=10000):
     return lam_grid[np.nonzero(lam_grid)][:-1]
 
 #grid = calc_lam_grid(2.,start=3050.,end=11232.) #chosen to correspond to min U filter and max z filter
-#wave_grid = np.load('wave_grid_2kms.npy')
-wave_grid = calc_lam_grid(0.35, start=3050., end=11232.)
-np.save('wave_grid_0.35kms.npy',wave_grid)
+#wave_grid = calc_lam_grid(0.35, start=3050., end=11232.) #this spacing encapsulates the maximal velocity resolution of the PHOENIX grid, and corresponds to Delta lambda = 0.006 Ang at 5000 Ang.
+
+#np.save('wave_grid_0.35kms.npy',wave_grid)
 #Truncate wave_grid to Dave's order
-#wave_grid = wave_grid[(wave_grid > 5122) & (wave_grid < 5218)]
-
-
-ones = np.ones((10,))
-def downsample(w_m,f_m,w_TRES):
-    out_flux = np.zeros_like(w_TRES)
-    len_mod = len(w_m)
-
-    #Determine the TRES bin edges
-    len_TRES = len(w_TRES)
-    edges = np.empty((len_TRES+1,))
-    difs = np.diff(w_TRES)/2.
-    edges[1:-1] = w_TRES[:-1] + difs
-    edges[0] = w_TRES[0] - difs[0]
[…]

### 2013-10-16 paper/paper.tex
Working marginalization, test to odyssey

@@
 %\end{figure}
+
+\section{Introduction}
+
+Other methods of stellar parameter inference: SPC\citep{blj+12}, MOOG, SME \citep{vp96}.
 
@@
 
+\subsection{Sampling with emcee}
+Citation to DFM \citep{fhl+12}, GW \citep{gw10}, and CosmoHAMMER \citep{asa+13}.
+
+\section{Validation tests}
+Comparison to main sequence stars of HAT-P-9, and WASP-14, from the TRES archive \citep{tfs+12}.
+
+Comparisons using HIRES data?
+
 \acknowledgments

### 2013-10-16 model.py

@@ -393,8 +393,57 @@ def marg_data():
     '''Do the same thing as `data` function before, but now analytically marginalize over nuisance coefficients.'''
     pass
 
+#Make this automatically detected
+xs = np.arange(2299)
+T0 = np.ones_like(xs)
+Ch1 = Ch([0,1], domain=[0,2298])
+T1 = Ch1(xs)
+Ch2 = Ch([0,0,1],domain=[0,2298])
+T2 = Ch2(xs)
+Ch3 = Ch([0,0,0,1],domain=[0,2298])
+T3 = Ch3(xs)
+T = np.array([T0,T1,T2,T3]) #multiply this by the flux and sigma vector for each order
+T = np.array([T0])
+TT = np.einsum("in,jn->ijn",T,T)
+mu = np.array([1,0,0,0])
+sigmac = 0.2
+D = sigmac**(-2) * np.eye(4)
+Dmu = np.einsum("ij,j->j",D,mu)
+muDmu = np.einsum("j,j->",mu,Dmu)
 
 def lnprob(p):
+    '''New lnprob, no nuisance coeffs'''
+    temp, logg, vsini, vz, flux_factor = p
+    if (logg < 0) or (logg > 6.0) or (vsini < 0) or (temp < 2300) or (temp > 10000): #or (Av < 0):
+        return -np.inf
+    else:
+        #shift TRES wavelengths
+        wlsz = wls * np.sqrt((c_kms + vz) / (c_kms - vz))
+        fmods = model(wlsz, temp, logg, vsini, flux_factor)
+
+        a= fmods**2/sigmas**2
+        A = np.einsum("in,jkn->ijk",a,TT)
+        Ap = A + D
+        detA = np.array(list(map(np.linalg.det, Ap)))
+        invA = np.array(list(map(np.linalg.inv, Ap)))
+
+        b = fmods * fls / sigmas**2
+        B = np.einsum("in,jn->ij",b,T)
+        Bp = B + Dmu
+
+        g = -0.5 * fls**2/sigmas**2
+        G = np.einsum("ij->i",g)
+        Gp = G - 0.5 * muDmu
+
+        invAB = np.einsum("ijk,ik->ij",invA,Bp)
+        BAB = np.einsum("ij,ij->i",Bp,invAB)
+
+        lnp = np.sum(0.5 * np.log((2. * np.pi)**norder/detA) + 0.5 * BAB + Gp)
+
+        return lnp
+
+
[…]

### 2013-10-16 plot_MCMC.py

@@ -266,6 +266,68 @@ def staircase_plot_thesis(flatchain):
     #plt.show()
     fig.savefig("plots/staircase.eps")
 
+def staircase_plot_proposal(flatchain):
+    '''flatchain has shape (N, M), where N is the number of samples and M is the number of parameters. Plot only the T_eff vs. log g'''
+
+
+    #Do a 2D histogram, get values
+    #Set a standard deviation, encloses 63%, 97, 99 or whatever.
+
+    margin = 0.1
+    bins = [np.linspace(5800, 6200, num=30), np.linspace(3.20, 3.85, num=30)] #left most edges of bins except [-1], which is right edge
+    margin = 0.05
+
+    fig = plt.figure(figsize=(6.5, 6.5))
+
+    H = np.histogramdd(flatchain[:,0:2],bins=bins)
+    H_mod = H[0] #2D array with hist on bins
+    Hshape = H_mod.shape
+
+    #flatten array
+    H_flat = H_mod.flatten()
+
+    #do argsort, cumsum, find values
+    args = np.argsort(H_flat)[::-1]
+    iargs = np.argsort(args)
+
+    print(args)
+
+    H_sort = H_flat[args] #this array is ranked from highest pixel to lowest pixel
+    stdvs = np.cumsum(H_sort)/np.sum(H_sort) #the array so we go outwards from the most dense point to the least, and normalize
+
+    print(stdvs)
+
+    sig_bounds = np.array([0.6827, 0.9545 , 0.9973])
+
+    #find all values where stdvs < sig_bounds
+    sig_1 = stdvs <= sig_bounds[0] #darkest color
+    sig_2 = (stdvs > sig_bounds[0]) & (stdvs <= sig_bounds[1])
+    sig_3 = (stdvs > sig_bounds[1]) & (stdvs <= sig_bounds[2])
+    sig_4 = (stdvs > sig_bounds[2]) #lightest color or white
+    print(sig_1,sig_2,sig_3,sig_4)
+
+    colors = np.array([(215, 48, 31),(252, 141, 89),(253, 204, 138),(254, 240, 217)])/256. #goes from darkest to lightest
+    #colors = np.array([(43,140,190),(166,189,219),(236,231,242),(256,256,256)])/256.
[…]

### 2013-10-16 MCMC.py
Working marginalization, test to odyssey

@@ -46,8 +46,8 @@ def main():
         sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, threads=config['threads'])
 
     # Choose an initial set of positions for the walkers, randomly distributed across a reasonable range of parameters.
-    temp = np.random.uniform(low=5800, high=6100, size=(nwalkers,))
-    logg = np.random.uniform(low=3.4, high=3.7, size=(nwalkers,))
+    temp = np.random.uniform(low=5800, high=6200, size=(nwalkers,))
+    logg = np.random.uniform(low=3.2, high=4.0, size=(nwalkers,))
     #M = np.random.uniform(low=0.1, high = 10, size=(nwalkers,))
     #R = np.random.uniform(low=0.1, high = 10, size=(nwalkers,))
     vsini = np.random.uniform(low=42, high=45, size=(nwalkers,))
@@ -55,8 +55,8 @@ def main():
     #Av = np.random.uniform(low=1, high = 5, size=(nwalkers,))
     flux_factor = np.random.uniform(low=1.e-27, high=2.e-27, size=(nwalkers,))
     #c0_21 = np.random.uniform(low=0.9, high = 1.1, size=(nwalkers,))
-    c1_21 = np.random.uniform(low=-0.05, high=-0.03, size=(nwalkers,))
-    c2_21 = np.random.uniform(low=-0.1, high=0.1, size=(nwalkers,))
+    #c1_21 = np.random.uniform(low=-0.05, high=-0.03, size=(nwalkers,))
+    #c2_21 = np.random.uniform(low=-0.1, high=0.1, size=(nwalkers,))
     #c0_22 = np.random.uniform(low=0.9, high = 1.1, size=(nwalkers,))
     #c1_22 = np.random.uniform(low=-0.1, high = 0.1, size=(nwalkers,))
     #c2_22 = np.random.uniform(low=-0.1, high = 0.1, size=(nwalkers,))
@@ -64,7 +64,7 @@ def main():
     #c1_23 = np.random.uniform(low=-0.1, high = 0.1, size=(nwalkers,))
     #c2_23 = np.random.uniform(low=-0.1, high = 0.1, size=(nwalkers,))
 
-    p0 = np.array([temp, logg, vsini, vz, flux_factor, c1_21, c2_21]).T#,c0_22,c1_22,c2_22,c0_23,c1_23,c2_23]).T
[…]

### 2013-10-16 cheb.py

@@ -1,6 +1,7 @@
 import matplotlib.pyplot as plt
 from numpy.polynomial import Chebyshev as Ch
 import numpy as np
+import model as m
 
 
 def test_chebyshev():
@@ -37,13 +38,61 @@ T3 = Ch3(xs)
 T = np.array([T0,T1,T2,T3]) #multiply this by the flux and sigma vector for each order
 TT = np.einsum("in,jn->ijn",T,T)
 
-wls = np.load("GWOri_cf_wls.npy")[22]
-fls = np.load("GWOri_cf_fls.npy")[22]
+c = np.array([1,0,0,0])
 
-sigmas = np.load("sigmas.npy")[22] #has shape (51, 2299), a sigma array for each order
 
-all = TT*wls*fls
-A = np.sum(all,axis=-1)
+orders = [21,22]
+
+wls = np.load("GWOri_cf_wls.npy")[orders]
+fls = np.load("GWOri_cf_fls.npy")[orders]
+
+sigmas = np.load("sigmas.npy")[orders] #has shape (51, 2299), a sigma array for each order
+
+fmods = m.model(wls,6001,3.5,42,2.1e-28)
+#fls = m.model(wls,6020,3.6,40,2e-27)
+
+TT = np.einsum("in,jn->ijn",T,T)
+
+mu = np.array([1,0,0,0])
+sigmac = 0.2
+D = sigmac**(-2) * np.eye(4)
+Dmu = np.einsum("ij,j->j",D,mu)
+muDmu = np.einsum("j,j->",mu,Dmu)
+
+a= fmods**2/sigmas**2
+A = np.einsum("in,jkn->ijk",a,TT)
+#add in prior
+A = A + D
+detA = np.array(list(map(np.linalg.det, A)))
+invA = np.array(list(map(np.linalg.inv, A)))
+
+b = fmods * fls / sigmas**2
+B = np.einsum("in,jn->ij",b,T)
+B = B + Dmu
+
+g = -0.5 * fls**2/sigmas**2
+G = np.einsum("ij->i",g)
+G = G - 0.5 * muDmu
+
+#A,B,G are correct
+
+invAB = np.einsum("ijk,ik->ij",invA,B)
+BAB = np.einsum("ij,ij->i",B,invAB)
+
+#these are now correct
+
+lnp = 0.5 * np.log((2. * np.pi)**len(orders)/detA) + 0.5 * BAB + G
[…]

### 2013-10-17 TRES_spectra/FLUX_STANDARDS/cal_plots.py
removed echelle_io and calibrate tools, included LkCa15

@@ -1,74 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-import os, sys
-
-parentdir = "/home/ian/Grad/Research/Disks/StellarSpectra/"
-sys.path.insert(0, parentdir)
-from echelle_io import rechelletxt
-
-from matplotlib.ticker import FormatStrFormatter as FSF
-import asciitable
-
-c_ang = 2.99792458e18 #A s^-1
-
-GW = rechelletxt("../GWOri/11LORI_2012-04-02_02h46m34s_cb.norm.crop.flux.spec")
-
-mod = asciitable.read('fhd93521.dat')
-w = mod['col1']
-f = mod['col2']
-
-
-def make_plots():
-    for i in range(51):
-        fig = plt.figure(figsize=(11, 8))
-        ax = fig.add_subplot(111)
-        wl, fl = GW[i]
-        ind = (w > (wl[0] - 2)) & (w < (wl[-1] + 2))
-        ax.plot(w[ind], f[ind])
-        ax.set_title("Order %s" % (i + 1))
-
-        ax.xaxis.set_major_formatter(FSF("%.0f"))
-        ax.set_xlabel(r"$\lambda\quad[\AA]$")
-        ax.set_ylabel(r"$F_\lambda\quad[\frac{{\rm ergs}}{{\rm cm}^2 /{\rm s}/\AA}]$")
-        fig.savefig("hd93521/{:0>2.0f}.png".format(i + 1))
-
-
-def trim_kurucz():
-    data = asciitable.read("vega_50000.dat")
-    nm = data['col1']
-    fl = data['col2']
-    ind = (nm > 300) & (nm < 950)
-    asciitable.write({"w": 10. * nm[ind], "fl": fl[ind]}, "vega_cut.dat", names=["w", "fl"])
-
-
-def vega_bp():
-    data = asciitable.read("mhr7001.dat")
-    wl = data['col1']
-    ab = data['col2']
-    #ab = -2.5 * np.log10((wl/10.)**2/c_ang * fl) - 13
-    bpass = np.empty((len(wl),))
-    bpass[:-1] = np.diff(wl)
-    bpass[-1] = bpass[-2]
-    asciitable.write({"w": wl, "ab": ab, "bp": bpass}, "hr7001bp.dat", names=["w", "ab", "bp"])
-
-
-def plot_vega():
-    data = asciitable.read("vega_cut.dat")
-    wl = data['w']
[…]

### 2013-10-17 echelle_io.py

@@ -1,143 +1,0 @@
-#import pyfits as pf
-from astropy.io import fits as pf
-import numpy as np
-from astropy.io import ascii
-
-#bname = "GWOri_c"
-#bname = "Vega_2012-04-02_12h50m48s_cb.norm.crop.spec"
-#bname = "Feige34_2012-04-29_03h42m55s_cb.norm.crop.spec"
-bname = "11LORI_2012-04-02_02h46m34s_cb.norm.spec"
-dname = "TRES_spectra/Vega/"
-bname = "11LORI_2012-04-02_02h46m34s_cb.norm.flux.spec"
-#fname = dname + bname + ".fits"
-#fname = "FLAT_2012-04-01_23h18m54s_all_cb.blaze.spec.fits"
-fname = "FLAT_2012-04-01_23h18m54s_all_cb.blaze.spec.crop.fits"
-#fname = "GWOri_nf.fits"
-fname = "GWOri_cf.fits"
-#odir = "GWOri_f"
-odir = "GWOri_cf"
-#odir = "Blaze"
-
-#Number of orders
-#f = pf.open(fname)
-#norder = f[0].data.shape[0]
-#f.close()
-#For TRES data this will always be 51
-norder = 51
-
-
-def wechelletxt():
-    from pyraf import iraf
-
-    for i in range(1, norder + 1):
-        inp = fname + "[*,{:d}]".format(i)
-        out = odir + "/{:0>2d}.txt".format(i)
-        print(inp, out)
-        iraf.wspectext(input=inp, output=out)
-
-
-def rechelletxt(bname):
-    espec = []
-    for i in range(1, norder + 1):
-        inp = bname + "/{:0>2d}.txt".format(i)
-        wl, fl = np.loadtxt(inp, unpack=True)
-        espec.append([wl, fl])
-    print("{:d} orders of echelle read".format(norder))
-    return espec
-
-
-def rechellenp(bname):
-    '''Reads text files similar to rechelletxt, but instead returns a 3D numpy array that takes on the shape (
-    norders, 2, len_wl). For example, GWOri has shape (51,2,2304). Assumes each order is the same length.'''
-    inp = bname + "/01.txt"
-    wl, fl = np.loadtxt(inp, unpack=True)
-    len_wl = len(wl)
-    spec = np.empty((norder, 2, len_wl))
-
-    for i in range(norder):
-        inp = bname + "/{:0>2d}.txt".format(i + 1)
[…]

### 2013-10-17 paper/paper.tex
removed echelle_io and calibrate tools, included LkCa15

@@
 
+\section{Processing TRES Data}
+We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel.
+By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
+\begin{equation}
+  S/N = \sqrt{ {\rm cts}}
+\end{equation}
+then the Noise-to-Signal ratio is the inverse of this
+\begin{equation}
+  N/S = \frac{1}{\sqrt{ {\rm cts}}}
+\end{equation}
+then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.
+
 \acknowledgments

### 2013-10-17 model.py

@@ -1,9 +1,6 @@
 import numpy as np
-from echelle_io import rechellenpflat, load_masks
 from scipy.interpolate import interp1d, LinearNDInterpolator,InterpolatedUnivariateSpline
-from scipy.integrate import trapz
 from scipy.ndimage.filters import convolve
-from scipy.optimize import leastsq, fmin
 from scipy.special import j1
 from deredden import av_points
 from numpy.polynomial import Chebyshev as Ch
@@ -55,13 +52,11 @@ logg_points = np.arange(0.0, 6.1, 0.5)
 T_points = T_points[16:-25]
 logg_points = logg_points[2:-2]
 
-#wls, fls = rechellenpflat("GWOri_cf") #each has shape (51,2299)
-wls = np.load("GWOri_cf_wls.npy")
-fls = np.load("GWOri_cf_fls.npy")
-
-sigmas = np.load("sigmas.npy") #has shape (51, 2299), a sigma array for each order
-
-masks = np.load("masks_array.npy")
+base = 'data/LkCa15//LkCa15_2013-10-13_09h37m31s_cb.flux.spec.'
+wls = np.load(base + "wls.npy")
+fls = np.load(base + "fls.npy")
+sigmas = np.load(base + "sigma.npy")
+masks = np.load(base + "mask.npy")
 
 orders = np.array(config['orders'])
 norder = len(orders)
@@ -72,8 +67,9 @@ fls = fls[orders]
 sigmas = sigmas[orders]
 masks = masks[orders]
 
-wave_grid = np.load("wave_grid_2kms.npy")
+len_wl = len(wls[0])
 
+wave_grid = np.load("wave_grid_2kms.npy")
 
 def load_flux(temp, logg):
     fname = "HiResNpy/PHOENIX-ACES-AGSS-COND-2011/Z-0.0/lte{temp:0>5.0f}-{logg:.2f}-0.0" \
@@ -383,27 +379,15 @@ def data(coefs_arr, wls, fls):
         #flsc[i] = Ch(coefs, domain=[wls[i][0], wls[i][-1]])(wls[i]) * fls[i]
     return flsc
 
-def calc_A(xs,n_c):
-    '''Given a pixel array and number of Chebyshev coefficients, calculate the A matrix'''
-    #For each x in xs,
-    pass
-
-
-def marg_data():
[…]

### 2013-10-17 mpi.py

@@ -1,51 +1,0 @@
-#!/usr/bin/env python
-"""
-Run this example with:
-mpirun -np 2 python examples/mpi.py
-
-"""
-
-from __future__ import print_function
-
-import sys
-import numpy as np
-import emcee
-from emcee.utils import MPIPool
-
-
-def lnprob(x):
-    return -0.5 * np.sum(x ** 2)
-
-# Initialize the MPI-based pool used for parallelization.
-pool = MPIPool(debug=True)
-
-if not pool.is_master():
-    # Wait for instructions from the master process.
-    pool.wait()
-    print("DONE")
-    sys.exit(0)
-
-ndim = 2
-nwalkers = 6
-p0 = [np.random.rand(ndim) for i in range(nwalkers)]
-
-# Initialize the sampler with the chosen specs.
-sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob, pool=pool)
-
-# Run 100 steps as a burn-in.
-pos, prob, state = sampler.run_mcmc(p0, 10)
-
-# Reset the chain to remove the burn-in samples.
-sampler.reset()
-
-# Starting from the final position in the burn-in chain, sample for 1000
-# steps.
-sampler.run_mcmc(pos, 10, rstate0=state)
-
-# Close the processes.
-pool.close()
-
-# Print out the mean acceptance fraction. In general, acceptance_fraction
-# has an entry for each walker so, in this case, it is a 250-dimensional
-# vector.
-print(u"Mean acceptance fraction: ", np.mean(sampler.acceptance_fraction))

### 2013-11-07 paper/paper.tex
Prior tests, lnprob implementation
Firmed up understanding of how priors work and then implemented lnprob
functions to normalize over everything. Fixed bug in metallicity
interpolator (reading wrong point indices), fixed bug in fake dataset
ceration for SNR=50, and halfway through implementing lnprob functions.

@@
 \begin{equation}
-  \chi^2 = \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - k(\lambda | \vN) \fD(\lambda) }{\sigma(\lambda)} \right ]^2
+  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
 \end{equation}
@@
 \begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - k(\lambda | \vN) \fD(\lambda) }{\sigma(\lambda)} \right ]^2
+  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
   \label{eqn:lnprob}
@@
 For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
-The data prefactor $k(\lambda | \vN)$ is a systematic error term that aims to account for errors in the blaze-correction or flux-calibration.
+The prefactor $k(\lambda | \vN)$ is a systematic error term that aims to account for errors in the blaze-correction or flux-calibration.
 In our case, $\vN = \{c_0, c_1, \ldots, c_N\}$ are a set of 4 Chebyshev polynomial coefficients for each order.
-This amounts to a ``re-fluxing'' following the techniques of \citet{elh+06}.
+This amounts to a ``re-fluxing'' of the model following the techniques of \citet{elh+06}.
 Additionally, we can add priors to limit the degree of re-fluxing (for example 20\%) to be consistent with a systematic noise floor for flux-calibration determined independently through calibration tests on TRES using standard stars.
+We scale the model and introduce non-physical warps rather than attempt to re-flux the data since this operation would also need to re-scale the noise, and the $\chi^2$ would be complicated function of $k$.
 
@@
 \begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fM(\lambda | \vstar) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ] \fD(\lambda) }{\sigma(\lambda)} \right ]^2
+  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ]\fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
   \label{eqn:lnprob2}
@@
 \begin{equation}
- \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fM^2 - 2 \fM \fD \sum_n c_n T_n + \fD^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
+ \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM \sum_n c_n T_n + \fM^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
  \label{eqn:lnprob_lambda}
  \end{equation}
- where $\sigma$, $\fM$, $\fD$, and $T_{n,m}$ are all evaluated at $\lambda_i$
+ where $\sigma$, $\fD$, $\fM$, and $T_{n,m}$ are all evaluated at $\lambda_i$
 
@@
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fM^2 - 2 \fM \fD\, \vec{T}^\trans \vc \; + \fD^2 \, \vc^\trans {\bm W}  \vc \bigr ]
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM\, \vec{T}^\trans \vc \; + \fM^2 \, \vc^\trans {\bm W}  \vc \bigr ]
 \end{equation}
@@
 \begin{align}
-  {\bm A} &= \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
-  \vec{B}^\trans(\vstar) &= \sum_\lambda \frac{\fM(\lambda | \vstar) \fD(\lambda)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
-  g(\vstar) &= -\frac{1}{2} \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)}\\
+  {\bm A} &= \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
+  \vec{B}^\trans(\vstar) &= \sum_\lambda \frac{\fD(\lambda) \fM(\lambda | \vstar) }{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
+  g(\vstar) &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}\\
 \end{align}
@@
 \begin{equation}
+  \boxed{
   p(\vt | \vD) = p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )
+}
   \label{eqn:posterior_vec}
 \end{equation}
-Because this is a multi-dimensional Gaussian in the Chebyshev coefficients and the parameters we are most interested in are $\vstar$, we can analytically marginalize out the nuisance parameters
+This is one of the \texttt{lnprob} options to sample in, provided you are interested in the results of the Chebyshev coefficients.
+However, because this is a multi-dimensional Gaussian in the Chebyshev coefficients and the parameters we are most interested in are $\vstar$, we can analytically marginalize out the nuisance parameters
 \begin{equation}
@@
 \begin{align}
-  {\bm A}^\prime({\bm D}) &= \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) + {\bm D}\\
-  \vec{B}^{\prime\trans}(\vstar | \vec{\mu}, {\bm D}) &= \sum_\lambda \frac{\fM(\lambda | \vstar) \fD(\lambda)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda) + ({\bm D} \vec{\mu})^\trans \\
-  g^\prime(\vstar | \vec{\mu}, {\bm D}) &= -\frac{1}{2} \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)}- \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}\\
+  {\bm A}^\prime({\bm D}) &= \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)} {\bm W}(\lambda) + {\bm D}\\
+  \vec{B}^{\prime\trans}(\vstar | \vec{\mu}, {\bm D}) &= \sum_\lambda \frac{\fD(\lambda)\fM(\lambda | \vstar)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda) + ({\bm D} \vec{\mu})^\trans \\
+  g^\prime(\vstar | \vec{\mu}, {\bm D}) &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}- \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}\\
 \end{align}
@@
 If we increase the permissible degree of re-fluxing by increasing $\sigma_i$, this will have the effect of smearing out or widening the posterior on a given $\vstar$.
+
+\subsection{Sampling a un-flux calibrated model}
+The previously derived formalism works well in the case where the correction to $c_0$ is small in each order and can be approximated by a Gaussian.
+In reality, the correct formulation of $k(\lambda | \vN)$ to use is
+\begin{equation}
+  k(\lambda | \vN) = c_0 T_0(\lambda) \left [1 + {c_1}^\prime T_1(\lambda) + \ldots + {c_N}^\prime T_N(\lambda) \right ]
+  \label{eqn:k_new}
+\end{equation}
+When applied to scale the model flux $\fM$, this becomes
+\begin{equation}
+  k(\lambda | \vN) \fM = c_0 \fM + {c_1}^\prime T_1 (c_0 \fM) + \ldots + {c_N}^\prime T_N (c_0 \fM)
+\end{equation}
+Now, $c_0$ is allowed to scale by large amounts while the higher order terms (denoted by $c_n$) are perturbations on top of the \emph{scaled} function.
+Previously, the higher order terms were only perturbations on the original, unscaled function, and if the $c_0$ term was large, then this gave an incorrect scaling.
+The correct priors to have on these small perturbations are Gaussian centered around 0
+\begin{equation}
+  p({c_n}^\prime) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^{\prime 2}}{2 \sigma_{ {c_n}^\prime}^2} \right)
+\end{equation}
+Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two. This means we should be using a log-normal prior on $c_0$
+\begin{equation}
+  p(c_0) = \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) 
+\end{equation}
+instead of a Gaussian prior.
+
+To gain insight, we can re-express this new version of $k$ in the original form as 
+\begin{equation}
+   k(\lambda | \vN) = c_0 T_0 + c_1 T_1 + \ldots + c_N T_N
+\end{equation}
+where $c_1 = c_0 {c_1}^\prime T_0$ and $c_N = c_0 {c_N}^\prime T_0$.
+Thus, if there needs to be a large perturbation in $c_0$, then there must also be a large perturbation in $c_n$.
+The prior on $c_1$ is now a product of a log-normal prior on $c_0$ and a Gaussian prior on ${c_1}^\prime$, which is not analytically tractable.
+Therefore, in the case of large perturbations, it is better to use Equation~\ref{eqn:k_new} and use a log-normal prior on $c_0$.
+When we use the original framework, we are saying that perturbations in $c_0$ are small enough that we can approximate the log-normal prior on $c_0$ as Gaussian and that the priors on $c_n$ can also be correctly approximated by a Gaussian.
+From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$.
+Beyond this, one should switch to this framework, for example to correct for un-flux calibrated data.
+
+From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (Figure?) flux-calibration function, or IRAF's ``sensitivity function.''   However, tests show that this effect is in fact small.
+
+This invalidates the previous section on normalization. However, for corrections of less than XX\%, this approximation is better than XX\% accurate.
+
+For the flux-calibration exercise, the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$.
+
+Unfortunately, using this more correct formalism means that we can no longer analytically marginalize over $c_0$, we must sample in this parameter.
+We can, however, still analytically marginalize over the ${c_n}^\prime$ as before.
+
+\begin{align}
+  {\bm A}_{ij} &= \frac{c_0^2 \fM^2}{\sigma^2} W_{ij}\\
+  B_i &= \frac{-\fM^2 c_0^2 + \fD \fM c_0}{\sigma^2} T_i\\
+  g & = -\frac{1}{2 \sigma^2} (\fM^2 c_0^2 - 2 \fD \fM c_0 + \fD^2)
+\end{align}
+and then 
+\begin{equation}
+  {\bm A}_{ij}^{-1} = \frac{\sigma^2}{\fM^2 c_0^2} W^{-1}_{ij}
+\end{equation}
+then, following the same integration as before, we have
+\begin{equation}
+  p(\vstar,c_0 | \vD) \propto \frac{1}{c_0^N} \sqrt{ \frac{(2 \pi)^N}{ {\rm det} |A_{\rm old}|}} \exp \left( -\frac{\fM^2 c_0^2}{\sigma^2} + \frac{\fM \fD}{\sigma^2} (T_i W^{-1}_{ij} T_j + 1 ) c_0 - \frac{\fD^2}{2 \sigma^2}  (T_i W^{-1}_{ij} T_j + 1 )\right)
+\end{equation}
+and then we can put our log-normal prior on $c_0$.
+Where $N$ is the number of higher order terms, in our case $N=3$.
+This function is not able to be integrated analytically, and so two-step marginalization is not feasible.
+
+The plot in plots/priortests shows that the log-normal prior is the correct choice.
+Since we care most about the inference on F, we want to know that the choice on prior of $c_0$ is not biasing the final outcome, while a Gaussian prior might if $\sigma_{c_0}$ was large.
+How to quantify this error and create a dividing line?
+For 5\% error?
+In peak?
+Do these plots need to be in the paper?
+
+Compare distributions by doing a K-S test on the cumulative distributions?
+
+\subsection{Hierarchical modelling}
+Can have a $c_1$, etc, that are marginalized over but we sample $c_0$ in a hierarchical manner which informs the overall flux-level.
 

### 2013-11-07 cheb.py
Prior tests, lnprob implementation
Firmed up understanding of how priors work and then implemented lnprob
functions to normalize over everything. Fixed bug in metallicity
interpolator (reading wrong point indices), fixed bug in fake dataset
ceration for SNR=50, and halfway through implementing lnprob functions.

@@ -43,12 +43,12 @@ c = np.array([1,0,0,0])
 
 orders = [21,22]
 
-wls = np.load("GWOri_cf_wls.npy")[orders]
-fls = np.load("GWOri_cf_fls.npy")[orders]
+wls = m.wls
+fls = m.fls
 
-sigmas = np.load("sigmas.npy")[orders] #has shape (51, 2299), a sigma array for each order
+sigmas = m.sigmas #has shape (51, 2299), a sigma array for each order
 
-fmods = m.model(wls,6001,3.5,42,2.1e-28)
+fmods = m.model(wls, 6001,3.5,0.0, 3, 0.0, 1e-10)
 #fls = m.model(wls,6020,3.6,40,2e-27)
 
 TT = np.einsum("in,jn->ijn",T,T)
@@ -91,7 +91,7 @@ Ac = np.einsum("ijk,k->ij",A,c)
 cAc = np.einsum("j,ij->i",c,Ac)
 Bc = np.einsum("ij,j->i",B,c)
 
-#to obtain the original, unnormalized P given c
+#to obtain the original, unnormalized, unmarginalized P given c
 print("Unmarginalized",np.sum(-0.5 * cAc + Bc + G))
 
 #plt.plot(xs,T0)

### 2013-11-07 model.py

@@ -51,19 +51,26 @@ Z_points = np.array([-0.5, 0.0, 0.5])
 
 #Limit grid size to relevant region
 grid_params = config['grid_params']
+
 T_low, T_high = grid_params['temp_range']
-T_points = T_points[(T_points > T_low) & (T_points < T_high)]
+T_ind = (T_points > T_low) & (T_points < T_high)
+T_points = T_points[T_ind]
+T_arg = np.where(T_ind)[0]
+
 g_low, g_high = grid_params['logg_range']
-logg_points = logg_points[(logg_points > g_low) & (logg_points < g_high)]
+logg_ind = (logg_points > g_low) & (logg_points < g_high)
+logg_points = logg_points[logg_ind]
+logg_arg = np.where(logg_ind)[0]
+
 Z_low, Z_high = grid_params['Z_range']
-Z_points = Z_points[(Z_points > Z_low) & (Z_points < Z_high)]
-print("Limited PHOENIX grid to temp: ", T_points, " logg: ", logg_points, " Z: ", Z_points)
+#Z_points = Z_points[(Z_points > Z_low) & (Z_points < Z_high)]
+print("Limiting PHOENIX grid to temp: ", T_points, " logg: ", logg_points, " Z: ", Z_points)
 
 base = 'data/' + config['dataset']
 wls = np.load(base + ".wls.npy")
 fls = np.load(base + ".fls.npy")
 #fls = np.load("fls_fake.npy")
-sigmas = 10 * np.load(base + ".sigma.npy") #3.8 gives chi^2_red = 1
+sigmas = np.load(base + ".sigma.npy") #3.8 gives chi^2_red = 1
 #sigmas = np.load('sigmas_fake.npy')
 masks = np.load(base + ".mask.npy")
 
@@ -95,13 +102,15 @@ def flux_interpolator_hdf5():
     LIB = fhdf5['LIB']
     param_combos = []
     var_combos = []
-    for t, temp in enumerate(T_points):
-        for l, logg in enumerate(logg_points):
+    for ti in range(len(T_points)):
+        for li in range(len(logg_points)):
             for z, Z in enumerate(Z_points):
-                param_combos.append([t, l, z])
-                var_combos.append([temp, logg, Z])
[…]

### 2013-11-07 run_all.py

@@ -1,13 +1,0 @@
-#!/usr/bin/env python3
-
-'''This script calls all the necessary files for a 'run' on Odyssey.'''
-
-__author__ = 'ian'
-
-#Might need to move mpi.master() call that is currently in MCMC.py here, so that we aren't copying things 64 times
-
-
-
-#Call MCMC.py script
-
-#When

### 2013-11-19 paper/paper.tex
Added model_p, working visualize script

@@
 \medskip
-Ian~Czekala\altaffilmark{1} and Sean M. Andrews\altaffilmark{1}
+Ian~Czekala\altaffilmark{1} et al.
 %Author2\altaffilmark{2},
@@
 
-Other methods of stellar parameter inference: SPC\citep{blj+12}, MOOG, SME \citep{vp96}.
+The fundamental parameters of stars, such as mass, luminosity, effective photospheric temperature, surface gravity, and metallicity are of supreme interest in astrophysics, for a wide variety of reasons.
+Exoplanet properties are found relative to the properties of the host star, and have important implications for models of planet formation.
+The properties are also important to the study of disks, for the incident radiation field sets photoevaporation, influences disk dispersal.
+Many methods exist for determination of stellar parameters: SPC \citep{blj+12}, MOOG, SME \citep{vp96}, and these are robust for FKG stars, where there is a well-defined continuum and the stars are ``normal.'' However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.
 
@@
 where the parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$.
-Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), v \sin i, v_z, A_v \}$ while the data vector $\vD = \fD(\lambda)$ represents the a high resolution spectrum.
-In our immediate case of our companion paper, this is a 51-order echelle spectrum from TRES at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$) and spanning the full optical range.
+Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum.
+For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.
 
-Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit.
-Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis.\footnote{We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).
-We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid.}
+Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}.
+Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis.
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 \end{equation}
-In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment.
-To start, we will assume flat priors on $\vstar$ and $\vN$ and then relax this assumption later.
-Then the logarithm of our posterior probability function is then
+With flat priors on $\vstar$ and $\vN$ (an assumption we will later relax) the logarithm of our posterior probability function is then
 \begin{equation}
@@
 \end{equation}
-To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating between grid points of a high-resolution PHOENIX model stellar spectra ($R = 500,000$), rotationally broadened, instrumentally broadened, Doppler shifted, extinction corrected, and downsampled to the exact pixels of the TRES spectrum.
-$\sigma(\lambda)$ is the RMS scatter of the continuum (measured in a line-free region) inversely scaled by the blaze function in each order, to account for Poisson photon counting errors.
+To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating in $T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}]$ space between grid points of instrumentally broadened, oversampled PHOENIX model stellar spectra, rotationally broadened using Fourier techniques, Doppler shifted, extinction corrected, and downsampled using spline interpolation to the exact pixels of the TRES spectrum.
+
+$\sigma(\lambda)$ is the Poisson photon counting errors.
 For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
-The prefactor $k(\lambda | \vN)$ is a systematic error term that aims to account for errors in the blaze-correction or flux-calibration.
+For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.
+
+The prefactor $k(\lambda | \vN)$ is a calibration functionm that aims to account for errors in the blaze-correction or flux-calibration by multiplying the model by a polynomial.
+The parameters describing this function are the coefficients of the polynomial.
+This approach of accounting for large-scale calibration errors in the spectrum is more justified than the traditional approach of normalizing to the stellar continuum.
+From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (Figure?) flux-calibration function, or IRAF's ``sensitivity function.'' These tests revealed that the largest source of error in flux-calibration was actually in the overall magnitude of individual order, this could be due to slit losses, poor correction due to parallactic angle, or scattered light within the instrument.
+These corrections are smaller than 10\%.
+But more importantly, the corrections in the higher order terms (ie, the shape of the bandpass calibration for each order) are small ($\lesssim 3\%$).
+As described in \S\ref{sec:priors}, we can set priors on these coefficients.
+This will be the optimal mode of operation for our model.
+However, there exists a wealth of archival data that is un-fluxcalibrated, since its primary purpose was for radial velocity determination of FGK stars, to which it is easy to normalize to a continuum.
+In this case, the scatter amongst the calibration can be as large as a factor of 2x.
+This prior information on the calibration parameters motivates our choice of functional parameterization for the nuisance parameters in \S\ref{sec:nuisance}.
+
+\subsection{Nuisance parameters}
+\label{sec:nuisance}
 In our case, $\vN = \{c_0, c_1, \ldots, c_N\}$ are a set of 4 Chebyshev polynomial coefficients for each order.
 This amounts to a ``re-fluxing'' of the model following the techniques of \citet{elh+06}.
-Additionally, we can add priors to limit the degree of re-fluxing (for example 20\%) to be consistent with a systematic noise floor for flux-calibration determined independently through calibration tests on TRES using standard stars.
-We scale the model and introduce non-physical warps rather than attempt to re-flux the data since this operation would also need to re-scale the noise, and the $\chi^2$ would be complicated function of $k$.
+We apply the re-fluxing to the model rather than attempt to re-flux the data since this operation would also need to re-scale the noise, making $\chi^2$ a complicated function of $k$.
 
-\subsection{Nuisance parameters}
-In order to understand the effects of the $k(\lambda | \vN)$ term, we now consider its effects on just one order. The first four Chebyshev polynomials are 
+In order to understand the effects of the $k(\lambda | \vN)$ term, we now consider its effects on just one order of the 51-order echelle spectrum. The first four Chebyshev polynomials are 
 \begin{align*}
@@
 \end{align*}
-In our implementation, we map the full range of a specific order $\lambda \in [\lambda_{\rm min}, \lambda_{\rm max}]$ to $x \in [-1, 1]$ (or we can map pixel number).
-Then, for a given set of Chebyshev coefficients, $\vN = \vc = \{ c_0, c_1, c_2, \ldots, c_N \}$, we have
-
+In our implementation, we map the full pixel range of a specific order $\lambda \in [\lambda_{\rm min}, \lambda_{\rm max}]$ to $x \in [-1, 1]$.
+By mapping pixel number in a given order, rather than wavelength, we are insensitive to wavelength shifts of the model.
+Then, for a given set of Chebyshev coefficients, $\vN = \{ c_0, c_1, c_2, \ldots, c_N \}$, we have
 \begin{align}
-  k(\lambda | \vN) &= c_0 T_0(\lambda) + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \\
-  k(\lambda | \vN) &= \sum^N_{i = 0} c_i T_i(\lambda)
+  k(\lambda | \vN) &= c_0 T_0(\lambda) \left [1 + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \right ]\\
+  k(\lambda | \vN) &= c_0 \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ]
+  \label{eqn:k_new}
+\end{align}
+When applied to scale the model flux $\fM$, this becomes
+\begin{align}
+  k(\lambda | \vN) \fM &=  \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ] (c_0 \fM)\\
+  k(\lambda | \vN) \fM &= c_0 \fM + c_1 T_1 (c_0 \fM) + \ldots + c_N T_N (c_0 \fM)
 \end{align}
 
-using this, we can rewrite Equation~\ref{eqn:lnprob} as 
+This functional form of $k$ allows $c_0$ to scale by large amounts while the higher order terms (denoted by $c_n$) are perturbations on top of the \emph{scaled} model ($c_0 \fM$).
+However, see \S\ref{sec:gaussian_simplification} for an alternative model when perturbations are small.
+We can rewrite Equation~\ref{eqn:lnprob} as
 \begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ]\fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
+  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [1 + \sum^N_{i = 1} c_i T_i(\lambda) \bigr ] c_0 \fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
   \label{eqn:lnprob2}
 \end{equation}
-To simplify the following discussion, we consider just one wavelength $\lambda_i$ and then generalize this to a sum over $\lambda$ later.
+To simplify the following discussion, we consider a single wavelength $\lambda_i$ and then generalize this to a sum over $\lambda$ later.
 If we expand out the square in Equation~\ref{eqn:lnprob2}, then for a given $\lambda_i$ we have
 \begin{equation}
- \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM \sum_n c_n T_n + \fM^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + \sum_{i=1}^N c_i T_i \right ]^2 \right ]
  \label{eqn:lnprob_lambda}
  \end{equation}
- where $\sigma$, $\fD$, $\fM$, and $T_{n,m}$ are all evaluated at $\lambda_i$
+where $\sigma$, $\fD$, $\fM$, and $T_{n,m}$ are all evaluated at $\lambda_i$. We can further expand this to
+\begin{equation}
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
+  \label{eqn:expanded}
+ \end{equation}
 
@@
   \begin{bmatrix}
-    c_0\\
     c_1\\
+    c_2\\
     \vdots\\
@@
 \begin{bmatrix}
-T_0(\lambda)\\
 T_1(\lambda)\\
+T_2(\lambda)\\
 \vdots\\
@@
 \end{equation}
-
 then 
 \begin{equation}
-  k(\lambda | \vN) = \sum^N_{i = 0} c_i T_i = \vec{T}^\trans \cdot \vc
+  k(\lambda | \vN) = c_0 \left [ 1 + \sum^N_{i = 0} c_i T_i \right ] = c_0 \left[ 1+ \vec{T}^\trans \cdot \vc \right]
 \end{equation}
@@
   \begin{bmatrix}
-T_0 T_0 & T_0 T_1 &  \hdots & T_0 T_N \\
-T_1 T_0 & T_1 T_1 &  \hdots & T_1 T_N \\
+T_1 T_1 & T_1 T_2 &  \hdots & T_1 T_N \\
+T_2 T_1 & T_2 T_2 &  \hdots & T_2 T_N \\
 \vdots  & \vdots  &  \ddots & \vdots \\
-T_N T_0 & T_N T_1 &  \hdots & T_N T_N \\
+T_N T_1 & T_N T_2 &  \hdots & T_N T_N \\
   \end{bmatrix}
@@
 \begin{equation}
-  \sum_n \sum_m c_n T_n c_m T_m = \vc^\trans \cdot {\bm W} \cdot \vc
+  \sum_{i =1}^N \sum_{j=1}^N c_i T_i c_j T_j = \vc^\trans \cdot {\bm W} \cdot \vc
 \end{equation}
-and we can rewrite Equation~\ref{eqn:lnprob_lambda} as 
+We can rewrite Equation~\ref{eqn:expanded} quadratically in $\vc$ as
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM\, \vec{T}^\trans \vc \; + \fM^2 \, \vc^\trans {\bm W}  \vc \bigr ]
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
+ \end{equation}
+\begin{equation}
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2} \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \sum_{i=1}^N c_i T_i  -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
+\end{equation}
+\begin{equation}
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2}  \vc^\trans {\bm W} \vc + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \vec{T}^\trans \vc -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
 \end{equation}
@@
 \begin{align}
-  {\bm A} &= \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
-  \vec{B}^\trans(\vstar) &= \sum_\lambda \frac{\fD(\lambda) \fM(\lambda | \vstar) }{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
-  g(\vstar) &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}\\
+  {\bm A} &= c_0^2 \sum_{\lambda} \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) \\
+  \vec{B} &= \sum_{\lambda} \frac{-\fM^2(\lambda) c_0^2 + \fD(\lambda) \fM (\lambda) c_0}{\sigma^2(\lambda)} \vec{T}(\lambda)\\
+  g &= -\frac{1}{2} \sum_{\lambda} \frac{c_0^2 \fM^2(\lambda) - 2 c_0 \fD(\lambda) \fM(\lambda) + \fD^2(\lambda)}{\sigma^2(\lambda)} 
 \end{align}
-Now we can rewrite Equation~\ref{eqn:lnprob2} as 
+Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
+\begin{align}
+  \ln \bigl [p(\vstar, \vN | \vD) \bigr] &\propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g\\
+  p(\vt | \vD) = p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
+  \label{eqn:lnprob_matrix}
+\end{align}
+Here we have the full posterior probability distribution defined in Equation~\ref{eqn:posterior}, assuming flat priors.
+This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$.
+Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters.
+This marginalization encapsulates the uncertainty due to the inference on the nuisance parameters.
+Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically, and then sample from the marginalized distribution.
 \begin{equation}
-  \ln \bigl (p(\vstar, \vN | \vD) \bigr) \propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar)
-  \label{eqn:lnprob_matrix}
-\end{equation}
-where remember $\vN = \{c_0, c_1, \ldots, c_N \}$. Now we have
-\begin{equation}
-  \boxed{
-  p(\vt | \vD) = p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )
-}
-  \label{eqn:posterior_vec}
-\end{equation}
-This is one of the \texttt{lnprob} options to sample in, provided you are interested in the results of the Chebyshev coefficients.
-However, because this is a multi-dimensional Gaussian in the Chebyshev coefficients and the parameters we are most interested in are $\vstar$, we can analytically marginalize out the nuisance parameters
-\begin{equation}
-  p(\vstar | \vD) = \int p(\vstar, \vN | \vD) \dd \vN
+  p(\vstar, c_0 | \vD) = \int p(\vstar, c_0, \vc\; | \vD) \dd \vc
 \end{equation}
@@
 \begin{align}
-  p(\vstar | \vD) &= \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right ) \dd \vc\\
-  p(\vstar | \vD) &= \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
+  p(\vstar, c_0 | \vD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
+  p(\vstar, c_0 | \vD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
 \end{align}
-This result means that for any given set of parameters $\vstar$, we only need to calculate $\vec{B}(\vstar)$ and $g(\vstar)$; ${\bm A}(\vD)$ can be computed once and stored for the remainder of the MCMC run.
-We have also reduced the dimensionality of our posterior space from $150+$ parameters ($\sim 6$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 6$, which will enable more rapid convergence of the Markov chain.
+Where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$.
+By doing this marginalization we have reduced the dimensionality of our posterior space from $200+$ parameters ($\sim 7$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 60$ ($\sim 7$ stellar parameters and one $c_0$ for each of the 51 echelle orders).
+This enables a dramatic speedup in the MCMC sampling.
+For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than XX\%), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters.
+This approach is detailed in \S\ref{sec:gaussian_simplification}.
 
 \subsection{Including Gaussian priors on nuisance parameters}
-In order to constrain the degree of ``re-fluxing'' to be consistent with an independent determination of the systematic error floor in flux-calibration (for example, 20\%), we can add Gaussian priors that limit the coefficient of each Chebyshev polynomial to within a percentage of no correction.
-For example, we might wish to limit the multiplicative linear term $c_1$ to within an amplitude of $\pm 0.2$.
-
-This corresponds to a prior on the nuisance parameters of
+\label{sec:priors}
+Using this formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters.
+From our flux-calibration experiment, we determined that the overall scaling could range by 10\%, while the higher order corrections were less than 3\%.
+Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two.
+This means we should be using a log-normal prior on $c_0$
+\begin{align}
+  p(c_0) &= \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) \\
+  \ln p(c_0) &= \ln \left( \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \right) - \frac{(\ln c_0)^2}{2 \sigma_{c_0}^2}
+\end{align}
+centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 f$), we assume Gaussian priors centered about 0
 \begin{equation}
-  p(\vN) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
+  p(c_n) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^2}{2 \sigma_{c_n^2}} \right)
+\end{equation}
+which we can express in vector form as
+\begin{equation}
+  p(\vc\;) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
   \label{eqn:nuisance_prior} 
@@
   \begin{bmatrix}
-    \sigma_0^{-2} & 0 & \hdots & 0 \\
-    0 & \sigma_1^{-2} & \hdots & 0 \\
+    \sigma_1^{-2} & 0 & \hdots & 0 \\
+    0 & \sigma_2^{-2} & \hdots & 0 \\
     \vdots & \vdots & \ddots & 0 \\
@@
 \end{equation}
-$\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient. We can rewrite Equation~\ref{eqn:posterior} with priors as
-\begin{align}
-  p(\vstar, \vN | \vD) &\propto p( \vD | \vstar, \vN) p(\vN) p(\vstar)\\
-  p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar)
+and $\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient, in our case $\sigma_i = \sigma_{c_n} \approx 0.03$.
+$\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction.
+Now, we can write Equation~\ref{eqn:posterior} with priors as
+\begin{equation}
+  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\; | \vD) \propto p( \vD | \vstar, c_0, \vc\; ) p(\vc\;) p(\vstar) p(c_0) 
+\end{equation}
+\begin{equation}
+  p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
   \label{eqn:posterior_prior}
-\end{align}
-we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form 
+\end{equation}
+we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$
 \begin{align}
@@
 \begin{align}
-  {\bm A}^\prime({\bm D}) &= \sum_\lambda \frac{\fM^2(\lambda|\vstar)}{\sigma^2(\lambda)} {\bm W}(\lambda) + {\bm D}\\
-  \vec{B}^{\prime\trans}(\vstar | \vec{\mu}, {\bm D}) &= \sum_\lambda \frac{\fD(\lambda)\fM(\lambda | \vstar)}{\sigma^2(\lambda)} \vec{T}^\trans(\lambda) + ({\bm D} \vec{\mu})^\trans \\
-  g^\prime(\vstar | \vec{\mu}, {\bm D}) &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}- \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}\\
+  {\bm A}^\prime &= {\bm A} + {\bm D}\\
+  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
+  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
 \end{align}
-The result is the posterior probability function for our stellar parameters $\vt$ that has already been marginalized over the nuisance parameters with Gaussian priors.
+in the case that $\vec{\mu} = \{0, 0, \ldots, 0\}$, there is only a non-zero correction to ${\bm A}$. The full posterior probability distribution is then
  \begin{equation}
    \boxed{
-  p(\vstar | \vD) = \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
+  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
 }
 \end{equation}
+This probability function appears in the code as \texttt{lnprob\_lognormal} and has $\sim 7$ stellar dimensions and $\sim 200$ nuisance dimensions.
+
+Using the previous multi-dimensional integral, we can marginalize over the $\vc$ including the Gaussian priors
+\begin{equation}
+  \boxed{
+  p(\vstar, c_0 | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) p(c_0)
+}
+\end{equation}
+where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$.
+This probability function appears in the code as \texttt{lnprob\_lognormal\_marg} and has $\sim 7$ stellar dimensions and $\sim 50$ nuisance dimensions.
 If we increase the permissible degree of re-fluxing by increasing $\sigma_i$, this will have the effect of smearing out or widening the posterior on a given $\vstar$.
@@
 \subsection{Sampling a un-flux calibrated model}
-The previously derived formalism works well in the case where the correction to $c_0$ is small in each order and can be approximated by a Gaussian.
-In reality, the correct formulation of $k(\lambda | \vN)$ to use is
-\begin{equation}
-  k(\lambda | \vN) = c_0 T_0(\lambda) \left [1 + {c_1}^\prime T_1(\lambda) + \ldots + {c_N}^\prime T_N(\lambda) \right ]
-  \label{eqn:k_new}
-\end{equation}
-When applied to scale the model flux $\fM$, this becomes
-\begin{equation}
-  k(\lambda | \vN) \fM = c_0 \fM + {c_1}^\prime T_1 (c_0 \fM) + \ldots + {c_N}^\prime T_N (c_0 \fM)
-\end{equation}
-Now, $c_0$ is allowed to scale by large amounts while the higher order terms (denoted by $c_n$) are perturbations on top of the \emph{scaled} function.
-Previously, the higher order terms were only perturbations on the original, unscaled function, and if the $c_0$ term was large, then this gave an incorrect scaling.
-The correct priors to have on these small perturbations are Gaussian centered around 0
-\begin{equation}
-  p({c_n}^\prime) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^{\prime 2}}{2 \sigma_{ {c_n}^\prime}^2} \right)
-\end{equation}
-Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two. This means we should be using a log-normal prior on $c_0$
-\begin{equation}
-  p(c_0) = \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) 
-\end{equation}
-instead of a Gaussian prior.
-
-To gain insight, we can re-express this new version of $k$ in the original form as 
-\begin{equation}
-   k(\lambda | \vN) = c_0 T_0 + c_1 T_1 + \ldots + c_N T_N
-\end{equation}
-where $c_1 = c_0 {c_1}^\prime T_0$ and $c_N = c_0 {c_N}^\prime T_0$.
-Thus, if there needs to be a large perturbation in $c_0$, then there must also be a large perturbation in $c_n$.
-The prior on $c_1$ is now a product of a log-normal prior on $c_0$ and a Gaussian prior on ${c_1}^\prime$, which is not analytically tractable.
-Therefore, in the case of large perturbations, it is better to use Equation~\ref{eqn:k_new} and use a log-normal prior on $c_0$.
-When we use the original framework, we are saying that perturbations in $c_0$ are small enough that we can approximate the log-normal prior on $c_0$ as Gaussian and that the priors on $c_n$ can also be correctly approximated by a Gaussian.
-From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$.
-Beyond this, one should switch to this framework, for example to correct for un-flux calibrated data.
-
-From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (Figure?) flux-calibration function, or IRAF's ``sensitivity function.''   However, tests show that this effect is in fact small.
-
 This invalidates the previous section on normalization. However, for corrections of less than XX\%, this approximation is better than XX\% accurate.
@@
 For the flux-calibration exercise, the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$.
-
-Unfortunately, using this more correct formalism means that we can no longer analytically marginalize over $c_0$, we must sample in this parameter.
-We can, however, still analytically marginalize over the ${c_n}^\prime$ as before.
-
-\begin{align}
-  {\bm A}_{ij} &= \frac{c_0^2 \fM^2}{\sigma^2} W_{ij}\\
-  B_i &= \frac{-\fM^2 c_0^2 + \fD \fM c_0}{\sigma^2} T_i\\
-  g & = -\frac{1}{2 \sigma^2} (\fM^2 c_0^2 - 2 \fD \fM c_0 + \fD^2)
-\end{align}
-and then 
-\begin{equation}
-  {\bm A}_{ij}^{-1} = \frac{\sigma^2}{\fM^2 c_0^2} W^{-1}_{ij}
-\end{equation}
-then, following the same integration as before, we have
-\begin{equation}
-  p(\vstar,c_0 | \vD) \propto \frac{1}{c_0^N} \sqrt{ \frac{(2 \pi)^N}{ {\rm det} |A_{\rm old}|}} \exp \left( -\frac{\fM^2 c_0^2}{\sigma^2} + \frac{\fM \fD}{\sigma^2} (T_i W^{-1}_{ij} T_j + 1 ) c_0 - \frac{\fD^2}{2 \sigma^2}  (T_i W^{-1}_{ij} T_j + 1 )\right)
-\end{equation}
-and then we can put our log-normal prior on $c_0$.
-Where $N$ is the number of higher order terms, in our case $N=3$.
-This function is not able to be integrated analytically, and so two-step marginalization is not feasible.
 
@@
 
-Compare distributions by doing a K-S test on the cumulative distributions?
+Compare distributions by doing a K-S test on the cumulative distributions between Gaussian and Log-normal?
 
@@
 Citation to DFM \citep{fhl+12}, GW \citep{gw10}, and CosmoHAMMER \citep{asa+13}.
+
+\subsection{Noise, Masking, Residuals, and Alternative likelihood functions}
+\label{sec:residuals}
+In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment.
+
+We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).
+We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid.
+
+Section on how to choose masked regions, testing the PHOENIX grid to see where the derivative is highest and thus gives the best leverage on stellar parameters.
+
+Section on interpolation errors.
 
@@
 
-%\appendix
-%
-%\section{}
-%
+\appendix
+
+\section{Simplification for small $c_0$ perturbations}
+\label{sec:gaussian_simplification}
+If the correction in $c_0$ is small, we can use an alternate formulation of $k$ and put Gaussian priors on all nuisance parameters.
+This allows analytic marginalization over all nuisance parameters (including $c_0$), and reduces the dimensionality of the posterior to only the $\sim 7$ stellar parameters, $\vstar$.
+
+\begin{align}
+  k(\lambda | \vN) &= c_0 T_0(\lambda) + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \\
+  k(\lambda | \vN) &= \sum^N_{i = 0} c_i T_i(\lambda)
+\end{align}
+Following \S\ref{sec:nuisance}
+\begin{equation}
+  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ]\fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
+\end{equation}
+\begin{equation}
+ \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM \sum_n c_n T_n + \fM^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
+ \end{equation}
+Now the column vectors $\vc$ and $\vec{T}(\lambda)$ include $c_0$ as well
+\begin{equation}
+  \vc = 
+  \begin{bmatrix}
+    c_0\\
+    c_1\\
+    \vdots\\
+    c_N
+  \end{bmatrix}
+  \hspace{3cm}
+\vec{T}(\lambda) = 
+\begin{bmatrix}
+T_0(\lambda)\\
+T_1(\lambda)\\
+\vdots\\
+T_N(\lambda)\\
+\end{bmatrix}
+\end{equation}
+\begin{equation}
+  k(\lambda | \vN) = \sum^N_{i = 0} c_i T_i = \vec{T}^\trans \cdot \vc
+\end{equation}
+\begin{equation}
+  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
+  \begin{bmatrix}
+T_0 T_0 & T_0 T_1 &  \hdots & T_0 T_N \\
+T_1 T_0 & T_1 T_1 &  \hdots & T_1 T_N \\
+\vdots  & \vdots  &  \ddots & \vdots \\
+T_N T_0 & T_N T_1 &  \hdots & T_N T_N \\
+  \end{bmatrix}
+\end{equation}
+\begin{equation}
+  \sum_n \sum_m c_n T_n c_m T_m = \vc^\trans \cdot {\bm W} \cdot \vc
+\end{equation}
+\begin{equation}
+  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM\, \vec{T}^\trans \vc \; + \fM^2 \, \vc^\trans {\bm W}  \vc \bigr ]
+\end{equation}
+\begin{align}
+  {\bm A} &= \sum_\lambda \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
+  \vec{B}^\trans &= \sum_\lambda \frac{\fD(\lambda) \fM(\lambda) }{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
+  g &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}\\
+\end{align}
+Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. If we include Gaussian priors, then we have 
+\begin{align}
+  {\bm A}^\prime &= {\bm A} + {\bm D}\\
+  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
+  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
+\end{align}
+where now ${\bm D}$ and $\vec{\mu}$ include $c_0$ as well.
+The main difference from before is that now the prior on $c_0$ is now Gaussian instead of log-normal.
+This is acceptable as long as $\sigma_{c_0}$ is small ($< XX\%$), otherwise this will bias the correction factor.
+This also means that we will usually have $\vec{\mu} = \{1, 0, \ldots, 0\}$.
+The full posterior probability function becomes
+ \begin{equation}
+   \boxed{
+  p(\vstar, \vN | \vD) = p(\vstar, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(\vc)
+}
+\end{equation}
+
+Using the previous multi-dimensional integral, we can marginalize over all $\vc$ including the Gaussian priors
+\begin{equation}
+  \boxed{
+  p(\vstar | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
+}
+\end{equation}
+
+To gain insight, we can re-express this new version of $k$ in the original form as 
+\begin{equation}
+   k(\lambda | \vN) = c_0 T_0 + c_1 T_1 + \ldots + c_N T_N
+\end{equation}
+where $c_1 = c_0 {c_1}^\prime T_0$ and $c_N = c_0 {c_N}^\prime T_0$.
+Thus, if there needs to be a large perturbation in $c_0$, then there must also be a large perturbation in $c_n$.
+The prior on $c_1$ is now a product of a log-normal prior on $c_0$ and a Gaussian prior on ${c_1}^\prime$, which is not analytically tractable.
+Therefore, in the case of large perturbations, it is better to use Equation~ and use a log-normal prior on $c_0$.
+When we use the original framework, we are saying that perturbations in $c_0$ are small enough that we can approximate the log-normal prior on $c_0$ as Gaussian and that the priors on $c_n$ can also be correctly approximated by a Gaussian.
+From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$.
+Beyond this, one should switch to this framework, for example to correct for un-flux calibrated data.
+
 %\begin{deluxetable}{ll}

### 2013-11-19 generate_webpage.py

@@ -1,0 +1,6 @@
+import jinja2
+
+'''Takes the output of a run and generates a viewable webpage. To be called by MCMC.py, likely.'''
+
+__author__ = 'ian'
+

### 2013-11-19 plot_MCMC.py

@@ -67,7 +67,6 @@ def hist_nuisance_param(flatchain):
         print(nuisance_dir, "already exists, overwriting.")
 
     cs = flatchain[:,nparams:]
-    print(cs.shape)
 
     if (config['lnprob'] == "lnprob_lognormal") or (config['lnprob'] == "lnprob_gaussian"):
         for i in range(norders):
@@ -119,6 +118,10 @@ def draw_chebyshev_samples():
     plt.show()
 
 def visualize_draws(flatchain, lnflatchain, sample_num=10):
+    '''Currently only implemented for the un-marginalized probability functions.'''
+
+    #TODO: expand to marginalized distributions by drawing from the conditionals (we can also display conditional prob function)
+
     visualize_dir = config['output_dir'] + "/" + config['name'] + '/visualize/'
     if not os.path.exists(visualize_dir):
         os.mkdir(visualize_dir)
@@ -130,7 +133,7 @@ def visualize_draws(flatchain, lnflatchain, sample_num=10):
 
     for i in range(sample_num):
         #For each sample from the posterior
-
+        sample_dir = visualize_dir + 'sample{i:0>2.0f}/'.format(i=i)
         if not os.path.exists(sample_dir):
             os.mkdir(sample_dir)
 
@@ -138,45 +141,55 @@ def visualize_draws(flatchain, lnflatchain, sample_num=10):
         ind = np.random.choice(all_inds)
         p = flatchain[ind]
         lnp = lnflatchain[ind]
-        sample_dir = visualize_dir + 'sample{i:0>2.0f}/'.format(i=i)
+
         #write p and lnp to a file in sample dir
-        f = open('params.txt',"w")
+        f = open(sample_dir + 'params.txt',"w")
         f.write("Parameters: %s \n" % (p,))
         f.write("lnp: %s \n" % (lnp,))
         f.close()
 
         #Get wavelength, flux, and sigmas
-        wl = m.wls[order]
-        fl = m.fls[order]
-        sigma = m.sigmas[order]
+        wls = m.wls
+        fls = m.fls
+        sigmas = m.sigmas
 
         #Reproduce the model spectrum for that parameter combo
-        wlsz = m.shift_TRES(2.)
-        f = m.model(wlsz, 5900., 3.5, 0.0, 5., 0.0, 1e-10)[order]
[…]

### 2013-11-19 MCMC.py
Added model_p, working visualize script

@@ -80,7 +80,7 @@ def main():
         print(outdir, "already exists, overwriting.")
 
     #Copy config.yaml and SLURM script to this directory
-    shutil.copy('config.yaml', outdir + 'config.yaml')
+    shutil.copy(confname, outdir + confname)
     shutil.copy('run', outdir + 'run')
 
     # Choose an initial set of positions for the walkers, randomly distributed across a reasonable range of parameters.
@@ -128,14 +128,20 @@ def main():
     np.save(outdir + "chain.npy", sampler.chain)
     #write lnprob to file
     np.save(outdir + "lnprobchain.npy", sampler.lnprobability)
+    #write flatchain to file
+    np.save(outdir + "flatchain.npy", sampler.flatchain)
+    #write flatlnprob to file
+    np.save(outdir + "flatlnprobchain.npy", sampler.flatlnprobability)
+
+
 
     ### if config['plots'] == True, Call routines to make plots of output ###
     if config['plots'] == True:
         plot_MCMC.auto_hist_param(sampler.flatchain)
         plot_MCMC.hist_nuisance_param(sampler.flatchain)
+        plot_MCMC.visualize_draws(sampler.flatchain, sampler.flatlnprobability)
     #Histograms of parameters
     #Walker positions as function of step position
-    #Samples from the posterior overlaid with the data
 
 
 if __name__ == "__main__":

### 2013-11-19 model.py

@@ -396,6 +396,32 @@ def model(wlsz, temp, logg, Z, vsini, Av, flux_factor):
     gc.collect() #necessary to prevent memory leak!
     return result
 
+def model_p(p):
+    '''Post processing routine that can take all parameter values and return the model.
+    Actual sampling does not require the use of this method since it is slow.'''
+    temp, logg, Z, vsini, vz, Av, flux_factor = p[:config['nparams']]
+
+    wlsz = wls * np.sqrt((c_kms + vz) / (c_kms - vz))
+    fmods = model(wlsz, temp, logg, Z, vsini, Av, flux_factor)
+
+    coefs = p[config['nparams']:]
+    # reshape to (norders, 4)
+    coefs_arr = coefs.reshape(len(orders), -1)
+    c0s = coefs_arr[:,0] #length norders
+    cns = coefs_arr[:,1:] #shape (norders, 3)
+
+    #now create polynomials for each order, and multiply through fls
+    #print("c0s.shape", c0s.shape)
+    #print("cns.shape", cns.shape)
+    #print("T.shape", T.shape)
+
+    Tc = np.einsum("jk,ij->ik", T,cns)
+    #print("Tc.shape",Tc.shape)
+    k = np.einsum("i,ij->ij",c0s, 1 + Tc)
+    #print("k.shape",k.shape)
+    #print("fmods.shape",fmods.shape)
+    refluxed = k * fmods
+    return [k, refluxed]
 
 def degrade_flux(wl, w, f_full):
     vsini = 40.
@@ -517,8 +543,6 @@ def lnprob_lognormal(p):
         a= fm2c2/sigmas**2
         A = np.einsum("in,jkn->ijk",a,TT)
         Ap = A + D
-        detA = np.array(list(map(np.linalg.det, Ap)))
-        invA = np.array(list(map(np.linalg.inv, Ap)))
 
         b = (-fm2c2 + fdfmc0) / sigmas**2
         B = np.einsum("in,jn->ij",b,T)
@@ -527,9 +551,6 @@ def lnprob_lognormal(p):
         g = -0.5/sigmas**2 * (fm2c2 - 2 * fdfmc0 + fls**2)
[…]

### 2014-01-15 model.py
Running lkca15 on kurucz

@@ -72,7 +72,7 @@ grid = grids[config['grid']]
 T_points = grid['T_points']
 logg_points = grid['logg_points']
 Z_points = grid['Z_points']
-alpha_points = grid['alpha_points']
+#alpha_points = grid['alpha_points']
 
 #Limit grid size to relevant region
 grid_params = config['grid_params']
@@ -92,10 +92,10 @@ Z_ind = (Z_points >= Z_low) & (Z_points <= Z_high)
 Z_points = Z_points[Z_ind]
 Z_arg = np.where(Z_ind)[0]
 
-A_low, A_high = grid_params['alpha_range']
-A_ind = (alpha_points >= A_low) & (alpha_points <= A_high)
-A_points = alpha_points[A_ind]
-A_arg = np.where(A_ind)[0]
+#A_low, A_high = grid_params['alpha_range']
+#A_ind = (alpha_points >= A_low) & (alpha_points <= A_high)
+#A_points = alpha_points[A_ind]
+#A_arg = np.where(A_ind)[0]
 
 #Load the data to fit
 base = 'data/' + config['dataset']
@@ -330,7 +330,7 @@ def trilinear_interpolator():
     return intp_func
 
 
-flux = quadlinear_interpolator()
+flux = trilinear_interpolator()
 
 ##################################################
 #Stellar Broadening
@@ -514,7 +514,57 @@ blended_real = pyfftw.n_byte_align_empty(chunk, 16, "float64")
 fft_object = pyfftw.FFTW(f_full, FF)
 ifft_object = pyfftw.FFTW(FF, blended, direction='FFTW_BACKWARD')
 
-def model(wlsz, temp, logg, Z, alpha, vsini, Av, flux_factor):
+def model(wlsz, temp, logg, Z, vsini, Av, flux_factor):
+    '''Given parameters, return the model, exactly sliced to match the format of the echelle spectra in `efile`.
+    `temp` is effective temperature of photosphere. vsini in km/s. vz is radial velocity, negative values imply
+    blueshift. Assumes M, R are in solar units, and that d is in parsecs'''
+    #wlsz has length norders
+
+    #M = M * M_sun #g
+    #R = R * R_sun #cm
+    #d = d * pc #cm
+
+    #logg = np.log10(G * M / R**2)
+    #flux_factor = R**2/d**2 #prefactor by which to multiply model flux (at surface of star) to get recieved TRES flux
+
[…]

### 2014-01-15 paper/paper.tex
Running lkca15 on kurucz

@@
-%\documentclass[manuscript]{aastex} %one-column, double-spaced document
 \documentclass[preprint]{aastex} %double-column, single-spaced document:
@@
 \newcommand{\vt}{\vec{\theta}}
+\newcommand{\vg}{\vt_{\star, {\rm grid}}}
 \newcommand{\vstar}{\vt_{\star}}
@@
 \newcommand{\trans}{\mathsf{T}}
+\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
+\newcommand{\A}{[\alpha/{\rm Fe}]}
 
@@
 
-\title{A Bayesian method for inference of fundamental stellar parameters}
+\title{A method for inference of fundamental stellar parameters using high resolution spectra}
 \author{\today{}\\
@@
 
-%\begin{abstract}
-%\end{abstract}
+\begin{abstract}
+  Matching observed spectra of stars against a grid synthetic spectra is a popular approach to determine fundamental stellar parameters such as effective temperature, surface gravity, and metallicity.
+However, a fundamental problem with this approach is that the synthetic spectra contain systematic errors such as wrong or missing spectral lines.
+In the presence of such errors, a simple $\chi^2$ fit between spectra will likely yield biased parameters and incorrect uncertainty estimates.
+This behaviour is analogous to the way a few strong outliers might severely bias the fit to a straight line.
+These spectral outliers are difficult to identify \emph{a priori}; procedures such as sigma clipping might be prone to their own systematic effects and discard useful information.
+We outline an approach to probabilistically identify and account for spectral features that deviate from reality.
+Such an approach is most needed when operating at high spectral resolution ($R >$ 10,000), since this is when systematic deviations begin to appear most striking, however our technique is still valid for lower resolution spectra as well.
+We also outline some approaches for spectral fits to brown dwarfs and planets, where instead of spectral lines there may be systematic deviations of a larger swath of the spectrum.
+\end{abstract}
+
 %\keywords{}
@@
 
-The fundamental parameters of stars, such as mass, luminosity, effective photospheric temperature, surface gravity, and metallicity are of supreme interest in astrophysics, for a wide variety of reasons.
-Exoplanet properties are found relative to the properties of the host star, and have important implications for models of planet formation.
-The properties are also important to the study of disks, for the incident radiation field sets photoevaporation, influences disk dispersal.
-Many methods exist for determination of stellar parameters: SPC \citep{blj+12}, MOOG, SME \citep{vp96}, and these are robust for FKG stars, where there is a well-defined continuum and the stars are ``normal.'' However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.
+The fundamental parameters of stars---mass, luminosity, effective photospheric temperature, surface gravity, and metallicity--are of supreme interest to astrophysicists for a wide variety of reasons.
+A salient example is that most newly discovered exoplanets must necessarily their mass and radius measured relative to the properties of their host stars (e.g.
+\citealt{tfs+12,blj+12,ssm+13}).
+The properties of stars are of course also important in their own right because they inform us about our ideas of stellar evolution \citep{dm97, bca+02}.
+Because stars are the fundamental units of the cosmos, techniques to infer the properties of stars date back to the dawn of astronomy.
+We briefly review some of the modern and popular techniques in the literature.
+Approaches can be broadly classified into techniques that compare stars by some broadly defined indices and techniques that attempt to forward-model an observed stellar spectrum by doing a pixel-to-pixel comparison.
 
-\section{The model}
+\subsection{Similarity techniques}
+Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra.
+I am calling them similarity techniques because they compare characteristics of spectra, not actual pixels.
 
-To summarize our model, we use Bayes rule to design a posterior probability function
+\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity.
+Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters.
+A recent example using M dwarfs is \citet{nci+14}.
+This technique works, but there are many things that could be improved.
+First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars.
+This is both a disadvantage and an advantage.
+One one hand by choosing a subset of lines to fit, one may be unwittingly throwing away useful information.
+On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters.
+Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.
+
+\paragraph{Principal component analysis} PCA reduces the dimensionality of the data space and searches for the most salient features of the spectra \citep{ptb14}.
+How does one properly estimate parameter uncertainties in such an approach?
+It seems like this approach exists because it is an easy thing to do.
+
+Matrix inversions: MATISSE \citep{rbd06}.
+This method is similar to PCA in that it derives basis vectors that describe the fundamental parameters of the star.
+In the PCA analysis, the principal components do not necessarily correspond to true physical quantities.
+The MATISSE algorithm determines weights that ensure the components are similar to the stellar parameters.
+They only test the algorithm against input synthetic stellar spectra to detect biases in the parameters, and against the Sun and Arcturus.
+They do acknowledge that there will be systematic deviations between the synthetic spectra and real spectra, but they do not account for the possible inflation of error estimates.
+
+Given the inherent systematic uncertainties involved in each method, some groups elect to combine the inference from different algorithms, such as (SLOAN SSPP) \citep{lbs+08}.
+This approach includes machine learning along with many of the previously mentioned techniques.
+
+\subsection{Pixel to pixel techniques using spectral libraries}
+Many high quality synthetic spectral grids are now available.
+Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement.
+We define this combination as
 \begin{equation}
-  \overbrace{p(\vt | \vD)}^{\rm posterior} \propto \underbrace{p(\vD | \vt)}_{\rm likelihood} \; \overbrace{p(\vt)}^{\rm prior}
-  \label{eqn:posterior}
+  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
 \end{equation}
-where the parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$.
+Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid.
+Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.
+
+\paragraph{SPC} A very popular method behind the Kepler follow up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
+This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method.
+The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient.
+It is uncertain how statistically justified this uncertainty estimate is.
+While this is a pixel-to-pixel method, it uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match  high resolution models of the Sun and Vega.
+
+Most other pixel-to-pixel techniques involve comparison via a $\chi^2$ likelihood function.
+
+\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations.
+Any spectral library can be used as a backend, including empirical spectral libraries, such as the one from ELODIE \citep{psk+07}.
+population uses an iterative kappa-sigma cleaning method to discard ``bad'' lines.
+They test for convergence to the same answer by starting guesses located at different parameters and seeing if the same lines are masked.
+
+Takes into account many of the things that need to be considered, including multiplication by a polynomial and rejection of bad data or model points.
+
+Bayesian methods exist, but they use a $\chi^2$ likelihood function \citep{sdm+07}.
+
+\citet{sb13} mention a comprehensive Bayesian method that can fit low or high resolution spectra.
+They too, mention that they use a $\chi^2$ likelihood.
+They allude to using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were identified.
+Also, some stellar grids focus on the most important parts of the spectrum, as identified by experts.
+Quoted from \citet{sb13}, grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).'' These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.
+
+\citet{mga13} fit BT-Settl PHOENIX models to M dwarfs.
+Surely these models have some serious problems with them.
+They use a $\chi^2$-like likelihood function, and iterate to clean bad pixels.
+They iterate the method against their grid of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars.
+Presumably, this would be due to a bad opacity for a certain line or molecular feature.
+Then, they mask these pixels and repeat the fitting.
+They claim that after two iterations the masks do not change.
+They note that rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures.
+However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.''
+
+\subsection{Line fitting using spectral synthesis}
+This is an improvement over the line index method, since now one includes the information contained in the line profile.
+With quality stellar spectral models, one can identify a few well-modelled lines to fit.
+By choosing a few good lines to fit, one could determine stellar parameters by matching line profiles for a few select lines \citep{dj03}.
+However, this is still subject to the need to identify promising regions of the spectrum.
+
+MOOG\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73}, SME\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96}, actually synthesize line profiles on the fly, compared to using a library of spectral models.
+This has the advantage of enabling individual tweaks to elemental abundances.
+And interpolation in the model atmosphere space, rather than the spectral line space.
+However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.
+Matching equivalent widths.
+Interestingly, MOOG sythesizes a spectrum on the fly.
+
+SME allows adjustment of oscillator strengths.
+We could download the synthesis code for SME and synthesize our own spectra on the fly.
+The biggest impediment here would be that the simplest atmospheres will not get certain lines correct.
+The powerful synthesis codes (PHOENIX) will get the atmospheres and lines correct, but they take a long time to run (~hours) per model, and the code is not publicly available.
+To do a full-spectrum fit for these sorts of codes would likely be too computationally difficult.
+However, it would allow access to fitting in the parameter space that we want.
+
+\subsection{Stellar parameter estimation in a utopia}
+
+Although the spectral synthesis domain is the correct domain to solve for stellar parameters, the radiation transfer is too slow.
+
+The fitting difficulties of the previous section are a result of trying to fit data to a model in a smaller parameter space.
+Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different.
+If we think back to first principles, the ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$.
+These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote some value relative to solar, a fraction of models allow changes in alpha abundances), and atomic constants describing the opacity contribution of a line.
+\emph{All} of these parameters have uncertainties in them, some of these are rather dramatic (one continually hears lamentation about the opacities and line lists\dots), yet for any given stellar model these parameters are \emph{fixed}.
+There may yet be unconsidered stellar physics necessary to explain a small fraction of lines (say for example the cores of Ca lines), but the field of stellar modelling has progressed to the state that I would have the confidence to say that if a group focused all of their attention on a single line, and using only their current physical models and making tweaks to the afore-mentioned lines, they could produce a satisfactory model.
+
+In reality, this has been done using the Kurucz models in an unpublished grid of stellar spectra, over the range $5000 - 5300$ \AA.
+
+The real problem is much more rich, as is the amount of information one could extract from a high resolution optical spectrum.
+One would want to fit for individual abundances, marginalize over oscillator strengths.
+
+The unfortunate reality is that sampling in the stellar model space for the full optical spectrum is not yet a computational reality.
+There do exist codes (MOOG) that can do this approach but only for a small ($\sim$ a few lines) at once, and using simplifications.
+Through collaboration with a stellar atmospheric expert, such an approach may soon be a possibility.
+If the computation of a given model could be reduced to less than a minute, efficient MCMC sampling parallelized on a cluster (using \texttt{emcee}, for example).
+However, as it stands without a collaborator, I do not have the expertise nor the time to pursue such an approach.
+Therefore, we are reduced to fitting in the space of $\vg$ and devising ways to deal with the issues.
+
+\section{General problems with the spectral library approach}
+
+Pixel-to-pixel approaches have the potential to be much more powerful than other historical stellar parameter estimation techniques.
+When comparing directly in the data space, we have the ability to add in other physics as we see fit.
+
+because we can fit multiple stars at once (say for unresolved binary or multiple systems), fit for velocity components, etc.
+Many of the most interesting systems for which we would want to determine stellar parameters are in multiple systems anyway, since we can dynamically determine masses for these systems and thus compare to evolutionary models.
+Determine veiling.
+And a perhaps wooden-headed belief that using the full spectrum \emph{should} give us a more accurate and less-biased posterior estimate.
+Such a belief may be foolish and in reality different parts of the spectrum could give completely different parameter estimates.
+
+Many, many methods exist for comparison of stellar spectra, including many ``Bayesian'' approaches in the pixel-to-pixel space.
+The nature of the Bayesian-ness, however, is usually limited to priors on temperature, surface gravity, and metallicity.
+These approaches do not explicitly state the likelihood function that they use, or if they do, it is a $\chi^2$ function.
+
+The more successful of these models use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000 ish). Or, a grid over a narrow range of wavelength space with finely tuned atomic constants.
+
+\textbf{The true fundamental problem with this approach is that the stellar spectra are not a perfect representation of reality}.
+Assumptions about Poisson errors break down, and fits become contaminated by systematics.
+There are types of errors which occur.
+
+FIGURE: 4 Panel Figure, from left to right, showing Class 0 through III errors.
+Use Mg b lines, other big line near Mg b triplet, order 24 line, and then Fe I/II line complex that doesn't fit properly.
+
+\paragraph{Class I} A spectral line appears in the data but not in the model spectrum over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$.
+Such a type of error is perhaps the easiest to understand and to forgive.
+For example, such a line could easily appear in the data due to interstellar absorption or telluric absorption, something which a stellar model could never produce.
+The model could also fail to reproduce a true stellar line due to a missing opacity source.
+
+\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$.
+This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident.
+Presumably this is a product of bad atmospheric structure (temperature or pressure is wrong at some point?) that excites a line when it shouldn't otherwise.
+
+\paragraph{Traditional ways to model Class I and Class II errors} are a form of sigma clipping.
+Mann et al.
+compares the stars to models over a wide range of parameters.
+One can track how these lines might be bad because Class I residuals (data - model)/sigma will always be negative (since we are dealing with absorption lines only) and not change significantly with $\vg$, since we the model can only produce continuum.
+Class II residuals will always be positive and should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.
+
+If a Class I or II error occurs on a line, that region of the spectrum is relatively useless for determining spectral parameters.
+Authors like Mann et al.
+or Kholeva et al.
+find these regions and simply mask them out of the final computation.
+Kholeva et al., and others approach in an iterative manner.
+Kholeva et al.
+approaches are more sophisticated because they track convergence.
+This is a sigma clipping approach, and while we could go and model these a la Hogg 2010, the approach is not super justified, result is probably not too wrong (similar to actual sigma clipping).
+Given that Kholeva et al.
+track convergence maps, I would place a higher degree of confidence in the answer.
+
+In our later approaches in Sec XX, how might we put a prior on $\Delta \vg$, the range of parameters over which the residuals will be correlated?
+Maybe we could Gibbs sample in the same way as the bad flags.
+We'd have to keep a bunch of spectra with varying $\vg$ in memory to make this comparison easy.
+
+\paragraph{Class III} In truth, this is just a superset of both Class I and Class II.
+Class III errors are when the line strengths between the data and model do not match at a $\vg$ that fits a majority of lines.
+What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, however this parameter combination would result in bad fits for the majority of the other lines.
+Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$.
+
+In addition, it is difficult to determine the ensemble of lines that suffer from Class III errors, since the strength of some lines will be under-predicted while others will be over-predicted.
+How do we know which lines to trust?
+If we take an approach similar to fixing Class I and II errors, and mask out bad regions, it would be very difficult to distinguish Class III errors from \emph{truthful} lines at an incorrect $\vg$.
+For example, if one imagines using a cleaning algorithm as described for Class I and Class II errors, the identification of Class III errors will be sensitive to the starting guesses of $\vg$.
+
+Because these Class III lines can still produce a significant residual spike, they tend to distort and in some cases artificially peak the posterior distribution.
+There is no one parameter combination that will fit all lines perfectly.
+Certainly there is a parameter combination that will produce a reasonable fit to 90\% of the lines to within 3 sigma residuals, but there will still be a fair number of both ``high'' and ``low'' regions.
+These lines need to be identified and dealt with, but how?
+
+How then, should the best fit be determined? Surely some of the remaining lines are ``bad.'' But in this regime, determination of which lines are bad will bias the results in a certain direction.
+
+Also, which lines should one actually fit? The stronger lines? Did the modelers themselves give more attention to the atomic constants which produce the strong lines?
+
+To summarize, the fundamental problem with comparing high resolution optical stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines.
+Deciding how to model the lines is then the difficult part.
+Different groups have sigma-clipped lines that are obviously bad, and this does improve the procedure dramatically.
+In such a case, the procedure of sigma-clipping doesn't gall with me too badly.
+However, the more difficult problem is dealing with Class III errors properly, something that no modeler has mentioned or successfully implemented in a statistically justified manner.
+
+In a spectrum fit, many lines do reasonably fit within the Poisson errors. For consistency, we will denote these as ``Class 0'' lines, where there is no error.
+
+\subsection{Testing the accuracy of stellar libraries}
+Gaia team members are panicking about the upcoming flood of data.
+
+\citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia.
+
+\citep{cbm+05} is the Coehlo grid.
+
+\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}
+
+The parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$.
 Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum.
 For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.
+
+The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
+\begin{equation}
+  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
+\end{equation}
+where the best model is defined as one that minimizes the residuals in some sense.
+We will discuss other goodness of fit functions soon.
+Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda)$.
+
+FIGURE showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000.
+This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species.
+For example, we could inject bad lines in this space.
+
+\subsection{Processing TRES Data}
+We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel.
+By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
+\begin{equation}
+  S/N = \sqrt{ {\rm cts}}
+\end{equation}
+then the Noise-to-Signal ratio is the inverse of this
+\begin{equation}
+  N/S = \frac{1}{\sqrt{ {\rm cts}}}
+\end{equation}
+then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.
+
+\subsection{Applied to a Gaussian likelihood function}
 
@@
 \end{align}
-Here we have the full posterior probability distribution defined in Equation~\ref{eqn:posterior}, assuming flat priors.
+Here we have the full posterior probability distribution assuming flat priors.
 This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$.
@@
 $\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction.
-Now, we can write Equation~\ref{eqn:posterior} with priors as
+Now, we can write the posterior with priors as
 \begin{equation}
@@
 
+\section{Moving beyond a perfect stellar model}
+\label{sec:residuals}
+
+\subsection{Additional sources of noise beyond Poisson}
+Presumably we shouldn't have to add an additional source of ``noise'' to the fit, because our nuisance parameters will find an exact flux calibration.
+
+Interpolation errors. Quantified via Husser et al. Ph.D. thesis as well as custom tests.
+
 \subsection{Noise, Masking, Residuals, and Alternative likelihood functions}
-\label{sec:residuals}
 In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment.
+
+\subsection{Masking}
 
@@
 
-Section on interpolation errors.
+Iterative masking to deal with Class I and II errors a la ULySS. They use a $\kappa - \sigma$ routine, where they iteratively sigma clip (Website link).
+
+\section{Gallery of likelihood functions}
+
+If you do a straight up $\chi^2$ fit to the data, you can get posteriors that look like $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
+This would be better than that obtained by astroseismology!
+Clearly there is something going wrong here.
+
+Lorentzian, linear combination of Gaussians, Sivia w/ prior on sigma. $\chi^2_Q$ model.
+
+These likelihood functions seem to be an improvement over the straight $\chi^2$ function, in the sense that they give parameter estimates that are
+
+\subsection{Bad data model}
+The bad data model \citep{pre97,dbl10} is a powerful framework for forward-modelling of data, allowing the modeler to account for the probability of outliers.
+
+Now, we have a $q_i$ for each data point, such that for the entire data set we have an ensemble of $\qN$.
+If this ensemble of $\qN$ is chosen appropriately, and the good data points are flagged as good and the bad data points are flagged as bad, the likelihood will be maximized.
+Written out explicitly, this is
+\begin{multline}
+  {\cal L} = p({\bm D} | m,b,\qN,Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma^2} \right] \right ]^{q_i}  \times \\
+  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma_{y_i}^2)}} \exp \left[ - \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{y_i}^2)} \right]
+  \right ]^{1 - q_i}
+\end{multline}
+
+using a Bernoulli prior
+
+\begin{equation}
+  p(q_i | P_b) = [1-P_b]^{q_i} P_b^{[1-q_i]}
+\end{equation}
+
+\begin{equation}
+  p(\qN | P_b) = \prod_{i=1}^N [1-P_b]^{q_i} P_b^{[1-q_i]}
+\end{equation}
+
+For this data point, we can integrate (in this case, discrete sum) over the two states of $q_i$: $q_i = 1$ and $q_i = 0$. 
+\begin{multline}
+  p(m, b, P_b, Y_b, V_b | D_i)  =  \int p( m, b, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
+  = p(m, b, q_i=1, P_b, Y_b, V_b | D_i) + p(m, b, q_i=0, P_b, Y_b, V_b | D_i) \\
+  = \Bigl[ p_{\rm good}(D_i | m,b)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(m,b,P_b, Y_b,V_b) 
+\end{multline}
+
+\begin{multline}
+  p(m,b,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i | m, b)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(m,b,P_b, Y_b, V_b)\\
+  \boxed{
+    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_{yi}^2}} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma_{yi}^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma_{yi}^2)}} \exp \left[- \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{yi}^2)} \right] \right\}
+  \times p(m,b,P_b,Y_b,V_b)}
+\end{multline}
+
+In it's simplest form (Hogg 2010), the bad data model results in a linear combination of Gaussians.
+
+A benefit of the bad data model over a chosen likelihood function is that it is motivated.
+It may still result in the linear combination of Gaussians (nice), but preserves real parameter correlations.
+You are not forcing the same mean.
+
+\subsection{Spatial correlations of pixels}
+We can do more.
+The bad model of the previous section assumes that each data point is independent.
+For example, the spectral lines are actually bad, and this means that if a single line is wrong, not just the core of the line (which may be a 10 - 20 sigma residual) will be wrong, but also the wings of the line, which may be only 5 or 3 sigma wrong.
+
+Now, let's come up with a model which includes spatial (wavelength) correlations in the bad data model.
+
+How could we accomplish this?
+If we identify one point as bad, then the flags around it will have a correlation length, ie, the prior on $q_i$ will be bad.
+This means that for however many nearby pixels that will be bad (most likely a constant amount for uncrowded lines), then this means that we could still integrate over the qis, it's just that we would now have to take a contribution from the nearest 5 pixels, say.
+That central pixel would then mark the ``line'' that is bad.
+
+That is one way to mark the ``badness of the line,'' that the qi are correlated. But we also know that the $y_i$ will be correlated too.
+
+This correlation length will probably depend on the type of star, and the contribution of the band.
+For example, if they're simple FGK stars, then the bad lines might be singular.
+But if they are M stars, then there is a chance that there might be a large ``pseudo-continuum'' region that might be incorrect.
+
+Once we have identified the center of the bad line, then there will be a decrease in the nearby pixels, similar to a Gaussian shape.
+Presumably we could come up with a correlation in this space too.
+(or, it could simply be that the correlation between $y_i$ is flat in the case of an M star.
+
+This might actually save the marginalization. We could still do the backwards ``what is the probability a given line is bad.''
+
+For Class I and II errors, we could have a hyper parameter describing the width of each line, and another hyper parameter describing the (distribution of) depth. These could also be sampled in.
+
+For Class III errors, describing how line strengths are off makes sense for stars with continuum, but as we move to later type stars (K stars) that do not have a nice continuum, we may run into trouble.
+For the later stars (M dwarfs), there might be a whole region that is incorrect.
+
+There are different ways to describe a correlation over pixels. Gaussian random fields. Autocorrelation function.
+
+\subsection{Simple correlation between pixels}
+For a simple correlation, say that if a given $q_i$ is bad, then there is also a high probability that the neighboring pixels are also bad.
+Say we know from inspection that it seems as though bad lines knock out 5 pixel chunks.
+
+Caveat: I think this assumes that bad lines will not occur within the same spread of each other. Maybe we should include a $P_B$ prior on the good counts too?
+
+\begin{equation}
+  p(q_i = 0 | q_{i \pm 5} = 0) = P_B
+\end{equation}
+
+\begin{align}
+ p(q_i = 0 | q_{i \pm 5} = 0) &= 1/5 \\
+ q_{i \pm 4} &= 2/5 \\
+ q_{i \pm 3} &= 3/5 \\
+ q_{i \pm 2} &= 4/5 \\
+\end{align}
+
+This might need more work with $P_B$.
+
+Then the probability functions would be
+\begin{multline}
+  p(D_{i + j} | \vt, q_i =0) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=0) + \\
+    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 0) \bigr ]
+\end{multline}
+
+\begin{multline}
+  p(D_{i + j} | \vt, q_i = 1) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=1) + \\
+    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 1) \bigr ]
+\end{multline}
+
+The marginalization could still be done analytically, but now the likelihood function involves a correlation between adjacent pixels.
+Could we use a sparse array for this?
+Or, we could evaluate all of the pixels, and then have the final likelihood function involve some correlation terms between the lnprobs evaluated at each pixel.
+We could do this using the logarithm so that everything is additive.
+
+Or, what if we evaluated the FFT of the $\log(p(D_i))$ and just the autocorrelation in Fourier space?
+
+\subsection{Correlation between $y_i$}
+Not only will the flags be spatially correlated (if a point is bad, the nearby $\sim$ 5 points are also likely to be bad) but for OBAFGK stars, the lines are likely to be Gaussian shape.
+That is if $q_i$ specifies the center of the line, then the neighbooring $y_{i+1}$ or $y_{i-1}$ will be of lower residual height.
+This means that the inside of the likelihood function should also have spatial correlations.
+
+Could this correlation be described by a Gaussian random process?
+
+\subsection{Different flags and weights}
+Instead of a good or bad flag, what if we had a flag that would identify which class of error the line became? For example we could have the option of $q_i = \{0, 1, 2, 3\}$, where $q_i=0$.
+
+How do we account for the fact that some parts of the spectrum carry much more information about $\vg$ than other parts of the spectrum?
+Is it important to weight this parts beforehand?
+Or should this contribution to the posterior naturally weight itself, the way certain points might heavily contribute to the slope and intercept of a linear fit?
+
+If so, we could imagine having a Gibbs sampler on the weights of each pixel.
+
+Proper Class III modeling is very important, following the discussion of \citep{mga13} Mann et al, and see recent example from Rajpurohit 2103.
+Some features will not fit perfectly, especially at the lower temperature scale.
+However, it's important to still take these features int account with the modelling.
+They also use a $\chi^2$ likelihood function, but perhaps even less appropriate is that they draw error bars by saying that the $\chi^2$ value changes by 5\%.
+In reality, they are checking the fits by eye and saying whether or not a spectrum matches.
+This approach is no more justified than a by-eye spectral classification.
+
+\subsection{Sampling such a model}
+If our bad model is sufficiently complex that we are not able to analytically marginalize over the $q_i$, it is still possible to sample in this huge parameter space.
+A stellar model with flags for each pixel would have $\vstar + N$ parameters, where for a typical optical spectrum $N >$ 100,000.
+Even using an ensemble sampler, sampling in this space would take a tremendous amount of time to converge.
+
+An alternative way to approach this problem is to do Gibbs sampling in the flags.
+It may be possible to do do Gibbs sampling where we draw from the conditional probability distribution, or at the very worst we could still do a Gibbs sampler with a Metropolis-Hastings backend, where we propose a jump in a single $q_i$ at a time.
+Such an sample requires none of the previous heavy lifting (spectral grid interpolation, FFT, or pixel interpolation), and so even though there are $\sim$ 100,000 flags to sample, this is likely to be computationally feasible.
+
+\section{Injection of fake lines}
+
+An alternative approach is instead of modelling the systematically wrong regions of the spectrum, actually add in lines to the spectrum.
+This can be thought of as correcting the models.
+We could add in a set number of lines.
+
+Presumably, each line would be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
+Strength could be described by a hyper prior, and FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM (in velocity space).
+
+In later type spectra, this might actually need to be a larger continuum feature that is added.
+
+We could penalize the model fits by having a hyperprior on the number of bad lines that can be created.
+
+In theory, it is possible to fit the profiles of only certain elements to determine the abundances of certain elements (see citation in conference proceedings of Husser).
+
+It may be possible to use a linelist from the group that produces the PHOENIX spectra, which describes the element that gave the highest contribution to the opacity in that bin.
+Now, if the vanadium lines are bad, then we could have a prior that the other vanadium lines might also be bad.
+We could assemble a piecewise spectrum, where certain pixels are drawn from a different combination of $\vg$, and these pixels had an associated normalization constant so that the piecewise transition was smooth (if, say the bolometric temperature was different).
+
+Class III errors might be difficult to fit in this approach. However, if the line width does not change dramatically, then we should be able to account for things this way.
 
@@
 Comparisons using HIRES data?
-
-\section{Processing TRES Data}
-We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel.
-By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
-\begin{equation}
-  S/N = \sqrt{ {\rm cts}}
-\end{equation}
-then the Noise-to-Signal ratio is the inverse of this
-\begin{equation}
-  N/S = \frac{1}{\sqrt{ {\rm cts}}}
-\end{equation}
-then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.
 
@@
 
-\bibliography{disks}
+\bibliography{disks,master}
 \bibliographystyle{hapj}

### 2014-01-15 paper/paper.tex
To run GMAur

@@
 
-\paragraph{SPC} A very popular method behind the Kepler follow up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
+The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
+\begin{equation}
+  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
+\end{equation}
+where the best model is defined as one that minimizes the residuals in some sense.
+We will discuss other goodness of fit functions soon.
+Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda | \vt)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda | \vt)$.
+
+Most pixel-to-pixel techniques use a $\chi^2$ likelihood function and seek to minimize the square of the residuals
+\begin{equation}
+  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
+\end{equation}
+It appears that for many techniques, $\sigma(\lambda)$ is the Poisson photon counting errors.
+For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
+For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.
+
+\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations.
+Any spectral library can be used as a backend, including empirical spectral libraries such as the one from ELODIE \citep{psk+07}.
+They explicitly take into account some of the necessary procedures when fitting spectra, such as continuum-multiplication by a polynomial to account for flux calibration errors and rejection of bad data or model points.
+They use an iterative kappa-sigma clipping method to discard ``bad'' points for the fit.
+They test that their clipping technique converges to the same answer by starting guesses located at different parameters $\vt$ and seeing if similar lines are masked and similar best fit parameters $\vt$ are obtained.
+
+\paragraph{\citet{mga13}} fit BT-Settl PHOENIX models to M dwarfs.
+Because high quality, high resolution M dwarf spectra are difficult to obtain owing to their intrinsic faintness, the development of stellar libraries for later spectral types has typically lagged behind.
+They use a $\chi^2$-like likelihood function and iterate to mask bad pixels.
+Unlike ULySS, however, they iterate the method against a sample of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars.
+Presumably, such a consistent residual would be due to an incorrect physical constant for a certain line or molecular feature.
+After masking, they repeat the fitting and claim that after two iterations the masks do not change.
+They note that in the future, rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures.
+\textbf{However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.}'' Thus, masking ``bad regions'' could very well destroy stellar information in the pursuit of a better ``fit.'' This is a nuanced point that masking schemes generally neglect.
+We aim to address later in this document.
+
+\paragraph{``Bayesian'' methods} do exist.
+\citet{sdm+07} developed a hierarchical Bayesian approach that purportedly accounts for both statistical and systematic errors in the estimation of stellar parameters.
+I need to work through their math more carefully, but it seems as though they use a hierarchical model for the scatter between ``true'' and observed spectrum, and the unknown scatter in the relationship between $\vg$ and the $\fM$.
+They claim to account for systematics, but it looks like these are systematics errors due to the performance of the instrument, such as a bad flux calibration.
+They use infrared spectra at low resolution ($R \sim $ 1,200), and a weighted $\chi^2$ function for goodness of fit.
+As far as I can ascertain, the actually pixel-to-pixel fitting does not account for systematic errors in regions of the model spectrum.
+
+\citet{sb13} present a comprehensive Bayesian method that can fit low or high resolution spectra in addition to photometry and parallax data.
+They too, mention that they use a $\chi^2$ likelihood.
+They mention using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were created or applied.
+Their unpublished stellar library focuses on parts of the optical spectrum which contain the most stellar information, as determined by experts.
+As noted by \citet{sb13}, the grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA\ (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).''
+
+\paragraph{SPC} A very popular method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
 This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method.
@@
 
-Most other pixel-to-pixel techniques involve comparison via a $\chi^2$ likelihood function.
-
-\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations.
-Any spectral library can be used as a backend, including empirical spectral libraries, such as the one from ELODIE \citep{psk+07}.
-population uses an iterative kappa-sigma cleaning method to discard ``bad'' lines.
-They test for convergence to the same answer by starting guesses located at different parameters and seeing if the same lines are masked.
-
-Takes into account many of the things that need to be considered, including multiplication by a polynomial and rejection of bad data or model points.
-
-Bayesian methods exist, but they use a $\chi^2$ likelihood function \citep{sdm+07}.
-
-\citet{sb13} mention a comprehensive Bayesian method that can fit low or high resolution spectra.
-They too, mention that they use a $\chi^2$ likelihood.
-They allude to using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were identified.
-Also, some stellar grids focus on the most important parts of the spectrum, as identified by experts.
-Quoted from \citet{sb13}, grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).'' These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.
-
-\citet{mga13} fit BT-Settl PHOENIX models to M dwarfs.
-Surely these models have some serious problems with them.
-They use a $\chi^2$-like likelihood function, and iterate to clean bad pixels.
-They iterate the method against their grid of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars.
-Presumably, this would be due to a bad opacity for a certain line or molecular feature.
-Then, they mask these pixels and repeat the fitting.
-They claim that after two iterations the masks do not change.
-They note that rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures.
-However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.''
+\paragraph{General trends}
+Successful pixel-to-pixel methods that are successful mask spectra lines if they are doing a full spectrum fit (although this brings up the concern of \citet{mga13}).
+Techniques which do not mask or just use a $\chi^2$ function are limited to a small region of the spectrum where presumably their stellar models are \emph{good enough}.
+By limiting themselves to These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.
 
@@
 For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.
-
-The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
-\begin{equation}
-  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
-\end{equation}
-where the best model is defined as one that minimizes the residuals in some sense.
-We will discuss other goodness of fit functions soon.
-Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda)$.
 

### 2014-01-16 paper/paper.tex
updated all the configs

@@
 It is uncertain how statistically justified this uncertainty estimate is.
-While this is a pixel-to-pixel method, it uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match  high resolution models of the Sun and Vega.
+Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match high resolution models of the Sun and Vega.
 
 \paragraph{General trends}
-Successful pixel-to-pixel methods that are successful mask spectra lines if they are doing a full spectrum fit (although this brings up the concern of \citet{mga13}).
-Techniques which do not mask or just use a $\chi^2$ function are limited to a small region of the spectrum where presumably their stellar models are \emph{good enough}.
-By limiting themselves to These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.
+Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors.
+This implicit masking brings up similar concerns to those of \citet{mga13}: how do you know that you are fitting the region of the spectrum that contains the most information about the stellar properties?
 

### 2014-01-19 plotting.py
Edited half the paper

@@ -283,11 +283,112 @@ def plot_line_residuals():
 
     plt.show()
 
+def line_classes():
+    fig, ax = plt.subplots(nrows=2, ncols=4, figsize=(6,4.))
+    #need to set config.yaml['orders'] = [22,23], WASP-14
+    fl23, fl24 = m.fls
+    sig23, sig24 = m.sigmas
+
+    wlsz23 = np.load("residuals/wlsz23.npy")[0]
+    f23 = np.load("residuals/fl23.npy")[0]
+    wlsz24 = np.load("residuals/wlsz24.npy")[0]
+    f24 = np.load("residuals/fl24.npy")[0]
+
+    #import matplotlib
+    #font = {'size' : 8}
+
+    #matplotlib.rc('font', **font)
+    #matplotlib.rc('labelsize', **font)
+    r23 = (fl23 - f23)/sig23
+    r24 = (fl24 - f24)/sig24
+    fl23 /= 2e-13
+    fl24 /= 2e-13
+    f23 /= 2e-13
+    f24 /= 2e-13
+
+
+    ax[0,0].plot(wlsz23, fl23, "b", label="data")
+    ax[0,0].plot(wlsz23, f23, "r", label="model")
+
+    ax[1,0].plot(wlsz23, r23, "g")
+    ax[0,0].set_xlim(5136.4, 5140.4)
+    ax[1,0].set_xlim(5136.4, 5140.4)
+    ax[1,0].set_ylim(-4,4)
+
+    ax[0,1].plot(wlsz23, fl23, "b")
+    ax[0,1].plot(wlsz23, f23, "r")
+    ax[1,1].plot(wlsz23, r23, "g")
+    ax[0,1].set_xlim(5188, 5189.5)
+    ax[1,1].set_xlim(5188, 5189.5)
+    #ax[1,1].set_ylim(-4,4)
+
+    ax[0,2].plot(wlsz24, fl24, "b", label='data')
+    ax[0,2].plot(wlsz24, f24, "r", label='model')
+    ax[0,2].legend(loc="lower center", prop={'size':10})
+
+    ax[1,2].plot(wlsz24, r24, "g")
+    ax[0,2].set_xlim(5258, 5260)
+    ax[1,2].set_xlim(5258, 5260)
+
+    ax[0,3].plot(wlsz24, fl24, "b")
+    ax[0,3].plot(wlsz24, f24, "r")
[…]

### 2014-01-19 paper/paper.tex
Edited half the paper

@@
 
-\section{Introduction}
+\section{Introduction and review of existing methods}
 
@@
 
-\subsection{Similarity techniques}
+\subsection{Classification techniques}
 Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra.
-I am calling them similarity techniques because they compare characteristics of spectra, not actual pixels.
+I am calling them similarity techniques because they focus on determining and recognizing patterns of spectra useful in classification.
 
@@
 This is both a disadvantage and an advantage.
-One one hand by choosing a subset of lines to fit, one may be unwittingly throwing away useful information.
+On one hand, by choosing a subset of lines to fit, one may be unwittingly throwing away useful information.
 On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters.
@@
 
-\subsection{Pixel to pixel techniques using spectral libraries}
+\paragraph{Machine learning}
+Many machine learning techniques exist and classify spectra by comparing patterns between observed spectra and a training set of spectra (which could be real spectra of stars classified by an expert or a synthetic library that has known $\vg$).
+However, my first impression is that these tools are designed to handle large collections of spectra from stellar surveys such as SDSS \citep{sam+10}.
+In our immediate project involving T Tauri stars, we only have a few ($\lesssim 20$) stars that we wish to study in exquisite detail, incorporating specific parameters such as $v \sin i$, extinction, and veiling.
+However, I am a novice when it comes to machine learning so I may be missing something here.
+
+\subsection{Pixel to pixel techniques using synthetic spectral libraries}
+\label{sec:pix}
 Many high quality synthetic spectral grids are now available.
@@
 \paragraph{``Bayesian'' methods} do exist.
-\citet{sdm+07} developed a hierarchical Bayesian approach that purportedly accounts for both statistical and systematic errors in the estimation of stellar parameters.
-I need to work through their math more carefully, but it seems as though they use a hierarchical model for the scatter between ``true'' and observed spectrum, and the unknown scatter in the relationship between $\vg$ and the $\fM$.
-They claim to account for systematics, but it looks like these are systematics errors due to the performance of the instrument, such as a bad flux calibration.
-They use infrared spectra at low resolution ($R \sim $ 1,200), and a weighted $\chi^2$ function for goodness of fit.
-As far as I can ascertain, the actually pixel-to-pixel fitting does not account for systematic errors in regions of the model spectrum.
+\citet{sdm+07} developed a hierarchical Bayesian approach to compare infrared spectra at low resolution ($R \sim $ 1,200) to a synthetic grid in order to determine stellar parameters $\vg$.
+Their model is hierarchical in that it has parameters describing the statistical and systematic uncertainties of the observed spectrum.
+The systematic errors are due to the imperfect performance of the instrument, for example a bad flux calibration.
+They use an ``empirical'' Bayesian model for the noise, meaning that they use data to estimate the prior parameters for the noise \citep{gcs+04}.
+They use a $\chi^2$ likelihood function and allow for ``shrinkage'' of their hierarchical parameters if there is good agreement between observed and synthetic spectra.
+Disappointingly, they do explicitly account for systematic errors in the synthesis of actual lines, for example due to wrong atomic constants.
+One could imagine having a systematic error parameter which describes the scatter in the relationship between $\fM(\lambda | \vg)$ and the ``true'' physical spectrum in the absence of noise $\fD$.
+Nor do they account for any systematic effect that likely exists between $\vg$ and the true physical stellar parameters (as far as $T_{\rm eff}$ has any physical significance) and is a function of which synthetic grid they use\footnote{For example, in our tests with Kurucz vs.
+PHOENIX spectra we often find a systematic offset in temperature of $\gtrsim 200$ K.}.
+However, they do account for interpolation errors by testing interpolating spectra in the pixel space and atmospheric interpolations.
+Then, they select the best model using squared error loss function and a penalty measure for model complexity.
 
@@
 
-\paragraph{SPC} A very popular method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
+\paragraph{SPC} A workhorse method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
 This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method.
@@
 It is uncertain how statistically justified this uncertainty estimate is.
-Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match high resolution models of the Sun and Vega.
+Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values} to match high resolution models of the Sun and Vega.
+
+\paragraph{Other methods} \citet{dj03} do not use a Bayesian model, but they do $\chi^2$ fitting to line profiles in the infrared.
+\citet{bai10} use a forward modelling approach that interpolates a synthetic grid to estimate stellar parameters from low resolution Gaia spectra.
+
+\paragraph{Comparison of synthetic stellar libraries}
+Gaia team members are in a mild panic about the upcoming flood of data.
+\citet{mc07} compare many spectral grids to determine their weaknesses.
+\citep{svt+11} discuss the range of stellar grids available for Gaia.
+\citep{cbm+05} is the Coehlo grid.
+Munari et al 2005, \citep{hwd+13}.
 
 \paragraph{General trends}
-Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors.
-This implicit masking brings up similar concerns to those of \citet{mga13}: how do you know that you are fitting the region of the spectrum that contains the most information about the stellar properties?
+Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors (and effectively masking the rest of the spectrum).
+Cherry picking regions of the spectrum like this evokes similar concerns to those of \citet{mga13}: how do you know that you are fitting the region(s) of the spectrum that deliver the right balance of information about the stellar properties?
 
 \subsection{Line fitting using spectral synthesis}
-This is an improvement over the line index method, since now one includes the information contained in the line profile.
-With quality stellar spectral models, one can identify a few well-modelled lines to fit.
-By choosing a few good lines to fit, one could determine stellar parameters by matching line profiles for a few select lines \citep{dj03}.
-However, this is still subject to the need to identify promising regions of the spectrum.
+Direct synthesis methods take a step closer towards stellar parameter estimation in the correct domain.
+For as large a region of the spectrum as is computationally tractable, these codes synthesize a spectrum for direct comparison with the data.
+Because this is an expensive operation, one must necessarily choose only a few good lines to fit.
 
-MOOG\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73}, SME\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96}, actually synthesize line profiles on the fly, compared to using a library of spectral models.
-This has the advantage of enabling individual tweaks to elemental abundances.
-And interpolation in the model atmosphere space, rather than the spectral line space.
-However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.
-Matching equivalent widths.
-Interestingly, MOOG sythesizes a spectrum on the fly.
+\textbf{MOOG}\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73} and \textbf{SME}\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96} are two popular synthesis codes.
+Given a set of stellar parameters, elemental abundances, and oscillator strengths, the program will interpolate a model atmosphere and then synthesize a spectrum (over a small region of wavelength).
+Then, they use a Levenberg-Marquartd algorithm to converge to the best-fit parameters.
+Their approach has the benefit of enabling individual tweaks to elemental abundances and oscillator strengths, something that is impossible in the aforementioned approaches.
 
-SME allows adjustment of oscillator strengths.
-We could download the synthesis code for SME and synthesize our own spectra on the fly.
-The biggest impediment here would be that the simplest atmospheres will not get certain lines correct.
-The powerful synthesis codes (PHOENIX) will get the atmospheres and lines correct, but they take a long time to run (~hours) per model, and the code is not publicly available.
-To do a full-spectrum fit for these sorts of codes would likely be too computationally difficult.
-However, it would allow access to fitting in the parameter space that we want.
+In order to make the direct synthesis approach computationally tractable, one must use the simplest atmospheric models and radiative transfer.
+Powerful codes like PHOENIX \citep{hwd+13} \emph{should} have more accurate atmospheres and physics, but they take a long time to run ($\sim$ hours) per model, and the code is not publicly available.
 
 \subsection{Stellar parameter estimation in a utopia}
+The difficulties encountered by the similarity techniques and synthetic grid comparisons are a result of fitting in the wrong parameter space.
+Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different.
+The ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$.
+These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote a value relative to solar ($\Z$) and some synthetic grids include alpha enhancements ($\A$)), and atomic constants describing the opacity contribution of a line.
+\emph{All} of these parameters have uncertainties in them, some of them quite dramatic (opacities and line lists are the frequently cited), yet for a synthetic grid these parameters are \emph{fixed}.
+There may still be some missing physics that would be needed to produce a perfect spectrum within the noise (say for example the cores of Ca lines, or dust settling at the latest spectral types), but my naive impression is that the field of stellar spectral synthesis is mature enough that nearly all lines could be properly reproduced given sufficient scrutiny to the atmosphere and atomic constants.
 
-Although the spectral synthesis domain is the correct domain to solve for stellar parameters, the radiation transfer is too slow.
+If we had infinite computational power, a tantalizing approach would be to use a model with stellar parameters $\vg$, atmospheric parameters (such as pressure and temperature profiles), and uncertain atomic constants such as oscillator strengths.
+One could fit for individual elemental abundances and marginalize over the atomic constants.
+This would extract a tremendous amount of information from a high resolution spectrum.
+Unfortunately, the relationship between $\vt$ and the spectrum $\fD(\lambda | \vt)$ is non-linear, meaning that a new spectrum would need to be synthesized for each new $\vt$.
+It may be possible to make an approximate a linear relationship between say an oscillator strength and the line depth as long as the tweak is minor.
+If no reasonable approximation can be found, then the radiative transfer becomes an insurmountable bottleneck.
 
-The fitting difficulties of the previous section are a result of trying to fit data to a model in a smaller parameter space.
-Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different.
-If we think back to first principles, the ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$.
-These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote some value relative to solar, a fraction of models allow changes in alpha abundances), and atomic constants describing the opacity contribution of a line.
-\emph{All} of these parameters have uncertainties in them, some of these are rather dramatic (one continually hears lamentation about the opacities and line lists\dots), yet for any given stellar model these parameters are \emph{fixed}.
-There may yet be unconsidered stellar physics necessary to explain a small fraction of lines (say for example the cores of Ca lines), but the field of stellar modelling has progressed to the state that I would have the confidence to say that if a group focused all of their attention on a single line, and using only their current physical models and making tweaks to the afore-mentioned lines, they could produce a satisfactory model.
-
-In reality, this has been done using the Kurucz models in an unpublished grid of stellar spectra, over the range $5000 - 5300$ \AA.
-
-The real problem is much more rich, as is the amount of information one could extract from a high resolution optical spectrum.
-One would want to fit for individual abundances, marginalize over oscillator strengths.
-
-The unfortunate reality is that sampling in the stellar model space for the full optical spectrum is not yet a computational reality.
-There do exist codes (MOOG) that can do this approach but only for a small ($\sim$ a few lines) at once, and using simplifications.
-Through collaboration with a stellar atmospheric expert, such an approach may soon be a possibility.
-If the computation of a given model could be reduced to less than a minute, efficient MCMC sampling parallelized on a cluster (using \texttt{emcee}, for example).
-However, as it stands without a collaborator, I do not have the expertise nor the time to pursue such an approach.
-Therefore, we are reduced to fitting in the space of $\vg$ and devising ways to deal with the issues.
+If the approximations could reduce the computation of synthetic spectrum, efficient MCMC sampling parallelized on a cluster might put this utopian vision within reach in the near future.
+However, without a collaborator with expert knowledge of spectral synthesis, I do not have the expertise to take such an approach.
+Many high quality synthetic spectral grids do exist, and so our current efforts are focused on addressing the shortfalls of the previously mentioned pixel-to-pixel methods.
 
@@
 
-Pixel-to-pixel approaches have the potential to be much more powerful than other historical stellar parameter estimation techniques.
-When comparing directly in the data space, we have the ability to add in other physics as we see fit.
+Forward modelling approaches (pixel-to-pixel) have the potential to be more powerful than other stellar parameter estimation techniques.
+When we model directly to the data space, we have the ability to account for instrumental and systematic effects, allowing us to derive a robust estimate of our uncertainty by marginalizing over the nuisance parameters.
+We can also derive more complicated models to fit unresolved binary or multiple star systems.
+In fact, many of the most interesting systems for which we might want to determine stellar parameters are in multiple systems since we can dynamically determine masses for these systems and thus compare to evolutionary models.
 
-because we can fit multiple stars at once (say for unresolved binary or multiple systems), fit for velocity components, etc.
-Many of the most interesting systems for which we would want to determine stellar parameters are in multiple systems anyway, since we can dynamically determine masses for these systems and thus compare to evolutionary models.
-Determine veiling.
-And a perhaps wooden-headed belief that using the full spectrum \emph{should} give us a more accurate and less-biased posterior estimate.
-Such a belief may be foolish and in reality different parts of the spectrum could give completely different parameter estimates.
+As discussed in \S\ref{sec:pix}, many forward modelling approaches exist, yet nearly all use a $\chi^2$ likelihood function and most employ some sort of masking scheme.
+The more successful pixel-to-pixel methods use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000) or a grid over a narrow range of wavelength space with finely tuned atomic constants, eschewing difficult decisions about what to mask.
+\textbf{The true fundamental problem with a pixel-to-pixel comparison is that the stellar spectra are not a perfect representation of reality}.
+This may be an obvious statement, but systematic errors in the synthetic spectra means that the previously valid assumption of Poisson errors and a $\chi^2$ likelihood function breaks down.
+In order to derive unbiased parameter estimates, systematic errors must be identified and properly handled.
 
-Many, many methods exist for comparison of stellar spectra, including many ``Bayesian'' approaches in the pixel-to-pixel space.
-The nature of the Bayesian-ness, however, is usually limited to priors on temperature, surface gravity, and metallicity.
-These approaches do not explicitly state the likelihood function that they use, or if they do, it is a $\chi^2$ function.
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics{lineclasses}
+\caption{The four classes of line behavior, from left to right: \textbf{Class 0}: the model behaves as expected.
+\textbf{Class I}: a spectral line is present in the data but missing from the model (in this case, the line is also blended).
+\textbf{Class II}: a spectral line exists in the model where there is only continuum in the data.
+\textbf{Class III}: lines exist in both the data and model, however they are of the wrong strength.
+Some lines are too strong while others are too weak.
+The lower panels show the residual spectra in units of the observational uncertainty (Poisson uncertainty only).}
+\label{fig:lineclasses}
+\end{center}
+\end{figure}
 
-The more successful of these models use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000 ish). Or, a grid over a narrow range of wavelength space with finely tuned atomic constants.
+\subsection{Classes of synthetic spectral line behavior}
+From our forward modelling experiments using the modified Kurucz grid used in \citet{blj+12} and the recent PHOENIX grid \citep{hwd+13}, we have identified ways that synthetic spectra diverge from the our observed data.
+To start, we will describe these classes of behaviour as being tied to specific spectral lines present against a well-defined continuum.
+Later on, we will use the classes to describe features more generally to cover the behavior of molecular features in later type spectra.
+See Figure~\ref{fig:lineclasses} for visual examples of each class of behavior.
 
-\textbf{The true fundamental problem with this approach is that the stellar spectra are not a perfect representation of reality}.
-Assumptions about Poisson errors break down, and fits become contaminated by systematics.
-There are types of errors which occur.
-
-FIGURE: 4 Panel Figure, from left to right, showing Class 0 through III errors.
-Use Mg b lines, other big line near Mg b triplet, order 24 line, and then Fe I/II line complex that doesn't fit properly.
+\paragraph{Class 0} These lines actually produce a good fit with a residual nearly consistent with the Poisson counting uncertainty (see \S\ref{sec:TRES} for how the data is processed).
+If every region of the spectrum behaved like this, we might be justified in using a $\chi^2$ likelihood function and calling it a day.
 
@@
 This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident.
-Presumably this is a product of bad atmospheric structure (temperature or pressure is wrong at some point?) that excites a line when it shouldn't otherwise.
+Presumably this might be a combination of incorrect atomic data coupled with incorrect atmospheric structure that incorrectly excites a line.
 
-\paragraph{Traditional ways to model Class I and Class II errors} are a form of sigma clipping.
-Mann et al.
-compares the stars to models over a wide range of parameters.
-One can track how these lines might be bad because Class I residuals (data - model)/sigma will always be negative (since we are dealing with absorption lines only) and not change significantly with $\vg$, since we the model can only produce continuum.
-Class II residuals will always be positive and should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.
+The traditional way to account for Class I and Class II errors is to sigma clip the bad lines in the spectrum \citep{kpb+09, mga13}.
+A region of the spectrum with Class I or II errors is useless for determining spectral parameters.
+While sigma-clipping routines are to be avoided, if they can correctly mask these regions, then at least they are moving in the right direction.
 
-If a Class I or II error occurs on a line, that region of the spectrum is relatively useless for determining spectral parameters.
-Authors like Mann et al.
-or Kholeva et al.
-find these regions and simply mask them out of the final computation.
-Kholeva et al., and others approach in an iterative manner.
-Kholeva et al.
-approaches are more sophisticated because they track convergence.
-This is a sigma clipping approach, and while we could go and model these a la Hogg 2010, the approach is not super justified, result is probably not too wrong (similar to actual sigma clipping).
-Given that Kholeva et al.
-track convergence maps, I would place a higher degree of confidence in the answer.
+There is additional information available to help determine which regions of the spectrum suffer from Class I or II errors.For a normal star with absorption lines against a continuum, Class I residuals will always be negative and not change significantly with $\vg$, since we the model can only produce continuum in that region.
+Class II residuals will always be positive and their magnitude should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.
+We could easily store the partial derivatives of an individual synthetic spectrum with respect to $\vg$
+\begin{equation}
+  \bigl \{ \frac{\partial \fM(\lambda | \vg)}{ \partial T_{\rm eff}}, \frac{\partial \fM(\lambda | \vg)}{ \partial \log(g)}, \frac{\partial \fM(\lambda | \vg)}{ \partial \Z}, \frac{\partial \fM(\lambda | \vg)}{ \partial \A} \bigr \}
+  \label{eqn:derivative}
+\end{equation}
+and coupled with a prior on the range of parameters over which the residuals will be correlated, we could shore up the classification of Class I and II errors.
 
-In our later approaches in Sec XX, how might we put a prior on $\Delta \vg$, the range of parameters over which the residuals will be correlated?
-Maybe we could Gibbs sample in the same way as the bad flags.
-We'd have to keep a bunch of spectra with varying $\vg$ in memory to make this comparison easy.
+\paragraph{Class III} errors occur when the line strengths between the data and model do not match at a $\vg$ that fits a majority of the other lines.
+What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, though this parameter combination would result in bad fits for the majority of the other lines.
+Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$ and have predictable behaviors with change in $\vg$.
+Class III errors, however, come in both directions: the line can be too strong or too weak.
+Identifying Class III errors requires making a choice about which lines to trust and how much.
 
-\paragraph{Class III} In truth, this is just a superset of both Class I and Class II.
-Class III errors are when the line strengths between the data and model do not match at a $\vg$ that fits a majority of lines.
-What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, however this parameter combination would result in bad fits for the majority of the other lines.
-Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$.
+Most of the Class III errors can probably be identified by finding the regions of the spectrum that have the highest residuals after dealing with Class I and II errors, although some mildly incorrect Class III lines will be difficult to distinguish from correct (Class 0) lines that are simply fit with a less probable $\vg$.
+Once identified, properly handling Class III errors requires careful thought because these regions of the spectrum \emph{still contain information about $\vg$}.
+If we take an masking approach as with Class I and II errors, we will be throwing it away.
+It might even be the case that Class III regions contain \emph{the most} information about $\vg$, since these ``delicate regions'' are the most sensitive to changes in $\vg$ at that particular $\vg$ and thus are tricky to tune correctly.
+In other words, Class III regions might also be regions where the spectrum derivative (Eqn~\ref{eqn:derivative}) is largest.
 
-In addition, it is difficult to determine the ensemble of lines that suffer from Class III errors, since the strength of some lines will be under-predicted while others will be over-predicted.
-How do we know which lines to trust?
-If we take an approach similar to fixing Class I and II errors, and mask out bad regions, it would be very difficult to distinguish Class III errors from \emph{truthful} lines at an incorrect $\vg$.
-For example, if one imagines using a cleaning algorithm as described for Class I and Class II errors, the identification of Class III errors will be sensitive to the starting guesses of $\vg$.
-
-Because these Class III lines can still produce a significant residual spike, they tend to distort and in some cases artificially peak the posterior distribution.
-There is no one parameter combination that will fit all lines perfectly.
-Certainly there is a parameter combination that will produce a reasonable fit to 90\% of the lines to within 3 sigma residuals, but there will still be a fair number of both ``high'' and ``low'' regions.
-These lines need to be identified and dealt with, but how?
-
-How then, should the best fit be determined? Surely some of the remaining lines are ``bad.'' But in this regime, determination of which lines are bad will bias the results in a certain direction.
-
-Also, which lines should one actually fit? The stronger lines? Did the modelers themselves give more attention to the atomic constants which produce the strong lines?
-
-To summarize, the fundamental problem with comparing high resolution optical stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines.
-Deciding how to model the lines is then the difficult part.
-Different groups have sigma-clipped lines that are obviously bad, and this does improve the procedure dramatically.
-In such a case, the procedure of sigma-clipping doesn't gall with me too badly.
-However, the more difficult problem is dealing with Class III errors properly, something that no modeler has mentioned or successfully implemented in a statistically justified manner.
-
-In a spectrum fit, many lines do reasonably fit within the Poisson errors. For consistency, we will denote these as ``Class 0'' lines, where there is no error.
-
-\subsection{Testing the accuracy of stellar libraries}
-Gaia team members are panicking about the upcoming flood of data.
-
-\citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia.
-
-\citep{cbm+05} is the Coehlo grid.
+To summarize, the fundamental problem with comparing high resolution stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines properly.
+One can either limit their fits to regions of the spectrum that they know their spectral libraries to be accurate (implicitly masking the rest and potentially discarding a large amount of information) or suck it up and forward model the regions where the models perform poorly.
 
 \section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}
+\label{sec:perfect}
+While this document aims to be general in scope, any forward-modelling approach must also discuss the characteristics of the instrument that performed the observations.
+We are currently modelling data obtained with the \emph{TRES}\footnote{\url{http://tdc-www.harvard.edu/instruments/tres/}} echelle spectrograph, and so we will describe the characteristics of the instrument and the calibration approaches we have taken, although we note that our approach is easily generalized for data obtained with other spectrographs.
 
-The parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$.
-Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum.
-For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.
+\emph{TRES} is a 51-order echelle spectrograph on the 1.5-m Tillinghast reflector at Whipple Observatory.
+It spans the full optical range ($3800-9000$\AA) at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$).
+Echelle spectrographs impart a strong blaze function upon each order, where transmission is largest in the center of the order and declines rapidly as one moves towards the edge of the order.
+The \emph{TRES} pipeline extracts orders such that they span a wavelength region until the transmission reaches $\lesssim$ 25\% of the central transmission in that order.
+Each order spans $\sim 100$\AA\ and the blue orders overlap while there are gaps between the wavelength coverage of the red orders.
+This spectra are blaze-corrected by dividing by a blaze function determined from fitting nightly exposures of flat lamps.
+Then, the spectrum is flux calibrated using sensitivity functions determined from nightly observations of spectrophotometric standards (O stars).
+The flux-calibrated spectrum is in units of ergs/s/${\rm cm}^2$/\AA.
 
-FIGURE showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000.
+\paragraph{Processing TRES Data}
+\label{sec:TRES}
+Because the \emph{TRES} reduction pipeline does not produce a ``sigma spectrum'' like that determined by a standard IRAF reduction, we instead calculate the Poisson uncertainty on each pixel.
+We can find the signal-to-noise ratio of each pixel using the raw (still-blazed) spectrum, which is in units of ``counts.''
+\begin{equation}
+  [S/N](\lambda) = \sqrt{ {\rm cts}(\lambda)}
+\end{equation}
+the Noise-to-Signal ratio is the inverse 
+\begin{equation}
+  [N/S](\lambda) = \frac{1}{\sqrt{ {\rm cts}(\lambda)}}
+\end{equation}
+To determine the uncertainty of the flux-calibrated spectrum, we multiply by N/S.
+The $[N/S](\lambda)$ ``spectrum'' is larger at the edges of each order due to the reduced transmission.
+Please note that statistic pixel uncertainties derived in this manner neglect regions of the spectrum that might have a higher variance due to sitting atop a night sky line, but short of implementing our own reduction pipeline for \emph{TRES}, there isn't much more we can do at the moment.
+
+\paragraph{Generating a model spectrum} We parameterize the fit to a star by $\vt = \{\vstar, \vN \}$, where stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), \Z, \A, v \sin i, v_z, A_v, R_\star^2/d^2 \}$.
+The data vector $\vD = \fD(\lambda)$ represents the \emph{TRES} spectrum.
+To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by first linearly interpolating in $T_{\rm eff}, \log(g), \Z, \A$ space between grid points of instrumentally broadened and oversampled synthetic stellar spectra.
+Then, the spectrum is rotationally broadened for $v \sin i$ using Fourier techniques, Doppler shifted by $v_z$, extinction corrected for $A_V$, multiplied by $R_\star^2/d^2$ for a solid angle correction, and finally downsampled to the exact pixels of the TRES spectrum using spline interpolation\footnote{Because we are dealing with flux-calibrated data and synthetic spectra, we \emph{resample} to lower resolution.
+We do not \emph{rebin} because we are dealing with flux-density, not photon counts}.
+
+FIGURE: showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000.
 This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species.
-For example, we could inject bad lines in this space.
-
-\subsection{Processing TRES Data}
-We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel.
-By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
-\begin{equation}
-  S/N = \sqrt{ {\rm cts}}
-\end{equation}
-then the Noise-to-Signal ratio is the inverse of this
-\begin{equation}
-  N/S = \frac{1}{\sqrt{ {\rm cts}}}
-\end{equation}
-then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.
+For example, we could inject bad lines in this high resolution space.
 
 \subsection{Applied to a Gaussian likelihood function}
-
+For the rest of \S\ref{sec:perfect}, we proceed as if there were no Class I, II, or III errors present in our model so that we may illustrate our technique for dealing with uncertainties in the flux-calibration.
 Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}.
@@
 \end{equation}
-To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating in $T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}]$ space between grid points of instrumentally broadened, oversampled PHOENIX model stellar spectra, rotationally broadened using Fourier techniques, Doppler shifted, extinction corrected, and downsampled using spline interpolation to the exact pixels of the TRES spectrum.
 
-$\sigma(\lambda)$ is the Poisson photon counting errors.
-For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
-For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.
-
-The prefactor $k(\lambda | \vN)$ is a calibration functionm that aims to account for errors in the blaze-correction or flux-calibration by multiplying the model by a polynomial.
+The prefactor $k(\lambda | \vN)$ is a calibration function that aims to account for errors in the blaze-correction or flux-calibration by multiplying the synthetic spectrum by a polynomial.
 The parameters describing this function are the coefficients of the polynomial.
 This approach of accounting for large-scale calibration errors in the spectrum is more justified than the traditional approach of normalizing to the stellar continuum.
-From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (Figure?) flux-calibration function, or IRAF's ``sensitivity function.'' These tests revealed that the largest source of error in flux-calibration was actually in the overall magnitude of individual order, this could be due to slit losses, poor correction due to parallactic angle, or scattered light within the instrument.
+From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (FIGURE showing the sensitivity function for a few representative orders) flux-calibration function, or IRAF's ``sensitivity function.'' These tests revealed that the largest source of error in flux-calibration was actually in the overall magnitude of individual order, this could be due to slit losses, poor correction due to parallactic angle, or scattered light within the instrument.
 These corrections are smaller than 10\%.
@@
 As described in \S\ref{sec:priors}, we can set priors on these coefficients.
-This will be the optimal mode of operation for our model.
-However, there exists a wealth of archival data that is un-fluxcalibrated, since its primary purpose was for radial velocity determination of FGK stars, to which it is easy to normalize to a continuum.
-In this case, the scatter amongst the calibration can be as large as a factor of 2x.
 This prior information on the calibration parameters motivates our choice of functional parameterization for the nuisance parameters in \S\ref{sec:nuisance}.
+The polynomial approach is also better than the normalization approach if the order only spans a small region of the spectrum.
+If a M star absorption band spanned the whole order, one might accidentally normalize to the feature rather than a pseudo-continuum.
 
@@
 This enables a dramatic speedup in the MCMC sampling.
-For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than XX\%), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters.
+For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than how much?), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters.
 This approach is detailed in \S\ref{sec:gaussian_simplification}.
@@
 \label{sec:priors}
-Using this formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters.
+Using this polynomial formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters.
 From our flux-calibration experiment, we determined that the overall scaling could range by 10\%, while the higher order corrections were less than 3\%.
@@
 \subsection{Sampling a un-flux calibrated model}
+There also exists a wealth of archival \emph{TRES} data that is not flux calibrated, since its primary purpose was for radial velocity determination of FGK stars.
+To fit this data we can use outdated sensitivity functions and soften the priors on the coefficients.
+
 This invalidates the previous section on normalization. However, for corrections of less than XX\%, this approximation is better than XX\% accurate.
@@
 
-Could this correlation be described by a Gaussian random process?
+Could this correlation be described by a Gaussian random process? \citep{rw05}.
 
@@
 This approach is no more justified than a by-eye spectral classification.
+
+\subsection{Hyperparameters}
+Difficulty of combining different data sets with different errors \citep{mb13}. Or a hyperparameter like \citep{sdm+07}. Shrinkage similar to \citet{kru10}.
+
+\citet{sdm+07} use ISO SWS infrared spectra.
+Hierarchical approach.
+Able to synthesize their own spectra.
+Use systematic error flags from the instrument, as well as statistical flags.
+Hierarchical model that uses shrinkage between the observed spectrum and the synthetic spectrum.
+They do a fit to the variance with a hierarchical linear mixed model (sec 5.2.2, and their figure 7) using design matricies and regression coefficients.
+This is probably to incorporate some sort of correlation length on the systematic errors, and ``smooth'' the function, as they say.
+I was having some difficulty understanding it, but it seems as though they were keeping the noise components fixed, and they did not have a contribution to the error budget from regions where the spectral synthesis \emph{itself} contained systematic errors.
+Perhaps these were masked by the systematic error from the instrument.
 
@@
 \acknowledgments
-IC would like to graciously acknowledge Gregory Green, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy.
+IC would like to graciously acknowledge Gregory Green, Kaisey Mandel, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy.
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{art+13}.
@@
 
-\bibliography{disks,master}
+\bibliography{disks,bayesian,master}
 \bibliographystyle{hapj}

### 2014-01-19 PHOENIX_tools.py

@@ -19,7 +19,7 @@ from functools import partial
 
 c_kms = 2.99792458e5 #km s^-1
 c_ang = 2.99792458e18 #A s^-1
-wl_file = pf.open("PHOENIX/HiResFITS/WAVE_PHOENIX-ACES-AGSS-COND-2011.fits")
+wl_file = pf.open("WAVE_PHOENIX-ACES-AGSS-COND-2011.fits")
 w_full = wl_file[0].data
 wl_file.close()
 ind = (w_full > 3000.) & (w_full < 13000.) #this corresponds to some extra space around the

### 2014-01-20 paper/paper.tex
Major paper update

@@
 \newcommand{\qN}{\{q_i\}_{i=1}^N}
+\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
 \newcommand{\yN}{\{y_i\}_{i=1}^N}
@@
 A salient example is that most newly discovered exoplanets must necessarily their mass and radius measured relative to the properties of their host stars (e.g.
-\citealt{tfs+12,blj+12,ssm+13}).
+\citealt{tfs+12,blj+12,ssm+13}) and uncertain stellar parameters translate into uncertain planet properties \citep{kan14}.
 The properties of stars are of course also important in their own right because they inform us about our ideas of stellar evolution \citep{dm97, bca+02}.
@@
 They use an ``empirical'' Bayesian model for the noise, meaning that they use data to estimate the prior parameters for the noise \citep{gcs+04}.
+They do a fit to the variance with a hierarchical linear mixed model (sec 5.2.2, and their figure 7) using design matricies and regression coefficients.
+This is probably to incorporate some sort of correlation length on the systematic errors, and ``smooth'' the function, as they say.
 They use a $\chi^2$ likelihood function and allow for ``shrinkage'' of their hierarchical parameters if there is good agreement between observed and synthetic spectra.
@@
 \end{equation}
-We can rewrite Equation~\ref{eqn:expanded} quadratically in $\vc$ as
+We can rewrite Equation~\ref{eqn:expanded} quthat should be accounted for? adratically in $\vc$ as
 \begin{equation}
@@
 }
+\label{eqn:lnprob_lognormal}
 \end{equation}
@@
 }
+\label{eqn:lnprob_lognormal_marg}
 \end{equation}
@@
 
+If one uses a marginalized posterior, but is still interested in obtaining a flux-corrected spectrum for the purposes of inspecting the residuals, it is still possible to do so.
+We can use Eqn~\ref{eqn:lnprob_lognormal} with $\vstar$ fixed at a sample from the posterior of Eqn~\ref{eqn:lnprob_lognormal_marg} and then use MCMC to draw samples from $p(\vN | \vstar, \vD)$.
+Now, the combination of $\vstar$ and $\vN$ completely describes a flux-corrected spectrum.
+
 \subsection{Sampling a un-flux calibrated model}
 There also exists a wealth of archival \emph{TRES} data that is not flux calibrated, since its primary purpose was for radial velocity determination of FGK stars.
-To fit this data we can use outdated sensitivity functions and soften the priors on the coefficients.
+To fit this data we can use newer sensitivity functions taken under different conditions and soften the priors on the coefficients.
+Tests indicate that this works rather well.
+The blaze function will might change significantly between nights and years, so this might be a source of error.
 
-This invalidates the previous section on normalization. However, for corrections of less than XX\%, this approximation is better than XX\% accurate.
-
-For the flux-calibration exercise, the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$.
-
-The plot in plots/priortests shows that the log-normal prior is the correct choice.
-Since we care most about the inference on F, we want to know that the choice on prior of $c_0$ is not biasing the final outcome, while a Gaussian prior might if $\sigma_{c_0}$ was large.
-How to quantify this error and create a dividing line?
-For 5\% error?
-In peak?
-Do these plots need to be in the paper?
-
-Compare distributions by doing a K-S test on the cumulative distributions between Gaussian and Log-normal?
+For non flux calibrated spectra, we determined that the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$.
+The plot in plots/priortests shows that log-normal prior (vs.
+Gaussian) on $c_0$ is the correct choice .
+Since we care most about the inference on F, we want to know that this prior is not biasing the final outcome.
+Using a Gaussian prior with a large $\sigma_{c_0}$ will bias the results, but if $\sigma_{c_0}$ is small ($\lesssim 0.2$), then we can get away with Gaussian priors (see \S\ref{sec:gaussian_simplification}).
+Is there a better way to determine these thresholds other than my current technique of testing and visual inspection?
 
 \subsection{Hierarchical modelling}
-Can have a $c_1$, etc, that are marginalized over but we sample $c_0$ in a hierarchical manner which informs the overall flux-level.
+Instead of the current framework where the nuisance coefficients of each order are independent, we could derive a hierarchical framework where each $c_0$ is drawn from a probability distribution of $c_0$ described by a hyperparameter $C_0$.
+If each order required a similar $c_0$, then this would enable shrinkage of $C_0$, however if each order required a widely disparate $c_0$, then $C_0$ would correspondingly increase.
+Higher order coefficients could still be analytically marginalized over.
 
-\subsection{Sampling with emcee}
-Citation to DFM \citep{fhl+12}, GW \citep{gw10}, and CosmoHAMMER \citep{asa+13}.
+\subsection{Sampling the posterior and performance of the $\chi^2$ likelihood}
+Currently we are sampling this posterior with the MCMC ensemble sampler \citep{gw10} implemented in \texttt{Python} by \citet{fhl+12} as \texttt{emcee}.
+See \citet[CosmoHAMMER]{asa+13} for a review of \texttt{emcee} parallel performance and scaling relationships on a computer cluster.
 
-\section{Moving beyond a perfect stellar model}
+In this section, we described our approach to take into account flux calibration errors assuming a perfect synthetic spectra.
+If one does a straight up $\chi^2$ fit to the data, however, you can get posteriors that look have 1-sigma contours of $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
+This would be as precise as stellar parameters obtained with astroseismology!
+Clearly there is something going wrong here.
+
+As our previous discussion has noted, these posteriors are unfortunately too good to be true and we must take a more sophisticated approach.
+However, many of the flux-calibration principles developed in this section should still apply.
+More advanced posterior functions in the following sections might require using more specialized MCMC techniques such as Gibbs sampling or expectation maximization (EM).
+
+\section{Baby steps beyond a perfect stellar model}
 \label{sec:residuals}
+In this section we describe a few simple things we might try (or have tried) in order to remedy the shortcomings of the previous section.
 
-\subsection{Additional sources of noise beyond Poisson}
-Presumably we shouldn't have to add an additional source of ``noise'' to the fit, because our nuisance parameters will find an exact flux calibration.
+\subsection{Additional sources of noise}
+Presumably we shouldn't have to add an additional source of instrumental ``noise'' to the fit, because our nuisance parameters will find an exact flux calibration.
+However, CCD readnoise and 1D spectrum extraction will almost definitely increase the noise in a pixel beyond theoretical Poisson limits.
+Normally this information will be encapsulated in a ``sigma spectrum,'' but unfortunately we do not have access to this unless we develop an alternative to the \emph{TRES} pipeline.
+We could try to fudge adjustments to the readout noise by adding a small error in quadrature.
 
-Interpolation errors. Quantified via Husser et al. Ph.D. thesis as well as custom tests.
+Even the stellar continuum can flicker/jitter (as shown by Kepler studies), which would produce a small departure from a perfect continuum.
+Even in the Class 0 region of our spectrum (Figure~\ref{fig:lineclasses}, left panel), the residuals are slightly larger than 1 $\sigma$ on average.
 
-\subsection{Noise, Masking, Residuals, and Alternative likelihood functions}
-In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment.
+When we interpolate a synthetic spectrum from the library to a specific $\vg$, we will introduce interpolation errors depending on how close we are to a grid point in the library.
+\citet{hus12} investigate the possible range of interpolation errors at specific $\vg$ by dropping out a spectrum from the grid, interpolating over it, and then comparing the true spectrum to the interpolated spectra.
+For hotter stars with well defined continua, we find the maximum error is $\lesssim 5$\%, while for M stars the error could be as large as $\sim 20$\% for some spectral features.
+By doing this interpolation test across the grid, we could determine an ``interpolation error spectrum'' that would encapsulate the maximum error possible for an interpolated spectrum, weighted by its distance from the nearest grid points.
+This uncertainty could be included in quadrature.
+Since most of our \emph{TRES} spectra are at signal/noise of $\sim 30$ per pixel (or noise/signal ratio of $1/30 \approx 0.03$), an interpolation noise contribution of $\sim 5$\% the level of the signal level would be significant.
+However, we tried arbitrarily increasing the Poisson errors by a factor of 2 or 3 and it didn't significantly change the results.
+While such an error spectrum should eventually be incorporated, we have bigger fish to fry.
 
 \subsection{Masking}
-
 We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).
 We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid.
+We also mask blatant Class I and II regions by eye, but currently do nothing for smaller errors or Class III errors.
 
-Section on how to choose masked regions, testing the PHOENIX grid to see where the derivative is highest and thus gives the best leverage on stellar parameters.
+\subsection{Gallery of likelihood functions}
+A problem with the $\chi^2$ likelihood function is that the Gaussian tails of the probability function rapidly decline, making any outliers extremely improbable, which will bias a fit.
 
-Iterative masking to deal with Class I and II errors a la ULySS. They use a $\kappa - \sigma$ routine, where they iteratively sigma clip (Website link).
+A common approach to make ``robust inference'' is to use a likelihood function with fatter tails.
+Although not rigorously motivated, this is can be a quick thing to try.
+We explore the following likelihood functions.
 
-\section{Gallery of likelihood functions}
+\paragraph{Combination of Gaussian with exponential wings}
+We fit a function of the form
+\begin{equation}
+  p(\fD(\lambda) | \fM(\lambda | \vt)) \propto \exp \left (\frac{(\fD - \fM)^2}{(\sigma_G \sigma_P)^2} \right ) + A \exp \left (\frac{|\fD - \fM|}{\sigma_E \sigma_P}  \right )
+\end{equation}
+to the residuals of a reasonable $\vt$ to determine $A$, $\sigma_G$, and $\sigma_E$.
+This functional form assumes that there will be a Gaussian core, similar to the $\chi^2$ likelihood, but also allows some contribution from an exponential tail, used to describe the residuals.
+I think the problem here is that the exponential is too steep to properly describe Class I and II residuals, although it might do well for Class III errors.
 
-If you do a straight up $\chi^2$ fit to the data, you can get posteriors that look like $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
-This would be better than that obtained by astroseismology!
-Clearly there is something going wrong here.
+\paragraph{Sivia's model}
+\citet{ss06} describe a likelihood function which treats the noise estimate for each data point suspiciously.
+The true noise $\sigma$ can be no smaller than some original estimate $\sigma_0$, but it may be much larger.
+By putting a prior on $\sigma$,
+\begin{equation}
+  p(\sigma | \sigma_0) = \left \{ \begin{array}{cc}
+    \sigma_0/\sigma^2 & \sigma  \leq \sigma_0\\
+    0                 & \textrm{else}
+  \end{array}
+    \right.
+\end{equation}
+and marginalizing over the $\chi^2$ likelihood, the new likelihood function becomes
+\begin{equation}
+  p(\fD | \fM, \sigma_0) = \frac{1}{\sigma_0 \sqrt{2 \pi}} \left [ \frac{1 - e^{-R^2/2}}{R^2} \right ] 
+\end{equation}
+where $R = (\fD - \fM)/\sigma_0$.
+This function has even fatter tails than the exponential model and seems to give more believable results.
+We have done the most testing using this new likelihood function, but it still does not give a truly satisfying result.
+For example, which lines contributed most to the posterior?
+What about the wings of class I and II errors?
+These individual pixels might still be given a weight of $\sigma_0$ because they are not nearly as bad as the line centers.
 
-Lorentzian, linear combination of Gaussians, Sivia w/ prior on sigma. $\chi^2_Q$ model.
+\paragraph{$\chi^2$ with flat tails}
+\citet{hbl10} describe a likelihood function that smoothly caps the total likelihood contribution any outlier might have at some constant $Q$ by using a $\chi^2_Q$ function
+\begin{equation}
+  \chi^2_Q = \sum_{i = 1}^N \frac{Q^2 [\fD - \fM]^2}{Q^2 \sigma_P^2 + [\fD - \fM]^2}
+\end{equation}
+In some ways, this is analogous to sigma-clipping. Once a residual reaches the flat tails, it fails to factor into the determination of parameters.
 
-These likelihood functions seem to be an improvement over the straight $\chi^2$ function, in the sense that they give parameter estimates that are
+Other functions we might try include a Student-t statistic, a Lorentzian, or a linear combination of Gaussians.
+While all of these likelihood function shenanigans are amusing, we would like to move beyond tweaking the fitting function towards an approach that properly incorporates knowledge about the specifics of our problem.
 
 \subsection{Bad data model}
-The bad data model \citep{pre97,dbl10} is a powerful framework for forward-modelling of data, allowing the modeler to account for the probability of outliers.
+The bad data model \citep{pre97,hbl10} is a powerful framework for forward modelling data, which allows the modeler to explicitly account for outliers using a generative model.
 
-Now, we have a $q_i$ for each data point, such that for the entire data set we have an ensemble of $\qN$.
+In the simplest mixture model, each data point has a binary flag $q_i$ which specifies whether that data point is good or bad.
+Then, for the entire data set we have an ensemble of $\qN$.
 If this ensemble of $\qN$ is chosen appropriately, and the good data points are flagged as good and the bad data points are flagged as bad, the likelihood will be maximized.
+We adopt $P_B$ for the \emph{a priori} probability that a given data point will be bad, $Y_b$ for the mean of the bad data distribution, and $V_b$ for its variance.
 Written out explicitly, this is
 \begin{multline}
-  {\cal L} = p({\bm D} | m,b,\qN,Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma^2} \right] \right ]^{q_i}  \times \\
-  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma_{y_i}^2)}} \exp \left[ - \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{y_i}^2)} \right]
+  {\cal L} = p({\bm D} | \vt,\qN, Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma} \exp \left[ - \frac{(\fD - \fM)^2}{2 \sigma^2} \right] \right ]^{q_i}  \times \\
+  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma^2)}} \exp \left[ - \frac{(\fD - Y_b)^2}{2 (V_b + \sigma^2)} \right]
   \right ]^{1 - q_i}
 \end{multline}
-
-using a Bernoulli prior
-
+We we a Bernoulli prior on the $q_i$
 \begin{equation}
@@
 \end{equation}
-
+then, the probability of any given ensemble of $\qN$ is 
 \begin{equation}
@@
 
-For this data point, we can integrate (in this case, discrete sum) over the two states of $q_i$: $q_i = 1$ and $q_i = 0$. 
+We can analytically marginalize over the single flag by a discrete sum over the two states of $q_i$: $q_i = 1$ and $q_i = 0$. 
 \begin{multline}
-  p(m, b, P_b, Y_b, V_b | D_i)  =  \int p( m, b, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
-  = p(m, b, q_i=1, P_b, Y_b, V_b | D_i) + p(m, b, q_i=0, P_b, Y_b, V_b | D_i) \\
-  = \Bigl[ p_{\rm good}(D_i | m,b)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(m,b,P_b, Y_b,V_b) 
+  p(\vt, P_b, Y_b, V_b | D_i)  =  \int p( \vt, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
+  = p(\vt, q_i=1, P_b, Y_b, V_b | D_i) + p(\vt, q_i=0, P_b, Y_b, V_b | D_i) \\
+  = \Bigl[ p_{\rm good}(D_i | \vt)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(\vt ,P_b, Y_b,V_b) 
 \end{multline}
@@
 \begin{multline}
-  p(m,b,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i | m, b)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(m,b,P_b, Y_b, V_b)\\
+  p(\vt,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i |\vt)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(\vt,P_b, Y_b, V_b)\\
   \boxed{
-    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_{yi}^2}} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma_{yi}^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma_{yi}^2)}} \exp \left[- \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{yi}^2)} \right] \right\}
-  \times p(m,b,P_b,Y_b,V_b)}
+    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_{yi}^2}} \exp \left[ - \frac{(\fD - \fM)^2}{2 \sigma^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma^2)}} \exp \left[- \frac{(\fD - Y_b)^2}{2 (V_b + \sigma^2)} \right] \right\}
+  \times p(\vt,P_b,Y_b,V_b)}
 \end{multline}
 
-In it's simplest form (Hogg 2010), the bad data model results in a linear combination of Gaussians.
-
-A benefit of the bad data model over a chosen likelihood function is that it is motivated.
-It may still result in the linear combination of Gaussians (nice), but preserves real parameter correlations.
-You are not forcing the same mean.
+In its simplest form, the bad data model results in a likelihood function that is a linear combination of Gaussians.
+Although the functional form you end up with is similar, the benefit of this bad data approach over a gallery of likelihood functions is that it is motivated by the characteristics of the bad data ($P_B$, $Y_b$, and $V_b$).
+If we wish to add additional information about the characteristics of the bad data, we can do so.
 
@@
 We can do more.
-The bad model of the previous section assumes that each data point is independent.
-For example, the spectral lines are actually bad, and this means that if a single line is wrong, not just the core of the line (which may be a 10 - 20 sigma residual) will be wrong, but also the wings of the line, which may be only 5 or 3 sigma wrong.
+We now discuss bad data models that leverage more of the information about our problem to more robustly handle systematic errors.
+In bad model of the previous section, we assumed that each data point is independent.
+However, we know that Class I, II, and III errors will be correlated across several pixels, where the correlation length is the typical width of a spectral line, set by the stellar rotation kernel $v \sin i$, the instrumental resolution kernel, or both.
+Some very strong lines (such as the Balmer lines, or the Mg b triplet) will have wings that are correlated across a wider length of pixels.
 
-Now, let's come up with a model which includes spatial (wavelength) correlations in the bad data model.
+\subsubsection{Mixture of multiple components}
+When we believe that a sample population might be composed of several populations with different behaviours (Class 0, I, II, and III), it makes sense to expand the bad data model to a multiple mixture model.
+Following \citet{gcs+04}, our likelihood function becomes
+\begin{equation}
+  p(\fD, \vec{Q} | \vt) = p(\fD | \vec{Q}, \theta) p(\vec{Q}) = \prod_\lambda \prod_{m = 0}^3  {\cal L}_m(\fD(\lambda | \theta_m))^{q_{im}}
+\end{equation}
+where we use $\vec{Q} = \qM$ to describe a single vector of all the individual $q_{im}$ flags.
+For example, when a Class I, II, or III error occurs, the core of the line may be a 20 sigma residual while the wings of the line will taper off to a 3 sigma residual, and eventually to a 0 residual once it reaches the continuum.
+If we treat pixels as uncorrelated, then the small-residual wings of the bad lines might actually have more probability if they were labelled as ``good'' pixels.
+To remedy this situation, we attempt to develop a bad data model which includes spatial (wavelength) correlations.
 
-How could we accomplish this?
-If we identify one point as bad, then the flags around it will have a correlation length, ie, the prior on $q_i$ will be bad.
-This means that for however many nearby pixels that will be bad (most likely a constant amount for uncrowded lines), then this means that we could still integrate over the qis, it's just that we would now have to take a contribution from the nearest 5 pixels, say.
-That central pixel would then mark the ``line'' that is bad.
-
-That is one way to mark the ``badness of the line,'' that the qi are correlated. But we also know that the $y_i$ will be correlated too.
-
-This correlation length will probably depend on the type of star, and the contribution of the band.
-For example, if they're simple FGK stars, then the bad lines might be singular.
-But if they are M stars, then there is a chance that there might be a large ``pseudo-continuum'' region that might be incorrect.
-
-Once we have identified the center of the bad line, then there will be a decrease in the nearby pixels, similar to a Gaussian shape.
-Presumably we could come up with a correlation in this space too.
-(or, it could simply be that the correlation between $y_i$ is flat in the case of an M star.
-
-This might actually save the marginalization. We could still do the backwards ``what is the probability a given line is bad.''
-
-For Class I and II errors, we could have a hyper parameter describing the width of each line, and another hyper parameter describing the (distribution of) depth. These could also be sampled in.
-
-For Class III errors, describing how line strengths are off makes sense for stars with continuum, but as we move to later type stars (K stars) that do not have a nice continuum, we may run into trouble.
+This correlation length will probably depend on the spectral type of star.
+For example, if they're simple F and G stars then the bad lines will likely be singular.
+But if they are M stars, then there is a chance that there might be a large ``pseudo-continuum'' region that would be incorrect.
+For Class I and II errors, we could include a hyper parameter describing the width of each line, and another hyper parameter describing the (distribution of) depth.
+For Class III errors, describing how line strengths might be wrong makes sense for stars with continuum, but as we move to later type stars (K stars) that do not have a nice continuum, we may run into trouble.
 For the later stars (M dwarfs), there might be a whole region that is incorrect.
 
-There are different ways to describe a correlation over pixels. Gaussian random fields. Autocorrelation function.
+Proper Class III modeling is very important, following the discussion of \citep{mga13}.
+Some spectral features will not fit perfectly, especially at the lower end of the temperature scale.
+However, it's important to still take these features into account when modelling since they still contain information.
+How should we account for the fact that some parts of the spectrum carry more information about $\vg$ than other parts of the spectrum?
+Is it important to weight this parts beforehand?
+Or should this contribution to the posterior naturally weight itself, the way certain points might heavily contribute to the slope and intercept of a linear fit?
 
-\subsection{Simple correlation between pixels}
-For a simple correlation, say that if a given $q_i$ is bad, then there is also a high probability that the neighboring pixels are also bad.
-Say we know from inspection that it seems as though bad lines knock out 5 pixel chunks.
+\subsubsection{Correlation of the $q_i$ flags}
 
-Caveat: I think this assumes that bad lines will not occur within the same spread of each other. Maybe we should include a $P_B$ prior on the good counts too?
-
+In the manner of Gaussian random processes, we develop a covariance function
 \begin{equation}
-  p(q_i = 0 | q_{i \pm 5} = 0) = P_B
+  {\rm cov}(q_i, q_j) = \rho(| \lambda_i - \lambda_j|)
 \end{equation}
-
-\begin{align}
- p(q_i = 0 | q_{i \pm 5} = 0) &= 1/5 \\
- q_{i \pm 4} &= 2/5 \\
- q_{i \pm 3} &= 3/5 \\
- q_{i \pm 2} &= 4/5 \\
-\end{align}
-
-This might need more work with $P_B$.
+or where the covariance matrix has components
+\begin{equation}
+  \Sigma_{ij} = \sigma^2 \rho(|\lambda_i - \lambda_j|)
+\end{equation}
+where
+\begin{equation}
+  \rho_{ij} = \frac{ {\rm cov}(i,j)}{\sigma_i \sigma_j} = \frac{E[(X - \mu_x)(Y - \mu_y)]}{\sigma_x \sigma_y}
+\end{equation}
+that describes the correlation between the state of the bad data flags.
+The simplest thing to do would be to take the common correlation length in the spectrum imposed by $l$.
+If we say that the typical feature will have five adjacent pixels wrong, then the correlation matrix would be
+\begin{equation}
+  \begin{bmatrix}
+    1 & 4/5 & 3/5 & 2/5 & 1/5 \\
+    4/5 & 1 & 4/5 & 3/5 & 2/5 \\
+    3/5 & 4/5 & 1 & 4/5 & 3/5 \\
+    2/5 & 3/5 & 4/5 & 1 & 4/5 \\
+    1/5 & 2/5 & 3/5 & 4/5 & 1\\
+  \end{bmatrix}
+\end{equation}
+(this is embedded in an N-pixel sparse matrix where everywhere but the diagonal is 0). And so
+\begin{equation}
+  {\rm cov}(q_i, q_j) = \left \{ \begin{array}{cc}
+    1 - \frac{|\lambda_i - \lambda_j|}{5} & |\lambda_i - \lambda_j| \leq 5 \\
+    0                                    & \textrm{else}
+\end{array}
+\right .
+\end{equation}
+Once we move beyond 5 pixels, any correlation vanishes.
 
@@
 The marginalization could still be done analytically, but now the likelihood function involves a correlation between adjacent pixels.
-Could we use a sparse array for this?
+Alternatively, we might be able to use a sparse array for this?
 Or, we could evaluate all of the pixels, and then have the final likelihood function involve some correlation terms between the lnprobs evaluated at each pixel.
 We could do this using the logarithm so that everything is additive.
+We might also be able to evaluate the FFT of the $\log(p(D_i))$ and just do the autocorrelation in Fourier space?
 
-Or, what if we evaluated the FFT of the $\log(p(D_i))$ and just the autocorrelation in Fourier space?
+\subsubsection{Correlation of adjacent $\fM(\lambda)$}
+We haven't yet mentioned what ${\cal L}_m$ should be for each $m$.
+Not only will the flags $q_i$ be spatially correlated but the values of $\fM$ will also be correlated on the scale of $v \sin i$ or the instrumental resolution kernel, whichever is larger.
 
-\subsection{Correlation between $y_i$}
-Not only will the flags be spatially correlated (if a point is bad, the nearby $\sim$ 5 points are also likely to be bad) but for OBAFGK stars, the lines are likely to be Gaussian shape.
-That is if $q_i$ specifies the center of the line, then the neighbooring $y_{i+1}$ or $y_{i-1}$ will be of lower residual height.
-This means that the inside of the likelihood function should also have spatial correlations.
-
-Could this correlation be described by a Gaussian random process? \citep{rw05}.
-
-\subsection{Different flags and weights}
-Instead of a good or bad flag, what if we had a flag that would identify which class of error the line became? For example we could have the option of $q_i = \{0, 1, 2, 3\}$, where $q_i=0$.
-
-How do we account for the fact that some parts of the spectrum carry much more information about $\vg$ than other parts of the spectrum?
-Is it important to weight this parts beforehand?
-Or should this contribution to the posterior naturally weight itself, the way certain points might heavily contribute to the slope and intercept of a linear fit?
-
-If so, we could imagine having a Gibbs sampler on the weights of each pixel.
-
-Proper Class III modeling is very important, following the discussion of \citep{mga13} Mann et al, and see recent example from Rajpurohit 2103.
-Some features will not fit perfectly, especially at the lower temperature scale.
-However, it's important to still take these features int account with the modelling.
-They also use a $\chi^2$ likelihood function, but perhaps even less appropriate is that they draw error bars by saying that the $\chi^2$ value changes by 5\%.
-In reality, they are checking the fits by eye and saying whether or not a spectrum matches.
-This approach is no more justified than a by-eye spectral classification.
+For OBAFGK stars, the lines are likely to be Gaussian shape.
+That is if $\fD(\lambda_i)$ specifies the center of the line, then the neighboring $\lambda_{i \pm 1,2,3,\ldots}$ pixels will be of lower residual height.
+For M stars, there might be a wider correlation length and the shape of the lines might not be well-described by a Gaussian.
+Whatever the case, this means that the inside of the likelihood function should also have spatial correlations.
+In this case, the likelihood function also includes the nearest 5 pixels
+\begin{equation}
+  {\cal L}(\lambda_i) = \exp\left ( 
+  \begin{bmatrix}
+x & x & \dots\\
+x & x & \dots\\
+  \end{bmatrix}
+  \right )
+\end{equation}
+or maybe evaluate the likelihood and then matrix multiply?
+This matrix math needs to be worked out.
+The covariance function between $\fD(\lambda_i)$ could be learned from the spectrum itself (a la Gaussian random processes \citep{rw05}), or specified to create a Gaussian shape.
+Then, the likelihood functions for each mixture component in the model would be custom tailored for the type of error that it expected to encounter.
+(Strength and shape of residuals).
 
 \subsection{Hyperparameters}
-Difficulty of combining different data sets with different errors \citep{mb13}. Or a hyperparameter like \citep{sdm+07}. Shrinkage similar to \citet{kru10}.
-
-\citet{sdm+07} use ISO SWS infrared spectra.
-Hierarchical approach.
-Able to synthesize their own spectra.
-Use systematic error flags from the instrument, as well as statistical flags.
-Hierarchical model that uses shrinkage between the observed spectrum and the synthetic spectrum.
-They do a fit to the variance with a hierarchical linear mixed model (sec 5.2.2, and their figure 7) using design matricies and regression coefficients.
-This is probably to incorporate some sort of correlation length on the systematic errors, and ``smooth'' the function, as they say.
-I was having some difficulty understanding it, but it seems as though they were keeping the noise components fixed, and they did not have a contribution to the error budget from regions where the spectral synthesis \emph{itself} contained systematic errors.
-Perhaps these were masked by the systematic error from the instrument.
+\citep{mb13} discuss the difficulty of combining different data sets for the combined inference on a single problem.
+They use a hyperparameter approach that can artificially inflate the variance of each data set to find a common posterior.
+They also allow correlations between data sets.
+We could imagine setting up a hyperparameter approach that can tune the sensitivity of certain lines, treating different lines as different datasets.
+Such a hyperparameter might also undergo ``shrinkage'' similar to \citet{kru10}.
 
@@
 If our bad model is sufficiently complex that we are not able to analytically marginalize over the $q_i$, it is still possible to sample in this huge parameter space.
-A stellar model with flags for each pixel would have $\vstar + N$ parameters, where for a typical optical spectrum $N >$ 100,000.
+A stellar model with flags for each pixel would have $\vstar + N$ parameters, where for a typical high resolution optical spectrum $N >$ 100,000.
 Even using an ensemble sampler, sampling in this space would take a tremendous amount of time to converge.
@@
 
+\citet{gcs+04} suggest to start with a crude estimate (such as $k$-means clustering) to tentatively identify which observations have been drawn from which mixture components.
+They also suggest that the expectation-maximize (EM) algorithm could also be used to sample this posterior.
+
 \section{Injection of fake lines}
+As an alternative approach to modelling the systematically wrong regions of the spectrum, we could actually inject fake lines into the spectrum to compensate for errors in the models.
+Presumably, each line would be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
+In M star spectra, this might actually need to be a larger continuum feature that is added.
 
-An alternative approach is instead of modelling the systematically wrong regions of the spectrum, actually add in lines to the spectrum.
-This can be thought of as correcting the models.
-We could add in a set number of lines.
+Line strength could be described by a hyper prior, and FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM (in velocity space).
+We could penalize the model fits by having a hyperprior on the number of bad lines that can be created.
+Given a bad line and the linelist that produced the synthetic models, it may be possible to identify other regions of the spectrum that might be bad.
+For example, if a vanadium line is bad, then we could check to see if the other vanadium lines might also be bad.
 
-Presumably, each line would be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
-Strength could be described by a hyper prior, and FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM (in velocity space).
+Class III errors might be difficult to address in this paradigm since we would be fitting the different of two lines, though we might still be able to get it to work.
 
-In later type spectra, this might actually need to be a larger continuum feature that is added.
+\subsection{Piecewise metallicity model}
+In theory, it is possible to fit the profiles of only certain elements to determine the abundances of certain elements.
+By masking all non-alpha elements, \citet{hus12} use PHOENIX spectra to derive individual alpha abundances of stars.
 
-We could penalize the model fits by having a hyperprior on the number of bad lines that can be created.
-
-In theory, it is possible to fit the profiles of only certain elements to determine the abundances of certain elements (see citation in conference proceedings of Husser).
-
-It may be possible to use a linelist from the group that produces the PHOENIX spectra, which describes the element that gave the highest contribution to the opacity in that bin.
-Now, if the vanadium lines are bad, then we could have a prior that the other vanadium lines might also be bad.
-We could assemble a piecewise spectrum, where certain pixels are drawn from a different combination of $\vg$, and these pixels had an associated normalization constant so that the piecewise transition was smooth (if, say the bolometric temperature was different).
-
-Class III errors might be difficult to fit in this approach. However, if the line width does not change dramatically, then we should be able to account for things this way.
+Rather than masking spectra, we could assemble a piecewise spectrum where regions of the spectrum might be drawn from a different combination of $\vg$, say with lower metallicity in order to account for a bad line strength.
+These regions would have to be normalized properly so that the composite spectrum is sufficiently smooth.
+Coupled with a linelist, one could tweak individual elemental abundances.
 
 \section{Validation tests}
-Comparison to main sequence stars of HAT-P-9, and WASP-14, from the TRES archive \citep{tfs+12}.
-
-Comparisons using HIRES data?
+Comparison to main sequence stars of HAT-P-9, and WASP-14, from the TRES archive \citep{tfs+12}. Possible comparisons using HIRES data.
 

### 2014-01-23 paper/paper.tex
Paper updates

--- 2014-01-16 paper/paper.tex
@@
 \newcommand{\vg}{\vt_{\star, {\rm grid}}}
+\newcommand{\vpp}{\vt_{\star, {\rm post}}}
 \newcommand{\vstar}{\vt_{\star}}
@@
 \newcommand{\vc}{\vec{c}}
-\newcommand{\fM}{f_{\rm M}}
-\newcommand{\fD}{f_{\rm D}}
-\newcommand{\vD}{\vec{D}}
+\newcommand{\fM}{ {\bm M}}
+\newcommand{\fMi}{M_i}
+\newcommand{\fD}{ {\bm D}}
+\newcommand{\fDi}{D_i}
 \newcommand{\dd}{\,{\rm d}}
@@
 
-\title{A method for inference of fundamental stellar parameters using high resolution spectra}
+\title{Methods for the spectroscopic inference of fundamental stellar parameters}
 \author{\today{}\\
@@
 
-\begin{abstract}
-  Matching observed spectra of stars against a grid synthetic spectra is a popular approach to determine fundamental stellar parameters such as effective temperature, surface gravity, and metallicity.
-However, a fundamental problem with this approach is that the synthetic spectra contain systematic errors such as wrong or missing spectral lines.
-In the presence of such errors, a simple $\chi^2$ fit between spectra will likely yield biased parameters and incorrect uncertainty estimates.
-This behaviour is analogous to the way a few strong outliers might severely bias the fit to a straight line.
-These spectral outliers are difficult to identify \emph{a priori}; procedures such as sigma clipping might be prone to their own systematic effects and discard useful information.
-We outline an approach to probabilistically identify and account for spectral features that deviate from reality.
-Such an approach is most needed when operating at high spectral resolution ($R >$ 10,000), since this is when systematic deviations begin to appear most striking, however our technique is still valid for lower resolution spectra as well.
-We also outline some approaches for spectral fits to brown dwarfs and planets, where instead of spectral lines there may be systematic deviations of a larger swath of the spectrum.
-\end{abstract}
+\section{Introduction}
+The fundamental properties of stars---effective photospheric temperature, surface gravity, and metallicity--are crucial for our understanding of pre-main sequence stellar evolution \citep{dm97, bca+02}.
+The exoplanet community is also intensely focused on deriving accurate stellar properties because the properties of newly discovered exoplanets must necessarily be quoted relative to those of their host stars (e.g.
+\citealt{tfs+12,blj+12,ssm+13}) and uncertain star properties translate into uncertain planet properties \citep{kan14}.
+Many techniques\footnote{If you are interested, I have also compiled a more verbose review of modern techniques described in the literature.
+However, for the sake of brevity I have omitted it from this document.} exist for stellar characterization such as line indexes, machine learning, and spectral synthesis\footnote{Given infinite computational power, we would prefer to fit \emph{all} of the parameters that completely describe a synthetic spectrum: atmospheric structure, individual elemental abundances, and atomic constants describing the opacity contribution of a line.
+Approximations to the synthesis and efficient MCMC sampling parallelized on a cluster might soon put this approach within reach.}.
+In order to derive accurate and robust stellar properties for our small sample ($\sim 20$) of T Tauri stars (some binary), we have decided to forward model the observed spectra by drawing from a grid of synthetic spectra.
 
-%\keywords{}
+\begin{table}[!htb]
+\begin{tabular}{ll}
+\hline
+\hline
+Symbol & Description\\
+\hline
+\hline
+$i$ & index specifying a pixel\\
+$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+  & that parameterize a synthetic spectrum from the grid\\
+$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+  & are applied during ``post processing'' of the synthetic spectrum\\
+$\vstar$ & $\{\vg,\vpp \}$\\
+$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
+$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
+$k(i | \vN)$  & polynomial function that multiplies a synthetic spectrum\\
+$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
+$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
+$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
+$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
+$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
+$\sigma_i$ & Poisson noise for a given pixel $i$\\
+$R_i$ & residuals $(\fDi - \fMi)/\sigma_i$\\
+\hline
+\end{tabular}
+\caption{Nomenclature used in this document}
+\label{tab:nomenclature}
+\end{table}
 
-%\begin{figure}[htb]
-%\begin{center}
-%\includegraphics{file}
-%\caption{}
-%\label{fig:}
-%\end{center}
-%\end{figure}
+\subsection{Pixel to pixel techniques using synthetic spectral libraries}
+\label{sec:pix}
+Most forward modeling techniques compare pixels directly using some form of
+\begin{equation}
+  \chi^2 = \sum_i \left [\frac{ \fDi - \fMi}{\sigma_i} \right ]^2 = \sum_i R_i^2
+\end{equation}
+where the variables are defined in Table~\ref{tab:nomenclature}.
+A serious problem with this approach is that synthetic spectra are not perfect: systematic errors invalidate the $\chi^2$ assumption and bias parameter estimates and their uncertainties.
 
-\section{Introduction}
+Current methods either circumvent this problem by using high quality synthetic grids that were painstakingly tuned over a narrow wavelength range \citep[$\lesssim 500$ \AA]{blj+12,sb13}, or they use less-perfect grids with full optical coverage and sigma clip the bad regions of the spectrum \citep{kpb+09,mga13}.
+Masking (either implicitly or explicitly, respectively) is troublesome because, in the words of \citet{mga13}, ``regions with [only] modest agreement between the real and synthetic spectra [that might be masked] may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.'' Masking could very well destroy stellar information in the pursuit of a better fit.
 
-The fundamental parameters of stars---mass, luminosity, effective photospheric temperature, surface gravity, and metallicity--are of supreme interest to astrophysicists for a wide variety of reasons.
-A salient example is that most newly discovered exoplanets must necessarily their mass and radius measured relative to the properties of their host stars (e.g.
-\citealt{tfs+12,blj+12,ssm+13}).
-The properties of stars are of course also important in their own right because they inform us about our ideas of stellar evolution \citep{dm97, bca+02}.
-Because stars are the fundamental units of the cosmos, techniques to infer the properties of stars date back to the dawn of astronomy.
-We briefly review some of the modern and popular techniques in the literature.
-Approaches can be broadly classified into techniques that compare stars by some broadly defined indices and techniques that attempt to forward-model an observed stellar spectrum by doing a pixel-to-pixel comparison.
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics{lineclasses}
+caption{The four classes of line behavior.
+\textbf{Class 0}: the model behaves as expected.
+\textbf{Class I}: a spectral line is present in the data but missing from the model (in this case, the line is also blended).
+\textbf{Class II}: a spectral line exists in the model where there is only continuum in the data.
+\textbf{Class III}: lines exist in both the data and model, but some are too strong while others are too weak.
+The lower panels show the residual spectra in units of the Poisson uncertainty.
+The data is a spectrum of WASP-14, an F star.
+The model is from the PHOENIX grid\citep{hwd+13}.}
+\label{fig:lineclasses}
+\end{center}
+\end{figure}
 
-\subsection{Similarity techniques}
-Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra.
-I am calling them similarity techniques because they compare characteristics of spectra, not actual pixels.
+\subsection{Classes of synthetic spectral line behavior}
+Illustrated in Figure~\ref{fig:lineclasses} are three ways that synthetic spectra often diverge from reality.
+When a spectral line appears in the data but not in the model, due to interstellar or telluric absorption or a missing opacity source, we call this a \textbf{Class I} error.
+When a spectral line appears in the model but not in the data, presumably due to a combination of incorrect atomic constants or atmospheric structure, we label it a \textbf{Class II} error.
+Both of these types of errors are present over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$) and there does not exist a $\vg$ that could ever fit these lines, making these regions useless for determining stellar parameters.
+We know that for a normal AFGK star Class I residuals will always be negative and will not change significantly with $\vg$, while Class II residuals will always be positive and their magnitude should decrease with decreasing metallicity.
+To help identify Class I and II errors,  we could use the partial derivatives of an individual synthetic spectrum with respect to $\vg$ and a prior on the range of parameters over which the residuals will be correlated to determine which lines show Class I or II behavior.
 
-\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity.
-Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters.
-A recent example using M dwarfs is \citet{nci+14}.
-This technique works, but there are many things that could be improved.
-First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars.
-This is both a disadvantage and an advantage.
-One one hand by choosing a subset of lines to fit, one may be unwittingly throwing away useful information.
-On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters.
-Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.
+When some line strengths do not match at a $\vg$ that properly fits most other lines, we term these \textbf{Class III} errors.
+These types of errors are pernicious because there does exist a $\vg$ that would fit each line individually, but no one $\vg$ can be chosen that fits all the lines.
+We must then decide which regions to trust and by how much, because Class III regions \emph{still contain information about $\vg$}.
+In fact, they might contain \emph{the most} information about $\vg$ since these delicate regions are the most sensitive to changes in $\vg$ and are consequently whey they are the hardest to model correctly.
+In other words, Class III regions might be where the spectrum gradient ($\sum_{j} \partial \fMi/ \partial \vt_{ \star {\rm grid}, j}$) is the largest.
 
-\paragraph{Principal component analysis} PCA reduces the dimensionality of the data space and searches for the most salient features of the spectra \citep{ptb14}.
-How does one properly estimate parameter uncertainties in such an approach?
-It seems like this approach exists because it is an easy thing to do.
+\section{In the limit of a perfect stellar model}
+To generate a model spectrum $\fM(\vg, \vpp)$, we use $\vg$ to linearly interpolate a synthetic spectrum from the grid, and then use $\vpp$ to apply instrumental and rotational broadening in the Fourier domain, Doppler shift, correct for extinction, and then downsample to the exact pixels of the data spectrum using spline interpolation\footnote{Because we are dealing with flux-calibrated data and synthetic spectra, we \emph{resample} to lower resolution.
+We do not \emph{rebin} because we are dealing with flux-density, not photon counts}.
 
-Matrix inversions: MATISSE \citep{rbd06}.
-This method is similar to PCA in that it derives basis vectors that describe the fundamental parameters of the star.
-In the PCA analysis, the principal components do not necessarily correspond to true physical quantities.
-The MATISSE algorithm determines weights that ensure the components are similar to the stellar parameters.
-They only test the algorithm against input synthetic stellar spectra to detect biases in the parameters, and against the Sun and Arcturus.
-They do acknowledge that there will be systematic deviations between the synthetic spectra and real spectra, but they do not account for the possible inflation of error estimates.
+To account for imperfect flux calibration, we extend the approach of \citet{elh+06} and multiply $\fM(\vg, \vpp)$ by $k(\vN)$, which is a Chebyshev polynomial described by a set of $N$ Chebyshev coefficients.
+We can either sample in the coefficients or analytically marginalize over them.
+We favor this approach over the traditional technique of normalizing to the stellar continuum due to concerns about correctly placing the pseudo-continuum in M stars.
+From now on we refer to $\fM(\vt) = k(\vN) \fM(\vg, \vpp)$.
+The mathematical details are in \S~\ref{sec:chebyshev}.
 
-Given the inherent systematic uncertainties involved in each method, some groups elect to combine the inference from different algorithms, such as (SLOAN SSPP) \citep{lbs+08}.
-This approach includes machine learning along with many of the previously mentioned techniques.
+Currently we are sampling this $\chi^2$ posterior with the MCMC ensemble sampler \citep{gw10} implemented in \texttt{Python} by \citet{fhl+12} as \texttt{emcee}.
+In this section, we described our approach to sample the posterior assuming perfect synthetic spectra.
+However, if one does a straight up $\chi^2$ fit to our spectra, one can get posteriors that have 1-sigma contours of $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
+This would be as precise as stellar parameters obtained with astroseismology!
+Clearly there is something going wrong here.
 
-\subsection{Pixel to pixel techniques using spectral libraries}
-Many high quality synthetic spectral grids are now available.
-Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement.
-We define this combination as
+\section{Beyond a perfect stellar model}
+\label{sec:residuals}
+\subsection{Additional sources of noise}
+\begin{itemize}
+  \item CCD readnoise and imperfect 1D spectrum extraction over night sky lines will almost definitely increase the noise in a pixel beyond theoretical Poisson limits.
+Normally this information will be encapsulated in a ``sigma spectrum'' output by IRAF, but unfortunately we do not have access to this for our current dataset.
+  \item The stellar continuum can flicker/jitter (as shown by Kepler), which would produce a small departure from a perfect continuum.
+This might explain why even in the Class 0 region of our spectrum (Figure~\ref{fig:lineclasses}, left panel) the residuals are slightly larger than 1 $\sigma$ on average.
+For our T Tauri stars this effect may be more pronounced.
+  \item When we interpolate a synthetic spectrum from the library to a specific $\vg$, we will introduce interpolation errors depending on how close we are to a grid point in the library.
+\citet{hus12} investigate the possible range of interpolation errors at specific $\vg$ by dropping out a spectrum from the grid, interpolating over it, and then comparing the true spectrum to the interpolated spectra.
+For hotter stars with well defined continua, we find the maximum error is $\lesssim 5$\%, while for M stars the error could be as large as $\sim 20$\% for some spectral features.
+By doing this interpolation test across the grid, we could determine an ``interpolation error spectrum'' that would encapsulate the maximum error possible for an interpolated spectrum, weighted by its distance from the nearest grid points.
+If a typical spectrum has signal/noise of $\sim 30$ per pixel (or noise/signal ratio of $1/30 \approx 3$\%), an interpolation noise contribution of $\sim 5$\% the level of the signal level would be significant.
+\end{itemize}
+However, we have tried arbitrarily increasing the Poisson errors by a factor of 2 or 3 and it didn't significantly change the results, suggesting that something more drastic is needed beyond simply adding in more noise.
+
+\subsection{Gallery of likelihood functions}
+\label{sec:gal}
+A problem with the $\chi^2$ likelihood function is that the Gaussian tails rapidly decline, making outliers extremely improbable.
+A common approach for ``robust regression'' is to use a likelihood function with fatter, more forgiving, tails.
+Although not rigorously motivated, this is can be a quick thing to try.
+We explored several likelihood functions including a combination of Gaussian with exponential wings, a Lorentzian (Cauchy) function, and a function described  by \citet{ss06}, which treats the noise estimate for each data point suspiciously.
+In Sivia's model, the true noise $\sigma$ can be no smaller than some original estimate $\sigma_0$, but it may be much larger.
+By putting a prior on $\sigma$,
 \begin{equation}
-  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
+  p(\sigma_i | {\sigma_0}_i) = \left \{ \begin{array}{cc}
+    {\sigma_0}_i/\sigma_i^2 & \sigma_i  \leq {\sigma_0}_i\\
+    0                 & \textrm{else}
+  \end{array}
+    \right.
 \end{equation}
-Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid.
-Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.
+and marginalizing over the $\chi^2$ likelihood, the new likelihood function becomes
+\begin{equation}
+p(\fDi | \fMi, {\sigma_0}_i) = \frac{1}{ {\sigma_0}_i \sqrt{2 \pi}} \left [ \frac{1 - e^{-R_i^2/2}}{R_i^2} \right ] 
+\end{equation}
+where $R_i = (\fDi - \fMi)/{\sigma_0}_i$.
+This function has even fatter tails than the exponential model and seems to give results less biased by outliers, however it is not truly satisfying.
+Which lines contributed most to the posterior?
+What about the wings of class I and II errors?
+These individual pixels might still be given a weight of ${\sigma_0}_i$ because they are not nearly as bad as the line centers.
 
-The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
+Other functions we might try in the future include a Student-$t$ distribution, a linear combination of Gaussians, or the $\chi^2_Q$ function described in \citet{hbl10}.
+While ``softer'' likelihood functions do help somewhat, we would like to move beyond tweaking the fitting function towards an approach that properly incorporates our knowledge of the systematic errors of the synthetic grids.
+In the remaining two sections of this document we outline potential ideas that we might try next, but have not yet implemented in code.
+
+\section{Bad data mixture model}
+The bad data model \citep{pre97,hbl10} is a powerful framework for forward modeling data, which allows the modeler to explicitly account for outliers using a generative model.
+In the simplest mixture model, each pixel has a binary flag $q_i$ which specifies whether that data point is good or bad.
+Then, for the entire data set we have an ensemble of $\qN$.
+If this ensemble of $\qN$ is chosen appropriately, and the good pixels are flagged as good and the bad pixels are flagged as bad, the posterior will be maximized.
+We adopt $P_B$ for the \emph{a priori} probability that a given pixel will be bad, $Y_b$ for the mean flux of the bad pixels, and $V_b$ for their variance.
+Written out explicitly, this is
+\begin{multline}
+  {\cal L} = p({\bm D} | \vt,\qN, Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma_i} \exp \left[ - \frac{(\fDi - \fMi)^2}{2 \sigma_i^2} \right] \right ]^{q_i}  \times \\
+  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma_i^2)}} \exp \left[ - \frac{(\fDi - Y_b)^2}{2 (V_b + \sigma_i^2)} \right]
+  \right ]^{1 - q_i}
+\end{multline}
+If we put a Bernoulli prior on the $q_i$
 \begin{equation}
-  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
+  p(q_i | P_b) = [1-P_b]^{q_i} P_b^{[1-q_i]}
 \end{equation}
-where the best model is defined as one that minimizes the residuals in some sense.
-We will discuss other goodness of fit functions soon.
-Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda | \vt)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda | \vt)$.
+the probability of any given ensemble of $\qN$ is 
+\begin{equation}
+  p(\qN | P_b) = \prod_{i=1}^N [1-P_b]^{q_i} P_b^{[1-q_i]}
+\end{equation}
+We can analytically marginalize over the settings for a single flag by a discrete sum over the two states of $q_i = 1$ and $q_i = 0$. 
+\begin{multline}
+  p(\vt, P_b, Y_b, V_b | D_i)  =  \int p( \vt, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
+  = p(\vt, q_i=1, P_b, Y_b, V_b | D_i) + p(\vt, q_i=0, P_b, Y_b, V_b | D_i) \\
+  = \Bigl[ p_{\rm good}(D_i | \vt)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(\vt ,P_b, Y_b,V_b) 
+\end{multline}
 
-Most pixel-to-pixel techniques use a $\chi^2$ likelihood function and seek to minimize the square of the residuals
+\begin{multline}
+  p(\vt,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i |\vt)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(\vt,P_b, Y_b, V_b)\\
+  \boxed{
+    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_i^2}} \exp \left[ - \frac{(\fDi - \fMi)^2}{2 \sigma_i^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma_i^2)}} \exp \left[- \frac{(\fDi - Y_b)^2}{2 (V_b + \sigma_i^2)} \right] \right\}
+  \times p(\vt,P_b,Y_b,V_b)}
+\end{multline}
+
+In its simplest form, the bad data model results in a likelihood function that is a linear combination of Gaussians.
+Although this functional form might be similar to some of the likelihood functions in \S~\ref{sec:gal}, the benefit of this bad data approach over a gallery of likelihood functions is that it is motivated by the characteristics of the bad data ($P_B$, $Y_b$, and $V_b$).
+If we wish to add additional information about the characteristics of the bad data, we can do so.
+
+\subsection{Mixture of multiple components}
+When we believe that a sample population might be composed of several populations with different behaviours (Class 0, I, II, and III), we can expand the bad data model to a multiple mixture model.
+Following \citet{gcs+04}, we develop a mixture model with $M=4$ components, one for each class of error.
+Our likelihood function is then
 \begin{equation}
-  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
+  p(\fD, \vec{Q} | \vt, J) = \prod_i^N \prod_{m=0}^M \left [ J_m {\cal L}_m(\fDi | \vt_m) \right ]^{q_{im}}
 \end{equation}
-It appears that for many techniques, $\sigma(\lambda)$ is the Poisson photon counting errors.
-For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
-For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.
+where $\vec{Q} = \qM$ is a two dimensional array ($N \times M$) of all the individual $q_{im}$ flags, and $J_m$ is a weight that represents the total fraction of each component in the mixture\footnote{The weights are normalized $\sum_m J_m = 1$}, and ${\cal L}_m$ is the likelihood function for $m$th mixture component.
+At this point, $\vt$ now contains all of the parameters that might be used by any of the likelihood functions.
+$\vt_m$ is used to denote only those parameters that will be used for ${\cal L}_m$.
+Given what we know about these synthetic spectral errors, we could specify the following likelihood functions
+\begin{equation}
+  \begin{array}{lcccc}
+ {\cal L}_0 &= p(\fDi, q_{i0} = 1) &\rightarrow & \textrm{Class 0} & \chi^2\\
+ {\cal L}_1 &= p(\fDi, q_{i1} = 1) &\rightarrow & \textrm{Class 1} & \textrm{flat lnprob}\\
+ {\cal L}_2 &= p(\fDi, q_{i2} = 1) &\rightarrow & \textrm{Class 2} & \textrm{flat lnprob}\\
+ {\cal L}_3 &= p(\fDi, q_{i3} = 1) &\rightarrow & \textrm{Class 3} & \textrm{soft } \chi^2\\
+\end{array}
+\end{equation}
+This would treat Class I and II errors as if they had been masked from the dataset but allow Class III to have a slight influence on the posterior since these regions still contain information about $\vstar$.
+The softening term of the $\chi^2$ could even be sampled for, much like the $V_b$ of the original bad data model.
 
-\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations.
-Any spectral library can be used as a backend, including empirical spectral libraries such as the one from ELODIE \citep{psk+07}.
-They explicitly take into account some of the necessary procedures when fitting spectra, such as continuum-multiplication by a polynomial to account for flux calibration errors and rejection of bad data or model points.
-They use an iterative kappa-sigma clipping method to discard ``bad'' points for the fit.
-They test that their clipping technique converges to the same answer by starting guesses located at different parameters $\vt$ and seeing if similar lines are masked and similar best fit parameters $\vt$ are obtained.
+\subsection{Spatial correlations of pixels}
+Previously we assumed that each data point is independent, but we know that the spectrum is spatially correlated across several pixels with a correlation length set by the width of the convolution of the stellar rotation kernel ($v \sin i$) and the instrumental resolution kernel, which is also the typical width of a spectral line\footnote{Some very strong lines (such as the Balmer lines or Mg b triplet) will have wings with a larger correlation length.}.
+This correlation length will have some dependence on the spectral type of star.
+For example, F and G stars will likely have singular, well-defined lines while M stars might have a large molecular feature covering a wide swath of the spectrum.
+Neighboring pixels are actually correlated in two coordinates: the $q_{im}$ flags and the residuals $R_i$.
 
-\paragraph{\citet{mga13}} fit BT-Settl PHOENIX models to M dwarfs.
-Because high quality, high resolution M dwarf spectra are difficult to obtain owing to their intrinsic faintness, the development of stellar libraries for later spectral types has typically lagged behind.
-They use a $\chi^2$-like likelihood function and iterate to mask bad pixels.
-Unlike ULySS, however, they iterate the method against a sample of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars.
-Presumably, such a consistent residual would be due to an incorrect physical constant for a certain line or molecular feature.
-After masking, they repeat the fitting and claim that after two iterations the masks do not change.
-They note that in the future, rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures.
-\textbf{However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.}'' Thus, masking ``bad regions'' could very well destroy stellar information in the pursuit of a better ``fit.'' This is a nuanced point that masking schemes generally neglect.
-We aim to address later in this document.
+\subsubsection{Correlation of the $q_i$ flags}
+The pixels at the center of a bad line will give a strong residual response, which will likely be flagged as bad.
+However, the pixels in the wings of the spectral line might not be sufficiently egregious to be properly flagged.
+To remedy this situation, we attempt to develop a bad data model which includes spatial (wavelength) correlations.
 
-\paragraph{``Bayesian'' methods} do exist.
-\citet{sdm+07} developed a hierarchical Bayesian approach that purportedly accounts for both statistical and systematic errors in the estimation of stellar parameters.
-I need to work through their math more carefully, but it seems as though they use a hierarchical model for the scatter between ``true'' and observed spectrum, and the unknown scatter in the relationship between $\vg$ and the $\fM$.
-They claim to account for systematics, but it looks like these are systematics errors due to the performance of the instrument, such as a bad flux calibration.
-They use infrared spectra at low resolution ($R \sim $ 1,200), and a weighted $\chi^2$ function for goodness of fit.
-As far as I can ascertain, the actually pixel-to-pixel fitting does not account for systematic errors in regions of the model spectrum.
-
-\citet{sb13} present a comprehensive Bayesian method that can fit low or high resolution spectra in addition to photometry and parallax data.
-They too, mention that they use a $\chi^2$ likelihood.
-They mention using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were created or applied.
-Their unpublished stellar library focuses on parts of the optical spectrum which contain the most stellar information, as determined by experts.
-As noted by \citet{sb13}, the grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA\ (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).''
-
-\paragraph{SPC} A very popular method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}).
-This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method.
-The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient.
-It is uncertain how statistically justified this uncertainty estimate is.
-Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match high resolution models of the Sun and Vega.
-
-\paragraph{General trends}
-Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors.
-This implicit masking brings up similar concerns to those of \citet{mga13}: how do you know that you are fitting the region of the spectrum that contains the most information about the stellar properties?
-
-\subsection{Line fitting using spectral synthesis}
-This is an improvement over the line index method, since now one includes the information contained in the line profile.
-With quality stellar spectral models, one can identify a few well-modelled lines to fit.
-By choosing a few good lines to fit, one could determine stellar parameters by matching line profiles for a few select lines \citep{dj03}.
-However, this is still subject to the need to identify promising regions of the spectrum.
-
-MOOG\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73}, SME\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96}, actually synthesize line profiles on the fly, compared to using a library of spectral models.
-This has the advantage of enabling individual tweaks to elemental abundances.
-And interpolation in the model atmosphere space, rather than the spectral line space.
-However, for many young T-Tauri stars, the spectrum is contaminated by accretion, veiling, and fractional star spot coverage, which can be a nuisance and cause systematic errors if traditional techniques are used.
-Matching equivalent widths.
-Interestingly, MOOG sythesizes a spectrum on the fly.
-
-SME allows adjustment of oscillator strengths.
-We could download the synthesis code for SME and synthesize our own spectra on the fly.
-The biggest impediment here would be that the simplest atmospheres will not get certain lines correct.
-The powerful synthesis codes (PHOENIX) will get the atmospheres and lines correct, but they take a long time to run (~hours) per model, and the code is not publicly available.
-To do a full-spectrum fit for these sorts of codes would likely be too computationally difficult.
-However, it would allow access to fitting in the parameter space that we want.
-
-\subsection{Stellar parameter estimation in a utopia}
-
-Although the spectral synthesis domain is the correct domain to solve for stellar parameters, the radiation transfer is too slow.
-
-The fitting difficulties of the previous section are a result of trying to fit data to a model in a smaller parameter space.
-Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different.
-If we think back to first principles, the ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$.
-These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote some value relative to solar, a fraction of models allow changes in alpha abundances), and atomic constants describing the opacity contribution of a line.
-\emph{All} of these parameters have uncertainties in them, some of these are rather dramatic (one continually hears lamentation about the opacities and line lists\dots), yet for any given stellar model these parameters are \emph{fixed}.
-There may yet be unconsidered stellar physics necessary to explain a small fraction of lines (say for example the cores of Ca lines), but the field of stellar modelling has progressed to the state that I would have the confidence to say that if a group focused all of their attention on a single line, and using only their current physical models and making tweaks to the afore-mentioned lines, they could produce a satisfactory model.
-
-In reality, this has been done using the Kurucz models in an unpublished grid of stellar spectra, over the range $5000 - 5300$ \AA.
-
-The real problem is much more rich, as is the amount of information one could extract from a high resolution optical spectrum.
-One would want to fit for individual abundances, marginalize over oscillator strengths.
-
-The unfortunate reality is that sampling in the stellar model space for the full optical spectrum is not yet a computational reality.
-There do exist codes (MOOG) that can do this approach but only for a small ($\sim$ a few lines) at once, and using simplifications.
-Through collaboration with a stellar atmospheric expert, such an approach may soon be a possibility.
-If the computation of a given model could be reduced to less than a minute, efficient MCMC sampling parallelized on a cluster (using \texttt{emcee}, for example).
-However, as it stands without a collaborator, I do not have the expertise nor the time to pursue such an approach.
-Therefore, we are reduced to fitting in the space of $\vg$ and devising ways to deal with the issues.
-
-\section{General problems with the spectral library approach}
-
-Pixel-to-pixel approaches have the potential to be much more powerful than other historical stellar parameter estimation techniques.
-When comparing directly in the data space, we have the ability to add in other physics as we see fit.
-
-because we can fit multiple stars at once (say for unresolved binary or multiple systems), fit for velocity components, etc.
-Many of the most interesting systems for which we would want to determine stellar parameters are in multiple systems anyway, since we can dynamically determine masses for these systems and thus compare to evolutionary models.
-Determine veiling.
-And a perhaps wooden-headed belief that using the full spectrum \emph{should} give us a more accurate and less-biased posterior estimate.
-Such a belief may be foolish and in reality different parts of the spectrum could give completely different parameter estimates.
-
-Many, many methods exist for comparison of stellar spectra, including many ``Bayesian'' approaches in the pixel-to-pixel space.
-The nature of the Bayesian-ness, however, is usually limited to priors on temperature, surface gravity, and metallicity.
-These approaches do not explicitly state the likelihood function that they use, or if they do, it is a $\chi^2$ function.
-
-The more successful of these models use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000 ish). Or, a grid over a narrow range of wavelength space with finely tuned atomic constants.
-
-\textbf{The true fundamental problem with this approach is that the stellar spectra are not a perfect representation of reality}.
-Assumptions about Poisson errors break down, and fits become contaminated by systematics.
-There are types of errors which occur.
-
-FIGURE: 4 Panel Figure, from left to right, showing Class 0 through III errors.
-Use Mg b lines, other big line near Mg b triplet, order 24 line, and then Fe I/II line complex that doesn't fit properly.
-
-\paragraph{Class I} A spectral line appears in the data but not in the model spectrum over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$.
-Such a type of error is perhaps the easiest to understand and to forgive.
-For example, such a line could easily appear in the data due to interstellar absorption or telluric absorption, something which a stellar model could never produce.
-The model could also fail to reproduce a true stellar line due to a missing opacity source.
-
-\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$.
-This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident.
-Presumably this is a product of bad atmospheric structure (temperature or pressure is wrong at some point?) that excites a line when it shouldn't otherwise.
-
-\paragraph{Traditional ways to model Class I and Class II errors} are a form of sigma clipping.
-Mann et al.
-compares the stars to models over a wide range of parameters.
-One can track how these lines might be bad because Class I residuals (data - model)/sigma will always be negative (since we are dealing with absorption lines only) and not change significantly with $\vg$, since we the model can only produce continuum.
-Class II residuals will always be positive and should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines.
-
-If a Class I or II error occurs on a line, that region of the spectrum is relatively useless for determining spectral parameters.
-Authors like Mann et al.
-or Kholeva et al.
-find these regions and simply mask them out of the final computation.
-Kholeva et al., and others approach in an iterative manner.
-Kholeva et al.
-approaches are more sophisticated because they track convergence.
-This is a sigma clipping approach, and while we could go and model these a la Hogg 2010, the approach is not super justified, result is probably not too wrong (similar to actual sigma clipping).
-Given that Kholeva et al.
-track convergence maps, I would place a higher degree of confidence in the answer.
-
-In our later approaches in Sec XX, how might we put a prior on $\Delta \vg$, the range of parameters over which the residuals will be correlated?
-Maybe we could Gibbs sample in the same way as the bad flags.
-We'd have to keep a bunch of spectra with varying $\vg$ in memory to make this comparison easy.
-
-\paragraph{Class III} In truth, this is just a superset of both Class I and Class II.
-Class III errors are when the line strengths between the data and model do not match at a $\vg$ that fits a majority of lines.
-What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, however this parameter combination would result in bad fits for the majority of the other lines.
-Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$.
-
-In addition, it is difficult to determine the ensemble of lines that suffer from Class III errors, since the strength of some lines will be under-predicted while others will be over-predicted.
-How do we know which lines to trust?
-If we take an approach similar to fixing Class I and II errors, and mask out bad regions, it would be very difficult to distinguish Class III errors from \emph{truthful} lines at an incorrect $\vg$.
-For example, if one imagines using a cleaning algorithm as described for Class I and Class II errors, the identification of Class III errors will be sensitive to the starting guesses of $\vg$.
-
-Because these Class III lines can still produce a significant residual spike, they tend to distort and in some cases artificially peak the posterior distribution.
-There is no one parameter combination that will fit all lines perfectly.
-Certainly there is a parameter combination that will produce a reasonable fit to 90\% of the lines to within 3 sigma residuals, but there will still be a fair number of both ``high'' and ``low'' regions.
-These lines need to be identified and dealt with, but how?
-
-How then, should the best fit be determined? Surely some of the remaining lines are ``bad.'' But in this regime, determination of which lines are bad will bias the results in a certain direction.
-
-Also, which lines should one actually fit? The stronger lines? Did the modelers themselves give more attention to the atomic constants which produce the strong lines?
-
-To summarize, the fundamental problem with comparing high resolution optical stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines.
-Deciding how to model the lines is then the difficult part.
-Different groups have sigma-clipped lines that are obviously bad, and this does improve the procedure dramatically.
-In such a case, the procedure of sigma-clipping doesn't gall with me too badly.
-However, the more difficult problem is dealing with Class III errors properly, something that no modeler has mentioned or successfully implemented in a statistically justified manner.
-
-In a spectrum fit, many lines do reasonably fit within the Poisson errors. For consistency, we will denote these as ``Class 0'' lines, where there is no error.
-
-\subsection{Testing the accuracy of stellar libraries}
-Gaia team members are panicking about the upcoming flood of data.
-
-\citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia.
-
-\citep{cbm+05} is the Coehlo grid.
-
-\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}
-
-The parameter vector is comprised of both stellar parameters and calibration parameters $\vt = \{\vstar, \vN \}$.
-Stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}], v \sin i, v_z, A_v, R_\star^2/d^2 \}$ while the data vector $\vD = \fD(\lambda)$ represents a high resolution spectrum.
-For demonstration purposes in this paper, we use a 51-order echelle spectrum from the \emph{TRES} spectrograph at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$), which spans the full optical range ($3800-9000$\AA), though our technique is applicable to any high resolution optical/near-IR spectrum.
-
-FIGURE showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000.
-This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species.
-For example, we could inject bad lines in this space.
-
-\subsection{Processing TRES Data}
-We flux calibrate TRES data, but because the reduction pipeline does not produce a ``sigma spectrum,'' we instead use the Poisson errors on each pixel.
-By taking the raw, still-blazed spectrum, in units of ``counts,'' we can find the Signal-to-Noise ratio by,
+In the manner of Gaussian random processes (GRP), we develop a covariance function
 \begin{equation}
-  S/N = \sqrt{ {\rm cts}}
-\end{equation}
-then the Noise-to-Signal ratio is the inverse of this
-\begin{equation}
-  N/S = \frac{1}{\sqrt{ {\rm cts}}}
-\end{equation}
-then the errors on the flux-calibrated spectrum (in ergs/s/A/cm2) is simply the flux spectrum multiplied by N/S.
-
-\subsection{Applied to a Gaussian likelihood function}
-
-Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}.
-Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis.
-\begin{equation}
- p(\vD | \vt) = {\cal L} \propto \exp \left(-\frac{\chi^2}{2} \right)
+  {\rm cov}(q_i, q_j) = \rho(| i - j|)
 \end{equation}
 \begin{equation}
-  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
+  \Sigma_{ij} = \sigma^2 \rho(|i - j|)
 \end{equation}
-With flat priors on $\vstar$ and $\vN$ (an assumption we will later relax) the logarithm of our posterior probability function is then
+that describes the correlation between the state of the bad data flags.
+The simplest thing to do would be to assume a common correlation length in the spectrum.
+If we say that the typical feature will have three contiguous pixels\footnote{Ideally the covariance function should be a function over $\lambda_i$ in velocity space, since all spectral features should have a similar velocity width.
+For the purpose of keeping the math simple for the moment, we use pixel index and assume the pixels are uniformly spaced.} wrong, then the correlation function would be
 \begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
-  \label{eqn:lnprob}
+  {\rm cov}(q_i, q_j) = \left \{ \begin{array}{cc}
+    1 - \frac{|i - j|}{3} & |i - j| < 3 \\
+    0                                    & \textrm{else}
+\end{array}
+\right .
 \end{equation}
-To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by linearly interpolating in $T_{\rm eff}, \log(g), [{\rm Fe}/{\rm H}]$ space between grid points of instrumentally broadened, oversampled PHOENIX model stellar spectra, rotationally broadened using Fourier techniques, Doppler shifted, extinction corrected, and downsampled using spline interpolation to the exact pixels of the TRES spectrum.
+and the matrix would be
+\begin{equation}
+  \begin{bmatrix}
+    1 & 2/3 & 1/3 & 0 & \cdots & 0\\
+    2/3 & 1 & 2/3 & \ddots & \ddots & \vdots\\
+    1/3 & 2/3 & \ddots & \ddots & \ddots & 0\\
+    0  & \ddots & \ddots & \ddots & 2/3 & 1/3\\
+    \vdots & \ddots & \ddots & 2/3 & 1 & 2/3\\
+    0      & \cdots & 0 & 1/3 & 2/3 & 1\\
+  \end{bmatrix}
+\end{equation}
+Because our correlation function has compact support, we now have a sparse covariance matrix. This can potentially speed up a matrix computation that would otherwise be very slow.
 
-$\sigma(\lambda)$ is the Poisson photon counting errors.
-For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead.
-For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.
+\subsubsection{Correlation of adjacent residuals}
+Not only will the flags $q_i$ be spatially correlated but so will the values of the residuals $R_i$.
+For OBAFGK stars, the residuals are likely to be shaped like Gaussians.
+If $i$ specifies the pixel at the line center of a large residual feature, then the neighboring $i \pm 1,2,\ldots$ pixels will have also have similar residuals but of declining magnitude $|R_{i\pm 1, 2, \ldots} | \lesssim |R_i|$.
+For M stars, there might be a wider correlation length and the shape of the lines might not be well-described by a Gaussian.
 
-The prefactor $k(\lambda | \vN)$ is a calibration functionm that aims to account for errors in the blaze-correction or flux-calibration by multiplying the model by a polynomial.
-The parameters describing this function are the coefficients of the polynomial.
-This approach of accounting for large-scale calibration errors in the spectrum is more justified than the traditional approach of normalizing to the stellar continuum.
-From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (Figure?) flux-calibration function, or IRAF's ``sensitivity function.'' These tests revealed that the largest source of error in flux-calibration was actually in the overall magnitude of individual order, this could be due to slit losses, poor correction due to parallactic angle, or scattered light within the instrument.
-These corrections are smaller than 10\%.
-But more importantly, the corrections in the higher order terms (ie, the shape of the bandpass calibration for each order) are small ($\lesssim 3\%$).
-As described in \S\ref{sec:priors}, we can set priors on these coefficients.
-This will be the optimal mode of operation for our model.
-However, there exists a wealth of archival data that is un-fluxcalibrated, since its primary purpose was for radial velocity determination of FGK stars, to which it is easy to normalize to a continuum.
-In this case, the scatter amongst the calibration can be as large as a factor of 2x.
-This prior information on the calibration parameters motivates our choice of functional parameterization for the nuisance parameters in \S\ref{sec:nuisance}.
+I am new to the field of GRP, but it seems like it might be possible to come up with a covariance function that would describe the predicted heights of the residuals and regress against them.
+The covariance function between $R_i$ could be learned from the spectrum itself.
+Then, the likelihood functions for each mixture component in the model would be custom tailored for the type of error that it expected to encounter.
+For example, Class 0 residuals would be flat while Class I, II, and III residuals would have some sort of Gaussian shape.
+Class I residuals would be negative while Class II residuals would be positive.
+I plan to further investigate covariance functions.
 
-\subsection{Nuisance parameters}
-\label{sec:nuisance}
-In our case, $\vN = \{c_0, c_1, \ldots, c_N\}$ are a set of 4 Chebyshev polynomial coefficients for each order.
-This amounts to a ``re-fluxing'' of the model following the techniques of \citet{elh+06}.
-We apply the re-fluxing to the model rather than attempt to re-flux the data since this operation would also need to re-scale the noise, making $\chi^2$ a complicated function of $k$.
+\subsubsection{Further ideas}
+\begin{itemize}
+  \item \citet[ch. 18]{gcs+04} list several references further describing inference for spatially-correlated data and I plan to work my way through these. 
+  \item Naively, it seems to me like there might be a way to use FFT methods to help speed up the correlation calculation, similar to Gaussian random fields.
+  \item My preliminary experiments with GRP reveal that there is an art in choosing the functional form of ${\rm cov}(i, j)$ and setting the hyperparameters.
+Perhaps these could also be parameters that we sample in.
+I am working through \citet{rw05} and \citet[Ch 21.]{gcs+13} to learn more.
+\end{itemize}
 
-In order to understand the effects of the $k(\lambda | \vN)$ term, we now consider its effects on just one order of the 51-order echelle spectrum. The first four Chebyshev polynomials are 
+\subsection{Hyperparameters}
+\citet{mb13} discuss a technique to deal with systematic errors when combining different data sets for the combined inference of a single problem.
+They use a hyperparameter approach that can artificially inflate the variance of each data set to find a common posterior and allow for correlations between data sets.
+We could set up a similar hyperparameter approach that can tune the likelihood contribution for certain lines, treating each line as they do different datasets.
+If all of the lines were in harmonious agreement, this hyperparameter would undergo ``shrinkage'' \citep{kru10} and tighten the posterior on $\vstar$.
+If the line fits were incompatible, this hyperparameter would increase while also broadening our posterior on $\vstar$.
+
+\subsection{Sampling ideas for the mixture model with spatial correlations}
+If we use multidimensional Gaussian likelihood functions and correlation functions with compact support, it might still be possible to analytically marginalize over the multiple mixture components.
+If we cannot marginalize, then we need alternate sampling methods to explore the parameter space in a computationally efficient manner.
+A stellar model with flags for each pixel would have $\vstar + (N \times M)$ parameters, where for a typical high resolution optical spectrum $N >$ 100,000.
+
+An alternative way to approach this problem is to do Gibbs sampling where we propose a jump in a single $q_{im}$ at a time.
+Such a sample requires none of the previous heavy lifting (spectral grid interpolation, rotational broadening, or pixel interpolation), and so even though there are $\sim$ 100,000 flags to sample, this might be computationally feasible.
+
+In order to ``trim the tree,'' \citet{gcs+04} suggest to start with a crude estimate (such as $k$-means clustering) to tentatively identify which observations have been drawn from which mixture components.
+They also suggest that the expectation-maximize (EM) algorithm could also be used to sample this posterior, although this is an algorithm that I need to research further.
+
+\section{Injection of fake lines}
+As an alternative approach to modeling the regions of the spectrum suffering from systematics, we could attempt to patch the models by injecting fake lines into the spectrum to compensate the errors.
+Each line could be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
+The line strength distribution could be described by a hyper prior, and the FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM.
+We could penalize the model fits by having a prior on the number of bad lines that can be created.
+Given a bad line and the linelist that produced the synthetic models, it may be possible to identify other regions of the spectrum that might be bad.
+For example, if a vanadium line is bad, then we could check to see if the other vanadium lines might also be bad.
+
+Class III errors might be difficult to address in this paradigm since we would be fitting the difference of two lines, but if they are both Gaussians of the same FWHM, then it should still work.
+In M star spectra, rather than adding a Gaussian, we might need to add a larger continuum feature, which might be difficult to parameterize.
+
+\subsection{Piecewise metallicity model}
+In theory, it is possible to fit the profiles of only certain elements to determine the abundances of these elements.
+By masking all non-alpha elements, \citet{hus12} use PHOENIX spectra \citep{hwd+13} to derive alpha abundances of stars.
+Rather than masking, we could assemble a piecewise spectrum where regions of the spectrum might be drawn from a different combination of $\vg$, say with lower metallicity for some lines in order to account for a bad line strength.
+These regions would have to be normalized properly so that the composite spectrum is sufficiently smooth.
+Coupled with a linelist, one could also tweak individual elemental abundances.
+
+\clearpage
+\appendix
+\section{A technique to account for imperfect flux calibration}
+\label{sec:chebyshev}
+
+The first four Chebyshev polynomials are 
 \begin{align*}
@@
 \end{align*}
-In our implementation, we map the full pixel range of a specific order $\lambda \in [\lambda_{\rm min}, \lambda_{\rm max}]$ to $x \in [-1, 1]$.
-By mapping pixel number in a given order, rather than wavelength, we are insensitive to wavelength shifts of the model.
-Then, for a given set of Chebyshev coefficients, $\vN = \{ c_0, c_1, c_2, \ldots, c_N \}$, we have
+In our implementation, we map the full pixel range of a specific order $i \in [i_{\rm min}, i_{\rm max}]$ to $x \in [-1, 1]$.
+By mapping to pixel number rather than wavelength, we are insensitive to Doppler shifts of the model.
+If the spectrum has multiple orders, like echelle spectra, we would have a set of Chebyshev coefficients $\vN = \{ c_0, c_1, c_2, \ldots, c_N \}$ for each order.
 \begin{align}
-  k(\lambda | \vN) &= c_0 T_0(\lambda) \left [1 + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \right ]\\
-  k(\lambda | \vN) &= c_0 \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ]
+  k(i | \vN) &= c_0 T_0(i) \left [1 + c_1 T_1(i) + \ldots + c_N T_N(i) \right ]\\
+  k(i | \vN) \fMi &=  \left [1 + \sum^N_{j = 1} c_j T_j(i) \right ] (c_0 \fMi)\\
   \label{eqn:k_new}
 \end{align}
-When applied to scale the model flux $\fM$, this becomes
-\begin{align}
-  k(\lambda | \vN) \fM &=  \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ] (c_0 \fM)\\
-  k(\lambda | \vN) \fM &= c_0 \fM + c_1 T_1 (c_0 \fM) + \ldots + c_N T_N (c_0 \fM)
-\end{align}
-
-This functional form of $k$ allows $c_0$ to scale by large amounts while the higher order terms (denoted by $c_n$) are perturbations on top of the \emph{scaled} model ($c_0 \fM$).
-However, see \S\ref{sec:gaussian_simplification} for an alternative model when perturbations are small.
-We can rewrite Equation~\ref{eqn:lnprob} as
+This functional form of $k$ allows $c_0$ to scale by large amounts while the higher order terms are perturbations on top of the \emph{scaled} model ($c_0 \fMi$).
+This formulation permits large rescalings up or down without biasing the higher order terms.
+If $c_0$ is expected to be small, we can use a simpler implementation described in \S\ref{sec:gaussian_simplification}.
+Then for the full spectrum
 \begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [1 + \sum^N_{i = 1} c_i T_i(\lambda) \bigr ] c_0 \fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
+  \ln \bigl [p(\vt | \fD) \bigr] \propto - \frac{1}{2} \sum_i \left [\frac{\fDi - \bigl [1 + \sum^N_{j = 1} c_j T_j(i) \bigr ] c_0 \fMi }{\sigma_i} \right ]^2
   \label{eqn:lnprob2}
 \end{equation}
-To simplify the following discussion, we consider a single wavelength $\lambda_i$ and then generalize this to a sum over $\lambda$ later.
-If we expand out the square in Equation~\ref{eqn:lnprob2}, then for a given $\lambda_i$ we have
+To simplify the following discussion, we consider a single pixel $i$ and then generalize this to a sum over all pixels later. We expand out the square in Equation~\ref{eqn:lnprob2}
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + \sum_{i=1}^N c_i T_i \right ]^2 \right ]
- \label{eqn:lnprob_lambda}
- \end{equation}
-where $\sigma$, $\fD$, $\fM$, and $T_{n,m}$ are all evaluated at $\lambda_i$. We can further expand this to
-\begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
+  \ln \bigl [p(\vt | \fDi) \bigr] \propto - \frac{1}{2 \sigma_i^2} \left [ \fDi^2 - 2 \fDi c_0 \fMi \left [1+ \sum_{j=1}^N c_j T_j(i) \right ]+ c_0^2 \fMi^2 \left [1 + 2 \sum_{j=1}^N c_j T_j(i) + \sum_{j=1}^N \sum_{k=1}^N c_j T_j(i) c_k T_k(i) \right ] \right ]
   \label{eqn:expanded}
  \end{equation}
-
-To simplify this math, we can rewrite the Chebyshev coefficients and polynomials as column vectors $\vc$ and $\vec{T}(\lambda)$, respectively
+To simplify the notation, we can rewrite the higher order Chebyshev coefficients and polynomials as column vectors $\vc$ and $\vec{T}(i)$, respectively
 \begin{equation}
@@
   \hspace{3cm}
-\vec{T}(\lambda) = 
+\vec{T}(i) = 
 \begin{bmatrix}
-T_1(\lambda)\\
-T_2(\lambda)\\
+T_1(i)\\
+T_2(i)\\
 \vdots\\
-T_N(\lambda)\\
+T_N(i)\\
 \end{bmatrix}
@@
 \begin{equation}
-  k(\lambda | \vN) = c_0 \left [ 1 + \sum^N_{i = 0} c_i T_i \right ] = c_0 \left[ 1+ \vec{T}^\trans \cdot \vc \right]
+  k(i| \vN) = c_0 \left [ 1 + \sum^N_{j = 0} c_j T_j(i) \right ] = c_0 \left[ 1+ \vec{T}^\trans \cdot \vc \right]
 \end{equation}
@@
 \begin{equation}
-  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
+  {\bm W}(i) = \vec{T} \cdot \vec{T}^\trans = 
   \begin{bmatrix}
@@
 \begin{equation}
-  \sum_{i =1}^N \sum_{j=1}^N c_i T_i c_j T_j = \vc^\trans \cdot {\bm W} \cdot \vc
+  \sum_{j =1}^N \sum_{k=1}^N c_j T_j(i) c_k T_k(i) = \vc^\trans \cdot {\bm W} \cdot \vc
 \end{equation}
@@
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
+  \ln \bigl [p(\vt | \fDi) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fMi^2}{\sigma_i^2}  \vc^\trans {\bm W} \vc + \left (\frac{- c_0^2 \fMi^2 + c_0 \fDi \fMi}{\sigma_i^2} \right ) \vec{T}^\trans \vc -  \left( \frac{c_0^2 \fMi^2 - 2 c_0 \fDi \fMi + \fDi^2}{2 \sigma_i^2} \right)
 \end{equation}
+Because matrix multiplication is associative, we can sum ${\bm W(i)}$ and $\vec{T}^\trans(i)$ across all pixels and define
+\begin{align}
+  {\bm A} &= c_0^2 \sum_{i} \frac{\fMi^2}{\sigma_i^2} {\bm W}(i) \\
+  \vec{B} &= \sum_{i} \frac{-\fMi^2 c_0^2 + \fDi \fMi c_0}{\sigma_i^2} \vec{T}(i)\\
+  g &= -\frac{1}{2} \sum_{i} \frac{c_0^2 \fMi^2 - 2 c_0 \fDi \fMi + \fDi^2}{\sigma_i^2} 
+\end{align}
+Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\fD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2} \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \sum_{i=1}^N c_i T_i  -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
+  p(\vt | \fD) = p(\vstar, \vN | \fD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
+  \label{eqn:lnprob_matrix}
 \end{equation}
+Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters.
+Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically \citep{sgd+09}, and then sample from the marginalized distribution.
 \begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2}  \vc^\trans {\bm W} \vc + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \vec{T}^\trans \vc -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
+  p(\vstar, c_0 | \fD) = \int p(\vstar, c_0, \vc\; | \fD) \dd \vc
 \end{equation}
+\begin{align}
+  p(\vstar, c_0 | \fD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
+  p(\vstar, c_0 | \fD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
+\end{align}
+By doing this marginalization we drastically reduce the dimensionality of our parameter space, especially if our echelle spectrum has multiple orders.
 
-Because matrix multiplication is associative, we can sum ${\bm W(\lambda)}$ and $\vec{T}^\trans(\lambda)$ across all $\lambda$, and define
-\begin{align}
-  {\bm A} &= c_0^2 \sum_{\lambda} \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) \\
-  \vec{B} &= \sum_{\lambda} \frac{-\fM^2(\lambda) c_0^2 + \fD(\lambda) \fM (\lambda) c_0}{\sigma^2(\lambda)} \vec{T}(\lambda)\\
-  g &= -\frac{1}{2} \sum_{\lambda} \frac{c_0^2 \fM^2(\lambda) - 2 c_0 \fD(\lambda) \fM(\lambda) + \fD^2(\lambda)}{\sigma^2(\lambda)} 
-\end{align}
-Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
-\begin{align}
-  \ln \bigl [p(\vstar, \vN | \vD) \bigr] &\propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g\\
-  p(\vt | \vD) = p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
-  \label{eqn:lnprob_matrix}
-\end{align}
-Here we have the full posterior probability distribution assuming flat priors.
-This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$.
-Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters.
-This marginalization encapsulates the uncertainty due to the inference on the nuisance parameters.
-Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically, and then sample from the marginalized distribution.
-\begin{equation}
-  p(\vstar, c_0 | \vD) = \int p(\vstar, c_0, \vc\; | \vD) \dd \vc
-\end{equation}
-The analytic multi-dimensional Gaussian integral with linear term \citep{sgd+09} yields 
-\begin{align}
-  p(\vstar, c_0 | \vD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
-  p(\vstar, c_0 | \vD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
-\end{align}
-Where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$.
-By doing this marginalization we have reduced the dimensionality of our posterior space from $200+$ parameters ($\sim 7$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 60$ ($\sim 7$ stellar parameters and one $c_0$ for each of the 51 echelle orders).
-This enables a dramatic speedup in the MCMC sampling.
-For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than XX\%), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters.
-This approach is detailed in \S\ref{sec:gaussian_simplification}.
-
-\subsection{Including Gaussian priors on nuisance parameters}
+\subsection{Including Gaussian priors on the nuisance parameters}
 \label{sec:priors}
-Using this formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters.
-From our flux-calibration experiment, we determined that the overall scaling could range by 10\%, while the higher order corrections were less than 3\%.
+Using this polynomial formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters, determined by observing several spectrophotometric standard stars.
 Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two.
 This means we should be using a log-normal prior on $c_0$
-\begin{align}
-  p(c_0) &= \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) \\
-  \ln p(c_0) &= \ln \left( \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \right) - \frac{(\ln c_0)^2}{2 \sigma_{c_0}^2}
-\end{align}
-centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 f$), we assume Gaussian priors centered about 0
+\begin{equation}
+  p(c_0) = \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right)
+\end{equation}
+centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 \fMi$), we assume Gaussian priors centered about 0
 \begin{equation}
@@
 \end{equation}
-and $\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient, in our case $\sigma_i = \sigma_{c_n} \approx 0.03$.
+and $\sigma_j$ represents the width of the Gaussian prior on the j-th Chebyshev coefficient.
 $\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction.
-Now, we can write the posterior with priors as
+Now, we can write the full posterior with priors as
 \begin{equation}
-  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\; | \vD) \propto p( \vD | \vstar, c_0, \vc\; ) p(\vc\;) p(\vstar) p(c_0) 
-\end{equation}
-\begin{equation}
-  p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
+  p(\vstar, \vN | \fD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
   \label{eqn:posterior_prior}
 \end{equation}
-we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$
-\begin{align}
-  -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) &= \frac{1}{2}\left ( -\vc^\trans {\bm D} \vc + \vc^\trans {\bm D} \vec{\mu} + \vec{\mu}^\trans {\bm D} \vc - \vec{\mu}^\trans {\bm D} \vec{\mu} \right )\\
-  &= -\frac{1}{2} \vc^\trans {\bm D} \vc + ({\bm D} \vec{\mu})^\trans \vc - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}
-\end{align}
-then we can rewrite
+We can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$ rewrite
 \begin{align}
@@
    \boxed{
-  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
+  p(\vstar, \vN | \fD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
 }
+\label{eqn:lnprob_lognormal}
 \end{equation}
-This probability function appears in the code as \texttt{lnprob\_lognormal} and has $\sim 7$ stellar dimensions and $\sim 200$ nuisance dimensions.
-
-Using the previous multi-dimensional integral, we can marginalize over the $\vc$ including the Gaussian priors
+and the margin form is 
 \begin{equation}
   \boxed{
-  p(\vstar, c_0 | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) p(c_0)
+  p(\vstar, c_0 | \fD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) p(c_0)
 }
+\label{eqn:lnprob_lognormal_marg}
 \end{equation}
-where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$.
-This probability function appears in the code as \texttt{lnprob\_lognormal\_marg} and has $\sim 7$ stellar dimensions and $\sim 50$ nuisance dimensions.
-If we increase the permissible degree of re-fluxing by increasing $\sigma_i$, this will have the effect of smearing out or widening the posterior on a given $\vstar$.
+If one uses MCMC to sample the marginalized posterior, but desires a flux-corrected spectrum for the purposes of inspecting the residuals, it is still possible to do.
+We can draw a random sample $\vstar$ from the MCMC chain and then resample the unmarginalized posterior (Eqn~\ref{eqn:lnprob_lognormal}) conditional on $p(\vN | \vstar, \fD)$.
+Now, the combination of $\vstar$ and $\vN$ completely describes a flux-corrected spectrum.
 
-\subsection{Sampling a un-flux calibrated model}
-This invalidates the previous section on normalization. However, for corrections of less than XX\%, this approximation is better than XX\% accurate.
-
-For the flux-calibration exercise, the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$.
-
-The plot in plots/priortests shows that the log-normal prior is the correct choice.
-Since we care most about the inference on F, we want to know that the choice on prior of $c_0$ is not biasing the final outcome, while a Gaussian prior might if $\sigma_{c_0}$ was large.
-How to quantify this error and create a dividing line?
-For 5\% error?
-In peak?
-Do these plots need to be in the paper?
-
-Compare distributions by doing a K-S test on the cumulative distributions between Gaussian and Log-normal?
-
-\subsection{Hierarchical modelling}
-Can have a $c_1$, etc, that are marginalized over but we sample $c_0$ in a hierarchical manner which informs the overall flux-level.
-
-\subsection{Sampling with emcee}
-Citation to DFM \citep{fhl+12}, GW \citep{gw10}, and CosmoHAMMER \citep{asa+13}.
-
-\section{Moving beyond a perfect stellar model}
-\label{sec:residuals}
-
-\subsection{Additional sources of noise beyond Poisson}
-Presumably we shouldn't have to add an additional source of ``noise'' to the fit, because our nuisance parameters will find an exact flux calibration.
-
-Interpolation errors. Quantified via Husser et al. Ph.D. thesis as well as custom tests.
-
-\subsection{Noise, Masking, Residuals, and Alternative likelihood functions}
-In the future, we could modify our likelihood function to use a Student-t statistic instead of a Gaussian, such that it would be more robust in the presence of outlier pixels, however this does not seem necessary at the moment.
-
-\subsection{Masking}
-
-We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).
-We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid.
-
-Section on how to choose masked regions, testing the PHOENIX grid to see where the derivative is highest and thus gives the best leverage on stellar parameters.
-
-Iterative masking to deal with Class I and II errors a la ULySS. They use a $\kappa - \sigma$ routine, where they iteratively sigma clip (Website link).
-
-\section{Gallery of likelihood functions}
-
-If you do a straight up $\chi^2$ fit to the data, you can get posteriors that look like $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
-This would be better than that obtained by astroseismology!
-Clearly there is something going wrong here.
-
-Lorentzian, linear combination of Gaussians, Sivia w/ prior on sigma. $\chi^2_Q$ model.
-
-These likelihood functions seem to be an improvement over the straight $\chi^2$ function, in the sense that they give parameter estimates that are
-
-\subsection{Bad data model}
-The bad data model \citep{pre97,dbl10} is a powerful framework for forward-modelling of data, allowing the modeler to account for the probability of outliers.
-
-Now, we have a $q_i$ for each data point, such that for the entire data set we have an ensemble of $\qN$.
-If this ensemble of $\qN$ is chosen appropriately, and the good data points are flagged as good and the bad data points are flagged as bad, the likelihood will be maximized.
-Written out explicitly, this is
-\begin{multline}
-  {\cal L} = p({\bm D} | m,b,\qN,Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma^2} \right] \right ]^{q_i}  \times \\
-  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma_{y_i}^2)}} \exp \left[ - \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{y_i}^2)} \right]
-  \right ]^{1 - q_i}
-\end{multline}
-
-using a Bernoulli prior
-
-\begin{equation}
-  p(q_i | P_b) = [1-P_b]^{q_i} P_b^{[1-q_i]}
-\end{equation}
-
-\begin{equation}
-  p(\qN | P_b) = \prod_{i=1}^N [1-P_b]^{q_i} P_b^{[1-q_i]}
-\end{equation}
-
-For this data point, we can integrate (in this case, discrete sum) over the two states of $q_i$: $q_i = 1$ and $q_i = 0$. 
-\begin{multline}
-  p(m, b, P_b, Y_b, V_b | D_i)  =  \int p( m, b, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
-  = p(m, b, q_i=1, P_b, Y_b, V_b | D_i) + p(m, b, q_i=0, P_b, Y_b, V_b | D_i) \\
-  = \Bigl[ p_{\rm good}(D_i | m,b)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(m,b,P_b, Y_b,V_b) 
-\end{multline}
-
-\begin{multline}
-  p(m,b,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i | m, b)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(m,b,P_b, Y_b, V_b)\\
-  \boxed{
-    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_{yi}^2}} \exp \left[ - \frac{(y_i - m x_i - b)^2}{2 \sigma_{yi}^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma_{yi}^2)}} \exp \left[- \frac{(y_i - Y_b)^2}{2 (V_b + \sigma_{yi}^2)} \right] \right\}
-  \times p(m,b,P_b,Y_b,V_b)}
-\end{multline}
-
-In it's simplest form (Hogg 2010), the bad data model results in a linear combination of Gaussians.
-
-A benefit of the bad data model over a chosen likelihood function is that it is motivated.
-It may still result in the linear combination of Gaussians (nice), but preserves real parameter correlations.
-You are not forcing the same mean.
-
-\subsection{Spatial correlations of pixels}
-We can do more.
-The bad model of the previous section assumes that each data point is independent.
-For example, the spectral lines are actually bad, and this means that if a single line is wrong, not just the core of the line (which may be a 10 - 20 sigma residual) will be wrong, but also the wings of the line, which may be only 5 or 3 sigma wrong.
-
-Now, let's come up with a model which includes spatial (wavelength) correlations in the bad data model.
-
-How could we accomplish this?
-If we identify one point as bad, then the flags around it will have a correlation length, ie, the prior on $q_i$ will be bad.
-This means that for however many nearby pixels that will be bad (most likely a constant amount for uncrowded lines), then this means that we could still integrate over the qis, it's just that we would now have to take a contribution from the nearest 5 pixels, say.
-That central pixel would then mark the ``line'' that is bad.
-
-That is one way to mark the ``badness of the line,'' that the qi are correlated. But we also know that the $y_i$ will be correlated too.
-
-This correlation length will probably depend on the type of star, and the contribution of the band.
-For example, if they're simple FGK stars, then the bad lines might be singular.
-But if they are M stars, then there is a chance that there might be a large ``pseudo-continuum'' region that might be incorrect.
-
-Once we have identified the center of the bad line, then there will be a decrease in the nearby pixels, similar to a Gaussian shape.
-Presumably we could come up with a correlation in this space too.
-(or, it could simply be that the correlation between $y_i$ is flat in the case of an M star.
-
-This might actually save the marginalization. We could still do the backwards ``what is the probability a given line is bad.''
-
-For Class I and II errors, we could have a hyper parameter describing the width of each line, and another hyper parameter describing the (distribution of) depth. These could also be sampled in.
-
-For Class III errors, describing how line strengths are off makes sense for stars with continuum, but as we move to later type stars (K stars) that do not have a nice continuum, we may run into trouble.
-For the later stars (M dwarfs), there might be a whole region that is incorrect.
-
-There are different ways to describe a correlation over pixels. Gaussian random fields. Autocorrelation function.
-
-\subsection{Simple correlation between pixels}
-For a simple correlation, say that if a given $q_i$ is bad, then there is also a high probability that the neighboring pixels are also bad.
-Say we know from inspection that it seems as though bad lines knock out 5 pixel chunks.
-
-Caveat: I think this assumes that bad lines will not occur within the same spread of each other. Maybe we should include a $P_B$ prior on the good counts too?
-
-\begin{equation}
-  p(q_i = 0 | q_{i \pm 5} = 0) = P_B
-\end{equation}
-
-\begin{align}
- p(q_i = 0 | q_{i \pm 5} = 0) &= 1/5 \\
- q_{i \pm 4} &= 2/5 \\
- q_{i \pm 3} &= 3/5 \\
- q_{i \pm 2} &= 4/5 \\
-\end{align}
-
-This might need more work with $P_B$.
-
-Then the probability functions would be
-\begin{multline}
-  p(D_{i + j} | \vt, q_i =0) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=0) + \\
-    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 0) \bigr ]
-\end{multline}
-
-\begin{multline}
-  p(D_{i + j} | \vt, q_i = 1) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=1) + \\
-    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 1) \bigr ]
-\end{multline}
-
-The marginalization could still be done analytically, but now the likelihood function involves a correlation between adjacent pixels.
-Could we use a sparse array for this?
-Or, we could evaluate all of the pixels, and then have the final likelihood function involve some correlation terms between the lnprobs evaluated at each pixel.
-We could do this using the logarithm so that everything is additive.
-
-Or, what if we evaluated the FFT of the $\log(p(D_i))$ and just the autocorrelation in Fourier space?
-
-\subsection{Correlation between $y_i$}
-Not only will the flags be spatially correlated (if a point is bad, the nearby $\sim$ 5 points are also likely to be bad) but for OBAFGK stars, the lines are likely to be Gaussian shape.
-That is if $q_i$ specifies the center of the line, then the neighbooring $y_{i+1}$ or $y_{i-1}$ will be of lower residual height.
-This means that the inside of the likelihood function should also have spatial correlations.
-
-Could this correlation be described by a Gaussian random process?
-
-\subsection{Different flags and weights}
-Instead of a good or bad flag, what if we had a flag that would identify which class of error the line became? For example we could have the option of $q_i = \{0, 1, 2, 3\}$, where $q_i=0$.
-
-How do we account for the fact that some parts of the spectrum carry much more information about $\vg$ than other parts of the spectrum?
-Is it important to weight this parts beforehand?
-Or should this contribution to the posterior naturally weight itself, the way certain points might heavily contribute to the slope and intercept of a linear fit?
-
-If so, we could imagine having a Gibbs sampler on the weights of each pixel.
-
-Proper Class III modeling is very important, following the discussion of \citep{mga13} Mann et al, and see recent example from Rajpurohit 2103.
-Some features will not fit perfectly, especially at the lower temperature scale.
-However, it's important to still take these features int account with the modelling.
-They also use a $\chi^2$ likelihood function, but perhaps even less appropriate is that they draw error bars by saying that the $\chi^2$ value changes by 5\%.
-In reality, they are checking the fits by eye and saying whether or not a spectrum matches.
-This approach is no more justified than a by-eye spectral classification.
-
-\subsection{Sampling such a model}
-If our bad model is sufficiently complex that we are not able to analytically marginalize over the $q_i$, it is still possible to sample in this huge parameter space.
-A stellar model with flags for each pixel would have $\vstar + N$ parameters, where for a typical optical spectrum $N >$ 100,000.
-Even using an ensemble sampler, sampling in this space would take a tremendous amount of time to converge.
-
-An alternative way to approach this problem is to do Gibbs sampling in the flags.
-It may be possible to do do Gibbs sampling where we draw from the conditional probability distribution, or at the very worst we could still do a Gibbs sampler with a Metropolis-Hastings backend, where we propose a jump in a single $q_i$ at a time.
-Such an sample requires none of the previous heavy lifting (spectral grid interpolation, FFT, or pixel interpolation), and so even though there are $\sim$ 100,000 flags to sample, this is likely to be computationally feasible.
-
-\section{Injection of fake lines}
-
-An alternative approach is instead of modelling the systematically wrong regions of the spectrum, actually add in lines to the spectrum.
-This can be thought of as correcting the models.
-We could add in a set number of lines.
-
-Presumably, each line would be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
-Strength could be described by a hyper prior, and FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM (in velocity space).
-
-In later type spectra, this might actually need to be a larger continuum feature that is added.
-
-We could penalize the model fits by having a hyperprior on the number of bad lines that can be created.
-
-In theory, it is possible to fit the profiles of only certain elements to determine the abundances of certain elements (see citation in conference proceedings of Husser).
-
-It may be possible to use a linelist from the group that produces the PHOENIX spectra, which describes the element that gave the highest contribution to the opacity in that bin.
-Now, if the vanadium lines are bad, then we could have a prior that the other vanadium lines might also be bad.
-We could assemble a piecewise spectrum, where certain pixels are drawn from a different combination of $\vg$, and these pixels had an associated normalization constant so that the piecewise transition was smooth (if, say the bolometric temperature was different).
-
-Class III errors might be difficult to fit in this approach. However, if the line width does not change dramatically, then we should be able to account for things this way.
-
-\section{Validation tests}
-Comparison to main sequence stars of HAT-P-9, and WASP-14, from the TRES archive \citep{tfs+12}.
-
-Comparisons using HIRES data?
-
-\acknowledgments
-IC would like to graciously acknowledge Gregory Green, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy.
-This research made use of Astropy, a community-developed core Python package for Astronomy \citep{art+13}.
-
-\appendix
-
-\section{Simplification for small $c_0$ perturbations}
+\subsection{Simplification for small $c_0$ perturbations}
 \label{sec:gaussian_simplification}
 If the correction in $c_0$ is small, we can use an alternate formulation of $k$ and put Gaussian priors on all nuisance parameters.
-This allows analytic marginalization over all nuisance parameters (including $c_0$), and reduces the dimensionality of the posterior to only the $\sim 7$ stellar parameters, $\vstar$.
-
+This allows analytic marginalization over all nuisance parameters (including $c_0$), and reduces the dimensionality of the posterior to only the stellar parameters, $\vstar$.
 \begin{align}
@@
 \end{align}
-Following \S\ref{sec:nuisance}
-\begin{equation}
-  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ]\fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
-\end{equation}
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM\, \vec{T}^\trans \vc \; + \fM^2 \, \vc^\trans {\bm W}  \vc \bigr ]
-\end{equation}
 \begin{align}
@@
 \end{align}
-Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. If we include Gaussian priors, then we have 
+If we include Gaussian priors, then we have 
 \begin{align}
@@
    \boxed{
-  p(\vstar, \vN | \vD) = p(\vstar, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(\vc)
+  p(\vstar, \vN | \fD) = p(\vstar, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(\vc)
 }
 \end{equation}
-
-Using the previous multi-dimensional integral, we can marginalize over all $\vc$ including the Gaussian priors
+and the marginal form is
 \begin{equation}
   \boxed{
-  p(\vstar | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
+  p(\vstar | \fD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
 }
@@
 From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$.
-Beyond this, one should switch to this framework, for example to correct for un-flux calibrated data.
 

### 2014-03-16 gauss_rand_proc.py
Commit before merging OOP to master.
Rewrote grid_tools.py and model.py and spectrum.py to use object
oriented programming.

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+wls = np.load("data/WASP14/WASP-14_2009-06-15_04h13m57s_cb.spec.flux.wls.npy")
+fls = np.load("data/WASP14/WASP-14_2009-06-15_04h13m57s_cb.spec.flux.fls.npy")
+sigmas = np.load("data/WASP14/WASP-14_2009-06-15_04h13m57s_cb.spec.flux.sigma.npy")
+
+ind = (wls[22] >= 5140) & (wls[22] <= 5190)
+wl = wls[22][ind]
+fl = fls[22][ind]
+sigma = sigmas[22][ind]
+std = np.std(fl)**2
+print(std)
+
+xind = (wl >= 5170.9) & (wl <= 5171.3)
+wl2 = wl[~xind]
+fl2 = fl[~xind]
+
+l = 0.17 #correlation length
+nsamples = 10
+
+def func(x0i, x1i, x0v=None, x1v=None):
+    return std * np.exp(-0.5 * np.abs((x0v[x0i] -x1v[x1i])/l)**2)
+
+#cov = np.fromfunction(func, (npoints,npoints), xv=np.arange(npoints), yv=np.arange(npoints),dtype=np.int)
+
+
+#training set
+#x = np.linspace(0,10,num=10)
+#y = np.cos(x/2.)
+#std = np.std(y)**2
+
+x = wl2
+y = fl2
+#plt.plot(x,y)
+#plt.show()
+
+xs = np.arange(5170.5,5171.5,0.05)
+#xs = np.linspace(0,5, num = 10)
+
+
+Kxx = np.fromfunction(func, (len(x),len(x)), x0v=x, x1v=x, dtype=np.int)
+Kxxs = np.fromfunction(func, (len(x),len(xs)), x0v=x, x1v=xs, dtype=np.int)
+Kxsx = np.fromfunction(func, (len(xs),len(x)), x0v=xs, x1v=x, dtype=np.int)
+Kxsxs = np.fromfunction(func, (len(xs),len(xs)), x0v=xs, x1v=xs, dtype=np.int)
+#print(Kxx.shape, Kxxs.shape, Kxsx.shape, Kxsxs.shape) 
+
+var = np.average(sigma)**2
+Kxx += var * np.eye(len(x))
+
+inv = np.linalg.inv(Kxx)
+#print(inv)
[…]

### 2014-03-16 tests/profile_evaluate.py

@@ -1,0 +1,32 @@
+import numpy as np
+
+from StellarSpectra.model import Model
+from StellarSpectra.spectrum import DataSpectrum
+from StellarSpectra.grid_tools import TRES, HDF5Interface
+
+myDataSpectrum = DataSpectrum.open("/home/ian/Grad/Research/Disks/StellarSpectra/tests/WASP14/WASP-14_2009-06-15_04h13m57s_cb.spec.flux", orders=np.array([22]))
+myInstrument = TRES()
+myHDF5Interface = HDF5Interface("/home/ian/Grad/Research/Disks/StellarSpectra/libraries/PHOENIX_submaster.hdf5")
+
+myModel = Model(myDataSpectrum, myInstrument, myHDF5Interface, stellar_tuple=("temp", "logg", "Z", "vsini", "vz", "logOmega"), 
+                cov_tuple=("sigAmp", "logAmp", "l"))
+
+params = {"temp":6200, "logg":4.0, "Z":-0.2, "vsini":10, "vz":15, "logOmega":-22}
+myModel.update_Model(params)
+myModel.update_Cheb(np.array([0, 0, 0]))
+
+def eval0():
+    myModel.evaluate()
+
+def eval1():
+    myModel.evaluate()
+
+def main():
+    #print("evaluating 0")
+    #eval0()
+    print("evaluating 1")
+    myModel.update_Cov({"sigAmp": 1, "logAmp":-15 , "l":1})
+    eval1()
+
+if __name__=="__main__":
+    main()

### 2014-03-16 tests/test_model.py

@@ -18,18 +18,18 @@ class TestModel:
         self.DataSpectrum = DataSpectrum.open("/home/ian/Grad/Research/Disks/StellarSpectra/tests/WASP14/WASP-14_2009-06-15_04h13m57s_cb.spec.flux", orders=np.array([21,22,23]))
         self.Instrument = TRES()
         self.HDF5Interface = HDF5Interface("/home/ian/Grad/Research/Disks/StellarSpectra/libraries/PHOENIX_submaster.hdf5")
-        self.model = Model(self.DataSpectrum, self.Instrument, self.HDF5Interface, ("temp", "logg", "Z", "alpha", "vsini", "vz", "Av", "log_Omega"))
+        self.model = Model(self.DataSpectrum, self.Instrument, self.HDF5Interface, ("temp", "logg", "Z", "alpha", "vsini", "vz", "Av", "logOmega"), cov_tuple=("sigAmp", "logAmp", "l"))
 
     def test_evaluate(self):
         print(self.model.evaluate())
 
     def test_update_model(self):
-        self.model.update_Model({"temp":6102, "logg":3.9, "Z":-0.5, "alpha":0.2, "vsini":6, "vz":13.7, "Av":0.0, "log_Omega":-19.7})
+        self.model.update_Model({"temp":6102, "logg":3.9, "Z":-0.5, "alpha":0.2, "vsini":6, "vz":13.7, "Av":0.0, "logOmega":-19.7})
         print(self.model.evaluate())
 
     def test_no_alpha(self):
-        model = Model(self.DataSpectrum, self.Instrument, self.HDF5Interface, ("temp", "logg", "Z", "vsini", "vz", "Av", "log_Omega"))
-        model.update_Model({"temp":6102, "logg":3.9, "Z":-0.5, "alpha":0.2, "vsini":6, "vz":13.7, "Av":0.0, "log_Omega":-19.7})
+        model = Model(self.DataSpectrum, self.Instrument, self.HDF5Interface, ("temp", "logg", "Z", "vsini", "vz", "Av", "logOmega"), cov_tuple=("sigAmp", "logAmp", "l"))
+        model.update_Model({"temp":6102, "logg":3.9, "Z":-0.5, "alpha":0.2, "vsini":6, "vz":13.7, "Av":0.0, "logOmega":-19.7})
         print(model.evaluate())
 
 
@@ -38,8 +38,8 @@ class TestModel:
         print(self.model.evaluate())
 
[…]

### 2014-03-16 paper/notes.tex
Commit before merging OOP to master.
Rewrote grid_tools.py and model.py and spectrum.py to use object
oriented programming.

\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

\usepackage{hyperref}
%\usepackage{graphicx}
%\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{bm}

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{f_{\rm M}}
\newcommand{\fD}{f_{\rm D}}
\newcommand{\vD}{\vec{D}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}

%\slugcomment{}
%\shorttitle{}
%\shortauthors{}

\begin{document}

\title{A method for inference of fundamental stellar parameters using high resolution spectra}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} et al.
%Author2\altaffilmark{2},
}

\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street MS 10, Cambridge, MA 02138}
%\altaffiltext{2}{Institution 2}
\email{iczekala@cfa.harvard.edu}

\begin{abstract}
  Matching observed spectra of stars against a grid synthetic spectra is a popular approach to determine fundamental stellar parameters such as effective temperature, surface gravity, and metallicity. However, a fundamental problem with this approach is that the synthetic spectra contain systematic errors such as wrong or missing spectral lines. In the presence of such errors, a simple $\chi^2$ fit between spectra will likely yield biased parameters and incorrect uncertainty estimates. This behaviour is analogous to the way a few strong outliers might severely bias the fit to a straight line. These spectral outliers are difficult to identify \emph{a priori}; procedures such as sigma clipping might be prone to their own systematic effects and discard useful information. We outline an approach to probabilistically identify and account for spectral features that deviate from reality. Such an approach is most needed when operating at high spectral resolution ($R >$ 10,000), since this is when systematic deviations begin to appear most striking, however our technique is still valid for lower resolution spectra as well. We also outline some approaches for spectral fits to brown dwarfs and planets, where instead of spectral lines there may be systematic deviations of a larger swath of the spectrum.
\end{abstract}

%\keywords{}

%\begin{figure}[htb]
%\begin{center}
%\includegraphics{file}
%\caption{}
%\label{fig:}
%\end{center}
%\end{figure}

\section{Introduction and review of existing methods}

The fundamental parameters of stars---mass, luminosity, effective photospheric temperature, surface gravity, and metallicity--are of supreme interest to astrophysicists for a wide variety of reasons. A salient example is that most newly discovered exoplanets must necessarily their mass and radius measured relative to the properties of their host stars (e.g. \citealt{tfs+12,blj+12,ssm+13}) and uncertain stellar parameters translate into uncertain planet properties \citep{kan14}. The properties of stars are of course also important in their own right because they inform us about our ideas of stellar evolution \citep{dm97, bca+02}. Because stars are the fundamental units of the cosmos, techniques to infer the properties of stars date back to the dawn of astronomy. We briefly review some of the modern and popular techniques in the literature. Approaches can be broadly classified into techniques that compare stars by some broadly defined indices and techniques that attempt to forward-model an observed stellar spectrum by doing a pixel-to-pixel comparison.

\subsection{Classification techniques}
Many of existing techniques were written to analyze low to medium resolution ($R < $ 10,000) spectra. I am calling them similarity techniques because they focus on determining and recognizing patterns of spectra useful in classification.

\paragraph{Line indexes} By cataloguing the equivalent widths of chosen spectral lines into indexes, one can correlate the index with stellar properties and identify features that are sensitive temperature, surface gravity, or metallicity. Then, one matches the indecies of a new star against these relationships and derives a set of stellar parameters. A recent example using M dwarfs is \citet{nci+14}. This technique works, but there are many things that could be improved. First, the selection of which lines to measure is arduous and based upon expert knowledge of previously catalogued stars. This is both a disadvantage and an advantage. On one hand, by choosing a subset of lines to fit, one may be unwittingly throwing away useful information. On the other hand, it allows the modeller to discard regions of the spectrum that do not follow a simple trend with stellar parameters. Second, while measuring equivalent widths is easy to do, this discards any useful information contained in the shape of the line profile.

\paragraph{Principal component analysis} PCA reduces the dimensionality of the data space and searches for the most salient features of the spectra \citep{ptb14}. How does one properly estimate parameter uncertainties in such an approach? It seems like this approach exists because it is an easy thing to do.

Matrix inversions: MATISSE \citep{rbd06}. This method is similar to PCA in that it derives basis vectors that describe the fundamental parameters of the star. In the PCA analysis, the principal components do not necessarily correspond to true physical quantities. The MATISSE algorithm determines weights that ensure the components are similar to the stellar parameters. They only test the algorithm against input synthetic stellar spectra to detect biases in the parameters, and against the Sun and Arcturus. They do acknowledge that there will be systematic deviations between the synthetic spectra and real spectra, but they do not account for the possible inflation of error estimates.

Given the inherent systematic uncertainties involved in each method, some groups elect to combine the inference from different algorithms, such as (SLOAN SSPP) \citep{lbs+08}. This approach includes machine learning along with many of the previously mentioned techniques.

\paragraph{Machine learning}
Many machine learning techniques exist and classify spectra by comparing patterns between observed spectra and a training set of spectra (which could be real spectra of stars classified by an expert or a synthetic library that has known $\vg$). However, my first impression is that these tools are designed to handle large collections of spectra from stellar surveys such as SDSS \citep{sam+10}. In our immediate project involving T Tauri stars, we only have a few ($\lesssim 20$) stars that we wish to study in exquisite detail, incorporating specific parameters such as $v \sin i$, extinction, and veiling. However, I am a novice when it comes to machine learning so I may be missing something here.

\subsection{Pixel to pixel techniques using synthetic spectral libraries}
\label{sec:pix}
Many high quality synthetic spectral grids are now available. Nearly all synthetic stellar libraries are parameterized by the ``fundamental'' stellar parameters of effective temperature, surface gravity, metallicity, and alpha enhancement. We define this combination as 
\begin{equation}
  \vg = \{T_{\rm eff}, \log(g), \Z, \A \}
\end{equation}
Naively, it makes sense to compare an observed spectrum against spectra from the synthetic grid. Using some goodness of fit statistic one can identify the $\vg$ that best reproduce the observed spectrum.

The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
\begin{equation}
  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
\end{equation}
where the best model is defined as one that minimizes the residuals in some sense. We will discuss other goodness of fit functions soon. Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda | \vt)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda | \vt)$.

Most pixel-to-pixel techniques use a $\chi^2$ likelihood function and seek to minimize the square of the residuals
\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
\end{equation}
It appears that for many techniques, $\sigma(\lambda)$ is the Poisson photon counting errors. For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead. For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.

\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations. Any spectral library can be used as a backend, including empirical spectral libraries such as the one from ELODIE \citep{psk+07}. They explicitly take into account some of the necessary procedures when fitting spectra, such as continuum-multiplication by a polynomial to account for flux calibration errors and rejection of bad data or model points. They use an iterative kappa-sigma clipping method to discard ``bad'' points for the fit. They test that their clipping technique converges to the same answer by starting guesses located at different parameters $\vt$ and seeing if similar lines are masked and similar best fit parameters $\vt$ are obtained.

\paragraph{\citet{mga13}} fit BT-Settl PHOENIX models to M dwarfs. Because high quality, high resolution M dwarf spectra are difficult to obtain owing to their intrinsic faintness, the development of stellar libraries for later spectral types has typically lagged behind. They use a $\chi^2$-like likelihood function and iterate to mask bad pixels. Unlike ULySS, however, they iterate the method against a sample of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars. Presumably, such a consistent residual would be due to an incorrect physical constant for a certain line or molecular feature. After masking, they repeat the fitting and claim that after two iterations the masks do not change. They note that in the future, rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures. \textbf{However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.}'' Thus, masking ``bad regions'' could very well destroy stellar information in the pursuit of a better ``fit.'' This is a nuanced point that masking schemes generally neglect. We aim to address later in this document.

\paragraph{``Bayesian'' methods} do exist. \citet{sdm+07} developed a hierarchical Bayesian approach to compare infrared spectra at low resolution ($R \sim $ 1,200) to a synthetic grid in order to determine stellar parameters $\vg$. Their model is hierarchical in that it has parameters describing the statistical and systematic uncertainties of the observed spectrum. The systematic errors are due to the imperfect performance of the instrument, for example a bad flux calibration. They use an ``empirical'' Bayesian model for the noise, meaning that they use data to estimate the prior parameters for the noise \citep{gcs+04}. They do a fit to the variance with a hierarchical linear mixed model (sec 5.2.2, and their figure 7) using design matricies and regression coefficients. This is probably to incorporate some sort of correlation length on the systematic errors, and ``smooth'' the function, as they say. They use a $\chi^2$ likelihood function and allow for ``shrinkage'' of their hierarchical parameters if there is good agreement between observed and synthetic spectra. Disappointingly, they do not explicitly account for systematic errors in the synthesis of actual lines, for example due to wrong atomic constants. One could imagine having a systematic error parameter which describes the scatter in the relationship between $\fM(\lambda | \vg)$ and the ``true'' physical spectrum in the absence of noise $\fD$. Nor do they account for any systematic effect that likely exists between $\vg$ and the true physical stellar parameters (as far as $T_{\rm eff}$ has any physical significance) and is a function of which synthetic grid they use\footnote{For example, in our tests with Kurucz vs. PHOENIX spectra we often find a systematic offset in temperature of $\gtrsim 200$ K.}. However, they do account for interpolation errors by testing interpolating spectra in the pixel space and atmospheric interpolations. Then, they select the best model using squared error loss function and a penalty measure for model complexity.

\citet{sb13} present a comprehensive Bayesian method that can fit low or high resolution spectra in addition to photometry and parallax data. They too, mention that they use a $\chi^2$ likelihood. They mention using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were created or applied. Their unpublished stellar library focuses on parts of the optical spectrum which contain the most stellar information, as determined by experts. As noted by \citet{sb13}, the grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA\ (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).''

\paragraph{SPC} A workhorse method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. Currently SPC uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values} to match high resolution models of the Sun and Vega.

\paragraph{Other methods} \citet{dj03} do not use a Bayesian model, but they do $\chi^2$ fitting to line profiles in the infrared. \citet{bai10} use a forward modelling approach that interpolates a synthetic grid to estimate stellar parameters from low resolution Gaia spectra.

\paragraph{Comparison of synthetic stellar libraries}
Gaia team members are in a mild panic about the upcoming flood of data. \citet{mc07} compare many spectral grids to determine their weaknesses. \citep{svt+11} discuss the range of stellar grids available for Gaia. \citep{cbm+05} is the Coehlo grid. Munari et al 2005, \citep{hwd+13}.

\paragraph{General trends}
Successful pixel-to-pixel methods either mask bad spectral lines or limit themselves to a narrow region of the spectrum, where they can be relatively certain that the spectral library does not contain gross systematical errors (and effectively masking the rest of the spectrum). Cherry picking regions of the spectrum like this evokes similar concerns to those of \citet{mga13}: how do you know that you are fitting the region(s) of the spectrum that deliver the right balance of information about the stellar properties?

\subsection{Line fitting using spectral synthesis}
Direct synthesis methods take a step closer towards stellar parameter estimation in the correct domain. For as large a region of the spectrum as is computationally tractable, these codes synthesize a spectrum for direct comparison with the data. Because this is an expensive operation, one must necessarily choose only a few good lines to fit.

\textbf{MOOG}\footnote{\url{http://www.as.utexas.edu/~chris/moog.html}} \citep{sne73} and \textbf{SME}\footnote{\url{http://www.stsci.edu/~valenti/sme.html}} \citep{vp96} are two popular synthesis codes. Given a set of stellar parameters, elemental abundances, and oscillator strengths, the program will interpolate a model atmosphere and then synthesize a spectrum (over a small region of wavelength). Then, they use a Levenberg-Marquartd algorithm to converge to the best-fit parameters. Their approach has the benefit of enabling individual tweaks to elemental abundances and oscillator strengths, something that is impossible in the aforementioned approaches.

In order to make the direct synthesis approach computationally tractable, one must use the simplest atmospheric models and radiative transfer. Powerful codes like PHOENIX \citep{hwd+13} \emph{should} have more accurate atmospheres and physics, but they take a long time to run ($\sim$ hours) per model, and the code is not publicly available.

\subsection{Stellar parameter estimation in a utopia}
The difficulties encountered by the similarity techniques and synthetic grid comparisons are a result of fitting in the wrong parameter space. Yes, the combination of $\vg$ is an excellent proxy to describe a star, but in truth (and at high resolution) each star is slightly different. The ``model'' that is actually producing the synthetic spectrum has many, many parameters that have been collapsed down into $\vg$. These parameters include the atmospheric structure (relations of temperature and pressure as a function of stellar depth), individual elemental abundances (all models quote a value relative to solar ($\Z$) and some synthetic grids include alpha enhancements ($\A$)), and atomic constants describing the opacity contribution of a line. \emph{All} of these parameters have uncertainties in them, some of them quite dramatic (opacities and line lists are the frequently cited), yet for a synthetic grid these parameters are \emph{fixed}. There may still be some missing physics that would be needed to produce a perfect spectrum within the noise (say for example the cores of Ca lines, or dust settling at the latest spectral types), but my naive impression is that the field of stellar spectral synthesis is mature enough that nearly all lines could be properly reproduced given sufficient scrutiny to the atmosphere and atomic constants.

If we had infinite computational power, a tantalizing approach would be to use a model with stellar parameters $\vg$, atmospheric parameters (such as pressure and temperature profiles), and uncertain atomic constants such as oscillator strengths. One could fit for individual elemental abundances and marginalize over the atomic constants. This would extract a tremendous amount of information from a high resolution spectrum. Unfortunately, the relationship between $\vt$ and the spectrum $\fD(\lambda | \vt)$ is non-linear, meaning that a new spectrum would need to be synthesized for each new $\vt$. It may be possible to make an approximate a linear relationship between say an oscillator strength and the line depth as long as the tweak is minor. If no reasonable approximation can be found, then the radiative transfer becomes an insurmountable bottleneck.

If the approximations could reduce the computation of synthetic spectrum, efficient MCMC sampling parallelized on a cluster might put this utopian vision within reach in the near future. However, without a collaborator with expert knowledge of spectral synthesis, I do not have the expertise to take such an approach. Many high quality synthetic spectral grids do exist, and so our current efforts are focused on addressing the shortfalls of the previously mentioned pixel-to-pixel methods.

\section{General problems with the spectral library approach}

Forward modelling approaches (pixel-to-pixel) have the potential to be more powerful than other stellar parameter estimation techniques. When we model directly to the data space, we have the ability to account for instrumental and systematic effects, allowing us to derive a robust estimate of our uncertainty by marginalizing over the nuisance parameters. We can also derive more complicated models to fit unresolved binary or multiple star systems. In fact, many of the most interesting systems for which we might want to determine stellar parameters are in multiple systems since we can dynamically determine masses for these systems and thus compare to evolutionary models.

As discussed in \S\ref{sec:pix}, many forward modelling approaches exist, yet nearly all use a $\chi^2$ likelihood function and most employ some sort of masking scheme. The more successful pixel-to-pixel methods use low or moderate resolution spectra ($R \sim$ 1,000 to 10,000) or a grid over a narrow range of wavelength space with finely tuned atomic constants, eschewing difficult decisions about what to mask. \textbf{The true fundamental problem with a pixel-to-pixel comparison is that the stellar spectra are not a perfect representation of reality}. This may be an obvious statement, but systematic errors in the synthetic spectra means that the previously valid assumption of Poisson errors and a $\chi^2$ likelihood function breaks down. In order to derive unbiased parameter estimates, systematic errors must be identified and properly handled.

\begin{figure}[!htb]
\begin{center}
\includegraphics{lineclasses}
\caption{The four classes of line behavior, from left to right: \textbf{Class 0}: the model behaves as expected. \textbf{Class I}: a spectral line is present in the data but missing from the model (in this case, the line is also blended). \textbf{Class II}: a spectral line exists in the model where there is only continuum in the data. \textbf{Class III}: lines exist in both the data and model, however they are of the wrong strength. Some lines are too strong while others are too weak. The lower panels show the residual spectra in units of the observational uncertainty (Poisson uncertainty only).}
\label{fig:lineclasses}
\end{center}
\end{figure}

\subsection{Classes of synthetic spectral line behavior}
From our forward modelling experiments using the modified Kurucz grid used in \citet{blj+12} and the recent PHOENIX grid \citep{hwd+13}, we have identified ways that synthetic spectra diverge from the our observed data. To start, we will describe these classes of behaviour as being tied to specific spectral lines present against a well-defined continuum. Later on, we will use the classes to describe features more generally to cover the behavior of molecular features in later type spectra. See Figure~\ref{fig:lineclasses} for visual examples of each class of behavior.

\paragraph{Class 0} These lines actually produce a good fit with a residual nearly consistent with the Poisson counting uncertainty (see \S\ref{sec:TRES} for how the data is processed). If every region of the spectrum behaved like this, we might be justified in using a $\chi^2$ likelihood function and calling it a day.

\paragraph{Class I} A spectral line appears in the data but not in the model spectrum over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. Such a type of error is perhaps the easiest to understand and to forgive. For example, such a line could easily appear in the data due to interstellar absorption or telluric absorption, something which a stellar model could never produce. The model could also fail to reproduce a true stellar line due to a missing opacity source.

\paragraph{Class II} A spectral line appears in the model but not in the data over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$. This is a more puzzling error, because presumably atomic constants would be much more likely to be \emph{missing} than added in by accident. Presumably this might be a combination of incorrect atomic data coupled with incorrect atmospheric structure that incorrectly excites a line.

The traditional way to account for Class I and Class II errors is to sigma clip the bad lines in the spectrum \citep{kpb+09, mga13}. A region of the spectrum with Class I or II errors is useless for determining spectral parameters. While sigma-clipping routines are to be avoided, if they can correctly mask these regions, then at least they are moving in the right direction.

There is additional information available to help determine which regions of the spectrum suffer from Class I or II errors.For a normal star with absorption lines against a continuum, Class I residuals will always be negative and not change significantly with $\vg$, since we the model can only produce continuum in that region. Class II residuals will always be positive and their magnitude should decrease with decreasing metallicity, since lower metallicity will decrease the strength of all lines. We could easily store the partial derivatives of an individual synthetic spectrum with respect to $\vg$
\begin{equation}
  \bigl \{ \frac{\partial \fM(\lambda | \vg)}{ \partial T_{\rm eff}}, \frac{\partial \fM(\lambda | \vg)}{ \partial \log(g)}, \frac{\partial \fM(\lambda | \vg)}{ \partial \Z}, \frac{\partial \fM(\lambda | \vg)}{ \partial \A} \bigr \}
  \label{eqn:derivative}
\end{equation}
and coupled with a prior on the range of parameters over which the residuals will be correlated, we could shore up the classification of Class I and II errors.

\paragraph{Class III} errors occur when the line strengths between the data and model do not match at a $\vg$ that fits a majority of the other lines. What makes Class III errors truly pernicious is that for a specific line, there does exist a set of $\vg$ that could properly fit the line, though this parameter combination would result in bad fits for the majority of the other lines. Class I and II errors are easy to identify since the lines will always be bad across a wide range of $\vg$ and have predictable behaviors with change in $\vg$. Class III errors, however, come in both directions: the line can be too strong or too weak. Identifying Class III errors requires making a choice about which lines to trust and how much.

Most of the Class III errors can probably be identified by finding the regions of the spectrum that have the highest residuals after dealing with Class I and II errors, although some mildly incorrect Class III lines will be difficult to distinguish from correct (Class 0) lines that are simply fit with a less probable $\vg$. Once identified, properly handling Class III errors requires careful thought because these regions of the spectrum \emph{still contain information about $\vg$}. If we take an masking approach as with Class I and II errors, we will be throwing it away. It might even be the case that Class III regions contain \emph{the most} information about $\vg$, since these ``delicate regions'' are the most sensitive to changes in $\vg$ at that particular $\vg$ and thus are tricky to tune correctly. In other words, Class III regions might also be regions where the spectrum derivative (Eqn~\ref{eqn:derivative}) is largest.

To summarize, the fundamental problem with comparing high resolution stellar spectra to grids of synthetic stellar spectra is that no one model $\vg$ will fit all of the lines properly. One can either limit their fits to regions of the spectrum that they know their spectral libraries to be accurate (implicitly masking the rest and potentially discarding a large amount of information) or suck it up and forward model the regions where the models perform poorly.

\section{Instrumental effects and ``nuisance'' parameters in the limit of a perfect stellar model}
\label{sec:perfect}
While this document aims to be general in scope, any forward-modelling approach must also discuss the characteristics of the instrument that performed the observations. We are currently modelling data obtained with the \emph{TRES}\footnote{\url{http://tdc-www.harvard.edu/instruments/tres/}} echelle spectrograph, and so we will describe the characteristics of the instrument and the calibration approaches we have taken, although we note that our approach is easily generalized for data obtained with other spectrographs.

\emph{TRES} is a 51-order echelle spectrograph on the 1.5-m Tillinghast reflector at Whipple Observatory. It spans the full optical range ($3800-9000$\AA) at a resolution of $R=48,000$ ($6.8\;{\rm km/s}$). Echelle spectrographs impart a strong blaze function upon each order, where transmission is largest in the center of the order and declines rapidly as one moves towards the edge of the order. The \emph{TRES} pipeline extracts orders such that they span a wavelength region until the transmission reaches $\lesssim$ 25\% of the central transmission in that order. Each order spans $\sim 100$\AA\ and the blue orders overlap while there are gaps between the wavelength coverage of the red orders. This spectra are blaze-corrected by dividing by a blaze function determined from fitting nightly exposures of flat lamps. Then, the spectrum is flux calibrated using sensitivity functions determined from nightly observations of spectrophotometric standards (O stars). The flux-calibrated spectrum is in units of ergs/s/${\rm cm}^2$/\AA.

\paragraph{Processing TRES Data}
\label{sec:TRES}
Because the \emph{TRES} reduction pipeline does not produce a ``sigma spectrum'' like that determined by a standard IRAF reduction, we instead calculate the Poisson uncertainty on each pixel. We can find the signal-to-noise ratio of each pixel using the raw (still-blazed) spectrum, which is in units of ``counts.''
\begin{equation}
  [S/N](\lambda) = \sqrt{ {\rm cts}(\lambda)}
\end{equation}
the Noise-to-Signal ratio is the inverse 
\begin{equation}
  [N/S](\lambda) = \frac{1}{\sqrt{ {\rm cts}(\lambda)}}
\end{equation}
To determine the uncertainty of the flux-calibrated spectrum, we multiply by N/S. The $[N/S](\lambda)$ ``spectrum'' is larger at the edges of each order due to the reduced transmission. Please note that statistic pixel uncertainties derived in this manner neglect regions of the spectrum that might have a higher variance due to sitting atop a night sky line, but short of implementing our own reduction pipeline for \emph{TRES}, there isn't much more we can do at the moment.

\paragraph{Generating a model spectrum} We parameterize the fit to a star by $\vt = \{\vstar, \vN \}$, where stellar parameters include $\vstar = \{T_{\rm eff}, \log(g), \Z, \A, v \sin i, v_z, A_v, R_\star^2/d^2 \}$. The data vector $\vD = \fD(\lambda)$ represents the \emph{TRES} spectrum. To generate a model spectrum given a set of parameters $\fM(\vt)$, a model spectrum is created by first linearly interpolating in $T_{\rm eff}, \log(g), \Z, \A$ space between grid points of instrumentally broadened and oversampled synthetic stellar spectra. Then, the spectrum is rotationally broadened for $v \sin i$ using Fourier techniques, Doppler shifted by $v_z$, extinction corrected for $A_V$, multiplied by $R_\star^2/d^2$ for a solid angle correction, and finally downsampled to the exact pixels of the TRES spectrum using spline interpolation\footnote{Because we are dealing with flux-calibrated data and synthetic spectra, we \emph{resample} to lower resolution. We do not \emph{rebin} because we are dealing with flux-density, not photon counts}.

FIGURE: showing the high resolution, raw $R =$ 500,000 spectrum compared to the one convolved with a kernel down to $R =$ 50,000. This shows that lines are blended, but it is important to keep in mind that we do have access to the high resolution models when talking about bad pixels or bad line-species. For example, we could inject bad lines in this high resolution space.

\subsection{Applied to a Gaussian likelihood function}
For the rest of \S\ref{sec:perfect}, we proceed as if there were no Class I, II, or III errors present in our model so that we may illustrate our technique for dealing with uncertainties in the flux-calibration. Although per-pixel photon counting errors are Poisson, we can safely assume that we have enough counts that we are in the Gaussian limit, although see the discussion in \S\ref{sec:residuals}. Then our likelihood function becomes a pixel-by-pixel $\chi^2$ comparison between the data spectrum $\fD$, and the model spectrum $\fM$, summed over the wavelength axis. 
\begin{equation}
 p(\vD | \vt) = {\cal L} \propto \exp \left(-\frac{\chi^2}{2} \right)
\end{equation}
\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
\end{equation}
With flat priors on $\vstar$ and $\vN$ (an assumption we will later relax) the logarithm of our posterior probability function is then
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{ \fD(\lambda) - k(\lambda | \vN) \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob}
\end{equation}

The prefactor $k(\lambda | \vN)$ is a calibration function that aims to account for errors in the blaze-correction or flux-calibration by multiplying the synthetic spectrum by a polynomial. The parameters describing this function are the coefficients of the polynomial. This approach of accounting for large-scale calibration errors in the spectrum is more justified than the traditional approach of normalizing to the stellar continuum. From a flux-calibration experiment where we observed spectrophotometric standards BD+28 and BD+25 multiple times, we calibrated the dispersion in (FIGURE showing the sensitivity function for a few representative orders) flux-calibration function, or IRAF's ``sensitivity function.'' These tests revealed that the largest source of error in flux-calibration was actually in the overall magnitude of individual order, this could be due to slit losses, poor correction due to parallactic angle, or scattered light within the instrument. These corrections are smaller than 10\%. But more importantly, the corrections in the higher order terms (ie, the shape of the bandpass calibration for each order) are small ($\lesssim 3\%$). As described in \S\ref{sec:priors}, we can set priors on these coefficients. This prior information on the calibration parameters motivates our choice of functional parameterization for the nuisance parameters in \S\ref{sec:nuisance}.
The polynomial approach is also better than the normalization approach if the order only spans a small region of the spectrum. If a M star absorption band spanned the whole order, one might accidentally normalize to the feature rather than a pseudo-continuum.

\subsection{Nuisance parameters}
\label{sec:nuisance}
In our case, $\vN = \{c_0, c_1, \ldots, c_N\}$ are a set of 4 Chebyshev polynomial coefficients for each order. This amounts to a ``re-fluxing'' of the model following the techniques of \citet{elh+06}. We apply the re-fluxing to the model rather than attempt to re-flux the data since this operation would also need to re-scale the noise, making $\chi^2$ a complicated function of $k$.

In order to understand the effects of the $k(\lambda | \vN)$ term, we now consider its effects on just one order of the 51-order echelle spectrum. The first four Chebyshev polynomials are 
\begin{align*}
  T_0(x) &= 1\\
  T_1(x) &= x\\
  T_2(x) &= 2 x^2 - 1\\
  T_3(x) &= 4 x^3 - 3x\\
\end{align*}
In our implementation, we map the full pixel range of a specific order $\lambda \in [\lambda_{\rm min}, \lambda_{\rm max}]$ to $x \in [-1, 1]$. By mapping pixel number in a given order, rather than wavelength, we are insensitive to wavelength shifts of the model. Then, for a given set of Chebyshev coefficients, $\vN = \{ c_0, c_1, c_2, \ldots, c_N \}$, we have 
\begin{align}
  k(\lambda | \vN) &= c_0 T_0(\lambda) \left [1 + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \right ]\\
  k(\lambda | \vN) &= c_0 \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ]
  \label{eqn:k_new}
\end{align}
When applied to scale the model flux $\fM$, this becomes
\begin{align}
  k(\lambda | \vN) \fM &=  \left [1 + \sum^N_{i = 1} c_i T_i(\lambda) \right ] (c_0 \fM)\\
  k(\lambda | \vN) \fM &= c_0 \fM + c_1 T_1 (c_0 \fM) + \ldots + c_N T_N (c_0 \fM)
\end{align}

This functional form of $k$ allows $c_0$ to scale by large amounts while the higher order terms (denoted by $c_n$) are perturbations on top of the \emph{scaled} model ($c_0 \fM$). However, see \S\ref{sec:gaussian_simplification} for an alternative model when perturbations are small. We can rewrite Equation~\ref{eqn:lnprob} as 
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [1 + \sum^N_{i = 1} c_i T_i(\lambda) \bigr ] c_0 \fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
  \label{eqn:lnprob2}
\end{equation}
To simplify the following discussion, we consider a single wavelength $\lambda_i$ and then generalize this to a sum over $\lambda$ later. If we expand out the square in Equation~\ref{eqn:lnprob2}, then for a given $\lambda_i$ we have
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + \sum_{i=1}^N c_i T_i \right ]^2 \right ]
 \label{eqn:lnprob_lambda}
 \end{equation}
where $\sigma$, $\fD$, $\fM$, and $T_{n,m}$ are all evaluated at $\lambda_i$. We can further expand this to
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
  \label{eqn:expanded}
 \end{equation}

To simplify this math, we can rewrite the Chebyshev coefficients and polynomials as column vectors $\vc$ and $\vec{T}(\lambda)$, respectively
\begin{equation}
  \vc = 
  \begin{bmatrix}
    c_1\\
    c_2\\
    \vdots\\
    c_N
  \end{bmatrix}
  \hspace{3cm}
\vec{T}(\lambda) = 
\begin{bmatrix}
T_1(\lambda)\\
T_2(\lambda)\\
\vdots\\
T_N(\lambda)\\
\end{bmatrix}
\end{equation}
then 
\begin{equation}
  k(\lambda | \vN) = c_0 \left [ 1 + \sum^N_{i = 0} c_i T_i \right ] = c_0 \left[ 1+ \vec{T}^\trans \cdot \vc \right]
\end{equation}
If we let 
\begin{equation}
  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
  \begin{bmatrix}
T_1 T_1 & T_1 T_2 &  \hdots & T_1 T_N \\
T_2 T_1 & T_2 T_2 &  \hdots & T_2 T_N \\
\vdots  & \vdots  &  \ddots & \vdots \\
T_N T_1 & T_N T_2 &  \hdots & T_N T_N \\
  \end{bmatrix}
\end{equation}
then we have
\begin{equation}
  \sum_{i =1}^N \sum_{j=1}^N c_i T_i c_j T_j = \vc^\trans \cdot {\bm W} \cdot \vc
\end{equation}
We can rewrite Equation~\ref{eqn:expanded} quthat should be accounted for? adratically in $\vc$ as
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2 \sigma^2} \left [ \fD^2 - 2 \fD c_0 \fM \left [1+ \sum_{i=1}^N c_i T_i \right ]+ c_0^2 \fM^2 \left [1 + 2 \sum_{i=1}^N c_i T_i + \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j \right ] \right ]
 \end{equation}
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2} \sum_{i=1}^N \sum_{j=1}^N c_i T_i c_j T_j + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \sum_{i=1}^N c_i T_i  -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
\end{equation}
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fM^2}{\sigma^2}  \vc^\trans {\bm W} \vc + \left (\frac{- c_0^2 \fM^2 + c_0 \fD \fM}{\sigma^2} \right ) \vec{T}^\trans \vc -  \left( \frac{c_0^2 \fM^2 - 2 c_0 \fD \fM + \fD^2}{2 \sigma^2} \right)
\end{equation}

Because matrix multiplication is associative, we can sum ${\bm W(\lambda)}$ and $\vec{T}^\trans(\lambda)$ across all $\lambda$, and define
\begin{align}
  {\bm A} &= c_0^2 \sum_{\lambda} \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda) \\
  \vec{B} &= \sum_{\lambda} \frac{-\fM^2(\lambda) c_0^2 + \fD(\lambda) \fM (\lambda) c_0}{\sigma^2(\lambda)} \vec{T}(\lambda)\\
  g &= -\frac{1}{2} \sum_{\lambda} \frac{c_0^2 \fM^2(\lambda) - 2 c_0 \fD(\lambda) \fM(\lambda) + \fD^2(\lambda)}{\sigma^2(\lambda)} 
\end{align}
Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
\begin{align}
  \ln \bigl [p(\vstar, \vN | \vD) \bigr] &\propto - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g\\
  p(\vt | \vD) = p(\vstar, \vN | \vD) &\propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
  \label{eqn:lnprob_matrix}
\end{align}
Here we have the full posterior probability distribution assuming flat priors. This posterior is a function of both the stellar parameters $\vstar$ and the nuisance parameters $\vN$. Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters. This marginalization encapsulates the uncertainty due to the inference on the nuisance parameters. Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically, and then sample from the marginalized distribution. 
\begin{equation}
  p(\vstar, c_0 | \vD) = \int p(\vstar, c_0, \vc\; | \vD) \dd \vc
\end{equation}
The analytic multi-dimensional Gaussian integral with linear term \citep{sgd+09} yields 
\begin{align}
  p(\vstar, c_0 | \vD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
  p(\vstar, c_0 | \vD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
\end{align}
Where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$. By doing this marginalization we have reduced the dimensionality of our posterior space from $200+$ parameters ($\sim 7$ stellar parameters plus 3 Chebyshev coefficients for each of 51 echelle orders) to $\sim 60$ ($\sim 7$ stellar parameters and one $c_0$ for each of the 51 echelle orders). This enables a dramatic speedup in the MCMC sampling. For flux-calibrated spectra where the variation in $c_0$ is expected to be small (less than how much?), we can approximate $k$ in a different manner which allows marginalization over the $c_0$ coefficients as well, reducing the dimensionality of the space to only the $\sim 7$ stellar parameters. This approach is detailed in \S\ref{sec:gaussian_simplification}.

\subsection{Including Gaussian priors on nuisance parameters}
\label{sec:priors}
Using this polynomial formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters. From our flux-calibration experiment, we determined that the overall scaling could range by 10\%, while the higher order corrections were less than 3\%. Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two. This means we should be using a log-normal prior on $c_0$
\begin{align}
  p(c_0) &= \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right) \\
  \ln p(c_0) &= \ln \left( \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \right) - \frac{(\ln c_0)^2}{2 \sigma_{c_0}^2}
\end{align}
centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 f$), we assume Gaussian priors centered about 0
\begin{equation}
  p(c_n) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^2}{2 \sigma_{c_n^2}} \right)
\end{equation}
which we can express in vector form as
\begin{equation}
  p(\vc\;) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
  \label{eqn:nuisance_prior} 
\end{equation}
where 
\begin{equation}
  {\bm D} = 
  \begin{bmatrix}
    \sigma_1^{-2} & 0 & \hdots & 0 \\
    0 & \sigma_2^{-2} & \hdots & 0 \\
    \vdots & \vdots & \ddots & 0 \\
    0 & 0 & \hdots & \sigma_N^{-2} \\
  \end{bmatrix}
\end{equation}
and $\sigma_i$ represents the width of the Gaussian prior on the i-th Chebyshev coefficient, in our case $\sigma_i = \sigma_{c_n} \approx 0.03$. $\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction. Now, we can write the posterior with priors as
\begin{equation}
  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\; | \vD) \propto p( \vD | \vstar, c_0, \vc\; ) p(\vc\;) p(\vstar) p(c_0) 
\end{equation}
\begin{equation}
  p(\vstar, \vN | \vD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
  \label{eqn:posterior_prior}
\end{equation}
we can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$
\begin{align}
  -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) &= \frac{1}{2}\left ( -\vc^\trans {\bm D} \vc + \vc^\trans {\bm D} \vec{\mu} + \vec{\mu}^\trans {\bm D} \vc - \vec{\mu}^\trans {\bm D} \vec{\mu} \right )\\
  &= -\frac{1}{2} \vc^\trans {\bm D} \vc + ({\bm D} \vec{\mu})^\trans \vc - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu}
\end{align}
then we can rewrite
\begin{align}
  {\bm A}^\prime &= {\bm A} + {\bm D}\\
  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
\end{align}
in the case that $\vec{\mu} = \{0, 0, \ldots, 0\}$, there is only a non-zero correction to ${\bm A}$. The full posterior probability distribution is then
 \begin{equation}
   \boxed{
  p(\vstar, \vN | \vD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
}
\label{eqn:lnprob_lognormal}
\end{equation}
This probability function appears in the code as \texttt{lnprob\_lognormal} and has $\sim 7$ stellar dimensions and $\sim 200$ nuisance dimensions.

Using the previous multi-dimensional integral, we can marginalize over the $\vc$ including the Gaussian priors
\begin{equation}
  \boxed{
  p(\vstar, c_0 | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) p(c_0)
}
\label{eqn:lnprob_lognormal_marg}
\end{equation}
where $N$ is the number of Chebyshev coefficients marginalized over, in our case $N=3$. This probability function appears in the code as \texttt{lnprob\_lognormal\_marg} and has $\sim 7$ stellar dimensions and $\sim 50$ nuisance dimensions. If we increase the permissible degree of re-fluxing by increasing $\sigma_i$, this will have the effect of smearing out or widening the posterior on a given $\vstar$.

If one uses a marginalized posterior, but is still interested in obtaining a flux-corrected spectrum for the purposes of inspecting the residuals, it is still possible to do so. We can use Eqn~\ref{eqn:lnprob_lognormal} with $\vstar$ fixed at a sample from the posterior of Eqn~\ref{eqn:lnprob_lognormal_marg} and then use MCMC to draw samples from $p(\vN | \vstar, \vD)$. Now, the combination of $\vstar$ and $\vN$ completely describes a flux-corrected spectrum.

\subsection{Sampling a un-flux calibrated model}
There also exists a wealth of archival \emph{TRES} data that is not flux calibrated, since its primary purpose was for radial velocity determination of FGK stars. To fit this data we can use newer sensitivity functions taken under different conditions and soften the priors on the coefficients. Tests indicate that this works rather well. The blaze function will might change significantly between nights and years, so this might be a source of error.

For non flux calibrated spectra, we determined that the overall spread in $c_0$ across all orders should not be more than $\sigma_{c_0} \leq 2$, or the one-sigma deviation of $c_0$ is between $0.125 \leq c_0 \leq 8.0$. The plot in plots/priortests shows that log-normal prior (vs. Gaussian) on $c_0$ is the correct choice . Since we care most about the inference on F, we want to know that this prior is not biasing the final outcome. Using a Gaussian prior with a large $\sigma_{c_0}$ will bias the results, but if $\sigma_{c_0}$ is small ($\lesssim 0.2$), then we can get away with Gaussian priors (see \S\ref{sec:gaussian_simplification}). Is there a better way to determine these thresholds other than my current technique of testing and visual inspection?

\subsection{Hierarchical modelling}
Instead of the current framework where the nuisance coefficients of each order are independent, we could derive a hierarchical framework where each $c_0$ is drawn from a probability distribution of $c_0$ described by a hyperparameter $C_0$. If each order required a similar $c_0$, then this would enable shrinkage of $C_0$, however if each order required a widely disparate $c_0$, then $C_0$ would correspondingly increase. Higher order coefficients could still be analytically marginalized over.

\subsection{Sampling the posterior and performance of the $\chi^2$ likelihood}
Currently we are sampling this posterior with the MCMC ensemble sampler \citep{gw10} implemented in \texttt{Python} by \citet{fhl+12} as \texttt{emcee}. See \citet[CosmoHAMMER]{asa+13} for a review of \texttt{emcee} parallel performance and scaling relationships on a computer cluster.

In this section, we described our approach to take into account flux calibration errors assuming a perfect synthetic spectra. If one does a straight up $\chi^2$ fit to the data, however, you can get posteriors that look have 1-sigma contours of $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity. This would be as precise as stellar parameters obtained with astroseismology! Clearly there is something going wrong here.

As our previous discussion has noted, these posteriors are unfortunately too good to be true and we must take a more sophisticated approach. However, many of the flux-calibration principles developed in this section should still apply. More advanced posterior functions in the following sections might require using more specialized MCMC techniques such as Gibbs sampling or expectation maximization (EM).

\section{Baby steps beyond a perfect stellar model}
\label{sec:residuals}
In this section we describe a few simple things we might try (or have tried) in order to remedy the shortcomings of the previous section.

\subsection{Additional sources of noise}
Presumably we shouldn't have to add an additional source of instrumental ``noise'' to the fit, because our nuisance parameters will find an exact flux calibration. However, CCD readnoise and 1D spectrum extraction will almost definitely increase the noise in a pixel beyond theoretical Poisson limits. Normally this information will be encapsulated in a ``sigma spectrum,'' but unfortunately we do not have access to this unless we develop an alternative to the \emph{TRES} pipeline. We could try to fudge adjustments to the readout noise by adding a small error in quadrature.

Even the stellar continuum can flicker/jitter (as shown by Kepler studies), which would produce a small departure from a perfect continuum. Even in the Class 0 region of our spectrum (Figure~\ref{fig:lineclasses}, left panel), the residuals are slightly larger than 1 $\sigma$ on average.

When we interpolate a synthetic spectrum from the library to a specific $\vg$, we will introduce interpolation errors depending on how close we are to a grid point in the library. \citet{hus12} investigate the possible range of interpolation errors at specific $\vg$ by dropping out a spectrum from the grid, interpolating over it, and then comparing the true spectrum to the interpolated spectra. For hotter stars with well defined continua, we find the maximum error is $\lesssim 5$\%, while for M stars the error could be as large as $\sim 20$\% for some spectral features. By doing this interpolation test across the grid, we could determine an ``interpolation error spectrum'' that would encapsulate the maximum error possible for an interpolated spectrum, weighted by its distance from the nearest grid points. This uncertainty could be included in quadrature. Since most of our \emph{TRES} spectra are at signal/noise of $\sim 30$ per pixel (or noise/signal ratio of $1/30 \approx 0.03$), an interpolation noise contribution of $\sim 5$\% the level of the signal level would be significant. However, we tried arbitrarily increasing the Poisson errors by a factor of 2 or 3 and it didn't significantly change the results. While such an error spectrum should eventually be incorporated, we have bigger fish to fry.

\subsection{Masking}
We must necessarily mask certain wavelength regions of the T Tauri spectra that are contaminated (one might also say ``made more interesting'') by the astrophysical realities of accretion or stellar spots (such as the Balmer lines, which are in emission).  We also accept the assumption that rotation or accretion onto the star does not fundamentally alter the structure of the model atmosphere that generates the model spectrum and that a spectral comparison of the ``clean'' regions of the spectrum is valid. We also mask blatant Class I and II regions by eye, but currently do nothing for smaller errors or Class III errors.

\subsection{Gallery of likelihood functions}
A problem with the $\chi^2$ likelihood function is that the Gaussian tails of the probability function rapidly decline, making any outliers extremely improbable, which will bias a fit.

A common approach to make ``robust inference'' is to use a likelihood function with fatter tails. Although not rigorously motivated, this is can be a quick thing to try. We explore the following likelihood functions.

\paragraph{Combination of Gaussian with exponential wings}
We fit a function of the form
\begin{equation}
  p(\fD(\lambda) | \fM(\lambda | \vt)) \propto \exp \left (\frac{(\fD - \fM)^2}{(\sigma_G \sigma_P)^2} \right ) + A \exp \left (\frac{|\fD - \fM|}{\sigma_E \sigma_P}  \right )
\end{equation}
to the residuals of a reasonable $\vt$ to determine $A$, $\sigma_G$, and $\sigma_E$. This functional form assumes that there will be a Gaussian core, similar to the $\chi^2$ likelihood, but also allows some contribution from an exponential tail, used to describe the residuals. I think the problem here is that the exponential is too steep to properly describe Class I and II residuals, although it might do well for Class III errors.

\paragraph{Sivia's model}
\citet{ss06} describe a likelihood function which treats the noise estimate for each data point suspiciously. The true noise $\sigma$ can be no smaller than some original estimate $\sigma_0$, but it may be much larger. By putting a prior on $\sigma$, 
\begin{equation}
  p(\sigma | \sigma_0) = \left \{ \begin{array}{cc}
    \sigma_0/\sigma^2 & \sigma  \leq \sigma_0\\
    0                 & \textrm{else}
  \end{array}
    \right.
\end{equation}
and marginalizing over the $\chi^2$ likelihood, the new likelihood function becomes
\begin{equation}
  p(\fD | \fM, \sigma_0) = \frac{1}{\sigma_0 \sqrt{2 \pi}} \left [ \frac{1 - e^{-R^2/2}}{R^2} \right ] 
\end{equation}
where $R = (\fD - \fM)/\sigma_0$. This function has even fatter tails than the exponential model and seems to give more believable results. We have done the most testing using this new likelihood function, but it still does not give a truly satisfying result. For example, which lines contributed most to the posterior? What about the wings of class I and II errors? These individual pixels might still be given a weight of $\sigma_0$ because they are not nearly as bad as the line centers.

\paragraph{$\chi^2$ with flat tails}
\citet{hbl10} describe a likelihood function that smoothly caps the total likelihood contribution any outlier might have at some constant $Q$ by using a $\chi^2_Q$ function
\begin{equation}
  \chi^2_Q = \sum_{i = 1}^N \frac{Q^2 [\fD - \fM]^2}{Q^2 \sigma_P^2 + [\fD - \fM]^2}
\end{equation}
In some ways, this is analogous to sigma-clipping. Once a residual reaches the flat tails, it fails to factor into the determination of parameters.

Other functions we might try include a Student-t statistic, a Lorentzian, or a linear combination of Gaussians. While all of these likelihood function shenanigans are amusing, we would like to move beyond tweaking the fitting function towards an approach that properly incorporates knowledge about the specifics of our problem.

\subsection{Bad data model}
The bad data model \citep{pre97,hbl10} is a powerful framework for forward modelling data, which allows the modeler to explicitly account for outliers using a generative model.

In the simplest mixture model, each data point has a binary flag $q_i$ which specifies whether that data point is good or bad. Then, for the entire data set we have an ensemble of $\qN$. If this ensemble of $\qN$ is chosen appropriately, and the good data points are flagged as good and the bad data points are flagged as bad, the likelihood will be maximized. We adopt $P_B$ for the \emph{a priori} probability that a given data point will be bad, $Y_b$ for the mean of the bad data distribution, and $V_b$ for its variance. Written out explicitly, this is
\begin{multline}
  {\cal L} = p({\bm D} | \vt,\qN, Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma} \exp \left[ - \frac{(\fD - \fM)^2}{2 \sigma^2} \right] \right ]^{q_i}  \times \\
  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma^2)}} \exp \left[ - \frac{(\fD - Y_b)^2}{2 (V_b + \sigma^2)} \right]
  \right ]^{1 - q_i}
\end{multline}
We we a Bernoulli prior on the $q_i$
\begin{equation}
  p(q_i | P_b) = [1-P_b]^{q_i} P_b^{[1-q_i]}
\end{equation}
then, the probability of any given ensemble of $\qN$ is 
\begin{equation}
  p(\qN | P_b) = \prod_{i=1}^N [1-P_b]^{q_i} P_b^{[1-q_i]}
\end{equation}

We can analytically marginalize over the single flag by a discrete sum over the two states of $q_i$: $q_i = 1$ and $q_i = 0$. 
\begin{multline}
  p(\vt, P_b, Y_b, V_b | D_i)  =  \int p( \vt, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
  = p(\vt, q_i=1, P_b, Y_b, V_b | D_i) + p(\vt, q_i=0, P_b, Y_b, V_b | D_i) \\
  = \Bigl[ p_{\rm good}(D_i | \vt)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(\vt ,P_b, Y_b,V_b) 
\end{multline}

\begin{multline}
  p(\vt,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i |\vt)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(\vt,P_b, Y_b, V_b)\\
  \boxed{
    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_{yi}^2}} \exp \left[ - \frac{(\fD - \fM)^2}{2 \sigma^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma^2)}} \exp \left[- \frac{(\fD - Y_b)^2}{2 (V_b + \sigma^2)} \right] \right\}
  \times p(\vt,P_b,Y_b,V_b)}
\end{multline}

In its simplest form, the bad data model results in a likelihood function that is a linear combination of Gaussians. Although the functional form you end up with is similar, the benefit of this bad data approach over a gallery of likelihood functions is that it is motivated by the characteristics of the bad data ($P_B$, $Y_b$, and $V_b$). If we wish to add additional information about the characteristics of the bad data, we can do so.

\subsection{Spatial correlations of pixels}
We can do more. We now discuss bad data models that leverage more of the information about our problem to more robustly handle systematic errors. In bad model of the previous section, we assumed that each data point is independent. However, we know that Class I, II, and III errors will be correlated across several pixels, where the correlation length is the typical width of a spectral line, set by the stellar rotation kernel $v \sin i$, the instrumental resolution kernel, or both. Some very strong lines (such as the Balmer lines, or the Mg b triplet) will have wings that are correlated across a wider length of pixels.

\subsubsection{Mixture of multiple components}
When we believe that a sample population might be composed of several populations with different behaviours (Class 0, I, II, and III), it makes sense to expand the bad data model to a multiple mixture model. Following \citet{gcs+04}, our likelihood function becomes
\begin{equation}
  p(\fD, \vec{Q} | \vt) = p(\fD | \vec{Q}, \theta) p(\vec{Q}) = \prod_\lambda \prod_{m = 0}^3  {\cal L}_m(\fD(\lambda | \theta_m))^{q_{im}}
\end{equation}
where we use $\vec{Q} = \qM$ to describe a single vector of all the individual $q_{im}$ flags. For example, when a Class I, II, or III error occurs, the core of the line may be a 20 sigma residual while the wings of the line will taper off to a 3 sigma residual, and eventually to a 0 residual once it reaches the continuum. If we treat pixels as uncorrelated, then the small-residual wings of the bad lines might actually have more probability if they were labelled as ``good'' pixels. To remedy this situation, we attempt to develop a bad data model which includes spatial (wavelength) correlations.

This correlation length will probably depend on the spectral type of star. For example, if they're simple F and G stars then the bad lines will likely be singular. But if they are M stars, then there is a chance that there might be a large ``pseudo-continuum'' region that would be incorrect. For Class I and II errors, we could include a hyper parameter describing the width of each line, and another hyper parameter describing the (distribution of) depth. For Class III errors, describing how line strengths might be wrong makes sense for stars with continuum, but as we move to later type stars (K stars) that do not have a nice continuum, we may run into trouble. For the later stars (M dwarfs), there might be a whole region that is incorrect.

Proper Class III modeling is very important, following the discussion of \citep{mga13}. Some spectral features will not fit perfectly, especially at the lower end of the temperature scale. However, it's important to still take these features into account when modelling since they still contain information. How should we account for the fact that some parts of the spectrum carry more information about $\vg$ than other parts of the spectrum? Is it important to weight this parts beforehand? Or should this contribution to the posterior naturally weight itself, the way certain points might heavily contribute to the slope and intercept of a linear fit?

\subsubsection{Correlation of the $q_i$ flags}

In the manner of Gaussian random processes, we develop a covariance function
\begin{equation}
  {\rm cov}(q_i, q_j) = \rho(| \lambda_i - \lambda_j|)
\end{equation}
or where the covariance matrix has components
\begin{equation}
  \Sigma_{ij} = \sigma^2 \rho(|\lambda_i - \lambda_j|)
\end{equation}
where
\begin{equation}
  \rho_{ij} = \frac{ {\rm cov}(i,j)}{\sigma_i \sigma_j} = \frac{E[(X - \mu_x)(Y - \mu_y)]}{\sigma_x \sigma_y}
\end{equation}
that describes the correlation between the state of the bad data flags. The simplest thing to do would be to take the common correlation length in the spectrum imposed by $l$. If we say that the typical feature will have five adjacent pixels wrong, then the correlation matrix would be
\begin{equation}
  \begin{bmatrix}
    1 & 4/5 & 3/5 & 2/5 & 1/5 \\
    4/5 & 1 & 4/5 & 3/5 & 2/5 \\
    3/5 & 4/5 & 1 & 4/5 & 3/5 \\
    2/5 & 3/5 & 4/5 & 1 & 4/5 \\
    1/5 & 2/5 & 3/5 & 4/5 & 1\\
  \end{bmatrix}
\end{equation}
(this is embedded in an N-pixel sparse matrix where everywhere but the diagonal is 0). And so
\begin{equation}
  {\rm cov}(q_i, q_j) = \left \{ \begin{array}{cc}
    1 - \frac{|\lambda_i - \lambda_j|}{5} & |\lambda_i - \lambda_j| \leq 5 \\
    0                                    & \textrm{else}
\end{array}
\right .
\end{equation}
Once we move beyond 5 pixels, any correlation vanishes.

Then the probability functions would be
\begin{multline}
  p(D_{i + j} | \vt, q_i =0) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=0) + \\
    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 0) \bigr ]
\end{multline}

\begin{multline}
  p(D_{i + j} | \vt, q_i = 1) = \prod_{j = -4, \ne 0}^4 \bigl [ p(D_{i+j} | \vt, q_{i +j}=0) p(q_{i+j} = 0 | q_i=1) + \\
    p(D_{i + j} | \vt, q_{i + j} = 1) p(q_{i +j} = 1 | q_i = 1) \bigr ]
\end{multline}

The marginalization could still be done analytically, but now the likelihood function involves a correlation between adjacent pixels. Alternatively, we might be able to use a sparse array for this? Or, we could evaluate all of the pixels, and then have the final likelihood function involve some correlation terms between the lnprobs evaluated at each pixel. We could do this using the logarithm so that everything is additive. We might also be able to evaluate the FFT of the $\log(p(D_i))$ and just do the autocorrelation in Fourier space?

\subsubsection{Correlation of adjacent $\fM(\lambda)$}
We haven't yet mentioned what ${\cal L}_m$ should be for each $m$. Not only will the flags $q_i$ be spatially correlated but the values of $\fM$ will also be correlated on the scale of $v \sin i$ or the instrumental resolution kernel, whichever is larger.

For OBAFGK stars, the lines are likely to be Gaussian shape. That is if $\fD(\lambda_i)$ specifies the center of the line, then the neighboring $\lambda_{i \pm 1,2,3,\ldots}$ pixels will be of lower residual height. For M stars, there might be a wider correlation length and the shape of the lines might not be well-described by a Gaussian. Whatever the case, this means that the inside of the likelihood function should also have spatial correlations. In this case, the likelihood function also includes the nearest 5 pixels
\begin{equation}
  {\cal L}(\lambda_i) = \exp\left ( 
  \begin{bmatrix}
x & x & \dots\\
x & x & \dots\\
  \end{bmatrix}
  \right )
\end{equation}
or maybe evaluate the likelihood and then matrix multiply? This matrix math needs to be worked out. The covariance function between $\fD(\lambda_i)$ could be learned from the spectrum itself (a la Gaussian random processes \citep{rw05}), or specified to create a Gaussian shape. Then, the likelihood functions for each mixture component in the model would be custom tailored for the type of error that it expected to encounter. (Strength and shape of residuals).

\subsection{Hyperparameters}
\citep{mb13} discuss the difficulty of combining different data sets for the combined inference on a single problem. They use a hyperparameter approach that can artificially inflate the variance of each data set to find a common posterior. They also allow correlations between data sets. We could imagine setting up a hyperparameter approach that can tune the sensitivity of certain lines, treating different lines as different datasets. Such a hyperparameter might also undergo ``shrinkage'' similar to \citet{kru10}.

\subsection{Sampling such a model}
If our bad model is sufficiently complex that we are not able to analytically marginalize over the $q_i$, it is still possible to sample in this huge parameter space. A stellar model with flags for each pixel would have $\vstar + N$ parameters, where for a typical high resolution optical spectrum $N >$ 100,000. Even using an ensemble sampler, sampling in this space would take a tremendous amount of time to converge.

An alternative way to approach this problem is to do Gibbs sampling in the flags. It may be possible to do do Gibbs sampling where we draw from the conditional probability distribution, or at the very worst we could still do a Gibbs sampler with a Metropolis-Hastings backend, where we propose a jump in a single $q_i$ at a time. Such an sample requires none of the previous heavy lifting (spectral grid interpolation, FFT, or pixel interpolation), and so even though there are $\sim$ 100,000 flags to sample, this is likely to be computationally feasible.

\citet{gcs+04} suggest to start with a crude estimate (such as $k$-means clustering) to tentatively identify which observations have been drawn from which mixture components. They also suggest that the expectation-maximize (EM) algorithm could also be used to sample this posterior.

\section{Injection of fake lines}
As an alternative approach to modelling the systematically wrong regions of the spectrum, we could actually inject fake lines into the spectrum to compensate for errors in the models. Presumably, each line would be a Gaussian parameterized by $\lambda_0$, strength, and FWHM. In M star spectra, this might actually need to be a larger continuum feature that is added.

Line strength could be described by a hyper prior, and FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM (in velocity space). We could penalize the model fits by having a hyperprior on the number of bad lines that can be created. Given a bad line and the linelist that produced the synthetic models, it may be possible to identify other regions of the spectrum that might be bad. For example, if a vanadium line is bad, then we could check to see if the other vanadium lines might also be bad.

Class III errors might be difficult to address in this paradigm since we would be fitting the different of two lines, though we might still be able to get it to work.

\subsection{Piecewise metallicity model}
In theory, it is possible to fit the profiles of only certain elements to determine the abundances of certain elements. By masking all non-alpha elements, \citet{hus12} use PHOENIX spectra to derive individual alpha abundances of stars.

Rather than masking spectra, we could assemble a piecewise spectrum where regions of the spectrum might be drawn from a different combination of $\vg$, say with lower metallicity in order to account for a bad line strength. These regions would have to be normalized properly so that the composite spectrum is sufficiently smooth. Coupled with a linelist, one could tweak individual elemental abundances.

\section{Validation tests}
Comparison to main sequence stars of HAT-P-9, and WASP-14, from the TRES archive \citep{tfs+12}. Possible comparisons using HIRES data.

\acknowledgments
IC would like to graciously acknowledge Gregory Green, Kaisey Mandel, Doug Finkbeiner, and Daniel Eisenstein for many fruitful discussions about statistics and spectroscopy. This research made use of Astropy, a community-developed core Python package for Astronomy \citep{art+13}.

\appendix

\section{Simplification for small $c_0$ perturbations}
\label{sec:gaussian_simplification}
If the correction in $c_0$ is small, we can use an alternate formulation of $k$ and put Gaussian priors on all nuisance parameters. This allows analytic marginalization over all nuisance parameters (including $c_0$), and reduces the dimensionality of the posterior to only the $\sim 7$ stellar parameters, $\vstar$.

\begin{align}
  k(\lambda | \vN) &= c_0 T_0(\lambda) + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \\
  k(\lambda | \vN) &= \sum^N_{i = 0} c_i T_i(\lambda)
\end{align}
Following \S\ref{sec:nuisance}
\begin{equation}
  \ln \bigl [p(\vt | \vD) \bigr] \propto - \frac{1}{2} \sum_\lambda \left [\frac{\fD(\lambda) - \bigl [ \sum^N_{i = 0} c_i T_i(\lambda) \bigr ]\fM(\lambda | \vstar)  }{\sigma(\lambda)} \right ]^2
\end{equation}
\begin{equation}
 \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM \sum_n c_n T_n + \fM^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
 \end{equation}
Now the column vectors $\vc$ and $\vec{T}(\lambda)$ include $c_0$ as well
\begin{equation}
  \vc = 
  \begin{bmatrix}
    c_0\\
    c_1\\
    \vdots\\
    c_N
  \end{bmatrix}
  \hspace{3cm}
\vec{T}(\lambda) = 
\begin{bmatrix}
T_0(\lambda)\\
T_1(\lambda)\\
\vdots\\
T_N(\lambda)\\
\end{bmatrix}
\end{equation}
\begin{equation}
  k(\lambda | \vN) = \sum^N_{i = 0} c_i T_i = \vec{T}^\trans \cdot \vc
\end{equation}
\begin{equation}
  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
  \begin{bmatrix}
T_0 T_0 & T_0 T_1 &  \hdots & T_0 T_N \\
T_1 T_0 & T_1 T_1 &  \hdots & T_1 T_N \\
\vdots  & \vdots  &  \ddots & \vdots \\
T_N T_0 & T_N T_1 &  \hdots & T_N T_N \\
  \end{bmatrix}
\end{equation}
\begin{equation}
  \sum_n \sum_m c_n T_n c_m T_m = \vc^\trans \cdot {\bm W} \cdot \vc
\end{equation}
\begin{equation}
  \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM\, \vec{T}^\trans \vc \; + \fM^2 \, \vc^\trans {\bm W}  \vc \bigr ]
\end{equation}
\begin{align}
  {\bm A} &= \sum_\lambda \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
  \vec{B}^\trans &= \sum_\lambda \frac{\fD(\lambda) \fM(\lambda) }{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
  g &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}\\
\end{align}
Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\vD$. If we include Gaussian priors, then we have 
\begin{align}
  {\bm A}^\prime &= {\bm A} + {\bm D}\\
  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
\end{align}
where now ${\bm D}$ and $\vec{\mu}$ include $c_0$ as well. The main difference from before is that now the prior on $c_0$ is now Gaussian instead of log-normal. This is acceptable as long as $\sigma_{c_0}$ is small ($< XX\%$), otherwise this will bias the correction factor. This also means that we will usually have $\vec{\mu} = \{1, 0, \ldots, 0\}$. The full posterior probability function becomes
 \begin{equation}
   \boxed{
  p(\vstar, \vN | \vD) = p(\vstar, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(\vc)
}
\end{equation}

Using the previous multi-dimensional integral, we can marginalize over all $\vc$ including the Gaussian priors
\begin{equation}
  \boxed{
  p(\vstar | \vD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) 
}
\end{equation}

To gain insight, we can re-express this new version of $k$ in the original form as 
\begin{equation}
   k(\lambda | \vN) = c_0 T_0 + c_1 T_1 + \ldots + c_N T_N
\end{equation}
where $c_1 = c_0 {c_1}^\prime T_0$ and $c_N = c_0 {c_N}^\prime T_0$. Thus, if there needs to be a large perturbation in $c_0$, then there must also be a large perturbation in $c_n$. The prior on $c_1$ is now a product of a log-normal prior on $c_0$ and a Gaussian prior on ${c_1}^\prime$, which is not analytically tractable. Therefore, in the case of large perturbations, it is better to use Equation~ and use a log-normal prior on $c_0$. When we use the original framework, we are saying that perturbations in $c_0$ are small enough that we can approximate the log-normal prior on $c_0$ as Gaussian and that the priors on $c_n$ can also be correctly approximated by a Gaussian. From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$. Beyond this, one should switch to this framework, for example to correct for un-flux calibrated data.

%\begin{deluxetable}{ll}
%\tablecaption{\label{table:} Title}
%\tablehead{\colhead{Col1} & \colhead{Col2}}
%\startdata
%\enddata
%\tablecomments{}
%\end{deluxetable}

\bibliography{disks,bayesian,master}
\bibliographystyle{hapj}
\end{document}

### 2014-03-16 lnprobs/base_lnprob.py

@@ -10,11 +10,11 @@ myDataSpectrum = DataSpectrum.open("/home/ian/Grad/Research/Disks/StellarSpectra
 myInstrument = TRES()
 myHDF5Interface = HDF5Interface("/home/ian/Grad/Research/Disks/StellarSpectra/libraries/PHOENIX_submaster.hdf5")
 
-stellar_Starting = {"temp":(6000, 6100), "logg":(3.9, 4.2), "Z":(-0.6, -0.3), "vsini":(4, 6), "vz":(12.8, 14), "logOmega":(-19.68, -19.65)}
+stellar_Starting = {"temp":(6000, 6100), "logg":(3.9, 4.2), "Z":(-0.6, -0.3), "vsini":(4, 6), "vz":(15.0, 16.0), "logOmega":(-19.665, -19.664)}
 
 stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
 
-cheb_Starting = {"c1": (-.1, 0), "c2": (-.1, .0), "c3": (-.05, .0)}
+cheb_Starting = {"c1": (-.02, -0.015), "c2": (-.0195, -0.0165), "c3": (-.005, .0)}
 cov_Starting = {"sigAmp":(0.8, 1.2), "logAmp":(-15, -13), "l":(0.1, 2.0)}
 cov_tuple = C.dictkeys_to_covtuple(cov_Starting)
 
@@ -34,8 +34,11 @@ def lnprob_Cheb(p):
 
 def lnprob_Cov(p):
     params = myModel.zip_Cov_p(p)
-    myModel.update_Cov(params)
-    return myModel.evaluate()
+    try:
+        myModel.update_Cov(params)
+        return myModel.evaluate()
+    except C.ModelError:
+        return -np.inf
 
 pool = MPIPool()
 if not pool.is_master():
@@ -46,10 +49,10 @@ myStellarSampler = StellarSampler(lnprob_Model, stellar_Starting, pool=pool)
 myChebSampler = ChebSampler(lnprob_Cheb, cheb_Starting, pool=pool)
 myCovSampler = CovSampler(lnprob_Cov, cov_Starting, pool=pool)
 
-myMegaSampler = MegaSampler(samplers=(myStellarSampler, myChebSampler, myCovSampler), burnInCadence=(1, 1, 2), cadence=(10, 10, 10))
[…]

### 2014-03-16 model.py

@@ -94,6 +94,18 @@ class Model:
         lnp = self.CovarianceMatrix.evaluate(model_fls)
         return lnp
 
+    def evaluate_cho(self):
+        '''
+        Compare the different data and models.
+        '''
+        #Incorporate priors using self.ModelSpectrum.params, self.ChebyshevSpectrum.c0s, cns, self.CovarianceMatrix.params, etc...
+
+        model_fls = self.ChebyshevSpectrum.k * self.ModelSpectrum.downsampled_fls
+
+        #CovarianceMatrix will do the math from DFM
+        lnp = self.CovarianceMatrix.evaluate_cho(model_fls)
+        return lnp
+
 ##How to use ProgressBar
 from progressbar import ProgressBar, Percentage, Bar, ETA
 

### 2014-03-16 spectrum.py

@@ -5,6 +5,7 @@ from scipy.special import j1
 import scipy.sparse as sp
 from astropy.io import ascii,fits
 from scipy.sparse.linalg import spsolve
+from scipy.linalg import cho_factor, cho_solve
 import gc
 import pyfftw
 import warnings
@@ -877,6 +878,8 @@ class CovarianceMatrix:
         #Because sparse matrices only come in 2D, we have a list of sparse matrices.
         self.sigma_matrices = [sp.diags([sigma**2], [0], dtype=np.float64, format="csc") for sigma in self.DataSpectrum.sigmas]
         self.matrices = [sp.diags([sigma**2], [0], dtype=np.float64, format="csc") for sigma in self.DataSpectrum.sigmas]
+        self.logdets = np.empty((self.norders,))
+        self.calculate_logdets()
 
 
     def sparse_k_3_2(self, xs, amp, l):
@@ -905,6 +908,16 @@ class CovarianceMatrix:
         offsets = [i for i in range(-offset + 1, offset)]
         return sp.diags(diags, offsets, format="csc")
 
+    def calculate_logdets(self):
+        #Calculate logdet for each matrix
+        for i in range(self.norders):
+            s = self.matrices[i]
+            sign, logdet = np.linalg.slogdet(s.todense())
+            if sign <= 0:
+                raise C.ModelError("The determinant is negative.")
+            self.logdets[i] = logdet
+
+
     def update(self, params):
         '''
         Update the covariance matrix using the parameters.
@@ -918,9 +931,16 @@ class CovarianceMatrix:
         l = params['l']
         sigAmp = params['sigAmp']
 
+        if (l <= 0) or (sigAmp < 0):
+            raise C.ModelError("l {} and sigAmp {} must be positive.".format(l, sigAmp))
+
+
         for i in range(self.norders):
             wl = self.DataSpectrum.wls[i]
             self.matrices[i] = self.sparse_k_3_2(wl, amp, l) + sigAmp**2 * self.sigma_matrices[i]
+
+        self.calculate_logdets()
+
 
 
     def gauss_func(x0i, x1i, x0v=None, x1v=None, amp=None, mu=None, sigma=None):
@@ -973,15 +993,29 @@ class CovarianceMatrix:
         for i in range(self.norders):
             a = A[i]
             s = self.matrices[i]
-            sign, logdet = np.linalg.slogdet(s.todense())
[…]

### 2014-03-16 tests/test_spectrum.py

@@ -380,7 +380,8 @@ class TestCovarianceMatrix:
         self.cov = CovarianceMatrix(myDataSpectrum)
 
     def test_update(self):
-        self.cov.update({"sig_amp":1, "amp":1, "l":1})
+        self.cov.update({"sigAmp":1, "logAmp":1, "l":1})
+
 
 
     #def test_chi2(self):

### 2014-04-23 simple_model.py
updated response.txt.

@@ -1,36 +1,0 @@
-import numpy as np
-
-class Data:
-    def __init__(self, x, y):
-        self.x = x
-        self.y = y
-
-class LineModel:
-    def __init__(self, x, params):
-        self.params = params
-        self.x = x
-        self.model()
-
-    def model(self):
-        self.y = self.params[0] * self.x + self.params[1]
-
-    def set_params(self,params):
-        self.params = params
-        self.model()
-
-class LineModelFlag:
-    def __init__(self, x, params):
-        self.params = params
-        self.x = x
-        self.flags = np.ones_like(self.x, dtype='b') #boolean flags to mark a data point as "good" (=1) or "bad" (=0)
-        self.model()                                 #start with all points considered good.
-
-    def model(self):
-        self.y = self.params[0] * self.x + self.params[1]
-
-    def set_params(self,params):
-        self.params = params
-        self.model()
-
-    def set_flags(self,flags):
-        self.flags = flags

### 2014-04-23 tests/interpolation_tests.py

@@ -1,106 +1,0 @@
-def interpolate_raw_test_temp():
-    base = 'data/LkCa15//LkCa15_2013-10-13_09h37m31s_cb.flux.spec.'
-    wls = np.load(base + "wls.npy")
-    fls = np.load(base + "fls.npy")
-    wl = wls[22]
-    ind2 = (m.w_full > wl[0]) & (m.w_full < wl[-1])
-    w = m.w_full[ind2]
-    f58 = load_flux_npy(5800, 3.5)[ind2]
-    f59 = load_flux_npy(5900, 3.5)[ind2]
-    f60 = load_flux_npy(6000, 3.5)[ind2]
-
-    bit = np.array([5800, 6000])
-    f = np.array([f58, f60]).T
-    func = interp1d(bit, f)
-    f59i = func(5900)
-
-    fig = plt.figure(figsize=(11, 8))
-    ax = fig.add_subplot(111)
-    ax.axhline(0, color="k")
-    ax.plot(w, (f59 - f59i) * 100 / f59)
-    ax.set_xlabel(r"$\lambda\quad[\AA]$")
-    ax.xaxis.set_major_formatter(FSF("%.0f"))
-    ax.set_ylabel("Fractional Error [\%]")
-    fig.savefig("plots/interp_tests/5800_5900_6000_logg3.5.png")
-
-
-def interpolate_raw_test_logg():
-    base = 'data/LkCa15//LkCa15_2013-10-13_09h37m31s_cb.flux.spec.'
-    wls = np.load(base + "wls.npy")
-    fls = np.load(base + "fls.npy")
-
-    wl = wls[22]
-    ind2 = (m.w_full > wl[0]) & (m.w_full < wl[-1])
-    w = m.w_full[ind2]
-    f3 = load_flux_npy(5900, 3.0)[ind2]
-    f3_5 = load_flux_npy(5900, 3.5)[ind2]
-    f4 = load_flux_npy(5900, 4.0)[ind2]
-
-    bit = np.array([3.0, 4.0])
-    f = np.array([f3, f4]).T
-    func = interp1d(bit, f)
-    f3_5i = func(3.5)
-
-    fig = plt.figure(figsize=(11, 8))
-    ax = fig.add_subplot(111)
-    ax.axhline(0, color="k")
-    ax.plot(w, (f3_5 - f3_5i) * 100 / f3_5)
-    ax.set_xlabel(r"$\lambda\quad[\AA]$")
[…]

### 2014-04-23 plotting.py
updated response.txt.

@@ -26,21 +26,6 @@ masks = np.load(base + ".mask.npy")
 #k.xaxis.set_major_formatter(FSF("%.0f"))
 #k.locator_params(axis='x', nbins=5)
 
-def plot_logg_grid():
-    fig, ax = plt.subplots(nrows=13, sharex=True, sharey=True, figsize=(11, 8))
-    #fig = plt.figure(figsize=(11,8))
-    #ax = fig.add_subplot(111)
-    ax[0].plot(m.wl, m.fl)
-    ax[0].set_ylabel("GW Ori")
-    for i, j in enumerate(np.arange(0.5, 6.1, 0.5)):
-        f = m.model(5900, -30, logg=j)
-        ax[i + 1].plot(m.wl, f / f[0])
-        ax[i + 1].set_ylabel("%.2f" % j)
-    ax[-1].locator_params(axis='y', nbins=3)
-    ax[-1].set_xlabel(r"$\lambda\quad[\AA]$")
-    ax[-1].xaxis.set_major_formatter(FSF("%.0f"))
-    fig.subplots_adjust(top=0.96, right=0.96)
-    plt.show()
 
 def compare_kurucz():
     wg_full = np.load('wave_grids/PHOENIX_2kms_air.npy')

### 2014-04-23 tests/comp_kurucz.py
updated response.txt.

@@ -1,7 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-wl, fl = np.loadtxt("GWOri_c/23.txt", unpack=True)
-#wl = 10.**wl
-
-#np.arange(len(flux),C

### 2014-04-23 StellarSpectra/model.py
added bokeh plotter. long phoenix run

@@ -123,6 +123,9 @@ class Model:
                     print("creating region ", i, region, regions_dict[str(region)])
                     CovMatrix.create_region(regions_dict[str(region)])
 
+        #Now update the stellar model again so it accounts for the Chebyshevs when downsampling
+        model.update_Model(read['stellar_params'])
+
         return model
 
     def __init__(self, DataSpectrum, Instrument, HDF5Interface, stellar_tuple, cheb_tuple, cov_tuple, region_tuple, outdir=""):
@@ -216,13 +219,22 @@ class OrderModel:
         #We can expose the RegionMatrices from the self.CovarianceMatrix, or keep track as they are added
         self.region_list = []
 
+    def get_data(self):
+        return (self.wl, self.fl)
+
     def update_Cheb(self, params):
         self.ChebyshevSpectrum.update(params)
         self.cheb_params = params
 
+    def get_Cheb(self):
+        return self.ChebyshevSpectrum.k
+
     def update_Cov(self, params):
         self.CovarianceMatrix.update_global(params)
         self.global_cov_params = params
+
+    def get_Cov(self):
+        return self.CovarianceMatrix.cholmod_to_scipy_sparse()
 
     def get_regions_dict(self):
         return self.CovarianceMatrix.get_regions_dict()

### 2014-04-23 StellarSpectra/spectrum.py
added bokeh plotter. long phoenix run

@@ -837,8 +837,8 @@ class ModelSpectrum:
 
         '''
         #First set stellar parameters using _update_grid_params
-        grid_params = {'temp':params['temp'], 'logg': 4.29, 'Z':params['Z']}
-        #grid_params = {key:params[key] for key in self.interpolator.parameters}
+        # grid_params = {'temp':params['temp'], 'logg': 4.29, 'Z':params['Z']}
+        grid_params = {key:params[key] for key in self.interpolator.parameters}
 
         self._update_grid_params(grid_params)
 
@@ -1030,9 +1030,8 @@ class ModelSpectrumLog:
         #self.interpolator.parameters automatically contains alpha or not
 
         #Kludgey code to account for Willie setting logg = 4.29
-
-        grid_params = {'temp':params['temp'], 'logg': 4.29, 'Z':params['Z']}
-        #grid_params = {key:params[key] for key in self.interpolator.parameters}
+        # grid_params = {'temp':params['temp'], 'logg': 4.29, 'Z':params['Z']}
+        grid_params = {key:params[key] for key in self.interpolator.parameters}
 
 
         self._update_grid_params(grid_params)

### 2014-04-23 StellarSpectra/plot_model.py
added bokeh plotter. long phoenix run

@@ -1,0 +1,128 @@
+#!/usr/bin/env python
+from StellarSpectra.model import Model
+from StellarSpectra.spectrum import DataSpectrum
+from StellarSpectra.grid_tools import TRES, HDF5Interface
+import StellarSpectra.constants as C
+import numpy as np
+import yaml
+import json
+
+import bokeh
+from bokeh.plotting import *
+from bokeh.objects import Range1d
+
+#Use argparse to determine if we've specified a config file
+import argparse
+parser = argparse.ArgumentParser(prog="plot_model.py", description="Plot the model.")
+parser.add_argument("json", help="*.json file describing the model.")
+parser.add_argument("params", help="*.yaml file specifying run parameters.")
+parser.add_argument("-o", "--output", help="*.html file for output")
+args = parser.parse_args()
+
+
+if args.json: #
+    #assert that we actually specified a *.json file
+    if ".json" not in args.json:
+        import sys
+        sys.exit("Must specify a *.json file.")
+
+if args.params: #
+    #assert that we actually specified a *.yaml file
+    if ".yaml" in args.params:
+        yaml_file = args.params
+        f = open(args.params)
+        config = yaml.load(f)
+        f.close()
+
+    else:
+        import sys
+        sys.exit("Must specify a *.yaml file.")
+        yaml_file = args.params
+
+
+myDataSpectrum = DataSpectrum.open(config['data_dir'], orders=config['orders'])
+myInstrument = TRES()
+myHDF5Interface = HDF5Interface(config['HDF5_path'])
+
+stellar_Starting = config['stellar_params']
+stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
+
+cheb_Starting = config['cheb_params']
+cheb_tuple = ("logc0", "c1", "c2", "c3")
+
+cov_Starting = config['cov_params']
+cov_tuple = C.dictkeys_to_cov_global_tuple(cov_Starting)
+
+region_tuple = ("h", "loga", "mu", "sigma")
+region_MH_cov = np.array([0.05, 0.04, 0.02, 0.02])**2 * np.identity(len(region_tuple))
+
+
+myModel = Model.from_json(args.json, myDataSpectrum, myInstrument, myHDF5Interface)
+
+model = myModel.OrderModels[0]
+
+#Get the data
+wl, fl = model.get_data()
+
[…]

### 2014-04-23 correlated.py

@@ -1,7 +1,7 @@
 import scipy.sparse as sp
 import numpy as np
-from numpy.linalg import inv
-from scipy.sparse.linalg import inv as sinv
+from numpy.linalg import slogdet
+from scipy.sparse.linalg import spsolve
 import matplotlib.pyplot as plt
 
 def gauss(x, amp, mu, sigma):
@@ -10,60 +10,70 @@ def gauss(x, amp, mu, sigma):
 def line(x, b, m):
     return b + m * x
 
-xs = np.linspace(-10,10)
+xs = np.linspace(-20,20)
 npoints = len(xs)
 
 #Create a continuum with a gaussian absorption line superimposed. Add Gaussian noise.
-ys = line(xs, 10, 0.2) - gauss(xs, amp=15, mu=0, sigma=1) + np.random.normal(size=npoints)
+ys = line(xs, 10, 0.05) - gauss(xs, amp=15, mu=0, sigma=1) + np.random.normal(size=npoints)
 
-def plot_line():
-    fig = plt.figure()
-    ax = fig.add_subplot(111)
-    ax.plot(xs,ys)
-    ax.set_xlabel(r"$x$")
-    ax.set_ylabel(r"$y$")
-    fig.savefig("plots/correlated_gaussian.png")
-plot_line()
+def gauss_func(x0i, x1i, x0v=None, x1v=None, amp=None, mu=None, sigma=None):
+    x0 = x0v[x0i]
+    x1 = x1v[x1i]
+    return amp**2/(2 * np.pi * sigma**2) * np.exp(-((x0 - mu)**2 + (x1 - mu)**2)/(2 * sigma**2))
+
+def Cregion(xs, amp, mu, sigma, var=1):
+    '''Create a sparse covariance matrix using identity and block_diagonal'''
+    #In the region of the Gaussian, the matrix will be dense, so just create it as `fromfunction`
+    #and then later turn it into a sparse matrix with size xs x xs
+
+    #Given mu, and the extent of sigma, estimate the data points that are above, in Gaussian, and below
+    n_above = np.sum(xs < (mu - 4 * sigma))
+    n_below = np.sum(xs > (mu + 4 * sigma))
+
+    #Create dense matrix and indexes, then convert to lists so that you can pack things in as:
+
+    #csc_matrix((data, ij), [shape=(M, N)])
[…]

### 2014-04-23 scripts/stars/base_lnprob.py

@@ -40,7 +40,7 @@ myHDF5Interface = HDF5Interface(config['HDF5_path'])
 
 stellar_Starting = config['stellar_params']
 #Note that these values are sigma^2!!
-stellar_MH_cov = np.array([2, 0.02, 0.02, 0.02, 1e-4])**2 * np.identity(len(stellar_Starting))
+stellar_MH_cov = np.array([2, 0.02, 0.02, 0.02, 0.02, 1e-4])**2 * np.identity(len(stellar_Starting))
 stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
 
 #Tuning the correlations should depend on how the models were normalized, right?
@@ -123,7 +123,7 @@ mySampler.reset()
 
 mySampler.run(config["samples"])
 mySampler.acceptance_fraction()
-myModel.to_json()
+myModel.to_json("model_final.json")
 mySampler.write()
 mySampler.plot()
 

### 2014-04-30 paper/paper.tex
Caption error
Revisions and added text to paper

@@
-\documentclass[preprint]{aastex} %double-column, single-spaced document:
-%\documentclass[iop,floatfix]{emulateapj}
+%\documentclass[preprint]{aastex} %double-column, single-spaced document:
+\documentclass[iop,floatfix]{emulateapj}
 
@@
 \usepackage{bm}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 \newcommand{\vstar}{\vt_{\star}}
+\newcommand{\finst}{f_{\lambda, {\rm inst}}}
+\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
 \newcommand{\vN}{\vt_{\rm N}}
@@
 \newcommand{\fDi}{D_i}
+\newcommand{\fR}{ {\bm R}}
 \newcommand{\dd}{\,{\rm d}}
@@
 \newcommand{\A}{[\alpha/{\rm Fe}]}
+
+\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}
+
+%% Nomenclature guide
+%%%%%%%%%%%%%%%%%%%%%
+% * collections of synthetic spectra are generically called ``libraries.'' The term 
+% ``grid'' is used when referring specifically to the spacing of spectral parameters
+% in the library
 
@@
 
-\title{Methods for the spectroscopic inference of fundamental stellar parameters}
+\title{A Method for the spectroscopic inference of fundamental stellar parameters}
 \author{\today{}\\
@@
 \section{Introduction}
-The fundamental properties of stars---effective photospheric temperature, surface gravity, and metallicity--are crucial for our understanding of pre-main sequence stellar evolution \citep{dm97, bca+02}.
-The exoplanet community is also intensely focused on deriving accurate stellar properties because the properties of newly discovered exoplanets must necessarily be quoted relative to those of their host stars (e.g.
-\citealt{tfs+12,blj+12,ssm+13}) and uncertain star properties translate into uncertain planet properties \citep{kan14}.
-Many techniques\footnote{If you are interested, I have also compiled a more verbose review of modern techniques described in the literature.
-However, for the sake of brevity I have omitted it from this document.} exist for stellar characterization such as line indexes, machine learning, and spectral synthesis\footnote{Given infinite computational power, we would prefer to fit \emph{all} of the parameters that completely describe a synthetic spectrum: atmospheric structure, individual elemental abundances, and atomic constants describing the opacity contribution of a line.
-Approximations to the synthesis and efficient MCMC sampling parallelized on a cluster might soon put this approach within reach.}.
-In order to derive accurate and robust stellar properties for our small sample ($\sim 20$) of T Tauri stars (some binary), we have decided to forward model the observed spectra by drawing from a grid of synthetic spectra.
+\begin{itemize}
+  \item Where our technique fits in the ecosystem of stellar methods.
+  \item Applicable to any kind of spectrum (long-slit, echelle, infrared, flux-calibrated or not)
+  \item Examples of fields where accurate and unbiased stellar parameters are crucial. Exoplanets. T Tauri stars.
+\end{itemize}
+
+\section{Method}
+Typically, the stellar properties we most desire are effective temperature $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
+Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz, more being developed for GAIA) and are parameterized by these fundamental stellar parameters, which together, we call $\vg$.
+Synthetic libraries provide high-resolution spectra which span a range of $\vg$ covering the main sequence of spectral types.
+The library is typically specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
+
+In addition to these fundamental parameters, a star also has several observed properties that are a function of its kinematics, geometry, and location in our galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$, total extinction $A_V$, and solid angle $\Omega$.
+We call these secondary parameters $\vpp$, because we can model their effects on the synthetic spectra in a ``post-processing'' step by convolution with an appropriate kernel or multiplication with a smooth function.
+Together, we call $\vstar = \{\vg, \vpp\}$.
+
+Using a synthetic stellar library as a backend and applying post-processing techniques, we attempt to forward-model the observed stellar spectra in order to determine the best-fit $\vstar$.
+Because properly sampled spectra have correlated pixels, we present a framework which self-consistently models these correlations.
 
@@
 $\vstar$ & $\{\vg,\vpp \}$\\
+$\finst(\lambda)$ & data spectrum\\
+$\fsynth(\lambda)$ & synthetic spectrum\\
 $c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
@@
 $\sigma_i$ & Poisson noise for a given pixel $i$\\
-$R_i$ & residuals $(\fDi - \fMi)/\sigma_i$\\
+$R_i$ & residuals $\fDi - \fMi$\\
+$\fR$ & residual vector $\fD - \fM$\\
 \hline
@@
 
-\subsection{Pixel to pixel techniques using synthetic spectral libraries}
-\label{sec:pix}
-Most forward modeling techniques compare pixels directly using some form of
+\subsection{Interpolation from synthetic spectral libraries}
+Creating a new spectrum with a specific $\vg$ requires either synthesizing a new spectrum using radiative transfer codes or interpolating from nearby grid points in the library.
+Because it is computationally expensive to synthesize a spectrum at high resolution over a wide wavelength range, we choose to interpolate.
+Spline interpolation for spectra \citep{hus12}.
+\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
+\todo{Investigate Tom Loredo's mentions of ``Cosmological Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
+
+\subsection{Post-processing synthetic spectra to match the data}
+Libraries of synthetic spectra are generally computed at high resolution ($R \geq $100,000), sampled with many pixels per resolution element, and do not include any rotational broadening or account for any instrumental effects.
+In order to transform a raw synthetic spectrum interpolated from the grid into one that matches the spectrum of a real star, we must post-process the spectrum to account for these secondary effects.
+The projected rotational velocity of the star, parameterized by $v \sin i$, broadens spectral line profiles and is mathematically described by convolution with a kernel $g_{v \sin i}(\lambda)$.
+UV, optical, and infrared spectra are acquired using a spectrograph which disperses light onto a CCD, defining a specific resolution and sampling rate for the spectrum.
+Because the resolution of the spectrograph and spacing of the pixels are different from the synthetic spectral library, we must convolve the raw spectrum with the line spread function (LSF) of the instrument $g_{\rm LSF}(\lambda)$ and resample to the exact pixels of the CCD.
+In wavelength space, the $v \sin i$ and LSF operations are represented by the convolution of the synthetic spectrum with these two kernels
 \begin{equation}
-  \chi^2 = \sum_i \left [\frac{ \fDi - \fMi}{\sigma_i} \right ]^2 = \sum_i R_i^2
+  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) \otimes \fsynth(\lambda)
+  \label{eqn:convolution}
 \end{equation}
-where the variables are defined in Table~\ref{tab:nomenclature}.
-A serious problem with this approach is that synthetic spectra are not perfect: systematic errors invalidate the $\chi^2$ assumption and bias parameter estimates and their uncertainties.
+Using the convolution theorem, we can rewrite these operations as multiplications in Fourier space 
+\begin{equation}
+  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
+\end{equation}
+where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform pairs.
+In order to use the fast Fourier transform (FFT) to execute these operations using discrete samples of the synthetic spectrum $\fsynth$, we must first resample $\fsynth$ so that it is on a uniform grid.
+In the case of a spectrum, the natural uniform grid is one that is equally sampled in velocity space, such there is an equal velocity shift $\Delta v$ between pixels.
+This results in a wavelength grid that is linearly sampled in the logarithm of wavelength.
+Therefore, the Fourier coordinate $s$ is the number of cycles per sampling interval, having units of inverse velocity $[{\rm s}/{\rm km}]$.
+Next, we multiply the Fourier transform of the synthetic spectrum by the Fourier transforms of the rotational velocity and line spread function kernels.
+Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda, {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength locations of the pixels in the detector ($\finst$).
+When resampling the spectrum, we are careful to ensure band-limited interpolation using splines in order to prevent introducing non-physical high frequency structure into the spectrum.
 
-Current methods either circumvent this problem by using high quality synthetic grids that were painstakingly tuned over a narrow wavelength range \citep[$\lesssim 500$ \AA]{blj+12,sb13}, or they use less-perfect grids with full optical coverage and sigma clip the bad regions of the spectrum \citep{kpb+09,mga13}.
-Masking (either implicitly or explicitly, respectively) is troublesome because, in the words of \citet{mga13}, ``regions with [only] modest agreement between the real and synthetic spectra [that might be masked] may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.'' Masking could very well destroy stellar information in the pursuit of a better fit.
+Finally we apply corrections for interstellar extinction and Doppler shift.
+Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which is wavelength-dependent.
+We use the (name here) dust-extinction model, which is parameterized by extinction in the $V$ photometric band, $A_V$.
+We correct for a radial velocity difference $v_z$ between the star and the earth by Doppler shifting the model spectrum.
 
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics{lineclasses}
-caption{The four classes of line behavior.
-\textbf{Class 0}: the model behaves as expected.
-\textbf{Class I}: a spectral line is present in the data but missing from the model (in this case, the line is also blended).
-\textbf{Class II}: a spectral line exists in the model where there is only continuum in the data.
-\textbf{Class III}: lines exist in both the data and model, but some are too strong while others are too weak.
-The lower panels show the residual spectra in units of the Poisson uncertainty.
-The data is a spectrum of WASP-14, an F star.
-The model is from the PHOENIX grid\citep{hwd+13}.}
-\label{fig:lineclasses}
-\end{center}
-\end{figure}
+\todo{Convolution in wavelength space is a possibility. Allows LSF to change with wavelength.}
+Pixel width effects are not important as long as the LSF of the instrument is adequately sampled ($\gtrsim 3$ pixels across the LSF).
+If our spectrum were not properly sampled, we would need to additionally convolve with a boxcar the width of the pixels.
 
-\subsection{Classes of synthetic spectral line behavior}
-Illustrated in Figure~\ref{fig:lineclasses} are three ways that synthetic spectra often diverge from reality.
-When a spectral line appears in the data but not in the model, due to interstellar or telluric absorption or a missing opacity source, we call this a \textbf{Class I} error.
-When a spectral line appears in the model but not in the data, presumably due to a combination of incorrect atomic constants or atmospheric structure, we label it a \textbf{Class II} error.
-Both of these types of errors are present over a wide range of probable stellar parameters (1000 K in $T_{\rm eff}$, and 1 dex in $\log(g)$, $\Z$, and $\A$) and there does not exist a $\vg$ that could ever fit these lines, making these regions useless for determining stellar parameters.
-We know that for a normal AFGK star Class I residuals will always be negative and will not change significantly with $\vg$, while Class II residuals will always be positive and their magnitude should decrease with decreasing metallicity.
-To help identify Class I and II errors,  we could use the partial derivatives of an individual synthetic spectrum with respect to $\vg$ and a prior on the range of parameters over which the residuals will be correlated to determine which lines show Class I or II behavior.
+\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in Fourier space, but it's pretty fast as-is.}
 
-When some line strengths do not match at a $\vg$ that properly fits most other lines, we term these \textbf{Class III} errors.
-These types of errors are pernicious because there does exist a $\vg$ that would fit each line individually, but no one $\vg$ can be chosen that fits all the lines.
-We must then decide which regions to trust and by how much, because Class III regions \emph{still contain information about $\vg$}.
-In fact, they might contain \emph{the most} information about $\vg$ since these delicate regions are the most sensitive to changes in $\vg$ and are consequently whey they are the hardest to model correctly.
-In other words, Class III regions might be where the spectrum gradient ($\sum_{j} \partial \fMi/ \partial \vt_{ \star {\rm grid}, j}$) is the largest.
+Due to small imperfections in the flux-calibration of the data spectrum, there may be broad regions of the spectrum which are offset from the true continuum level by a slight amount.
+A traditional approach to address this problem is to normalize both the data and the spectrum to the continuum level before comparison.
+While this may work well for hotter stars with well-defined continuua, for cooler stars with large molecular features, normalization may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.
+Instead, we choose to multiply the model spectrum by a low-order Chebyshev polynomial that is normalized to unity.
+The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters in our model that we will solve for.
+If the behavior of the spectrograph is characterized by observations of spectrophotometric standards, then we can put priors on the coefficients that will prevent overfitting of the polynomials.
 
-\section{In the limit of a perfect stellar model}
-To generate a model spectrum $\fM(\vg, \vpp)$, we use $\vg$ to linearly interpolate a synthetic spectrum from the grid, and then use $\vpp$ to apply instrumental and rotational broadening in the Fourier domain, Doppler shift, correct for extinction, and then downsample to the exact pixels of the data spectrum using spline interpolation\footnote{Because we are dealing with flux-calibrated data and synthetic spectra, we \emph{resample} to lower resolution.
-We do not \emph{rebin} because we are dealing with flux-density, not photon counts}.
+With more uncertainty, the polynomial multiplication can also be used as a substitute for flux-calibration since the sensitivity function of an instrument is usually a smooth function with wavelength.
+When spectra are not flux calibrated, there is greater uncertainty about the properties of the spectrograph and wider priors must be used on the Chebyshev coefficients.
+In this case, the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by the flux calibration for the model.
 
-To account for imperfect flux calibration, we extend the approach of \citet{elh+06} and multiply $\fM(\vg, \vpp)$ by $k(\vN)$, which is a Chebyshev polynomial described by a set of $N$ Chebyshev coefficients.
-We can either sample in the coefficients or analytically marginalize over them.
-We favor this approach over the traditional technique of normalizing to the stellar continuum due to concerns about correctly placing the pseudo-continuum in M stars.
-From now on we refer to $\fM(\vt) = k(\vN) \fM(\vg, \vpp)$.
-The mathematical details are in \S~\ref{sec:chebyshev}.
+\subsection{Likelihood function}
+To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we use a standard multidimensional Gaussian likelihood function which allows for covariance between data points.
+If the residuals vector is
+\begin{equation}
+  \fR(\vstar) = \fD - \fM(\vstar)
+\end{equation}
+with $N$ data points, then the likelihood function is
+\begin{equation}
+  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2} \fR^T C^{-1} \fR \right ) 
+\end{equation}
+and the log-likelihood function is
+\begin{equation}
+    \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  - \frac{N}{2} \ln 2 \pi 
+    \label{eqn:lnprob}
+\end{equation}
+$C$ is the covariance matrix which describes the covariance between pixels in the spectrum. For independent noise with noise per pixel $\sigma_i$, this matrix is
+\begin{equation}
+  C = 
+  \begin{bmatrix}
+    \sigma_0^2 & 0  & \cdots & 0\\
+    0 & \sigma_1^2 & \cdots & 0\\
+    \vdots  &   & \ddots  & \vdots \\
+    0 & 0 & \cdots & \sigma_N^2\\
+  \end{bmatrix}
+\end{equation}
+and Equation~\ref{eqn:lnprob} reduces to the familiar $\chi^2$ form of a sum over the square of the residuals, weighted by the inverse variance of each data point
+\begin{equation}
+  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
+\end{equation}
+However, residuals from a spectroscopic fit are often correlated due to the convolution of the stellar rotation and LSF kernels.
+Additionally, systematic errors in the synthetic spectra (such as incorrect line strengths) will resulted in regions of highly correlated residuals, which requires that we use a covariance matrix with off-diagonal terms $\sigma_{ij}$ and Equation~\ref{eqn:lnprob}, which includes a matrix product instead of a sum over independent pixels.
 
-Currently we are sampling this $\chi^2$ posterior with the MCMC ensemble sampler \citep{gw10} implemented in \texttt{Python} by \citet{fhl+12} as \texttt{emcee}.
-In this section, we described our approach to sample the posterior assuming perfect synthetic spectra.
-However, if one does a straight up $\chi^2$ fit to our spectra, one can get posteriors that have 1-sigma contours of $\pm 1$ K and $\pm 0.01$ dex in gravity and metallicity.
-This would be as precise as stellar parameters obtained with astroseismology!
-Clearly there is something going wrong here.
+In order properly account for these correlated residuals, we must understand the covariance structure and parameterize it in our model.
+We seek to understand two types of covariance structure: the large scale global covariance structure that results from the stellar rotation and LSF kernels, and a regional covariance structure that is localized to areas where spectral line mismatch results in small patches of highly correlated residuals.
 
-\section{Beyond a perfect stellar model}
-\label{sec:residuals}
-\subsection{Additional sources of noise}
-\begin{itemize}
-  \item CCD readnoise and imperfect 1D spectrum extraction over night sky lines will almost definitely increase the noise in a pixel beyond theoretical Poisson limits.
-Normally this information will be encapsulated in a ``sigma spectrum'' output by IRAF, but unfortunately we do not have access to this for our current dataset.
-  \item The stellar continuum can flicker/jitter (as shown by Kepler), which would produce a small departure from a perfect continuum.
-This might explain why even in the Class 0 region of our spectrum (Figure~\ref{fig:lineclasses}, left panel) the residuals are slightly larger than 1 $\sigma$ on average.
-For our T Tauri stars this effect may be more pronounced.
-  \item When we interpolate a synthetic spectrum from the library to a specific $\vg$, we will introduce interpolation errors depending on how close we are to a grid point in the library.
-\citet{hus12} investigate the possible range of interpolation errors at specific $\vg$ by dropping out a spectrum from the grid, interpolating over it, and then comparing the true spectrum to the interpolated spectra.
-For hotter stars with well defined continua, we find the maximum error is $\lesssim 5$\%, while for M stars the error could be as large as $\sim 20$\% for some spectral features.
-By doing this interpolation test across the grid, we could determine an ``interpolation error spectrum'' that would encapsulate the maximum error possible for an interpolated spectrum, weighted by its distance from the nearest grid points.
-If a typical spectrum has signal/noise of $\sim 30$ per pixel (or noise/signal ratio of $1/30 \approx 3$\%), an interpolation noise contribution of $\sim 5$\% the level of the signal level would be significant.
-\end{itemize}
-However, we have tried arbitrarily increasing the Poisson errors by a factor of 2 or 3 and it didn't significantly change the results, suggesting that something more drastic is needed beyond simply adding in more noise.
+\todo{Tom Loredo was saying that if we have a model that perfectly fits the data then our residuals will be uncorrelated, regardless of the PSF/LSF.
+But is this true?
+Shouldn't all noise be correlated on some scale by the LSF?
+Correspond by email to clarify and show an example.}
 
-\subsection{Gallery of likelihood functions}
-\label{sec:gal}
-A problem with the $\chi^2$ likelihood function is that the Gaussian tails rapidly decline, making outliers extremely improbable.
-A common approach for ``robust regression'' is to use a likelihood function with fatter, more forgiving, tails.
-Although not rigorously motivated, this is can be a quick thing to try.
-We explored several likelihood functions including a combination of Gaussian with exponential wings, a Lorentzian (Cauchy) function, and a function described  by \citet{ss06}, which treats the noise estimate for each data point suspiciously.
-In Sivia's model, the true noise $\sigma$ can be no smaller than some original estimate $\sigma_0$, but it may be much larger.
-By putting a prior on $\sigma$,
+To account for the global covariance structure, we use a Matern 3/2 kernel.
+\todo{Explore difference between using 3/2 Matern kernel and squared exponential kernel.}
 \begin{equation}
-  p(\sigma_i | {\sigma_0}_i) = \left \{ \begin{array}{cc}
-    {\sigma_0}_i/\sigma_i^2 & \sigma_i  \leq {\sigma_0}_i\\
-    0                 & \textrm{else}
-  \end{array}
-    \right.
+  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp \left (- \frac{\sqrt{3} r}{l} \right )
 \end{equation}
-and marginalizing over the $\chi^2$ likelihood, the new likelihood function becomes
+
+tapered for compact support using a Hann window with width $r_0 = 6 l$ (Why? Because this looked good).
+
+Specific line covariance. Non-stationary covariance kernel.
 \begin{equation}
-p(\fDi | \fMi, {\sigma_0}_i) = \frac{1}{ {\sigma_0}_i \sqrt{2 \pi}} \left [ \frac{1 - e^{-R_i^2/2}}{R_i^2} \right ] 
-\end{equation}
-where $R_i = (\fDi - \fMi)/{\sigma_0}_i$.
-This function has even fatter tails than the exponential model and seems to give results less biased by outliers, however it is not truly satisfying.
-Which lines contributed most to the posterior?
-What about the wings of class I and II errors?
-These individual pixels might still be given a weight of ${\sigma_0}_i$ because they are not nearly as bad as the line centers.
-
-Other functions we might try in the future include a Student-$t$ distribution, a linear combination of Gaussians, or the $\chi^2_Q$ function described in \citet{hbl10}.
-While ``softer'' likelihood functions do help somewhat, we would like to move beyond tweaking the fitting function towards an approach that properly incorporates our knowledge of the systematic errors of the synthetic grids.
-In the remaining two sections of this document we outline potential ideas that we might try next, but have not yet implemented in code.
-
-\section{Bad data mixture model}
-The bad data model \citep{pre97,hbl10} is a powerful framework for forward modeling data, which allows the modeler to explicitly account for outliers using a generative model.
-In the simplest mixture model, each pixel has a binary flag $q_i$ which specifies whether that data point is good or bad.
-Then, for the entire data set we have an ensemble of $\qN$.
-If this ensemble of $\qN$ is chosen appropriately, and the good pixels are flagged as good and the bad pixels are flagged as bad, the posterior will be maximized.
-We adopt $P_B$ for the \emph{a priori} probability that a given pixel will be bad, $Y_b$ for the mean flux of the bad pixels, and $V_b$ for their variance.
-Written out explicitly, this is
-\begin{multline}
-  {\cal L} = p({\bm D} | \vt,\qN, Y_b,V_b) = \prod_{i=1}^N \left [ \frac{1}{\sqrt{2 \pi} \sigma_i} \exp \left[ - \frac{(\fDi - \fMi)^2}{2 \sigma_i^2} \right] \right ]^{q_i}  \times \\
-  \left [ \frac{1}{\sqrt{2 \pi (V_b + \sigma_i^2)}} \exp \left[ - \frac{(\fDi - Y_b)^2}{2 (V_b + \sigma_i^2)} \right]
-  \right ]^{1 - q_i}
-\end{multline}
-If we put a Bernoulli prior on the $q_i$
-\begin{equation}
-  p(q_i | P_b) = [1-P_b]^{q_i} P_b^{[1-q_i]}
-\end{equation}
-the probability of any given ensemble of $\qN$ is 
-\begin{equation}
-  p(\qN | P_b) = \prod_{i=1}^N [1-P_b]^{q_i} P_b^{[1-q_i]}
-\end{equation}
-We can analytically marginalize over the settings for a single flag by a discrete sum over the two states of $q_i = 1$ and $q_i = 0$. 
-\begin{multline}
-  p(\vt, P_b, Y_b, V_b | D_i)  =  \int p( \vt, q_i, P_b, Y_b, V_b| D_i) {\rm d} q_i \\
-  = p(\vt, q_i=1, P_b, Y_b, V_b | D_i) + p(\vt, q_i=0, P_b, Y_b, V_b | D_i) \\
-  = \Bigl[ p_{\rm good}(D_i | \vt)\, p(q_i=1 | P_b) +  p_{\rm bad}(D_i | Y_b, V_b)\, p(q_i=0 | P_b)\Bigr] \times p(\vt ,P_b, Y_b,V_b) 
-\end{multline}
-
-\begin{multline}
-  p(\vt,P_b,Y_b,V_b|\,{\bm D}) \propto \prod_{i=1}^N \Bigl [ [1 - P_b] p_{\rm good}(D_i |\vt)  + P_b p_{\rm bad}(D_i | Y_b, V_b)\Bigr ] \times p(\vt,P_b, Y_b, V_b)\\
-  \boxed{
-    = \prod_{i=1}^N \left\{ \frac{1 - P_b}{\sqrt{2 \pi \sigma_i^2}} \exp \left[ - \frac{(\fDi - \fMi)^2}{2 \sigma_i^2} \right] + \frac{P_b}{\sqrt{2 \pi (V_b + \sigma_i^2)}} \exp \left[- \frac{(\fDi - Y_b)^2}{2 (V_b + \sigma_i^2)} \right] \right\}
-  \times p(\vt,P_b,Y_b,V_b)}
-\end{multline}
-
-In its simplest form, the bad data model results in a likelihood function that is a linear combination of Gaussians.
-Although this functional form might be similar to some of the likelihood functions in \S~\ref{sec:gal}, the benefit of this bad data approach over a gallery of likelihood functions is that it is motivated by the characteristics of the bad data ($P_B$, $Y_b$, and $V_b$).
-If we wish to add additional information about the characteristics of the bad data, we can do so.
-
-\subsection{Mixture of multiple components}
-When we believe that a sample population might be composed of several populations with different behaviours (Class 0, I, II, and III), we can expand the bad data model to a multiple mixture model.
-Following \citet{gcs+04}, we develop a mixture model with $M=4$ components, one for each class of error.
-Our likelihood function is then
-\begin{equation}
-  p(\fD, \vec{Q} | \vt, J) = \prod_i^N \prod_{m=0}^M \left [ J_m {\cal L}_m(\fDi | \vt_m) \right ]^{q_{im}}
-\end{equation}
-where $\vec{Q} = \qM$ is a two dimensional array ($N \times M$) of all the individual $q_{im}$ flags, and $J_m$ is a weight that represents the total fraction of each component in the mixture\footnote{The weights are normalized $\sum_m J_m = 1$}, and ${\cal L}_m$ is the likelihood function for $m$th mixture component.
-At this point, $\vt$ now contains all of the parameters that might be used by any of the likelihood functions.
-$\vt_m$ is used to denote only those parameters that will be used for ${\cal L}_m$.
-Given what we know about these synthetic spectral errors, we could specify the following likelihood functions
-\begin{equation}
-  \begin{array}{lcccc}
- {\cal L}_0 &= p(\fDi, q_{i0} = 1) &\rightarrow & \textrm{Class 0} & \chi^2\\
- {\cal L}_1 &= p(\fDi, q_{i1} = 1) &\rightarrow & \textrm{Class 1} & \textrm{flat lnprob}\\
- {\cal L}_2 &= p(\fDi, q_{i2} = 1) &\rightarrow & \textrm{Class 2} & \textrm{flat lnprob}\\
- {\cal L}_3 &= p(\fDi, q_{i3} = 1) &\rightarrow & \textrm{Class 3} & \textrm{soft } \chi^2\\
-\end{array}
-\end{equation}
-This would treat Class I and II errors as if they had been masked from the dataset but allow Class III to have a slight influence on the posterior since these regions still contain information about $\vstar$.
-The softening term of the $\chi^2$ could even be sampled for, much like the $V_b$ of the original bad data model.
-
-\subsection{Spatial correlations of pixels}
-Previously we assumed that each data point is independent, but we know that the spectrum is spatially correlated across several pixels with a correlation length set by the width of the convolution of the stellar rotation kernel ($v \sin i$) and the instrumental resolution kernel, which is also the typical width of a spectral line\footnote{Some very strong lines (such as the Balmer lines or Mg b triplet) will have wings with a larger correlation length.}.
-This correlation length will have some dependence on the spectral type of star.
-For example, F and G stars will likely have singular, well-defined lines while M stars might have a large molecular feature covering a wide swath of the spectrum.
-Neighboring pixels are actually correlated in two coordinates: the $q_{im}$ flags and the residuals $R_i$.
-
-\subsubsection{Correlation of the $q_i$ flags}
-The pixels at the center of a bad line will give a strong residual response, which will likely be flagged as bad.
-However, the pixels in the wings of the spectral line might not be sufficiently egregious to be properly flagged.
-To remedy this situation, we attempt to develop a bad data model which includes spatial (wavelength) correlations.
-
-In the manner of Gaussian random processes (GRP), we develop a covariance function
-\begin{equation}
-  {\rm cov}(q_i, q_j) = \rho(| i - j|)
+  k_{\rm G}(x, x^\prime | a, \mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{[(x - \mu)^2 + (x^\prime - \mu)^2]}{2 \sigma^2}\right )
 \end{equation}
 \begin{equation}
-  \Sigma_{ij} = \sigma^2 \rho(|i - j|)
+k(x, x^\prime | h, a, \mu, \sigma) = \exp \left ( \frac{-( x - x^\prime)^2 }{2 h^2} \right ) k_{\rm G}(x, x^\prime | a, \mu, \sigma)
 \end{equation}
-that describes the correlation between the state of the bad data flags.
-The simplest thing to do would be to assume a common correlation length in the spectrum.
-If we say that the typical feature will have three contiguous pixels\footnote{Ideally the covariance function should be a function over $\lambda_i$ in velocity space, since all spectral features should have a similar velocity width.
-For the purpose of keeping the math simple for the moment, we use pixel index and assume the pixels are uniformly spaced.} wrong, then the correlation function would be
-\begin{equation}
-  {\rm cov}(q_i, q_j) = \left \{ \begin{array}{cc}
-    1 - \frac{|i - j|}{3} & |i - j| < 3 \\
-    0                                    & \textrm{else}
-\end{array}
-\right .
-\end{equation}
-and the matrix would be
-\begin{equation}
-  \begin{bmatrix}
-    1 & 2/3 & 1/3 & 0 & \cdots & 0\\
-    2/3 & 1 & 2/3 & \ddots & \ddots & \vdots\\
-    1/3 & 2/3 & \ddots & \ddots & \ddots & 0\\
-    0  & \ddots & \ddots & \ddots & 2/3 & 1/3\\
-    \vdots & \ddots & \ddots & 2/3 & 1 & 2/3\\
-    0      & \cdots & 0 & 1/3 & 2/3 & 1\\
-  \end{bmatrix}
-\end{equation}
-Because our correlation function has compact support, we now have a sparse covariance matrix. This can potentially speed up a matrix computation that would otherwise be very slow.
+$h$ is a "bandwidth." If $h$ is small, then there will be high-frequency structure. If $h$ is large, then only low-frequency structure will remain.
 
-\subsubsection{Correlation of adjacent residuals}
-Not only will the flags $q_i$ be spatially correlated but so will the values of the residuals $R_i$.
-For OBAFGK stars, the residuals are likely to be shaped like Gaussians.
-If $i$ specifies the pixel at the line center of a large residual feature, then the neighboring $i \pm 1,2,\ldots$ pixels will have also have similar residuals but of declining magnitude $|R_{i\pm 1, 2, \ldots} | \lesssim |R_i|$.
-For M stars, there might be a wider correlation length and the shape of the lines might not be well-described by a Gaussian.
+These are all parameters that we can sample in. Also, there would be a set of these parameters for each line in the model.
 
-I am new to the field of GRP, but it seems like it might be possible to come up with a covariance function that would describe the predicted heights of the residuals and regress against them.
-The covariance function between $R_i$ could be learned from the spectrum itself.
-Then, the likelihood functions for each mixture component in the model would be custom tailored for the type of error that it expected to encounter.
-For example, Class 0 residuals would be flat while Class I, II, and III residuals would have some sort of Gaussian shape.
-Class I residuals would be negative while Class II residuals would be positive.
-I plan to further investigate covariance functions.
+The benefit mainly comes from simply modelling these systematic residuals to begin with, since a model with covariance is far more likely than forcing the fit, and far more justified than masking regions which do not fit.
 
-\subsubsection{Further ideas}
-\begin{itemize}
-  \item \citet[ch. 18]{gcs+04} list several references further describing inference for spatially-correlated data and I plan to work my way through these. 
-  \item Naively, it seems to me like there might be a way to use FFT methods to help speed up the correlation calculation, similar to Gaussian random fields.
-  \item My preliminary experiments with GRP reveal that there is an art in choosing the functional form of ${\rm cov}(i, j)$ and setting the hyperparameters.
-Perhaps these could also be parameters that we sample in.
-I am working through \citet{rw05} and \citet[Ch 21.]{gcs+13} to learn more.
-\end{itemize}
-
-\subsection{Hyperparameters}
-\citet{mb13} discuss a technique to deal with systematic errors when combining different data sets for the combined inference of a single problem.
-They use a hyperparameter approach that can artificially inflate the variance of each data set to find a common posterior and allow for correlations between data sets.
-We could set up a similar hyperparameter approach that can tune the likelihood contribution for certain lines, treating each line as they do different datasets.
-If all of the lines were in harmonious agreement, this hyperparameter would undergo ``shrinkage'' \citep{kru10} and tighten the posterior on $\vstar$.
-If the line fits were incompatible, this hyperparameter would increase while also broadening our posterior on $\vstar$.
-
-\subsection{Sampling ideas for the mixture model with spatial correlations}
-If we use multidimensional Gaussian likelihood functions and correlation functions with compact support, it might still be possible to analytically marginalize over the multiple mixture components.
-If we cannot marginalize, then we need alternate sampling methods to explore the parameter space in a computationally efficient manner.
-A stellar model with flags for each pixel would have $\vstar + (N \times M)$ parameters, where for a typical high resolution optical spectrum $N >$ 100,000.
-
-An alternative way to approach this problem is to do Gibbs sampling where we propose a jump in a single $q_{im}$ at a time.
-Such a sample requires none of the previous heavy lifting (spectral grid interpolation, rotational broadening, or pixel interpolation), and so even though there are $\sim$ 100,000 flags to sample, this might be computationally feasible.
-
-In order to ``trim the tree,'' \citet{gcs+04} suggest to start with a crude estimate (such as $k$-means clustering) to tentatively identify which observations have been drawn from which mixture components.
-They also suggest that the expectation-maximize (EM) algorithm could also be used to sample this posterior, although this is an algorithm that I need to research further.
-
-\section{Injection of fake lines}
-As an alternative approach to modeling the regions of the spectrum suffering from systematics, we could attempt to patch the models by injecting fake lines into the spectrum to compensate the errors.
-Each line could be a Gaussian parameterized by $\lambda_0$, strength, and FWHM.
-The line strength distribution could be described by a hyper prior, and the FWHM described by a hyper parameter, since all lines should have pretty much the same FWHM.
-We could penalize the model fits by having a prior on the number of bad lines that can be created.
-Given a bad line and the linelist that produced the synthetic models, it may be possible to identify other regions of the spectrum that might be bad.
-For example, if a vanadium line is bad, then we could check to see if the other vanadium lines might also be bad.
-
-Class III errors might be difficult to address in this paradigm since we would be fitting the difference of two lines, but if they are both Gaussians of the same FWHM, then it should still work.
-In M star spectra, rather than adding a Gaussian, we might need to add a larger continuum feature, which might be difficult to parameterize.
-
-\subsection{Piecewise metallicity model}
-In theory, it is possible to fit the profiles of only certain elements to determine the abundances of these elements.
-By masking all non-alpha elements, \citet{hus12} use PHOENIX spectra \citep{hwd+13} to derive alpha abundances of stars.
-Rather than masking, we could assemble a piecewise spectrum where regions of the spectrum might be drawn from a different combination of $\vg$, say with lower metallicity for some lines in order to account for a bad line strength.
-These regions would have to be normalized properly so that the composite spectrum is sufficiently smooth.
-Coupled with a linelist, one could also tweak individual elemental abundances.
-
-\clearpage
-\appendix
-\section{A technique to account for imperfect flux calibration}
-\label{sec:chebyshev}
-
-The first four Chebyshev polynomials are 
-\begin{align*}
-  T_0(x) &= 1\\
-  T_1(x) &= x\\
-  T_2(x) &= 2 x^2 - 1\\
-  T_3(x) &= 4 x^3 - 3x\\
-\end{align*}
-In our implementation, we map the full pixel range of a specific order $i \in [i_{\rm min}, i_{\rm max}]$ to $x \in [-1, 1]$.
-By mapping to pixel number rather than wavelength, we are insensitive to Doppler shifts of the model.
-If the spectrum has multiple orders, like echelle spectra, we would have a set of Chebyshev coefficients $\vN = \{ c_0, c_1, c_2, \ldots, c_N \}$ for each order.
-\begin{align}
-  k(i | \vN) &= c_0 T_0(i) \left [1 + c_1 T_1(i) + \ldots + c_N T_N(i) \right ]\\
-  k(i | \vN) \fMi &=  \left [1 + \sum^N_{j = 1} c_j T_j(i) \right ] (c_0 \fMi)\\
-  \label{eqn:k_new}
-\end{align}
-This functional form of $k$ allows $c_0$ to scale by large amounts while the higher order terms are perturbations on top of the \emph{scaled} model ($c_0 \fMi$).
-This formulation permits large rescalings up or down without biasing the higher order terms.
-If $c_0$ is expected to be small, we can use a simpler implementation described in \S\ref{sec:gaussian_simplification}.
-Then for the full spectrum
-\begin{equation}
-  \ln \bigl [p(\vt | \fD) \bigr] \propto - \frac{1}{2} \sum_i \left [\frac{\fDi - \bigl [1 + \sum^N_{j = 1} c_j T_j(i) \bigr ] c_0 \fMi }{\sigma_i} \right ]^2
-  \label{eqn:lnprob2}
-\end{equation}
-To simplify the following discussion, we consider a single pixel $i$ and then generalize this to a sum over all pixels later. We expand out the square in Equation~\ref{eqn:lnprob2}
-\begin{equation}
-  \ln \bigl [p(\vt | \fDi) \bigr] \propto - \frac{1}{2 \sigma_i^2} \left [ \fDi^2 - 2 \fDi c_0 \fMi \left [1+ \sum_{j=1}^N c_j T_j(i) \right ]+ c_0^2 \fMi^2 \left [1 + 2 \sum_{j=1}^N c_j T_j(i) + \sum_{j=1}^N \sum_{k=1}^N c_j T_j(i) c_k T_k(i) \right ] \right ]
-  \label{eqn:expanded}
- \end{equation}
-To simplify the notation, we can rewrite the higher order Chebyshev coefficients and polynomials as column vectors $\vc$ and $\vec{T}(i)$, respectively
-\begin{equation}
-  \vc = 
-  \begin{bmatrix}
-    c_1\\
-    c_2\\
-    \vdots\\
-    c_N
-  \end{bmatrix}
-  \hspace{3cm}
-\vec{T}(i) = 
-\begin{bmatrix}
-T_1(i)\\
-T_2(i)\\
-\vdots\\
-T_N(i)\\
-\end{bmatrix}
-\end{equation}
-then 
-\begin{equation}
-  k(i| \vN) = c_0 \left [ 1 + \sum^N_{j = 0} c_j T_j(i) \right ] = c_0 \left[ 1+ \vec{T}^\trans \cdot \vc \right]
-\end{equation}
-If we let 
-\begin{equation}
-  {\bm W}(i) = \vec{T} \cdot \vec{T}^\trans = 
-  \begin{bmatrix}
-T_1 T_1 & T_1 T_2 &  \hdots & T_1 T_N \\
-T_2 T_1 & T_2 T_2 &  \hdots & T_2 T_N \\
-\vdots  & \vdots  &  \ddots & \vdots \\
-T_N T_1 & T_N T_2 &  \hdots & T_N T_N \\
-  \end{bmatrix}
-\end{equation}
-then we have
-\begin{equation}
-  \sum_{j =1}^N \sum_{k=1}^N c_j T_j(i) c_k T_k(i) = \vc^\trans \cdot {\bm W} \cdot \vc
-\end{equation}
-We can rewrite Equation~\ref{eqn:expanded} quadratically in $\vc$ as
-\begin{equation}
-  \ln \bigl [p(\vt | \fDi) \bigr] \propto - \frac{1}{2} \frac{c_0^2 \fMi^2}{\sigma_i^2}  \vc^\trans {\bm W} \vc + \left (\frac{- c_0^2 \fMi^2 + c_0 \fDi \fMi}{\sigma_i^2} \right ) \vec{T}^\trans \vc -  \left( \frac{c_0^2 \fMi^2 - 2 c_0 \fDi \fMi + \fDi^2}{2 \sigma_i^2} \right)
-\end{equation}
-Because matrix multiplication is associative, we can sum ${\bm W(i)}$ and $\vec{T}^\trans(i)$ across all pixels and define
-\begin{align}
-  {\bm A} &= c_0^2 \sum_{i} \frac{\fMi^2}{\sigma_i^2} {\bm W}(i) \\
-  \vec{B} &= \sum_{i} \frac{-\fMi^2 c_0^2 + \fDi \fMi c_0}{\sigma_i^2} \vec{T}(i)\\
-  g &= -\frac{1}{2} \sum_{i} \frac{c_0^2 \fMi^2 - 2 c_0 \fDi \fMi + \fDi^2}{\sigma_i^2} 
-\end{align}
-Where ${\bm A}$, $\vec{B}$, and $g$ are each a function of $\vstar$, $\vN$, and $\fD$. Now we can rewrite Equation~\ref{eqn:lnprob2} as 
-\begin{equation}
-  p(\vt | \fD) = p(\vstar, \vN | \fD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right )
-  \label{eqn:lnprob_matrix}
-\end{equation}
-Generally, we are most interested in the stellar parameters after they have been marginalized over the nuisance parameters.
-Because this probability function is a multi-dimensional Gaussian in the higher order Chebyshev coefficients $\vc = \{c_1, c_2, \ldots, c_N \}$, we can do this marginalization analytically \citep{sgd+09}, and then sample from the marginalized distribution.
-\begin{equation}
-  p(\vstar, c_0 | \fD) = \int p(\vstar, c_0, \vc\; | \fD) \dd \vc
-\end{equation}
-\begin{align}
-  p(\vstar, c_0 | \fD) &\propto \int \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans \vc + g \right ) \dd \vc\\
-  p(\vstar, c_0 | \fD) &\propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A} \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^\trans {\bm A}^{-1} \vec{B} + g \right )}
-\end{align}
-By doing this marginalization we drastically reduce the dimensionality of our parameter space, especially if our echelle spectrum has multiple orders.
-
-\subsection{Including Gaussian priors on the nuisance parameters}
-\label{sec:priors}
-Using this polynomial formalism, it is possible to incorporate prior knowledge about the degree of necessary ``re-fluxing'' through priors on the nuisance parameters, determined by observing several spectrophotometric standard stars.
-Because $c_0$ is a scale factor, it should be equally probable to scale up by a factor of two as it is to scale down by a factor of two.
-This means we should be using a log-normal prior on $c_0$
-\begin{equation}
-  p(c_0) = \frac{1}{\sqrt{2 \pi} \sigma_{c_0} c_0} \exp \left( -\frac{(\ln c_0)^2}{2 \sigma_{c_0}^2} \right)
-\end{equation}
-centered on $c_0 = 1$. Because the higher order polynomial terms are perturbations on the scaled model ($c_0 \fMi$), we assume Gaussian priors centered about 0
-\begin{equation}
-  p(c_n) = \frac{1}{\sqrt{2 \pi} \sigma_{c_n}} \exp \left( - \frac{c_n^2}{2 \sigma_{c_n^2}} \right)
-\end{equation}
-which we can express in vector form as
-\begin{equation}
-  p(\vc\;) \propto \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right )
-  \label{eqn:nuisance_prior} 
-\end{equation}
-where 
-\begin{equation}
-  {\bm D} = 
-  \begin{bmatrix}
-    \sigma_1^{-2} & 0 & \hdots & 0 \\
-    0 & \sigma_2^{-2} & \hdots & 0 \\
-    \vdots & \vdots & \ddots & 0 \\
-    0 & 0 & \hdots & \sigma_N^{-2} \\
-  \end{bmatrix}
-\end{equation}
-and $\sigma_j$ represents the width of the Gaussian prior on the j-th Chebyshev coefficient.
-$\vec{\mu}$ is the mean of the Gaussian prior, in nearly all cases we expect to set this to $\vec{\mu} = \{0, 0, \ldots, 0\}$, unless we have prior knowledge that the flux-calibration is systematically warped in one direction.
-Now, we can write the full posterior with priors as
-\begin{equation}
-  p(\vstar, \vN | \fD) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A} \vc + \vec{B}^\trans(\vstar) \vc + g(\vstar) \right )  \exp \left ( -\frac{1}{2} (\vc - \vec{\mu})^\trans {\bm D} (\vc - \vec{\mu}) \right ) p(\vstar) p(c_0)
-  \label{eqn:posterior_prior}
-\end{equation}
-We can expand and rearrange the argument of Equation~\ref{eqn:nuisance_prior} to a similar form that is quadratic in $\vc$ rewrite
-\begin{align}
-  {\bm A}^\prime &= {\bm A} + {\bm D}\\
-  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
-  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
-\end{align}
-in the case that $\vec{\mu} = \{0, 0, \ldots, 0\}$, there is only a non-zero correction to ${\bm A}$. The full posterior probability distribution is then
- \begin{equation}
-   \boxed{
-  p(\vstar, \vN | \fD) = p(\vstar, c_0, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(c_0) p(\vc)
-}
-\label{eqn:lnprob_lognormal}
-\end{equation}
-and the margin form is 
-\begin{equation}
-  \boxed{
-  p(\vstar, c_0 | \fD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar) p(c_0)
-}
-\label{eqn:lnprob_lognormal_marg}
-\end{equation}
-If one uses MCMC to sample the marginalized posterior, but desires a flux-corrected spectrum for the purposes of inspecting the residuals, it is still possible to do.
-We can draw a random sample $\vstar$ from the MCMC chain and then resample the unmarginalized posterior (Eqn~\ref{eqn:lnprob_lognormal}) conditional on $p(\vN | \vstar, \fD)$.
-Now, the combination of $\vstar$ and $\vN$ completely describes a flux-corrected spectrum.
-
-\subsection{Simplification for small $c_0$ perturbations}
-\label{sec:gaussian_simplification}
-If the correction in $c_0$ is small, we can use an alternate formulation of $k$ and put Gaussian priors on all nuisance parameters.
-This allows analytic marginalization over all nuisance parameters (including $c_0$), and reduces the dimensionality of the posterior to only the stellar parameters, $\vstar$.
-\begin{align}
-  k(\lambda | \vN) &= c_0 T_0(\lambda) + c_1 T_1(\lambda) + \ldots + c_N T_N(\lambda) \\
-  k(\lambda | \vN) &= \sum^N_{i = 0} c_i T_i(\lambda)
-\end{align}
-\begin{equation}
- \ln \bigl [p(\vt | \lambda_i) \bigr] \propto -\frac{1}{2 \sigma^2} \bigl [ \fD^2 - 2 \fD \fM \sum_n c_n T_n + \fM^2 \sum_n \sum_m c_n T_n c_m T_m \bigr ]
- \end{equation}
-Now the column vectors $\vc$ and $\vec{T}(\lambda)$ include $c_0$ as well
-\begin{equation}
-  \vc = 
-  \begin{bmatrix}
-    c_0\\
-    c_1\\
-    \vdots\\
-    c_N
-  \end{bmatrix}
-  \hspace{3cm}
-\vec{T}(\lambda) = 
-\begin{bmatrix}
-T_0(\lambda)\\
-T_1(\lambda)\\
-\vdots\\
-T_N(\lambda)\\
-\end{bmatrix}
-\end{equation}
-\begin{equation}
-  k(\lambda | \vN) = \sum^N_{i = 0} c_i T_i = \vec{T}^\trans \cdot \vc
-\end{equation}
-\begin{equation}
-  {\bm W}(\lambda) = \vec{T} \cdot \vec{T}^\trans = 
-  \begin{bmatrix}
-T_0 T_0 & T_0 T_1 &  \hdots & T_0 T_N \\
-T_1 T_0 & T_1 T_1 &  \hdots & T_1 T_N \\
-\vdots  & \vdots  &  \ddots & \vdots \\
-T_N T_0 & T_N T_1 &  \hdots & T_N T_N \\
-  \end{bmatrix}
-\end{equation}
-\begin{equation}
-  \sum_n \sum_m c_n T_n c_m T_m = \vc^\trans \cdot {\bm W} \cdot \vc
-\end{equation}
-\begin{align}
-  {\bm A} &= \sum_\lambda \frac{\fM^2(\lambda)}{\sigma^2(\lambda)} {\bm W}(\lambda)\\
-  \vec{B}^\trans &= \sum_\lambda \frac{\fD(\lambda) \fM(\lambda) }{\sigma^2(\lambda)} \vec{T}^\trans(\lambda)\\
-  g &= -\frac{1}{2} \sum_\lambda \frac{\fD(\lambda)}{\sigma^2(\lambda)}\\
-\end{align}
-If we include Gaussian priors, then we have 
-\begin{align}
-  {\bm A}^\prime &= {\bm A} + {\bm D}\\
-  \vec{B}^\prime &= \vec{B} + ({\bm D} \vec{\mu})^\trans\\
-  g^\prime &= g - \frac{1}{2} \vec{\mu}^\trans {\bm D} \vec{\mu} 
-\end{align}
-where now ${\bm D}$ and $\vec{\mu}$ include $c_0$ as well.
-The main difference from before is that now the prior on $c_0$ is now Gaussian instead of log-normal.
-This is acceptable as long as $\sigma_{c_0}$ is small ($< XX\%$), otherwise this will bias the correction factor.
-This also means that we will usually have $\vec{\mu} = \{1, 0, \ldots, 0\}$.
-The full posterior probability function becomes
- \begin{equation}
-   \boxed{
-  p(\vstar, \vN | \fD) = p(\vstar, \vc\;) \propto \exp \left ( - \frac{1}{2} \vc^\trans {\bm A}^\prime \vc + \vec{B}^{\prime \trans} \vc + g^\prime \right ) p(\vstar) p(\vc)
-}
-\end{equation}
-and the marginal form is
-\begin{equation}
-  \boxed{
-  p(\vstar | \fD) \propto \sqrt{\frac{(2 \pi )^N}{ {\rm det} \bigl |{\bm A}^\prime \bigr |}} \exp{ \left ( \frac{1}{2} \vec{B}^{\prime\trans} {\bm A}^{\prime -1} \vec{B}^\prime + g^\prime \right )} p(\vstar)
-}
-\end{equation}
-
-To gain insight, we can re-express this new version of $k$ in the original form as 
-\begin{equation}
-   k(\lambda | \vN) = c_0 T_0 + c_1 T_1 + \ldots + c_N T_N
-\end{equation}
-where $c_1 = c_0 {c_1}^\prime T_0$ and $c_N = c_0 {c_N}^\prime T_0$.
-Thus, if there needs to be a large perturbation in $c_0$, then there must also be a large perturbation in $c_n$.
-The prior on $c_1$ is now a product of a log-normal prior on $c_0$ and a Gaussian prior on ${c_1}^\prime$, which is not analytically tractable.
-Therefore, in the case of large perturbations, it is better to use Equation~ and use a log-normal prior on $c_0$.
-When we use the original framework, we are saying that perturbations in $c_0$ are small enough that we can approximate the log-normal prior on $c_0$ as Gaussian and that the priors on $c_n$ can also be correctly approximated by a Gaussian.
-From tests, the original framework gives acceptable accuracy (using $\sigma_{c_n} = 0.05$) if the corrections in $c_0$ are $\sigma_{c_0} \leq 0.35$, or the one-sigma deviation of $c_0$ is between $0.7 \leq c_0 \leq 1.42$.
-
-%\begin{deluxetable}{ll}
-%\tablecaption{\label{table:} Title}
-%\tablehead{\colhead{Col1} & \colhead{Col2}}
-%\startdata
-%\enddata
-%\tablecomments{}
-%\end{deluxetable}
+\subsection{Applications}
+Learnt covariance structure can be used to correct the models.
 

### 2014-05-02 paper/paper.tex
paper additions. updated figs 1 + 2

@@
-%\documentclass[preprint]{aastex} %double-column, single-spaced document:
-\documentclass[iop,floatfix]{emulateapj}
+\documentclass[preprint]{aastex} %double-column, single-spaced document:
+%\documentclass[iop,floatfix]{emulateapj}
 
 \usepackage{hyperref}
-%\usepackage{graphicx}
+\usepackage{graphicx}
 %\usepackage{apjfonts}
@@
 \usepackage{amsmath,amssymb}
-\usepackage{geometry}
 \usepackage{bm}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
+\usepackage[utf8]{inputenc}
 
@@
 \newcommand{\A}{[\alpha/{\rm Fe}]}
+\newcommand{\matern}{Mat\'{e}rn}
 
@@
 
-%\slugcomment{}
-%\shorttitle{}
-%\shortauthors{}
+\shorttitle{Spectroscopic inference}
+\shortauthors{Czekala et al.}
 
@@
 $\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
-$k(i | \vN)$  & polynomial function that multiplies a synthetic spectrum\\
 $\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
@@
 $\fR$ & residual vector $\fD - \fM$\\
+$C$ & covariance matrix\\
+$k_{\rm global}$  & global covariance kernel\\
+$k_{\rm region}$  & regional covariance kernel\\
 \hline
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{model_data}
+  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum to account for inaccuracies in the flux-calibration.
+\textbf{Middle}: The data spectrum and model spectrum, after it has been interpolated, post-processed, and multiplied by the Chebyshev polynomial.
+\textbf{Bottom}: Residuals from the model fit.
+Note the large residual at $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.
+\protect \todo{Model spectrum will be updated once I properly burn in the chain.
+Do these line styles look OK?
+I was going for something legible that might also print well in B/W}.}
+\label{fig:model_data}
+\end{center}
+\end{figure*}
+
 \subsection{Interpolation from synthetic spectral libraries}
@@
 \todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
-\todo{Investigate Tom Loredo's mentions of ``Cosmological Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
+\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
 
@@
 We correct for a radial velocity difference $v_z$ between the star and the earth by Doppler shifting the model spectrum.
+\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in Fourier space, but it's pretty fast as-is.}
 
-\todo{Convolution in wavelength space is a possibility. Allows LSF to change with wavelength.}
+\todo{We might try implementing the convolution in wavelength space, rather than the FFT. This would allow the LSF to change with wavelength.\\}
 Pixel width effects are not important as long as the LSF of the instrument is adequately sampled ($\gtrsim 3$ pixels across the LSF).
@@
 
-\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in Fourier space, but it's pretty fast as-is.}
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[width=0.9\textwidth]{badlines}
+  \caption{The variety of residual behaviour, depending on the quality of model fit.
+From left to right: \textbf{Class 0} covariance results from slight model mismatch, and correlates nearby residuals.
+\textbf{Class I}: A missing absorption line in the model leaves a large, highly correlated patch of negative residuals.
+\textbf{Class II}: An extraneous line in the model leaves a large, highly correlated patch of positive residual.
+\textbf{Class III}: If lines are present in the model but of the wrong strength, many correlated residuals of moderate amplitude will result.
+The difficulty with class III lines is that for any specific line, there might exist a $\vstar$ that will fit the line, but there does not exist a $\vstar$ that will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure*}
 
@@
 If the behavior of the spectrograph is characterized by observations of spectrophotometric standards, then we can put priors on the coefficients that will prevent overfitting of the polynomials.
+For an example of a typical spectrum, and Chebyshev polynomial, see Figure~\ref{fig:model_data}.
 
@@
 To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we use a standard multidimensional Gaussian likelihood function which allows for covariance between data points.
-If the residuals vector is
+If the vector of residuals is
 \begin{equation}
@@
 \end{equation}
-with $N$ data points, then the likelihood function is
+with a length of $N$ data points, then the likelihood function is
 \begin{equation}
@@
 \end{equation}
-$C$ is the covariance matrix which describes the covariance between pixels in the spectrum. For independent noise with noise per pixel $\sigma_i$, this matrix is
+$C$ is the covariance matrix which describes the covariance between pixels in the spectrum. For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
 \begin{equation}
@@
   \end{bmatrix}
+  \label{eqn:covariance_diagonal}
 \end{equation}
-and Equation~\ref{eqn:lnprob} reduces to the familiar $\chi^2$ form of a sum over the square of the residuals, weighted by the inverse variance of each data point
+with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the familiar $\chi^2$ form of a sum over the square of the residuals, weighted by the inverse variance of each data point
 \begin{equation}
@@
 \end{equation}
-However, residuals from a spectroscopic fit are often correlated due to the convolution of the stellar rotation and LSF kernels.
-Additionally, systematic errors in the synthetic spectra (such as incorrect line strengths) will resulted in regions of highly correlated residuals, which requires that we use a covariance matrix with off-diagonal terms $\sigma_{ij}$ and Equation~\ref{eqn:lnprob}, which includes a matrix product instead of a sum over independent pixels.
 
-In order properly account for these correlated residuals, we must understand the covariance structure and parameterize it in our model.
-We seek to understand two types of covariance structure: the large scale global covariance structure that results from the stellar rotation and LSF kernels, and a regional covariance structure that is localized to areas where spectral line mismatch results in small patches of highly correlated residuals.
+However, residuals from a spectroscopic fit are often correlated due to \todo{DUE TO WHAT, exactly??
+I'm confused.
+Wait for response from Tom Loredo}.
+Additionally, systematic errors in the synthetic spectra (such as incorrect line strengths) will result in regions of highly correlated residuals.
+Properly accounting for correlations in the residuals requires that we use a covariance matrix with off-diagonal terms $\sigma_{ij}$ and a likelihood function (Equation~\ref{eqn:lnprob}) which uses a matrix product instead of a sum over independent pixels.
 
-\todo{Tom Loredo was saying that if we have a model that perfectly fits the data then our residuals will be uncorrelated, regardless of the PSF/LSF.
-But is this true?
-Shouldn't all noise be correlated on some scale by the LSF?
-Correspond by email to clarify and show an example.}
+We seek to understand and parameterize two types of covariance structure in our model, exemplified in Figure~\ref{fig:badlines}: the large scale but generally mild global covariance structure that results from [still to be finalized cause], and a regional covariance structure that is localized to areas where spectral line mismatch results in small patches of highly correlated residuals.
+Ignoring the complexity of spectroscopic residuals will bias the estimates of the model parameters $\vt$.
 
-To account for the global covariance structure, we use a Matern 3/2 kernel.
-\todo{Explore difference between using 3/2 Matern kernel and squared exponential kernel.}
+As a simple analogy, consider the fit of a straight line to a dataset.
+If the noise in the dataset is correlated, then adjacent data points might be offset from the linear trend in the same direction by a similar amount.
+If the covariant structure of the noise is ignored, a simple $\chi^2$ will treat these correlated offsets as part of the linear trend, which will result in a biased determination of the slope and intercept of the line, typically with uncertainties that are too small.
+
+\subsection{Parameterizing the covariance structure}
+We parameterize the covariance structure using covariance kernels, which sets the covariance between two pixels $\lambda_i$ and $\lambda_j$.
+This behavior is analogous to the two-point correlation function used in cosmology, where the distance between two galaxies is used instead of wavelength.
+
+\paragraph{Global covariance structure}
+A global covariance structure may result from a slight mismatch in the continuum or mild mismatch of spectral lines like that shown as class 0 in Figure~\ref{fig:badlines}.
+To account for this structure, which is generally mild and typically exists across only a few pixels, we use a \emph{stationary} covariance kernel, which satisfies the property that the degree of correlation only depends on the distance between the two pixels $r$.
+Functions like this are also called \emph{radial basis functions}.
+For spectra, we map $r$ to the velocity difference between two pixels
+\begin{equation}
+  r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i - \lambda_j}{ \lambda_i + \lambda_j} \right |
+\end{equation}
+We choose the \matern\ $3/2$ kernel because\ldots
+\todo{Explore difference between using \matern\ kernel and squared exponential kernel.
+Why \matern\ and why Hann?
+Because they are the vanilla covariance functions that are tried and tested in the machine learning communities.}
 \begin{equation}
@@
 \end{equation}
+In order to keep the covariance matrix sparse for computational efficiency, we taper for compact support using a Hann window
+\begin{equation}
+  w(r, r_0) = \frac{1}{2} + \frac{1}{2} \cos \left( \frac{\pi r}{r_0} \right) 
+\end{equation}
+for a global covariance kernel of 
+\begin{equation}
+  k_{\rm global} (r, a, l, r_0) = w(r, r_0) k_{\nu = 3/2}(r, a, l) 
+  \label{eqn:global}
+\end{equation}
+We add this covariance kernel to the independent variance already present along the matrix diagonal
+\begin{equation}
+  k(\lambda_i, \lambda_j) = k_{\rm global}(r, a, l, r_0) + \delta_{ij} \sigma_{ij}^2
+\end{equation}
+where $\delta_{ij}$ is the Kroenecker delta function.
 
-tapered for compact support using a Hann window with width $r_0 = 6 l$ (Why? Because this looked good).
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{matern_matrix}
+\includegraphics[width=0.4\textwidth]{matern_draw}
+\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel and typical parameters for our dataset.
+\textbf{Right} general spectroscopic residuals near a stellar continuum overlaid with a sample draw using the covariance matrix, showing that the two look similar in structure and amplitude.
+\protect \todo{Prettify these figures.
+Consistent labelling.
+Change color stretch so Matrix zeros are white.}}
+\label{fig:matern}
+\end{center}
+\end{figure*}
 
-Specific line covariance. Non-stationary covariance kernel.
+To visualize what the covariance matrix $C$ might look like when parameterized by Equation~\ref{eqn:global}, see the left panel of Figure~\ref{fig:matern}.
+To explore the relationship between the parameters of the covariance kernel and the properties of the noise, first consider simulating uncorrelated random Gaussian noise by drawing many independent samples from a Gaussian.
+This same uncorrelated noise could also be represented as a single draw from a $N$-dimensional Gaussian with a diagonal covariance matrix, like in Equation~\ref{eqn:covariance_diagonal}.
+If instead of a diagonal covariance matrix, we were to draw samples from a non-trivial covariance matrix like the one in the left panel of Figure~\ref{fig:matern}, then we will see correlated noise like in the right panel.
+\todo{Do you think examples like Figure~\ref{fig:matern} are helpful or a waste of space?}
+
+\paragraph{Specific line covariance} For the large and highly correlated residuals that result from class I, II, and III line errors, we use a \emph{non-stationary} covariance kernel, which means that the covariance explicitly depends on $\lambda_i$ and $\lambda_j$.
+\todo{Explain how we come up with this kernel, which is the covariance resulting from a Gaussian function.}
 \begin{equation}
-  k_{\rm G}(x, x^\prime | a, \mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{[(x - \mu)^2 + (x^\prime - \mu)^2]}{2 \sigma^2}\right )
+  k_{\rm G}(\lambda_i, \lambda_j | a, \lambda_\mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{r^2(\lambda_i, \lambda_\mu) + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
 \end{equation}
+The squared exponential kernel is
+\todo{Why do we use the squared exponential and not \matern?}
 \begin{equation}
-k(x, x^\prime | h, a, \mu, \sigma) = \exp \left ( \frac{-( x - x^\prime)^2 }{2 h^2} \right ) k_{\rm G}(x, x^\prime | a, \mu, \sigma)
+  k_{ \rm exp}(r, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
 \end{equation}
-$h$ is a "bandwidth." If $h$ is small, then there will be high-frequency structure. If $h$ is large, then only low-frequency structure will remain.
+$h$ is a bandwidth.
+If $h$ is small, then there will be high-frequency structure.
+If $h$ is large, then only low-frequency structure will remain.
+We also use the Hann window to taper the kernel for compact support
+\begin{equation}
+  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = w(r, r_0)\; k_{ \rm exp}(r, h) \; k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
+\end{equation}
 
-These are all parameters that we can sample in. Also, there would be a set of these parameters for each line in the model.
+\todo{Similar paragraph about visualizing the covariance matrix, drawing samples. References to Figure~\ref{fig:region}.}
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{matrix_region_covariance}
+\includegraphics[width=0.4\textwidth]{random_draw}
+\caption{\textbf{Left} a covariance matrix generated with the region kernel.
+\textbf{Right} general spectroscopic residuals near a mismatched stellar line overlaid with a sample draw using the covariance matrix, showing that the two look similar in structure and amplitude.
+\protect \todo{Similar improvements to Figure~\ref{fig:matern}.}}
+\label{fig:region}
+\end{center}
+\end{figure*}
+
+\begin{itemize}
+  \item non-stationary kernel localizes increased variance to a specific spectral line
+  \item better than just masking out the region, since these contain info
+  \item there will be many sets of regions ($\sim$1-4\% of all lines might be covered by a region)
+  \item there can be hyperparameters describing the \emph{population} of regions in a hierarchical model (mostly width, amplitude).
+\end{itemize}
+
+Covariance kernel parameters are parameters in our model that we explicitly sample for.
 
 The benefit mainly comes from simply modelling these systematic residuals to begin with, since a model with covariance is far more likely than forcing the fit, and far more justified than masking regions which do not fit.
+
+\todo{Show how the kernel can accounts for complicated M dwarf structure where there is large, messy mismatch.}
+
+\subsection{Exploring the posterior}
+Markov Chain Monte Carlo
+\todo{Given this is 2014 and not 2010, how much do you think I need to explain?}
 
@@
 
+\section{Tests}
+\todo{Which tests to show?}
+
 \bibliography{disks,bayesian,master}

### 2014-05-05 tex/hogg1.tex
Reformatted tex directory

\documentclass[preprint]{aastex} %double-column, single-spaced document:
%\documentclass[iop,floatfix]{emulateapj}

\usepackage{hyperref}
%\usepackage{graphicx}
%\usepackage{apjfonts}
\usepackage{enumerate}
\usepackage{amsmath,amssymb}
\usepackage{geometry}
\usepackage{bm}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\newcommand{\prob}{{\rm prob}}
\newcommand{\qN}{\{q_i\}_{i=1}^N}
\newcommand{\qM}{\{q_{im}\}_{i=1,m=0}^{N,M}}
\newcommand{\yN}{\{y_i\}_{i=1}^N}
\newcommand{\vt}{\vec{\theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\vstar}{\vt_{\star}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vc}{\vec{c}}
\newcommand{\fM}{ {\bm M}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ {\bm D}}
\newcommand{\fDi}{D_i}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}

\newcommand{\hcom}[1]{ \textcolor{Blue}{#1}}

%\slugcomment{}
%\shorttitle{}
%\shortauthors{}

\begin{document}

\title{Methods for the spectroscopic inference of fundamental stellar parameters, continued.}
\author{\today{}\\
\medskip
Ian~Czekala\altaffilmark{1} et al.
%Author2\altaffilmark{2},
}

\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street MS 10, Cambridge, MA 02138}
%\altaffiltext{2}{Institution 2}
\email{iczekala@cfa.harvard.edu}

\section{Downsampling}

\hcom{section 2: why ``linearly interpolate?'' It would be good to see more specificity here; there are lots of ways to do this smooth and downsample slightly wrong.}

There are a few interpolations that are going on in this method. I will explain them in order.

\subsection{Interpolating a spectrum from a grid of synthetic spectra}
If we desire a raw, synthetic spectrum that has stellar properties $\vg = \{T_{\rm eff}, \log g, [{\rm Fe}/{\rm H}]\}$ that are not one of the grid points, then we must interpolate. At the moment, this is done as a linear interpolation because it is quick and memory efficient. I do agree with you that it would be better to use a more sophisticated interpolator. For example, \citet{hwd+13} use a spline interpolator that we plan to implement soon. This has currently been a low priority because our interpolation tests show that the expected error due to this interpolation should be small. However, for each raw pixel of the spectrum, we \emph{should} be doing a band-limited interpolation with respect to $T_{\rm eff}$, $\log g$, and $[{\rm Fe}/{\rm H}]$.

After interpolation, the new spectrum we just interpolated from the grid has the same pixel spacing and resolution as the original synthetic spectra.

\hcom{When you do the drop and interpolate over testing, can you build an empirical interpolation error model and then use it?}
Point taken! It's on the to-do list.

\subsection{Post-processing the interpolated raw spectrum into one that can be compared to our data spectrum}

We flux-calibrate our data spectrum by comparison with a spectrophotometric standard. Using the IRAF tasks \texttt{standard}, \texttt{sensfunc}, and \texttt{calibrate}, we convert the raw spectrum (in units of counts/pixel) into a flux-density ($f_\lambda$ in units of ${\rm erg}/{\rm s}/{\rm cm}^2/$\AA) by dividing by the sensitivity function determined from the spectrophotometric standard.

The raw synthetic spectra are also given in spectral flux density ($f_\lambda$), sampled at $R\approx$500,000. I prefer to think of the synthetic spectrum as a band-limited function that is sampled at many points, hopefully fine enough such that it is Nyquist sampled. Stellar spectra in particular are band-limited by several effects. The first is that intrinsic stellar phenomenon like microturbulence, macroturbulence, and rotation all act to suppress high frequency features in the spectrum.  Second, the line-spread function of the observing instrument also removes high frequency stellar features. Finally, the width of the pixel on the CCD also destroys spectral information of the highest frequencies, if any remain. The result is that a band-limited spectrum has a Nyquist frequency above which we need not sample, but below which we sample at our peril.

To transform our raw interpolated spectrum into a format that matches our data spectrum, we first resample the uniformly sampled (in angstroms) raw spectrum to a log-linear wavelength grid using spline interpolation. We take care to use a fine enough spacing such that no spectral information is lost. We test that we can interpolate from the log-linear grid \emph{back} to the original grid and recover the same flux values to within floating point accuracy. A log-linear grid has constant spacing in velocity space, and so it is the appropriate grid to use for an FFT. Once in Fourier space, we then multiply the spectrum by the PSF kernel of the instrument and a $v \sin i$ kernel for rotational broadening. For an illustration of the effects on the spectrum in Fourier space, see Figure~\ref{fig:pixel_response}.

\begin{figure}[!htb]
\begin{center}
  \includegraphics[width=0.48\textwidth]{pixel_response}
  \caption{What happens to the spectrum in the Fourier domain during post processing. The first panel shows the Fourier transform of the raw spectrum interpolated from the grid. The second panel shows the kernels for a PSF with ${\rm FWHM} = 6.8$ km/s and a rotation kernel for $v\sin i = 5$ km/s. The third panel is the raw spectrum multiplied by these kernels. The final panel shows the sinc functions representing the frequency response of the smallest and largest pixel on our detector.}
\label{fig:pixel_response}
\end{center}
\end{figure}

Should we be accounting for pixel effects? In the image plane, an individual pixel could be represented as a box-car or top-hat function which convolves the spectrum $f_\lambda$ and is then resampled at the pixel-center $\lambda_0$. In the Fourier domain, the effect of the pixel is represented as multiplication by a sinc function. Because the width of the pixel is small, the main lobe of the sinc function is large. Figure~\ref{fig:pixel_response} shows that all of the stellar information in Fourier space is within this main lobe, and so as an approximation it seems like we do not need to worry about pixel-convolutions. I believe the situation would be different if our pixels undersampled the PSF.

Ignoring the pixel width effects for now, we do an IFFT back to the wavelength domain and then use splines to resample the flux-density at the exact pixel centers of our dataset. Here, we also test the spline interpolation using differently spaced wavelength grids to check that we can still recover the same spectrum to within numerical accuracy. The pixels of our particular CCD are roughly log-linear spaced, and there are at least 2.5 pixels across the FWHM of the spectrograph PSF. Summarizing the previous statements in math, we have discrete samples of some synthetic stellar spectrum $f_{\lambda, {\rm synth}}$, and we would like to morph them into samples of a different spectrum $f_{\lambda, {\rm inst}}$, which correspond to the CCD pixels of our spectrograph. In wavelength space this operation is written as
\begin{equation}
  f_{\lambda, {\rm inst}}(\lambda) = \Pi(\lambda) \otimes g_{\rm PSF}(\lambda) \otimes g_{v \sin i}(\lambda) \otimes f_{\lambda, {\rm synth}}(\lambda)
  \label{eqn:convolution}
\end{equation}
and in Fourier space as
\begin{equation}
  F_{\lambda, {\rm inst}}(s) = {\rm sinc}(s) G_{\rm PSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
\end{equation}

In the future, we may consider doing away with the FFT routine entirely in favour of a sliding convolution approach in the wavelength domain (Equation~\ref{eqn:convolution}). With this approach, we can account for the variation of PSF and pixel size with wavelength. The difficulty in implementation comes with implementing a continuous convolution using discrete samples of the spectrum while preserving accuracy. After convolution, we would still need to take care to ensure band-limited interpolation to the CCD pixel.

\subsection{Pixel-convolved PSF}

The previous approach is not very rigorous about the distinction between intensity and flux. In your blog and papers you've mentioned the ``pixel-convolved'' PSF at the focal plane, or the ``pixel-beam.'' I was wondering if you could point me to a reference that makes prodigious use of equations? I've worked out the following\ldots could you point out anywhere I have gone wrong?

The intensity field $I(\vec{x}, \vec{n})$ is a function of three dimensional space and direction, denoted by a position vector $\vec{x}$ and direction vector $\vec{n}$.  In a spectrograph, this intensity field is modified by the system response (PSF) of the dispersive mechanism. Assuming monochromatic light, the intensity field behind the spectrograph can be written as
\begin{equation}
  I^\prime_\nu(\vec{x}, \vec{n}) = PSF_\nu(\vec{x}, \vec{n}) \otimes I_\nu(\vec{x}, \vec{n})
\end{equation}
where this convolution takes place in three-dimensions. What we care most about is the intensity field at the focal plane, 
\begin{equation}
  I_{\nu, {\rm FP}}(x, y, \vec{n}) = I^\prime_\nu(x, y, z=0, \vec{n})
\end{equation}
where $x$ and $y$ denote position on the 2D focal plane.

If we have a infinitesimal pixel with size $dA$, then in time $dt$ a source at an angle of $\theta$ and with angular size $d\Omega$ will deliver the following energy to the pixel
\begin{equation}
  dE = I_\nu dt (dA \cos \theta) d\Omega d\nu
\end{equation}

My impression is that the ``pixel-beam'' becomes important when we consider that each pixel is not infinitesimal in size nor does it have a uniform response throughout the pixel nor to all directions of incoming light. If we define a directional pixel efficiency as $\eta_\nu(x, y, \vec{n})$, then the energy received by a pixel with finite area is
\begin{equation}
  E = \int \iint \int \int I_{\nu, {\rm FP}} (x, y, \vec{n})\, dt\, (dx\,dy \cos \theta) \eta_\nu(x, y, \vec{n}) \cdot \vec{n} \, d\vec{n}\, d\nu
\end{equation}

where the spatial integral over $dx, dy$ spans the size of the pixel. By ``pixel-beam,'' do you mean something similar to $\eta_\nu$? I am interested in learning more about how this is done, but have been unable to find a detailed mathematical example. For spectra, there is also the additional problem of ``optimal'' extraction, or converting the 2D spectrum into a 1D spectrum.

\section{Correlated noise}
\begin{itemize}
  \item \hcom{systematic errors don't necessarily invalidate chi-squared if they can be treated as a noise with a non-trivial covariance matrix; they just make chi-squared not an independent sum of residuals weighted by inverse variances but because a dot-product through a non-trivial inverse covariance tensor.}
  \item \hcom{``more drastic than adding in noise'' -- what about ``adding in highly covariant and correlated noise''? That takes no more equipment (everything is still Gaussian), but can handle far worse situations.}
  \item \hcom{Where you mention $\chi^2_Q$, you are treating the pixels as independent. But they aren't if there is model mismatch. That is, there are correlations among the residuals.}
\end{itemize}

Thank you for pointing this out. Your suggestions spurred a flurry of math on my part and I arrived at an interesting implementation of the correlated noise idea. When Dan Foreman-Mackey visited the CfA, I showed him my ideas and he gave me some great suggestions about to implement them in code.

\subsection{Gaussian residual}
My first approach was to deal with the systematically bad regions of the spectrum, where there is a large, correlated residual due to a line strength mis-match. By approximating the residual as a Gaussian, I derived a covariance kernel which might parameterize this ``patch'' of noise.

If the mean function is
\begin{equation}
m(x) = E[f(x)]
\end{equation}
and the covariance function is
\begin{equation}
k(x, x^\prime) = E[(f(x) - m(x))(f(x^\prime) - m(x^\prime))]
\end{equation}
then we can compute the non-stationary covariance function describing a Gaussian as
\begin{equation}
k(x, x^\prime | a, \mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{[(x - \mu)^2 + (x^\prime - \mu)^2]}{2 \sigma^2}\right )
\end{equation}
If we have a line residual as in Figure~\ref{fig:fake_line}, then we can use this kernel to create a matrix like in Figure~\ref{fig:matrix_region} in order to downweight the residual. To check that we could reproduce a residual with a covariance matrix like this, we can draw random samples from this covariance matrix and plot these against the residuals, as in Figure~\ref{fig:random_draw}.

\begin{figure}[!htb]
\begin{center}
  \includegraphics[width=0.5\textwidth]{fake_line}
\caption{A fake Gaussian residual resulting from an imperfect model.}
\label{fig:fake_line}
\end{center}
\end{figure}

\begin{figure}[!htb]
\begin{center}
\includegraphics{matrix_region}
\caption{Covariance matrix of a single patch of Gaussian noise. Left: sparsity diagram. Right: covariance matrix.}
\label{fig:matrix_region}
\end{center}
\end{figure}

\begin{figure}[!htb]
\begin{center}
\includegraphics{random_draw}
\caption{15 random draws from the covariance matrix, overplotted with the residuals from Figure~\ref{fig:fake_line}.}
\label{fig:random_draw}
\end{center}
\end{figure}

\subsection{Global covariance}
DFM recommended that we also implement a stationary covariance kernel to account for the general covariance structure of the data. For this, we use a Matern kernel 
\begin{equation}
  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp \left (- \frac{\sqrt{3} r}{l} \right ) 
\end{equation}
and add compact support by tapering with a Haan window
\begin{equation}
  w(r) = 0.5 + 0.5 \cos \left( \frac{\pi r}{r_0} \right)
\end{equation}
This kernel does a decent job of matching the covariance structure that still remains in a region of stellar continuum. We sample the parameters describing these kernels in our MCMC code. We also have a bit of logic that adds the patches of correlated noise until we have addressed all of the egregious residuals.

\hcom{good on the stuff about there being important information where the models and data disagree.}

I am fairly new to the Gaussian processes kernel game, but I would imagine that different covariance functions could be useful for different spectral problems, such as M dwarf continuua (Figure~\ref{fig:mann}). My first attempt is to create a new kernel with a finite bandwith taper to replace the original Gaussian kernel for a bad spectral line. This covariance function is a squared exponential with a Gaussian taper
\begin{equation}
k(x, x^\prime | h, a, \mu, \sigma) = \exp \left ( \frac{-( x - x^\prime)^2 }{2 h^2} \right ) \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{[(x - \mu)^2 + (x^\prime - \mu)^2]}{2 \sigma^2}\right )
\end{equation}
here $h$ is a "bandwidth" that controls the power of the oscillations. If $h$ is small, then there will be high-frequency structure (such as in an M dwarf residual). If $h$ is large, then only low-frequency structure will remain (such as in an FGK residual).

\begin{figure}[!htb]
\begin{center}
\includegraphics{mann}
\caption{Figure from \citet{mga13} showing the residuals (bottom panel) commonly encountered when fitting M dwarf spectra.}
\label{fig:mann}
\end{center}
\end{figure}

\hcom{in your four classes of line behavior, there is none which is dominated by continuum issues? Or is there? That is, continuum fitting can cause problems, no?}
In principle, there should be areas where continuum fitting can be a problem. However, in the stars I'm looking at (over a narrow echelle range), the Chebyshev formalism essentially assumes that any continuum mis-match is actually a flux-calibration (instrumental) issue, and accounts for it. This means that if we were fitting hotter stars with few lines, any information contained in the continuum shape would be lost.

\hcom{Section 5: You might even be able to learn changes to the models. This is the holy grail.}

If we were to run against a large spectroscopic library of many stars (say, the \emph{Kepler} follow-up sample, or some SLOAN sample), we could use the learnt covariance of the spectrum to identify the bad regions and then determine how the models would need to be changed in order to fit all of the observed stars the best, similar to using Gaussian processes to predict the value of a function where data is missing.

\section{Chebyshev}
\hcom{Can you really marginalize analytically over the Chebyshev polynomials? And doesn't that make a huge, correlated uncertainty tensor? And what priors do you use?}
\hcom{In re +/- 0.01 dex: You have tons of pixels so of course you get amazing constraints! It would be surprising if you didn't. But do you get such amazing constraints if you marginalize analytically over thousands of Chebyshevs?}
Because the residual echelle blaze is smooth and small ($\le 5\%$ error) we use only a fourth-order Chebyshevs to fit each order of the spectrum. For a visual description of what this looks like in practice, please see \url{http://iancze.github.io/image.html}.

\bibliography{disks,bayesian,master}
\bibliographystyle{hapj}
\end{document}

### 2014-05-05 tex/paper.tex
Reformatted tex directory

--- 2014-05-02 paper/paper.tex
@@
 \usepackage{graphicx}
-%\usepackage{apjfonts}
+\usepackage{apjfonts}
 \usepackage{enumerate}
@@
 \usepackage[utf8]{inputenc}
+
+%version-control tagging based off of github.com/bd-j/speccal
+\input{vc}
 
@@
 \section{Method}
-Typically, the stellar properties we most desire are effective temperature $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
-Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz, more being developed for GAIA) and are parameterized by these fundamental stellar parameters, which together, we call $\vg$.
-Synthetic libraries provide high-resolution spectra which span a range of $\vg$ covering the main sequence of spectral types.
-The library is typically specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
+Typically, the stellar properties we most desire are effective temperature
+ $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
+Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz,
+ more being developed for GAIA) and are parameterized by these fundamental
+ stellar parameters, which together, we call $\vg$. 
+Synthetic libraries provide high-resolution spectra which span a range of $\vg$
+ covering the main sequence of spectral types. The library is typically
+ specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
 
-In addition to these fundamental parameters, a star also has several observed properties that are a function of its kinematics, geometry, and location in our galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$, total extinction $A_V$, and solid angle $\Omega$.
-We call these secondary parameters $\vpp$, because we can model their effects on the synthetic spectra in a ``post-processing'' step by convolution with an appropriate kernel or multiplication with a smooth function.
+In addition to these fundamental parameters, a star also has several observed
+ properties that are a function of its kinematics, geometry, and location in our
+ galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
+ total extinction $A_V$, and solid angle $\Omega$. 
+We call these secondary parameters $\vpp$, because we can model their effects
+ on the synthetic spectra in a ``post-processing'' step by convolution with an
+ appropriate kernel or multiplication with a smooth function. 
 Together, we call $\vstar = \{\vg, \vpp\}$.
 
-Using a synthetic stellar library as a backend and applying post-processing techniques, we attempt to forward-model the observed stellar spectra in order to determine the best-fit $\vstar$.
-Because properly sampled spectra have correlated pixels, we present a framework which self-consistently models these correlations.
+Using a synthetic stellar library as a backend and applying post-processing
+ techniques, we attempt to forward-model the observed stellar spectra in order
+ to determine the best-fit $\vstar$. 
+Because properly sampled spectra have correlated pixels, we present a framework
+ which self-consistently models these correlations.
 
@@
 \begin{center}
-  \includegraphics{model_data}
-  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum to account for inaccuracies in the flux-calibration.
-\textbf{Middle}: The data spectrum and model spectrum, after it has been interpolated, post-processed, and multiplied by the Chebyshev polynomial.
-\textbf{Bottom}: Residuals from the model fit.
-Note the large residual at $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.
-\protect \todo{Model spectrum will be updated once I properly burn in the chain.
-Do these line styles look OK?
-I was going for something legible that might also print well in B/W}.}
+  \includegraphics{figs/model_data}
+  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
+   to account for inaccuracies in the flux-calibration. 
+  \textbf{Middle}: The data spectrum and model spectrum, after it has been
+   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
+  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
+   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}. 
+  \protect \todo{Model spectrum will be updated once I properly burn in the
+   chain. Do these line styles look OK? I was going for something legible
+   that might also print well in B/W}.}
 \label{fig:model_data}
@@
 \subsection{Interpolation from synthetic spectral libraries}
-Creating a new spectrum with a specific $\vg$ requires either synthesizing a new spectrum using radiative transfer codes or interpolating from nearby grid points in the library.
-Because it is computationally expensive to synthesize a spectrum at high resolution over a wide wavelength range, we choose to interpolate.
+Creating a new spectrum with a specific $\vg$ requires either synthesizing a
+ new spectrum using radiative transfer codes or interpolating from nearby grid
+ points in the library. 
+Because it is computationally expensive to synthesize a spectrum at high
+ resolution over a wide wavelength range, we choose to interpolate. 
 Spline interpolation for spectra \citep{hus12}.
@@
 \subsection{Post-processing synthetic spectra to match the data}
-Libraries of synthetic spectra are generally computed at high resolution ($R \geq $100,000), sampled with many pixels per resolution element, and do not include any rotational broadening or account for any instrumental effects.
-In order to transform a raw synthetic spectrum interpolated from the grid into one that matches the spectrum of a real star, we must post-process the spectrum to account for these secondary effects.
-The projected rotational velocity of the star, parameterized by $v \sin i$, broadens spectral line profiles and is mathematically described by convolution with a kernel $g_{v \sin i}(\lambda)$.
-UV, optical, and infrared spectra are acquired using a spectrograph which disperses light onto a CCD, defining a specific resolution and sampling rate for the spectrum.
-Because the resolution of the spectrograph and spacing of the pixels are different from the synthetic spectral library, we must convolve the raw spectrum with the line spread function (LSF) of the instrument $g_{\rm LSF}(\lambda)$ and resample to the exact pixels of the CCD.
-In wavelength space, the $v \sin i$ and LSF operations are represented by the convolution of the synthetic spectrum with these two kernels
+Libraries of synthetic spectra are generally computed at high resolution ($R
+ \geq $100,000), sampled with many pixels per resolution element, and do not
+ include any rotational broadening or account for any instrumental effects. 
+In order to transform a raw synthetic spectrum interpolated from the grid into
+ one that matches the spectrum of a real star, we must post-process the spectrum
+ to account for these secondary effects. 
+The projected rotational velocity of the star, parameterized by $v \sin i$,
+ broadens spectral line profiles and is mathematically described by convolution
+ with a kernel $g_{v \sin i}(\lambda)$.  
+UV, optical, and infrared spectra are acquired using a spectrograph which
+ disperses light onto a CCD, defining a specific resolution and sampling rate
+ for the spectrum. 
+Because the resolution of the spectrograph and spacing of the pixels are
+ different from the synthetic spectral library, we must convolve the raw
+ spectrum with the line spread function (LSF) of the instrument $g_{\rm
+ LSF}(\lambda)$ and resample to the exact pixels of the CCD. In wavelength
+ space, the $v \sin i$ and LSF operations are represented by the convolution of
+ the synthetic spectrum with these two kernels
 \begin{equation}
@@
 \end{equation}
-Using the convolution theorem, we can rewrite these operations as multiplications in Fourier space 
+Using the convolution theorem, we can rewrite these operations as
+ multiplications in Fourier space 
 \begin{equation}
@@
 \end{equation}
-where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform pairs.
-In order to use the fast Fourier transform (FFT) to execute these operations using discrete samples of the synthetic spectrum $\fsynth$, we must first resample $\fsynth$ so that it is on a uniform grid.
-In the case of a spectrum, the natural uniform grid is one that is equally sampled in velocity space, such there is an equal velocity shift $\Delta v$ between pixels.
-This results in a wavelength grid that is linearly sampled in the logarithm of wavelength.
-Therefore, the Fourier coordinate $s$ is the number of cycles per sampling interval, having units of inverse velocity $[{\rm s}/{\rm km}]$.
-Next, we multiply the Fourier transform of the synthetic spectrum by the Fourier transforms of the rotational velocity and line spread function kernels.
-Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda, {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength locations of the pixels in the detector ($\finst$).
-When resampling the spectrum, we are careful to ensure band-limited interpolation using splines in order to prevent introducing non-physical high frequency structure into the spectrum.
+ where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform
+ pairs. 
+In order to use the fast Fourier transform (FFT) to execute these
+ operations using discrete samples of the synthetic spectrum $\fsynth$, we must
+ first resample $\fsynth$ so that it is on a uniform grid. 
+In the case of a spectrum, the natural uniform grid is one that is equally
+ sampled in velocity space, such there is an equal velocity shift $\Delta v$
+ between pixels. This results in a wavelength grid that is linearly sampled in
+ the logarithm of wavelength. 
+Therefore, the Fourier coordinate $s$ is the number of cycles per sampling
+ interval, having units of inverse velocity $[{\rm s}/{\rm km}]$. 
+Next, we multiply the Fourier transform of the synthetic spectrum by the
+ Fourier transforms of the rotational velocity and line spread function kernels.
+Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda,
+ {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength
+ locations of the pixels in the detector ($\finst$). 
+When resampling the spectrum, we are careful to ensure band-limited
+ interpolation using splines in order to prevent introducing non-physical high
+ frequency structure into the spectrum.
 
 Finally we apply corrections for interstellar extinction and Doppler shift.
-Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which is wavelength-dependent.
-We use the (name here) dust-extinction model, which is parameterized by extinction in the $V$ photometric band, $A_V$.
-We correct for a radial velocity difference $v_z$ between the star and the earth by Doppler shifting the model spectrum.
-\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in Fourier space, but it's pretty fast as-is.}
+Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which
+is wavelength-dependent. 
+We use the (name here) dust-extinction model, which is parameterized by
+ extinction in the $V$ photometric band, $A_V$. 
+We correct for a radial velocity difference $v_z$ between the star and the
+ earth by Doppler shifting the model spectrum. 
+\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
+ Fourier space, but it's pretty fast as-is.}
 
-\todo{We might try implementing the convolution in wavelength space, rather than the FFT. This would allow the LSF to change with wavelength.\\}
-Pixel width effects are not important as long as the LSF of the instrument is adequately sampled ($\gtrsim 3$ pixels across the LSF).
-If our spectrum were not properly sampled, we would need to additionally convolve with a boxcar the width of the pixels.
+\todo{We might try implementing the convolution in wavelength space, rather
+ than the FFT. 
+This would allow the LSF to change with wavelength.\\} 
+Pixel width effects are not important as long as the LSF of the instrument is
+ adequately sampled ($\gtrsim 3$ pixels across the LSF). 
+If our spectrum were not properly sampled, we would need to additionally
+ convolve with a boxcar the width of the pixels.
 
@@
 \begin{center}
-  \includegraphics[width=0.9\textwidth]{badlines}
-  \caption{The variety of residual behaviour, depending on the quality of model fit.
-From left to right: \textbf{Class 0} covariance results from slight model mismatch, and correlates nearby residuals.
-\textbf{Class I}: A missing absorption line in the model leaves a large, highly correlated patch of negative residuals.
-\textbf{Class II}: An extraneous line in the model leaves a large, highly correlated patch of positive residual.
-\textbf{Class III}: If lines are present in the model but of the wrong strength, many correlated residuals of moderate amplitude will result.
-The difficulty with class III lines is that for any specific line, there might exist a $\vstar$ that will fit the line, but there does not exist a $\vstar$ that will properly fit \emph{all} the lines.}
+  \includegraphics[width=0.9\textwidth]{figs/badlines}
+  \caption{The variety of residual behaviour, depending on the quality of model
+    fit. 
+    From left to right: \textbf{Class 0} covariance results from slight
+     model mismatch, and correlates nearby residuals. 
+    \textbf{Class I}: A missing absorption line in the model leaves a large,
+     highly correlated patch of negative residuals. 
+    \textbf{Class II}: An extraneous line in the model leaves a large, highly
+     correlated patch of positive residual.  
+    \textbf{Class III}: If lines are present in the model but of the wrong
+     strength, many correlated residuals of moderate amplitude will result. 
+    The difficulty with class III lines is that for any specific line, there might
+     exist a $\vstar$ that will fit the line, but there does not exist a
+     $\vstar$ that will properly fit \emph{all} the lines.}
 \label{fig:badlines}
@@
 
-Due to small imperfections in the flux-calibration of the data spectrum, there may be broad regions of the spectrum which are offset from the true continuum level by a slight amount.
-A traditional approach to address this problem is to normalize both the data and the spectrum to the continuum level before comparison.
-While this may work well for hotter stars with well-defined continuua, for cooler stars with large molecular features, normalization may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.
-Instead, we choose to multiply the model spectrum by a low-order Chebyshev polynomial that is normalized to unity.
-The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters in our model that we will solve for.
-If the behavior of the spectrograph is characterized by observations of spectrophotometric standards, then we can put priors on the coefficients that will prevent overfitting of the polynomials.
-For an example of a typical spectrum, and Chebyshev polynomial, see Figure~\ref{fig:model_data}.
+Due to small imperfections in the flux-calibration of the data spectrum, there
+ may be broad regions of the spectrum which are offset from the true continuum
+ level by a slight amount. 
+A traditional approach to address this problem is to normalize both the data
+ and the spectrum to the continuum level before comparison. 
+While this may work well for hotter stars with well-defined continuua, for
+ cooler stars with large molecular features, normalization may introduce
+ artificial features into the spectrum when the pseudo-continuum is incorrectly
+ placed. 
+Instead, we choose to multiply the model spectrum by a low-order Chebyshev
+ polynomial that is normalized to unity. 
+The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters
+ in our model that we will solve for. 
+If the behavior of the spectrograph is characterized by observations of
+ spectrophotometric standards, then we can put priors on the coefficients that
+ will prevent overfitting of the polynomials. 
+For an example of a typical spectrum, and Chebyshev polynomial, see
+ Figure~\ref{fig:model_data}.
 
-With more uncertainty, the polynomial multiplication can also be used as a substitute for flux-calibration since the sensitivity function of an instrument is usually a smooth function with wavelength.
-When spectra are not flux calibrated, there is greater uncertainty about the properties of the spectrograph and wider priors must be used on the Chebyshev coefficients.
-In this case, the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by the flux calibration for the model.
+With more uncertainty, the polynomial multiplication can also be used as a
+ substitute for flux-calibration since the sensitivity function of an instrument
+ is usually a smooth function with wavelength. 
+When spectra are not flux calibrated, there is greater uncertainty about the
+ properties of the spectrograph and wider priors must be used on the Chebyshev
+ coefficients. 
+In this case, the polynomial might destroy broad-scale information present in
+ the spectrum by fitting it out, something that would be avoided with more
+ accurate priors provided by the flux calibration for the model.
 
 \subsection{Likelihood function}
-To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we use a standard multidimensional Gaussian likelihood function which allows for covariance between data points.
+To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
+ use a standard multidimensional Gaussian likelihood function which allows for
+ covariance between data points. 
 If the vector of residuals is 
@@
 \end{equation}
-$C$ is the covariance matrix which describes the covariance between pixels in the spectrum. For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
+$C$ is the covariance matrix which describes the covariance between pixels in
+ the spectrum. 
+For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
 \begin{equation}
@@
 \end{equation}
-with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the familiar $\chi^2$ form of a sum over the square of the residuals, weighted by the inverse variance of each data point
+with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the
+ familiar $\chi^2$ form of a sum over the square of the residuals, weighted by
+ the inverse variance of each data point
 \begin{equation}
@@
 
-However, residuals from a spectroscopic fit are often correlated due to \todo{DUE TO WHAT, exactly??
-I'm confused.
+However, residuals from a spectroscopic fit are often correlated due to
+ \todo{DUE TO WHAT, exactly?? I'm confused. 
 Wait for response from Tom Loredo}. 
-Additionally, systematic errors in the synthetic spectra (such as incorrect line strengths) will result in regions of highly correlated residuals.
-Properly accounting for correlations in the residuals requires that we use a covariance matrix with off-diagonal terms $\sigma_{ij}$ and a likelihood function (Equation~\ref{eqn:lnprob}) which uses a matrix product instead of a sum over independent pixels.
+Additionally, systematic errors in the synthetic spectra (such as incorrect
+ line strengths) will result in regions of highly correlated residuals. 
+Properly accounting for correlations in the residuals requires that we use a
+ covariance matrix with off-diagonal terms $\sigma_{ij}$ and a likelihood
+ function (Equation~\ref{eqn:lnprob}) which uses a matrix product instead of a
+ sum over independent pixels.
 
-We seek to understand and parameterize two types of covariance structure in our model, exemplified in Figure~\ref{fig:badlines}: the large scale but generally mild global covariance structure that results from [still to be finalized cause], and a regional covariance structure that is localized to areas where spectral line mismatch results in small patches of highly correlated residuals.
-Ignoring the complexity of spectroscopic residuals will bias the estimates of the model parameters $\vt$.
+We seek to understand and parameterize two types of covariance structure in our
+ model, exemplified in Figure~\ref{fig:badlines}: the large scale but generally
+ mild global covariance structure that results from [still to be finalized
+ cause], and a regional covariance structure that is localized to areas where
+ spectral line mismatch results in small patches of highly correlated residuals. 
+Ignoring the complexity of spectroscopic residuals will bias the estimates of
+ the model parameters $\vt$.
 
 As a simple analogy, consider the fit of a straight line to a dataset. 
-If the noise in the dataset is correlated, then adjacent data points might be offset from the linear trend in the same direction by a similar amount.
-If the covariant structure of the noise is ignored, a simple $\chi^2$ will treat these correlated offsets as part of the linear trend, which will result in a biased determination of the slope and intercept of the line, typically with uncertainties that are too small.
+If the noise in the dataset is correlated, then adjacent data points might be
+ offset from the linear trend in the same direction by a similar amount. 
+If the covariant structure of the noise is ignored, a simple $\chi^2$ will
+ treat these correlated offsets as part of the linear trend, which will result
+ in a biased determination of the slope and intercept of the line, typically
+ with uncertainties that are too small.
 
 \subsection{Parameterizing the covariance structure}
-We parameterize the covariance structure using covariance kernels, which sets the covariance between two pixels $\lambda_i$ and $\lambda_j$.
-This behavior is analogous to the two-point correlation function used in cosmology, where the distance between two galaxies is used instead of wavelength.
+We parameterize the covariance structure using covariance kernels, which sets
+ the covariance between two pixels $\lambda_i$ and $\lambda_j$. 
+This behavior is analogous to the two-point correlation function used in
+ cosmology, where the distance between two galaxies is used instead of
+ wavelength.
 
 \paragraph{Global covariance structure}
-A global covariance structure may result from a slight mismatch in the continuum or mild mismatch of spectral lines like that shown as class 0 in Figure~\ref{fig:badlines}.
-To account for this structure, which is generally mild and typically exists across only a few pixels, we use a \emph{stationary} covariance kernel, which satisfies the property that the degree of correlation only depends on the distance between the two pixels $r$.
+A global covariance structure may result from a slight mismatch in the
+ continuum or mild mismatch of spectral lines like that shown as class 0 in
+ Figure~\ref{fig:badlines}. 
+To account for this structure, which is generally mild and typically exists
+ across only a few pixels, we use a \emph{stationary} covariance kernel, which
+ satisfies the property that the degree of correlation only depends on the
+ distance between the two pixels $r$. 
 Functions like this are also called \emph{radial basis functions}. 
@@
 We choose the \matern\ $3/2$ kernel because\ldots
-\todo{Explore difference between using \matern\ kernel and squared exponential kernel.
+\todo{Explore difference between using \matern\ kernel and squared exponential
+ kernel. 
 Why \matern\ and why Hann? 
-Because they are the vanilla covariance functions that are tried and tested in the machine learning communities.}
+Because they are the vanilla covariance functions that are tried and tested in
+ the machine learning communities.}
 \begin{equation}
@@
 \end{equation}
-In order to keep the covariance matrix sparse for computational efficiency, we taper for compact support using a Hann window
+In order to keep the covariance matrix sparse for computational efficiency, we
+ taper for compact support using a Hann window
 \begin{equation}
@@
 \end{equation}
-We add this covariance kernel to the independent variance already present along the matrix diagonal
+We add this covariance kernel to the independent variance already present along
+ the matrix diagonal
 \begin{equation}
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{matern_matrix}
-\includegraphics[width=0.4\textwidth]{matern_draw}
-\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel and typical parameters for our dataset.
-\textbf{Right} general spectroscopic residuals near a stellar continuum overlaid with a sample draw using the covariance matrix, showing that the two look similar in structure and amplitude.
-\protect \todo{Prettify these figures.
-Consistent labelling.
-Change color stretch so Matrix zeros are white.}}
+\includegraphics[width=0.4\textwidth]{figs/matern_matrix}
+\includegraphics[width=0.4\textwidth]{figs/matern_draw}
+\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
+ and typical parameters for our dataset.  
+ \textbf{Right} general spectroscopic residuals near a stellar continuum
+  overlaid with a sample draw using the covariance matrix, showing that the two
+  look similar in structure and amplitude. 
+  \protect \todo{Prettify these figures. Consistent labelling. Change
+  color stretch so Matrix zeros are white.}}
 \label{fig:matern}
@@
 
-To visualize what the covariance matrix $C$ might look like when parameterized by Equation~\ref{eqn:global}, see the left panel of Figure~\ref{fig:matern}.
-To explore the relationship between the parameters of the covariance kernel and the properties of the noise, first consider simulating uncorrelated random Gaussian noise by drawing many independent samples from a Gaussian.
-This same uncorrelated noise could also be represented as a single draw from a $N$-dimensional Gaussian with a diagonal covariance matrix, like in Equation~\ref{eqn:covariance_diagonal}.
-If instead of a diagonal covariance matrix, we were to draw samples from a non-trivial covariance matrix like the one in the left panel of Figure~\ref{fig:matern}, then we will see correlated noise like in the right panel.
-\todo{Do you think examples like Figure~\ref{fig:matern} are helpful or a waste of space?}
+To visualize what the covariance matrix $C$ might look like when parameterized
+ by Equation~\ref{eqn:global}, see the left panel of Figure~\ref{fig:matern}. 
+To explore the relationship between the parameters of the covariance kernel and
+ the properties of the noise, first consider simulating uncorrelated random
+ Gaussian noise by drawing many independent samples from a Gaussian. 
+This same uncorrelated noise could also be represented as a single draw from a
+ $N$-dimensional Gaussian with a diagonal covariance matrix, like in
+ Equation~\ref{eqn:covariance_diagonal}. 
+If instead of a diagonal covariance matrix, we were to draw samples from a
+ non-trivial covariance matrix like the one in the left panel of
+ Figure~\ref{fig:matern}, then we will see correlated noise like in the right
+ panel. 
+\todo{Do you think examples like Figure~\ref{fig:matern} are helpful or a waste
+ of space?}
 
-\paragraph{Specific line covariance} For the large and highly correlated residuals that result from class I, II, and III line errors, we use a \emph{non-stationary} covariance kernel, which means that the covariance explicitly depends on $\lambda_i$ and $\lambda_j$.
-\todo{Explain how we come up with this kernel, which is the covariance resulting from a Gaussian function.}
+\paragraph{Specific line covariance} For the large and highly correlated
+ residuals that result from class I, II, and III line errors, we use a
+ \emph{non-stationary} covariance kernel, which means that the covariance
+ explicitly depends on $\lambda_i$ and $\lambda_j$.
+\todo{Explain how we come up with this kernel, which is the covariance
+ resulting from a Gaussian function.}
 \begin{equation}
@@
 $h$ is a bandwidth. 
-If $h$ is small, then there will be high-frequency structure.
-If $h$ is large, then only low-frequency structure will remain.
+If $h$ is small, then there will be high-frequency structure. If $h$ is large,
+ then only low-frequency structure will remain. 
 We also use the Hann window to taper the kernel for compact support
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{matrix_region_covariance}
-\includegraphics[width=0.4\textwidth]{random_draw}
+\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance}
+\includegraphics[width=0.4\textwidth]{figs/random_draw}
 \caption{\textbf{Left} a covariance matrix generated with the region kernel.
-\textbf{Right} general spectroscopic residuals near a mismatched stellar line overlaid with a sample draw using the covariance matrix, showing that the two look similar in structure and amplitude.
-\protect \todo{Similar improvements to Figure~\ref{fig:matern}.}}
+\textbf{Right} general spectroscopic residuals near a mismatched stellar line
+ overlaid with a sample draw using the covariance matrix, showing that the two
+ look similar in structure and amplitude. 
+\protect \todo{Similar improvements to
+  Figure~\ref{fig:matern}.}}
 \label{fig:region}
@@
 \begin{itemize}
-  \item non-stationary kernel localizes increased variance to a specific spectral line
+  \item non-stationary kernel localizes increased variance to a specific
+    spectral line
   \item better than just masking out the region, since these contain info
-  \item there will be many sets of regions ($\sim$1-4\% of all lines might be covered by a region)
-  \item there can be hyperparameters describing the \emph{population} of regions in a hierarchical model (mostly width, amplitude).
+  \item there will be many sets of regions ($\sim$1-4\% of all lines might be
+    covered by a region)
+  \item there can be hyperparameters describing the \emph{population} of
+    regions in a hierarchical model (mostly width, amplitude).
 \end{itemize}
 
-Covariance kernel parameters are parameters in our model that we explicitly sample for.
+Covariance kernel parameters are parameters in our model that we explicitly
+ sample for.
 
-The benefit mainly comes from simply modelling these systematic residuals to begin with, since a model with covariance is far more likely than forcing the fit, and far more justified than masking regions which do not fit.
+The benefit mainly comes from simply modelling these systematic residuals to
+ begin with, since a model with covariance is far more likely than forcing the
+ fit, and far more justified than masking regions which do not fit.
 
-\todo{Show how the kernel can accounts for complicated M dwarf structure where there is large, messy mismatch.}
+\todo{Show how the kernel can accounts for complicated M dwarf structure where
+ there is large, messy mismatch.}
 
@@
 
-\bibliography{disks,bayesian,master}
+This version of the paper was generated
+ from a git repository available at \url{http://github.com/iancze/StellarSpectra/}
+ with git hash \texttt{\githash\,(\gitdate)}.
+
+\bibliography{stellarspectra}
 \bibliographystyle{hapj}

### 2014-05-05 tex/paper.tex
better word wrapping
more word wrapping
blacked otu table

@@
 
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street MS 10, Cambridge, MA 02138}
+\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street 
+ MS 10, Cambridge, MA 02138}
 %\altaffiltext{2}{Institution 2}
@@
   \item Where our technique fits in the ecosystem of stellar methods.
-  \item Applicable to any kind of spectrum (long-slit, echelle, infrared, flux-calibrated or not)
-  \item Examples of fields where accurate and unbiased stellar parameters are crucial. Exoplanets. T Tauri stars.
+  \item Applicable to any kind of spectrum (long-slit, echelle, infrared,
+    flux-calibrated or not)
+  \item Examples of fields where accurate and unbiased stellar parameters are
+    crucial. Exoplanets. T Tauri stars.
 \end{itemize}
@@
 
-\begin{table}[!htb]
-\begin{tabular}{ll}
-\hline
-\hline
-Symbol & Description\\
-\hline
-\hline
-$i$ & index specifying a pixel\\
-$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-  & that parameterize a synthetic spectrum from the grid\\
-$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-  & are applied during ``post processing'' of the synthetic spectrum\\
-$\vstar$ & $\{\vg,\vpp \}$\\
-$\finst(\lambda)$ & data spectrum\\
-$\fsynth(\lambda)$ & synthetic spectrum\\
-$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
-$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
-$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
-$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
-$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
-$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
-$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
-$\sigma_i$ & Poisson noise for a given pixel $i$\\
-$R_i$ & residuals $\fDi - \fMi$\\
-$\fR$ & residual vector $\fD - \fM$\\
-$C$ & covariance matrix\\
-$k_{\rm global}$  & global covariance kernel\\
-$k_{\rm region}$  & regional covariance kernel\\
-\hline
-\end{tabular}
-\caption{Nomenclature used in this document}
-\label{tab:nomenclature}
-\end{table}
+%\begin{table}[!htb]
+%\begin{tabular}{ll}
+%\hline
+%\hline
+%Symbol & Description\\
+%\hline
+%\hline
+%$i$ & index specifying a pixel\\
+%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+%$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+%  & that parameterize a synthetic spectrum from the grid\\
+%$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+%  & are applied during ``post processing'' of the synthetic spectrum\\
+%$\vstar$ & $\{\vg,\vpp \}$\\
+%$\finst(\lambda)$ & data spectrum\\
+%$\fsynth(\lambda)$ & synthetic spectrum\\
+%$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
+%$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
+%$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
+%$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
+%$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
+%$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
+%$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
+%$\sigma_i$ & Poisson noise for a given pixel $i$\\
+%$R_i$ & residuals $\fDi - \fMi$\\
+%$\fR$ & residual vector $\fD - \fM$\\
+%$C$ & covariance matrix\\
+%$k_{\rm global}$  & global covariance kernel\\
+%$k_{\rm region}$  & regional covariance kernel\\
+%\hline
+%\end{tabular}
+%\caption{Nomenclature used in this document}
+%\label{tab:nomenclature}
+%\end{table}
 
@@
 Spline interpolation for spectra \citep{hus12}.
-\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
-\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
+\todo{Implement band-limited interpolation from the synthetic grid using spline
+ interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be
+ Nyquist sampling.}
+\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for
+ interpolating from a simulation grid.}
 
@@
 \begin{equation}
-  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) \otimes \fsynth(\lambda)
+  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) 
+   \otimes \fsynth(\lambda)
   \label{eqn:convolution}
@@
 \begin{equation}
-  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
+  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, 
+    {\rm synth}}(s)
 \end{equation}
@@
 \begin{equation}
-  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2} \fR^T C^{-1} \fR \right ) 
+  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
+   \fR^T C^{-1} \fR \right ) 
 \end{equation}
@@
 \begin{equation}
-    \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  - \frac{N}{2} \ln 2 \pi 
+  \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
+   - \frac{N}{2} \ln 2 \pi 
   \label{eqn:lnprob}
@@
 \begin{equation}
-  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
+  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i
+   \frac{R_i^2}{\sigma_i^2}
 \end{equation}
@@
 \begin{equation}
-  r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i - \lambda_j}{ \lambda_i + \lambda_j} \right |
+  r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
+   - \lambda_j}{ \lambda_i + \lambda_j} \right |
 \end{equation}
@@
 \begin{equation}
-  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp \left (- \frac{\sqrt{3} r}{l} \right )
+  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
+   \left (- \frac{\sqrt{3} r}{l} \right )
 \end{equation}
@@
 \begin{equation}
-  k_{\rm G}(\lambda_i, \lambda_j | a, \lambda_\mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{r^2(\lambda_i, \lambda_\mu) + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
+  k_{\rm G}(\lambda_i, \lambda_j | a, \lambda_\mu, \sigma) = 
+   \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{r^2(\lambda_i, \lambda_\mu) 
+    + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
 \end{equation}
@@
 \begin{equation}
-  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = w(r, r_0)\; k_{ \rm exp}(r, h) \; k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
+  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = 
+   w(r, r_0)\; k_{ \rm exp}(r, h) \; 
+   k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
 \end{equation}
 
-\todo{Similar paragraph about visualizing the covariance matrix, drawing samples. References to Figure~\ref{fig:region}.}
+\todo{Similar paragraph about visualizing the covariance matrix, drawing
+  samples. References to Figure~\ref{fig:region}.}
 

### 2014-05-06 tex/paper.tex
trying to get figures to work on CF
changing figs
file format
more figs
added yahapj.bst
added table back in
this is a test
merged from upstream

--- 2014-05-05 tex/paper.tex
@@
-%\documentclass[preprint]{aastex} %double-column, single-spaced document:
-\documentclass[iop,floatfix]{emulateapj}
+\documentclass[preprint]{aastex} %double-column, single-spaced document:
+%\documentclass[iop,floatfix]{emulateapj}
 
-\usepackage{hyperref}
+\usepackage[breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
 \usepackage{graphicx}
@@
 \newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}
+\newcommand{\comm}[1]{ \textcolor{Red}{SA: #1}}
 
@@
 
+\graphicspath{{figs/}}
+
 \title{A Method for the spectroscopic inference of fundamental stellar parameters}
@@
 \section{Method}
+
+\comm{This section needs a true introduction of its own.  So far, you just sort 
+of fly right into the details.  The introduction should basically summarize the 
+underlying approach you are going to take, and then literally lay out a basic 
+roadmap for the following subsections.  Some suggested paragraphs below.}
+
+\comm{(1) describe the {\it philosophy} behind the approach: you want to make 
+inferences on fundamental stellar parameters using the vast amount of 
+information available in a spectrum, but at the same time you want to 
+acknowledge and pro-actively account for both intrinsic parametric degeneracies 
+and any practical covariances introduced into the data and the way its 
+treated.  Ideally, this can be done in a flexible, modular way to easily 
+incorporate additional complexity to treat different kinds of problems.  By 
+casting the problem in a specific (Bayesian) framework, you can do that while 
+also simultaneously learning in a data-driven way about how the underlying 
+models of stellar spectra might be improved.  [Note that this paragraph might 
+seem repetitive in the paper, but it sort of encapsulates all the key points so 
+it merits reinforcement.]}
+
+\comm{(2) now give an overview of the basic ``steps" involved in the process; 
+in essence this is a $\sim$1-2 sentence summary of each subsequent subsection.  
+This is to give the reader a ``big picture" overview of the methodology.  If we 
+exclude this, then the following subsections become very confusing and 
+incoherent, even for an expert in the subject.  At the same time, you want to 
+be specific enough that its clear what you are going to do: (a) for a given set 
+of stellar parameters, you generate (somehow) a synthetic model spectrum 
+(Section 2.1); (b) you then post-process that spectrum for realism (using 
+additional parameters) and to match the data (using nuisance parameters) 
+(Section 2.2); (c) you then perform a pixel-by-pixel comparison with the data 
+using a prescribed likelihood function (Section 2.3) and a parametric treatment 
+of the covariance matrix (Section 2.4); and then (d) you iterate this procedure 
+in an MCMC simulation to explore the posterior PDF (Section 2.5).  At the end, 
+you want to be specific about what you ``get" from this procedure: robust 
+estimates of the stellar parameters (and associated nuisance parameters), and a 
+characterization of the covariance matrix that could (in principle) be used to 
+identify and quantify the key discrepancies between the data and the models.}
+
+\subsection{Synthetic Model Spectra (\comm{note change})}
+
+\comm{This subsection needs to make a little more effort to be generic first.  
+Your initial instinct is fine: first introduce the (three) fundamental 
+parameters you care about (but note that these could be expanded in principle, 
+to include things like B-fields, etc.).  But then, we do not want to state a 
+``preferred" way of generating a synthetic model spectrum.  Instead, it would 
+be interesting to make a few comments here harking back to the different 
+approaches you have outlined (or will outline) in Sect 1 -- you can (a) run a 
+stellar structure + radiative transfer code to generate a spectrum 
+corresponding to these parameters (brute force; impractical computationally); 
+(b) interpolate from a grid of stellar structures and run a simple radiative 
+transfer code (i.e., SME); or (c) interpolate from a library of synthetic 
+spectra (i.e., SPC, maybe others...).  Comment on the trade-off for speed 
+versus accuracy (very briefly), and then say that we are going to focus here on 
+the latter approach (although fundamentally the methodology is applicable to 
+all three equally).}
+
 Typically, the stellar properties we most desire are effective temperature
@@
  specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
+\comm{Note the re-arrangement here (below); I think this subsection should deal 
+only with ``obtaining" a full-resolution synthetic model spectrum with the key 
+stellar parameters.  So, I've moved stuff about $v \sin i$, etc. into Sect 2.2.}
 
-In addition to these fundamental parameters, a star also has several observed
- properties that are a function of its kinematics, geometry, and location in our
- galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
- total extinction $A_V$, and solid angle $\Omega$. 
-We call these secondary parameters $\vpp$, because we can model their effects
- on the synthetic spectra in a ``post-processing'' step by convolution with an
- appropriate kernel or multiplication with a smooth function. 
-Together, we call $\vstar = \{\vg, \vpp\}$.
-
+\comm{These sentences should be recast and moved into the Sect 2 introduction.}
 Using a synthetic stellar library as a backend and applying post-processing
@@
 \end{tabular}
-\caption{Nomenclature used in this document}
+\caption{Nomenclature used in this document \comm{Obviously this table does not 
+belong here...}}
 \label{tab:nomenclature}
@@
 \begin{center}
-  \includegraphics{figs/model_data}
+  \includegraphics{figs/model_data.pdf}
   \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
@@
 
-\subsection{Interpolation from synthetic spectral libraries}
 Creating a new spectrum with a specific $\vg$ requires either synthesizing a
@@
 Spline interpolation for spectra \citep{hus12}.
-\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
-\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
+\todo{Implement band-limited interpolation from the synthetic grid using spline
+ interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be
+ Nyquist sampling. \comm{I think this is an important thing to do {\it now}.}}
+\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for
+ interpolating from a simulation grid.  \comm{I do not think this is essential, 
+so it should be put off until the paper is basically done, if even then.}}
 
-\subsection{Post-processing synthetic spectra to match the data}
+\subsection{Post-Processing}
+
+\comm{You need to incorporate these statements (moved from previous subsection) 
+somewhere into the paragraph below.}
+In addition to these fundamental parameters, a star also has several observed
+ properties that are a function of its kinematics, geometry, and location in our
+ galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
+ total extinction $A_V$, and solid angle $\Omega$. 
+We call these secondary parameters $\vpp$, because we can model their effects
+ on the synthetic spectra in a ``post-processing'' step by convolution with an
+ appropriate kernel or multiplication with a smooth function. 
+Together, we call $\vstar = \{\vg, \vpp\}$.
+
 Libraries of synthetic spectra are generally computed at high resolution ($R
@@
 
+\comm{It seems odd to me to do the extinction correction here, after all the 
+work on velocities.  I suppose it does not really matter, but it would seem 
+more harmonious if you finished discussion of all velocity-related manipulation 
+and then applied both the extinction correction and the overall luminosity (or 
+solid angle) scaling after that.  So, maybe re-arrange a bit for a more logical 
+flow (regardless of what the code itself is doing).}
 Finally we apply corrections for interstellar extinction and Doppler shift.
@@
 \todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
- Fourier space, but it's pretty fast as-is.}
+ Fourier space, but it's pretty fast as-is.  \comm{Is there a reason that would 
+be better?  If this is not a bottleneck, I would ignore this.}}
 
@@
 If our spectrum were not properly sampled, we would need to additionally
- convolve with a boxcar the width of the pixels.
+ convolve with a boxcar the width of the pixels.  \comm{I interpret this as a 
+comment to yourself...either integrate into the text in a clear way, or remove.}
 
@@
 \begin{center}
-  \includegraphics[width=0.9\textwidth]{figs/badlines}
+  \includegraphics[width=0.9\textwidth]{figs/badlines.pdf}
   \caption{The variety of residual behaviour, depending on the quality of model
@@
 
+\comm{I feel like this (important) part of the section is rushed into...it is 
+lacking some kind of introductory, pedagogical lead-in to make the reader 
+recognize its power and flexibility.  You are essentially touching on the big 
+issue of parameterizing your calibration.  Now, we don't want to go down that 
+path in detail, but it would be worth discussing by way of a better motivation 
+for using these polynomials.}
 Due to small imperfections in the flux-calibration of the data spectrum, there
@@
 For an example of a typical spectrum, and Chebyshev polynomial, see
- Figure~\ref{fig:model_data}.
+ Figure~\ref{fig:model_data}.  \comm{One thing that has never really been clear 
+to me is how you differentiate the coefficients from the solid angle scaling 
+(or {\it if} you do that.  It would be worthwhile to describe that if necessary 
+especially because I currently see no mention of $L_{\ast}$ or $R_{\ast}/d$ 
+anywhere in this section (and of course, it should be here somewhere).}
 
@@
 
-\subsection{Likelihood function}
+\subsection{Likelihood Calculation}
+
+\comm{I will reserve some comments about notation and symbols for another 
+iteration, but I think you at least want to be very careful about writing the 
+conditional probabilities properly in Eq (5) and (7) -- it should always be 
+$p(D | \theta)$ (and note that its $\theta$, not just $\vstar$).  Ultimately, I 
+would like to see a brief synposis first here as well, primarily to describe 
+the nomenclature up front.  E.g., what do you mean by $D$ and $M$?  I know what 
+you mean, but you cannot assume the reader does without telling them.}
+
 To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
@@
 \begin{equation}
-  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
+  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i
+   \frac{R_i^2}{\sigma_i^2}
 \end{equation}
+\comm{I am not sure how I feel about the dividing line between Sect 2.3 and 
+2.4.  It might make more sense to end right after Eq (5), and then begin Sect 
+2.4 with a discussion of $C$ and the corresponding likelihood in the sample 
+(independent) case, and then progress into the adopted parameterization.  I 
+think I would prefer that, so please wrap up briefly after Eq (5) and push all 
+this text into a re-write of Sect 2.4.  Like we discussed, you ought to just 
+write as if your explanation to Tom Loredo is correct; you can always revisit 
+that if need be when you hear back from him.}
 
@@
 \subsection{Parameterizing the covariance structure}
+
+\comm{I am leaving this alone for now, since it obviously needs more work to be 
+edit-able.  Give it a shot.}
+
 We parameterize the covariance structure using covariance kernels, which sets
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matern_matrix}
-\includegraphics[width=0.4\textwidth]{figs/matern_draw}
+\includegraphics[width=0.4\textwidth]{figs/matern_matrix.png}
+\includegraphics[width=0.4\textwidth]{figs/matern_draw.png}
 \caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance}
-\includegraphics[width=0.4\textwidth]{figs/random_draw}
+\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance.png}
+\includegraphics[width=0.4\textwidth]{figs/random_draw.png}
 \caption{\textbf{Left} a covariance matrix generated with the region kernel.
@@
 
+\bibliographystyle{yahapj}
 \bibliography{stellarspectra}
-\bibliographystyle{hapj}
 \end{document}

### 2014-05-14 tex/paper.tex
halfway done addressing first round of SA edits

--- 2014-05-05 tex/paper.tex
@@
 \newcommand{\vc}{\vec{c}}
-\newcommand{\fM}{ {\bm M}}
+\newcommand{\fM}{ \vec{{\bm M}}}
 \newcommand{\fMi}{M_i}
-\newcommand{\fD}{ {\bm D}}
+\newcommand{\fD}{ \vec{{\bm D}}}
 \newcommand{\fDi}{D_i}
@@
 \section{Method}
-Typically, the stellar properties we most desire are effective temperature
- $T_{\rm eff}$, radius or surface gravity $\log g$, and metallicity $\Z$.
-Several high-quality libraries of synthetic spectra now exist (PHOENIX, Kurucz,
- more being developed for GAIA) and are parameterized by these fundamental
- stellar parameters, which together, we call $\vg$. 
-Synthetic libraries provide high-resolution spectra which span a range of $\vg$
- covering the main sequence of spectral types. The library is typically
- specified on a grid of equal spacing in $T_{\rm eff}$, $\log g$, and $\Z$.
 
+A single spectrum provides a wealth of information that can be used to infer the fundamental properties of a star: effective temperature $T_{\rm eff}$, surface gravity $\log g$, and metallicity $\Z$; together $\vg$.
+We wish to extract the maximal amount information from a given spectrum while also recognizing the limitations imposed by our stellar model and exposing any degeneracies between the parameters.
+For example, if we model a large range of the spectrum at once (e.g., the full optical range at high resolution), then we must explicitly account for any covariances and biases introduced by improper calibration or inaccuracies in the model spectrum.
+We present a generative model which uses a synthetic stellar library as a backend and apply post-processing techniques to mimic the observed spectrum as a function of several stellar parameters and calibration parameters in order to determine the best fit.
+By forward modelling the data spectrum in a Bayesian framework, our modular approach can easily incorporate additional calibration artifacts and stellar complexity (e.g., veiling or Zeeman splitting) while still being true to the limitations of the dataset.
+This data driven approach will enable us to learn how the underlying synthetic models may be improved and refined in an iterative manner.
+
+The roadmap for our method is as follows.
+ First, for a given set of fundamental stellar parameters $\vg$, we generate a synthetic
+ spectrum by interpolating from a grid of synthetic spectra (\S\ref{subsec:synthetic}). 
+Then, we post-process the interpolated spectrum for realism by using additional
+ parameters to describe Doppler shift, projected rotational velocity,
+ interstellar extinction, and residual flux calibration errors
+ (\S\ref{subsec:postprocess}). 
+We perform a pixel-by-pixel comparison of the generated spectrum with the
+ data spectrum using a prescribed likelihood function
+ (\S\ref{subsec:likelihood}) and a parametric treatment of the covariance
+ between pixel residuals (\S\ref{subsec:covariance}). 
+We use a Markov chain Monte Carlo (MCMC) Gibbs sampler to numerically explore
+the posterior probability density and determine constraints on the stellar
+and nuisance parameters (\S\ref{subsec:MCMC}). 
+The benefit of our method is that it provides robust estimates of the stellar and nuisance
+parameters with well-quantified uncertainties and parameter degeneracies. 
+Our parameterization of the covariance between pixel residuals also yields a means to identify
+ and quantify the key discrepancies between the data and model, and provides a path forward to 
+ improve the stellar models in an iterative fashion (\S\ref{subsec:learning}).
+ \todo{I really want to overload the word ``covariance'' and use it to describe both the covariance between $T_{\rm eff}$ and $\log g$ as well as the covariance structure of the pixel residuals themselves.
+However I think this might be confusing so I am using ``degeneracies'' for the former and reserving ``covariance'' for the later.}
+
+\subsection{Synthetic Model Spectra}
+\label{subsec:synthetic}
+Many methods exist to generate a synthetic model of a stellar spectrum. Generally, when
+a high quality spectrum is needed for a specific set of stellar parameters, one
+runs a stellar atmosphere and radiative transfer code to synthesize the
+spectrum. Unfortunately, it is computationally expensive to accurately generate
+a spectrum over a wide range of wavelengths, and so these codes are generally
+employed to generate libraries of stellar spectra, which exist over a range of
+$\vg$ at specified points on a regular grid, typically covering the main sequence of stellar
+types. Several high-quality libraries of synthetic spectra now exist
+(\citealt{hwd+13}, Kurucz, and more are being developed for GAIA).
+
+In a forward-modelling approach, which may make many calls to generate a
+spectra of varying $\vg$, it is currently computationally impractical to
+iterate the structure of the stellar atmosphere and then run the stellar
+synthesis for each call of $\vg$.  Some methods compromise by taking a halfway
+approach, whereby a precomputed library of stellar atmospheres is interpolated
+for a given $\vg$, and then a simple radiative transfer code is run to
+synthesize a spectrum (e.g.  SME; \citealt{vp96}). This technique has the
+advantage that interpolating atmospheres and then running the spectral
+synthesis for each $\vg$ might be more accurate than directly interpolating
+synthetic spectra. However, if a large region of wavelength is required at high
+resolution, then the synthesis step will be slow.  A different approach might
+be instead to eschew forward modelling (and thus repeated synthesis or grid
+interpolation) of the spectra entirely and evaluate the posterior probability
+density only at the grid points of the spectral library.  Then, these discrete
+samples of the posterior density can be interpolated to find the posterior
+surface and draw confidence intervals on the parameters (e.g. SPC;
+\citealt{blj+12}). However, the difficulty with this approach is that the
+parameter uncertainty might be smaller than the grid spacing, and then
+interpolation will not properly capture degeneracies between parameters.
+
+Because we desire a generative model of the spectra without the added
+complexity or computational overhead of running a stellar synthesis, we opt to
+use a \todo{Implement a more accurate but still fast interpolator for the grid.
+Likely this will be a combination of using splines to pre-interpolate a finer
+grid (~20 K, 0.1 dex spacing in gravity and metallicity and then doing linear
+interpolation off of this new grid during the actual MCMC run. It will be described here.} However, the
+techniques we develop in this paper to account for and parameterize the
+correlated, systematic pixel residuals are generally applicable to spectra fit
+by any of these methods.
+
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
+  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
+   to account for inaccuracies in the flux-calibration. 
+  \textbf{Middle}: The data spectrum and model spectrum, after it has been
+   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
+  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
+   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.}
+\label{fig:model_data}
+\end{center}
+\end{figure*}
+
+\subsection{Post-Processing}
+\label{subsec:postprocess}
+
+Libraries of synthetic spectra are generally computed at high resolution ($R
+ \geq $100,000), sampled with many pixels per resolution element, and do not
+ include any rotational broadening or account for any instrumental effects. 
+In order to transform a raw synthetic spectrum interpolated from the grid into
+ one that matches the spectrum of a real star, we must post-process the spectrum
+ to account for these secondary effects. 
 In addition to these fundamental parameters, a star also has several observed
@@
 
-Using a synthetic stellar library as a backend and applying post-processing
- techniques, we attempt to forward-model the observed stellar spectra in order
- to determine the best-fit $\vstar$. 
-Because properly sampled spectra have correlated pixels, we present a framework
- which self-consistently models these correlations.
+The projected rotational velocity of the star, parameterized by $v \sin i$,
+broadens spectral line profiles and is mathematically described by convolution
+with a specific kernel \citep[Equation 18.14]{gra08}. UV, optical, and infrared
+spectra are acquired using a spectrograph which disperses light onto a CCD,
+defining a specific resolution and sampling rate for the spectrum.  Because the
+resolution of the spectrograph and pixel sampling are different from the
+synthetic spectral library, we must convolve the raw spectrum with the line
+spread function (LSF) of the instrument and resample to the exact pixels of the
+CCD. In wavelength space, the $v \sin i$ and LSF operations are represented by
+the convolution of the synthetic spectrum with these two kernels.  Using the
+convolution theorem, we can rewrite these convolutions as multiplications in
+Fourier space.  In order to use the fast Fourier transform (FFT) with discrete
+samples of the synthetic spectrum, we must first resample it
+onto a uniform grid.  In the case of a spectrum, the natural
+uniform grid is one that is equally sampled in velocity space, such there is an
+equal velocity shift $\Delta v$ between pixels. This results in a wavelength
+grid that is linearly sampled in the logarithm of wavelength.  Next, we
+multiply the Fourier transform of the synthetic spectrum by the Fourier
+transforms of the rotational velocity and line spread function kernels.
+Lastly, we do an inverse FFT to transform the modified spectrum back to
+wavelength space, where it is sampled at the wavelength locations of the pixels
+in the detector.  When resampling the spectrum, we are careful to ensure
+band-limited interpolation using splines in order to prevent introducing
+non-physical high frequency structure into the spectrum.
+
+We scale the raw synthetic flux (often reported at the stellar surface) to an
+observed flux by multiplying by the solid angle of the star, $\Omega =
+R_\star^2/d^2$, where $R_\star$ is the stellar radius and $d$ is the distance
+to the star from earth. Then, we correct for a radial velocity difference $v_z$ between the
+star and the earth by Doppler shifting the model spectrum.  Finally, we correct
+for interstellar extinction by dividing by a dust extinction model
+$A_\lambda$, which is parameterized by extinction in the $V$ photometric band,
+$A_V$. 
+\todo{We can add in convolution with photometric filter
+profiles to use photometry and provide a better constraint on the luminosity of
+the star}
+\todo{We might try implementing the convolution in wavelength space, rather
+ than the FFT. 
+This would allow the LSF to change with wavelength.\\}
+
+\paragraph{Calibration polynomials} In contrast to a collection of photometric points in a spectral energy distribution, or any ensemble of data for which each data point is acquired by an individual measurement, adjacent wavelength points in a spectrum are measured simultaneously, with the same detector and--more importantly--the same calibration.
+Typically, spectra are flux calibrated by dividing by a ``sensitivity function,'' which is determined by comparing an observation of a spectrophotometric standard to a previously calibrated model spectrum.
+Therefore, any error in the sensitivity function, which is typically a smooth function with wavelength, will create a region of the data spectrum which is systematically offset by a small amount from what the true flux should be.
+The flux calibration error is then the residual difference in the empirically determined sensitivity function from the what the ``true'' sensitivity function would be if we had a perfectly calibrated instrument and spectrophotometric standard.
+Therefore, it is necessary to introduce additional degrees of freedom into our model in order to account for the calibration error and allow the model spectrum to properly match the data.
+By adding the residual calibration as a parameter in our model, we can explore the distribution of possible calibrations (\S\ref{subsec:MCMC}) and then marginalize the calibration parameters out of the final posterior distribution of the stellar parameters.
+This means that we can derive estimates of stellar parameters that properly account for the intrinsic uncertainty in the flux calibration.
+
+A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.
+While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.
+Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity.
+The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm.
+If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.
+For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}.
+Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts.
+The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$.
+Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$.
+If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$.
+If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order.
+This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.
+
+This polynomial formalism can in principle be used to model raw (not flux calibrated) spectra by using the Chebyshev polynomial to solve directly for the sensitivity function itself, rather than a residual calibration error as is currently done.
+When used in this manner, there will be greater uncertainty about the properties of the spectrograph, which will translate into wider priors on the Chebyshev coefficients.
+In this case, there is more of a chance that the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by a flux calibration.
+
+\subsection{Likelihood Calculation}
+\label{subsec:likelihood}
+
+To evaluate which parameters of the model spectrum fit the data spectrum best,
+we use a likelihood function to compare individual pixels between the two
+spectra. A likelihood function provides the probability of the data spectrum
+$\fD$, given the model spectrum $\fM(\vt)$, which is parameterized by the
+stellar and calibration parameters $\vt$. The data spectrum and model spectrum are both composed of $N$ pixels, and so $\fD$ and $\fM$ represent the collection of these pixels in vector form, 
+$\fD = \{\fDi\}^N_{i=1}$ and $\fM = \{\fMi\}^N_{i=1} $, respectively. The vector of residuals is the difference between the data spectrum and the model spectrum
+\begin{equation}
+  \fR(\vt) = \fD - \fM(\vt)
+\end{equation}
+In the language of probability calculus, the likelihood function is written as
+$p(\fD | \fM(\vt))$, or $p(\fD | \vt)$ for short. We choose a standard
+multidimensional Gaussian likelihood function which allows for covariance
+between the pixel residuals 
+\begin{equation}
+  p(\fD | \vt) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
+   \fR^T C^{-1} \fR \right ) 
+   \label{eqn:prob}
+\end{equation}
+which penalizes models that yield larger residuals. $C$ is the covariance
+matrix which describes the covariance between the pixel residuals. The structure of the covariance matrix is discussed in detail in the next section (\S\ref{subsec:covariance}). Because
+the likelihood function generally exhibits a large dynamic range, it is numerically
+preferred to use the log-likelihood function
+\begin{equation}
+  \ln[p(\fD | \vt)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
+   - \frac{N}{2} \ln 2 \pi 
+  \label{eqn:lnprob}
+\end{equation}
+
+\subsection{Parameterizing the covariance structure}
+\label{subsec:covariance}
+The covariance matrix $C$ describes the measurement noise in each pixel of the spectrum as well as the covariance between pixel residuals.
+It is fundamental to the likelihood calculation described in \S\ref{subsec:likelihood}.
+For most problems, such as fitting a spectral energy distribution to photometric measurements, each data point is acquired in an independent measurement from the other data points, which means that the noise for each data point, $\sigma_i$, is also independent from the other data points.
+This results in a diagonal covariance matrix
+\begin{equation}
+  C = 
+  \begin{bmatrix}
+    \sigma_0^2 & 0  & \cdots & 0\\
+    0 & \sigma_1^2 & \cdots & 0\\
+    \vdots  &   & \ddots  & \vdots \\
+    0 & 0 & \cdots & \sigma_N^2\\
+  \end{bmatrix}
+  \label{eqn:covariance_diagonal}
+\end{equation}
+with $\sigma_{ij} = \delta_{ij} \sigma_{ij}$, where $\delta_{ij}$ is the Kronecker delta function. When the covariance matrix is diagonal, 
+the likelihood function (Equation~\ref{eqn:lnprob}) reduces to the familiar
+$\chi^2$ form of a sum over the square of the residuals, weighted by the
+inverse variance of each data point
+\begin{equation}
+  \ln[p(\fD | \vt)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i^N
+   \frac{R_i^2}{\sigma_i^2}
+   \label{eqn:chi2}
+\end{equation}
+
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines.pdf}
+  \caption{A collection of spectral lines induced by imperfect model fits.
+    From left to right: \textbf{Class 0} The majority of spectral lines
+    ($\gtrsim 60$\%) will have minor differences in strength between data and
+    model spectrum, which produce low-amplitude correlations in the residuals
+    on the length scale of the width of a typical spectral line.  \textbf{Class
+    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+    the model (in this case a line-blended \ion{Ca}{1}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+    of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residual.  \textbf{Class III}: If the line strengths are
+    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+    correlated residuals of moderate amplitude.  The difficulty
+    with class III lines is that for any specific line, there might exist a
+    $\vstar$ that will fit the line, but there does not exist a $\vstar$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure*}
+
+However, the pixel residuals from a spectroscopic fit will often be correlated and will invalidate Equation~\ref{eqn:chi2}.
+These correlations arise because the CCD pixels of a spectrograph do not sample independent wavelengths of the spectrum.
+Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum.
+Because of this, when a spectral feature differs between the model and the data it creates a residual which spans several adjacent data points.
+For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.
+
+\begin{figure*}[!htb]
+\begin{center}
+  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
+  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
+  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but shown at greater zoom to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\sim 3 -4$ pixels.}
+\label{fig:class0}
+\end{center}
+\end{figure*}
+
+There is an important distinction between astrophysical or instrumental ``noise'' and the amplitude and correlation length of the residuals from a spectroscopic fit.
+Noise introduced to a spectrum by astrophysical or instrumental effects is generally uncorrelated with wavelength.
+This is because the arrival of each photon to the detector is a discrete event, and while photons are scattered by the LSF, the particular scattering direction of each photon is independent from any other photon.
+We can verify this by measuring the autocorrelation of residuals in a region of well-fit stellar continuum, and show that it is flat.
+However, when we compare two spectra that have many absorption lines, even small differences between the data and model spectra will produce correlated \emph{residuals}.
+We can verify this correlation by estimating the autocorrelation of the residuals
+\begin{equation}
+  \hat{\rho}\,(k) = \frac{1}{(n - k)\,\sigma^2} \sum_{i = 1}^{n - k} (R_i - \mu_R)\,(R_{i+k} - \mu_R)
+  \label{eqn:autocorrelation}
+\end{equation}
+over a region of spectrum of length $n$.
+Figure~\ref{fig:class0} shows a general residual structure induced by slight model mismatch and the autocorrelation of this series.
+There is significant correlation on length scales of $\sim 3 -4$ pixels, which is also the typical width of a spectral line for this spectrum.
+
+Although the correlation in the residuals is artificially induced by an
+imperfect model fit rather than originating from a physical source of
+correlated noise, we can treat these two effects as mathematically similar. By
+using a non-trivial covariance matrix with off-diagonal terms we can describe
+the small-scale covariance between pixel residuals shown in
+Figures~\ref{fig:badlines}~and~\ref{fig:class0}. A non-diagonal covariance also
+requires that we use  a likelihood function (Equation~\ref{eqn:lnprob}) which
+uses a matrix product instead of a sum over independent pixels.
+
+We seek to understand and parameterize the structure of covariance in the
+spectroscopic residuals, because ignoring this complexity will bias the estimates of
+the model parameters $\vt$. As a simple analogy, consider the fit of a straight
+line to a dataset.  If the noise in the dataset is correlated, then adjacent
+data points might be offset from the linear trend in the same direction by a
+similar amount.  If the covariant structure of the noise is ignored, a simple
+$\chi^2$ will treat these correlated offsets as part of the linear trend, which
+will result in a biased determination of the slope and intercept of the line,
+typically with uncertainties that are too small. We face a mathematically similar problem with 
+correlated pixel residuals from a spectroscopic fit.
+
+We parameterize the covariance structure using covariance kernels, taken from
+the canon of Gaussian processes. Covariance kernels describe the covariance
+between two pixels $\lambda_i$ and $\lambda_j$.  These kernels are analogous to
+the cosmological two-point correlation function, where the distance
+between two galaxies is used instead of wavelength of light.
+
+\paragraph{Global and mildly covariant structure}
+To account for the mild covariance structure shown in Figure~\ref{fig:class0},
+which is generally low amplitude and has a correlation length of only a few
+pixels but is present throughout the majority of the spectrum, we use a
+\emph{stationary} covariance kernel. Stationary covariance kernels, also called \emph{radial basis functions}, satisfy the property that the
+amount of correlation only depends on the distance between the two pixels $r$.  For spectra,
+we map $r$ to the velocity difference between two pixels
+\begin{equation}
+  r\,(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
+   - \lambda_j}{ \lambda_i + \lambda_j} \right |
+\end{equation}
+Then the covariance kernel describes the expected covariance between pixel residuals
+\begin{equation}
+  k(\lambda_i, \lambda_j) =  \langle R_i \; R_j \rangle
+  \label{eqn:expectation}
+\end{equation}
+In theory, we could chose from a variety of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance.
+In practice, there are a few covariance kernels which are well-respected within the machine-learning community and frequently employed to parameterize covariance of this nature \citep{rw05}.
+We choose to use the \matern\ $\nu = 3/2$ kernel because it is one of the most commonly used kernels for this purpose and has an extensive history of use (cite?).
+\begin{equation}
+  k_{\nu = 3/2}(r,\, a,\, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
+   \left (- \frac{\sqrt{3} r}{l} \right )
+\end{equation}
+This kernel is an exponential tapered by a polynomial, which makes for a smooth transition from small scale covariance to negligible covariance at longer scales.
+In order to ensure sparsity of the covariance matrix for computational efficiency, we
+ taper the \matern\ kernel for compact support using a Hann window
+\begin{equation}
+  w\,(r,\, r_0) = \left \{ 
+    \begin{array}{cc}
+    0.5 + 0.5 \cos (\pi r/ r_0 ) & r \le r_0 \\
+    0 & r > r_0\\
+  \end{array}
+  \right .
+  \label{eqn:Hann}
+\end{equation}
+for a final covariance kernel of 
+\begin{equation}
+  k_{\rm global} (r,\, a,\, l,\, r_0) = w(r,\, r_0)\; k_{\nu = 3/2}(r,\, a,\, l) 
+  \label{eqn:global}
+\end{equation}
+We add this covariance kernel to the independent pixel noise $\delta_{ij} \sigma_{ij}$ already present along
+ the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Thus, each element of the covariance matrix is completely described by
+\begin{equation}
+  k(\lambda_i, \lambda_j |\, r,\, a,\, l,\, r_0,\, b) = k_{\rm global}(r, a, l, r_0) + b\,\delta_{ij} \sigma_{ij}
+\end{equation}
+Therefore, the \emph{hyperparameters} of this covariance kernel directly dictate the structure of the covariance matrix.
+By varying the hyperparameters, we can create a covariance matrix which accurately mimics the correlated structure of the pixel residuals.
+Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the \emph{likelihood function will be maximized} (Equation~\ref{eqn:prob}) while encapsulating the \emph{true uncertainty in the model-data fit}.
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
+\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global}.
+There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off such that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
+This matrix can be used to model spectrum which is correlated on a $\sim$few pixel scale.
+\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
+Top: the same residuals shown in Figure~\ref{fig:class0}, plotted against three sets of simulated residuals.
+The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
+\label{fig:matern}
+\end{center}
+\end{figure*}
+
+The left panel of Figure~\ref{fig:matern} shows what the covariance matrix might look like for reasonable choices of the kernel hyperparameters.
+To verify that this covariance matrix indeed mimics the correlated structure seen in the residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our noise model and compare it to the actual pixel residuals.
+Because we are using a multidimensional Gaussian as our likelihood function (Equation~\ref{eqn:prob}), we have assumed that the underlying residual structure in a spectroscopic fit is described by a multidimensional Gaussian.
+This is directly analogous to the implicit assumption of independent Gaussian noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}).
+If one wished to compare the residuals that resulted from a $\chi^2$ fit to the noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$ random samples from a univariate normal distribution with zero mean and variance $\sigma_i^2$.
+We can make the same test with our multidimensional Gaussian likelihood function, however in this case the mean is a vector $\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance matrix $C$.
+The residual vector is then a random draw from this multivariate normal distribution
+\begin{equation}
+  R_{\rm rand} \sim {\cal N}\left( \vec{\mu}=0,\, C\right)
+\end{equation}
+Functions to generate random samples from a multidimensional Gaussian are implemented in most numerical programming languages.
+
+\paragraph{Specific line covariance} Even after the mildly covariant structure has been accounted for by the global covariance kernel (Equation~\ref{eqn:global}), there still exist a significant number large amplitude pixel residuals that are \emph{highly correlated}.
+These patches of large amplitude residuals are caused by incorrect lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}.
+To parameterize the covariance in a given patch of these residuals, we introduce an additional \emph{non-stationary} covariance kernel.
+In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$, and not simply the distance between them, therefore we can now individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.
+
+Because the shape of a spectral line can be generally approximated by a Gaussian function, and the difference of two Gaussians with the same mean is also another Gaussian--the same shape of residuals that are shown in Figure~\ref{fig:badlines}--we use Equation~\ref{eqn:expectation} to design a new covariance kernel about the Gaussian function.
+If we expect the shape of the residuals to be a Gaussian function about the line center $\lambda_\mu$
+\begin{equation}
+  R(\lambda |\, a, \lambda_\mu, \sigma) = \frac{a}{\sqrt{2 \pi} \sigma} \exp \left (-\; \frac{r^2(\lambda, \lambda_\mu)}{2 \sigma^2} \right )
+\end{equation}
+then the Gaussian covariance kernel is
+\begin{align}
+k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma) &= \langle R(\lambda_i)\; R(\lambda_j) \rangle \\
+&= \frac{a^2}{2 \pi \sigma} \exp \left ( - \, \frac{r^2(\lambda_i, \lambda_\mu) 
+    + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
+\end{align}
+We again enforce compact support by tapering the kernel with a Hann window (Equation~\ref{eqn:Hann}), and so the covariance kernel for the highly correlated residuals induced by a poor line fit becomes
+\begin{equation}
+  k_{\rm line}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma, r, r_0) = w(r,\, r_0)\, k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma)
+  \label{eqn:line}
+\end{equation}
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance.png}
+\includegraphics[width=0.4\textwidth]{figs/random_draw.png}
+\caption{\textbf{Left} a covariance matrix generated with the region kernel.
+\textbf{Right} general spectroscopic residuals near a mismatched stellar line
+ overlaid with a sample draw using the covariance matrix, showing that the two
+ look similar in structure and amplitude. 
+\protect \todo{Similar improvements to
+  Figure~\ref{fig:matern}.}}
+\label{fig:region}
+\end{center}
+\end{figure*}
+
+\todo{Similar paragraph about visualizing the covariance matrix, drawing
+  samples. References to Figure~\ref{fig:region}.}
+
+non-stationary kernel localizes increased variance to a specific
+    spectral line
+
+better than just masking out the region, since these contain info
+
+here will be many sets of regions ($\sim$1-4\% of all lines might be
+    covered by a region)
+
+Covariance kernel parameters are parameters in our model that we explicitly sample for.
+There can be hyperparameters describing the \emph{population} of regions in a hierarchical model (mostly width, amplitude).
+
+The benefit mainly comes from simply modelling these systematic residuals to
+ begin with, since a model with covariance is far more likely than forcing the
+ fit, and far more justified than masking regions which do not fit.
+
+Where to draw the line between specific region and global covariance? When we reach some threshold, say three times the global amplitude?
+
+\paragraph{M dwarf residual}
+The squared exponential kernel is
+\todo{Why do we use the squared exponential and not \matern?}
+\begin{equation}
+  k_{ \rm exp}(r, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
+\end{equation}
+$h$ is a bandwidth. 
+If $h$ is small, then there will be high-frequency structure. If $h$ is large,
+ then only low-frequency structure will remain. 
+We also use the Hann window to taper the kernel for compact support
+\begin{equation}
+  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = 
+   w(r, r_0)\; k_{ \rm exp}(r, h) \; 
+   k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
+\end{equation}
+
+\todo{Show how the kernel can accounts for complicated M dwarf structure where
+ there is large, messy mismatch.}
+
+\subsection{Exploring the posterior}
+\label{subsec:MCMC}
+Markov Chain Monte Carlo.
+
+Gibbs sampler.
+Sparse matrix.
+Implemented using SuiteSparse and the Cholesky factorization, means that the RCR step can be computed quickly and reused.
+
+Show figure using dfm/triangle.py that shows recovered stellar parameters and degeneracies. Separately for the regions.
+
+\subsection{Applications}
+\label{subsec:learning}
+Learnt covariance structure can be used to correct the models.
+
+\section{Tests}
+\todo{Which tests to show?}
+
+\bibliographystyle{yahapj}
+\bibliography{stellarspectra}
+
+\appendix
 
@@
 \end{tabular}
-\caption{Nomenclature used in this document}
+\caption{Nomenclature used in this document.}
 \label{tab:nomenclature}
 \end{table}
-
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/model_data}
-  \caption{\textbf{Top}: Chebyshev polynomial which modifies the model spectrum
-   to account for inaccuracies in the flux-calibration. 
-  \textbf{Middle}: The data spectrum and model spectrum, after it has been
-   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
-  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
-   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}. 
-  \protect \todo{Model spectrum will be updated once I properly burn in the
-   chain. Do these line styles look OK? I was going for something legible
-   that might also print well in B/W}.}
-\label{fig:model_data}
-\end{center}
-\end{figure*}
-
-\subsection{Interpolation from synthetic spectral libraries}
-Creating a new spectrum with a specific $\vg$ requires either synthesizing a
- new spectrum using radiative transfer codes or interpolating from nearby grid
- points in the library. 
-Because it is computationally expensive to synthesize a spectrum at high
- resolution over a wide wavelength range, we choose to interpolate. 
-Spline interpolation for spectra \citep{hus12}.
-\todo{Implement band-limited interpolation from the synthetic grid using spline interpolation. Caution: As is, 100K/0.5 logg spacing in $\vg$ may not be Nyquist sampling.}
-\todo{Investigate a ``Bayesian Emulator'' for using Gaussian processes for interpolating from a simulation grid.}
-
-\subsection{Post-processing synthetic spectra to match the data}
-Libraries of synthetic spectra are generally computed at high resolution ($R
- \geq $100,000), sampled with many pixels per resolution element, and do not
- include any rotational broadening or account for any instrumental effects. 
-In order to transform a raw synthetic spectrum interpolated from the grid into
- one that matches the spectrum of a real star, we must post-process the spectrum
- to account for these secondary effects. 
-The projected rotational velocity of the star, parameterized by $v \sin i$,
- broadens spectral line profiles and is mathematically described by convolution
- with a kernel $g_{v \sin i}(\lambda)$.  
-UV, optical, and infrared spectra are acquired using a spectrograph which
- disperses light onto a CCD, defining a specific resolution and sampling rate
- for the spectrum. 
-Because the resolution of the spectrograph and spacing of the pixels are
- different from the synthetic spectral library, we must convolve the raw
- spectrum with the line spread function (LSF) of the instrument $g_{\rm
- LSF}(\lambda)$ and resample to the exact pixels of the CCD. In wavelength
- space, the $v \sin i$ and LSF operations are represented by the convolution of
- the synthetic spectrum with these two kernels
-\begin{equation}
-  \finst(\lambda) = g_{v \sin i}(\lambda) \otimes g_{\rm LSF}(\lambda) \otimes \fsynth(\lambda)
-  \label{eqn:convolution}
-\end{equation}
-Using the convolution theorem, we can rewrite these operations as
- multiplications in Fourier space 
-\begin{equation}
-  F_{\lambda, {\rm inst}}(s) = G_{\rm LSF}(s) G_{v \sin i}(s) F_{\lambda, {\rm synth}}(s)
-\end{equation}
- where $G \leftrightarrow g$ and $F \leftrightarrow f$ are Fourier transform
- pairs. 
-In order to use the fast Fourier transform (FFT) to execute these
- operations using discrete samples of the synthetic spectrum $\fsynth$, we must
- first resample $\fsynth$ so that it is on a uniform grid. 
-In the case of a spectrum, the natural uniform grid is one that is equally
- sampled in velocity space, such there is an equal velocity shift $\Delta v$
- between pixels. This results in a wavelength grid that is linearly sampled in
- the logarithm of wavelength. 
-Therefore, the Fourier coordinate $s$ is the number of cycles per sampling
- interval, having units of inverse velocity $[{\rm s}/{\rm km}]$. 
-Next, we multiply the Fourier transform of the synthetic spectrum by the
- Fourier transforms of the rotational velocity and line spread function kernels.
-Lastly, we do an inverse FFT to transform the modified spectrum $F_{\lambda,
- {\rm inst}}$ back to wavelength space, where it is sampled at the wavelength
- locations of the pixels in the detector ($\finst$). 
-When resampling the spectrum, we are careful to ensure band-limited
- interpolation using splines in order to prevent introducing non-physical high
- frequency structure into the spectrum.
-
-Finally we apply corrections for interstellar extinction and Doppler shift.
-Interstellar extinction attenuates the spectrum by an amount $A_\lambda$, which
-is wavelength-dependent. 
-We use the (name here) dust-extinction model, which is parameterized by
- extinction in the $V$ photometric band, $A_V$. 
-We correct for a radial velocity difference $v_z$ between the star and the
- earth by Doppler shifting the model spectrum. 
-\todo{Doppler shifting the wavelengths could be implemented as a phase-shift in
- Fourier space, but it's pretty fast as-is.}
-
-\todo{We might try implementing the convolution in wavelength space, rather
- than the FFT. 
-This would allow the LSF to change with wavelength.\\} 
-Pixel width effects are not important as long as the LSF of the instrument is
- adequately sampled ($\gtrsim 3$ pixels across the LSF). 
-If our spectrum were not properly sampled, we would need to additionally
- convolve with a boxcar the width of the pixels.
-
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=0.9\textwidth]{figs/badlines}
-  \caption{The variety of residual behaviour, depending on the quality of model
-    fit. 
-    From left to right: \textbf{Class 0} covariance results from slight
-     model mismatch, and correlates nearby residuals. 
-    \textbf{Class I}: A missing absorption line in the model leaves a large,
-     highly correlated patch of negative residuals. 
-    \textbf{Class II}: An extraneous line in the model leaves a large, highly
-     correlated patch of positive residual.  
-    \textbf{Class III}: If lines are present in the model but of the wrong
-     strength, many correlated residuals of moderate amplitude will result. 
-    The difficulty with class III lines is that for any specific line, there might
-     exist a $\vstar$ that will fit the line, but there does not exist a
-     $\vstar$ that will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure*}
-
-Due to small imperfections in the flux-calibration of the data spectrum, there
- may be broad regions of the spectrum which are offset from the true continuum
- level by a slight amount. 
-A traditional approach to address this problem is to normalize both the data
- and the spectrum to the continuum level before comparison. 
-While this may work well for hotter stars with well-defined continuua, for
- cooler stars with large molecular features, normalization may introduce
- artificial features into the spectrum when the pseudo-continuum is incorrectly
- placed. 
-Instead, we choose to multiply the model spectrum by a low-order Chebyshev
- polynomial that is normalized to unity. 
-The coefficients of the polynomial, $c_0, c_1, \ldots$, will be free-parameters
- in our model that we will solve for. 
-If the behavior of the spectrograph is characterized by observations of
- spectrophotometric standards, then we can put priors on the coefficients that
- will prevent overfitting of the polynomials. 
-For an example of a typical spectrum, and Chebyshev polynomial, see
- Figure~\ref{fig:model_data}.
-
-With more uncertainty, the polynomial multiplication can also be used as a
- substitute for flux-calibration since the sensitivity function of an instrument
- is usually a smooth function with wavelength. 
-When spectra are not flux calibrated, there is greater uncertainty about the
- properties of the spectrograph and wider priors must be used on the Chebyshev
- coefficients. 
-In this case, the polynomial might destroy broad-scale information present in
- the spectrum by fitting it out, something that would be avoided with more
- accurate priors provided by the flux calibration for the model.
-
-\subsection{Likelihood function}
-To evaluate which parameters of the model $\fM$ fit the data set $\fD$ best we
- use a standard multidimensional Gaussian likelihood function which allows for
- covariance between data points. 
-If the vector of residuals is 
-\begin{equation}
-  \fR(\vstar) = \fD - \fM(\vstar)
-\end{equation}
- with a length of $N$ data points, then the likelihood function is
-\begin{equation}
-  p(\fD | \vstar) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2} \fR^T C^{-1} \fR \right ) 
-\end{equation}
- and the log-likelihood function is
-\begin{equation}
-    \ln[p(\vstar)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  - \frac{N}{2} \ln 2 \pi 
-    \label{eqn:lnprob}
-\end{equation}
-$C$ is the covariance matrix which describes the covariance between pixels in
- the spectrum. 
-For independent noise with noise per pixel $\sigma_i$, this matrix is diagonal
-\begin{equation}
-  C = 
-  \begin{bmatrix}
-    \sigma_0^2 & 0  & \cdots & 0\\
-    0 & \sigma_1^2 & \cdots & 0\\
-    \vdots  &   & \ddots  & \vdots \\
-    0 & 0 & \cdots & \sigma_N^2\\
-  \end{bmatrix}
-  \label{eqn:covariance_diagonal}
-\end{equation}
-with $\sigma_{ij} = 0$ everywhere and equation~\ref{eqn:lnprob} reduces to the
- familiar $\chi^2$ form of a sum over the square of the residuals, weighted by
- the inverse variance of each data point
-\begin{equation}
-  \ln[p(\vstar)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i \frac{R_i^2}{\sigma_i^2}
-\end{equation}
-
-However, residuals from a spectroscopic fit are often correlated due to
- \todo{DUE TO WHAT, exactly?? I'm confused. 
-Wait for response from Tom Loredo}. 
-Additionally, systematic errors in the synthetic spectra (such as incorrect
- line strengths) will result in regions of highly correlated residuals. 
-Properly accounting for correlations in the residuals requires that we use a
- covariance matrix with off-diagonal terms $\sigma_{ij}$ and a likelihood
- function (Equation~\ref{eqn:lnprob}) which uses a matrix product instead of a
- sum over independent pixels.
-
-We seek to understand and parameterize two types of covariance structure in our
- model, exemplified in Figure~\ref{fig:badlines}: the large scale but generally
- mild global covariance structure that results from [still to be finalized
- cause], and a regional covariance structure that is localized to areas where
- spectral line mismatch results in small patches of highly correlated residuals. 
-Ignoring the complexity of spectroscopic residuals will bias the estimates of
- the model parameters $\vt$.
-
-As a simple analogy, consider the fit of a straight line to a dataset. 
-If the noise in the dataset is correlated, then adjacent data points might be
- offset from the linear trend in the same direction by a similar amount. 
-If the covariant structure of the noise is ignored, a simple $\chi^2$ will
- treat these correlated offsets as part of the linear trend, which will result
- in a biased determination of the slope and intercept of the line, typically
- with uncertainties that are too small.
-
-\subsection{Parameterizing the covariance structure}
-We parameterize the covariance structure using covariance kernels, which sets
- the covariance between two pixels $\lambda_i$ and $\lambda_j$. 
-This behavior is analogous to the two-point correlation function used in
- cosmology, where the distance between two galaxies is used instead of
- wavelength.
-
-\paragraph{Global covariance structure}
-A global covariance structure may result from a slight mismatch in the
- continuum or mild mismatch of spectral lines like that shown as class 0 in
- Figure~\ref{fig:badlines}. 
-To account for this structure, which is generally mild and typically exists
- across only a few pixels, we use a \emph{stationary} covariance kernel, which
- satisfies the property that the degree of correlation only depends on the
- distance between the two pixels $r$. 
-Functions like this are also called \emph{radial basis functions}. 
-For spectra, we map $r$ to the velocity difference between two pixels
-\begin{equation}
-  r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i - \lambda_j}{ \lambda_i + \lambda_j} \right |
-\end{equation}
-We choose the \matern\ $3/2$ kernel because\ldots
-\todo{Explore difference between using \matern\ kernel and squared exponential
- kernel. 
-Why \matern\ and why Hann? 
-Because they are the vanilla covariance functions that are tried and tested in
- the machine learning communities.}
-\begin{equation}
-  k_{\nu = 3/2}(r, a, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp \left (- \frac{\sqrt{3} r}{l} \right )
-\end{equation}
-In order to keep the covariance matrix sparse for computational efficiency, we
- taper for compact support using a Hann window
-\begin{equation}
-  w(r, r_0) = \frac{1}{2} + \frac{1}{2} \cos \left( \frac{\pi r}{r_0} \right) 
-\end{equation}
-for a global covariance kernel of 
-\begin{equation}
-  k_{\rm global} (r, a, l, r_0) = w(r, r_0) k_{\nu = 3/2}(r, a, l) 
-  \label{eqn:global}
-\end{equation}
-We add this covariance kernel to the independent variance already present along
- the matrix diagonal
-\begin{equation}
-  k(\lambda_i, \lambda_j) = k_{\rm global}(r, a, l, r_0) + \delta_{ij} \sigma_{ij}^2
-\end{equation}
-where $\delta_{ij}$ is the Kroenecker delta function.
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matern_matrix}
-\includegraphics[width=0.4\textwidth]{figs/matern_draw}
-\caption{\textbf{Left} a covariance matrix generated with the \matern\ kernel
- and typical parameters for our dataset.  
- \textbf{Right} general spectroscopic residuals near a stellar continuum
-  overlaid with a sample draw using the covariance matrix, showing that the two
-  look similar in structure and amplitude. 
-  \protect \todo{Prettify these figures. Consistent labelling. Change
-  color stretch so Matrix zeros are white.}}
-\label{fig:matern}
-\end{center}
-\end{figure*}
-
-To visualize what the covariance matrix $C$ might look like when parameterized
- by Equation~\ref{eqn:global}, see the left panel of Figure~\ref{fig:matern}. 
-To explore the relationship between the parameters of the covariance kernel and
- the properties of the noise, first consider simulating uncorrelated random
- Gaussian noise by drawing many independent samples from a Gaussian. 
-This same uncorrelated noise could also be represented as a single draw from a
- $N$-dimensional Gaussian with a diagonal covariance matrix, like in
- Equation~\ref{eqn:covariance_diagonal}. 
-If instead of a diagonal covariance matrix, we were to draw samples from a
- non-trivial covariance matrix like the one in the left panel of
- Figure~\ref{fig:matern}, then we will see correlated noise like in the right
- panel. 
-\todo{Do you think examples like Figure~\ref{fig:matern} are helpful or a waste
- of space?}
-
-\paragraph{Specific line covariance} For the large and highly correlated
- residuals that result from class I, II, and III line errors, we use a
- \emph{non-stationary} covariance kernel, which means that the covariance
- explicitly depends on $\lambda_i$ and $\lambda_j$.
-\todo{Explain how we come up with this kernel, which is the covariance
- resulting from a Gaussian function.}
-\begin{equation}
-  k_{\rm G}(\lambda_i, \lambda_j | a, \lambda_\mu, \sigma) = \frac{a^2}{2 \pi \sigma} \exp \left ( - \frac{r^2(\lambda_i, \lambda_\mu) + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
-\end{equation}
-The squared exponential kernel is
-\todo{Why do we use the squared exponential and not \matern?}
-\begin{equation}
-  k_{ \rm exp}(r, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
-\end{equation}
-$h$ is a bandwidth. 
-If $h$ is small, then there will be high-frequency structure. If $h$ is large,
- then only low-frequency structure will remain. 
-We also use the Hann window to taper the kernel for compact support
-\begin{equation}
-  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = w(r, r_0)\; k_{ \rm exp}(r, h) \; k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
-\end{equation}
-
-\todo{Similar paragraph about visualizing the covariance matrix, drawing samples. References to Figure~\ref{fig:region}.}
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance}
-\includegraphics[width=0.4\textwidth]{figs/random_draw}
-\caption{\textbf{Left} a covariance matrix generated with the region kernel.
-\textbf{Right} general spectroscopic residuals near a mismatched stellar line
- overlaid with a sample draw using the covariance matrix, showing that the two
- look similar in structure and amplitude. 
-\protect \todo{Similar improvements to
-  Figure~\ref{fig:matern}.}}
-\label{fig:region}
-\end{center}
-\end{figure*}
-
-\begin{itemize}
-  \item non-stationary kernel localizes increased variance to a specific
-    spectral line
-  \item better than just masking out the region, since these contain info
-  \item there will be many sets of regions ($\sim$1-4\% of all lines might be
-    covered by a region)
-  \item there can be hyperparameters describing the \emph{population} of
-    regions in a hierarchical model (mostly width, amplitude).
-\end{itemize}
-
-Covariance kernel parameters are parameters in our model that we explicitly
- sample for.
-
-The benefit mainly comes from simply modelling these systematic residuals to
- begin with, since a model with covariance is far more likely than forcing the
- fit, and far more justified than masking regions which do not fit.
-
-\todo{Show how the kernel can accounts for complicated M dwarf structure where
- there is large, messy mismatch.}
-
-\subsection{Exploring the posterior}
-Markov Chain Monte Carlo
-\todo{Given this is 2014 and not 2010, how much do you think I need to explain?}
-
-\subsection{Applications}
-Learnt covariance structure can be used to correct the models.
-
-\section{Tests}
-\todo{Which tests to show?}
 
@@
 
-\bibliography{stellarspectra}
-\bibliographystyle{hapj}
 \end{document}

### 2014-05-21 tex/paper.tex
Updated paper for SA review

@@
 \newcommand{\yN}{\{y_i\}_{i=1}^N}
-\newcommand{\vt}{\vec{\theta}}
+
+%\vt stands for ``vector theta''
+\newcommand{\vt}{ {\bm \theta}}
 \newcommand{\vg}{\vt_{\star, {\rm grid}}}
 \newcommand{\vpp}{\vt_{\star, {\rm post}}}
-\newcommand{\vstar}{\vt_{\star}}
 \newcommand{\finst}{f_{\lambda, {\rm inst}}}
@@
 \newcommand{\vN}{\vt_{\rm N}}
-\newcommand{\vc}{\vec{c}}
+\newcommand{\vtstar}{\vt_{\star}}
+\newcommand{\vtcheb}{\vt_{\rm Cheb}}
+\newcommand{\vtglobal}{\vt_{\rm global}}
+\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
+\newcommand{\vtorders}{\vt_{\rm orders}}
+\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
+\newcommand{\vtlines}{\vt_{\rm lines}}
+\newcommand{\vtcov}{\vt_{\rm cov}}
 \newcommand{\fM}{ \vec{{\bm M}}}
@@
   \textbf{Bottom}: Residuals from the model fit. Note the large residual at
-   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{1}.}
+   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{2}.}
 \label{fig:model_data}
@@
  appropriate kernel or multiplication with a smooth function. 
-Together, we call $\vstar = \{\vg, \vpp\}$.
+Together, we call $\vtstar = \{\vg, \vpp\}$.
 
@@
 Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity.
-The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm.
+The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm, $\vtcheb = \{c_0, c_1, \ldots, c_N \}$.
 If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.
@@
   \includegraphics{figs/badlines.pdf}
-  \caption{A collection of spectral lines induced by imperfect model fits.
+  \caption{A collection of spectral lines which have imperfect model fits.
     From left to right: \textbf{Class 0} The majority of spectral lines
@@
     I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{1}) leaves a large, highly correlated
+    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
     patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
     of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residual.  \textbf{Class III}: If the line strengths are
+    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
     substantially discrepant ($\lesssim 10$\% of all lines), there will be many
@@
     with class III lines is that for any specific line, there might exist a
-    $\vstar$ that will fit the line, but there does not exist a $\vstar$ that
+    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
     will properly fit \emph{all} the lines.}
@@
 Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum.
-Because of this, when a spectral feature differs between the model and the data it creates a residual which spans several adjacent data points.
+Because of this, when a spectral feature differs between the model and the data, it creates a residual which spans several adjacent data points.
 For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.
@@
   $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
-  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but shown at greater zoom to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\sim 3 -4$ pixels.}
+  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
 \label{fig:class0}
@@
 
-There is an important distinction between astrophysical or instrumental ``noise'' and the amplitude and correlation length of the residuals from a spectroscopic fit.
+There is an important distinction between astrophysical or instrumental ``noise'' and the residuals from a spectroscopic fit.
 Noise introduced to a spectrum by astrophysical or instrumental effects is generally uncorrelated with wavelength.
@@
 We can verify this by measuring the autocorrelation of residuals in a region of well-fit stellar continuum, and show that it is flat.
-However, when we compare two spectra that have many absorption lines, even small differences between the data and model spectra will produce correlated \emph{residuals}.
-We can verify this correlation by estimating the autocorrelation of the residuals
+If the model spectrum is \emph{perfect}, then the residuals from a spectroscopic fit will be uncorrelated and equivalent to the astrophysical and instrumental noise introduced into the measurement.
+However, when we compare a data spectrum and a model spectrum that have many absorption lines, even small differences will produce correlated \emph{residuals}.
+We can verify this correlation by estimating the autocorrelation of the residuals as a function of pixel offset $k$
 \begin{equation}
@@
 \end{equation}
-over a region of spectrum of length $n$.
+over a region of spectrum of length $n$ pixels.
 Figure~\ref{fig:class0} shows a general residual structure induced by slight model mismatch and the autocorrelation of this series.
-There is significant correlation on length scales of $\sim 3 -4$ pixels, which is also the typical width of a spectral line for this spectrum.
+There is significant correlation for offsets of $\lesssim 4$ pixels, which is also the typical width of a spectral line for this spectrum.
 
@@
 Figures~\ref{fig:badlines}~and~\ref{fig:class0}. A non-diagonal covariance also
-requires that we use  a likelihood function (Equation~\ref{eqn:lnprob}) which
-uses a matrix product instead of a sum over independent pixels.
+requires that we use  a likelihood function  which uses a matrix product instead of a sum over independent pixels (Equation~\ref{eqn:lnprob}).
 
@@
 the model parameters $\vt$. As a simple analogy, consider the fit of a straight
-line to a dataset.  If the noise in the dataset is correlated, then adjacent
+line to a dataset.  If the noise in the dataset is correlated, then nearby
 data points might be offset from the linear trend in the same direction by a
 similar amount.  If the covariant structure of the noise is ignored, a simple
-$\chi^2$ will treat these correlated offsets as part of the linear trend, which
+$\chi^2$ will treat these correlated offsets as due to the linear trend, which
 will result in a biased determination of the slope and intercept of the line,
@@
 We parameterize the covariance structure using covariance kernels, taken from
-the canon of Gaussian processes. Covariance kernels describe the covariance
+the field of Gaussian processes. Covariance kernels describe the covariance
 between two pixels $\lambda_i$ and $\lambda_j$.  These kernels are analogous to
 the cosmological two-point correlation function, where the distance
-between two galaxies is used instead of wavelength of light.
+between two galaxies is used instead of the wavelength of light.
 
@@
 \end{equation}
-In theory, we could chose from a variety of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance.
+In theory, we could chose from a plethora of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance.
 In practice, there are a few covariance kernels which are well-respected within the machine-learning community and frequently employed to parameterize covariance of this nature \citep{rw05}.
@@
 \begin{equation}
-  k_{\nu = 3/2}(r,\, a,\, l) = a \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
+  k_{\nu = 3/2}(r\, |\, a_{\rm g},\, l) = a_{\rm g} \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
    \left (- \frac{\sqrt{3} r}{l} \right )
@@
 \begin{equation}
-  w\,(r,\, r_0) = \left \{ 
+  w\,(r |\, r_0) = \left \{ 
     \begin{array}{cc}
@@
 \end{equation}
-for a final covariance kernel of 
+for which we generally fix $r_0$ to be some fixed multiple of the length scale, say $r_0 = 4 l$.
+Then, if we define the global parameters $\vtglobal = \{a_g, l \}$, we have a final covariance kernel of
 \begin{equation}
-  k_{\rm global} (r,\, a,\, l,\, r_0) = w(r,\, r_0)\; k_{\nu = 3/2}(r,\, a,\, l) 
+  k_{\rm global} (r\,|\,\vtglobal) = w(r |\, r_0 = 4l)\; k_{\nu = 3/2}(r |\, a_{\rm g},\, l) 
   \label{eqn:global}
 \end{equation}
+\todo{This might be a good place to actually include a simple plot of $k_{\rm global}$ as a function of $r$, as long as we aren't providing too many figures.
+Maybe it could be combined with Figure~\ref{fig:matern}.}
+
 We add this covariance kernel to the independent pixel noise $\delta_{ij} \sigma_{ij}$ already present along
- the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Thus, each element of the covariance matrix is completely described by
+ the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Then, each element of the covariance matrix is completely described by
 \begin{equation}
-  k(\lambda_i, \lambda_j |\, r,\, a,\, l,\, r_0,\, b) = k_{\rm global}(r, a, l, r_0) + b\,\delta_{ij} \sigma_{ij}
+  k(\lambda_i, \lambda_j |\vtglobal, \, b) = k_{\rm global}(r | \vtglobal) + b\,\delta_{ij} \sigma_{ij}
 \end{equation}
-Therefore, the \emph{hyperparameters} of this covariance kernel directly dictate the structure of the covariance matrix.
+Therefore, the \emph{hyperparameters} of this covariance kernel directly determine the structure of the covariance matrix.
 By varying the hyperparameters, we can create a covariance matrix which accurately mimics the correlated structure of the pixel residuals.
-Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the \emph{likelihood function will be maximized} (Equation~\ref{eqn:prob}) while encapsulating the \emph{true uncertainty in the model-data fit}.
+Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the likelihood function will be maximized (Equation~\ref{eqn:prob}) while also encapsulating the true uncertainty in the spectroscopic fit.
 
@@
 \includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
-\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global}.
-There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off such that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
-This matrix can be used to model spectrum which is correlated on a $\sim$few pixel scale.
+\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
+There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
+This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
 \textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
-Top: the same residuals shown in Figure~\ref{fig:class0}, plotted against three sets of simulated residuals.
+The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
 The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
@@
 
-The left panel of Figure~\ref{fig:matern} shows what the covariance matrix might look like for reasonable choices of the kernel hyperparameters.
-To verify that this covariance matrix indeed mimics the correlated structure seen in the residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our noise model and compare it to the actual pixel residuals.
-Because we are using a multidimensional Gaussian as our likelihood function (Equation~\ref{eqn:prob}), we have assumed that the underlying residual structure in a spectroscopic fit is described by a multidimensional Gaussian.
-This is directly analogous to the implicit assumption of independent Gaussian noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}).
-If one wished to compare the residuals that resulted from a $\chi^2$ fit to the noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$ random samples from a univariate normal distribution with zero mean and variance $\sigma_i^2$.
-We can make the same test with our multidimensional Gaussian likelihood function, however in this case the mean is a vector $\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance matrix $C$.
-The residual vector is then a random draw from this multivariate normal distribution
+The left panel of Figure~\ref{fig:matern} shows what the covariance matrix
+might look like for reasonable choices of the kernel hyperparameters. To verify
+that this covariance matrix indeed mimics the correlated structure seen in the
+residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our
+noise model and compare it to the actual pixel residuals. Because we are using
+a multidimensional Gaussian as our likelihood function
+(Equation~\ref{eqn:prob}), we have assumed that the underlying residual
+structure in a spectroscopic fit is described by a multidimensional Gaussian.
+This is directly analogous to the implicit assumption of independent Gaussian
+noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}). If
+one wished to compare the residuals that resulted from a $\chi^2$ fit to the
+noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$
+random samples from a univariate normal distribution with zero mean and
+variance $\sigma_i^2$. We can make the same test with our multidimensional
+Gaussian likelihood function, however in this case the mean is a vector
+$\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance
+matrix $C$. The residual vector is then a random draw from this multivariate
+normal distribution 
 \begin{equation}
@@
 Functions to generate random samples from a multidimensional Gaussian are implemented in most numerical programming languages.
+As we see in the right panel of Figure~\ref{fig:matern}, the structure and amplitude of the fake residuals closely mimics that of the actual data residuals, suggesting that the covariance kernel faithfully reproduces the covariance induced by the data-model mismatch.
 
 \paragraph{Specific line covariance} Even after the mildly covariant structure has been accounted for by the global covariance kernel (Equation~\ref{eqn:global}), there still exist a significant number large amplitude pixel residuals that are \emph{highly correlated}.
-These patches of large amplitude residuals are caused by incorrect lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}.
+These patches of large amplitude residuals are caused by incorrect synthetic lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}.
 To parameterize the covariance in a given patch of these residuals, we introduce an additional \emph{non-stationary} covariance kernel.
-In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$, and not simply the distance between them, therefore we can now individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.
+In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$ and not simply the distance between them, enabling us to individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.
 
-Because the shape of a spectral line can be generally approximated by a Gaussian function, and the difference of two Gaussians with the same mean is also another Gaussian--the same shape of residuals that are shown in Figure~\ref{fig:badlines}--we use Equation~\ref{eqn:expectation} to design a new covariance kernel about the Gaussian function.
-If we expect the shape of the residuals to be a Gaussian function about the line center $\lambda_\mu$
+Because the shape of a spectral line can be generally approximated by a Gaussian function, and the difference of two Gaussians with the same mean is also another Gaussian, we use Equation~\ref{eqn:expectation} to design a new covariance kernel about the Gaussian function.
+Consider the residuals from a single spectral line, labeled by the subscript $0$.
+If we expect the shape of the residuals to be a Gaussian function about the line center with parameters $\vt_{{\rm line}_1} = \{ a_1, \lambda_{\mu_1}, \sigma_1 \}$
 \begin{equation}
-  R(\lambda |\, a, \lambda_\mu, \sigma) = \frac{a}{\sqrt{2 \pi} \sigma} \exp \left (-\; \frac{r^2(\lambda, \lambda_\mu)}{2 \sigma^2} \right )
+  R(\lambda | \vtline{1} ) = \frac{a_1}{\sqrt{2 \pi} \sigma_1} \exp \left (-\; \frac{r^2(\lambda, \lambda_{\mu_1})}{2 \sigma_1^2} \right )
 \end{equation}
-then the Gaussian covariance kernel is
+then the covariance of any two pixels along this Gaussian residual is 
 \begin{align}
-k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma) &= \langle R(\lambda_i)\; R(\lambda_j) \rangle \\
-&= \frac{a^2}{2 \pi \sigma} \exp \left ( - \, \frac{r^2(\lambda_i, \lambda_\mu) 
-    + r^2(\lambda_j, \lambda_\mu)}{2 \sigma^2}\right )
+  k_{\rm G}(\lambda_i, \lambda_j |\vtline{1} ) &= \langle R(\lambda_i)\; R(\lambda_j) \rangle \\
+  &= \frac{a_1^2}{2 \pi \sigma_1} \exp \left ( - \, \frac{r^2(\lambda_i, \lambda_{\mu_1}) 
+  + r^2(\lambda_j, \lambda_{\mu_1})}{2 \sigma_1^2}\right )
 \end{align}
-We again enforce compact support by tapering the kernel with a Hann window (Equation~\ref{eqn:Hann}), and so the covariance kernel for the highly correlated residuals induced by a poor line fit becomes
+We again taper the kernel with a Hann window (Equation~\ref{eqn:Hann}) for compact support, and so the covariance kernel for the highly correlated residuals induced by a poor line fit becomes
 \begin{equation}
-  k_{\rm line}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma, r, r_0) = w(r,\, r_0)\, k_{\rm G}(\lambda_i, \lambda_j |\, a, \lambda_\mu, \sigma)
+  k_{\rm line}(\lambda_i, \lambda_j |\vtline{1} ) = w(r,\, r_0)\, k_{\rm G}(\lambda_i, \lambda_j | \vt_{{\rm line}_1})
   \label{eqn:line}
+\end{equation}
+To generate the covariance matrix, we now sum the effects of all the kernels, including the global covariance kernel, the specific line kernel, and the individual pixel noises
+\begin{equation}
+  k(\lambda_i, \lambda_j | \vtline{1}, \vtglobal, b) = k_{\rm line}(\lambda_i, \lambda_j |\vtline{1}) + k_{\rm global}(r | \vtglobal) + b\,\delta_{ij} \sigma_{ij}
+  \label{eqn:line_total}
 \end{equation}
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matrix_region_covariance.png}
-\includegraphics[width=0.4\textwidth]{figs/random_draw.png}
-\caption{\textbf{Left} a covariance matrix generated with the region kernel.
-\textbf{Right} general spectroscopic residuals near a mismatched stellar line
- overlaid with a sample draw using the covariance matrix, showing that the two
- look similar in structure and amplitude. 
-\protect \todo{Similar improvements to
-  Figure~\ref{fig:matern}.}}
+\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
+The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
+These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
+\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
+The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
 \label{fig:region}
@@
 
-\todo{Similar paragraph about visualizing the covariance matrix, drawing
-  samples. References to Figure~\ref{fig:region}.}
+In Figure~\ref{fig:region}, we see that the non-stationary Gaussian line kernel
+localizes the increased variance to a narrow range of the spectrum,
+corresponding to the residual from a specific poorly modeled spectral line.
+This increased covariance has the effect of down-weighting the influence of a
+specific spectral line on the likelihood function, the same way we might
+down-weight an outlier while fitting a straight line to a collection of
+data points.
 
-non-stationary kernel localizes increased variance to a specific
-    spectral line
+When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
+In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:line}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
+Along with the global covariance parameters and Chebyshev parameters for this order, we call the collection of nuisance hyperparameters for a specific order  (e.g., order $1$) $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$.
+Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.
 
-better than just masking out the region, since these contain info
+\todo{We can (should?) add an additional level to the hierarchy of
+  hyperparameters and add parameters to describe the \emph{population
+  characteristics} of poorly modeled spectral lines (mostly the typical width,
+  amplitude of these lines).  This will tell us about the frequency and
+distribution of spectral modelling errors.} 
+We explicitly fit for the hyperparemeters of the covariance kernels at the same
+time we fit for the stellar parameters. While including these extra
+parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
+regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
 
-here will be many sets of regions ($\sim$1-4\% of all lines might be
-    covered by a region)
+The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, it may be the case that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
 
-Covariance kernel parameters are parameters in our model that we explicitly sample for.
-There can be hyperparameters describing the \emph{population} of regions in a hierarchical model (mostly width, amplitude).
+Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
+This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
+However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
 
-The benefit mainly comes from simply modelling these systematic residuals to
- begin with, since a model with covariance is far more likely than forcing the
- fit, and far more justified than masking regions which do not fit.
-
-Where to draw the line between specific region and global covariance? When we reach some threshold, say three times the global amplitude?
+The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
 
 \paragraph{M dwarf residual}
-The squared exponential kernel is
-\todo{Why do we use the squared exponential and not \matern?}
+\todo{This third and final covariance kernel example might push this section length over the top.
+However, I think this application is the most exciting and could be the ``killer app'' that would drive people to adopt our framework given the computational overhead.}
+\todo{I think to properly design and test this kernel, I will need to actually explore fitting an M dwarf spectrum.
+For now I have just been looking at the figure by Andrew Mann.
+However, I have high hopes!}
+
+A powerful application of covariance kernels is that data-model mismatch can be accounted for even if it takes on a messy, not-easily-parameterized form like the Gaussian line example.
+We can not just inflate the noise in a specific region, but we also taper the kernel for more complicated structure.
+For example, consider a typical residual pattern from the spectroscopic fit of an M dwarf.
+There is a complicated pattern of high and low residuals, or there might even be a region of large offset, such that features are correlated on a large scale but not so well on the near scale.
+Therefore we need a kernel which can transition between these two.
+We multiply the existing kernel by the squared exponential kernel
 \begin{equation}
-  k_{ \rm exp}(r, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
+  k_{ \rm exp}(r |\, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
 \end{equation}
-$h$ is a bandwidth. 
-If $h$ is small, then there will be high-frequency structure. If $h$ is large,
- then only low-frequency structure will remain. 
-We also use the Hann window to taper the kernel for compact support
+where $h$ is a bandwidth. If $h$ is small, then there will be high-frequency
+structure. If $h$ is large, then only low-frequency structure will remain. 
 \begin{equation}
-  k(\lambda_i, \lambda_j | h, a, \mu, \sigma, r_0) = 
-   w(r, r_0)\; k_{ \rm exp}(r, h) \; 
-   k_{\rm G}(\lambda_i, \lambda_j | a, \mu, \sigma)
+  k(\lambda_i, \lambda_j | h, \vtline{0}) = 
+   k_{\rm exp}(r |\, h) \; 
+   k_{\rm line}(\lambda_i, \lambda_j | \vtline{0})
 \end{equation}
@@
 \todo{Show how the kernel can accounts for complicated M dwarf structure where
- there is large, messy mismatch.}
+  there is large, messy mismatch. Produce figures analogous to Figures~\ref{fig:matern}~and~\ref{fig:region}. Also add an example of M dwarf mismatch to Figure~\ref{fig:badlines}.}
 
@@
 \label{subsec:MCMC}
-Markov Chain Monte Carlo.
+We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}).
+The MCMC routine efficiently explores a high dimensional probability space using a stochastic, iterative approach.
+The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters.
+For more on MCMC and Gibbs samplers see \citet[ch 11]{gcs+13} and the references therein.
 
-Gibbs sampler.
-Sparse matrix.
-Implemented using SuiteSparse and the Cholesky factorization, means that the RCR step can be computed quickly and reused.
+In addition to the stellar parameters $\vtstar$, we have introduced several nuisance parameters for calibration and residual modeling.
+These nuisance parameters have a logical hierarchical structure.
+At the lowest level of the hierarchy, a single order of an echelle spectrum has $N$ regions of high covariance $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
+If we wish to evaluate the probability of the parameters of a specific line (e.g., line $1$) conditional on the current values of all other lines, then we denote this by $p(\vt_{\textrm{line} = 1} | \vt_{\textrm{lines} \ne 1})$.
+The collection of nuisance parameters for a specific order  (e.g., order $1$) is the aggregate parameter $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$.
+Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.
+Because the nuisance parameters for a single echelle order are independent from the nuisance parameters for any other echelle order, we have $p(\vt_{\textrm{order} = 1} | \vt_{\textrm{orders} \ne 1}) = p(\vt_{\textrm{order} = 1})$.
 
-Show figure using dfm/triangle.py that shows recovered stellar parameters and degeneracies. Separately for the regions.
+\begin{figure*}[!htb]
+\begin{minipage}{\textwidth}
+\begin{center}
+  \includegraphics[width=5in]{figs/stellar_triangle.png}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+These stellar parameters are marginalized over the Chebyshev and noise parameters.
+Made with \texttt{triangle.py}\protect\footnotemark
+  \protect\todo{In the updated figure I won't show $v_z$ or $\Omega$ so the font is easier to see. Many more samples so posteriors look smooth.}
+ }
+\label{fig:stellar_posterior}
+\end{center}
+\footnotetext{\url{https://github.com/dfm/triangle.py}}
+\end{minipage}
+\end{figure*}
+
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
+  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
+  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
+\textbf{Right}: The posterior probability function of the line parameters, marginalized over the stellar and global noise parameters.}
+\label{fig:noise_posterior}
+\end{center}
+\end{figure*}
+
+The Gibbs sampler iteratively explores this hierarchy of parameters.
+At each level of the hierarchy, corresponding to a different subset of nuisance parameters, we use the Metropolis-Hastings algorithm to propose a new subset of parameters and then either accept or reject them.
+This process of proposal and acceptance/rejection is called ``sampling.'' The Gibbs sampler rotates between sampling in different subsets of parameters, eventually sampling all of the nuisance parameters at a certain cadence.
+To initialize the MCMC algorithm, we make a reasonable guess for the starting parameters.
+We use existing spectral types in the literature for the stellar parameters, flat Chebyshev polynomials (no flux-calibration correction), and no covariance structure in the residuals (the global covariance parameters are set to zero and no line kernels are instantiated).
+We use superscripts to denote iterations of the MCMC algorithm.
+$i$ denotes the parameters from the current iteration and $i - 1$ denotes the previous iteration.
+The Gibbs sampler rotates among subsets of the parameters as follows
+
+\begin{enumerate}
+  \item Sample in the stellar parameters.
+For each proposal of the Metropolis-Hastings algorithm, generate a model spectrum following the steps in \S\ref{subsec:postprocess}.
+When deciding whether or not to accept the parameters, the Gibbs sampler evaluates $p(\vtstar^i | \vtorders^{i-1})$.
+  \item For each order of the echelle spectrum
+    \begin{enumerate}
+      \item Sample in the Chebyshev polynomial parameters.
+Adjust the spectrum as detailed in \S\ref{subsec:postprocess}.Gibbs sampler evaluates $p(\vtcheb^{i} | \vtstar^{i}, \vtglobal^{i -1}, \vtlines^{i-1})$.
+      \item Sample in the global covariance parameters.
+Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}.
+Gibbs sampler evaluates $p(\vtglobal^{i}| \vtstar^{i}, \vtcheb^{i}, \vtlines^{i -1})$.
+
+      \item Check to see whether the algorithm should instantiate/delete new/old line kernels
+      \item For each line kernel $k$
+	\begin{enumerate}
+	  \item Sample in the parameters for each line, conditional on the parameters for all the other lines.
+Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}.
+Gibbs sampler evaluates $p(\vtline{k}^i| \vt_{\textrm{lines}\ne k}^{i - 1})$.
+	\end{enumerate}
+    \end{enumerate}
+\end{enumerate}
+
+At each stage of the Gibbs sampler, the likelihood function (Equation~\ref{eqn:lnprob}) is evaluated for a proposal of a particular subset of parameters, conditional on the current values of all the other parameters.
+This hierarchical parameter structure and the kernel parameterization of the covariance matrix influences how the algorithm may be efficiently implemented in code.
+For a typical optical spectrum with $\gtrsim$1,000 pixels, the most computationally intensive step of the likelihood evaluation is usually the matrix product $\fR^T C^{-1} \fR$.
+Since we have designed the covariance matrix to be sparse, we can use optimized sparse matrix algorithms which are much faster and memory efficient than dense matrix operations.
+Because we are not interested in the matrix inverse $C^{-1}$ by itself, but rather the product with the residual vectors, we can use efficient routines for solving linear systems to bypass the computationally difficult step of matrix inversion.
+Additionally, because the covariance matrix is positive semi-definite, we can use the Cholesky factorization of the matrix to optimize the evaluation of the matrix product.
+Once the covariance matrix is factorized, any subsequent evaluation of the matrix product for different residual vectors $\fR$ is extremely rapid.
+This makes the $\vtstar$ and $\vtcheb$ steps of the Gibbs sampler extremely fast.
+When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update.
+However, because we designed the kernels to deliver a sparse matrix these operations are efficient enough to be used with an MCMC algorithm.
+We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{cdh+08, dh09} and the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{fhl+12}.
+
+We run the MCMC Gibbs sampler for many iterations until the estimate of the posterior distribution has converged.
+To check that the chain is not stuck in a local maximum of the posterior, we redo the MCMC run many times with different starting parameters, to ensure that the algorithm converges to the same global maximum.
+A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
+Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
+This enables us to present a posterior of the stellar parameters $\vtstar$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
+This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
+
+The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
+If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and then the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
+This leads to estimates of stellar parameters with more uncertainty than necessary.
+\todo{Since this point is better carried with an example or test using the global covariance structure and not using it (see Test 1), we might want to defer this discussion to the testing section.}
 
@@
 \label{subsec:learning}
-Learnt covariance structure can be used to correct the models.
+By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
+After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
+The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
+Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
+
+Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
+Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
+Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
+Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
+In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
+This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kru10}.
+
+Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct defects in the synthetic models.
+\todo{more delicate phrasing?} Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
+This fundamental application of machine learning would enable us to create our own library of semi-empirical stellar models.
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
 
 \section{Tests}
-\todo{Which tests to show?}
+Here are some ideas of which tests we might want to show.
+
+\paragraph{Test 1}: Fitting with and without the global covariance kernel to show how the width of the posteriors nicely inflates to reasonable uncertainties.
+
+\paragraph{Test 2}: I think we should fit both an F/G star (WASP-14) and a K/M star (TBD) with both the Kurucz and PHOENIX spectra. This will show a few things
+\begin{itemize}
+  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
+  \item Both spectral libraries have stars that they perform better and worse on.
+  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
+\end{itemize}
 
@@
   & are applied during ``post processing'' of the synthetic spectrum\\
-$\vstar$ & $\{\vg,\vpp \}$\\
+$\vtstar$ & $\{\vg,\vpp \}$\\
 $\finst(\lambda)$ & data spectrum\\
 $\fsynth(\lambda)$ & synthetic spectrum\\
-$c_0, c_1, \ldots$ & Chebyshev polynomial coefficients for residual flux calibration\\
-$\vN$ & the set of nuisance parameters composed of $\{c_0, c_1, \ldots, c_N\}$\\
+$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
+$\vtline{1}$ & \\
+$\vtlines$ & \\
+$\vtorder{1}$ & \\
+$\vtorders$ & \\
 $\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
 $\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
-$\fD$ & data vector comprised of all $\fDi$, $i = \{0, \ldots, N\}$\\
+$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
 $\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
-$\fM$ & model vector comprised of all $\fMi$, $i = \{0, \ldots, N\}$\\
+$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
 $\sigma_i$ & Poisson noise for a given pixel $i$\\
@@
 $C$ & covariance matrix\\
+$\sigma_{ij}$ & element in the covariance matrix\\
+$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
 $k_{\rm global}$ & global covariance kernel\\
-$k_{\rm region}$  & regional covariance kernel\\
+$k_{\rm line}$ & regional covariance kernel\\
 \hline

### 2014-06-14 tex/paper.tex

@@
-\documentclass[preprint]{aastex} %double-column, single-spaced document:
+\documentclass[preprint]{aastex} 
 %\documentclass[iop,floatfix]{emulateapj}
@@
 \usepackage{graphicx}
-\usepackage{apjfonts}
+%\usepackage{apjfonts}
 \usepackage{enumerate}
@@
 \newcommand{\vt}{ {\bm \theta}}
+\newcommand{\vT}{ {\bm \Theta}}
 \newcommand{\vg}{\vt_{\star, {\rm grid}}}
@@
 \newcommand{\vtcov}{\vt_{\rm cov}}
+\newcommand{\vM}{\mathsf{M}}
+\newcommand{\vD}{\mathsf{D}}
+\newcommand{\vR}{\mathsf{R}}
+\newcommand{\vC}{\mathsf{C}}
 \newcommand{\fM}{ \vec{{\bm M}}}
@@
 
-\title{A Method for the spectroscopic inference of fundamental stellar parameters}
-\author{\today{}\\
-\medskip
-Ian~Czekala\altaffilmark{1} et al.
-%Author2\altaffilmark{2},
-}
+\slugcomment{draft: \today{}}
 
-\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street 
- MS 10, Cambridge, MA 02138}
-%\altaffiltext{2}{Institution 2}
-\email{iczekala@cfa.harvard.edu}
+\title{A Method for the Spectroscopic Inference of Stellar Parameters}
+\author{Ian Czekala, Sean M. Andrews, et al.}
+\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
+%\email{iczekala@cfa.harvard.edu}
 
-\section{Introduction}
-\begin{itemize}
-  \item Where our technique fits in the ecosystem of stellar methods.
-  \item Applicable to any kind of spectrum (long-slit, echelle, infrared,
-    flux-calibrated or not)
-  \item Examples of fields where accurate and unbiased stellar parameters are
-    crucial. Exoplanets. T Tauri stars.
-\end{itemize}
+\section{Introduction \label{sec:intro}}
 
-\section{Method}
+\comm{The following are some suggested paragraph outlines for a concise introduction:}
 
-A single spectrum provides a wealth of information that can be used to infer the fundamental properties of a star: effective temperature $T_{\rm eff}$, surface gravity $\log g$, and metallicity $\Z$; together $\vg$.
-We wish to extract the maximal amount information from a given spectrum while also recognizing the limitations imposed by our stellar model and exposing any degeneracies between the parameters.
-For example, if we model a large range of the spectrum at once (e.g., the full optical range at high resolution), then we must explicitly account for any covariances and biases introduced by improper calibration or inaccuracies in the model spectrum.
-We present a generative model which uses a synthetic stellar library as a backend and apply post-processing techniques to mimic the observed spectrum as a function of several stellar parameters and calibration parameters in order to determine the best fit.
-By forward modelling the data spectrum in a Bayesian framework, our modular approach can easily incorporate additional calibration artifacts and stellar complexity (e.g., veiling or Zeeman splitting) while still being true to the limitations of the dataset.
-This data driven approach will enable us to learn how the underlying synthetic models may be improved and refined in an iterative manner.
+\comm{(1) start general: vast amount of information in an astronomical spectrum, and we want to 
+make the most of that to learn about various physical parameters of interest.  accurate, unbiased 
+stellar parameters (like temperature, gravity, metallicity, etc.) are of crucial importance in many 
+astrophysics problems; e.g., exoplanets (and explain why), etc. (pick another one or two).}
 
-The roadmap for our method is as follows.
- First, for a given set of fundamental stellar parameters $\vg$, we generate a synthetic
- spectrum by interpolating from a grid of synthetic spectra (\S\ref{subsec:synthetic}). 
-Then, we post-process the interpolated spectrum for realism by using additional
- parameters to describe Doppler shift, projected rotational velocity,
- interstellar extinction, and residual flux calibration errors
- (\S\ref{subsec:postprocess}). 
-We perform a pixel-by-pixel comparison of the generated spectrum with the
- data spectrum using a prescribed likelihood function
- (\S\ref{subsec:likelihood}) and a parametric treatment of the covariance
- between pixel residuals (\S\ref{subsec:covariance}). 
-We use a Markov chain Monte Carlo (MCMC) Gibbs sampler to numerically explore
-the posterior probability density and determine constraints on the stellar
-and nuisance parameters (\S\ref{subsec:MCMC}). 
-The benefit of our method is that it provides robust estimates of the stellar and nuisance
-parameters with well-quantified uncertainties and parameter degeneracies. 
-Our parameterization of the covariance between pixel residuals also yields a means to identify
- and quantify the key discrepancies between the data and model, and provides a path forward to 
- improve the stellar models in an iterative fashion (\S\ref{subsec:learning}).
- \todo{I really want to overload the word ``covariance'' and use it to describe both the covariance between $T_{\rm eff}$ and $\log g$ as well as the covariance structure of the pixel residuals themselves.
-However I think this might be confusing so I am using ``degeneracies'' for the former and reserving ``covariance'' for the later.}
+\comm{(2) point out the challenge: quantitative extraction of these physical parameters is not 
+easy.  go through some of the basic underlying limitations (practical, computational, limited 
+models, etc.)}
 
-\subsection{Synthetic Model Spectra}
-\label{subsec:synthetic}
-Many methods exist to generate a synthetic model of a stellar spectrum. Generally, when
-a high quality spectrum is needed for a specific set of stellar parameters, one
-runs a stellar atmosphere and radiative transfer code to synthesize the
-spectrum. Unfortunately, it is computationally expensive to accurately generate
-a spectrum over a wide range of wavelengths, and so these codes are generally
-employed to generate libraries of stellar spectra, which exist over a range of
-$\vg$ at specified points on a regular grid, typically covering the main sequence of stellar
-types. Several high-quality libraries of synthetic spectra now exist
-(\citealt{hwd+13}, Kurucz, and more are being developed for GAIA).
+\comm{(3) highlight various techniques/approaches being used (lots of references here).}
 
-In a forward-modelling approach, which may make many calls to generate a
-spectra of varying $\vg$, it is currently computationally impractical to
-iterate the structure of the stellar atmosphere and then run the stellar
-synthesis for each call of $\vg$.  Some methods compromise by taking a halfway
-approach, whereby a precomputed library of stellar atmospheres is interpolated
-for a given $\vg$, and then a simple radiative transfer code is run to
-synthesize a spectrum (e.g.  SME; \citealt{vp96}). This technique has the
-advantage that interpolating atmospheres and then running the spectral
-synthesis for each $\vg$ might be more accurate than directly interpolating
-synthetic spectra. However, if a large region of wavelength is required at high
-resolution, then the synthesis step will be slow.  A different approach might
-be instead to eschew forward modelling (and thus repeated synthesis or grid
-interpolation) of the spectra entirely and evaluate the posterior probability
-density only at the grid points of the spectral library.  Then, these discrete
-samples of the posterior density can be interpolated to find the posterior
-surface and draw confidence intervals on the parameters (e.g. SPC;
-\citealt{blj+12}). However, the difficulty with this approach is that the
-parameter uncertainty might be smaller than the grid spacing, and then
-interpolation will not properly capture degeneracies between parameters.
+\comm{(4) now, without criticizing those approaches, just launch into a brief summary of our goals 
+and a basic outline of the paper.}
 
-Because we desire a generative model of the spectra without the added
-complexity or computational overhead of running a stellar synthesis, we opt to
-use a \todo{Implement a more accurate but still fast interpolator for the grid.
-Likely this will be a combination of using splines to pre-interpolate a finer
-grid (~20 K, 0.1 dex spacing in gravity and metallicity and then doing linear
-interpolation off of this new grid during the actual MCMC run. It will be described here.} However, the
-techniques we develop in this paper to account for and parameterize the
-correlated, systematic pixel residuals are generally applicable to spectra fit
-by any of these methods.
+\comm{I have made a lot of edits in Section 2, but have tried to restrict comments.  When there is 
+something useful to say, I have added some commented-out text in the .tex file.}
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=0.85\textwidth]{figs/model_data.pdf}
-  \caption{\textbf{Top}: Chebyshev polynomial which multiplies the model spectrum
-   to account for inaccuracies in the flux-calibration. 
-  \textbf{Middle}: The data spectrum and model spectrum, after it has been
-   interpolated, post-processed, and multiplied by the Chebyshev polynomial.
-  \textbf{Bottom}: Residuals from the model fit. Note the large residual at
-   $\lambda5189$\AA\ due to a missing opacity source from \ion{Ca}{2}.}
-\label{fig:model_data}
-\end{center}
-\end{figure*}
+\comm{Important: Note how I have changed the naming convention in the .bib file, so that a paper is 
+referenced by the first author's last name and the year, like czekala14.  Your haphazard formatting 
+was driving me nuts, and you can't find anything in the bib file (plus, if you read inside the .tex 
+file, my approach makes things a hell of a lot easier).}
 
-\subsection{Post-Processing}
-\label{subsec:postprocess}
+\section{Methodology \label{sec:method}}
 
-Libraries of synthetic spectra are generally computed at high resolution ($R
- \geq $100,000), sampled with many pixels per resolution element, and do not
- include any rotational broadening or account for any instrumental effects. 
-In order to transform a raw synthetic spectrum interpolated from the grid into
- one that matches the spectrum of a real star, we must post-process the spectrum
- to account for these secondary effects. 
-In addition to these fundamental parameters, a star also has several observed
- properties that are a function of its kinematics, geometry, and location in our
- galaxy: projected rotational velocity $v \sin i$, line of sight velocity $v_z$,
- total extinction $A_V$, and solid angle $\Omega$. 
-We call these secondary parameters $\vpp$, because we can model their effects
- on the synthetic spectra in a ``post-processing'' step by convolution with an
- appropriate kernel or multiplication with a smooth function. 
-Together, we call $\vtstar = \{\vg, \vpp\}$.
+All astronomers recognize that spectroscopy offers a wealth of information that can help 
+characterize the fundamental properties of the observing target.  However, as noted in Section 
+\ref{sec:intro}, the reliable and statistically robust inference of those physical (or empirical) 
+parameters from an observed spectrum can be extraordinarily challenging.  Here we describe a 
+generative Bayesian modeling framework that confronts some of the key obstacles in that process.  
+The goal of this approach is to conservatively extract the maximal amount of information about a 
+prescribed (and degenerate) parameter set by forward-modeling an observed spectrum, while also 
+recognizing and explicitly accounting for the covariances and biases introduced by pathologically 
+imperfect models or calibrations.  The method is modular, and therefore can easily incorporate 
+additional physical or nuisance parameters as desired without sacrificing an accurate reflection of 
+the limitations in the data.  Moreover, with a well-crafted observational sample, this data-driven 
+approach should ultimately enable us to systematically learn how synthetic spectral models can be 
+improved.  The specific applications discussed here are related to the spectra of individual stars, 
+but the methodology is generic (and could be used for composite spectra of unresolved stellar 
+clusters, galaxies, etc.).
 
-The projected rotational velocity of the star, parameterized by $v \sin i$,
-broadens spectral line profiles and is mathematically described by convolution
-with a specific kernel \citep[Equation 18.14]{gra08}. UV, optical, and infrared
-spectra are acquired using a spectrograph which disperses light onto a CCD,
-defining a specific resolution and sampling rate for the spectrum.  Because the
-resolution of the spectrograph and pixel sampling are different from the
-synthetic spectral library, we must convolve the raw spectrum with the line
-spread function (LSF) of the instrument and resample to the exact pixels of the
-CCD. In wavelength space, the $v \sin i$ and LSF operations are represented by
-the convolution of the synthetic spectrum with these two kernels.  Using the
-convolution theorem, we can rewrite these convolutions as multiplications in
-Fourier space.  In order to use the fast Fourier transform (FFT) with discrete
-samples of the synthetic spectrum, we must first resample it
-onto a uniform grid.  In the case of a spectrum, the natural
-uniform grid is one that is equally sampled in velocity space, such there is an
-equal velocity shift $\Delta v$ between pixels. This results in a wavelength
-grid that is linearly sampled in the logarithm of wavelength.  Next, we
-multiply the Fourier transform of the synthetic spectrum by the Fourier
-transforms of the rotational velocity and line spread function kernels.
-Lastly, we do an inverse FFT to transform the modified spectrum back to
-wavelength space, where it is sampled at the wavelength locations of the pixels
-in the detector.  When resampling the spectrum, we are careful to ensure
-band-limited interpolation using splines in order to prevent introducing
-non-physical high frequency structure into the spectrum.
+The remainder of this section describes the mechanics of this modeling framework.  First, a model 
+spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
+then post-processed to mimic reality using a set of observational and practical nuisance parameters 
+(Section \ref{subsec:postprocess}).  Next, a direct, pixel-by-pixel comparison between the data and 
+model spectra is made with a prescribed likelihood function and a parametric treatment of the 
+covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
+under the guise of hierarchical Monte Carlo Markov Chain (MCMC) simulations to numerically explore 
+the posterior probability density of the model conditioned on the data, and thereby to determine 
+constraints on the parameters of interest (Section \ref{subsec:MCMC}).  Along the way, these 
+procedures are illustrated with real observations of the high resolution optical spectrum from a 
+nearby F star.  That specific application, along with some alternative demonstrations of the 
+method, are discussed in more detail in Section \ref{sec:examples}.
 
-We scale the raw synthetic flux (often reported at the stellar surface) to an
-observed flux by multiplying by the solid angle of the star, $\Omega =
-R_\star^2/d^2$, where $R_\star$ is the stellar radius and $d$ is the distance
-to the star from earth. Then, we correct for a radial velocity difference $v_z$ between the
-star and the earth by Doppler shifting the model spectrum.  Finally, we correct
-for interstellar extinction by dividing by a dust extinction model
-$A_\lambda$, which is parameterized by extinction in the $V$ photometric band,
-$A_V$. 
-\todo{We can add in convolution with photometric filter
-profiles to use photometry and provide a better constraint on the luminosity of
-the star}
-\todo{We might try implementing the convolution in wavelength space, rather
- than the FFT. 
-This would allow the LSF to change with wavelength.\\}
+\subsection{Generating a Model Spectrum \label{subsec:synthetic}}
 
-\paragraph{Calibration polynomials} In contrast to a collection of photometric points in a spectral energy distribution, or any ensemble of data for which each data point is acquired by an individual measurement, adjacent wavelength points in a spectrum are measured simultaneously, with the same detector and--more importantly--the same calibration.
-Typically, spectra are flux calibrated by dividing by a ``sensitivity function,'' which is determined by comparing an observation of a spectrophotometric standard to a previously calibrated model spectrum.
-Therefore, any error in the sensitivity function, which is typically a smooth function with wavelength, will create a region of the data spectrum which is systematically offset by a small amount from what the true flux should be.
-The flux calibration error is then the residual difference in the empirically determined sensitivity function from the what the ``true'' sensitivity function would be if we had a perfectly calibrated instrument and spectrophotometric standard.
-Therefore, it is necessary to introduce additional degrees of freedom into our model in order to account for the calibration error and allow the model spectrum to properly match the data.
-By adding the residual calibration as a parameter in our model, we can explore the distribution of possible calibrations (\S\ref{subsec:MCMC}) and then marginalize the calibration parameters out of the final posterior distribution of the stellar parameters.
-This means that we can derive estimates of stellar parameters that properly account for the intrinsic uncertainty in the flux calibration.
+There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
+parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
+subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
+However, in general this approach is still computationally prohibitive for any iterative method of 
+probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
+structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
+grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
+that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{vp96}.  A more 
+common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
+$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
+technically the former approach is most similar to the ideal case, the computational cost of 
+repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
+(particularly for data with a large spectral range) considerably less appealing.  A related, but 
+different approach is to eschew forward modeling entirely (and therefore repeated spectral 
+syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
+points of the library.  Then, these discretized samples of the posterior probability density can be 
+interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
+\citep[e.g., the method of SPC;][]{blj+12}.  The difficulty with this latter approach is that the 
+parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
+that this interpolation might not accurately recover intrinsic parameter degeneracies.
 
-A traditional way to deal with errors in the flux calibration is to avoid flux-calibration entirely and simply normalize both the data and the spectrum to the continuum level before comparison.
-While this procedure may work well for hotter stars with well-defined continuua, normalizing spectra of cooler stars with large molecular features may introduce artificial features into the spectrum when the pseudo-continuum is incorrectly placed.
-Instead, we choose to parameterize the residual flux calibration error by a low-order Chebyshev polynomial which is normalized to unity.
-The coefficients of the polynomial are parameters in the model that will be determined by the fitting algorithm, $\vtcheb = \{c_0, c_1, \ldots, c_N \}$.
-If we have multiple observations of spectrophotometric standards, then we can characterize the typical flux-calibration error of the spectrograph and put priors on the coefficients that will prevent overfitting of the polynomials.
-For an example of the Chebyshev polynomial determined for a typical spectrum, see Figure~\ref{fig:model_data}.
-Using priors on the Chebyshev coefficients ensures that real spectral features, such as deep molecular bands in later type stars, are not accidentally treated as calibration artifacts.
-The overall normalization of the polynomial $c_0$ is degenerate with the solid angle of the star $\Omega$.
-Therefore we apply an additional constraint and enforce that the mean of the Chebyshev polynomial be equal to $1$.
-If there is only one order in the spectrum, such as with a long-slit spectrum or a single order of an echelle spectrum, then this means setting $c_0 = 1$.
-If we are fitting more than one order of an echelle spectrum then we adopt an independent Chebyshev polynomial for each order.
-This means that we chose the $c_0$ of a specific order to anchor to 1, while the $c_0$ for other orders are allowed to take on perturbations about $1$.
+Here we opt to take the computationally expedient approach that employs a library of model spectra, 
+$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
+(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
+it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
+generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
+$\vt_{\ast}$ must be interpolated from among the spectral library, 
+\begin{equation}
+f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
+\label{eqn:interp}
+\end{equation}
+where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
+interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
+is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
+undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
+can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
+calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
+with the corresponding library spectrum.  After an extensive exploration of such calculations 
+\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
+achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
+a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
+tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
+be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
+could be avoided if the spectral library was computed over a refined grid (with a substantial 
+up-front computational investment); but for the time being, we can empirically propagate these 
+interpolation uncertainties into the likelihood calculations (as will be described in 
+Section~\ref{subsec:likelihood}).
 
-This polynomial formalism can in principle be used to model raw (not flux calibrated) spectra by using the Chebyshev polynomial to solve directly for the sensitivity function itself, rather than a residual calibration error as is currently done.
-When used in this manner, there will be greater uncertainty about the properties of the spectrograph, which will translate into wider priors on the Chebyshev coefficients.
-In this case, there is more of a chance that the polynomial might destroy broad-scale information present in the spectrum by fitting it out, something that would be avoided with more accurate priors provided by a flux calibration.
+\subsection{Post-Processing \label{subsec:postprocess}}
 
-\subsection{Likelihood Calculation}
-\label{subsec:likelihood}
+Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
+to a typical observed spectrum, and does not account for several additional observational and 
+instrumental effects that become important in comparisons with real data.  Therefore, a certain 
+amount of post-processing is required before assessing the model quality.  We treat that 
+post-processing in two stages: the first deals with an additional set of ``observational" 
+parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
+location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
+n}$, designed to mitigate an imperfect data calibration.
 
-To evaluate which parameters of the model spectrum fit the data spectrum best,
-we use a likelihood function to compare individual pixels between the two
-spectra. A likelihood function provides the probability of the data spectrum
-$\fD$, given the model spectrum $\fM(\vt)$, which is parameterized by the
-stellar and calibration parameters $\vt$. The data spectrum and model spectrum are both composed of $N$ pixels, and so $\fD$ and $\fM$ represent the collection of these pixels in vector form, 
-$\fD = \{\fDi\}^N_{i=1}$ and $\fM = \{\fMi\}^N_{i=1} $, respectively. The vector of residuals is the difference between the data spectrum and the model spectrum
+We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
+spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
+former, we consider three kernels that contribute to the observed line-of-sight velocity 
+distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
+spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
+with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
+sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
+broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
+\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
+incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
+parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
+a convolution in velocity-space,
 \begin{equation}
-  \fR(\vt) = \fD - \fM(\vt)
+f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
+\label{eqn:broadening}
 \end{equation}
-In the language of probability calculus, the likelihood function is written as
-$p(\fD | \fM(\vt))$, or $p(\fD | \vt)$ for short. We choose a standard
-multidimensional Gaussian likelihood function which allows for covariance
-between the pixel residuals 
+and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
 \begin{equation}
-  p(\fD | \vt) = \frac{1}{\sqrt{(2 \pi)^N \det(C)}} \exp\left ( -\frac{1}{2}
-   \fR^T C^{-1} \fR \right ) 
+f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
+\label{eqn:resampling}
+\end{equation}
+where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
+$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
+For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
+post-processing steps.
+% SA: My idea for this figure is sort of 1-column wide (~3.5"), with a set of spectral panels going 
+% down vertically.  The top panel shows a segment of the full resolution model.  The next one down 
+% can show the individual kernels described above, as well as their composite, \varphi_v.  Then the 
+% next panel down can show the convolved (but still full-resolution) spectrum.  Finally, the bottom 
+% panel will show the spectrum re-sampled onto the observed pixels.  It need not be a large figure 
+% (maybe 5" long?).
+
+At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
+is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
+diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
+radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
+account for interstellar extinction, assuming a 
+derived extinction law $A_{\lambda}$ \citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  
+The parameters $\vt_{{\rm obs}, f} = [\Omega, \,\, A_V]$ are applied as
+\begin{equation}
+\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
+\label{eqn:scaling}
+\end{equation}
+where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.
+
+So far, the procedure summarized in Eq.~\ref{eqn:interp}-\ref{eqn:scaling} is composed of 
+relatively straightforward operations demanded by practical astronomical and computing issues.  If 
+the observed spectrum was {\it perfectly} calibrated, we could proceed to a likelihood calculation 
+(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vt)$ at this point.  
+However, that is unlikely to be the case.  The primary concern is that an imperfect calibration 
+produces (presumably low-level) mismatches in the underlying shape of the observed spectrum on 
+relatively broad wavelength scales.  When compared with the model, these mismatches might represent 
+a non-trivial contribution to the likelihood, and thereby bias our estimates of the desired 
+physical parameters.  Often, this concern is treated externally to any modeling procedure, usually 
+by dividing the observed spectrum (and model spectrum) by an appropriate polynomial function.  But 
+that ``normalization" procedure implicitly assumes that there is no relevant information content on 
+those mismatched scales; if $\vt$ also contributes to the broad spectral shape, then adopting this 
+approach will corrupt the inferences of these parameters.  Moreover, in practice this approach is 
+limited, since defining an appropriate polynomial becomes difficult in cases where the spectral 
+line density is high (e.g., molecular bands for cool stars).
+
+We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
+modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
+freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
+the model spectrum, enabling us to explore the distribution of possible calibrations (see 
+Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
+In essence, the inferences for the relevant stellar parameters will properly account for the 
+underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
+model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{elh+06,koleva09}, 
+\begin{equation}
+\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
+\end{equation}
+where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
+set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
+spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
+unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
+artifacts is negligible (see Secton~\ref{sec:examples} for examples).  The lowest-order (scaling) 
+coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
+Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
+unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
+model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
+the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
+formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
+rather than the residual calibration mismatches as described here (with a suitable relaxation of 
+the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:cheb} offers a practical demonstration of how 
+these nuisance parameters are applied. 
+% SA: I envision this figure much like the previous one, in a 1-column vertical distribution of 
+% panels.  At top, you would show the data spectrum at an overlap of two spectral orders, along 
+% with the post-processed model with no Chebyshev polynomials applied.  Next, you would show the 
+% two Chebyshev polynomials appropriate for the two orders.  And then in the final panel you would 
+% revisit the top panel with the polynomials applied, demonstrating the improvement.
+
+\subsection{Model Evaluation \label{subsec:likelihood}}
+
+The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
+likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
+$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
+\begin{equation}
+\vR \equiv \vR(\vt, \vT_{\rm cheb}) \equiv \vD-\vM(\vt, \vT_{\rm cheb}).
+\end{equation}
+To quantify the probability of the data conditioned on the model, we adopt a standard 
+multi-dimensional Gaussian likelihood function,
+\begin{equation}
+p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
+   \vR^\trans \vC^{-1} \vR \right ), 
    \label{eqn:prob}
 \end{equation}
-which penalizes models that yield larger residuals. $C$ is the covariance
-matrix which describes the covariance between the pixel residuals. The structure of the covariance matrix is discussed in detail in the next section (\S\ref{subsec:covariance}). Because
-the likelihood function generally exhibits a large dynamic range, it is numerically
-preferred to use the log-likelihood function
+which penalizes models that yield larger residuals and explicitly allows for covariances in the 
+residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
+numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
+this case is
 \begin{equation}
-  \ln[p(\fD | \vt)] = -\frac{1}{2} \fR^T C^{-1} \fR - \frac{1}{2} \ln \det C  
-   - \frac{N}{2} \ln 2 \pi 
+  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
   \label{eqn:lnprob}
@@
 
-\subsection{Parameterizing the covariance structure}
-\label{subsec:covariance}
-The covariance matrix $C$ describes the measurement noise in each pixel of the spectrum as well as the covariance between pixel residuals.
-It is fundamental to the likelihood calculation described in \S\ref{subsec:likelihood}.
-For most problems, such as fitting a spectral energy distribution to photometric measurements, each data point is acquired in an independent measurement from the other data points, which means that the noise for each data point, $\sigma_i$, is also independent from the other data points.
-This results in a diagonal covariance matrix
+The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; 
+``noise") in each pixel and the intrinsic covariance between pixels.  The special case where each 
+pixel represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
+\delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
+the Kronecker delta function, and Eq.~\ref{eqn:lnprob} reduces to the familiar
 \begin{equation}
-  C = 
-  \begin{bmatrix}
-    \sigma_0^2 & 0  & \cdots & 0\\
-    0 & \sigma_1^2 & \cdots & 0\\
-    \vdots  &   & \ddots  & \vdots \\
-    0 & 0 & \cdots & \sigma_N^2\\
-  \end{bmatrix}
-  \label{eqn:covariance_diagonal}
+\ln{p(\vD | \vM)} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
+\label{eqn:chisq}
 \end{equation}
-with $\sigma_{ij} = \delta_{ij} \sigma_{ij}$, where $\delta_{ij}$ is the Kronecker delta function. When the covariance matrix is diagonal, 
-the likelihood function (Equation~\ref{eqn:lnprob}) reduces to the familiar
-$\chi^2$ form of a sum over the square of the residuals, weighted by the
-inverse variance of each data point
+the sum of the square of the residuals weighted by the inverse variances (squared uncertainties).  
+However, the problem being addressed here necessitates the use of a more complex covariance matrix; 
+additional off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances 
+imposed by the discrete over-sampling of the line-spread function, and (2) highly correlated 
+residuals as manifestations of the still-imperfect model library are required to avoid biasing our 
+inferences of the physically interesting parameters ($\vt$).  The following subsections describe 
+how these issues are addressed in the practical implementation of $\vC$.
+
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics{figs/badlines.pdf}
+%  \caption{A collection of spectral lines which have imperfect model fits.
+%    From left to right: \textbf{Class 0} The majority of spectral lines
+%    ($\gtrsim 60$\%) will have minor differences in strength between data and
+%    model spectrum, which produce low-amplitude correlations in the residuals
+%    on the length scale of the width of a typical spectral line.  \textbf{Class
+%    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+%    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+%    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+%    of all lines), an extraneous line in the model leaves a large, highly
+%    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
+%    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+%    correlated residuals of moderate amplitude.  The difficulty
+%    with class III lines is that for any specific line, there might exist a
+%    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
+%    will properly fit \emph{all} the lines.}
+%\label{fig:badlines}
+%\end{center}
+%\end{figure*}
+
+\subsubsection{Global Covariance Structure}
+
+Astronomical spectrographs are designed so that the detector over-samples the instrumental line 
+broadening function with at least a few pixels.  Therefore, adjacent pixels are never completely 
+independent samples of the observed spectrum.  In that case, a difference between an observed and 
+modeled spectral feature will create a residual that spans multiple pixels.  This can be 
+demonstrated in practice by examining the autocorrelation of the residual spectrum: a slight model 
+mismatch will produce correlated residuals over a characteristic scale similar to the observational 
+line-broadening kernel width.  Figure \ref{fig:class0} highlights a specific example of these 
+correlated residuals in real data, where an imperfect model generates residuals which exhibit a 
+significant autocorrelation signal on the scale of $\sim$4 pixels, which corresponds to the typical 
+line width in this spectrum.  
+% SA: It may be true that the figure you've made to demonstrate this is accurate, but it certainly 
+% does not seem obvious to me what is going on in it.  I like the idea of the figure, but I think 
+% it would be better demonstrated on a relatively smooth (continuum) region with a small spectral 
+% mismatch, rather than deep lines everywhere.  Can we do that instead?
+
+%\begin{figure*}[!htb]
+%\begin{center}
+%  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
+%  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
+%  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+%  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
+%\label{fig:class0}
+%\end{center}
+%\end{figure*}
+
+It seems important to distinguish here between ``noise" and the fit residuals.  Noise introduced to 
+the spectrograph by astrophysical or instrumental effects is generally uncorrelated with 
+wavelength.  The arrival of each photon to the detector is an independent event; while these 
+photons are scattered by the instrumental line-spread function, the magnitude and direction of that 
+scatter is independent for each such event.  In essence, the noise itself is uncorrelated, but the 
+fit residuals likely are correlated.  However, from a mathematical perspective the correlated 
+residuals can be treated in the same way as correlated noise, by constructing a non-trivial 
+covariance matrix with off-diagonal terms.  In practice, this is achieved by parameterizing $\vC$ 
+with a kernel that describes the covariance between any pair of pixels, representing wavelengths 
+$\lambda_i$ and $\lambda_j$ (in many ways analogous to the standard two-point spatial correlation 
+function used in cosmology).
+
+For a well-designed spectrograph, this {\it global} (i.e., present throughout the spectrum) 
+covariant structure should have a relatively low amplitude and small correlation length.  To 
+describe that structure, we assume a stationary covariance kernel (also called a radial basis 
+function) with an amplitude that depends only on the velocity distance between two pixels, 
 \begin{equation}
-  \ln[p(\fD | \vt)] \propto - \frac{1}{2} \chi^2 = - \frac{1}{2} \sum_i^N
-   \frac{R_i^2}{\sigma_i^2}
-   \label{eqn:chi2}
+  r_{ij} = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
+   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
 \end{equation}
-
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    From left to right: \textbf{Class 0} The majority of spectral lines
-    ($\gtrsim 60$\%) will have minor differences in strength between data and
-    model spectrum, which produce low-amplitude correlations in the residuals
-    on the length scale of the width of a typical spectral line.  \textbf{Class
-    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-    of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-    correlated residuals of moderate amplitude.  The difficulty
-    with class III lines is that for any specific line, there might exist a
-    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure*}
-
-However, the pixel residuals from a spectroscopic fit will often be correlated and will invalidate Equation~\ref{eqn:chi2}.
-These correlations arise because the CCD pixels of a spectrograph do not sample independent wavelengths of the spectrum.
-Instead, spectrographs are designed so that the LSF is properly sampled by at least a few pixels, meaning that adjacent pixels do not sample completely independent parts of the spectrum.
-Because of this, when a spectral feature differs between the model and the data, it creates a residual which spans several adjacent data points.
-For a collection of correlated residuals produced by a typical spectroscopic fit, see Figure~\ref{fig:badlines}.
-
-\begin{figure*}[!htb]
-\begin{center}
-  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
-  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
-  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
-\label{fig:class0}
-\end{center}
-\end{figure*}
-
-There is an important distinction between astrophysical or instrumental ``noise'' and the residuals from a spectroscopic fit.
-Noise introduced to a spectrum by astrophysical or instrumental effects is generally uncorrelated with wavelength.
-This is because the arrival of each photon to the detector is a discrete event, and while photons are scattered by the LSF, the particular scattering direction of each photon is independent from any other photon.
-We can verify this by measuring the autocorrelation of residuals in a region of well-fit stellar continuum, and show that it is flat.
-If the model spectrum is \emph{perfect}, then the residuals from a spectroscopic fit will be uncorrelated and equivalent to the astrophysical and instrumental noise introduced into the measurement.
-However, when we compare a data spectrum and a model spectrum that have many absorption lines, even small differences will produce correlated \emph{residuals}.
-We can verify this correlation by estimating the autocorrelation of the residuals as a function of pixel offset $k$
+where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
+between pixel residuals, such that
 \begin{equation}
-  \hat{\rho}\,(k) = \frac{1}{(n - k)\,\sigma^2} \sum_{i = 1}^{n - k} (R_i - \mu_R)\,(R_{i+k} - \mu_R)
-  \label{eqn:autocorrelation}
-\end{equation}
-over a region of spectrum of length $n$ pixels.
-Figure~\ref{fig:class0} shows a general residual structure induced by slight model mismatch and the autocorrelation of this series.
-There is significant correlation for offsets of $\lesssim 4$ pixels, which is also the typical width of a spectral line for this spectrum.
-
-Although the correlation in the residuals is artificially induced by an
-imperfect model fit rather than originating from a physical source of
-correlated noise, we can treat these two effects as mathematically similar. By
-using a non-trivial covariance matrix with off-diagonal terms we can describe
-the small-scale covariance between pixel residuals shown in
-Figures~\ref{fig:badlines}~and~\ref{fig:class0}. A non-diagonal covariance also
-requires that we use  a likelihood function  which uses a matrix product instead of a sum over independent pixels (Equation~\ref{eqn:lnprob}).
-
-We seek to understand and parameterize the structure of covariance in the
-spectroscopic residuals, because ignoring this complexity will bias the estimates of
-the model parameters $\vt$. As a simple analogy, consider the fit of a straight
-line to a dataset.  If the noise in the dataset is correlated, then nearby
-data points might be offset from the linear trend in the same direction by a
-similar amount.  If the covariant structure of the noise is ignored, a simple
-$\chi^2$ will treat these correlated offsets as due to the linear trend, which
-will result in a biased determination of the slope and intercept of the line,
-typically with uncertainties that are too small. We face a mathematically similar problem with 
-correlated pixel residuals from a spectroscopic fit.
-
-We parameterize the covariance structure using covariance kernels, taken from
-the field of Gaussian processes. Covariance kernels describe the covariance
-between two pixels $\lambda_i$ and $\lambda_j$.  These kernels are analogous to
-the cosmological two-point correlation function, where the distance
-between two galaxies is used instead of the wavelength of light.
-
-\paragraph{Global and mildly covariant structure}
-To account for the mild covariance structure shown in Figure~\ref{fig:class0},
-which is generally low amplitude and has a correlation length of only a few
-pixels but is present throughout the majority of the spectrum, we use a
-\emph{stationary} covariance kernel. Stationary covariance kernels, also called \emph{radial basis functions}, satisfy the property that the
-amount of correlation only depends on the distance between the two pixels $r$.  For spectra,
-we map $r$ to the velocity difference between two pixels
-\begin{equation}
-  r\,(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
-   - \lambda_j}{ \lambda_i + \lambda_j} \right |
-\end{equation}
-Then the covariance kernel describes the expected covariance between pixel residuals
-\begin{equation}
-  k(\lambda_i, \lambda_j) =  \langle R_i \; R_j \rangle
+  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
   \label{eqn:expectation}
 \end{equation}
-In theory, we could chose from a plethora of covariance kernels to parameterize this mildly covariant structure, as long as they satisfy the property that the degree of covariance smoothly declines with increasing radial distance.
-In practice, there are a few covariance kernels which are well-respected within the machine-learning community and frequently employed to parameterize covariance of this nature \citep{rw05}.
-We choose to use the \matern\ $\nu = 3/2$ kernel because it is one of the most commonly used kernels for this purpose and has an extensive history of use (cite?).
+A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
+mildly covariant structure \citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
+$\nu = 3/2$,
 \begin{equation}
-  k_{\nu = 3/2}(r\, |\, a_{\rm g},\, l) = a_{\rm g} \left(1 + \frac{\sqrt{3} r}{l} \right ) \exp 
-   \left (- \frac{\sqrt{3} r}{l} \right )
+  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
 \end{equation}
-This kernel is an exponential tapered by a polynomial, which makes for a smooth transition from small scale covariance to negligible covariance at longer scales.
-In order to ensure sparsity of the covariance matrix for computational efficiency, we
- taper the \matern\ kernel for compact support using a Hann window
+which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
+negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
+enables computational expendiency, we employ a Hann window function
 \begin{equation}
-  w\,(r |\, r_0) = \left \{ 
+  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
     \begin{array}{cc}
-    0.5 + 0.5 \cos (\pi r/ r_0 ) & r \le r_0 \\
-    0 & r > r_0\\
+    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
+    0 & r_{ij} > r_0 \\
   \end{array}
@@
 \end{equation}
-for which we generally fix $r_0$ to be some fixed multiple of the length scale, say $r_0 = 4 l$.
-Then, if we define the global parameters $\vtglobal = \{a_g, l \}$, we have a final covariance kernel of
-\begin{equation}
-  k_{\rm global} (r\,|\,\vtglobal) = w(r |\, r_0 = 4l)\; k_{\nu = 3/2}(r |\, a_{\rm g},\, l) 
-  \label{eqn:global}
-\end{equation}
-\todo{This might be a good place to actually include a simple plot of $k_{\rm global}$ as a function of $r$, as long as we aren't providing too many figures.
-Maybe it could be combined with Figure~\ref{fig:matern}.}
+to taper the kernel (so the effective kernel is $w \, \mathcal{K}_g$).  The truncation distance 
+$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
+The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
+structure in the residual spectrum that is similar to those seen in a typical data--model 
+comparison.
 
-We add this covariance kernel to the independent pixel noise $\delta_{ij} \sigma_{ij}$ already present along
- the matrix diagonal, and introduce an additional parameter to allow a constant rescaling of the independent noise. Then, each element of the covariance matrix is completely described by
-\begin{equation}
-  k(\lambda_i, \lambda_j |\vtglobal, \, b) = k_{\rm global}(r | \vtglobal) + b\,\delta_{ij} \sigma_{ij}
-\end{equation}
-Therefore, the \emph{hyperparameters} of this covariance kernel directly determine the structure of the covariance matrix.
-By varying the hyperparameters, we can create a covariance matrix which accurately mimics the correlated structure of the pixel residuals.
-Most importantly, when the hyperparameters are correctly chosen to properly model this correlated structure, the likelihood function will be maximized (Equation~\ref{eqn:prob}) while also encapsulating the true uncertainty in the spectroscopic fit.
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
-\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
+%\begin{figure*}[!htb]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
+%\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
+%\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
 There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
 This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
-\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
+%\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
 The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
 The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
-\label{fig:matern}
-\end{center}
-\end{figure*}
+%\label{fig:matern}
+%\end{center}
+%\end{figure*}
 
-The left panel of Figure~\ref{fig:matern} shows what the covariance matrix
-might look like for reasonable choices of the kernel hyperparameters. To verify
-that this covariance matrix indeed mimics the correlated structure seen in the
-residuals, we can draw a random vector of residuals $R_{\rm rand}$ from our
-noise model and compare it to the actual pixel residuals. Because we are using
-a multidimensional Gaussian as our likelihood function
-(Equation~\ref{eqn:prob}), we have assumed that the underlying residual
-structure in a spectroscopic fit is described by a multidimensional Gaussian.
-This is directly analogous to the implicit assumption of independent Gaussian
-noise when using a standard $\chi^2$ likelihood (Equation~\ref{eqn:chi2}). If
-one wished to compare the residuals that resulted from a $\chi^2$ fit to the
-noise in the data set, then one could simulate $R_{\rm rand}$ by drawing $N$
-random samples from a univariate normal distribution with zero mean and
-variance $\sigma_i^2$. We can make the same test with our multidimensional
-Gaussian likelihood function, however in this case the mean is a vector
-$\vec{\mu} = 0$ with length $N$ and the variance is described by the covariance
-matrix $C$. The residual vector is then a random draw from this multivariate
-normal distribution 
+\subsubsection{Local Covariance Structure}
+
+Aside from the global covariance structure described above, there are likely also local regions of 
+strong, highly correlated residuals that need to be treated in the modeling framework.  These large 
+amplitude residual regions are usually produced by imperfect spectral lines in the models (e.g., 
+missing opacity sources, uncertain oscillator strengths, etc.); some representative examples are 
+highlighted in Figure \ref{fig:badlines}.  To parameterize such regions in the covariance matrix, 
+we introduce a sequence of non-stationary kernels that explicitly depend on the actual wavelength 
+values of a pair of pixels (on $\lambda_i$ and $\lambda_j$), and not simply the distance between 
+them ($r_{ij}$).
+
+Assuming that these local residuals are produced primarily by pathological differences in the 
+spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
+model.  In that case, the $k^{\rm th}$ such local residual can be described as
 \begin{equation}
-  R_{\rm rand} \sim {\cal N}\left( \vec{\mu}=0,\, C\right)
+\vR_{\lambda}(a_k, \mu_k, \sigma_k) = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
 \end{equation}
-Functions to generate random samples from a multidimensional Gaussian are implemented in most numerical programming languages.
-As we see in the right panel of Figure~\ref{fig:matern}, the structure and amplitude of the fake residuals closely mimics that of the actual data residuals, suggesting that the covariance kernel faithfully reproduces the covariance induced by the data-model mismatch.
+with an amplitude $a_k$, a mean wavelength $\mu_k$, and a width $\sigma_k$.  Following 
+Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to 
+the $k^{\rm th}$ residual region is
+\begin{equation}
+  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
+% SA: There was an error in this equation; the sigma in the denominator should be squared, but was 
+%     not -- please check and make sure this is not manifested as a code bug.
+\label{eqn:klocal}
+\end{equation}
+We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
+efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
+to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  
+Figure \ref{fig:region} demonstrates how this non-stationary Gaussian 
+kernel generates a localized region of enhanced variance that successfully mimics the kind of 
+residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
+influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
+they might induce on inferences of the interesting parameters ($\vt$).  In essence, this is a 
+robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the 
+integrity of the probabilistic framework being developed (as opposed to manual or threshold-based 
+clipping/masking).  
+% SA: The basic structure of this figure is nice.  But we should talk about making it presentation 
+% quality once the other figures have been iterated as I've described (well, requested).
 
-\paragraph{Specific line covariance} Even after the mildly covariant structure has been accounted for by the global covariance kernel (Equation~\ref{eqn:global}), there still exist a significant number large amplitude pixel residuals that are \emph{highly correlated}.
-These patches of large amplitude residuals are caused by incorrect synthetic lines like those in the class I, II, and III examples in Figure~\ref{fig:badlines}.
-To parameterize the covariance in a given patch of these residuals, we introduce an additional \emph{non-stationary} covariance kernel.
-In this context, non-stationary means that the degree of covariance explicitly depends on the values of $\lambda_i$ and $\lambda_j$ and not simply the distance between them, enabling us to individually target the large residuals induced by the mismatch of a \emph{specific} spectral line.
+These local kernels can be further modified to account for more complex residual structures.  For 
+example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
+pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
+spectral scale.  This phenomenologically different local covariance behavior can still be treated 
+in this framework, if we permit the kernel in Eq.~\ref{eqn:klocal} to be modified by another 
+(stationary) function.  Here we assume that function is a squared exponential,
+\begin{equation}
+  \mathcal{S}_k(\lambda_i, \lambda_j |\, h_k) = \exp \left ( - \, \frac{r_{ij}^2}{2\,h_k^2}\right ),
+\end{equation}
+where $h_k$ is a bandwidth parameter; small $h_k$ generates high-frequency structure, and vice 
+versa.  The functionality of this kernel modification (where the appropriate kernel is now $w_k \, 
+\mathcal{S}_k \, \mathcal{K}_k$) is demonstrated for the worked example presented in Section 
+\ref{subsec:mdwarf}.  
+% SA: I have to admit that I do not really understand this very well.  I think some appropriate 
+% figures in Sect 3.2 will really help, though, so I'll wait for that before decided whether or not 
+% we need to change this part of the text.
 
-Because the shape of a spectral line can be generally approximated by a Gaussian function, and the difference of two Gaussians with the same mean is also another Gaussian, we use Equation~\ref{eqn:expectation} to design a new covariance kernel about the Gaussian function.
-Consider the residuals from a single spectral line, labeled by the subscript $0$.
-If we expect the shape of the residuals to be a Gaussian function about the line center with parameters $\vt_{{\rm line}_1} = \{ a_1, \lambda_{\mu_1}, \sigma_1 \}$
-\begin{equation}
-  R(\lambda | \vtline{1} ) = \frac{a_1}{\sqrt{2 \pi} \sigma_1} \exp \left (-\; \frac{r^2(\lambda, \lambda_{\mu_1})}{2 \sigma_1^2} \right )
-\end{equation}
-then the covariance of any two pixels along this Gaussian residual is 
-\begin{align}
-  k_{\rm G}(\lambda_i, \lambda_j |\vtline{1} ) &= \langle R(\lambda_i)\; R(\lambda_j) \rangle \\
-  &= \frac{a_1^2}{2 \pi \sigma_1} \exp \left ( - \, \frac{r^2(\lambda_i, \lambda_{\mu_1}) 
-  + r^2(\lambda_j, \lambda_{\mu_1})}{2 \sigma_1^2}\right )
-\end{align}
-We again taper the kernel with a Hann window (Equation~\ref{eqn:Hann}) for compact support, and so the covariance kernel for the highly correlated residuals induced by a poor line fit becomes
-\begin{equation}
-  k_{\rm line}(\lambda_i, \lambda_j |\vtline{1} ) = w(r,\, r_0)\, k_{\rm G}(\lambda_i, \lambda_j | \vt_{{\rm line}_1})
-  \label{eqn:line}
-\end{equation}
-To generate the covariance matrix, we now sum the effects of all the kernels, including the global covariance kernel, the specific line kernel, and the individual pixel noises
-\begin{equation}
-  k(\lambda_i, \lambda_j | \vtline{1}, \vtglobal, b) = k_{\rm line}(\lambda_i, \lambda_j |\vtline{1}) + k_{\rm global}(r | \vtglobal) + b\,\delta_{ij} \sigma_{ij}
-  \label{eqn:line_total}
-\end{equation}
+\subsubsection{Composite Covariance Matrix}
 
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
+We can now compute the covariance matrix employed in the likelihood calculation 
+(Eq.~\ref{eqn:lnprob}) as the linear combination of these kernels and the trivial pixel-by-pixel 
+noise matrix, 
+\begin{eqnarray}
+\vC_{ij}(\vT_{\rm cov}) & = & b \, \delta_{ij}\sigma_i + \\ \nonumber
+         &   &  w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ \nonumber
+         &   &  \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
+\end{eqnarray}
+where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
+h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions (see below for 
+details on how this is determined), and $b$ is assumed to be a fixed parameter that scales up the 
+pixel noise values to account for interpolation uncertainties (see Section \ref{subsec:synthetic}; 
+reasonable values are $b \approx 1.02$--1.10 for optical spectra; see Section \ref{sec:examples} 
+for examples).
+
+\comm{No idea if this is your intention for $b$, but it was the only way I could rationalize it.}
+
+%\begin{figure*}[!htb]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
 The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
 These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
-\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
 The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
 The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
-\label{fig:region}
-\end{center}
-\end{figure*}
+%\label{fig:region}
+%\end{center}
+%\end{figure*}
 
-In Figure~\ref{fig:region}, we see that the non-stationary Gaussian line kernel
-localizes the increased variance to a narrow range of the spectrum,
-corresponding to the residual from a specific poorly modeled spectral line.
-This increased covariance has the effect of down-weighting the influence of a
-specific spectral line on the likelihood function, the same way we might
-down-weight an outlier while fitting a straight line to a collection of
-data points.
+\subsection{Exploring the Posterior \label{subsec:MCMC}}
+
+\comm{Not much done here...want to meet and discuss before I try editing.}
 
@@
 
-\todo{We can (should?) add an additional level to the hierarchy of
-  hyperparameters and add parameters to describe the \emph{population
-  characteristics} of poorly modeled spectral lines (mostly the typical width,
-  amplitude of these lines).  This will tell us about the frequency and
-distribution of spectral modelling errors.} 
-We explicitly fit for the hyperparemeters of the covariance kernels at the same
-time we fit for the stellar parameters. While including these extra
-parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
-regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
+%\todo{We can (should?) add an additional level to the hierarchy of
+%  hyperparameters and add parameters to describe the \emph{population
+%  characteristics} of poorly modeled spectral lines (mostly the typical width,
+%  amplitude of these lines).  This will tell us about the frequency and
+%distribution of spectral modelling errors.} 
+%We explicitly fit for the hyperparemeters of the covariance kernels at the same
+%time we fit for the stellar parameters. While including these extra
+%parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
+%regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
 
-The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+%The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 
-Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+%Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
 When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
@@
 
-The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
+%The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
 
-\paragraph{M dwarf residual}
-\todo{This third and final covariance kernel example might push this section length over the top.
-However, I think this application is the most exciting and could be the ``killer app'' that would drive people to adopt our framework given the computational overhead.}
-\todo{I think to properly design and test this kernel, I will need to actually explore fitting an M dwarf spectrum.
-For now I have just been looking at the figure by Andrew Mann.
-However, I have high hopes!}
-
-A powerful application of covariance kernels is that data-model mismatch can be accounted for even if it takes on a messy, not-easily-parameterized form like the Gaussian line example.
-We can not just inflate the noise in a specific region, but we also taper the kernel for more complicated structure.
-For example, consider a typical residual pattern from the spectroscopic fit of an M dwarf.
-There is a complicated pattern of high and low residuals, or there might even be a region of large offset, such that features are correlated on a large scale but not so well on the near scale.
-Therefore we need a kernel which can transition between these two.
-We multiply the existing kernel by the squared exponential kernel
-\begin{equation}
-  k_{ \rm exp}(r |\, h) = \exp \left ( \frac{- r^2 }{2 h^2} \right )
-\end{equation}
-where $h$ is a bandwidth. If $h$ is small, then there will be high-frequency
-structure. If $h$ is large, then only low-frequency structure will remain. 
-\begin{equation}
-  k(\lambda_i, \lambda_j | h, \vtline{0}) = 
-   k_{\rm exp}(r |\, h) \; 
-   k_{\rm line}(\lambda_i, \lambda_j | \vtline{0})
-\end{equation}
-
-\todo{Show how the kernel can accounts for complicated M dwarf structure where
-  there is large, messy mismatch. Produce figures analogous to Figures~\ref{fig:matern}~and~\ref{fig:region}. Also add an example of M dwarf mismatch to Figure~\ref{fig:badlines}.}
-
-\subsection{Exploring the posterior}
-\label{subsec:MCMC}
 We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}).
@@
 
-\begin{figure*}[!htb]
-\begin{minipage}{\textwidth}
-\begin{center}
-  \includegraphics[width=5in]{figs/stellar_triangle.png}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+%\begin{figure*}[!htb]
+%\begin{minipage}{\textwidth}
+%\begin{center}
+%  \includegraphics[width=5in]{figs/stellar_triangle.png}
+%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
 These stellar parameters are marginalized over the Chebyshev and noise parameters.
 Made with \texttt{triangle.py}\protect\footnotemark
-  \protect\todo{In the updated figure I won't show $v_z$ or $\Omega$ so the font is easier to see. Many more samples so posteriors look smooth.}
- }
-\label{fig:stellar_posterior}
-\end{center}
-\footnotetext{\url{https://github.com/dfm/triangle.py}}
-\end{minipage}
-\end{figure*}
+%  \protect\todo{In the updated figure I won't show $v_z$ or $\Omega$ so the font is easier to see. Many more samples so posteriors look smooth.}
+% }
+%\label{fig:stellar_posterior}
+%\end{center}
+%\footnotetext{\url{https://github.com/dfm/triangle.py}}
+%\end{minipage}
+%\end{figure*}
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
-  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
-  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
+%  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
+%  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
 \textbf{Right}: The posterior probability function of the line parameters, marginalized over the stellar and global noise parameters.}
-\label{fig:noise_posterior}
-\end{center}
-\end{figure*}
+%\label{fig:noise_posterior}
+%\end{center}
+%\end{figure*}
 
@@
 
-\section{Tests}
+\section{Worked Examples \label{sec:examples}}
 Here are some ideas of which tests we might want to show.

### 2014-06-14 tex/paper.tex
changed bib tags

@@
 
-%% Nomenclature guide
-%%%%%%%%%%%%%%%%%%%%%
-% * collections of synthetic spectra are generically called ``libraries.'' The term 
-% ``grid'' is used when referring specifically to the spacing of spectral parameters
-% in the library
-
 \shorttitle{Spectroscopic inference}
@@
 something useful to say, I have added some commented-out text in the .tex file.}
-
-\comm{Important: Note how I have changed the naming convention in the .bib file, so that a paper is 
-referenced by the first author's last name and the year, like czekala14.  Your haphazard formatting 
-was driving me nuts, and you can't find anything in the bib file (plus, if you read inside the .tex 
-file, my approach makes things a hell of a lot easier).}
 
@@
 grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
-that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{vp96}.  A more 
+that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{valenti96}.  A more 
 common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
@@
 interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
-\citep[e.g., the method of SPC;][]{blj+12}.  The difficulty with this latter approach is that the 
+\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
 parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
@@
 underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
-model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{elh+06,koleva09}, 
+model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein+06,koleva09}, 
 \begin{equation}
@@
 
-For a well-designed spectrograph, this {\it global} (i.e., present throughout the spectrum) 
-covariant structure should have a relatively low amplitude and small correlation length.  To 
-describe that structure, we assume a stationary covariance kernel (also called a radial basis 
-function) with an amplitude that depends only on the velocity distance between two pixels, 
+For a well-designed spectrograph and sufficiently accurate model spectrum, this
+{\it global} (i.e., present throughout the spectrum) covariant structure should
+have a relatively low amplitude and small correlation length.  To describe that
+structure, we assume a stationary covariance kernel (also called a radial basis
+function) with an amplitude that depends only on the velocity distance between
+two pixels, 
 \begin{equation}
-  r_{ij} = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
+  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
    - \lambda_j}{ \lambda_i + \lambda_j} \right |,
@@
 \end{equation}
-A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
-mildly covariant structure \citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
+A variety of kernels have been used in the field of Gaussian
+processes to parameterize such a mildly covariant structure
+\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
 $\nu = 3/2$,
@@
 The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters.
-For more on MCMC and Gibbs samplers see \citet[ch 11]{gcs+13} and the references therein.
+For more on MCMC and Gibbs samplers see \citet[ch 11]{gelman13} and the references therein.
 
@@
 However, because we designed the kernels to deliver a sparse matrix these operations are efficient enough to be used with an MCMC algorithm.
-We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{cdh+08, dh09} and the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{fhl+12}.
+We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.
 
@@
 In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
-This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kru10}.
+This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
 

### 2014-06-26 StellarSpectra/model.py
Chebyshev arbitrary degree set by config

@@ -145,7 +145,8 @@ class Model:
         self.stellar_params = None
 
         #Now create a a list which contains an OrderModel for each order
-        self.OrderModels = [OrderModel(self.ModelSpectrum, self.DataSpectrum, index) for index in range(self.norders)]
+        self.OrderModels = [OrderModel(self.ModelSpectrum, self.DataSpectrum, index, npoly=len(self.cheb_tuple))
+                            for index in range(self.norders)]
 
     def zip_stellar_p(self, p):
         return dict(zip(self.stellar_tuple, p))
@@ -342,7 +343,7 @@ class ModelHA:
         f.close()
 
 class OrderModel:
-    def __init__(self, ModelSpectrum, DataSpectrum, index):
+    def __init__(self, ModelSpectrum, DataSpectrum, index, npoly=4):
         print("Creating OrderModel {}".format(index))
         self.index = index
         self.DataSpectrum = DataSpectrum
@@ -351,7 +352,7 @@ class OrderModel:
         self.mask = self.DataSpectrum.masks[self.index]
         self.order = self.DataSpectrum.orders[self.index]
         self.ModelSpectrum = ModelSpectrum
-        self.ChebyshevSpectrum = ChebyshevSpectrum(self.DataSpectrum, self.index)
+        self.ChebyshevSpectrum = ChebyshevSpectrum(self.DataSpectrum, self.index, npoly=npoly)
         self.CovarianceMatrix = CovarianceMatrix(self.DataSpectrum, self.index)
         self.global_cov_params = None
         self.cheb_params = None
@@ -360,7 +361,7 @@ class OrderModel:
         self.region_list = []
 
     def get_data(self):
-        return (self.wl, self.fl)
+        return (self.wl, self.fl, self.mask)
 
     def update_Cheb(self, params):
         self.ChebyshevSpectrum.update(params)
@@ -573,8 +574,9 @@ class ChebSampler(Sampler):
         #From a simple param dict, create a more complex param_dict
 
         #Then set param_tuple
+        nparams = len(starting_param_dict)
         #For now it is just set manually
-        self.param_tuple = ("logc0", "c1", "c2", "c3")
+        self.param_tuple = ("logc0",) + tuple(["c{}".format(i) for i in range(1, nparams)])
         #self.param_tuple = ("c1", "c2", "c3")
         self.order_index = order_index
 

### 2014-06-26 scripts/stars/base_lnprob.py
Chebyshev arbitrary degree set by config

@@ -43,7 +43,7 @@ myHDF5Interface = HDF5Interface(config['HDF5_path'])
 
 stellar_Starting = config['stellar_params']
 #Note that these values are sigma^2!!
-stellar_MH_cov = np.array([5, 0.02, 0.02, 0.02, 0.02, 2e-3])**2 * np.identity(len(stellar_Starting))
+stellar_MH_cov = np.array([5, 0.02, 0.02, 0.5, 0.1, 2e-3])**2 * np.identity(len(stellar_Starting))
 # stellar_MH_cov = np.array([0.1, 0.001, 0.001, 0.001, 0.001, 1e-5])**2 * np.identity(len(stellar_Starting))
 stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
 
@@ -66,10 +66,15 @@ stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
 #We could test to see if these jumps are being executed in the right direction by checking to see what the 2D pairwise
 # chain positions look like
 
+cheb_degree = config['cheb_degree']
+cheb_MH_cov = (5e-3)**2 * np.identity(cheb_degree)
+cheb_tuple = ("logc0",)
+#add in new coefficients
+for i in range(1, cheb_degree):
+    cheb_tuple += ("c{}".format(i),)
+#set starting position to 0
+cheb_Starting = {k:0.0 for k in cheb_tuple}
 
-cheb_Starting = config['cheb_params']
-cheb_MH_cov = np.array([5e-3, 5e-3, 5e-3, 5e-3])**2 * np.identity(len(cheb_Starting))
-cheb_tuple = ("logc0", "c1", "c2", "c3")
 
 cov_Starting = config['cov_params']
 cov_MH_cov = np.array([0.02, 0.02, 0.4])**2 * np.identity(len(cov_Starting))
@@ -113,6 +118,7 @@ for i in range(len(config['orders'])):
     samplerList.append(CovGlobalSampler(myModel, cov_MH_cov, cov_Starting, order_index=i, outdir=outdir))
     samplerList.append(RegionsSampler(myModel, region_MH_cov, max_regions=config['max_regions'], order_index=i, outdir=outdir))
     cadenceList += [6, 6, 2]
+    # cadenceList += [6, 2]
 
 mySampler = MegaSampler(myModel, samplers=[myStellarSampler] + samplerList,
                         burnInCadence=[10] + cadenceList, cadence=[10] + cadenceList)

### 2014-06-26 StellarSpectra/spectrum.py
Chebyshev arbitrary degree set by config

@@ -1389,17 +1389,29 @@ class ChebyshevSpectrum:
         self.fix_c0 = True if index == (len(DataSpectrum.wls) - 1) else False #Fix the last c0
 
         xs = np.arange(len_wl)
-        T0 = np.ones_like(xs)
-        Ch1 = Ch([0, 1], domain=[0, len_wl - 1])
-        T1 = Ch1(xs)
-        Ch2 = Ch([0, 0, 1], domain=[0, len_wl - 1])
-        T2 = Ch2(xs)
-        Ch3 = Ch([0, 0, 0, 1], domain=[0, len_wl - 1])
-        T3 = Ch3(xs)
 
-        self.T = np.array([T1, T2, T3])
+        #Create Ch1, etc... for each coefficient in npoly excepting logc0
+        #Evaluate these and stuff them into self.T
+        coeff = [1]
+        T = []
+        for i in range(1, npoly):
+            # print("i = ", i)
+            coeff = [0] + coeff
+            Chtemp = Ch(coeff, domain=[0, len_wl - 1])
+            Ttemp = Chtemp(xs)
+            T += [Ttemp]
+
+        # Ch1 = Ch([0, 1], domain=[0, len_wl - 1])
+        # T1 = Ch1(xs)
+        # Ch2 = Ch([0, 0, 1], domain=[0, len_wl - 1])
+        # T2 = Ch2(xs)
+        # Ch3 = Ch([0, 0, 0, 1], domain=[0, len_wl - 1])
+        # T3 = Ch3(xs)
+
+        # self.T = np.array([T1, T2, T3])
+        self.T = np.array(T)
         self.npoly = npoly
-        assert self.npoly == 4, "Only handling order 4 Chebyshev for now."
+        # assert self.npoly == 4, "Only handling order 4 Chebyshev for now."
 
         #Dummy holders
         self.k = np.ones(len_wl)
@@ -1425,7 +1437,10 @@ class ChebyshevSpectrum:
         else:
             logc0 = params["logc0"]
 
-        cns = np.array([params["c1"], params["c2"], params["c3"]])
+        #Convert params dict to a 1d array of coefficients
+        cns = np.array([params["c{}".format(i)] for i in range(1, self.npoly)])
+
+        # cns = np.array([params["c1"], params["c2"], params["c3"]])
 
         #if c0 < 0:
         #    raise C.ModelError("Negative c0s are not allowed.")

### 2014-06-29 tex/paper.tex
added flot submodule
Chebyshev arbitrary degree set by config
WASP run

@@
 \newcommand{\yN}{\{y_i\}_{i=1}^N}
+
+\newcommand{\kms}{ \textrm{km s}^{-1} }
 
@@
 
-\comm{The following are some suggested paragraph outlines for a concise introduction:}
+A single stellar spectrum contains a vast amount of information about the physical properties of the star.
+It is of supreme interest to extract a maximal amount of this information from the spectrum in order to study the fundamental stellar parameters such as effective temperature, surface gravity, and metallicity.
+Accurate and unbiased estimates of stellar parameters are crucial to the study of numerous fields of astrophysics.
+For example, the mass, radius, and temperature of every newly discovered exoplanet (save for directly imaged planets) are measured relative to the properties of its host star.
+Large uncertainties or biases in the stellar parameters will translate into incorrect estimates of the habitable zone and the occurrence rate of earth-like planets \citep{kane14}.
+High quality stellar parameters are also necessary for the study of stellar evolution.
+Modelers of stellar evolution rely upon accurate samples of stellar parameters to chart the evolutionary tracks of stars through the Hertsprung-Russell diagram.
 
-\comm{(1) start general: vast amount of information in an astronomical spectrum, and we want to 
-make the most of that to learn about various physical parameters of interest.  accurate, unbiased 
-stellar parameters (like temperature, gravity, metallicity, etc.) are of crucial importance in many 
-astrophysics problems; e.g., exoplanets (and explain why), etc. (pick another one or two).}
+However, extracting physical parameters of stars from a stellar spectrum is often difficult, due to several factors.
+First, generating synthetic stellar spectra to use as a benchmark for comparison requires a sophisticated model of the stellar atmosphere and radiative transfer, along with an extensive, high-quality database of opacity sources.
+Accurately calibrating the atomic constants in this database is a laborious task, and requires high resolution, high signal-to noise spectra for verification, which limits targets mainly to extremely bright stars like the Sun and Vega.
+This makes it difficult to achieve the same precision when extending synthetic stellar models to intrinsically fainter stars, like M dwarfs, which have additional sources of molecular opacity.
+Second,  synthesizing a spectrum with such a sophisticated model requires a large amount of computational power, such that a full-optical spectrum at high resolution will take a day or more to process on a supercomputer.
 
-\comm{(2) point out the challenge: quantitative extraction of these physical parameters is not 
-easy.  go through some of the basic underlying limitations (practical, computational, limited 
-models, etc.)}
+Several, well-tested techniques exist that use spectra to infer stellar parameters.
+The most widely used technique relies upon measuring ratios of lines, or line indexes, which are matched to a set of indices measured from a synthetic spectrum.
+More sophisticated codes exist that synthesize spectra for direct comparison with certain well-studied lines, (e.g., \ion{Fe}{0}, \ion{Mg}{0}, and \ion{Ca}{0}) such as MOOG \citep{sneden73} and SME \citep{valenti96}.
+Other codes such as SPC \citep{buchhave12} cross-correlate the observed spectrum against a grid of pre-computed synthetic spectra and then fit a surface to the normalized cross-correlation coefficients to determine parameter uncertainties.
 
-\comm{(3) highlight various techniques/approaches being used (lots of references here).}
-
-\comm{(4) now, without criticizing those approaches, just launch into a brief summary of our goals 
-and a basic outline of the paper.}
-
-\comm{I have made a lot of edits in Section 2, but have tried to restrict comments.  When there is 
-something useful to say, I have added some commented-out text in the .tex file.}
+Owing to the high quality of libraries of pre-computed synthetic stellar spectra, it is now possible to generate an accurate forward model of a stellar spectrum.
+We seek to place this model in a Bayesian framework and construct a likelihood function suitable for comparing synthetic spectral models to real spectral data.
+We use a non-trivial covariance matrix, parameterized by a Gaussian process covariance kernel, to account for the residual pixel correlations that arise from a spectroscopic fit, which is a more complex problem than can be addressed with a simple $\chi^2$ best-fit metric.
+While designed for fitting stellar spectra, our method is general for all comparisons of spectra, and could be used to fit any type of spectra with a synthetic model, (e.g., galaxies or unresolved stellar clusters).
+In \S\ref{sec:method}, we describe the methodology, including a treatment of the covariance introduced by a spectroscopic fit.
+In \S\ref{sec:examples}, we present applications of our method to two examples, using optical spectra of an F type star and infrared spectra of an M star.
+In \S\ref{sec:discussion} we discuss the implications of our method and potential applications to develop data-driven models of stellar spectra.
 
@@
 post-processing steps.
+
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{figs/kernels.pdf}
+  \includegraphics{figs/high2low.pdf}
+  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite.
+Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000.
+Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$.
+Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
+\label{fig:broadening}
+\end{center}
+\end{figure*}
+
 % SA: My idea for this figure is sort of 1-column wide (~3.5"), with a set of spectral panels going 
@@
 underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
-model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein+06,koleva09}, 
+model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
 \begin{equation}
@@
 these nuisance parameters are applied. 
-% SA: I envision this figure much like the previous one, in a 1-column vertical distribution of 
-% panels.  At top, you would show the data spectrum at an overlap of two spectral orders, along 
-% with the post-processed model with no Chebyshev polynomials applied.  Next, you would show the 
-% two Chebyshev polynomials appropriate for the two orders.  And then in the final panel you would 
-% revisit the top panel with the polynomials applied, demonstrating the improvement.
+\comm{I envision this figure much like the previous one, in a 1-column vertical distribution of 
+ panels.  At top, you would show the data spectrum at an overlap of two spectral orders, along 
+ with the post-processed model with no Chebyshev polynomials applied.  Next, you would show the 
+ two Chebyshev polynomials appropriate for the two orders.  And then in the final panel you would 
+ revisit the top panel with the polynomials applied, demonstrating the improvement.}
 
@@
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics{figs/badlines.pdf}
-%  \caption{A collection of spectral lines which have imperfect model fits.
-%    From left to right: \textbf{Class 0} The majority of spectral lines
-%    ($\gtrsim 60$\%) will have minor differences in strength between data and
-%    model spectrum, which produce low-amplitude correlations in the residuals
-%    on the length scale of the width of a typical spectral line.  \textbf{Class
-%    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-%    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-%    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-%    of all lines), an extraneous line in the model leaves a large, highly
-%    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-%    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-%    correlated residuals of moderate amplitude.  The difficulty
-%    with class III lines is that for any specific line, there might exist a
-%    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
-%    will properly fit \emph{all} the lines.}
-%\label{fig:badlines}
-%\end{center}
-%\end{figure*}
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines.pdf}
+  \caption{A collection of spectral lines which have imperfect model fits.
+    From left to right: \textbf{Class 0} The majority of spectral lines
+    ($\gtrsim 60$\%) will have minor differences in strength between data and
+    model spectrum, which produce low-amplitude correlations in the residuals
+    on the length scale of the width of a typical spectral line.  \textbf{Class
+    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+    of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
+    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+    correlated residuals of moderate amplitude.  The difficulty
+    with class III lines is that for any specific line, there might exist a
+    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure*}
 
@@
 line width in this spectrum.
-% SA: It may be true that the figure you've made to demonstrate this is accurate, but it certainly 
-% does not seem obvious to me what is going on in it.  I like the idea of the figure, but I think 
-% it would be better demonstrated on a relatively smooth (continuum) region with a small spectral 
-% mismatch, rather than deep lines everywhere.  Can we do that instead?
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
-%  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
-%  \caption{\textbf{Left} The same residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-%  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
-%\label{fig:class0}
-%\end{center}
-%\end{figure*}
+\begin{figure*}[!htb]
+\begin{center}
+  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
+  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
+  \caption{\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
+\label{fig:class0}
+\end{center}
+\end{figure*}
 
@@
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
-%\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
-%\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
+\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
 There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
 This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
-%\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
+\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
 The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
 The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
-%\label{fig:matern}
-%\end{center}
-%\end{figure*}
+\label{fig:matern}
+\end{center}
+\end{figure*}
 
@@
   \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
-% SA: There was an error in this equation; the sigma in the denominator should be squared, but was 
-%     not -- please check and make sure this is not manifested as a code bug.
 \label{eqn:klocal}
@@
 clipping/masking).
-% SA: The basic structure of this figure is nice.  But we should talk about making it presentation 
-% quality once the other figures have been iterated as I've described (well, requested).
 
@@
 \ref{subsec:mdwarf}.  
-% SA: I have to admit that I do not really understand this very well.  I think some appropriate 
-% figures in Sect 3.2 will really help, though, so I'll wait for that before decided whether or not 
-% we need to change this part of the text.
+\comm{I have to admit that I do not really understand this very well.  I think some appropriate 
+figures in Sect 3.2 will really help, though, so I'll wait for that before decided whether or not 
+we need to change this part of the text.}
 
@@
 where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
-h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions (see below for 
-details on how this is determined), and $b$ is assumed to be a fixed parameter that scales up the 
-pixel noise values to account for interpolation uncertainties (see Section \ref{subsec:synthetic}; 
-reasonable values are $b \approx 1.02$--1.10 for optical spectra; see Section \ref{sec:examples} 
-for examples).
+h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
+(see below for details on how this is determined), and $b$ is assumed to be a
+fixed parameter that scales up the pixel noise values to account for read
+noise, noise added during the 2D to 1D spectral extraction procedure, and
+interpolation uncertainties (see Section \ref{subsec:synthetic}; reasonable
+values are $b \approx 1.02$--1.10 for well-calibrated optical spectra; see Section
+\ref{sec:examples} for examples).
 
@@
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
-%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
 The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
 These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
-%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
 The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
 The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
-%\label{fig:region}
-%\end{center}
-%\end{figure*}
+\label{fig:region}
+\end{center}
+\end{figure*}
 
@@
 
-%\todo{We can (should?) add an additional level to the hierarchy of
-%  hyperparameters and add parameters to describe the \emph{population
-%  characteristics} of poorly modeled spectral lines (mostly the typical width,
-%  amplitude of these lines).  This will tell us about the frequency and
-%distribution of spectral modelling errors.} 
-%We explicitly fit for the hyperparemeters of the covariance kernels at the same
-%time we fit for the stellar parameters. While including these extra
-%parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
-%regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
+\todo{We can (should?) add an additional level to the hierarchy of
+  hyperparameters and add parameters to describe the \emph{population
+  characteristics} of poorly modeled spectral lines (mostly the typical width,
+  amplitude of these lines).  This will tell us about the frequency and
+distribution of spectral modelling errors.} 
+We explicitly fit for the hyperparameters of the covariance kernels at the same
+time we fit for the stellar parameters. While including these extra
+parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
+regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
 
-%The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 
-%Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
 When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
@@
 
-%The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
+The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
@@
 
-%\begin{figure*}[!htb]
-%\begin{minipage}{\textwidth}
-%\begin{center}
-%  \includegraphics[width=5in]{figs/stellar_triangle.png}
-%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft, width=5in]{figs/stellar_triangle.png}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
 These stellar parameters are marginalized over the Chebyshev and noise parameters.
-Made with \texttt{triangle.py}\protect\footnotemark
-%  \protect\todo{In the updated figure I won't show $v_z$ or $\Omega$ so the font is easier to see. Many more samples so posteriors look smooth.}
-% }
-%\label{fig:stellar_posterior}
-%\end{center}
-%\footnotetext{\url{https://github.com/dfm/triangle.py}}
-%\end{minipage}
-%\end{figure*}
+Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+}
+\label{fig:stellar_posterior}
+\end{center}
+\end{figure*}
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
-%  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
-%  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
+  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
+  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
 \textbf{Right}: The posterior probability function of the line parameters, marginalized over the stellar and global noise parameters.}
-%\label{fig:noise_posterior}
-%\end{center}
-%\end{figure*}
+\label{fig:noise_posterior}
+\end{center}
+\end{figure*}
 
@@
 
+\section{Discussion \label{sec:discussion}}
+
+\section{Conclusion \label{sec:conclusion}}
+
 \bibliographystyle{yahapj}

### 2014-06-29 scripts/stars/base_lnprob.py
WASP run

@@ -10,21 +10,25 @@ import shutil
 
 import argparse
 parser = argparse.ArgumentParser(prog="base_lnprob.py", description="Run StellarSpectra fitting model.")
-parser.add_argument("-p", "--params", help="*.yaml file specifying parameters.")
+parser.add_argument("-i", "--input", help="*.yaml file specifying parameters.")
+parser.add_argument("-r", "--run_index", help="Which run (of those running concurrently) is this? All data will "
+                                        "be written into this directory, overwriting any that exists.")
+parser.add_argument("-p", "--perturb", action="store_true", help="Randomly perturb the starting position of the "
+                                                                 "chain, as a multiple of the jump parameters.")
 args = parser.parse_args()
 
-if args.params: #
+if args.input: #
     #assert that we actually specified a *.yaml file
-    if ".yaml" in args.params:
-        yaml_file = args.params
-        f = open(args.params)
+    if ".yaml" in args.input:
+        yaml_file = args.input
+        f = open(args.input)
         config = yaml.load(f)
         f.close()
 
     else:
         import sys
         sys.exit("Must specify a *.yaml file.")
-        yaml_file = args.params
+
 else:
     #load the default config file
     yaml_file = "scripts/stars/input.yaml"
@@ -84,12 +88,24 @@ region_MH_cov = np.array([0.04, 0.02, 0.1])**2 * np.identity(len(region_tuple))
 outdir = config['outdir']
 name = config['name']
 base = outdir + name + "run{:0>2}/"
-run_num = 0
-while os.path.exists(base.format(run_num)) and (run_num < 20):
-    print(base.format(run_num), "exists")
-    run_num += 1
 
-outdir = base.format(run_num)
+#This code is necessary for multiple simultaneous runs on odyssey, so that different runs do not write into the same
+#output directory
+if args.run_index == None:
+    run_index = 0
+    while os.path.exists(base.format(run_index)) and (run_index < 20):
+        print(base.format(run_index), "exists")
+        run_index += 1
+    outdir = base.format(run_index)
+
+else:
+    run_index = args.run_index
+    outdir = base.format(run_index)
+    #Delete this outdir, if it exists
+    if os.path.exists(outdir):
+        print("Deleting", outdir)
+        shutil.rmtree(outdir)
+
 print("Creating ", outdir)
 os.makedirs(outdir)
 
[…]

### 2014-06-29 tex/paper.tex
updated path

@@
-\documentclass[preprint]{aastex} 
-%\documentclass[iop,floatfix]{emulateapj}
+%\documentclass[preprint]{aastex} 
+\documentclass[iop,floatfix]{emulateapj}
 
@@
 
-\section{Methodology \label{sec:method}}
+\section{Methodology} \label{sec:method}
 
@@
 
-\subsection{Post-Processing \label{subsec:postprocess}}
+\subsection{Post-Processing}
+\label{subsec:postprocess}
 
@@
 
-\begin{figure*}[!htb]
+\begin{figure}[!htb]
 \begin{center}
@@
 \end{center}
-\end{figure*}
-
-% SA: My idea for this figure is sort of 1-column wide (~3.5"), with a set of spectral panels going 
-% down vertically.  The top panel shows a segment of the full resolution model.  The next one down 
-% can show the individual kernels described above, as well as their composite, \varphi_v.  Then the 
-% next panel down can show the convolved (but still full-resolution) spectrum.  Finally, the bottom 
-% panel will show the spectrum re-sampled onto the observed pixels.  It need not be a large figure 
-% (maybe 5" long?).
+\end{figure}
 
@@
 rather than the residual calibration mismatches as described here (with a suitable relaxation of 
-the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:cheb} offers a practical demonstration of how 
+the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
 these nuisance parameters are applied.
-\comm{I envision this figure much like the previous one, in a 1-column vertical distribution of 
- panels.  At top, you would show the data spectrum at an overlap of two spectral orders, along 
- with the post-processed model with no Chebyshev polynomials applied.  Next, you would show the 
- two Chebyshev polynomials appropriate for the two orders.  And then in the final panel you would 
- revisit the top panel with the polynomials applied, demonstrating the improvement.}
 
-\subsection{Model Evaluation \label{subsec:likelihood}}
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics{figs/chebyshev.pdf}
+  \caption{A region of the echelle spectrum at the overlap between two adjacent orders of the echelle spectrum.
+Panel 1: In green and blue, we have the flux-calibrated dataset.
+The adjacent orders overlap but there is a slight discrepancy of $\lesssim 3$\%.
+In the background in red is a model spectrum generated for reasonable values of stellar parameters, $\vt$.
+Panel 2: In order to account for residual errors in the flux calibration, we allow the model to be multiplied by a Chebyshev polynomial.
+Here are the polynomials determined as best fit, to multiply the model.
+Panels 3 \& 4: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4).
+This has the advantage of preserving the original dataset as a fixed quantity and makes the model linear in the Chebyshev polynomial coefficients.}
+\label{fig:chebyshev}
+\end{center}
+\end{figure}
+
+\subsection{Model Evaluation}
+\label{subsec:likelihood}
 
@@
 
-\comm{No idea if this is your intention for $b$, but it was the only way I could rationalize it.}
-
 \begin{figure*}[!htb]
@@
 
-\subsection{Exploring the Posterior \label{subsec:MCMC}}
+\subsection{Exploring the Posterior }
+\label{subsec:MCMC}
 

### 2014-06-29 scripts/test_arr.py
updated path

@@ -1,0 +1,9 @@
+import argparse
+
+parser = argparse.ArgumentParser(prog="test_arr.py", description="Identify Job Array variables")
+parser.add_argument("--littlea")
+parser.add_argument("--biga")
+args = parser.parse_args()
+
+print("a =", args.littlea)
+print("A =", args.biga)

### 2014-06-29 tex/paper.tex
hdfcat

@@
   amplitude of these lines).  This will tell us about the frequency and
-distribution of spectral modelling errors.} 
-We explicitly fit for the hyperparameters of the covariance kernels at the same
-time we fit for the stellar parameters. While including these extra
-parameters does increase the dimensionality and complexity of our model, they do provide several advantages. \emph{A priori} we do not know which
-regions of the spectrum are improperly modeled. The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
+distribution of spectral modelling errors.} We explicitly fit for the
+hyperparameters of the covariance kernels at the same time we fit for the
+stellar parameters. While including these extra parameters does increase the
+dimensionality and complexity of our model, they do provide several advantages.
+\emph{A priori} we do not know which regions of the spectrum are improperly
+modeled. The covariance hyperparameters provide a framework to identify these
+regions iteratively and in a self-consistent manner.
 
@@
 
-\paragraph{Test 2}: I think we should fit both an F/G star (WASP-14) and a K/M star (TBD) with both the Kurucz and PHOENIX spectra. This will show a few things
+\paragraph{Test 2}: I think we should fit an optical spectrum of an F star (WASP-14) and a near-IR spectrum of an M star (Gl51).
+The WASP will be fit with both the Kurucz and PHOENIX spectra, while the M star will be fit with just the PHOENIX.
+This will show a few things
 \begin{itemize}

### 2014-06-30 flot_model.py
new input file

@@ -81,8 +81,7 @@ def main():
 
     from StellarSpectra.model import Model
     from StellarSpectra.spectrum import DataSpectrum
-    from StellarSpectra.\
-        grid_tools import TRES, HDF5Interface
+    from StellarSpectra.grid_tools import TRES, HDF5Interface
 
     #Figure out what the relative path is to base
     import StellarSpectra
@@ -95,6 +94,13 @@ def main():
     myModel = Model.from_json(args.json, myDataSpectrum, myInstrument, myHDF5Interface)
 
     for model in myModel.OrderModels:
+
+        #If an order has regions, read these out from model_final.json
+        region_dict = model.get_regions_dict()
+        print("Region dict", region_dict)
+        #loop through these to determine the wavelength of each
+        wl_regions = [value["mu"] for value in region_dict.values()]
+        print(wl_regions)
 
         #Get the data, sigmas, and mask
         wl, fl, sigma, mask = model.get_data()
@@ -110,9 +116,6 @@ def main():
         plot_data = order_json(wl, fl, sigma, mask, flm, cheb)
 
         render_template(base, plot_data)
-
-        #Get the Chebyshev spectrum
-        # cheb = model.get_Cheb()
 
         #Get the covariance matrix
         # S = model.get_Cov()

### 2014-06-30 scripts/stars/base_lnprob.py

@@ -132,7 +132,7 @@ for order in config['orders']:
     os.makedirs(order_dir)
 
 #Copy yaml file to outdir
-shutil.copy(yaml_file, outdir)
+shutil.copy(yaml_file, outdir + "/input.yaml")
 
 myModel = Model(myDataSpectrum, myInstrument, myHDF5Interface, stellar_tuple=stellar_tuple, cheb_tuple=cheb_tuple,
                 cov_tuple=cov_tuple, region_tuple=region_tuple, outdir=outdir)

### 2014-06-30 tex/paper.tex
new input file
added nocov ability

@@
 
-\section{Introduction \label{sec:intro}}
+\section{Introduction}
+\label{sec:intro}
 
@@
 unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
-artifacts is negligible (see Secton~\ref{sec:examples} for examples).  The lowest-order (scaling) 
+artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-order (scaling) 
 coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
@@
 \begin{center}
-  \includegraphics[draft, width=5in]{figs/stellar_triangle.png}
+  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
   \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
@@
 \label{fig:stellar_posterior}
+\end{center}
+\end{figure*}
+
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[width=5in]{figs/stellar_triangle_WASP14_Kurucz.png}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, logg fixed to match \citep{torres12}. Using three orders, Kurucz grid. Same assumptions.}
+\label{fig:stellar_posterior_logg}
 \end{center}
@@
 
-\section{Worked Examples \label{sec:examples}}
+\section{Worked Examples}
+\label{sec:examples}
 Here are some ideas of which tests we might want to show.
@@
 
-\section{Discussion \label{sec:discussion}}
+\section{Discussion}
+\label{sec:discussion}
 
-\section{Conclusion \label{sec:conclusion}}
+\section{Conclusion}
+\label{sec:conclusion}
 

### 2014-06-30 flot_model.py

@@ -100,7 +100,8 @@ def main():
         print("Region dict", region_dict)
         #loop through these to determine the wavelength of each
         wl_regions = [value["mu"] for value in region_dict.values()]
-        print(wl_regions)
+
+        #Make vertical markings at the location of the wl_regions.
 
         #Get the data, sigmas, and mask
         wl, fl, sigma, mask = model.get_data()

### 2014-06-30 tex/paper.tex
PHOENIX run

@@
 \usepackage{bm}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 
+\usepackage{color}
 \usepackage[utf8]{inputenc}
@@
 
-\newcommand{\todo}[1]{ \textcolor{Blue}{\\TODO: #1}}
-\newcommand{\comm}[1]{ \textcolor{Red}{SA: #1}}
+\newcommand{\todo}[1]{ \textcolor{blue}{\\TODO: #1}}
+\newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}
 
@@
 
+\section{Tables}
+
+\begin{deluxetable}{rr}
+\tablecaption{Nomenclature used in this document}
+\tablehead{
+  \colhead{Symbol} & \colhead{Description}
+}
+\startdata
+%Symbol & Description\\
+%\hline
+%\hline
+$i$ & index specifying a pixel\\
+%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+%$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+%  & that parameterize a synthetic spectrum from the grid\\
+%$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+%  & are applied during ``post processing'' of the synthetic spectrum\\
+%$\vtstar$ & $\{\vg,\vpp \}$\\
+%$\finst(\lambda)$ & data spectrum\\
+%$\fsynth(\lambda)$ & synthetic spectrum\\
+%$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
+%$\vtline{1}$ & \\
+%$\vtlines$ & \\
+%$\vtorder{1}$ & \\
+%$\vtorders$ & \\
+%$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
+%$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
+%$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
+%$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
+%$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
+%$\sigma_i$ & Poisson noise for a given pixel $i$\\
+%$R_i$ & residuals $\fDi - \fMi$\\
+%$\fR$ & residual vector $\fD - \fM$\\
+%$C$ & covariance matrix\\
+%$\sigma_{ij}$ & element in the covariance matrix\\
+%$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
+%$k_{\rm global}$ & global covariance kernel\\
+%$k_{\rm line}$ & regional covariance kernel\\
+%\hline
+%\caption{Nomenclature used in this document.}
+%\label{tab:nomenclature}
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{rr} 
+\tablecaption{\label{table:tests} Tests}
+\tablehead{\colhead{Object} & \colhead{Orders}}
+\startdata
+WASP-14 & 22-24 \\
+WASP-14 & 22-24 \\
+WASP-14 & 22-24 \\
+WASP-14 & 22-24 \\
+\enddata
+\end{deluxetable}
+
 \appendix
-
-\begin{table}[!htb]
-\begin{tabular}{ll}
-\hline
-\hline
-Symbol & Description\\
-\hline
-\hline
-$i$ & index specifying a pixel\\
-$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-  & that parameterize a synthetic spectrum from the grid\\
-$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-  & are applied during ``post processing'' of the synthetic spectrum\\
-$\vtstar$ & $\{\vg,\vpp \}$\\
-$\finst(\lambda)$ & data spectrum\\
-$\fsynth(\lambda)$ & synthetic spectrum\\
-$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
-$\vtline{1}$ & \\
-$\vtlines$ & \\
-$\vtorder{1}$ & \\
-$\vtorders$ & \\
-$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
-$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
-$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
-$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
-$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
-$\sigma_i$ & Poisson noise for a given pixel $i$\\
-$R_i$ & residuals $\fDi - \fMi$\\
-$\fR$ & residual vector $\fD - \fM$\\
-$C$ & covariance matrix\\
-$\sigma_{ij}$ & element in the covariance matrix\\
-$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-$k_{\rm global}$ & global covariance kernel\\
-$k_{\rm line}$ & regional covariance kernel\\
-\hline
-\end{tabular}
-\caption{Nomenclature used in this document.}
-\label{tab:nomenclature}
-\end{table}
 
@@
  with git hash \texttt{\githash\,(\gitdate)}.
-
 \end{document}

### 2014-07-01 tex/paper.tex
new phoenix

@@
 
-\begin{deluxetable}{rr}
+\begin{deluxetable}{ll}
 \tablecaption{Nomenclature used in this document}
@@
 \startdata
-%Symbol & Description\\
-%\hline
-%\hline
 $i$ & index specifying a pixel\\
-%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-%$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-%  & that parameterize a synthetic spectrum from the grid\\
-%$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-%  & are applied during ``post processing'' of the synthetic spectrum\\
-%$\vtstar$ & $\{\vg,\vpp \}$\\
-%$\finst(\lambda)$ & data spectrum\\
-%$\fsynth(\lambda)$ & synthetic spectrum\\
-%$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
-%$\vtline{1}$ & \\
-%$\vtlines$ & \\
-%$\vtorder{1}$ & \\
-%$\vtorders$ & \\
-%$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
-%$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
-%$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
-%$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
-%$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
-%$\sigma_i$ & Poisson noise for a given pixel $i$\\
-%$R_i$ & residuals $\fDi - \fMi$\\
-%$\fR$ & residual vector $\fD - \fM$\\
-%$C$ & covariance matrix\\
-%$\sigma_{ij}$ & element in the covariance matrix\\
-%$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-%$k_{\rm global}$ & global covariance kernel\\
-%$k_{\rm line}$ & regional covariance kernel\\
-%\hline
-%\caption{Nomenclature used in this document.}
-%\label{tab:nomenclature}
+$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+  & that parameterize a synthetic spectrum from the grid\\
+$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+  & are applied during ``post processing'' of the synthetic spectrum\\
+$\vtstar$ & $\{\vg,\vpp \}$\\
+$\finst(\lambda)$ & data spectrum\\
+$\fsynth(\lambda)$ & synthetic spectrum\\
+$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
+$\vtline{1}$ & \\
+$\vtlines$ & \\
+$\vtorder{1}$ & \\
+$\vtorders$ & \\
+$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
+$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
+$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
+$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
+$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
+$\sigma_i$ & Poisson noise for a given pixel $i$\\
+$R_i$ & residuals $\fDi - \fMi$\\
+$\fR$ & residual vector $\fD - \fM$\\
+$C$ & covariance matrix\\
+$\sigma_{ij}$ & element in the covariance matrix\\
+$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
+$k_{\rm global}$ & global covariance kernel\\
+$k_{\rm line}$ & regional covariance kernel\\
 \enddata
@@
 
-\begin{deluxetable}{rr} 
+\begin{deluxetable}{llrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
 \tablecaption{\label{table:tests} Tests}
-\tablehead{\colhead{Object} & \colhead{Orders}}
+\tablehead{
+  \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
+  \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
+  \multicolumn{2}{c}{$\Z \pm \sigma$} & \colhead{comments}
+}
 \startdata
-WASP-14 & 22-24 \\
-WASP-14 & 22-24 \\
-WASP-14 & 22-24 \\
-WASP-14 & 22-24 \\
+WASP-14 & 22-24 & 5060-5315 & Kurucz &  &  &  &  &  & & No covariance\\
+WASP-14 & 22-24 & 5060-5315& Kurucz & 6314 & 50 & 3.92 & 0.08 & -0.39 & 0.03 & Matern only\\
+WASP-14 & 22-24 & 5060-5315& Kurucz & 6512 & 32 & \multicolumn{2}{c}{\nodata} & -0.27 & 0.02 & Matern + fixed $\log g = 4.29$\\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & No covariance \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Matern only \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Matern + fixed $\log g = 4.29$ \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Regions included \\
 \enddata

### 2014-07-02 tex/paper.tex
running just order 24

@@
 
+\subsection{Fitting data}
+
+[This may need to be eventually moved to \S2, but here it is for now] Cite papers that acquired the data.
+Describe the instruments.
+Describe when it was taken.
+Describe the noise properties, etc.
+Maybe a table might be worthwhile to describe the spectra.
+Here also describe the WASP-14 fit using the Kurucz spectra, using a fixed log g, and state that they are the same as Willie's paper.
+Talk about putting things in a Bayesian framework, using a sensitivity to a prior, to a $\delta$-function prior.
+Point out that this Bayesian approach allows you to do this.
+
+\subsection{Generic Tests}
+
+Designed to show off what the model framework can actually do.
+
+Talk about giving more reasonable errors using regions, in the biased case.
+
 \paragraph{Test 1}: Fitting with and without the global covariance kernel to show how the width of the posteriors nicely inflates to reasonable uncertainties.
@@
 \startdata
-WASP-14 & 22-24 & 5060-5315 & Kurucz &  &  &  &  &  & & No covariance\\
+WASP-14 & 22-24 & 5060-5315 & Kurucz & 6310 & 52 & 3.91  & 0.08 & -0.38  & 0.03 & Poisson only\\
 WASP-14 & 22-24 & 5060-5315& Kurucz & 6314 & 50 & 3.92 & 0.08 & -0.39 & 0.03 & Matern only\\
 WASP-14 & 22-24 & 5060-5315& Kurucz & 6512 & 32 & \multicolumn{2}{c}{\nodata} & -0.27 & 0.02 & Matern + fixed $\log g = 4.29$\\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & No covariance \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Matern only \\
+WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.3 & -0.50 & 0.01 & Poisson only \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & 5865 & 43 & 3.2 & 0.08 & -0.85 & 0.03 & Matern only. Because of 24. \\
 WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Matern + fixed $\log g = 4.29$ \\

### 2014-07-02 tex/paper.tex
set up for regions

@@
 
+Summary end of this section is that we are
+
+1) validating other techniques (WASP14, Kurucz, etc)
+2) and two different synthetic libraries give different answers
+
+this can feed into a discussion about tweaking models.
+
 \section{Discussion}
@@
 WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.3 & -0.50 & 0.01 & Poisson only \\
+WASP-14 & 24 & & PHOENIX & & & & & & & Matern only, no regions \\
 WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\

### 2014-07-02 scripts/stars/base_lnprob.py
set up for regions

@@ -98,8 +98,9 @@ if args.perturb:
     perturb(cov_Starting, config['cov_jump'], factor=args.perturb)
 
 
+region_Starting = config['region_params']
 region_tuple = ("loga", "mu", "sigma")
-region_MH_cov = np.array([0.04, 0.02, 0.1])**2 * np.identity(len(region_tuple))
+region_MH_cov = np.array([float(config["region_jump"][key]) for key in region_tuple])**2 * np.identity(len(region_tuple))
 
 
 outdir = config['outdir']
@@ -151,8 +152,9 @@ cadenceList = []
 for i in range(len(config['orders'])):
     samplerList.append(ChebSampler(myModel, cheb_MH_cov, cheb_Starting, order_index=i, outdir=outdir))
     samplerList.append(CovGlobalSampler(myModel, cov_MH_cov, cov_Starting, order_index=i, outdir=outdir))
-    samplerList.append(RegionsSampler(myModel, region_MH_cov, max_regions=config['max_regions'], order_index=i, outdir=outdir))
-    cadenceList += [6, 6, 2]
+    samplerList.append(RegionsSampler(myModel, region_MH_cov, max_regions=config['max_regions'],
+                    default_param_dict=region_Starting, order_index=i, outdir=outdir))
+    cadenceList += [6, 6, 1]
     # cadenceList += [6, 2]
 
 mySampler = MegaSampler(myModel, samplers=[myStellarSampler] + samplerList,

### 2014-07-04 tex/paper.tex
updaded run sheets

@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    From left to right: \textbf{Class 0} The majority of spectral lines
-    ($\gtrsim 60$\%) will have minor differences in strength between data and
-    model spectrum, which produce low-amplitude correlations in the residuals
-    on the length scale of the width of a typical spectral line.  \textbf{Class
-    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-    of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-    correlated residuals of moderate amplitude.  The difficulty
-    with class III lines is that for any specific line, there might exist a
-    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure*}
-
 \subsubsection{Global Covariance Structure}
@@
 \begin{center}
-  $\vcenter{\hbox{\includegraphics{figs/class0_residuals.pdf}}}$
-  $\vcenter{\hbox{\includegraphics{figs/class0_autocorrelation.pdf}}}$
-  \caption{\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+  \includegraphics{correlation_panel.pdf}
+  \caption{\textbf{Top}: the big spectrum.
+\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
   \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines.pdf}
+  \caption{A collection of spectral lines which have imperfect model fits.
+    From left to right: \textbf{Class 0} The majority of spectral lines
+    ($\gtrsim 60$\%) will have minor differences in strength between data and
+    model spectrum, which produce low-amplitude correlations in the residuals
+    on the length scale of the width of a typical spectral line.  \textbf{Class
+    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+    of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
+    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+    correlated residuals of moderate amplitude.  The difficulty
+    with class III lines is that for any specific line, there might exist a
+    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure*}
+
 \section{Discussion}
@@
 WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.3 & -0.50 & 0.01 & Poisson only \\
-WASP-14 & 24 & & PHOENIX & & & & & & & Matern only, no regions \\
+WASP-14 & 23 & & PHOENIX & & & & & & & Matern only, no regions \\
 WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\

### 2014-07-06 StellarSpectra/covariance.pyx
PHOENIX regions

@@ -138,6 +138,7 @@ cdef class CovarianceMatrix:
     cdef GlobalCovarianceMatrix GCM
     cdef RegionList
     cdef current_region
+    cdef logPrior
 
     #Both sum_regions and partial_sum_regions are initially NULL pointers, and will be until
     #any elements in the RegionList have been declared.
@@ -161,6 +162,7 @@ cdef class CovarianceMatrix:
         self.fl = DataSpectrum.fls[self.order_index][mask]
         self.npoints = len(self.fl)
         self.logdet = 0.0
+        self.logPrior = 0.0 #neutral prior, reflects the sum of GlobalCovariance.logPrior and all region.logPrior's
 
         self.GCM = GlobalCovarianceMatrix(self.DataSpectrum, self.order_index,self.common)
         self.RegionList = []
@@ -190,6 +192,7 @@ cdef class CovarianceMatrix:
         '''
         self.GCM.update(params)
         self.update_factorization()
+        self.update_logPrior()
 
     def create_region(self, params):
         '''
@@ -242,6 +245,7 @@ cdef class CovarianceMatrix:
             #Because we are creating an extra, COPY the matrix
             self.sum_regions = cholmod_copy_sparse(get_A(region), self.c)
         self.update_factorization()
+        self.update_logPrior()
 
     def print_all_regions(self):
         string = ""
@@ -265,7 +269,6 @@ cdef class CovarianceMatrix:
         Update the sparse matrix that is the sum of all region matrices. Used before starting global
         covariance sampling.
         '''
-        #print("updating sum_regions")
         cdef cholmod_sparse *R = cholmod_copy_sparse(get_A(self.RegionList[0]), self.c)
         cdef cholmod_sparse *temp
 
@@ -318,6 +321,13 @@ cdef class CovarianceMatrix:
                 coverage = temp_ind | coverage
 
             return coverage
+
+    def update_logPrior(self):
+        '''
+        Go through all of the regions in RegionList and sum together their logPrior along 
+        with the GlobalCovariance logPrior
+        '''
+        self.logPrior = self.GCM.logPrior + np.sum([region.get_prior() for region in self.RegionList])
 
     def update_factorization(self):
         '''
@@ -372,7 +382,7 @@ cdef class CovarianceMatrix:
 
         #logdet does not depend on the residuals, so it is pre-computed
         #evaluate lnprob with logdet and chi2
[…]

### 2014-07-06 tex/paper.tex
PHOENIX regions

@@
 \vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
+\label{eqn:chebyshev}
 \end{equation}
@@
   \includegraphics{figs/chebyshev.pdf}
-  \caption{A region of the echelle spectrum at the overlap between two adjacent orders of the echelle spectrum.
-Panel 1: In green and blue, we have the flux-calibrated dataset.
-The adjacent orders overlap but there is a slight discrepancy of $\lesssim 3$\%.
-In the background in red is a model spectrum generated for reasonable values of stellar parameters, $\vt$.
-Panel 2: In order to account for residual errors in the flux calibration, we allow the model to be multiplied by a Chebyshev polynomial.
-Here are the polynomials determined as best fit, to multiply the model.
-Panels 3 \& 4: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4).
-This has the advantage of preserving the original dataset as a fixed quantity and makes the model linear in the Chebyshev polynomial coefficients.}
+  \caption{The spectrum at the overlap of two echelle orders (23 \& 24).
+\textbf{Panel 1}: The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders.
+\textbf{Panel 2}: To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model.
+\textbf{Panels 3 \& 4}: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4).
+This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.}
 \label{fig:chebyshev}
@@
 the $k^{\rm th}$ residual region is
-\begin{equation}
-  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
+\begin{multline}
+  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
+  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
 \label{eqn:klocal}
-\end{equation}
+\end{multline}
 We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
@@
 noise matrix, 
-\begin{eqnarray}
-\vC_{ij}(\vT_{\rm cov}) & = & b \, \delta_{ij}\sigma_i + \\ \nonumber
-         &   &  w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ \nonumber
-         &   &  \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
-\end{eqnarray}
+\begin{multline}
+\vC_{ij}(\vT_{\rm cov})  = b \, \delta_{ij}\sigma_i + \\
+         w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ 
+         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
+\end{multline}
 where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
@@
 
-\comm{Not much done here...want to meet and discuss before I try editing.}
-
 When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
@@
 
-\todo{We can (should?) add an additional level to the hierarchy of
-  hyperparameters and add parameters to describe the \emph{population
-  characteristics} of poorly modeled spectral lines (mostly the typical width,
-  amplitude of these lines).  This will tell us about the frequency and
-distribution of spectral modelling errors.} We explicitly fit for the
+ We explicitly fit for the
 hyperparameters of the covariance kernels at the same time we fit for the
@@
 
+We can (should?) add an additional level to the hierarchy of
+  hyperparameters and add parameters to describe the \emph{population
+  characteristics} of poorly modeled spectral lines (mostly the typical width,
+  amplitude of these lines).  This will tell us about the frequency and
+distribution of spectral modelling errors.
+
 \section{Conclusion}
@@
 WASP-14 & 22-24 & 5060-5315& Kurucz & 6512 & 32 & \multicolumn{2}{c}{\nodata} & -0.27 & 0.02 & Matern + fixed $\log g = 4.29$\\
-WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.3 & -0.50 & 0.01 & Poisson only \\
-WASP-14 & 23 & & PHOENIX & & & & & & & Matern only, no regions \\
+WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.03 & -0.50 & 0.01 & Poisson only \\
+WASP-14 & 23 & & PHOENIX & 5965 & 70 & 3.78 & 0.09 & -0.68 & 0.07 & Matern only, no regions \\
+WASP-14 & 23 & & PHOENIX & & & & & & & Matern and regions \\
 WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\

### 2014-07-06 StellarSpectra/model.py

@@ -432,7 +432,6 @@ class OrderModel:
         lnp = self.CovarianceMatrix.evaluate(model_fl)
         return lnp
 
-
 class Sampler:
     '''
     Helper class designed to be overwritten for StellarSampler, ChebSampler, CovSampler.C

### 2014-07-06 tex/paper.tex
mor eburn in

@@
 \subsubsection{Global Covariance Structure}
+\label{subsec:global_covariance}
 
@@
 \textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
+  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
 \label{fig:class0}
@@
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
+   \label{eqn:global}
 \end{equation}
@@
 \subsubsection{Local Covariance Structure}
+\label{subsec:local_covariance}
 
@@
 spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework, if we permit the kernel in Eq.~\ref{eqn:klocal} to be modified by another 
-(stationary) function.  Here we assume that function is a squared exponential,
-\begin{equation}
-  \mathcal{S}_k(\lambda_i, \lambda_j |\, h_k) = \exp \left ( - \, \frac{r_{ij}^2}{2\,h_k^2}\right ),
-\end{equation}
-where $h_k$ is a bandwidth parameter; small $h_k$ generates high-frequency structure, and vice 
-versa.  The functionality of this kernel modification (where the appropriate kernel is now $w_k \, 
-\mathcal{S}_k \, \mathcal{K}_k$) is demonstrated for the worked example presented in Section 
-\ref{subsec:mdwarf}.  
-\comm{I have to admit that I do not really understand this very well.  I think some appropriate 
-figures in Sect 3.2 will really help, though, so I'll wait for that before decided whether or not 
-we need to change this part of the text.}
+in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well. 
+\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}
 
@@
 \includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}).
+\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
 The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
@@
 Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
-In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:line}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
+In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:klocal}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
 Along with the global covariance parameters and Chebyshev parameters for this order, we call the collection of nuisance hyperparameters for a specific order  (e.g., order $1$) $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$.
@@
 
-The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, it may be the case that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
 
@@
 However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
-
-The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum.
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
@@
 The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters.
-For more on MCMC and Gibbs samplers see \citet[ch 11]{gelman13} and the references therein.
+For more on MCMC and Gibbs samplers see \citet[Ch.
+11]{gelman13} and the references therein.
 
@@
   \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev and noise parameters.
+These stellar parameters are marginalized over the Chebyshev ($\vT_{\rm cheb}$) and noise parameters ($\vT_{\rm cov}$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
@@
 \begin{center}
-  \includegraphics[width=5in]{figs/stellar_triangle_WASP14_Kurucz.png}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, logg fixed to match \citep{torres12}. Using three orders, Kurucz grid. Same assumptions.}
-\label{fig:stellar_posterior_logg}
-\end{center}
-\end{figure*}
-
-\begin{figure*}[!htb]
-\begin{center}
   \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
@@
   \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
-\textbf{Right}: The posterior probability function of the line parameters, marginalized over the stellar and global noise parameters.}
+\textbf{Right}: The posterior probability function of the parameters for a single line kernel, marginalized over the stellar and global noise parameters.}
 \label{fig:noise_posterior}
@@
       \item Sample in the global covariance parameters.
-Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}.
+Adjust the covariance matrix $C$ as described in \S\ref{subsec:global_covariance}.
 Gibbs sampler evaluates $p(\vtglobal^{i}| \vtstar^{i}, \vtcheb^{i}, \vtlines^{i -1})$.
@@
 	  \item Sample in the parameters for each line, conditional on the parameters for all the other lines.
-Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}.
+Adjust the covariance matrix $C$ as described in \S\ref{subsec:local_covariance}.
 Gibbs sampler evaluates $p(\vtline{k}^i| \vt_{\textrm{lines}\ne k}^{i - 1})$.
@@
 At each stage of the Gibbs sampler, the likelihood function (Equation~\ref{eqn:lnprob}) is evaluated for a proposal of a particular subset of parameters, conditional on the current values of all the other parameters.
-This hierarchical parameter structure and the kernel parameterization of the covariance matrix influences how the algorithm may be efficiently implemented in code.
+The hierarchical structure of the parameters influences how the algorithm may be efficiently implemented in code.
 For a typical optical spectrum with $\gtrsim$1,000 pixels, the most computationally intensive step of the likelihood evaluation is usually the matrix product $\fR^T C^{-1} \fR$.
@@
 When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update.
-However, because we designed the kernels to deliver a sparse matrix these operations are efficient enough to be used with an MCMC algorithm.
-We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.
+However, because we designed the kernels to deliver a sparse matrix these operations are still moderately efficient.
+We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and utilize the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.
 
@@
 The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
-If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and then the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
-This leads to estimates of stellar parameters with more uncertainty than necessary.
-\todo{Since this point is better carried with an example or test using the global covariance structure and not using it (see Test 1), we might want to defer this discussion to the testing section.}
+If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
 
@@
 
-Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct defects in the synthetic models.
-\todo{more delicate phrasing?} Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
-This fundamental application of machine learning would enable us to create our own library of semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
+Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
+Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
+This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
 
-\section{Worked Examples}
+\section{Examples}
 \label{sec:examples}
-Here are some ideas of which tests we might want to show.
 
@@
 
-[This may need to be eventually moved to \S2, but here it is for now] Cite papers that acquired the data.
+To illustrate the performance of our technique, we use two datasets with varying properties.
+The first is a high-resolution optical spectrum of an F-star, WASP-14.
+The second is a moderate resolution, mid-infrared spectrum of Gl 51.
+
+WASP-14 is a metal-poor F star acquired with the \emph{TRES} instrument \citep{mink11} on the 1.5m telescope at FLWO.
+The data was previously analyzed in the context of transiting exoplanet hosts by \citet{torres12}.
+Automated reduction pipeline by Buchhave.
+Has XX S/N per resolution element in the \ion{Mg}{b} order ($\sim 5100$ \AA).
+51 echelle orders spanning from XX to XX.
+
+Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}.
+CITE (Rojas-Ayala) use the H20 -K index, defined over what wl range?
+We do this as well as more general tests.
+
+The SPC analysis within \citet{torres12} focused on a region of spectrum from $5000 5300$ \AA and used many methods.
+The other methods are not directly comparable because they use different underlying assumptions.
+Additionally, there are systematics introduced by metallicity assumptions and radiative transfer.
+Using this same spectral library as SPC, we recover the same parameters.
+
 Describe the instruments.
@@
 Talk about giving more reasonable errors using regions, in the biased case.
+Show a figure of the regions determined, plotted with the increased variance envelope over the residuals.
 

### 2014-07-06 tex/paper.tex
updated text

@@
 
-To illustrate the performance of our technique, we use two datasets with varying properties.
-The first is a high-resolution optical spectrum of an F-star, WASP-14.
-The second is a moderate resolution, mid-infrared spectrum of Gl 51.
-
-WASP-14 is a metal-poor F star acquired with the \emph{TRES} instrument \citep{mink11} on the 1.5m telescope at FLWO.
-The data was previously analyzed in the context of transiting exoplanet hosts by \citet{torres12}.
-Automated reduction pipeline by Buchhave.
-Has XX S/N per resolution element in the \ion{Mg}{b} order ($\sim 5100$ \AA).
-51 echelle orders spanning from XX to XX.
-
-Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}.
-CITE (Rojas-Ayala) use the H20 -K index, defined over what wl range?
-We do this as well as more general tests.
+To demonstrate our technique, we fit spectra of two different stars with varying stellar properties and acquired with different instruments.
+The first object is WASP-14, an F5 star \citep{joshi09}.
+The spectrum was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}) \citep{mink11} on the FLWO 1.5m telescope on June 14th, 2009.
+\emph{TRES} delivers a 51 echelle spectrum spanning the full optical wavelength spectrum $3860 - 9100$ \AA\ at a resolution of $R \sim$ 44,000.
+\citet{torres12} acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts.
+The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.
+Buchhave.
+In the \ion{Mg}{0}b order ($\sim 5100$ \AA), the spectrum has a signal to noise of 150 per resolution element.
 
@@
 
-Describe the instruments.
-Describe when it was taken.
-Describe the noise properties, etc.
-Maybe a table might be worthwhile to describe the spectra.
-Here also describe the WASP-14 fit using the Kurucz spectra, using a fixed log g, and state that they are the same as Willie's paper.
-Talk about putting things in a Bayesian framework, using a sensitivity to a prior, to a $\delta$-function prior.
-Point out that this Bayesian approach allows you to do this.
+When we fit this spectrum using the same assumptions, (amounting to a $\delta$-function prior on $\log g$.
+global noise amplitude is so low, almost non-existent.
+This demonstrates sensitivity to a prior as well as the fact that surface gravity estimates from spectroscopy may be biased against fits derived from planetary transits.
+
+The second object is Gl 51, an M5 star.
+The spectrum was obtained with the \emph{SPEX} instrument at IRTF.
+The spectrum is at moderate resolution ($R \sim $ 2,000), mid-infrared spectrum from $0.7 - 5.5\mu$m.
+Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}.
+\citep{rojas-ayala12} use the H20 -K index (defined over what wl range?).
+
+A Bayesian approach allows you to explore sensitivity to prior assumptions.
 
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
+  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.}
+\label{fig:regions}
+\end{center}
+\end{figure*}
+
 Talk about giving more reasonable errors using regions, in the biased case.
@@
 
-\paragraph{Test 1}: Fitting with and without the global covariance kernel to show how the width of the posteriors nicely inflates to reasonable uncertainties.
-
-\paragraph{Test 2}: I think we should fit an optical spectrum of an F star (WASP-14) and a near-IR spectrum of an M star (Gl51).
-The WASP will be fit with both the Kurucz and PHOENIX spectra, while the M star will be fit with just the PHOENIX.
-This will show a few things
 \begin{itemize}
@@
 WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.03 & -0.50 & 0.01 & Poisson only \\
-WASP-14 & 23 & & PHOENIX & 5965 & 70 & 3.78 & 0.09 & -0.68 & 0.07 & Matern only, no regions \\
-WASP-14 & 23 & & PHOENIX & & & & & & & Matern and regions \\
+WASP-14 & 23 & & PHOENIX & 5965 & 70 & 3.78 & 0.09 & -0.68 & 0.07 & Matern only, no regions (conv?) \\
+WASP-14 & 23 & & PHOENIX & 5883 & 55 & 3.71 & 0.07 & -0.74 & 0.05 & Matern and regions (conv?) \\
 WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\

### 2014-07-17 tex/paper.tex
Redone Exploring posterior

@@
-%\documentclass[preprint]{aastex} 
-\documentclass[iop,floatfix]{emulateapj}
+\documentclass[preprint]{aastex} 
+%\documentclass[iop,floatfix]{emulateapj}
 
@@
 
-%\vt stands for ``vector theta''
-\newcommand{\vt}{ {\bm \theta}}
-\newcommand{\vT}{ {\bm \Theta}}
-\newcommand{\vg}{\vt_{\star, {\rm grid}}}
-\newcommand{\vpp}{\vt_{\star, {\rm post}}}
-\newcommand{\finst}{f_{\lambda, {\rm inst}}}
-\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
-\newcommand{\vN}{\vt_{\rm N}}
-\newcommand{\vtstar}{\vt_{\star}}
-\newcommand{\vtcheb}{\vt_{\rm Cheb}}
-\newcommand{\vtglobal}{\vt_{\rm global}}
-\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
-\newcommand{\vtorders}{\vt_{\rm orders}}
-\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
-\newcommand{\vtlines}{\vt_{\rm lines}}
-\newcommand{\vtcov}{\vt_{\rm cov}}
 \newcommand{\vM}{\mathsf{M}}
@@
 
+%flux
+\newcommand{\flam}{f_\lambda}
+
+%\vt stands for ``vector theta''
+\newcommand{\vt}{ {\bm \theta}}
+\newcommand{\vT}{ {\bm \Theta}}
+
+%\vp stands for ``vector phi''
+\newcommand{\vp}{ {\bm \phi}}
+\newcommand{\vP}{ {\bm \Phi}}
+
+\newcommand{\fundamentalParameters}{\vt_{\ast}}
+\newcommand{\gridParameters}{\{\fundamentalParameters\}^{\rm 
+grid}}
+\newcommand{\observationalParameters}{\vt_{\textrm{obs}}}
+\newcommand{\stellarParameters}{\vt}
+
+%\phi and \Phi are used for nuisance parameters
+\newcommand{\cheb}{ \vp_{\textrm{Cheb}}}
+\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}} %for signifying a specific order
+\newcommand{\Cheb}{ \vP_{\textrm{Cheb}}}
+\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} %for specifying everything but a specific order
+\newcommand{\cov}{ \vp_{\textrm{cov}}}
+\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} %for signifying a specific order
+\newcommand{\Cov}{ \vP_{\textrm{cov}}}
+\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}} %for specifying everything but a specific order
+
+\newcommand{\allParameters}{\vT} %all parameters
+\newcommand{\nuisanceParameters}{\vP} %all nuisance parameters
+
+%kernel notation
+\newcommand{\KK}{\mathcal{K}}
+\newcommand{\Kglobal}{\KK_{\textrm{g}}}
+\newcommand{\Klocal}{\KK_k}
+
 \newcommand{\todo}[1]{ \textcolor{blue}{\\TODO: #1}}
 \newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}
+
+\newcommand{\vtcheb}{\vt_{\rm Cheb}}
+\newcommand{\vtglobal}{\vt_{\rm global}}
+\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
+\newcommand{\vtorders}{\vt_{\rm orders}}
+\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
+\newcommand{\vtlines}{\vt_{\rm lines}}
+\newcommand{\vtcov}{\vt_{\rm cov}}
 
@@
 
-There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
-parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
+There are various approaches to synthesizing a spectrum, $\flam$, for a specific set of model
+parameters, $\fundamentalParameters$.  In an ideal case, a model stellar atmosphere is constructed and then 
 subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
@@
 probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
-structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
-grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
-that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{valenti96}.  A more 
+structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
+that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for SME;][]{valenti96}.  A more 
 common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
-$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
+$\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
 technically the former approach is most similar to the ideal case, the computational cost of 
@@
 points of the library.  Then, these discretized samples of the posterior probability density can be 
-interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
+interpolated to an arbitrary $\fundamentalParameters$ to construct appropriate confidence intervals 
 \citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
@@
 Here we opt to take the computationally expedient approach that employs a library of model spectra, 
-$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
+$\flam(\gridParameters)$, where $\fundamentalParameters = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
 (in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
@@
 \begin{equation}
-f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
+\flam(\gridParameters) \leadsto \flam(\fundamentalParameters),
 \label{eqn:interp}
@@
 can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
-calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
+calculated location in $\gridParameters$, and then comparing the interpolated spectrum 
 with the corresponding library spectrum.  After an extensive exploration of such calculations 
@@
 achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
-a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
+a $\gridParameters$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
 tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
@@
 
-Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
+Generally, the ``raw" model spectrum $\flam(\fundamentalParameters)$ will be highly over-sampled compared 
 to a typical observed spectrum, and does not account for several additional observational and 
@@
 post-processing in two stages: the first deals with an additional set of ``observational" 
-parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
-location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
-n}$, designed to mitigate an imperfect data calibration.
+parameters, $\observationalParameters$, that incorporate dynamical effects, geometry, and the relative 
+location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\Cheb$, designed to mitigate an imperfect data calibration.
 
-We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
-spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
+We can further divide $\observationalParameters$ into those parameters that impact the model primarily in the 
+spectral or flux dimensions; $\observationalParameters = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
 former, we consider three kernels that contribute to the observed line-of-sight velocity 
@@
 \begin{equation}
-f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
+\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) = \flam(\fundamentalParameters) \otimes \varphi_v = \flam(\fundamentalParameters) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
 \label{eqn:broadening}
@@
 \begin{equation}
-f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
+\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) \mapsto \vM(\fundamentalParameters, \vt_{{\rm obs}, v}),
 \label{eqn:resampling}
@@
 In essence, the inferences for the relevant stellar parameters will properly account for the 
-underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
-model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
+underlying uncertainty in the calibration process.  For a continuous spectrum,
+such as a long-slit spectrum or a single spectral order of an echelle spectrum,
+this is achieved in practice by distorting the model spectrum with a
+(low-degree) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
 \begin{equation}
-\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
+\vM(\stellarParameters, \cheb) = \vM(\stellarParameters) \times \sum_n c_n \, T_n,
 \label{eqn:chebyshev}
 \end{equation}
-where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
-set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
-spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
+where $T_n$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated as a
+set of nuisance (hyper-)parameters, $\cheb = \{c_0, c_1, \ldots, c_{n-1} \}$.
+For an echelle spectrum with $N$ spectral orders, there is a Chebsyhev polynomial for each order of the spectrum $i$, such that the total collection of Chebyshev parameters is $\Cheb = \{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$.
+With judicious priors on $\Cheb$, we can ensure that the
 unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
-artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-order (scaling) 
+artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-degree (scaling) 
 coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
@@
 \begin{equation}
-\vR \equiv \vR(\vt, \vT_{\rm cheb}) \equiv \vD-\vM(\vt, \vT_{\rm cheb}).
+\vR \equiv \vR(\stellarParameters, \cheb) \equiv \vD-\vM(\stellarParameters, \cheb).
 \end{equation}
@@
    \vR^\trans \vC^{-1} \vR \right ), 
-   \label{eqn:prob}
+   \label{eqn:likelihood}
 \end{equation}
@@
   \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
-  \label{eqn:lnprob}
+  \label{eqn:lnlikelihood}
 \end{equation}
@@
 \delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
-the Kronecker delta function, and Eq.~\ref{eqn:lnprob} reduces to the familiar
+the Kronecker delta function, and Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
 \begin{equation}
@@
 \begin{equation}
-  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
+  \Kglobal(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
   \label{eqn:expectation}
@@
 \begin{equation}
-  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+  \Kglobal(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
@@
 \end{equation}
-to taper the kernel (so the effective kernel is $w \, \mathcal{K}_g$).  The truncation distance 
+to taper the kernel (so the effective kernel is $w \, \Kglobal$).  The truncation distance 
 $r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
@@
 \begin{multline}
-  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
+  \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
   \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
@@
 influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
-they might induce on inferences of the interesting parameters ($\vt$).  In essence, this is a 
+they might induce on inferences of the interesting parameters ($\stellarParameters$).  In essence, this is a 
 robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the 
@@
 We can now compute the covariance matrix employed in the likelihood calculation 
-(Eq.~\ref{eqn:lnprob}) as the linear combination of these kernels and the trivial pixel-by-pixel 
-noise matrix, 
+(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel 
+noise matrix and these kernels, 
 \begin{multline}
-\vC_{ij}(\vT_{\rm cov})  = b \, \delta_{ij}\sigma_i + \\
-         w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ 
-         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
+\vC_{ij}(\cov)  = b \, \delta_{ij}\sigma_i + \\
+         w(r_{ij}|r_0 = 4\ell) \,\, \Kglobal(\lambda_i, \lambda_j|a_g, \ell) + \\ 
+         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
 \end{multline}
-where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
+where the covariance hyperparameters $\cov = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
 h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
@@
 \ref{sec:examples} for examples).
+If the echelle spectrum contains $N$ spectral orders, then the collection of covariance parameters for the whole spectrum are $\Cov = \{\covi{1}, \covi{2}, \ldots, \covi{N} \}$.
+
+If the spectrum is either long-slit or a single echelle order, then with the covariance nuisance parameters the log-likelihood (Equation~\ref{eqn:likelihood}) is now
+\begin{multline}
+  \ln{p(\vD | \vM(\stellarParameters, \cheb), \cov)} = \\
+  -\frac{1}{2} \Bigl( \vR(\stellarParameters, \cheb)^\trans \vC(\cov)^{-1} \vR(\stellarParameters, \cheb) + \\
+  \ln{\det{\vC(\cov)}} + N_{\rm pix} \ln{2\pi} \Bigr),
+\end{multline}
+If the spectrum contains multiple spectral orders, then the total log-likelihood is
+\begin{multline}
+  \ln{p(\vD | \vM(\stellarParameters, \Cheb), \Cov)} = \\
+  -\frac{1}{2} \sum_i^{N_\textrm{orders}} \Bigl( \vR_i(\stellarParameters, \chebi{i})^\trans \vC(\covi{i})^{-1} \vR_i(\stellarParameters, \chebi{i}) + \\
+  \ln{\det{\vC(\covi{i})}} + N_{ {\rm pix}, i} \ln{2\pi} \Bigr) .
+\end{multline}
 
@@
 
-When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
-Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
-In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:klocal}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
-Along with the global covariance parameters and Chebyshev parameters for this order, we call the collection of nuisance hyperparameters for a specific order  (e.g., order $1$) $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$.
-Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.
-
- We explicitly fit for the
-hyperparameters of the covariance kernels at the same time we fit for the
-stellar parameters. While including these extra parameters does increase the
-dimensionality and complexity of our model, they do provide several advantages.
-\emph{A priori} we do not know which regions of the spectrum are improperly
-modeled. The covariance hyperparameters provide a framework to identify these
-regions iteratively and in a self-consistent manner.
-
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
-When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
-This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
-However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
-
-We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}).
-The MCMC routine efficiently explores a high dimensional probability space using a stochastic, iterative approach.
-The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters.
+Using Bayes' rule, we write down the posterior probability distribution as proportional to the likelihood times the prior
+\begin{equation}
+  p(\allParameters | \vD) =  p(\stellarParameters, \Cheb, \Cov | \vD ) \propto p(\vD | \allParameters) p(\allParameters).
+\end{equation}
+Given the logical hierarchical structure of the collection of stellar and nuisance parameters, $\allParameters$, it is convenient to use a blocked Gibbs sampler with the Metropolis-Hastings algorithm to explore this posterior distribution.
+The sampler works by using the Metropolis-Hastings algorithm to sample a sub-collection of the parameters conditional on the current values of all the other parameters.
+After each iteration, the Gibbs sampler updates the values of the sub-collection, and then moves to a new sub-collection of parameters.
+By rotating through all of the sub-collections of parameters, the Gibbs sampler ensures that it samples the full posterior probability distribution.
 For more on MCMC and Gibbs samplers see \citet[Ch.
 11]{gelman13} and the references therein.
-
-In addition to the stellar parameters $\vtstar$, we have introduced several nuisance parameters for calibration and residual modeling.
-These nuisance parameters have a logical hierarchical structure.
-At the lowest level of the hierarchy, a single order of an echelle spectrum has $N$ regions of high covariance $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$.
-If we wish to evaluate the probability of the parameters of a specific line (e.g., line $1$) conditional on the current values of all other lines, then we denote this by $p(\vt_{\textrm{line} = 1} | \vt_{\textrm{lines} \ne 1})$.
-The collection of nuisance parameters for a specific order  (e.g., order $1$) is the aggregate parameter $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$.
-Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.
-Because the nuisance parameters for a single echelle order are independent from the nuisance parameters for any other echelle order, we have $p(\vt_{\textrm{order} = 1} | \vt_{\textrm{orders} \ne 1}) = p(\vt_{\textrm{order} = 1})$.
 
@@
 
-The Gibbs sampler iteratively explores this hierarchy of parameters.
-At each level of the hierarchy, corresponding to a different subset of nuisance parameters, we use the Metropolis-Hastings algorithm to propose a new subset of parameters and then either accept or reject them.
-This process of proposal and acceptance/rejection is called ``sampling.'' The Gibbs sampler rotates between sampling in different subsets of parameters, eventually sampling all of the nuisance parameters at a certain cadence.
-To initialize the MCMC algorithm, we make a reasonable guess for the starting parameters.
-We use existing spectral types in the literature for the stellar parameters, flat Chebyshev polynomials (no flux-calibration correction), and no covariance structure in the residuals (the global covariance parameters are set to zero and no line kernels are instantiated).
-We use superscripts to denote iterations of the MCMC algorithm.
-$i$ denotes the parameters from the current iteration and $i - 1$ denotes the previous iteration.
-The Gibbs sampler rotates among subsets of the parameters as follows
+The following procedure illustrates how the Gibbs sampler rotates among subsets of the parameters when sampling a multi-order echelle spectrum.
+We use superscripts to denote iterations of the MCMC algorithm, where $j$ denotes the parameters from the current iteration and $j - 1$ denotes the previous iteration.
 
 \begin{enumerate}
-  \item Sample in the stellar parameters.
-For each proposal of the Metropolis-Hastings algorithm, generate a model spectrum following the steps in \S\ref{subsec:postprocess}.
-When deciding whether or not to accept the parameters, the Gibbs sampler evaluates $p(\vtstar^i | \vtorders^{i-1})$.
-  \item For each order of the echelle spectrum
+  \item Choose starting parameters for $\allParameters^0$.
+Generally, these will be $\stellarParameters$ drawn from the literature, flat Chebyshev polynomials ($\Cheb$), and only Poisson noise ($\Cov$).
+  \item Start a new iteration of the Gibbs sampler with the parameters of the current proposal denoted with the superscript $j$ and the previous iteration denoted by $j-1$. 
+  \item Sample the stellar parameters $p(\stellarParameters^j | \Cheb^{j-1}, \Cov^{j-1})$. 
+  \item For each order of the echelle spectrum, $i$
     \begin{enumerate}
-      \item Sample in the Chebyshev polynomial parameters.
-Adjust the spectrum as detailed in \S\ref{subsec:postprocess}.Gibbs sampler evaluates $p(\vtcheb^{i} | \vtstar^{i}, \vtglobal^{i -1}, \vtlines^{i-1})$.
-      \item Sample in the global covariance parameters.
-Adjust the covariance matrix $C$ as described in \S\ref{subsec:global_covariance}.
-Gibbs sampler evaluates $p(\vtglobal^{i}| \vtstar^{i}, \vtcheb^{i}, \vtlines^{i -1})$.
-
-      \item Check to see whether the algorithm should instantiate/delete new/old line kernels
-      \item For each line kernel $k$
-	\begin{enumerate}
-	  \item Sample in the parameters for each line, conditional on the parameters for all the other lines.
-Adjust the covariance matrix $C$ as described in \S\ref{subsec:local_covariance}.
-Gibbs sampler evaluates $p(\vtline{k}^i| \vt_{\textrm{lines}\ne k}^{i - 1})$.
+      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \Chebi{i}^{j-1}, \stellarParameters^j, \Cov^{j-1} )$
+      \item Check to see whether the algorithm should instantiate/delete new/old local kernels
+      \item Sample the covariance parameters $p(\covi{i}^j | \Covi{i}^{j-1}, \stellarParameters^j, \Cheb^{j-1})$ 
     \end{enumerate}
-    \end{enumerate}
+    \item return to step 2 until Markov Chain is converged
 \end{enumerate}
 
-At each stage of the Gibbs sampler, the likelihood function (Equation~\ref{eqn:lnprob}) is evaluated for a proposal of a particular subset of parameters, conditional on the current values of all the other parameters.
 The hierarchical structure of the parameters influences how the algorithm may be efficiently implemented in code.
@@
 Once the covariance matrix is factorized, any subsequent evaluation of the matrix product for different residual vectors $\fR$ is extremely rapid.
-This makes the $\vtstar$ and $\vtcheb$ steps of the Gibbs sampler extremely fast.
+This makes the $\stellarParameters$ and $\cheb$ steps of the Gibbs sampler extremely fast.
 When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update.
@@
 Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
-This enables us to present a posterior of the stellar parameters $\vtstar$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
+This enables us to present a posterior of the stellar parameters $\stellarParameters$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
 This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
@@
 If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
+
+When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
+As demonstrated by the Gibbs sampler routine, we explicitly fit for the
+hyperparameters of the covariance kernels in parallel with the
+stellar parameters. While including these extra parameters does increase the
+dimensionality and complexity of our model, they do provide several advantages.
+\emph{A priori} we do not know which regions of the spectrum are improperly
+modeled. The covariance hyperparameters provide a framework to identify these
+regions iteratively and in a self-consistent manner.
+
+The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+
+Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
+This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
+However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
 
@@
     with class III lines is that for any specific line, there might exist a
-    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
+    $\stellarParameters$ that will fit the line, but there does not exist a $\stellarParameters$ that
     will properly fit \emph{all} the lines.}
@@
 \startdata
-$i$ & index specifying a pixel\\
 $\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\vg$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-  & that parameterize a synthetic spectrum from the grid\\
-$\vpp$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+$\fundamentalParameters$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+$\gridParameters$ & stellar parameters that specify a spectrum from a synthetic library\\
+$\observationalParameters$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
   & are applied during ``post processing'' of the synthetic spectrum\\
-$\vtstar$ & $\{\vg,\vpp \}$\\
-$\finst(\lambda)$ & data spectrum\\
-$\fsynth(\lambda)$ & synthetic spectrum\\
-$\vtcheb$ & the set of Chebyshev polynomial coefficients $\{c_0, c_1, \ldots, c_N\}$\\
-$\vtline{1}$ & \\
-$\vtlines$ & \\
-$\vtorder{1}$ & \\
-$\vtorders$ & \\
-$\vt$ & the parameters $\{\vg, \vpp, \vN\}$ that completely describe a model spectrum\\
-$\fDi$ & data flux for a given pixel, $D(\lambda_i)$\\
-$\fD$ & data vector comprised of all $\fDi$, $i = \{1, \ldots, N\}$\\
-$\fMi$ & model flux for a given pixel, $M(\lambda_i | \vt)$\\
-$\fM$ & model vector comprised of all $\fMi$, $i = \{1, \ldots, N\}$\\
-$\sigma_i$ & Poisson noise for a given pixel $i$\\
-$R_i$ & residuals $\fDi - \fMi$\\
-$\fR$ & residual vector $\fD - \fM$\\
-$C$ & covariance matrix\\
-$\sigma_{ij}$ & element in the covariance matrix\\
+$\stellarParameters$ & $\{\fundamentalParameters,\observationalParameters \}$, all stellar parameters\\
+
+$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
+ & a single spectral order $i$ \\
+$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
+  & a multi-order echelle spectrum with $N$ spectral orders  \\
+ $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
+ & except those for order $2$ \\
+$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
+$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
+ & a multi-order echelle spectrum with $N$ spectral orders  \\
+ $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
+ & except those for order $2$ \\
+
+$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
+$\vT$ & $\{\stellarParameters, \vP\}$, the collection of all parameters \\
+$\vD$ & data spectrum \\
+$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
+$\vR$ & $\vD - \vM $, residuals  \\
+$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
+$\vC_{ij}$ & an individual element in the covariance matrix\\
 $r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-$k_{\rm global}$ & global covariance kernel\\
-$k_{\rm line}$ & regional covariance kernel\\
+
+$\Kglobal$ & global covariance kernel\\
+$\Klocal$ & local covariance kernel\\
+$w$ & Hanning taper for kernels \\
 \enddata
@@
 WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.03 & -0.50 & 0.01 & Poisson only \\
+WASP-14 & 23 & & PHOENIX & & & & & & & Poisson only, masked \\
 WASP-14 & 23 & & PHOENIX & 5965 & 70 & 3.78 & 0.09 & -0.68 & 0.07 & Matern only, no regions (conv?) \\

### 2014-07-17 tex/paper.tex
sampling edition updated

@@
   \item Choose starting parameters for $\allParameters^0$.
-Generally, these will be $\stellarParameters$ drawn from the literature, flat Chebyshev polynomials ($\Cheb$), and only Poisson noise ($\Cov$).
+Generally, these will be $\stellarParameters^0$ drawn from the literature, flat Chebyshev polynomials ($\Cheb^0$), and only Poisson noise ($\Cov^0$).
   \item Start a new iteration of the Gibbs sampler with the parameters of the current proposal denoted with the superscript $j$ and the previous iteration denoted by $j-1$.
+If this is the first iteration of the Gibbs sampler, then $j - 1 = 0$.
   \item Sample the stellar parameters $p(\stellarParameters^j | \Cheb^{j-1}, \Cov^{j-1})$. 
@@
     \begin{enumerate}
-      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \Chebi{i}^{j-1}, \stellarParameters^j, \Cov^{j-1} )$
+      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \stellarParameters^j, \covi{i}^{j-1} )$
       \item Check to see whether the algorithm should instantiate/delete new/old local kernels
-      \item Sample the covariance parameters $p(\covi{i}^j | \Covi{i}^{j-1}, \stellarParameters^j, \Cheb^{j-1})$ 
+      \item Sample the covariance parameters $p(\covi{i}^j | \stellarParameters^j, \chebi{i}^{j})$ 
     \end{enumerate}
+  after step 4 is completed, all Chebyshev and covariance parameters have been updated: $\Cheb^{j-1} \rightarrow \Cheb^j$ and $\Cov^{j-1} \rightarrow \Cov^j$
     \item return to step 2 until Markov Chain is converged

### 2014-07-17 tex/paper.tex
sean's intermediate edits

--- 2014-07-06 tex/paper.tex
@@
 \title{A Method for the Spectroscopic Inference of Stellar Parameters}
-\author{Ian Czekala, Sean M. Andrews, et al.}
+\author{Ian Czekala, Sean M.~Andrews, et al.}
 \affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 
-\section{Introduction}
-\label{sec:intro}
+\section{Introduction} \label{sec:intro}
 
-A single stellar spectrum contains a vast amount of information about the physical properties of the star.
-It is of supreme interest to extract a maximal amount of this information from the spectrum in order to study the fundamental stellar parameters such as effective temperature, surface gravity, and metallicity.
-Accurate and unbiased estimates of stellar parameters are crucial to the study of numerous fields of astrophysics.
-For example, the mass, radius, and temperature of every newly discovered exoplanet (save for directly imaged planets) are measured relative to the properties of its host star.
-Large uncertainties or biases in the stellar parameters will translate into incorrect estimates of the habitable zone and the occurrence rate of earth-like planets \citep{kane14}.
-High quality stellar parameters are also necessary for the study of stellar evolution.
-Modelers of stellar evolution rely upon accurate samples of stellar parameters to chart the evolutionary tracks of stars through the Hertsprung-Russell diagram.
+All astronomers recognize that spectroscopy offers a wealth of information that can help
+characterize the properties of the observing target.  In the context of stellar astrophysics, 
+spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption 
+lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface 
+gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the 
+masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and 
+molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]), 
+facilitating study of the chemical hallmarks of different stellar populations.  With sufficient 
+spectral resolution, the velocity content of a spectrum can convey crucial information about 
+stellar rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through 
+the radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of 
+these stellar properties, there is acute interest in the exoplanet community.  There, all estimates 
+of the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host 
+mass or radius {\it ratio} is measured with the radial velocity or transit techniques, 
+respectively).  Moreover, essential clues to the planet formation process are encapsulated in the 
+dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
+\citep[e.g.,][]{fischer05,buchhave14}.   
+% SA: we ought to check with exoplanet people on these references. Can do later...
 
-However, extracting physical parameters of stars from a stellar spectrum is often difficult, due to several factors.
-First, generating synthetic stellar spectra to use as a benchmark for comparison requires a sophisticated model of the stellar atmosphere and radiative transfer, along with an extensive, high-quality database of opacity sources.
-Accurately calibrating the atomic constants in this database is a laborious task, and requires high resolution, high signal-to noise spectra for verification, which limits targets mainly to extremely bright stars like the Sun and Vega.
-This makes it difficult to achieve the same precision when extending synthetic stellar models to intrinsically fainter stars, like M dwarfs, which have additional sources of molecular opacity.
-Second,  synthesizing a spectrum with such a sophisticated model requires a large amount of computational power, such that a full-optical spectrum at high resolution will take a day or more to process on a supercomputer.
+That said, the robust and quantitative extraction of physical (or empirical) parameters from an 
+observed spectrum can be an extraordinary challenge.  Stellar models are ultimately required as a 
+comparative benchmark to associate observed spectral features with the parameters of interest.  
+Generating a synthetic model spectrum requires a complex numerical treatment of the stellar 
+structure and radiative transfer through the atmosphere \citep[e.g.,][]{kurucz93,castelli04,
+hauschildt99,husser13,paxton11}.  Detailed models calibrated to individual stars are important, but 
+rare (e.g., the Sun, Vega); as such, these stellar models are relatively untested in large swaths 
+of parameter-space.  Moreover, they necessarily include simplifications to treat complicated 
+physical processes (e.g., convection) or computational limitations (e.g., geometry, boundary 
+conditions), and often must rely on incomplete or inaccurate atomic and molecular information 
+(e.g., oscillator strengths, opacities).  In principle, the models could be further improved with 
+appropriate reference to spectroscopic datasets.  Nevertheless, they achieve a remarkable level of 
+success in reproducing many key diagnostic features in stellar spectra.
 
-Several, well-tested techniques exist that use spectra to infer stellar parameters.
-The most widely used technique relies upon measuring ratios of lines, or line indexes, which are matched to a set of indices measured from a synthetic spectrum.
-More sophisticated codes exist that synthesize spectra for direct comparison with certain well-studied lines, (e.g., \ion{Fe}{0}, \ion{Mg}{0}, and \ion{Ca}{0}) such as MOOG \citep{sneden73} and SME \citep{valenti96}.
-Other codes such as SPC \citep{buchhave12} cross-correlate the observed spectrum against a grid of pre-computed synthetic spectra and then fit a surface to the normalized cross-correlation coefficients to determine parameter uncertainties.
+There are various well-tested techniques being used to compare these models with observed spectra 
+and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
+ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
+subset of the spectral data, usually in the form of spectral line equivalent widths and/or local 
+continuum shapes.  A combined sequence of the ratios of these quantities are found to be especially 
+sensitive to one key parameter in the models, and therefore can be used to quantify it 
+\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  The line index comparison approach has 
+the major advantage of being trivially fast, but each index relationship is only informative over 
+a limited range of parameter-space.  A second technique exploits the cross-correlation of an 
+observed spectrum with a suite of model templates to identify an optimized set of parameters, 
+usually with some preferential weighting applied to specific spectral regions \citep[e.g., {\tt 
+SPC};][]{buchhave12}.  In this case, the speed advantage is maintained while taking advantage of 
+more data content (particularly in the spectral dimension), thereby achieving high precision even 
+for data with comparatively low sensitivity; the disadvantage is that the model quality and 
+parameter inferences are assessed in an empirical, rather than probabilistic, framework.  The third 
+approach of interest prefers a direct, pixel-by-pixel comparison between model and observed 
+spectra, and is usually associated with a spectral synthesis front-end (e.g., {\tt MOOG}; 
+\citealt{sneden73}; {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric 
+flexibility (e.g., one can fit for atomic abundances or arbitrary structures) and a proper 
+inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian 
+format;][]{shkedy07,schoenrich13}, but it is computationally expensive and can be biased by model 
+systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a 
+pre-selected subsample of the data.
 
-Owing to the high quality of libraries of pre-computed synthetic stellar spectra, it is now possible to generate an accurate forward model of a stellar spectrum.
-We seek to place this model in a Bayesian framework and construct a likelihood function suitable for comparing synthetic spectral models to real spectral data.
-We use a non-trivial covariance matrix, parameterized by a Gaussian process covariance kernel, to account for the residual pixel correlations that arise from a spectroscopic fit, which is a more complex problem than can be addressed with a simple $\chi^2$ best-fit metric.
-While designed for fitting stellar spectra, our method is general for all comparisons of spectra, and could be used to fit any type of spectra with a synthetic model, (e.g., galaxies or unresolved stellar clusters).
-In \S\ref{sec:method}, we describe the methodology, including a treatment of the covariance introduced by a spectroscopic fit.
-In \S\ref{sec:examples}, we present applications of our method to two examples, using optical spectra of an F type star and infrared spectra of an M star.
-In \S\ref{sec:discussion} we discuss the implications of our method and potential applications to develop data-driven models of stellar spectra.
+In this article, we design a flexible forward-modeling approach to the general spectroscopic 
+inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
+highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
+(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
+residual pixel-by-pixel correlations that arise when comparing observed spectra to intrinsically 
+imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
+systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
+observations to guide substantially improvements in the models themselves.  An overview of the 
+methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
+matrix formalism.  Some relevant tests and example applications (specifically for a high resolution 
+optical spectrum of an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of 
+the method are described in Section \ref{sec:examples}.  Finally, a discussion of the potential 
+utility of the technique, and the possibility of extending it to develop data-driven spectral 
+models, is provided in Section \ref{sec:discussion}.
 
@@
 
-All astronomers recognize that spectroscopy offers a wealth of information that can help 
-characterize the fundamental properties of the observing target.  However, as noted in Section 
-\ref{sec:intro}, the reliable and statistically robust inference of those physical (or empirical) 
-parameters from an observed spectrum can be extraordinarily challenging.  Here we describe a 
-generative Bayesian modeling framework that confronts some of the key obstacles in that process.  
-The goal of this approach is to conservatively extract the maximal amount of information about a 
-prescribed (and degenerate) parameter set by forward-modeling an observed spectrum, while also 
-recognizing and explicitly accounting for the covariances and biases introduced by pathologically 
-imperfect models or calibrations.  The method is modular, and therefore can easily incorporate 
-additional physical or nuisance parameters as desired without sacrificing an accurate reflection of 
-the limitations in the data.  Moreover, with a well-crafted observational sample, this data-driven 
-approach should ultimately enable us to systematically learn how synthetic spectral models can be 
-improved.  The specific applications discussed here are related to the spectra of individual stars, 
-but the methodology is generic (and could be used for composite spectra of unresolved stellar 
-clusters, galaxies, etc.).
+Here we describe a generative Bayesian modeling framework that confronts some of the key obstacles 
+in the spectroscopic inference problem.  The goal of this approach is to conservatively extract the 
+maximal amount of information about a prescribed (and degenerate) parameter set by forward-modeling 
+an observed spectrum, while also recognizing and explicitly accounting for the covariances and 
+biases introduced by pathologically imperfect models or calibrations.  The method is modular, and 
+therefore can easily incorporate additional physical or nuisance parameters as desired without 
+sacrificing an accurate reflection of the limitations in the data.  Moreover, with a well-crafted 
+observational sample, this data-driven approach should ultimately enable us to systematically learn 
+how synthetic spectral models can be improved.  The specific applications discussed here are 
+related to the spectra of individual stars, but the methodology is generic (and could be used for 
+composite spectra of unresolved stellar clusters, galaxies, etc.).
 
@@
 grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
-that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{valenti96}.  A more 
-common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
-$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
-technically the former approach is most similar to the ideal case, the computational cost of 
-repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
-(particularly for data with a large spectral range) considerably less appealing.  A related, but 
-different approach is to eschew forward modeling entirely (and therefore repeated spectral 
-syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
-points of the library.  Then, these discretized samples of the posterior probability density can be 
-interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
-\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
-parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
-that this interpolation might not accurately recover intrinsic parameter degeneracies.
+that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for {\tt 
+SME};][]{valenti96}.  A more common variant is to instead rely on interpolation over a library of 
+pre-synthesized model spectra, $f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,
+allard12,husser13}.  While technically the former approach is most similar to the ideal case, the 
+computational cost of repeated spectral synthesis is sufficiently high to make a detailed 
+exploration of parameter space (particularly for data with a large spectral range) considerably 
+less appealing.  A related, but different approach is to eschew forward modeling entirely (and 
+therefore repeated spectral syntheses and/or library interpolations), and instead evaluate the 
+models only at the discrete grid points of the library.  Then, these discretized samples of the 
+posterior probability density can be interpolated to an arbitrary $\vt_{\ast}$ to construct 
+appropriate confidence intervals \citep[similar to the method of {\tt SPC};][]{buchhave12}.  The 
+difficulty with this latter approach is that the parameter uncertainties can be smaller than the 
+grid spacing; in that case, there is valid concern that this interpolation might not accurately 
+recover intrinsic parameter degeneracies.
 
@@
 $f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
-(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
-it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
-generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
+(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
+that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
+generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
 $\vt_{\ast}$ must be interpolated from among the spectral library, 
@@
 Section~\ref{subsec:likelihood}).  
+\comm{I see nothing was done here regarding your updated approach; how should we handle that?}
 
-\subsection{Post-Processing}
-\label{subsec:postprocess}
+\subsection{Post-Processing} \label{subsec:postprocess}
 
-Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
-to a typical observed spectrum, and does not account for several additional observational and 
+Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ is highly over-sampled compared to a 
+typical observed spectrum, and does not account for several additional observational and 
 instrumental effects that become important in comparisons with real data.  Therefore, a certain 
@@
 We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
-spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
-former, we consider three kernels that contribute to the observed line-of-sight velocity 
+spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{{\rm obs,} v}, \,\, \vt_{{\rm obs,} f}]$.  For 
+the former, we consider three kernels that contribute to the observed line-of-sight velocity 
 distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
@@
 parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
-a convolution in velocity-space,
+a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
+multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
+the mathematical formalism is presented here for clarity.}
 \begin{equation}
@@
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
 \begin{center}
@@
 radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
-account for interstellar extinction, assuming a 
-derived extinction law $A_{\lambda}$ \citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  
-The parameters $\vt_{{\rm obs}, f} = [\Omega, \,\, A_V]$ are applied as
+account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
+\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $\vt_{{\rm obs}, f} = 
+[\Omega, \,\, A_V]$ are applied as
 \begin{equation}
@@
 
-We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
+We adopt a somewhat analogous approach to deal with this issue, but cast it {\it internal} to the 
 modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
@@
 rather than the residual calibration mismatches as described here (with a suitable relaxation of 
-the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
-these nuisance parameters are applied.
+the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of 
+how these nuisance parameters are applied.
 
@@
 
-\subsection{Model Evaluation}
-\label{subsec:likelihood}
+\subsection{Model Evaluation} \label{subsec:likelihood}
 
 The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
-likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
-$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
+likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
+spectrum (of $N_{\rm pix}$ pixels) can be defined for any input parameter set,
 \begin{equation}
@@
 \begin{equation}
-p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
+p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
    \vR^\trans \vC^{-1} \vR \right ), 
@@
 
-The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; 
-``noise") in each pixel and the intrinsic covariance between pixels.  The special case where each 
-pixel represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
+The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; ``noise") in 
+each pixel and the intrinsic covariance between pixels.  The special case where each pixel 
+represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
 \delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
@@
 
-\subsubsection{Global Covariance Structure}
-\label{subsec:global_covariance}
+\subsubsection{Global Covariance Structure} \label{subsec:global_covariance}
 
@@
 
-For a well-designed spectrograph and sufficiently accurate model spectrum, this
-{\it global} (i.e., present throughout the spectrum) covariant structure should
-have a relatively low amplitude and small correlation length.  To describe that
-structure, we assume a stationary covariance kernel (also called a radial basis
-function) with an amplitude that depends only on the velocity distance between
-two pixels, 
+For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
+throughout the spectrum) covariant structure should have a relatively low amplitude and small 
+correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
+basis function) with an amplitude that depends only on the velocity distance between two pixels, 
 \begin{equation}
@@
 \end{equation}
-A variety of kernels have been used in the field of Gaussian
-processes to parameterize such a mildly covariant structure
-\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
-$\nu = 3/2$,
+A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
+mildly covariant structure \citep[e.g.,][]{rasmussen05}.  We adopt the \matern\ kernel with $\nu = 
+3/2$,
 \begin{equation}
@@
 
-\subsubsection{Local Covariance Structure}
-\label{subsec:local_covariance}
+\subsubsection{Local Covariance Structure} \label{subsec:local_covariance}
 
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines.pdf}
+  \caption{A collection of spectral lines which have imperfect model fits.
+    From left to right: \textbf{Class 0} The majority of spectral lines
+    ($\gtrsim 60$\%) will have minor differences in strength between data and
+    model spectrum, which produce low-amplitude correlations in the residuals
+    on the length scale of the width of a typical spectral line.  \textbf{Class
+    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+    of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
+    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+    correlated residuals of moderate amplitude.  The difficulty
+    with class III lines is that for any specific line, there might exist a
+    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure*}
+
 Assuming that these local residuals are produced primarily by pathological differences in the 
@@
 the $k^{\rm th}$ residual region is
-\begin{multline}
-  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
+\begin{equation}
+  \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = 
   \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
 \label{eqn:klocal}
-\end{multline}
+\end{equation}
 We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
 efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
-to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  
-Figure \ref{fig:region} demonstrates how this non-stationary Gaussian 
-kernel generates a localized region of enhanced variance that successfully mimics the kind of 
-residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
-influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
-they might induce on inferences of the interesting parameters ($\vt$).  In essence, this is a 
-robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the 
-integrity of the probabilistic framework being developed (as opposed to manual or threshold-based 
-clipping/masking).
-
-These local kernels can be further modified to account for more complex residual structures.  For 
-example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
-pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
-spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well. 
-\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}
-
-\subsubsection{Composite Covariance Matrix}
-
-We can now compute the covariance matrix employed in the likelihood calculation 
-(Eq.~\ref{eqn:lnprob}) as the linear combination of these kernels and the trivial pixel-by-pixel 
-noise matrix, 
-\begin{multline}
-\vC_{ij}(\vT_{\rm cov})  = b \, \delta_{ij}\sigma_i + \\
-         w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ 
-         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
-\end{multline}
-where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
-h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
-(see below for details on how this is determined), and $b$ is assumed to be a
-fixed parameter that scales up the pixel noise values to account for read
-noise, noise added during the 2D to 1D spectral extraction procedure, and
-interpolation uncertainties (see Section \ref{subsec:synthetic}; reasonable
-values are $b \approx 1.02$--1.10 for well-calibrated optical spectra; see Section
-\ref{sec:examples} for examples).
+to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  Figure \ref{fig:region} 
+demonstrates how this non-stationary Gaussian kernel generates a localized region of enhanced 
+variance that successfully mimics the kind of residuals produced by an inaccurate spectral line 
+model.  These kernels down-weight the influence of such strong residuals in the likelihood 
+calculation, mitigating any potential bias they might induce on inferences of the interesting 
+parameters ($\vt$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
+outlier rejection that preserves the integrity of the probabilistic framework being developed (as 
+opposed to manual or threshold-based clipping or masking).
 
@@
 
-\subsection{Exploring the Posterior }
-\label{subsec:MCMC}
+\comm{These local kernels can be further modified to account for more complex residual structures.  For 
+example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
+pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
+spectral scale.  This phenomenologically different local covariance behavior can still be treated 
+in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.}
+\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}
 
-When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+\subsubsection{Composite Covariance Matrix}
+
+We can now compute the covariance matrix employed in the likelihood calculation 
+(Eq.~\ref{eqn:lnprob}) as the linear combination of these kernels and the trivial pixel-by-pixel 
+noise matrix, 
+\begin{eqnarray}
+\vC_{ij}(\vT_{\rm cov}) & = & b \, \delta_{ij}\sigma_i + \\ \nonumber
+         &   &  w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ \nonumber
+         &   &  \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
+\end{eqnarray}
+where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
+h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions (see below for details 
+on how this is determined), and $b$ is a fixed parameter that scales the pixel noise values to 
+account for additional error (e.g., introduced by read noise, order extraction, and the grid 
+interpolation noted in Section \ref{subsec:synthetic}; reasonable values are $b \approx 1.02$--1.10 
+for well-calibrated optical spectra, and examples are provided in Section \ref{sec:examples}).
+
+\subsection{Exploring the Posterior} \label{subsec:MCMC}
+
+%When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
 Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
@@
 
- We explicitly fit for the
-hyperparameters of the covariance kernels at the same time we fit for the
-stellar parameters. While including these extra parameters does increase the
-dimensionality and complexity of our model, they do provide several advantages.
-\emph{A priori} we do not know which regions of the spectrum are improperly
-modeled. The covariance hyperparameters provide a framework to identify these
-regions iteratively and in a self-consistent manner.
+% We explicitly fit for the
+%hyperparameters of the covariance kernels at the same time we fit for the
+%stellar parameters. While including these extra parameters does increase the
+%dimensionality and complexity of our model, they do provide several advantages.
+%\emph{A priori} we do not know which regions of the spectrum are improperly
+%modeled. The covariance hyperparameters provide a framework to identify these
+%regions iteratively and in a self-consistent manner.
 
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+%The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 
-We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}).
+%We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}).
 The MCMC routine efficiently explores a high dimensional probability space using a stochastic, iterative approach.
@@
 
-In addition to the stellar parameters $\vtstar$, we have introduced several nuisance parameters for calibration and residual modeling.
+%In addition to the stellar parameters $\vtstar$, we have introduced several nuisance parameters for calibration and residual modeling.
 These nuisance parameters have a logical hierarchical structure.
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
+%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
 These stellar parameters are marginalized over the Chebyshev ($\vT_{\rm cheb}$) and noise parameters ($\vT_{\rm cov}$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:stellar_posterior}
-\end{center}
-\end{figure*}
+%}
+%\label{fig:stellar_posterior}
+%\end{center}
+%\end{figure*}
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
-  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
-  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
+%  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
+%  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
 \textbf{Right}: The posterior probability function of the parameters for a single line kernel, marginalized over the stellar and global noise parameters.}
-\label{fig:noise_posterior}
-\end{center}
-\end{figure*}
+%\label{fig:noise_posterior}
+%\end{center}
+%\end{figure*}
+
+To explore the posterior probability distribution and make inferences on the parameters of 
+interest, we employ a hierarchical Monte Carlo Markov Chain (MCMC) approach that nests the standard 
+Metropolis-Hastings algorithm (for sampling the parameters $\vt$) inside the Gibbs sampler (which 
+rotates through subsets of the hyper-parameters $\vT$).
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    From left to right: \textbf{Class 0} The majority of spectral lines
-    ($\gtrsim 60$\%) will have minor differences in strength between data and
-    model spectrum, which produce low-amplitude correlations in the residuals
-    on the length scale of the width of a typical spectral line.  \textbf{Class
-    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-    of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-    correlated residuals of moderate amplitude.  The difficulty
-    with class III lines is that for any specific line, there might exist a
-    $\vtstar$ that will fit the line, but there does not exist a $\vtstar$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure*}
-
 \section{Discussion}

### 2014-07-17 tex/paper.tex

--- 2014-07-17 tex/paper.tex
@@
 
+%\vt stands for ``vector theta''
+\newcommand{\vt}{ {\bm \theta}}
+\newcommand{\vT}{ {\bm \Theta}}
+\newcommand{\vg}{\vt_{\star, {\rm grid}}}
+\newcommand{\vpp}{\vt_{\star, {\rm post}}}
+\newcommand{\finst}{f_{\lambda, {\rm inst}}}
+\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
+\newcommand{\vN}{\vt_{\rm N}}
+\newcommand{\vtstar}{\vt_{\star}}
+\newcommand{\vtcheb}{\vt_{\rm Cheb}}
+\newcommand{\vtglobal}{\vt_{\rm global}}
+\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
+\newcommand{\vtorders}{\vt_{\rm orders}}
+\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
+\newcommand{\vtlines}{\vt_{\rm lines}}
+\newcommand{\vtcov}{\vt_{\rm cov}}
 \newcommand{\vM}{\mathsf{M}}
@@
 
-There are various approaches to synthesizing a spectrum, $\flam$, for a specific set of model
-parameters, $\fundamentalParameters$.  In an ideal case, a model stellar atmosphere is constructed and then 
+There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
+parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
 subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
@@
 structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
-that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for SME;][]{valenti96}.  A more 
-common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
-$\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
-technically the former approach is most similar to the ideal case, the computational cost of 
-repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
-(particularly for data with a large spectral range) considerably less appealing.  A related, but 
-different approach is to eschew forward modeling entirely (and therefore repeated spectral 
-syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
-points of the library.  Then, these discretized samples of the posterior probability density can be 
-interpolated to an arbitrary $\fundamentalParameters$ to construct appropriate confidence intervals 
-\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
-parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
-that this interpolation might not accurately recover intrinsic parameter degeneracies.
+that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt 
+SME};][]{valenti96}.  A more common variant is to instead rely on interpolation over a library of 
+pre-synthesized model spectra, $\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,
+allard12,husser13}.  While technically the former approach is most similar to the ideal case, the 
+computational cost of repeated spectral synthesis is sufficiently high to make a detailed 
+exploration of parameter space (particularly for data with a large spectral range) considerably 
+less appealing.  A related, but different approach is to eschew forward modeling entirely (and 
+therefore repeated spectral syntheses and/or library interpolations), and instead evaluate the 
+models only at the discrete grid points of the library.  Then, these discretized samples of the 
+posterior probability density can be interpolated to an arbitrary $\fundamentalParameters$ to construct 
+appropriate confidence intervals \citep[similar to the method of {\tt SPC};][]{buchhave12}.  The 
+difficulty with this latter approach is that the parameter uncertainties can be smaller than the 
+grid spacing; in that case, there is valid concern that this interpolation might not accurately 
+recover intrinsic parameter degeneracies.
 
 Here we opt to take the computationally expedient approach that employs a library of model spectra, 
-$\flam(\gridParameters)$, where $\fundamentalParameters = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
-(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
-it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
-generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
+$\flam(\gridParameters)$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
+(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
+that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
+generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
 $\vt_{\ast}$ must be interpolated from among the spectral library, 
@@
 can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
-calculated location in $\gridParameters$, and then comparing the interpolated spectrum 
+calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
 with the corresponding library spectrum.  After an extensive exploration of such calculations 
@@
 achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
-a $\gridParameters$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
+a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
 tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
@@
 Section~\ref{subsec:likelihood}).  
+\comm{I see nothing was done here regarding your updated approach; how should we handle that?}
 
-\subsection{Post-Processing}
-\label{subsec:postprocess}
+\subsection{Post-Processing} \label{subsec:postprocess}
 
@@
 efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
-to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  
-Figure \ref{fig:region} demonstrates how this non-stationary Gaussian 
-kernel generates a localized region of enhanced variance that successfully mimics the kind of 
-residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
-influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
-they might induce on inferences of the interesting parameters ($\stellarParameters$).  In essence, this is a 
-robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the 
-integrity of the probabilistic framework being developed (as opposed to manual or threshold-based 
-clipping/masking).
+to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  Figure \ref{fig:region} 
+demonstrates how this non-stationary Gaussian kernel generates a localized region of enhanced 
+variance that successfully mimics the kind of residuals produced by an inaccurate spectral line 
+model.  These kernels down-weight the influence of such strong residuals in the likelihood 
+calculation, mitigating any potential bias they might induce on inferences of the interesting 
+parameters ($\stellarParameters$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
+outlier rejection that preserves the integrity of the probabilistic framework being developed (as 
+opposed to manual or threshold-based clipping or masking).
 
-These local kernels can be further modified to account for more complex residual structures.  For 
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
+The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
+These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
+\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
+The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
+\label{fig:region}
+\end{center}
+\end{figure*}
+
+\comm{These local kernels can be further modified to account for more complex residual structures.  For 
 example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
@@
 spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well. 
+in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.}
 \todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}
@@
 
+\subsection{Exploring the Posterior} \label{subsec:MCMC}
+
 \subsection{Exploring the Posterior }
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
-  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
-  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters.
-\textbf{Right}: The posterior probability function of the parameters for a single line kernel, marginalized over the stellar and global noise parameters.}
-\label{fig:noise_posterior}
-\end{center}
-\end{figure*}
+To explore the posterior probability distribution and make inferences on the parameters of 
+interest, we employ a hierarchical Monte Carlo Markov Chain (MCMC) approach that nests the standard 
+Metropolis-Hastings algorithm (for sampling the parameters $\vt$) inside the Gibbs sampler (which 
+rotates through subsets of the hyper-parameters $\vT$).
 
@@
 \startdata
+$i$ & index specifying a pixel\\
 $\lambda_i$ & wavelength corresponding to a given pixel $i$\\

### 2014-07-17 tex/paper.tex
Merged SA intermediate comments and IC 2.4 text

@@
 \newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}
-
-\newcommand{\vtcheb}{\vt_{\rm Cheb}}
-\newcommand{\vtglobal}{\vt_{\rm global}}
-\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
-\newcommand{\vtorders}{\vt_{\rm orders}}
-\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
-\newcommand{\vtlines}{\vt_{\rm lines}}
-\newcommand{\vtcov}{\vt_{\rm cov}}
 
@@
 
-\subsection{Exploring the Posterior }
-\label{subsec:MCMC}
-
 Using Bayes' rule, we write down the posterior probability distribution as proportional to the likelihood times the prior
@@
   \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\vT_{\rm cheb}$) and noise parameters ($\vT_{\rm cov}$).
+These stellar parameters are marginalized over the Chebyshev ($\Cheb$) and noise parameters ($\Cov$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
@@
 \end{figure*}
-
-To explore the posterior probability distribution and make inferences on the parameters of 
-interest, we employ a hierarchical Monte Carlo Markov Chain (MCMC) approach that nests the standard 
-Metropolis-Hastings algorithm (for sampling the parameters $\vt$) inside the Gibbs sampler (which 
-rotates through subsets of the hyper-parameters $\vT$).
 
@@
 \startdata
-$i$ & index specifying a pixel\\
 $\lambda_i$ & wavelength corresponding to a given pixel $i$\\

### 2014-07-21 tex/paper.tex
updated sect 2, some notation

@@
 
-\newcommand{\fundamentalParameters}{\vt_{\ast}}
-\newcommand{\gridParameters}{\{\fundamentalParameters\}^{\rm 
-grid}}
-\newcommand{\observationalParameters}{\vt_{\textrm{obs}}}
-\newcommand{\stellarParameters}{\vt}
+% Ian, this was ridiculous...don't type out giant words that are longer than the required tex.  It 
+% actually takes a lot more time and is a huge pain.  
+%\newcommand{\fundamentalParameters}{\vt_{\ast}}
+%\newcommand{\gridParameters}{\{\fundamentalParameters\}^{\rm 
+%grid}}
+%\newcommand{\observationalParameters}{\vt_{\textrm{obs}}}
+%\newcommand{\stellarParameters}{\vt}
 
 %\phi and \Phi are used for nuisance parameters
-\newcommand{\cheb}{ \vp_{\textrm{Cheb}}}
+\newcommand{\cheb}{ \vp_{\mathsf{P}}}
 \newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}} %for signifying a specific order
@@
 \newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} %for specifying everything but a specific order
-\newcommand{\cov}{ \vp_{\textrm{cov}}}
+\newcommand{\cov}{ \vp_{\mathsf{C}}}
 \newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} %for signifying a specific order
@@
 \newcommand{\Kglobal}{\KK_{\textrm{g}}}
-\newcommand{\Klocal}{\KK_k}
+\newcommand{\Klocal}{\KK_l}
 
@@
 of the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host 
-mass or radius {\it ratio} is measured with the radial velocity or transit techniques, 
+mass or radius {\it ratio} is constrained with the radial velocity or transit techniques, 
 respectively).  Moreover, essential clues to the planet formation process are encapsulated in the 
@@
 covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
-under the guise of hierarchical Monte Carlo Markov Chain (MCMC) simulations to numerically explore 
+under the guise of hierarchical Markov Chain Monte Carlo (MCMC) simulations to numerically explore 
 the posterior probability density of the model conditioned on the data, and thereby to determine 
@@
 probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
-structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
-that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt 
-SME};][]{valenti96}.  A more common variant is to instead rely on interpolation over a library of 
-pre-synthesized model spectra, $\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,
-allard12,husser13}.  While technically the former approach is most similar to the ideal case, the 
-computational cost of repeated spectral synthesis is sufficiently high to make a detailed 
-exploration of parameter space (particularly for data with a large spectral range) considerably 
-less appealing.  A related, but different approach is to eschew forward modeling entirely (and 
-therefore repeated spectral syntheses and/or library interpolations), and instead evaluate the 
-models only at the discrete grid points of the library.  Then, these discretized samples of the 
-posterior probability density can be interpolated to an arbitrary $\fundamentalParameters$ to construct 
-appropriate confidence intervals \citep[similar to the method of {\tt SPC};][]{buchhave12}.  The 
-difficulty with this latter approach is that the parameter uncertainties can be smaller than the 
-grid spacing; in that case, there is valid concern that this interpolation might not accurately 
-recover intrinsic parameter degeneracies.
+structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
+grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
+that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt SME};][]{valenti96}.  A 
+more common variant is to instead rely on interpolation over a library of pre-synthesized model 
+spectra, $\flam(\vt_{\ast})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While technically the 
+former approach is most similar to the ideal case, the computational cost of repeated spectral 
+synthesis is sufficiently high to make a detailed exploration of parameter space (particularly for 
+data with a large spectral range) considerably less appealing.  A related, but different approach 
+is to eschew forward modeling entirely (and therefore repeated spectral syntheses and/or library 
+interpolations), and instead evaluate the models only at the discrete grid points of the library.  
+Then, these discretized samples of the posterior probability density can be interpolated to an 
+arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals \citep[similar to the method 
+of {\tt SPC};][]{buchhave12}.  The difficulty with this latter approach is that the parameter 
+uncertainties can be smaller than the grid spacing; in that case, there is valid concern that this 
+interpolation might not accurately recover intrinsic parameter degeneracies.
 
 Here we opt to take the computationally expedient approach that employs a library of model spectra, 
-$\flam(\gridParameters)$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
+$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
 (in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
@@
 \begin{equation}
-\flam(\gridParameters) \leadsto \flam(\fundamentalParameters),
+\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
 \label{eqn:interp}
@@
 Section~\ref{subsec:likelihood}).
-\comm{I see nothing was done here regarding your updated approach; how should we handle that?}
 
@@
 
-Generally, the ``raw" model spectrum $\flam(\fundamentalParameters)$ will be highly over-sampled compared 
-to a typical observed spectrum, and does not account for several additional observational and 
+Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ will be highly over-sampled compared to a 
+typical observed spectrum, and does not account for several additional observational and 
 instrumental effects that become important in comparisons with real data.  Therefore, a certain 
@@
 post-processing in two stages: the first deals with an additional set of ``observational" 
-parameters, $\observationalParameters$, that incorporate dynamical effects, geometry, and the relative 
-location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\Cheb$, designed to mitigate an imperfect data calibration.
+parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
+location of the target, while the second employs a suite of nuisance (hyper-)parameters ($\vp$) 
+designed to mitigate an imperfect data calibration.
 
-We can further divide $\observationalParameters$ into those parameters that impact the model primarily in the 
-spectral or flux dimensions; $\observationalParameters = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
+We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
+spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
 former, we consider three kernels that contribute to the observed line-of-sight velocity 
@@
 \begin{equation}
-\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) = \flam(\fundamentalParameters) \otimes \varphi_v = \flam(\fundamentalParameters) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
+\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) = \flam(\vt_{\ast}) \otimes \varphi_v = \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
 \label{eqn:broadening}
@@
 \begin{equation}
-\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) \mapsto \vM(\fundamentalParameters, \vt_{{\rm obs}, v}),
+\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
 \label{eqn:resampling}
@@
 where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
-$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
-For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
+$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
+spectrum).  For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
 post-processing steps.
@@
 \begin{equation}
-\vM(\vt) \equiv \vM(\vt_{\ast}, \vt_{\rm obs}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
+\vM(\vT) \equiv \vM(\vt_{\ast}, \vt_{{\rm obs},v}, \vt_{{\rm obs},f}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
 \label{eqn:scaling}
 \end{equation}
-where we have simplified the notation by writing $\vt \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.
+where we have simplified the notation by writing $\vT \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.
 
@@
 the observed spectrum was {\it perfectly} calibrated, we could proceed to a likelihood calculation 
-(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vt)$ at this point.  
+(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vT)$ at this point.  
 However, that is unlikely to be the case.  The primary concern is that an imperfect calibration 
@@
 that ``normalization" procedure implicitly assumes that there is no relevant information content on 
-those mismatched scales; if $\vt$ also contributes to the broad spectral shape, then adopting this 
+those mismatched scales; if $\vT$ also contributes to the broad spectral shape, then adopting this 
 approach will corrupt the inferences of these parameters.  Moreover, in practice this approach is 
@@
 In essence, the inferences for the relevant stellar parameters will properly account for the 
-underlying uncertainty in the calibration process.  For a continuous spectrum,
-such as a long-slit spectrum or a single spectral order of an echelle spectrum,
-this is achieved in practice by distorting the model spectrum with a
-(low-degree) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
+underlying uncertainty in the calibration process.  In practice, this is achieved by distorting 
+segments of the model spectrum with polynomials, $\mathsf{P}$ \citep[e.g.,][]{eisenstein06,
+koleva09}.  For data with $N_{\rm ord}$ spectral orders, each denoted with index $o$ (where $o$ 
+naturally has an implicit $\lambda$-dependence), we can write
 \begin{equation}
-\vM(\stellarParameters, \cheb) = \vM(\stellarParameters) \times \sum_n c_n \, T_n,
+\vM(\vT, \cheb) = \vM(\vT) \times \{\mathsf{P}_o\} \equiv \vM(\vT) \times \left\{ \sum_n c_o^{(n)} \, T_o^{(n)} \right\},
 \label{eqn:chebyshev}
 \end{equation}
-where $T_n$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated as a
-set of nuisance (hyper-)parameters, $\cheb = \{c_0, c_1, \ldots, c_{n-1} \}$.
-For an echelle spectrum with $N$ spectral orders, there is a Chebsyhev polynomial for each order of the spectrum $i$, such that the total collection of Chebyshev parameters is $\Cheb = \{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$.
-With judicious priors on $\Cheb$, we can ensure that the
-unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
-artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-degree (scaling) 
-coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
-Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
-unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
-model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
-the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
-formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
-rather than the residual calibration mismatches as described here (with a suitable relaxation of 
-the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of 
-how these nuisance parameters are applied.
+where $T^{(n)}$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated 
+as a set of $n \times N_{\rm ord}$ nuisance (hyper-)parameters, $\cheb = [\{\vp_{\mathsf{P}_o}\}] = 
+[\{c_o^{(0)}, c_o^{(1)}, \ldots, c_o^{(n-1)} \}]$.  With judicious priors on $\cheb$, we can ensure 
+that the unintended treatment of real spectral features (e.g., broad, deep molecular bands) as 
+calibration artifacts is negligible (see Section~\ref{sec:examples} for examples).  The 
+lowest-degree (scaling) coefficient, $c^{(0)}$, is by its nature degenerate with the solid angle 
+parameter, $\Omega$.  Therefore, we enforce an additional constraint by requiring that the mean of 
+the polynomial is unity.  For data with a single spectral order, this means simply setting $c^{(0)} 
+= 1$.  If the goal is to model multiple spectral orders, we assign $c^{(0)} = 1$ in an arbitrary 
+order as an anchor, but permit the $c^{(0)}$ values in other orders to be different (as 
+necessary).  It is worth noting that this formalism can, in principle, be extended to develop 
+models for completely uncalibrated spectra, rather than the residual calibration mismatches as 
+described here (with a suitable relaxation of the priors on $\cheb$).  Figure \ref{fig:chebyshev} 
+offers a practical demonstration of how these nuisance parameters are applied.
 
@@
 likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
-spectrum (of $N_{\rm pix}$ pixels) can be defined for any input parameter set,
+spectrum (vector of $N_{\rm pix}$ elements) can be defined for any input parameter set,
 \begin{equation}
-\vR \equiv \vR(\stellarParameters, \cheb) \equiv \vD-\vM(\stellarParameters, \cheb).
+\vR \equiv \vR(\vT, \cheb) \equiv \vD-\vM(\vT, \cheb).
 \end{equation}
@@
 residuals as manifestations of the still-imperfect model library are required to avoid biasing our 
-inferences of the physically interesting parameters ($\vt$).  The following subsections describe 
+inferences of the physically interesting parameters ($\vT$).  The following subsections describe 
 how these issues are addressed in the practical implementation of $\vC$.
@@
 \begin{equation}
-  \Kglobal(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+  \Kglobal(\lambda_i, \lambda_j \, |\, \vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
@@
 \end{equation}
-which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
+with the global covariance (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ are 
+characterized by an amplitude ($a_g$) and scale ($\ell$) designed to smoothly transition to 
 negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
@@
 with an amplitude $a_k$, a mean wavelength $\mu_k$, and a width $\sigma_k$.  Following 
-Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to 
-the $k^{\rm th}$ residual region is
-\begin{multline}
-  \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k) = \\
-  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ].
+Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to a 
+sequence of residual regions is the linear combination
+\begin{equation}
+  \Klocal(\lambda_i, \lambda_j | \vp_{{\mathsf C}, l}) = \sum_k
+  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
 \label{eqn:klocal}
-\end{multline}
+\end{equation}
+with the set of local covariance (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, 
+\sigma_k\}]$.
 We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
@@
 calculation, mitigating any potential bias they might induce on inferences of the interesting 
-parameters ($\stellarParameters$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
+parameters ($\vT$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
 outlier rejection that preserves the integrity of the probabilistic framework being developed (as 
@@
 
-\begin{figure*}[!htb]
+\begin{figure*}[!t]
 \begin{center}
@@
 We can now compute the covariance matrix employed in the likelihood calculation 
-(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel 
-noise matrix and these kernels, 
-\begin{multline}
-\vC_{ij}(\cov)  = b \, \delta_{ij}\sigma_i + \\
-         w(r_{ij}|r_0 = 4\ell) \,\, \Kglobal(\lambda_i, \lambda_j|a_g, \ell) + \\ 
-         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
-\end{multline}
-where the covariance hyperparameters $\cov = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
-h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
-(see below for details on how this is determined), and $b$ is assumed to be a
-fixed parameter that scales up the pixel noise values to account for read
-noise, noise added during the 2D to 1D spectral extraction procedure, and
-interpolation uncertainties (see Section \ref{subsec:synthetic}; reasonable
-values are $b \approx 1.02$--1.10 for well-calibrated optical spectra; see Section
-\ref{sec:examples} for examples).
-If the echelle spectrum contains $N$ spectral orders, then the collection of covariance parameters for the whole spectrum are $\Cov = \{\covi{1}, \covi{2}, \ldots, \covi{N} \}$.
+(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
+and these kernels, 
+\begin{eqnarray}
+\vC_{ij}(\cov)  &=& b \, \delta_{ij}\sigma_i + \\ \nonumber
+                &=& w(r_{ij}|r_0 = 4\ell) \,\, \Kglobal(\lambda_i, \lambda_j|a_g, \ell) + \\ \nonumber
+                &=& \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
+\end{eqnarray}
+\footnote{\comm{this needs adjustment for notation, pending what we do in the local covariance kernels for M dwarfs.}}
+where the covariance (hyper-)parameters are $\cov = [a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}]$, 
+and $b$ is assumed to be a fixed parameter that scales up the pixel noise values to account for 
+additional detector or data reduction uncertainties (e.g., read noise, uncertainties in the 
+spectral extraction procedure, and interpolation errors); reasonable values are $b \approx 
+1.02$--1.10 for well-calibrated optical spectra (see Section \ref{sec:examples} for examples).  If 
+there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
+determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance (hyper-)parameters, 
+$\cov$.
 
-If the spectrum is either long-slit or a single echelle order, then with the covariance nuisance parameters the log-likelihood (Equation~\ref{eqn:likelihood}) is now
-\begin{multline}
-  \ln{p(\vD | \vM(\stellarParameters, \cheb), \cov)} = \\
-  -\frac{1}{2} \Bigl( \vR(\stellarParameters, \cheb)^\trans \vC(\cov)^{-1} \vR(\stellarParameters, \cheb) + \\
-  \ln{\det{\vC(\cov)}} + N_{\rm pix} \ln{2\pi} \Bigr),
-\end{multline}
-If the spectrum contains multiple spectral orders, then the total log-likelihood is
-\begin{multline}
-  \ln{p(\vD | \vM(\stellarParameters, \Cheb), \Cov)} = \\
-  -\frac{1}{2} \sum_i^{N_\textrm{orders}} \Bigl( \vR_i(\stellarParameters, \chebi{i})^\trans \vC(\covi{i})^{-1} \vR_i(\stellarParameters, \chebi{i}) + \\
-  \ln{\det{\vC(\covi{i})}} + N_{ {\rm pix}, i} \ln{2\pi} \Bigr) .
-\end{multline}
-
-\begin{figure*}[!htb]
+\begin{figure*}[!t]
 \begin{center}
@@
 
+\subsection{Priors} \label{subsec:priors}
+
+\comm{We need a subsection describing what you assume for your priors.}
+
 \subsection{Exploring the Posterior} \label{subsec:MCMC}
 
-Using Bayes' rule, we write down the posterior probability distribution as proportional to the likelihood times the prior
+The inference problem as framed here has a natural hierarchical structure between the collections 
+of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
+$\vP = [\cheb, \cov]$.  To explore the posterior probability distribution of the model conditioned 
+on an observed spectrum in this case, it is convenient to use MCMC simulations with a blocked Gibbs 
+sampler coupled to the Metropolis-Hastings algorithm.  In essence, this procedure works by sampling 
+in a subset of parameters with Metropolis-Hastings conditioned on the current (fixed) values of the 
+other parameters; after each iteration, the Gibbs sampler then updates the values of that parameter 
+subset and then cycles through all the different parameter subsets in the same way (for a more 
+nuanced mathematical description of the process, see Chapter 11 of \citealt{gelman13}).  The 
+step-by-step prescription is as follows, where we denote the $i^{\rm th}$ iteration of the Gibbs 
+sampler as a superscript index (not to be confused with the $i^{\rm th}$ pixel as in our previous 
+notation):
+\begin{enumerate}
+\item Initialize the parameters.  One might set $\vT^0$ based on estimates in the literature or 
+scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the Chebyshev 
+polynomial coefficients ($\cheb^0$) to be flat ($c_0 = 1$,  $c_{>0} = 0$ in all spectral orders) 
+and assume only the trivial noise spectrum contributes to the covariance matrix (all elements of 
+$\cov^0$ are zero).
+
+\item Start a new ($i^{\rm th}$; where $i \in [1,N_G]$) iteration of the Gibbs sampler.  For 
+$N_{MH}$ iterations of the Metropolis-Hastings algorithm, sample in $\vT$ to evaluate the posterior 
+probability 
 \begin{equation}
-  p(\allParameters | \vD) =  p(\stellarParameters, \Cheb, \Cov | \vD ) \propto p(\vD | \allParameters) p(\allParameters).
+p(\vT,\vP|\vD) \propto p(\vD | \vT, \vP = \vP^{i-1}) \,\, p(\vT,\vP)
 \end{equation}
-Given the logical hierarchical structure of the collection of stellar and nuisance parameters, $\allParameters$, it is convenient to use a blocked Gibbs sampler with the Metropolis-Hastings algorithm to explore this posterior distribution.
+following the framework laid out in Sections \ref{subsec:likelihood} and \ref{subsec:priors}, 
+representing a slice through the posterior space conditioned on the hyperparameters being fixed to 
+their $(i-1)^{\rm th}$ iteration values.  Now update the parameters $\vT^{i-1} \rightarrow \vT^i$.
+
+\item Cycle through and update the hyperparameters, $\vP^{i-1} \rightarrow \vP^i$:
+
+\begin{enumerate}
+\item For each spectral order, sample in the polynomial hyperparameters $\cheb$ using $N_{MH}$ 
+iterations of the Metropolis-Hastings algorithm, computing posterior probabilities
+\begin{equation}
+p(\vT,\vP|\vD) \propto p(\vD | \cheb, \cov = \cov^{i-1}, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
+\end{equation}
+along a subset of the posterior space where $\vT^i$ and $\cov^{i-1}$ are fixed.  Now update the 
+polynomial hyperparameters $\cheb^{i-1} \rightarrow \cheb^i$.
+
+\item Sample in the global (stationary) covariance kernel hyperparameters, $\vp_{{\mathsf C},g} = 
+[a_g, \ell]$, using $N_{MH}$ iterations of the Metropolis-Hastings algorithm to compute posteriors
+\begin{equation}
+p(\vT,\vP|\vD) \propto p(\vD | \vp_{{\mathsf C}, g}, \vp_{{\mathsf C}, k} = \vp_{{\mathsf C, k}}^{i-1}, \cheb = \cheb^i, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
+\end{equation}
+with $\vT^i$ and the local covariance kernel parameters, $\vp_{{\mathsf C}, k}^{i-1}$, fixed.  
+Update the global covariance hyperparameters $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf 
+C}, g}^i$.
+
+\item Decide whether to instantiate or delete any local (non-stationary) covariance kernels.  Here 
+we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
+$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this 
+will trigger the addition of a local kernel; if that same criterion is not met at the location of a 
+previously-instantiated kernel, that kernel is then deleted.
+
+\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using $N_{MH}$ 
+iterations of the Metropolis-Hastings algorithm to compute the posterior probabilities
+\begin{equation}
+p(\vT,\vP|\vD) \propto p(\vD | \vp_{{\mathsf C}, k}, \vp_{{\mathsf C}, g} = \vp_{{\mathsf C, g}}^i, \cheb = \cheb^i, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
+\end{equation}
+with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local covariance 
+hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.
+\end{enumerate}
+
+\item Return to Step 2 and iterate until a convergence criterion is met.
+
+\item Repeat this procedure with different initializations (Step 1), to ensure that a global 
+maximum in the posterior probability distribution is being sampled.
+\end{enumerate}
+The number of Metropolis-Hastings iterations in each step is somewhat arbitrary, and can be tuned 
+for optimization purposes; we typically set $N_{MH} \approx 10$.  The number of Gibbs sampler 
+iterations, $N_G$, are determined empirically with a convergence criterion .... \comm{how exactly?}
+
+This procedure can be a computational challenge.  A typical spectrum might have $N_{\rm 
+pix}$$>$1000, and therefore the many repeated evaluations of the matrix product $\vR^{\trans} 
+\vC^{-1} \vR$ in the likelihood function calculation can be numerically expensive; prohibitively so 
+in the brute force sense, since $\vC$ is non-trivial by design.  However, there are two aspects of 
+these calculations that enable us to take advantage of clever, efficient algorithms designed for 
+specifically these kinds of issues.  First, since we are interested in the matrix product 
+$\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can exploit efficient routines for 
+solving linear systems that bypass the computational expense of matrix inversion.\footnote{\comm{(This seems vague.)}}  And second, because $\vC$ is a positive self-definite 
+matrix, we can employ Cholesky factorization to optimize the evaluation of the matrix product.  The 
+makes Step 2 and Steps 3a and 3b in the procedure outlined above extremely rapid.  Although this 
+factorization must be re-done for each update of the covariance hyperparameters, such operations 
+are still modestly efficient because the covariance kernels were specifically designed to deliver a 
+sparse matrix (via the Hann windows).  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse 
+matrix and Cholesky factorization operations \citep{chen08,davis09}, and employ the 
+Metropolis-Hastings sampler included in the {\tt emcee} package \citep{foreman-mackey13}.
+
+%Using Bayes' rule, we write down the posterior probability distribution as proportional to the likelihood times the prior
+%\begin{equation}
+%  p(\vT | \vD) =  p(\vt_{\ast}, \cheb, \cov | \vD ) \propto p(\vD | \vT) p(\vT).
+%\end{equation}
+%Given the logical hierarchical structure of the collection of stellar and nuisance parameters, $\allParameters$, it is convenient to use a blocked Gibbs sampler with the Metropolis-Hastings algorithm to explore this posterior distribution.
 The sampler works by using the Metropolis-Hastings algorithm to sample a sub-collection of the parameters conditional on the current values of all the other parameters.
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
+%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
 These stellar parameters are marginalized over the Chebyshev ($\Cheb$) and noise parameters ($\Cov$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:stellar_posterior}
-\end{center}
-\end{figure*}
+%}
+%\label{fig:stellar_posterior}
+%\end{center}
+%\end{figure*}
 
-The following procedure illustrates how the Gibbs sampler rotates among subsets of the parameters when sampling a multi-order echelle spectrum.
+%The following procedure illustrates how the Gibbs sampler rotates among subsets of the parameters when sampling a multi-order echelle spectrum.
 We use superscripts to denote iterations of the MCMC algorithm, where $j$ denotes the parameters from the current iteration and $j - 1$ denotes the previous iteration.
 
-\begin{enumerate}
-  \item Choose starting parameters for $\allParameters^0$.
+%\begin{enumerate}
+%  \item Choose starting parameters for $\allParameters^0$.
 Generally, these will be $\stellarParameters^0$ drawn from the literature, flat Chebyshev polynomials ($\Cheb^0$), and only Poisson noise ($\Cov^0$).
-  \item Start a new iteration of the Gibbs sampler with the parameters of the current proposal denoted with the superscript $j$ and the previous iteration denoted by $j-1$.
+%  \item Start a new iteration of the Gibbs sampler with the parameters of the current proposal denoted with the superscript $j$ and the previous iteration denoted by $j-1$.
 If this is the first iteration of the Gibbs sampler, then $j - 1 = 0$.
-  \item Sample the stellar parameters $p(\stellarParameters^j | \Cheb^{j-1}, \Cov^{j-1})$. 
-  \item For each order of the echelle spectrum, $i$
-    \begin{enumerate}
-      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \stellarParameters^j, \covi{i}^{j-1} )$
-      \item Check to see whether the algorithm should instantiate/delete new/old local kernels
-      \item Sample the covariance parameters $p(\covi{i}^j | \stellarParameters^j, \chebi{i}^{j})$ 
-    \end{enumerate}
-  after step 4 is completed, all Chebyshev and covariance parameters have been updated: $\Cheb^{j-1} \rightarrow \Cheb^j$ and $\Cov^{j-1} \rightarrow \Cov^j$
-    \item return to step 2 until Markov Chain is converged
-\end{enumerate}
+%  \item Sample the stellar parameters $p(\stellarParameters^j | \Cheb^{j-1}, \Cov^{j-1})$. 
+%  \item For each order of the echelle spectrum, $i$
+%    \begin{enumerate}
+%      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \stellarParameters^j, \covi{i}^{j-1} )$
+%      \item Check to see whether the algorithm should instantiate/delete new/old local kernels
+%      \item Sample the covariance parameters $p(\covi{i}^j | \stellarParameters^j, \chebi{i}^{j})$ 
+%    \end{enumerate}
+%  after step 4 is completed, all Chebyshev and covariance parameters have been updated: $\Cheb^{j-1} \rightarrow \Cheb^j$ and $\Cov^{j-1} \rightarrow \Cov^j$
+%    \item return to step 2 until Markov Chain is converged
+%\end{enumerate}
 
-The hierarchical structure of the parameters influences how the algorithm may be efficiently implemented in code.
-For a typical optical spectrum with $\gtrsim$1,000 pixels, the most computationally intensive step of the likelihood evaluation is usually the matrix product $\fR^T C^{-1} \fR$.
-Since we have designed the covariance matrix to be sparse, we can use optimized sparse matrix algorithms which are much faster and memory efficient than dense matrix operations.
-Because we are not interested in the matrix inverse $C^{-1}$ by itself, but rather the product with the residual vectors, we can use efficient routines for solving linear systems to bypass the computationally difficult step of matrix inversion.
-Additionally, because the covariance matrix is positive semi-definite, we can use the Cholesky factorization of the matrix to optimize the evaluation of the matrix product.
-Once the covariance matrix is factorized, any subsequent evaluation of the matrix product for different residual vectors $\fR$ is extremely rapid.
-This makes the $\stellarParameters$ and $\cheb$ steps of the Gibbs sampler extremely fast.
-When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update.
-However, because we designed the kernels to deliver a sparse matrix these operations are still moderately efficient.
-We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and utilize the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.
-
-We run the MCMC Gibbs sampler for many iterations until the estimate of the posterior distribution has converged.
+%We run the MCMC Gibbs sampler for many iterations until the estimate of the posterior distribution has converged.
 To check that the chain is not stuck in a local maximum of the posterior, we redo the MCMC run many times with different starting parameters, to ensure that the algorithm converges to the same global maximum.
@@
 Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
-This enables us to present a posterior of the stellar parameters $\stellarParameters$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
+This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
 This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
 
-The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
+%The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
 If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
 
-When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+%When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
 Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
 As demonstrated by the Gibbs sampler routine, we explicitly fit for the
-hyperparameters of the covariance kernels in parallel with the
-stellar parameters. While including these extra parameters does increase the
-dimensionality and complexity of our model, they do provide several advantages.
-\emph{A priori} we do not know which regions of the spectrum are improperly
-modeled. The covariance hyperparameters provide a framework to identify these
-regions iteratively and in a self-consistent manner.
+%hyperparameters of the covariance kernels in parallel with the
+%stellar parameters. While including these extra parameters does increase the
+%dimensionality and complexity of our model, they do provide several advantages.
+%\emph{A priori} we do not know which regions of the spectrum are improperly
+%modeled. The covariance hyperparameters provide a framework to identify these
+%regions iteratively and in a self-consistent manner.
 
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+%The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 
-\subsection{Applications}
-\label{subsec:learning}
-By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
+%\subsection{Applications \comm{untouched}}
+%\label{subsec:learning}
+%By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
 After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
@@
 
-Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
+%Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
 Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
@@
 
-Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
+%Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
 Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
@@
     with class III lines is that for any specific line, there might exist a
-    $\stellarParameters$ that will fit the line, but there does not exist a $\stellarParameters$ that
+    $\vT$ that will fit the line, but there does not exist a $\vT$ that
     will properly fit \emph{all} the lines.}
@@
 $\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\fundamentalParameters$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-$\gridParameters$ & stellar parameters that specify a spectrum from a synthetic library\\
-$\observationalParameters$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
+$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
   & are applied during ``post processing'' of the synthetic spectrum\\
-$\stellarParameters$ & $\{\fundamentalParameters,\observationalParameters \}$, all stellar parameters\\
+$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
 
@@
 $\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
-$\vT$ & $\{\stellarParameters, \vP\}$, the collection of all parameters \\
+$\vT$ & $\{\vT, \vP\}$, the collection of all parameters \\
 $\vD$ & data spectrum \\

### 2014-07-24 tex/paper.tex
Added revert capability to global covariance matrix

@@
 \newcommand{\matern}{Mat\'{e}rn}
+\newcommand{\HK}{$\textrm{H}_2$O-K2}
 
@@
 \newcommand{\vP}{ {\bm \Phi}}
-
-% Ian, this was ridiculous...don't type out giant words that are longer than the required tex.  It 
-% actually takes a lot more time and is a huge pain.  
-%\newcommand{\fundamentalParameters}{\vt_{\ast}}
-%\newcommand{\gridParameters}{\{\fundamentalParameters\}^{\rm 
-%grid}}
-%\newcommand{\observationalParameters}{\vt_{\textrm{obs}}}
-%\newcommand{\stellarParameters}{\vt}
 
@@
 Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}.
-\citep{rojas-ayala12} use the H20 -K index (defined over what wl range?).
+\citep{rojas-ayala12} use the 2.205~$\mu$m \ion{Na}{1} doublet and 2.263~$\mu$m \ion{Ca}{1} triplet and Na doublet lines (their table 2), and their \HK\ index, defined using three passbands: 2.070~$\mu$m-2.090~$\mu$m, 2.235~$\mu$m-2.255~$\mu$m, and 2.360~$\mu$m - 2.380~$\mu$m.
+The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water.
+For a useful visualization of which spectral features these indices track, see their figure 2.
+The \HK\ index is a monotonic function of temperature, showing more curvature in the spectrum towards cooler $T_{\rm eff}$ and the index is independent of $\Z$ in the range $3000\,{\rm K} \leq T_{\rm eff} \leq 4000\,{\rm K}$.
+Below $3000~{\rm K}$, the index may appear to be metallicity dependent, although they caution that this may be a computational artifact.
+They find that the \HK\ index has a negligible dependence on $\log g$.
+They find that the index correlates with spectral type.
+
+\citet{rojas-ayala12} use a custom grid of PHOENIX BT-Settl models run from the PHOENIX web simulator at.
+The note that the models fit the spectrum match the overall spectral shape well, although the models consistently underpredict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
+\citet{rajpurohit10} note that the oscillator strengths and opacities of the \ion{Na}{1} and \ion{Ca}{1} lines may be incorrect.
+In the range of stellar parameters for M dwarfs, \citet{rojas-ayala12} note that the strength of the \ion{Na}{1} doublet increases as $T_{\rm eff}$ decreases.
+The strength of the \ion{Ca}{1} triplet increases as $T_{\rm eff}$ and $\Z$ increase.
+
+Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl51.
 

### 2014-07-24 StellarSpectra/covariance.pyx
Added revert capability to global covariance matrix

@@ -126,10 +126,12 @@ cdef class CovarianceMatrix:
 
     cdef cholmod_sparse *A
     cdef cholmod_factor *L
+    cdef cholmod_factor *L_last
     cdef Common common
     cdef cholmod_common *c
     cdef double *one
     cdef double logdet
+    cdef double logdet_last
     cdef wl
     cdef fl
     cdef npoints
@@ -139,13 +141,15 @@ cdef class CovarianceMatrix:
     cdef RegionList
     cdef current_region
     cdef logPrior
+    cdef logPrior_last
+    cdef debug
 
     #Both sum_regions and partial_sum_regions are initially NULL pointers, and will be until
     #any elements in the RegionList have been declared.
     cdef cholmod_sparse *sum_regions #Holds a sparse matrix that is the sum of all regions
     cdef cholmod_sparse *partial_sum_regions #Holds a sparse matrix that is the sum of all regions but the one currently sampling (N -1)
 
-    def __init__(self, DataSpectrum, order_index):
+    def __init__(self, DataSpectrum, order_index, debug=False):
         self.common = Common()
         self.c = <cholmod_common *>&self.common.c
 
@@ -162,9 +166,12 @@ cdef class CovarianceMatrix:
         self.fl = DataSpectrum.fls[self.order_index][mask]
         self.npoints = len(self.fl)
         self.logdet = 0.0
+        self.logdet_last = self.logdet
         self.logPrior = 0.0 #neutral prior, reflects the sum of GlobalCovariance.logPrior and all region.logPrior's
+        self.logPrior_last = self.logPrior
+        self.debug = debug
 
-        self.GCM = GlobalCovarianceMatrix(self.DataSpectrum, self.order_index,self.common)
+        self.GCM = GlobalCovarianceMatrix(self.DataSpectrum, self.order_index,self.common, debug=self.debug)
         self.RegionList = []
         self.current_region = -1 #non physical value to force initialization
 
@@ -182,14 +189,25 @@ cdef class CovarianceMatrix:
             cholmod_free_sparse(&self.partial_sum_regions, self.c)
         if self.A != NULL:
             cholmod_free_sparse(&self.A, self.c)
-        if self.L != NULL:
-            cholmod_free_factor(&self.L, self.c)
+
+        # Due to revert functionality, there is a chance that __dealloc__ may be called in a state when self.L and
+        # self.L_last both point to the same piece of memory
+        if self.L != self.L_last:
+            if self.L != NULL:
+                cholmod_free_factor(&self.L, self.c)
[…]

### 2014-07-24 scripts/stars/base_lnprob_old.py
Added revert capability to global covariance matrix

@@ -1,5 +1,4 @@
-from StellarSpectra.model import Model, StellarSampler, ChebSampler, MegaSampler, CovGlobalSampler#, RegionsSampler,
-# MegaSampler
+from StellarSpectra.model import Model, StellarSampler, ChebSampler, CovGlobalSampler, RegionsSampler, MegaSampler
 from StellarSpectra.spectrum import DataSpectrum
 from StellarSpectra.grid_tools import TRES, HDF5Interface
 import StellarSpectra.constants as C
@@ -137,38 +136,37 @@ for order in config['orders']:
 shutil.copy(yaml_file, outdir + "/input.yaml")
 
 myModel = Model(myDataSpectrum, myInstrument, myHDF5Interface, stellar_tuple=stellar_tuple, cheb_tuple=cheb_tuple,
-                cov_tuple=cov_tuple, region_tuple=region_tuple, outdir=outdir, debug=False)
+                cov_tuple=cov_tuple, region_tuple=region_tuple, outdir=outdir)
 
 try:
     fix_logg = config['fix_logg']
 except KeyError:
     fix_logg = None
 
-myStellarSampler = StellarSampler(model=myModel, cov=stellar_MH_cov, starting_param_dict=stellar_Starting,
-                                  fix_logg=fix_logg, outdir=outdir, debug=False)
+myStellarSampler = StellarSampler(myModel, stellar_MH_cov, stellar_Starting, fix_logg=fix_logg,
+                                  outdir=outdir)
 
 #Create the three subsamplers for each order
 samplerList = []
-
+cadenceList = []
 for i in range(len(config['orders'])):
-    samplerList.append(ChebSampler(model=myModel, cov=cheb_MH_cov, starting_param_dict=cheb_Starting, order_index=i,
-                                   outdir=outdir, debug=False))
-    samplerList.append(CovGlobalSampler(model=myModel, cov=cov_MH_cov, starting_param_dict=cov_Starting, order_index=i,
-                                   outdir=outdir, debug=False))
-    #samplerList.append(RegionsSampler(myModel, region_MH_cov, max_regions=config['max_regions'],
-    #                default_param_dict=region_Starting, order_index=i, outdir=outdir))
-    #cadenceList += [1, 1, 1]
+    samplerList.append(ChebSampler(myModel, cheb_MH_cov, cheb_Starting, order_index=i, outdir=outdir))
[…]

### 2014-07-27 tex/paper.tex
works with multiple orders

@@
 for optimization purposes; we typically set $N_{MH} \approx 10$.  The number of Gibbs sampler 
-iterations, $N_G$, are determined empirically with a convergence criterion .... \comm{how exactly?}
+iterations, $N_G$, are determined empirically with the Gelman-Rubin convergence diagnostic ensuring it is below $1.1$, which compares inter-chain and intra-chain variance \citep[Eqn 11.4]{gelman13} and represents the scale by which the distribution might be reduced by continuing the simulation for infinity.
+Splitting chains and assessing mixing.
 

### 2014-07-27 hdfcat.py

@@ -14,6 +14,10 @@ parser.add_argument("--dir", action="store_true", help="Concatenate all of the f
 parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
 parser.add_argument("-o", "--output", default="combined.hdf5", help="Output HDF5 file.")
 parser.add_argument("--clobber", action="store_true", help="Overwrite existing file?")
+parser.add_argument("--burn", type=int, default=0, help="How many samples to discard from the beginning of the chain "
+                                                        "for burn in.")
+parser.add_argument("--thin", type=int, default=1, help="Thin the chain by this factor. E.g., --thin 100 will take "
+                                                        "every 100th sample.")
 args = parser.parse_args()
 
 #Check to see if output exists. If --clobber, overwrite, otherwise exit.
@@ -44,17 +48,70 @@ for stellar in stellarlist:
     assert stellar.attrs["parameters"] == stellarlist[0].attrs["parameters"], "Parameter lists do not match."
     assert stellar.shape[1] == stellarlist[0].shape[1], "Different number of parameters."
 
+
+def find_cov(name):
+    if name == "cov":
+        return True
+    return None
+
+def find_region(name):
+    if "cov_region" in name:
+        return True
+    return None
+
+#Determine how many orders, if there is global covariance, or regions
+#choose the first chain
+hdf5 = hdf5list[0]
+orders = [int(key) for key in hdf5.keys() if key != "stellar"]
+orders.sort()
+
+yes_cov = hdf5.visit(find_cov)
+yes_region = hdf5.visit(find_region)
+
+# give this a key relative from the top, and it will return a list of all flatchains
+def get_flatchains(key):
+    return [hdf5.get(key)[:] for hdf5 in hdf5list]
+
+#Order list will always be a 2D list, with the items being flatchains
+ordersList = []
+for order in orders:
+
+    temp = [get_flatchains("{}/cheb".format(order))]
+    if yes_cov:
+        temp += [get_flatchains("{}/cov".format(order))]
+
+    #TODO: do something about regions here
+
+    #accumulate all of the orders
+    ordersList += [temp]
+
+print("Thinning by ", args.thin)
+print("Burning out first {} samples".format(args.burn))
+stellarlist = [stellar[args.burn::args.thin] for stellar in stellarlist]
[…]

### 2014-07-28 tex/paper.tex
updated kurucz params

--- 2014-07-24 tex/paper.tex
@@
 
-\comm{These local kernels can be further modified to account for more complex residual structures.  For 
+These local kernels can be further modified to account for more complex residual structures.  For 
 example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
@@
 spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.}
-\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}
+in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.
 
@@
 \subsection{Priors} \label{subsec:priors}
-
-\comm{We need a subsection describing what you assume for your priors.}
+The final posterior probability distribution of the model parameters is defined by Bayes rule as proportional to the product of the likelihood function (Eqn~\ref{eqn:likelihood}) and the prior probabilities on the model parameters.
+Because a high quality spectrum provides so much information about $\vT$, in general the inference of the stellar parameters is not very sensitive to specific choices of priors.
+We generally assume uniform priors on the stellar parameters, which are flat for any $\vt_{\ast}$ within the range of $\{\vt_{\ast}\}^{\rm grid}$ and otherwise practically $0$.
+For stars with a well-defined continuum, it is generally acceptable to use flat priors on the polynomial parameters $\cheb$.
+However, if there is information about the systematic accuracy of flux calibration, for example from multiple observations of spectrophotometric standards, it is possible to encode this information with a Gaussian prior on the coefficients of $\cheb$, with a mean of $0$ and a width of the typical systematic variation of the flux calibration (e.g., $5\%$).
+For stars with a poorly defined continuum or pseudo-continuum, judicious choice of these priors will ensure that broad spectroscopic features are not destroyed by the polynomial fit.
+For the parameters of local covariance kernels, $\vp_{{\mathsf C},l}$, we assume a prior on the width of the local kernel that is flat below the scale of the instrumental PSF and rapidly Gaussian tapers to zero for widths larger than the PSF.
+This prior is designed to prevent the local kernels from diffusing to large widths and low amplitudes, effectively duplicating the role of the global kernels.
 
@@
 \item Start a new ($i^{\rm th}$; where $i \in [1,N_G]$) iteration of the Gibbs sampler.  For 
-$N_{MH}$ iterations of the Metropolis-Hastings algorithm, sample in $\vT$ to evaluate the posterior 
-probability 
-\begin{equation}
-p(\vT,\vP|\vD) \propto p(\vD | \vT, \vP = \vP^{i-1}) \,\, p(\vT,\vP)
-\end{equation}
-following the framework laid out in Sections \ref{subsec:likelihood} and \ref{subsec:priors}, 
+one iteration of the Metropolis-Hastings algorithm, sample in $\vT$ to evaluate the posterior 
+probability following the framework laid out in Sections \ref{subsec:likelihood} and \ref{subsec:priors}, 
 representing a slice through the posterior space conditioned on the hyperparameters being fixed to 
@@
 \begin{enumerate}
-\item For each spectral order, sample in the polynomial hyperparameters $\cheb$ using $N_{MH}$ 
-iterations of the Metropolis-Hastings algorithm, computing posterior probabilities
-\begin{equation}
-p(\vT,\vP|\vD) \propto p(\vD | \cheb, \cov = \cov^{i-1}, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
-\end{equation}
-along a subset of the posterior space where $\vT^i$ and $\cov^{i-1}$ are fixed.  Now update the 
+\item For each spectral order, sample in the polynomial hyperparameters $\cheb$ using one 
+iteration of the Metropolis-Hastings algorithm, computing posterior probabilities along a subset of the posterior space where $\vT^i$ and $\cov^{i-1}$ are fixed.  Now update the 
 polynomial hyperparameters $\cheb^{i-1} \rightarrow \cheb^i$.
@@
 \item Sample in the global (stationary) covariance kernel hyperparameters, $\vp_{{\mathsf C},g} = 
-[a_g, \ell]$, using $N_{MH}$ iterations of the Metropolis-Hastings algorithm to compute posteriors
-\begin{equation}
-p(\vT,\vP|\vD) \propto p(\vD | \vp_{{\mathsf C}, g}, \vp_{{\mathsf C}, k} = \vp_{{\mathsf C, k}}^{i-1}, \cheb = \cheb^i, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
-\end{equation}
+[a_g, \ell]$, using one iteration of the Metropolis-Hastings algorithm to compute posteriors
 with $\vT^i$ and the local covariance kernel parameters, $\vp_{{\mathsf C}, k}^{i-1}$, fixed.  
@@
 
-\item Decide whether to instantiate or delete any local (non-stationary) covariance kernels.  Here 
+\item Decide whether to instantiate any additional local (non-stationary) covariance kernels.  Here 
 we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
 $3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this 
-will trigger the addition of a local kernel; if that same criterion is not met at the location of a 
-previously-instantiated kernel, that kernel is then deleted.
+will trigger the addition of a local kernel.
 
-\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using $N_{MH}$ 
-iterations of the Metropolis-Hastings algorithm to compute the posterior probabilities
-\begin{equation}
-p(\vT,\vP|\vD) \propto p(\vD | \vp_{{\mathsf C}, k}, \vp_{{\mathsf C}, g} = \vp_{{\mathsf C, g}}^i, \cheb = \cheb^i, \vT = \vT^i) \,\, p(\vT, \cheb, \cov)
-\end{equation}
+\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using one 
+iteration of the Metropolis-Hastings algorithm to compute the posterior probabilities
 with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local covariance 
 hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.
+If the stellar parameters have moved to a region of parameter space where the local kernel is no longer required, then the amplitude of the kernel will be driven towards $0$ and the kernel would have a negligible contribution to the covariance matrix $\vC$.
 \end{enumerate}
 
-\item Return to Step 2 and iterate until a convergence criterion is met.
+\item Return to Step 2 and iterate for $N_G$ iterations.
 
-\item Repeat this procedure with different initializations (Step 1), to ensure that a global 
-maximum in the posterior probability distribution is being sampled.
+\item Repeat this procedure with different initializations (Step 1) and compute convergence diagnostics of the multiple chains, to ensure that all of the chains have converged to the same global 
+maximum in the posterior probability distribution and that each chain has converged.
 \end{enumerate}
-The number of Metropolis-Hastings iterations in each step is somewhat arbitrary, and can be tuned 
-for optimization purposes; we typically set $N_{MH} \approx 10$.  The number of Gibbs sampler 
-iterations, $N_G$, are determined empirically with a convergence criterion .... \comm{how exactly?}
+How many iterations to run the Gibbs sampler, $N_G$, is empirically determined through visual inspection of the Markov chains and with the Gelman-Rubin convergence diagnostic $\hat{R}$ \citep[Eqn 11.4]{gelman13}, which compares intra-chain and inter-chain variance to assess convergence.
+After a running a series of multiple chains, we compute the diagnostic, which summarizes how the inference of the posterior probability function would be improved if we were to run the Markov chain for an infinite number of iterations.
+If $\hat{R}$$>$1.1, then there is evidence that the chains have not fully converged and so we increase $N_G$ and repeat the simulations until we are confident of convergence.
 
-This procedure can be a computational challenge.  A typical spectrum might have $N_{\rm 
-pix}$$>$1000, and therefore the many repeated evaluations of the matrix product $\vR^{\trans} 
-\vC^{-1} \vR$ in the likelihood function calculation can be numerically expensive; prohibitively so 
-in the brute force sense, since $\vC$ is non-trivial by design.  However, there are two aspects of 
-these calculations that enable us to take advantage of clever, efficient algorithms designed for 
-specifically these kinds of issues.  First, since we are interested in the matrix product 
-$\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can exploit efficient routines for 
-solving linear systems that bypass the computational expense of matrix inversion.\footnote{\comm{(This seems vague.)}}  And second, because $\vC$ is a positive self-definite 
-matrix, we can employ Cholesky factorization to optimize the evaluation of the matrix product.  The 
-makes Step 2 and Steps 3a and 3b in the procedure outlined above extremely rapid.  Although this 
-factorization must be re-done for each update of the covariance hyperparameters, such operations 
-are still modestly efficient because the covariance kernels were specifically designed to deliver a 
-sparse matrix (via the Hann windows).  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse 
-matrix and Cholesky factorization operations \citep{chen08,davis09}, and employ the 
-Metropolis-Hastings sampler included in the {\tt emcee} package \citep{foreman-mackey13}.
+Gibbs sampling from the posterior with this procedure can be a computational
+challenge.  A typical spectrum might have $N_{\rm pix}$$>$1000, and therefore
+the many repeated evaluations of the matrix product $\vR^{\trans} \vC^{-1} \vR$
+in the likelihood function calculation can be numerically expensive;
+prohibitively so in the brute force sense, since $\vC$ is non-trivial by
+design.  However, there are two aspects of these calculations that enable us to
+take advantage of clever, efficient algorithms designed for specifically these
+kinds of issues.  First, since we are interested in the matrix product
+$\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can directly solve for the product using efficient algorithms and avoid the computational expense of
+matrix inversion. And second, because
+$\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky factorization
+to optimize the evaluation of the matrix product.  The makes Step 2 and Steps
+3a and 3b in the procedure outlined above extremely rapid.  Although this
+factorization must be re-done for each update of the covariance
+hyperparameters, such operations are still modestly efficient because the
+covariance kernels were specifically designed to deliver a sparse matrix (via
+the Hann windows).  We use the high-performance
+\texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}}
+library to implement the sparse matrix and Cholesky factorization operations
+\citep{chen08,davis09}, and extend the Metropolis-Hastings sampler included in
+the {\tt emcee} package \citep{foreman-mackey13} to function within a Gibbs sampler\footnote{\url{https://github.com/iancze/emcee}}.
 
-%Using Bayes' rule, we write down the posterior probability distribution as proportional to the likelihood times the prior
-%\begin{equation}
-%  p(\vT | \vD) =  p(\vt_{\ast}, \cheb, \cov | \vD ) \propto p(\vD | \vT) p(\vT).
-%\end{equation}
-%Given the logical hierarchical structure of the collection of stellar and nuisance parameters, $\allParameters$, it is convenient to use a blocked Gibbs sampler with the Metropolis-Hastings algorithm to explore this posterior distribution.
-The sampler works by using the Metropolis-Hastings algorithm to sample a sub-collection of the parameters conditional on the current values of all the other parameters.
-After each iteration, the Gibbs sampler updates the values of the sub-collection, and then moves to a new sub-collection of parameters.
-By rotating through all of the sub-collections of parameters, the Gibbs sampler ensures that it samples the full posterior probability distribution.
-For more on MCMC and Gibbs samplers see \citet[Ch.
-11]{gelman13} and the references therein.
-
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
-%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\Cheb$) and noise parameters ($\Cov$).
-Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-%}
-%\label{fig:stellar_posterior}
-%\end{center}
-%\end{figure*}
-
-%The following procedure illustrates how the Gibbs sampler rotates among subsets of the parameters when sampling a multi-order echelle spectrum.
-We use superscripts to denote iterations of the MCMC algorithm, where $j$ denotes the parameters from the current iteration and $j - 1$ denotes the previous iteration.
-
-%\begin{enumerate}
-%  \item Choose starting parameters for $\allParameters^0$.
-Generally, these will be $\stellarParameters^0$ drawn from the literature, flat Chebyshev polynomials ($\Cheb^0$), and only Poisson noise ($\Cov^0$).
-%  \item Start a new iteration of the Gibbs sampler with the parameters of the current proposal denoted with the superscript $j$ and the previous iteration denoted by $j-1$.
-If this is the first iteration of the Gibbs sampler, then $j - 1 = 0$.
-%  \item Sample the stellar parameters $p(\stellarParameters^j | \Cheb^{j-1}, \Cov^{j-1})$. 
-%  \item For each order of the echelle spectrum, $i$
-%    \begin{enumerate}
-%      \item Sample the Chebyshev polynomial parameters $p(\chebi{i}^j | \stellarParameters^j, \covi{i}^{j-1} )$
-%      \item Check to see whether the algorithm should instantiate/delete new/old local kernels
-%      \item Sample the covariance parameters $p(\covi{i}^j | \stellarParameters^j, \chebi{i}^{j})$ 
-%    \end{enumerate}
-%  after step 4 is completed, all Chebyshev and covariance parameters have been updated: $\Cheb^{j-1} \rightarrow \Cheb^j$ and $\Cov^{j-1} \rightarrow \Cov^j$
-%    \item return to step 2 until Markov Chain is converged
-%\end{enumerate}
-
-%We run the MCMC Gibbs sampler for many iterations until the estimate of the posterior distribution has converged.
-To check that the chain is not stuck in a local maximum of the posterior, we redo the MCMC run many times with different starting parameters, to ensure that the algorithm converges to the same global maximum.
-A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
+\todo{What should we do with the commented-out sections?}
+%A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
 Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
@@
 
-\subsection{Fitting data}
+\subsection{Datasets and historical analysis}
+% Paragraphs as follows:
+% Introduction to the two types of stars, in order to demonstrate the flexibility.
+To demonstrate the flexibility of our probabilistic framework, we model spectra of two different stars acquired with different instruments.
+The first data set is a high resolution optical spectrum of WASP-14, an F5 star \citep{joshi09}.
+The second data set is a moderate resolution mid-infrared spectrum of Gl~51, a M5 star \citep{rojas-ayala12}.
 
-To demonstrate our technique, we fit spectra of two different stars with varying stellar properties and acquired with different instruments.
-The first object is WASP-14, an F5 star \citep{joshi09}.
-The spectrum was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}) \citep{mink11} on the FLWO 1.5m telescope on June 14th, 2009.
-\emph{TRES} delivers a 51 echelle spectrum spanning the full optical wavelength spectrum $3860 - 9100$ \AA\ at a resolution of $R \sim$ 44,000.
-\citet{torres12} acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts.
-The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.
-Buchhave.
-In the \ion{Mg}{0}b order ($\sim 5100$ \AA), the spectrum has a signal to noise of 150 per resolution element.
+% Data description of WASP-14
+Our spectrum of WASP-14 was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}, \citealt{mink11}) on the Fred Laurence Whipple Observatory 1.5m telescope on June 14th, 2009.
+\emph{TRES} delivers a 51 echelle spectrum spanning the full optical spectrum (3860--9100\AA) at a resolution of $R \sim$44,000.
+\citet{torres12} originally acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts.
+The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave.
+In the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.
 
-The SPC analysis within \citet{torres12} focused on a region of spectrum from $5000 5300$ \AA and used many methods.
-The other methods are not directly comparable because they use different underlying assumptions.
+% What other people get when they fit this star. How does our inference compare?
+\citet{torres12} measured the stellar properties of WASP-14 using several different methods including MOOG, SME, and SPC.
+The MOOG and SME methods are not directly comparable to our technique because they measure individual line strengths.
+However, because SPC utilizes a large swath of the spectrum and a model spectrum, we can use the same spectral library and make the same assumptions.
+The SPC analysis within \citet{torres12} focused on a region of spectrum from 5050--5360\AA\ and used many methods.
 Additionally, there are systematics introduced by metallicity assumptions and radiative transfer.
 Using this same spectral library as SPC, we recover the same parameters.
+When we fit this spectrum using the same assumptions, (amounting to a $\delta$-function prior on $\log g$).
+This demonstrates sensitivity to a prior as well as the fact that surface gravity estimates from spectroscopy may be biased against fits derived from planetary transits.
+A Bayesian approach allows you to explore sensitivity to prior assumptions.
+Letting the covariance float, we achieve comparable estimates.
+See Table~\ref{table:tests} for our results.
 
-When we fit this spectrum using the same assumptions, (amounting to a $\delta$-function prior on $\log g$.
-global noise amplitude is so low, almost non-existent.
-This demonstrates sensitivity to a prior as well as the fact that surface gravity estimates from spectroscopy may be biased against fits derived from planetary transits.
+% Data description of Gl 51
+The spectrum of Gl~51 was obtained with the \emph{SPEX} instrument \citep{rayner03} on the IRTF.
+The spectrum is a moderate resolution (R $\sim$ 2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $\gt 200$.
+Gl~51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09}.
 
-The second object is Gl 51, an M5 star.
-The spectrum was obtained with the \emph{SPEX} instrument at IRTF.
-The spectrum is at moderate resolution ($R \sim $ 2,000), mid-infrared spectrum from $0.7 - 5.5\mu$m.
-Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}.
-\citep{rojas-ayala12} use the 2.205~$\mu$m \ion{Na}{1} doublet and 2.263~$\mu$m \ion{Ca}{1} triplet and Na doublet lines (their table 2), and their \HK\ index, defined using three passbands: 2.070~$\mu$m-2.090~$\mu$m, 2.235~$\mu$m-2.255~$\mu$m, and 2.360~$\mu$m - 2.380~$\mu$m.
+% What other people get when they fit this star. How does our inference compare?
+To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet and their \HK\ index, defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m.
 The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water.
@@
 They find that the index correlates with spectral type.
-
 \citet{rojas-ayala12} use a custom grid of PHOENIX BT-Settl models run from the PHOENIX web simulator at.
@@
 The strength of the \ion{Ca}{1} triplet increases as $T_{\rm eff}$ and $\Z$ increase.
-
-Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl51.
-
-A Bayesian approach allows you to explore sensitivity to prior assumptions.
+\todo{What does our analysis yield?} Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl 51.
 
 \subsection{Generic Tests}
+Designed to show off what the model framework can actually do.
 
-Designed to show off what the model framework can actually do.
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft, width=2.5in]{figs/stellar_triangle_Kurucz.png}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+These stellar parameters are marginalized over the Chebyshev ($\Cheb$) and noise parameters ($\Cov$).
+Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+}
+\label{fig:stellar_posterior}
+\end{center}
+\end{figure*}
 
@@
 Talk about giving more reasonable errors using regions, in the biased case.
-Show a figure of the regions determined, plotted with the increased variance envelope over the residuals.
+Show a (Figure~\ref{fig:regions}) of the regions determined, plotted with the increased variance envelope over the residuals.
 

### 2014-07-28 copy_HDF5.py
updated kurucz params

@@ -1,0 +1,46 @@
+#!python
+
+#Check to make sure the file has been copied to the node /scratch properly.
+
+#invoke with the following command line arguments
+#check_HDF5.sh existing.hdf5 destination.hdf5
+
+import sys
+import os
+import subprocess
+
+existing = sys.argv[1]
+destination = sys.argv[2]
+
+#compute MD5 sum of existing.hdf5
+def get_md5(file):
+    proc = subprocess.Popen(["md5sum", file], stdout=subprocess.PIPE)
+    out, err = proc.communicate()
+    out = out.decode('UTF-8')
+    #strip the filename and keep the code
+    return out.split()[0]
+
+md_existing = get_md5(existing)
+print("md5sum of {} is {}".format(existing, md_existing))
+
+#check if destination.hdf5 exists
+if os.path.exists(destination):
+    print("{} exists".format(destination))
+
+    md_destination = get_md5(destination)
+    print("md5sum of {} is {}".format(existing, md_destination))
+    #Do the md5sums match up?
+    if md_existing == md_destination:
+        print("md5 sums are equal")
+    else:
+        import time
+        tries = 0
+        while (md_existing != md_destination) and tries < 5:
+            print("md5 sums are not equal, waiting for 2 minutes")
+            time.sleep(120)
+            md_destination = get_md5(destination)
+            print("md5sum of {} is {}".format(existing, md_destination))
+            tries += 1
+else:
+    print("{} does not exist, copying".format(destination))
+    subprocess.call(["cp", existing, destination])

### 2014-07-28 tex/paper.tex
updated PHOENIX file

@@
 The spectrum of Gl~51 was obtained with the \emph{SPEX} instrument \citep{rayner03} on the IRTF.
-The spectrum is a moderate resolution (R $\sim$ 2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $\gt 200$.
+The spectrum is a moderate resolution (R $\sim$ 2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $> 200$.
 Gl~51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09}.
@@
 % What other people get when they fit this star. How does our inference compare?
-To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet and their \HK\ index, defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m.
-The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water.
-For a useful visualization of which spectral features these indices track, see their figure 2.
-The \HK\ index is a monotonic function of temperature, showing more curvature in the spectrum towards cooler $T_{\rm eff}$ and the index is independent of $\Z$ in the range $3000\,{\rm K} \leq T_{\rm eff} \leq 4000\,{\rm K}$.
-Below $3000~{\rm K}$, the index may appear to be metallicity dependent, although they caution that this may be a computational artifact.
-They find that the \HK\ index has a negligible dependence on $\log g$.
-They find that the index correlates with spectral type.
+To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet and their \HK\ index, defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization).
+The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules.
 \citet{rojas-ayala12} use a custom grid of PHOENIX BT-Settl models run from the PHOENIX web simulator at.
@@
 \citet{rajpurohit10} note that the oscillator strengths and opacities of the \ion{Na}{1} and \ion{Ca}{1} lines may be incorrect.
-In the range of stellar parameters for M dwarfs, \citet{rojas-ayala12} note that the strength of the \ion{Na}{1} doublet increases as $T_{\rm eff}$ decreases.
-The strength of the \ion{Ca}{1} triplet increases as $T_{\rm eff}$ and $\Z$ increase.
-\todo{What does our analysis yield?} Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl 51.
+When restricting the spectrum to the range of the indices that they use, we find that \todo{What does our analysis yield?
+It should be in the table.} Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl 51.
 

### 2014-07-28 tex/paper.tex
not picking up cov

@@
 $\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
-$\vT$ & $\{\vT, \vP\}$, the collection of all parameters \\
+$\vT$ & the collection of all stellar parameters \\
 $\vD$ & data spectrum \\

### 2014-07-30 tex/paper.tex
Ready for run on Odyssey

@@
 Section~\ref{subsec:likelihood}).  
+\todo{Talk about tests to show that it is within \% accuracy}.
 
@@
 former, we consider three kernels that contribute to the observed line-of-sight velocity 
-distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
-spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
+distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$,
+treats the instrumental spectral broadening\footnote{For computational
+efficiency, we first pre-convolve the entire raw spectral library with a slight
+under-prediction of $\mathcal{F}_v^{\rm inst}$ and resample the
+spectra to this lower resolution. Because the instrumental PSF acts as a
+low-pass filter, this procedure can reduce the size of the library by a factor
+of 10 or more while also preserving the final Fourier content of the
+spectrum. Then, when a given spectrum is post-processed, the actual $\mathcal{F}_v^{\rm
+inst}$ used is a correction to the original
+under-prediction.}.  For illustrative purposes we assume
+$\mathcal{F}_v^{\rm inst}$ is a Gaussian 
 with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
@@
 and assume only the trivial noise spectrum contributes to the covariance matrix (all elements of 
-$\cov^0$ are zero).
+$\cov^0$ are zero, $\vC$ is diagonal).
 
@@
 we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
-$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this 
-will trigger the addition of a local kernel.
+$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3
+\, a_g^i$), this will trigger the addition of a local kernel\footnote{For
+practical reasons, we do not delete local kernels from the parameter list once they have been
+instantiated. However, if the stellar parameters have moved to a region of
+parameter space where a given local kernel is no longer required, then the
+amplitude of that kernel will be driven towards $0$, the kernel would have a
+negligible contribution to the covariance matrix $\vC$, and the overall behavior of the model would
+be as if the kernel were deleted.}.
 
@@
 hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
-If the stellar parameters have moved to a region of parameter space where the local kernel is no longer required, then the amplitude of the kernel will be driven towards $0$ and the kernel would have a negligible contribution to the covariance matrix $\vC$.
 \end{enumerate}
@@
 
-\todo{What should we do with the commented-out sections?}
-%A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
-Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
-This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
-This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
-
-%The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
-If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
-
-%When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
-Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
-As demonstrated by the Gibbs sampler routine, we explicitly fit for the
-%hyperparameters of the covariance kernels in parallel with the
-%stellar parameters. While including these extra parameters does increase the
-%dimensionality and complexity of our model, they do provide several advantages.
-%\emph{A priori} we do not know which regions of the spectrum are improperly
-%modeled. The covariance hyperparameters provide a framework to identify these
-%regions iteratively and in a self-consistent manner.
-
-%The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-%Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
-When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
-This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
-However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
-
-%\subsection{Applications \comm{untouched}}
-%\label{subsec:learning}
-%By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
-After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
-The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
-Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
-
-%Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
-Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
-Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
-Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
-In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
-This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
-
-%Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
-Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
-This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
-
 \section{Examples}
@@
 The first data set is a high resolution optical spectrum of WASP-14, an F5 star \citep{joshi09}.
-The second data set is a moderate resolution mid-infrared spectrum of Gl~51, a M5 star \citep{rojas-ayala12}.
+The second data set is a moderate resolution mid-infrared spectrum of Gl~51, an M5 star \citep{rojas-ayala12}.
 
@@
 Our spectrum of WASP-14 was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}, \citealt{mink11}) on the Fred Laurence Whipple Observatory 1.5m telescope on June 14th, 2009.
-\emph{TRES} delivers a 51 echelle spectrum spanning the full optical spectrum (3860--9100\AA) at a resolution of $R \sim$44,000.
+\emph{TRES} delivers a 51 echelle spectrum spanning the full optical range (3860--9100\AA) at a resolution of $R$$\sim$44,000.
 \citet{torres12} originally acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts.
-The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave.
-In the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.
+The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave and in the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.
 
 % What other people get when they fit this star. How does our inference compare?
-\citet{torres12} measured the stellar properties of WASP-14 using several different methods including MOOG, SME, and SPC.
-The MOOG and SME methods are not directly comparable to our technique because they measure individual line strengths.
-However, because SPC utilizes a large swath of the spectrum and a model spectrum, we can use the same spectral library and make the same assumptions.
-The SPC analysis within \citet{torres12} focused on a region of spectrum from 5050--5360\AA\ and used many methods.
-Additionally, there are systematics introduced by metallicity assumptions and radiative transfer.
-Using this same spectral library as SPC, we recover the same parameters.
-When we fit this spectrum using the same assumptions, (amounting to a $\delta$-function prior on $\log g$).
-This demonstrates sensitivity to a prior as well as the fact that surface gravity estimates from spectroscopy may be biased against fits derived from planetary transits.
-A Bayesian approach allows you to explore sensitivity to prior assumptions.
-Letting the covariance float, we achieve comparable estimates.
-See Table~\ref{table:tests} for our results.
+\citet{torres12} determined the stellar properties of WASP-14 using several different methods including {\tt MOOG}, {\tt SME}, and {\tt SPC}.
+While the {\tt MOOG} and {\tt SME} methods are not directly comparable to our technique because they measure individual line strengths, {\tt SPC} utilizes a moderately-sized region of spectrum and a pre-computed synthetic library that enables a direct comparison with our technique.
+When we make the same assumptions in \citet{torres12}, such as using the same region of spectrum (5050--5360\AA), the same spectral library, and fixing $\log g$ to the value inferred from the exoplanet transit analysis by \citet{joshi09} (amounting to a $\delta$-function prior on $\log g$$=$4.29), our inference of stellar parameters agrees to within 1$\sigma$ (see Table~\ref{table:tests}).
+Interestingly, when we allow $\log g$ to float, we infer a lower $T_{\rm eff}$ and $\Z$.
+These results demonstrate the sensitivity to assumptions about $\log g$ as well as the fact that there may exist biases in surface gravity estimates from either spectroscopy, exoplanetary transits, or both.
 
 % Data description of Gl 51
-The spectrum of Gl~51 was obtained with the \emph{SPEX} instrument \citep{rayner03} on the IRTF.
-The spectrum is a moderate resolution (R $\sim$ 2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $> 200$.
-Gl~51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09}.
+The spectrum of Gl~51, an M5 star, was downloaded from \emph{IRTF} spectral library \citep{cushing05, rayner09}.
+The data was obtained with the \emph{SPEX} instrument \citep{rayner03} on the NASA \emph{IRTF}.
+The spectrum is a moderate resolution (R$\sim$2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $>$400 at 2.1$\mu$m.
 
 % What other people get when they fit this star. How does our inference compare?
-To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet and their \HK\ index, defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization).
-The index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules.
-\citet{rojas-ayala12} use a custom grid of PHOENIX BT-Settl models run from the PHOENIX web simulator at.
-The note that the models fit the spectrum match the overall spectral shape well, although the models consistently underpredict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
-\citet{rajpurohit10} note that the oscillator strengths and opacities of the \ion{Na}{1} and \ion{Ca}{1} lines may be incorrect.
-When restricting the spectrum to the range of the indices that they use, we find that \todo{What does our analysis yield?
-It should be in the table.} Because the PHOENIX grid by \citet{husser13} has more range/uniformity of parameters, we use this to fit Gl 51.
+To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the equivalent widths of the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet along with their \HK\ index, which is defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization).
+This index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules.
+In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
+They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
+\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
+When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{The results of our analysis will be in Table~\ref{table:tests}.}
 
 \subsection{Generic Tests}
-Designed to show off what the model framework can actually do.
+
+In addition to running tests under the same assumptions as historical analyses, we also allow the region kernels to play a role.
+
+%\begin{rant about why this method rocks}
+There are many advantages to placing spectroscopic inference within a Bayesian framework.
+Our technique uses as much as possible of the spectrum.
+You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
+True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a disparate gap between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
+These are coming up with ``fit metrics'' that are similar to distance functions or figures of merit that would be used in approximate Bayesian computing.
+
+These figures of merit make a lot of sense, true.
+But given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
+In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
+This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
+This is one of the great things about a \emph{forward-model}.
+
+Ultimately, with the current state of physical models, generating a spectrum from a limited grid, or only being able to tweak certain parameters, the posterior that one can derive is ultimately limited by the accuracy of the spectral synthesis, including any systematic biases that may exist (and they do, given the current spread of performance of stellar libraries).
+
+The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, allows the model-data mismatch to become a parameter in the model.
+That way there is shrinkage between estimates.
+
+%\end{rant}
+
+Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines.
+Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum.
+This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v.
+PHOENIX (-14.2) global cov amplitudes.
+
+A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
+Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
+This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
+This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
+
+The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
+If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
+
+When iteratively fitting a spectrum, we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
+As demonstrated by the Gibbs sampler routine, we explicitly fit for the
+hyperparameters of the covariance kernels in parallel with the
+stellar parameters. While including these extra parameters does increase the
+dimensionality and complexity of our model, they do provide several advantages.
+\emph{A priori} we do not know which regions of the spectrum are improperly
+modeled. The covariance hyperparameters provide a framework to identify these
+regions iteratively and in a self-consistent manner.
+
+The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+
+Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
+This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
+However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
 
@@
   \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
+And which technique, which may be a result of choosing certain indices, certain models, or both.
+Therefore it is in our interest to use as wide a region of spectrum as possible.
   \item Both spectral libraries have stars that they perform better and worse on.
@@
 \label{sec:discussion}
+
+Why is it a good idea to use our likelihood function for spectroscopic inference?
+
+By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
+After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
+The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
+Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
+
+Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
+Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
+Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
+Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
+In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
+This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
+
+Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
+Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
+This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
 
@@
 \begin{deluxetable}{llrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
-\tablecaption{\label{table:tests} Tests}
+\tablecaption{\label{table:tests} Results}
 \tablehead{
@@
 \startdata
-WASP-14 & 22-24 & 5060-5315 & Kurucz & 6310 & 52 & 3.91  & 0.08 & -0.38  & 0.03 & Poisson only\\
-WASP-14 & 22-24 & 5060-5315& Kurucz & 6314 & 50 & 3.92 & 0.08 & -0.39 & 0.03 & Matern only\\
-WASP-14 & 22-24 & 5060-5315& Kurucz & 6512 & 32 & \multicolumn{2}{c}{\nodata} & -0.27 & 0.02 & Matern + fixed $\log g = 4.29$\\
-WASP-14 & 23 & & PHOENIX & 6021 & 16 & 3.83 & 0.03 & -0.50 & 0.01 & Poisson only \\
-WASP-14 & 23 & & PHOENIX & & & & & & & Poisson only, masked \\
-WASP-14 & 23 & & PHOENIX & 5965 & 70 & 3.78 & 0.09 & -0.68 & 0.07 & Matern only, no regions (conv?) \\
-WASP-14 & 23 & & PHOENIX & 5883 & 55 & 3.71 & 0.07 & -0.74 & 0.05 & Matern and regions (conv?) \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & 6117 & 30 & 3.73 & 0.06 & -0.52 & 0.02 & Poisson only \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & 5865 & 43 & 3.2 & 0.08 & -0.85 & 0.03 & Matern only. Because of 24. \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Matern + fixed $\log g = 4.29$ \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Regions included \\
+WASP-14 & 23 & & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
+WASP-14 & 23 & & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
+WASP-14 & 23 & & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
+WASP-14 & 23 & & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
+WASP-14 & 23 & & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
+WASP-14 & 23 & & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
+WASP-14 & 23 & & PHOENIX & & & & & & & Global and local \\
+WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
+WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
+WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
+WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
 \enddata
+\tablenotetext{a}{Using the same assumptions in the \citet{torres12} {\tt SPC} analysis: same region of spectrum, same spectral library, and fixing $\log g$=$4.29$, the value determined by \citet{joshi09} from the exoplanet transit.
+For the WASP-14 spectrum acquired by \emph{TRES}, \citet{torres12} use {\tt SPC} to determine $T_{\rm eff}$=6507$\pm$50K and $\Z$=-0.15$\pm$0.08.
+With {\tt MOOG}, they determine $T_{\rm eff}$=6150$\pm$75K and $\Z$=-0.34$\pm$0.10.
+Because {\tt SPC} uses a cross-correlation technique which only accounts for the white noise and we model residual spectrum as correlated noise, there may be differences in the uncertainty estimate.}
 \end{deluxetable}

### 2014-07-31 tex/paper.tex
skip runs that haven't finished

@@
 
-\subsection{Generic Tests}
-
-In addition to running tests under the same assumptions as historical analyses, we also allow the region kernels to play a role.
-
-%\begin{rant about why this method rocks}
-There are many advantages to placing spectroscopic inference within a Bayesian framework.
-Our technique uses as much as possible of the spectrum.
-You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
-True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a disparate gap between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
-These are coming up with ``fit metrics'' that are similar to distance functions or figures of merit that would be used in approximate Bayesian computing.
-
-These figures of merit make a lot of sense, true.
-But given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
-In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
-This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
-This is one of the great things about a \emph{forward-model}.
-
-Ultimately, with the current state of physical models, generating a spectrum from a limited grid, or only being able to tweak certain parameters, the posterior that one can derive is ultimately limited by the accuracy of the spectral synthesis, including any systematic biases that may exist (and they do, given the current spread of performance of stellar libraries).
-
-The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, allows the model-data mismatch to become a parameter in the model.
-That way there is shrinkage between estimates.
-
-%\end{rant}
-
-Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines.
-Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum.
-This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v.
-PHOENIX (-14.2) global cov amplitudes.
+\subsection{WASP-14} \label{subsec:wasp}
+To evaluate the performance of our methodology, we perform a series of tests using WASP-14 and Gl~51.
+We use the PHOENIX grid (made with version XX PHOENIX, \citealt{husser13}), which spans wavelengths from XX to XX.
+The grid is very high quality and has been hand-calibrated against the Sun and Vega, therefore it performs very well for solar like stars.
+Because WASP-14 is an optical spectrum, we can use the high-quality CfA Kurucz grid, synthesized by Laird and Morse and used by {\tt SPC}.
+Because the spectrum of Gl~51 is the mid-infrared, we are unable to use the CfA grid for these comparisons.
 
@@
 
-When iteratively fitting a spectrum, we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
-Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
-As demonstrated by the Gibbs sampler routine, we explicitly fit for the
-hyperparameters of the covariance kernels in parallel with the
-stellar parameters. While including these extra parameters does increase the
-dimensionality and complexity of our model, they do provide several advantages.
-\emph{A priori} we do not know which regions of the spectrum are improperly
-modeled. The covariance hyperparameters provide a framework to identify these
-regions iteratively and in a self-consistent manner.
+For demonstrative purposes, the first fits we perform is using only the Poisson variances from the original spectrum (WK1 and WP1), ignoring any correlations that might arise in the residual spectrum.
+This is equivalent to setting $b=1$ and all other amplitudes of $\cov$ to 0, and allows our likelihood to simplify to the familiar form of $\chi^2$.
+While the uncertainty estimate returned from these fits is remarkable, it also leads to the situation of overconfidence.
+Such a small uncertainty arises because such a likelihood function assumes that the data set is generated from the model, meaning that the only way in which the data set might differ from the model is through the addition of white Poisson noise.
+Unfortunately, this is incorrect.
+While the synthetic stellar model is \emph{incredibly good}, it is not a perfect representation of the star, nor could it be given that a synthetic spectrum from a grid has only 3 $\vt_\ast$ to tweak.
+This means that some line strengths might be different, whether due to slight inaccuracies in the atomic constants or metallicity compositions that differ in a few elements different from what the Sun might be.
+Thus, the data set differs from the model not just from added Poisson uncertainties, but also \emph{systematically} because the model cannot be a perfect representation of reality.
+This is equivalent to fitting a straight line to a series of data points whose variance is drastically underestimated.
+A linear least squares fit $\chi^2$ routine might return a small uncertainty estimate, however the slope and intercept of the line might be drastically biased.
+Therefore, it is clear that they model requires a way to account for this additional source of ``noise.'' Make the distinction that the Kurucz model fits \emph{very} well, but still requires some global covariance simply because of the three parameter problem.
 
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
-When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
-This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
-However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
+Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
+And which technique, which may be a result of choosing certain indices, certain models, or both.
+Therefore it is in our interest to use as wide a region of spectrum as possible.
+Both spectral libraries have stars that they perform better and worse on.
+This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
+Eg, WASP 14 v.
+Gl~51.
 
@@
 
+Because the PHOENIX spectrum is computed over a wide range, there are often some egregious ``outlier'' lines that appear in the spectrum.
+To compare our approach with the common approach of ``sigma-clipping,'' for the next test (WP2), we manually mask any region of spectrum which has a more than 3 sigma residual near the best fit parameters.
+As you can see, this has the effect of actually changing the final stellar estimate significantly, while also reducing the uncertainty in the stellar parameters.
+Sigma-clipping like this is a procedural approach and is not justified in a probabilistic framework \citep{hogg10}.
+
+For the next tests, (WK2 and WP3), we introduce the global covariance kernel to the fit.
+This allow the model-data mismatch, and we can see that the uncertainty estimate naturally inflates to capture the discrepancy.
+Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines.
+Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum.
+This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v.
+PHOENIX (-14.2) global cov amplitudes.
+I would like to point out that this came about \emph{self-consistently} through the model, not something that was added in post-facto as is commonly done (cough, \citep{torres12, buchhave12}).
+Because our uncertainty is derived through the model itself, it \emph{preserves correlations} between parameter estimates.
+Here we see that $\vt_\ast$ are highly correlated.
+If you just slapped on 50K errorbars to everything, you would be destroying this information.
+
+For the next tests, (), we introduce the local covariance kernels to track outlier lines.
+These kernels can map out the regions of the spectrum which are outliers.
+The glory of these local kernels are that the parameters allow to play a role in data driven spectra.
+See figure XX for a cool visualization of what it looks like.
+
 \begin{figure*}[!htb]
@@
 
+\subsection{Gl~51}
+Gl~51 tests. Using no covariance, covariance, just the lines, just same region of spectrum. Adding local covariance kernels to Na and Ca lines that are underpredicted.
+
 Talk about giving more reasonable errors using regions, in the biased case.
 Show a (Figure~\ref{fig:regions}) of the regions determined, plotted with the increased variance envelope over the residuals.
-
-\begin{itemize}
-  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
-And which technique, which may be a result of choosing certain indices, certain models, or both.
-Therefore it is in our interest to use as wide a region of spectrum as possible.
-  \item Both spectral libraries have stars that they perform better and worse on.
-  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
-\end{itemize}
-
-Summary end of this section is that we are
 
@@
 Why is it a good idea to use our likelihood function for spectroscopic inference?
+
+%\begin{rant about why this method rocks}
+There are many advantages to placing spectroscopic inference within a Bayesian framework.
+Our technique uses as much as possible of the spectrum.
+You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
+True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a disparate gap between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
+These are coming up with ``fit metrics'' that are similar to distance functions or figures of merit that would be used in approximate Bayesian computing.
+
+These figures of merit make a lot of sense, true.
+But given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
+In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
+This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
+This is one of the great things about a \emph{forward-model}.
+
+Ultimately, with the current state of physical models, generating a spectrum from a limited grid, or only being able to tweak certain parameters, the posterior that one can derive is ultimately limited by the accuracy of the spectral synthesis, including any systematic biases that may exist (and they do, given the current spread of performance of stellar libraries).
+
+The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, allows the model-data mismatch to become a parameter in the model.
+That way there is shrinkage between estimates.
+
+Adding in additional orders to the spectrum allows the posterior to broaden (if each order has disparate predictions about the stellar parameters) or to contract, since we are also fitting for the model-data mismatch.
+If all orders match well (case of Kurucz), then the global kernels will have low amplitude.
+However, if certain orders predict different stellar parameters, then for the one set of stellar parameters that is shared by the dataset, the global kernels will inflate for the other orders to absorb the model-data mismatch.
+
+%\end{rant}
+
+When iteratively fitting a spectrum, we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
+Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
+As demonstrated by the Gibbs sampler routine, we explicitly fit for the
+hyperparameters of the covariance kernels in parallel with the
+stellar parameters. While including these extra parameters does increase the
+dimensionality and complexity of our model, they do provide several advantages.
+\emph{A priori} we do not know which regions of the spectrum are improperly
+modeled. The covariance hyperparameters provide a framework to identify these
+regions iteratively and in a self-consistent manner.
+
+The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+
+Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
+When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
+This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
+However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
 
@@
 
-\begin{deluxetable}{llrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
+\begin{deluxetable}{lllrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
 \tablecaption{\label{table:tests} Results}
 \tablehead{
-  \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
+  \colhead{Test ID} & \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
   \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
@@
 \startdata
-WASP-14 & 23 & & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
-WASP-14 & 23 & & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
-WASP-14 & 23 & & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
-WASP-14 & 23 & & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
-WASP-14 & 23 & & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
-WASP-14 & 23 & & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
-WASP-14 & 23 & & PHOENIX & & & & & & & Global and local \\
-WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
-WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
-WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
-WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
+WK1 & WASP-14 & 23 & 5135-5235 & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
+WK2 & WASP-14 & 23 & 5135-5235 & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
+WK3 & WASP-14 & 23 & 5135-5235 & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
+WP1 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
+WP2 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
+WP3 & WASP-14 & 23 & 5135-5235 & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
+WP4 & WASP-14 & 23 & 5135-5235 & PHOENIX & & & & & & & Global and local \\
+&WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
+&WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
+&WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
+&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
+&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
+&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
+&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
+GP1 &Gl~51 & \nodata & IR & PHOENIX & & & & & & & Poisson\\
 \enddata

### 2014-07-31 hdfmultiple.py
skip runs that haven't finished

@@ -61,7 +61,15 @@ else:
     sys.exit("Must specify either --dir or --files")
 
 import h5py
-hdf5list = [h5py.File(file, "r") for file in files]
+#Because we are impatient and want to compute statistics before all the jobs are finished, there may be some
+# directories that do not have a flatchains.hdf5 file
+hdf5list = []
+for file in files:
+    try:
+        hdf5list += [h5py.File(file, "r")]
+    except OSError:
+        print("{} does not exist, skipping.".format(file))
+#hdf5list = [h5py.File(file, "r") for file in files]
 stellarlist = [hdf5.get("stellar") for hdf5 in hdf5list]
 
 for stellar in stellarlist:

### 2014-07-31 tex/paper.tex
New draft for Sean

--- 2014-07-30 tex/paper.tex
@@
 While the {\tt MOOG} and {\tt SME} methods are not directly comparable to our technique because they measure individual line strengths, {\tt SPC} utilizes a moderately-sized region of spectrum and a pre-computed synthetic library that enables a direct comparison with our technique.
-When we make the same assumptions in \citet{torres12}, such as using the same region of spectrum (5050--5360\AA), the same spectral library, and fixing $\log g$ to the value inferred from the exoplanet transit analysis by \citet{joshi09} (amounting to a $\delta$-function prior on $\log g$$=$4.29), our inference of stellar parameters agrees to within 1$\sigma$ (see Table~\ref{table:tests}).
-Interestingly, when we allow $\log g$ to float, we infer a lower $T_{\rm eff}$ and $\Z$.
+When we make the same assumptions in \citet{torres12}, such as using the same region of spectrum (5050--5360\AA), the same spectral library, and fixing $\log g$ to the value inferred from the exoplanet transit analysis by \citet{joshi09} (amounting to a $\delta$-function prior on $\log g$$=$4.29), our inference of stellar parameters agrees to within 1$\sigma$ (see Table~\ref{table:tests}, WKX).
+Interestingly, when we allow $\log g$ to float, we infer a lower $T_{\rm eff}$ and $\Z$ (WKX2).
 These results demonstrate the sensitivity to assumptions about $\log g$ as well as the fact that there may exist biases in surface gravity estimates from either spectroscopy, exoplanetary transits, or both.
@@
 
-\subsection{Generic Tests}
+\subsection{WASP-14} \label{subsec:wasp}
 
-In addition to running tests under the same assumptions as historical analyses, we also allow the region kernels to play a role.
+We use two synthetic libraries to fit WASP-14.
+The first is the recent publicly available PHOENIX grid (PHOENIX version 16, \citealt{husser13}), which spans wavelengths from 500\AA\ to 5.5$\mu$m and $T_\textrm{eff}$ from 2300-12,000K.
+The second is a custom ``CfA grid'' synthesized using Kurucz models by Laird and Morse and is the same grid used by {\tt SPC}.
+This grid is very high quality because it has been hand-calibrated against the Sun and Vega, therefore it performs very well for solar like stars, however it only spans the comparatively narrow wavelength range 5050--5360\AA.
 
-%\begin{rant about why this method rocks}
-There are many advantages to placing spectroscopic inference within a Bayesian framework.
-Our technique uses as much as possible of the spectrum.
-You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
-True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a disparate gap between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
-These are coming up with ``fit metrics'' that are similar to distance functions or figures of merit that would be used in approximate Bayesian computing.
+For demonstrative purposes, the first fits we perform use only the Poisson variances from the original data set (WK1 and WP1) and ignore any correlations that might arise in the residual spectrum.
+This is equivalent to setting $b=1$ and all other amplitudes of $\cov$ to 0, and simplifies the likelihood to the familiar form of $\chi^2$.
+While the parameter uncertainty returned from these fits appears remarkably good, it is also unrealistic.
+This small uncertainty results because any likelihood function assumes that the data set is generated from the model\footnote{For example, other ``robust'' likelihood functions like the Student-t distribution assume that there is some likelihood of non-Gaussian ``outlier'' points that will occur.
+The Gaussian process likelihood that we assume captures the type of ``outliers'' that we might expect from our data.}, meaning that in this paradigm the only way the data set can differ from the model is through the addition of white Poisson noise.
+While both the PHOENIX and CfA stellar grids are \emph{incredibly good}, this assumption is invalid because the three 3 $\vt_\ast$ parameters are not enough to account for the behavior of thousands of spectral lines that might be in a reasonably large swath of spectrum.
+Some line strengths might be different from the synthetic model due to slight inaccuracies in the atomic constants or metallicity compositions that are different from solar in a few elements.
+Thus, in practice, a data spectrum is different from a synthetic model not just by added Poisson noise, but is also \emph{systematically} different in many spectral lines because the model cannot be a perfect representation of reality.
+Thus, assuming only white Poisson noise for a spectroscopic fit is equivalent to fitting a straight line to a series of data points whose variance is drastically underestimated.
+A linear least squares fit $\chi^2$ routine might return a small uncertainty estimate, however the slope and intercept of the line might be drastically biased.
+Therefore, it is clear that they model requires a way to account for this additional source of systematic ``noise.''
 
-These figures of merit make a lot of sense, true.
-But given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
-In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
-This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
-This is one of the great things about a \emph{forward-model}.
+While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
+For demonstrative purposes, for the next test (WP2) we use the common technique of ``sigma-clipping,'' whereby we manually mask any region of the spectrum which has a more than a 3 sigma residual near the best fit parameters, refit, and then proceed until there are no more outliers.
+Masking out ``outlier'' spectral lines has the effect of actually changing the final stellar estimate significantly, while also reducing the uncertainty in the stellar parameters.
+However, several questions arise about the quality of the inference on the stellar parameters.
+Because sigma-clipping is a procedure, it is impossible to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
+For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
+Clipping these lines would discard important information and lead to an incorrect inference of stellar parameters.
 
-Ultimately, with the current state of physical models, generating a spectrum from a limited grid, or only being able to tweak certain parameters, the posterior that one can derive is ultimately limited by the accuracy of the spectral synthesis, including any systematic biases that may exist (and they do, given the current spread of performance of stellar libraries).
-
-The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, allows the model-data mismatch to become a parameter in the model.
-That way there is shrinkage between estimates.
-
-%\end{rant}
-
-Note that using global covariance actually lets the parameters float a bit more, and provides some cushion against the tyranny of too many lines.
-Only so many parameters, can't infinitely tweak the model spectrum for any large region of spectrum.
-This parameterizes the badness of fit, better models will have a lower covariance, as evidenced by the Kurucz (-14.6) v.
-PHOENIX (-14.2) global cov amplitudes.
+These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+For the next tests, (WK2 and WP3), we introduce the global covariance kernel (Eqn.~\ref{eqn:global}) to the fit.
+We can see that the amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum.
+Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the tyranny of too many lines, which cause the unreasonably low uncertainty of the previous tests.
+Therefore, including the global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance, as evidenced by the Kurucz ($10^{-14.6}$) v.
+PHOENIX ($10^{-14.2}$) global covariance amplitudes.
+\todo{Figure out a reasonable way to present nuisance parameter values}.
+It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
+Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
+Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
 
@@
 This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
-This posterior is the final estimate of the stellar parameters which \emph{incorporates} any inherent uncertainty due to model mismatch (via the covariance hyperparameters) and flux-calibration (via the Chebyshev polynomials).
-
-The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters.
-If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.
-
-When iteratively fitting a spectrum, we continue to add line covariance kernels until we have covered all of the high amplitude residuals.
-Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered.
-As demonstrated by the Gibbs sampler routine, we explicitly fit for the
-hyperparameters of the covariance kernels in parallel with the
-stellar parameters. While including these extra parameters does increase the
-dimensionality and complexity of our model, they do provide several advantages.
-\emph{A priori} we do not know which regions of the spectrum are improperly
-modeled. The covariance hyperparameters provide a framework to identify these
-regions iteratively and in a self-consistent manner.
-
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch.
-When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length.
-This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result.
-However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).
+This posterior is the final estimate of the stellar parameters which incorporates any inherent uncertainty due to flux-calibration (via $\cheb$) and model mismatch (via $\cov$).
 
@@
   \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\Cheb$) and noise parameters ($\Cov$).
+These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
@@
 
+For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
+These kernels can map out the regions of the spectrum which are outliers.
+The glory of these local kernels are that the parameters allow to play a role in data driven spectra.
+\emph{A priori} we do not know which regions of the spectrum are improperly modeled.
+The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
+See Figure~\ref{fig:regions} for a cool visualization of what it looks like.
+\todo{Runs are almost converged.
+This can feed into a discussion about tweaking models.}
+
 \begin{figure*}[!htb]
@@
   \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
-  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.}
+  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
+Plotted with the increased variance envelope over the residuals.
+}
 \label{fig:regions}
@@
 
-Talk about giving more reasonable errors using regions, in the biased case.
-Show a (Figure~\ref{fig:regions}) of the regions determined, plotted with the increased variance envelope over the residuals.
+In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
+Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.
 
-\begin{itemize}
-  \item Spectral parameters can vary by a large margin depending on which spectral library you use ($200~K$ or more).
-And which technique, which may be a result of choosing certain indices, certain models, or both.
-Therefore it is in our interest to use as wide a region of spectrum as possible.
-  \item Both spectral libraries have stars that they perform better and worse on.
-  \item This will be reflected in the increased level of global noise, and number of ``bad'' regions that have been instantiated.
-\end{itemize}
+\subsection{Gl~51}
+\todo{Tests run using no covariance, covariance, just the lines, just same region of spectrum. Adding local covariance kernels to Na and Ca lines that are underpredicted.}
 
-Summary end of this section is that we are
-
-1) validating other techniques (WASP14, Kurucz, etc)
-2) and two different synthetic libraries give different answers
-
-this can feed into a discussion about tweaking models.
+Because the spectrum of Gl~51 is the mid-infrared, we are unable to use the CfA grid for these comparisons. All Gl~51 tests were made using the PHOENIX grid.
 
@@
 \label{sec:discussion}
+%\begin{rant about why this method rocks}
+There are many advantages to placing spectroscopic inference within a Bayesian framework.
+Given the large costs of acquiring a high signal to noise spectrum, it would be advantageous to utilize as much as possible of the spectrum for performing inference.
+You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
+True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a gap in the chain of logic between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
+These ``fit metrics'' (line indices) are similar to distance functions or a figure of merit.
 
-Why is it a good idea to use our likelihood function for spectroscopic inference?
+These figures of merit do make a lot of sense, and one can see why they are used, but given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, instead of going through the intermediary step of a figure of merit, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
+In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
+This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
+This is one of the great things about a \emph{forward-model}.
 
+The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
+When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
+When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters.
+This hierarchical model also works within a single spectrum when fitting more than one spectral order.
+If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
+However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
+%\end{rant}
+
+The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
 By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
@@
 
-We can (should?) add an additional level to the hierarchy of
-  hyperparameters and add parameters to describe the \emph{population
-  characteristics} of poorly modeled spectral lines (mostly the typical width,
-  amplitude of these lines).  This will tell us about the frequency and
-distribution of spectral modelling errors.
+\todo{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
+This will tell us about the frequency and distribution of spectral modelling errors.
+Or we could just speculate about it.}
 

### 2014-08-19 tex/paper.tex
set up for kurucz
lots of minor Sect 1, 2 edits; first iteration of Sect 3 edits

@@
 \newcommand{\KK}{\mathcal{K}}
-\newcommand{\Kglobal}{\KK_{\textrm{g}}}
-\newcommand{\Klocal}{\KK_l}
+\newcommand{\Kglobal}{\KK^{\textrm{g}}}
+\newcommand{\Klocal}{\KK^l}
 
@@
 \newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}
+\newcommand{\hili}[1]{ \textcolor{green}{#1}}
 
@@
 \affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
-%\email{iczekala@cfa.harvard.edu}
+\email{iczekala@cfa.harvard.edu}
+
+\begin{abstract}
+An abstract needs to go here. \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd \\
+tbd
+\end{abstract}
+\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}
 
@@
 ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
-subset of the spectral data, usually in the form of spectral line equivalent widths and/or local 
-continuum shapes.  A combined sequence of the ratios of these quantities are found to be especially 
-sensitive to one key parameter in the models, and therefore can be used to quantify it 
-\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  The line index comparison approach has 
-the major advantage of being trivially fast, but each index relationship is only informative over 
-a limited range of parameter-space.  A second technique exploits the cross-correlation of an 
+subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
+shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
+to one key parameter in the models, and therefore can be used to quantify it 
+\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
+the advantage of being trivially fast, but each index relationship is only informative over a 
+limited region of parameter-space.  A second technique exploits the cross-correlation of an 
 observed spectrum with a suite of model templates to identify an optimized set of parameters, 
-usually with some preferential weighting applied to specific spectral regions \citep[e.g., {\tt 
-SPC};][]{buchhave12}.  In this case, the speed advantage is maintained while taking advantage of 
-more data content (particularly in the spectral dimension), thereby achieving high precision even 
-for data with comparatively low sensitivity; the disadvantage is that the model quality and 
-parameter inferences are assessed in an empirical, rather than probabilistic, framework.  The third 
-approach of interest prefers a direct, pixel-by-pixel comparison between model and observed 
-spectra, and is usually associated with a spectral synthesis front-end (e.g., {\tt MOOG}; 
-\citealt{sneden73}; {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric 
-flexibility (e.g., one can fit for atomic abundances or arbitrary structures) and a proper 
-inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian 
-format;][]{shkedy07,schoenrich13}, but it is computationally expensive and can be biased by model 
-systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a 
-pre-selected subsample of the data.
+usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
+SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
+used (particularly in the spectral dimension), thereby achieving higher precision even for data 
+with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
+inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
+employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
+a spectral synthesis front-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
+\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
+for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
+least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
+it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
+practice, this often requires an effort focused on a pre-selected subsample of the data.
 
@@
 (stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
-residual pixel-by-pixel correlations that arise when comparing observed spectra to intrinsically 
+residual pixel-to-pixel correlations that arise when comparing observed spectra to intrinsically 
 imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
@@
 methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
-matrix formalism.  Some relevant tests and example applications (specifically for a high resolution 
-optical spectrum of an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of 
-the method are described in Section \ref{sec:examples}.  Finally, a discussion of the potential 
-utility of the technique, and the possibility of extending it to develop data-driven spectral 
-models, is provided in Section \ref{sec:discussion}.
+matrix formalism.  Some tests and example applications (for a high resolution optical spectrum of 
+an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of the method are 
+described in Section \ref{sec:examples}.  Finally, a discussion of the potential utility of the 
+technique, and the possibility of extending it to develop data-driven spectral models, is provided 
+in Section \ref{sec:discussion}.  \\
 
@@
 
-Here we describe a generative Bayesian modeling framework that confronts some of the key obstacles 
-in the spectroscopic inference problem.  The goal of this approach is to conservatively extract the 
-maximal amount of information about a prescribed (and degenerate) parameter set by forward-modeling 
-an observed spectrum, while also recognizing and explicitly accounting for the covariances and 
-biases introduced by pathologically imperfect models or calibrations.  The method is modular, and 
-therefore can easily incorporate additional physical or nuisance parameters as desired without 
-sacrificing an accurate reflection of the limitations in the data.  Moreover, with a well-crafted 
-observational sample, this data-driven approach should ultimately enable us to systematically learn 
-how synthetic spectral models can be improved.  The specific applications discussed here are 
-related to the spectra of individual stars, but the methodology is generic (and could be used for 
-composite spectra of unresolved stellar clusters, galaxies, etc.).
-
-The remainder of this section describes the mechanics of this modeling framework.  First, a model 
+In this Section, we describe a generative Bayesian modeling framework that confronts some of the 
+key obstacles in the spectroscopic inference problem.  The goal of this approach is to 
+conservatively extract the maximal amount of information about a prescribed (and usually 
+degenerate) parameter set by forward-modeling an observed spectrum, while also recognizing and 
+explicitly accounting for the covariances and biases introduced by pathologically imperfect 
+models.  The method is modular, and therefore can easily incorporate additional physical or 
+nuisance parameters as desired without sacrificing an accurate reflection of the limitations in the 
+data.  Moreover, with a well-crafted observational sample, this data-driven approach should 
+ultimately enable us to systematically learn how synthetic spectral models can be improved.  The 
+specific applications discussed here are related to the spectra of individual stars, but the 
+methodology is generic (and could be used for the composite spectra of unresolved stellar clusters, 
+galaxies, etc.).
+
+The remainder of this Section describes the mechanics of this modeling framework.  First, a model 
 spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
@@
 synthesis is sufficiently high to make a detailed exploration of parameter space (particularly for 
-data with a large spectral range) considerably less appealing.  A related, but different approach 
-is to eschew forward modeling entirely (and therefore repeated spectral syntheses and/or library 
-interpolations), and instead evaluate the models only at the discrete grid points of the library.  
+data with a large spectral range) considerably less appealing.  An alternative approach eschews 
+forward modeling entirely (and therefore repeated spectral syntheses and/or library 
+interpolations), and instead evaluates the models only at the discrete grid points of the library.  
 Then, these discretized samples of the posterior probability density can be interpolated to an 
@@
 
-Here we opt to take the computationally expedient approach that employs a library of model spectra, 
-$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
-(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
-that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
-generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
-$\vt_{\ast}$ must be interpolated from among the spectral library, 
-\begin{equation}
+We opt for the computationally expedient approach that employs a library of model spectra, 
+$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
+[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
+{\it any} ``front-end" that generates a model spectrum.  In our adopted approach, the model 
+spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
+\begin{equation} \label{eqn:interp}
 \flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
-\label{eqn:interp}
 \end{equation}
@@
 is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
-undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
+undesirable level of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation quality 
 can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
-calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
-with the corresponding library spectrum.  After an extensive exploration of such calculations 
+$\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the interpolated spectrum with the 
+corresponding library spectrum.  After an extensive exploration of such calculations 
 \citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
@@
 tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
-be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
+be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
 could be avoided if the spectral library was computed over a refined grid (with a substantial 
@@
 Section~\ref{subsec:likelihood}).  
-\todo{Talk about tests to show that it is within \% accuracy}.
+%% \todo{tests regarding interpolation accuracy}.
 
@@
 
-Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ will be highly over-sampled compared to a 
-typical observed spectrum, and does not account for several additional observational and 
+Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
+observed spectrum, and does not account for several additional observational and 
 instrumental effects that become important in comparisons with real data.  Therefore, a certain 
@@
 We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
-spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
-former, we consider three kernels that contribute to the observed line-of-sight velocity 
-distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$,
-treats the instrumental spectral broadening\footnote{For computational
-efficiency, we first pre-convolve the entire raw spectral library with a slight
-under-prediction of $\mathcal{F}_v^{\rm inst}$ and resample the
-spectra to this lower resolution. Because the instrumental PSF acts as a
-low-pass filter, this procedure can reduce the size of the library by a factor
-of 10 or more while also preserving the final Fourier content of the
-spectrum. Then, when a given spectrum is post-processed, the actual $\mathcal{F}_v^{\rm
-inst}$ used is a correction to the original
-under-prediction.}.  For illustrative purposes we assume
-$\mathcal{F}_v^{\rm inst}$ is a Gaussian 
-with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
-sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
-broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
-\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
-incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
-parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
-a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
+spectral or flux dimensions.  For the former, we consider three kernels that contribute to the 
+line-of-sight velocity distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, 
+treats the instrumental spectral broadening.\footnote{For computational efficiency, we first 
+pre-convolve the entire raw spectral library with a slight under-prediction of $\mathcal{F}_v^{\rm 
+inst}$ and resample the spectra to a lower resolution.  Because the instrumental point spread 
+function acts as a low-pass filter, this procedure reduces the size of the library by a factor of 
+10 or more while still preserving the final Fourier content of the spectra.  Then, when any given 
+spectrum is post-processed, the actual $\mathcal{F}_v^{\rm inst}$ used is treated as a smaller 
+``correction" to this pre-processing kernel.}  For illustrative purposes, we assume 
+$\mathcal{F}_v^{\rm inst}$ is a Gaussian with mean $v = 0$ and constant width $\sigma_v$ at all 
+$\lambda$, although more sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm 
+rot}$, characterizes the broadening induced by (projected) stellar rotation, parameterized by 
+$v\sin{i}$ as described by \citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm 
+dop} = \delta(v-v_z)$, incorporates the radial velocity through a Doppler shift.  The model 
+spectrum is modified by the parameters $[\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, 
+using a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
 multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
-the mathematical formalism is presented here for clarity.}
-\begin{equation}
-\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) = \flam(\vt_{\ast}) \otimes \varphi_v = \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
-\label{eqn:broadening}
-\end{equation}
+the mathematical formalism is presented for clarity.}
+\begin{eqnarray} \label{eqn:broadening}
+\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \otimes \varphi_v \\
+                                           &=& \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop}, \nonumber
+\end{eqnarray} 
 and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
-\begin{equation}
-\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
-\label{eqn:resampling}
+\begin{equation} \label{eqn:resampling}
+\flam(\vt_{\ast},  \sigma_v, v\sin{i}, v_z) \mapsto \vM(\vt_{\ast},  \sigma_v, v\sin{i}, v_z),
 \end{equation}
@@
 $N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
-spectrum).  For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
-post-processing steps.
+spectrum).  Figure \ref{fig:broadening} shows a (condensed) graphical representation 
+of these post-processing steps.
 
@@
   \includegraphics{figs/high2low.pdf}
-  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite.
-Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000.
-Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$.
-Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
-\label{fig:broadening}
+  \figcaption{({\it top}) The line-of-sight velocity distribution function, $\varphi_v$, and its 
+decomposition into broadening kernels.  The instrumental kernel ({\it dotted}) is treated as 
+a Gaussian, the rotation kernel ({\it dashed}) is a Gaussian-like function of the projected 
+rotational velocity, and the Doppler kernel ({\it solid}) is a $\delta$-function that introduces 
+the radial velocity.  In this specific case, $\sigma_v = 2.9$\,km s$^{-1}$, $v\sin{i} = 5$\,km 
+s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{sec:examples}.x.
+({\it bottom}) A segment of a raw, full-resolution model spectrum and its post-processed equivalent 
+after convolution by $\varphi_v$ and re-sampling at the coarser resolution of the detector pixels. 
+\label{fig:broadening}}
 \end{center}
@@
 account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
-\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $\vt_{{\rm obs}, f} = 
-[\Omega, \,\, A_V]$ are applied as
-\begin{equation}
-\vM(\vT) \equiv \vM(\vt_{\ast}, \vt_{{\rm obs},v}, \vt_{{\rm obs},f}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
-\label{eqn:scaling}
-\end{equation}
-where we have simplified the notation by writing $\vT \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.
-
-So far, the procedure summarized in Eq.~\ref{eqn:interp}-\ref{eqn:scaling} is composed of 
-relatively straightforward operations demanded by practical astronomical and computing issues.  If 
-the observed spectrum was {\it perfectly} calibrated, we could proceed to a likelihood calculation 
-(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vT)$ at this point.  
-However, that is unlikely to be the case.  The primary concern is that an imperfect calibration 
-produces (presumably low-level) mismatches in the underlying shape of the observed spectrum on 
-relatively broad wavelength scales.  When compared with the model, these mismatches might represent 
-a non-trivial contribution to the likelihood, and thereby bias our estimates of the desired 
-physical parameters.  Often, this concern is treated externally to any modeling procedure, usually 
-by dividing the observed spectrum (and model spectrum) by an appropriate polynomial function.  But 
-that ``normalization" procedure implicitly assumes that there is no relevant information content on 
-those mismatched scales; if $\vT$ also contributes to the broad spectral shape, then adopting this 
-approach will corrupt the inferences of these parameters.  Moreover, in practice this approach is 
-limited, since defining an appropriate polynomial becomes difficult in cases where the spectral 
-line density is high (e.g., molecular bands for cool stars).
-
-We adopt a somewhat analogous approach to deal with this issue, but cast it {\it internal} to the 
-modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
-freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
-the model spectrum, enabling us to explore the distribution of possible calibrations (see 
-Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
-In essence, the inferences for the relevant stellar parameters will properly account for the 
-underlying uncertainty in the calibration process.  In practice, this is achieved by distorting 
-segments of the model spectrum with polynomials, $\mathsf{P}$ \citep[e.g.,][]{eisenstein06,
-koleva09}.  For data with $N_{\rm ord}$ spectral orders, each denoted with index $o$ (where $o$ 
-naturally has an implicit $\lambda$-dependence), we can write
-\begin{equation}
-\vM(\vT, \cheb) = \vM(\vT) \times \{\mathsf{P}_o\} \equiv \vM(\vT) \times \left\{ \sum_n c_o^{(n)} \, T_o^{(n)} \right\},
-\label{eqn:chebyshev}
-\end{equation}
-where $T^{(n)}$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated 
-as a set of $n \times N_{\rm ord}$ nuisance (hyper-)parameters, $\cheb = [\{\vp_{\mathsf{P}_o}\}] = 
-[\{c_o^{(0)}, c_o^{(1)}, \ldots, c_o^{(n-1)} \}]$.  With judicious priors on $\cheb$, we can ensure 
-that the unintended treatment of real spectral features (e.g., broad, deep molecular bands) as 
-calibration artifacts is negligible (see Section~\ref{sec:examples} for examples).  The 
-lowest-degree (scaling) coefficient, $c^{(0)}$, is by its nature degenerate with the solid angle 
-parameter, $\Omega$.  Therefore, we enforce an additional constraint by requiring that the mean of 
-the polynomial is unity.  For data with a single spectral order, this means simply setting $c^{(0)} 
-= 1$.  If the goal is to model multiple spectral orders, we assign $c^{(0)} = 1$ in an arbitrary 
-order as an anchor, but permit the $c^{(0)}$ values in other orders to be different (as 
-necessary).  It is worth noting that this formalism can, in principle, be extended to develop 
-models for completely uncalibrated spectra, rather than the residual calibration mismatches as 
-described here (with a suitable relaxation of the priors on $\cheb$).  Figure \ref{fig:chebyshev} 
-offers a practical demonstration of how these nuisance parameters are applied.
-
-\begin{figure}[!htb]
+\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $[\Omega, \,\, A_V]$ 
+are then applied as
+\begin{eqnarray} \label{eqn:scaling}
+\vM(\vT) &=& \vM(\vt_{\ast}, \vt_{\rm obs}) \\
+         &=& \vM(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) \times \Omega \times 10^{-0.4\,A_{\lambda}}, \nonumber
+\end{eqnarray}
+with simplified notation such that $\vT \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$, 
+where $\vt_{\rm obs} = [\sigma_v, v\sin{i}, v_z, \Omega, A_V]$.
+
+The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
+straightforward operations demanded by practical astronomical and computing issues.  If the data 
+were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
+comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
+imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
+Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
+they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
+the physical parameters.  This concern is treated externally to any modeling procedure, most often 
+by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
+``normalization" implicitly assumes that there is no relevant information content in the spectral 
+shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
+the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
+an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
+molecular bands for cool stars).
+
+We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
+the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
+in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as 
+explicit contributions to the model spectrum, enabling an exploration of the distribution of 
+possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance 
+parameters and then marginalized.  This way, the inferences for the stellar parameters will 
+properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
+practice, this is achieved by distorting segments of the model spectrum with polynomials, 
+$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
+orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
+\begin{eqnarray} \label{eqn:chebyshev}
+\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
+                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
+\end{eqnarray}
+and $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
+coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
+\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
+real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
+The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
+$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
+unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
+multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
+$c^{(0)}$ in other orders to be different.  It is worth noting that this formalism can, in 
+principle, be extended to develop models for completely uncalibrated spectra, rather than the 
+residual calibration mismatches described here (with a suitable relaxation of the priors on 
+$\cheb$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how these nuisance 
+parameters are applied.
+
+\begin{figure}[!b]
 \begin{center}
   \includegraphics{figs/chebyshev.pdf}
-  \caption{The spectrum at the overlap of two echelle orders (23 \& 24).
+  \figcaption{
+The spectrum at the overlap of two echelle orders (23 \& 24).
 \textbf{Panel 1}: The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders.
@@
 \textbf{Panels 3 \& 4}: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4).
-This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.}
-\label{fig:chebyshev}
+This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
+\label{fig:chebyshev}}
 \end{center}
@@
 likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding residual 
-spectrum (vector of $N_{\rm pix}$ elements) can be defined for any input parameter set,
+spectrum (an $N_{\rm pix}$-element vector) can be defined for any input parameter set,
 \begin{equation}
@@
 To quantify the probability of the data conditioned on the model, we adopt a standard 
-multi-dimensional Gaussian likelihood function,
+multi-dimensional Gaussian likelihood function
 \begin{equation}
 p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2} \,
-   \vR^\trans \vC^{-1} \vR \right ), 
+   \vR^\trans \vC^{-1} \vR \right )
    \label{eqn:likelihood}
 \end{equation}
-which penalizes models that yield larger residuals and explicitly allows for covariances in the 
-residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
-numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
-this case is
+that penalizes models which yield larger residuals and explicitly allows for covariances in the 
+residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix $\vC$.  For practical 
+reasons, the log-likelihood is used as the quality metric, where
 \begin{equation}
@@
 \end{equation}
-the sum of the square of the residuals weighted by the inverse variances (squared uncertainties).  
-However, the problem being addressed here necessitates the use of a more complex covariance matrix; 
-additional off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances 
-imposed by the discrete over-sampling of the line-spread function, and (2) highly correlated 
-residuals as manifestations of the still-imperfect model library are required to avoid biasing our 
-inferences of the physically interesting parameters ($\vT$).  The following subsections describe 
-how these issues are addressed in the practical implementation of $\vC$.
+the sum of the square of the residuals weighted by their inverse variances.  However, the problem 
+being addressed here necessitates the use of a more complex covariance matrix; additional 
+off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances imposed by the 
+discrete over-sampling of the line-spread function, and (2) highly correlated residuals as 
+manifestations of the still-imperfect model library are required to avoid biasing our inferences of 
+the physically interesting parameters ($\vT$).  The following subsections describe how these issues 
+are addressed in the practical implementation of $\vC$.
 
@@
 
-Astronomical spectrographs are designed so that the detector over-samples the instrumental line 
-broadening function with at least a few pixels.  Therefore, adjacent pixels are never completely 
-independent samples of the observed spectrum.  In that case, a difference between an observed and 
-modeled spectral feature will create a residual that spans multiple pixels.  This can be 
-demonstrated in practice by examining the autocorrelation of the residual spectrum: a slight model 
-mismatch will produce correlated residuals over a characteristic scale similar to the observational 
-line-broadening kernel width.  Figure \ref{fig:class0} highlights a specific example of these 
-correlated residuals in real data, where an imperfect model generates residuals which exhibit a 
-significant autocorrelation signal on the scale of $\sim$4 pixels, which corresponds to the typical 
-line width in this spectrum.
+Astronomical spectrographs are designed such that the detector over-samples the instrumental 
+line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
+completely independent samples of the true spectrum.  In that case, a difference between an 
+observed and modeled spectral feature will create a correlated residual that spans multiple 
+pixels.  This can be demonstrated practically through the autocorrelation of $\vR$: a slight model 
+mismatch will produce correlated residuals over a characteristic scale similar to the instrumental
+broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
+these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$4 
+pixel scale, corresponding to the 3\,km s$^{-1}$ width of $\mathcal{F}_v^{\rm inst}$.
 
@@
 the spectrograph by astrophysical or instrumental effects is generally uncorrelated with 
-wavelength.  The arrival of each photon to the detector is an independent event; while these 
-photons are scattered by the instrumental line-spread function, the magnitude and direction of that 
-scatter is independent for each such event.  In essence, the noise itself is uncorrelated, but the 
-fit residuals likely are correlated.  However, from a mathematical perspective the correlated 
+wavelength.  The arrival and propagation of each photon through the instrument and into the 
+detector can be considered an independent event.  In essence, the noise itself is not correlated, 
+but the fit residuals likely are.  However, from a mathematical perspective the correlated 
 residuals can be treated in the same way as correlated noise, by constructing a non-trivial 
 covariance matrix with off-diagonal terms.  In practice, this is achieved by parameterizing $\vC$ 
-with a kernel that describes the covariance between any pair of pixels, representing wavelengths 
-$\lambda_i$ and $\lambda_j$ (in many ways analogous to the standard two-point spatial correlation 
-function used in cosmology).
+with a kernel that describes the covariance between any pair of pixels, indexed $ij$, representing 
+wavelengths $\lambda_i$ and $\lambda_j$.
 
@@
 \end{equation}
-where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
-between pixel residuals, such that
+where $c$ is the speed of light.  This parametric kernel describes the covariance between pixel 
+residuals, 
 \begin{equation}
-  \Kglobal(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
+  \Kglobal_{ij} =  \langle \vR_i \; \vR_j \rangle.
   \label{eqn:expectation}
@@
 A variety of kernels have been used in the field of Gaussian processes to parameterize such a 
-mildly covariant structure \citep[e.g.,][]{rasmussen05}.  We adopt the \matern\ kernel with $\nu = 
-3/2$,
+covariant structure \citep[e.g.,][]{rasmussen05}.  Here, we adopt the commonly-employed \matern\ 
+kernel with $\nu = 3/2$,
 \begin{equation}
-  \Kglobal(\lambda_i, \lambda_j \, |\, \vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+  \Kglobal_{ij}(\vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
@@
 \end{equation}
-with the global covariance (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ are 
-characterized by an amplitude ($a_g$) and scale ($\ell$) designed to smoothly transition to 
-negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
+where the (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ include an amplitude ($a_g$) and 
+scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
 enables computational expendiency, we employ a Hann window function
 \begin{equation}
-  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
+  w_{ij}^{\rm g} \,(r_0) = \left \{ 
     \begin{array}{cc}
@@
 \end{equation}
-to taper the kernel (so the effective kernel is $w \, \Kglobal$).  The truncation distance 
-$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
-The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
-structure in the residual spectrum that is similar to those seen in a typical data--model 
-comparison.
+to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
+scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matern} demonstrates that random draws from such a 
+kernel readily produce correlated structure similar to those seen in a typical residual spectrum.
 
@@
 
-Aside from the global covariance structure described above, there are likely also local regions of 
-strong, highly correlated residuals that need to be treated in the modeling framework.  These large 
-amplitude residual regions are usually produced by imperfect spectral lines in the models (e.g., 
-missing opacity sources, uncertain oscillator strengths, etc.); some representative examples are 
-highlighted in Figure \ref{fig:badlines}.  To parameterize such regions in the covariance matrix, 
-we introduce a sequence of non-stationary kernels that explicitly depend on the actual wavelength 
-values of a pair of pixels (on $\lambda_i$ and $\lambda_j$), and not simply the distance between 
-them ($r_{ij}$).
+Aside from the global covariance structure, there can also be local regions of strong (highly 
+correlated) residuals that need to be treated in the modeling framework.  These patches of large 
+$\vR$ are usually produced by imperfect spectral lines in the models (missing opacity sources, 
+uncertain oscillator strengths, etc.); some representative examples are highlighted in Figure 
+\ref{fig:badlines}.  To parameterize such regions in $\vC$, we introduce a sequence of 
+non-stationary kernels that explicitly depend on the actual wavelength values of a pair of pixels 
+(on $\lambda_i$ and $\lambda_j$), and not simply the distance between them ($r_{ij}$).
 
@@
 spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
-model.  In that case, the $k^{\rm th}$ such local residual can be described as
+model.  In that case, the $k^{\rm th}$ such local residual could be described as
 \begin{equation}
-\vR_{\lambda}(a_k, \mu_k, \sigma_k) = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
-\end{equation}
-with an amplitude $a_k$, a mean wavelength $\mu_k$, and a width $\sigma_k$.  Following 
-Eq.~\ref{eqn:expectation}, the kernel that describes the covariance of any two pixels related to a 
-sequence of residual regions is the linear combination
-\begin{equation}
-  \Klocal(\lambda_i, \lambda_j | \vp_{{\mathsf C}, l}) = \sum_k
-  \frac{1}{2\pi} \left(\frac{a_k}{\sigma_k}\right)^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
-\label{eqn:klocal}
-\end{equation}
-with the set of local covariance (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, 
-\sigma_k\}]$.
-We again taper this kernel with a Hann window (Eq.~\ref{eqn:Hann}) to ensure computational 
-efficiency with a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set 
-to some multiple of the Gaussian width (we set $r_0 = 4\sigma_k$).  Figure \ref{fig:region} 
-demonstrates how this non-stationary Gaussian kernel generates a localized region of enhanced 
-variance that successfully mimics the kind of residuals produced by an inaccurate spectral line 
-model.  These kernels down-weight the influence of such strong residuals in the likelihood 
-calculation, mitigating any potential bias they might induce on inferences of the interesting 
-parameters ($\vT$).  In essence, this is a robust, flexible, and unbiased method for (correlated) 
-outlier rejection that preserves the integrity of the probabilistic framework being developed (as 
-opposed to manual or threshold-based clipping or masking).
+\vR = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
+\end{equation}
+with amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
+Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
+residual regions can be written
+\begin{equation} \label{eqn:kregion}
+  \mathcal{K}^{l,k}_{ij} = \frac{a_k^2}{2\pi\sigma_k^2} \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
+\end{equation}
+such that the full local covariance kernel composed of a sequence of residual regions is the linear 
+combination,
+\begin{equation} \label{eqn:klocal}
+\Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{l,k}_{ij},
+\end{equation}
+with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
+again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure that a sparse covariance 
+matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
+width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:region} shows how this non-stationary Gaussian 
+kernel generates a localized region of enhanced variance that successfully mimics the kind of 
+residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
+influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
+they might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
+flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
+the probabilistic framework (as opposed to manual or threshold-based clipping or masking).
 
@@
 spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well.
+in this framework if we design an appropriate kernel.
 
@@
 (Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
-and these kernels, 
-\begin{eqnarray}
-\vC_{ij}(\cov)  &=& b \, \delta_{ij}\sigma_i + \\ \nonumber
-                &=& w(r_{ij}|r_0 = 4\ell) \,\, \Kglobal(\lambda_i, \lambda_j|a_g, \ell) + \\ \nonumber
-                &=& \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
-\end{eqnarray}
-\footnote{\comm{this needs adjustment for notation, pending what we do in the local covariance kernels for M dwarfs.}}
-where the covariance (hyper-)parameters are $\cov = [a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}]$, 
-and $b$ is assumed to be a fixed parameter that scales up the pixel noise values to account for 
-additional detector or data reduction uncertainties (e.g., read noise, uncertainties in the 
-spectral extraction procedure, and interpolation errors); reasonable values are $b \approx 
-1.02$--1.10 for well-calibrated optical spectra (see Section \ref{sec:examples} for examples).  If 
-there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
-determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance (hyper-)parameters, 
-$\cov$.
-
-\begin{figure*}[!t]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
+and these global and local kernels, 
+\begin{equation}
+\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \,\, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
+                  \Klocal_{ij}(\vp_{{\mathsf C}, l}), 
+\end{equation}
+with (hyper-)parameters $\cov = [\vp_{{\mathsf C}, g}, \vp_{{\mathsf C}, l}]$.  The factor $b$ is 
+a (constant) parameter that scales up the pixel noise values to account for additional detector or 
+data reduction uncertainties (e.g., read noise, uncertainties in the spectral extraction procedure, 
+and interpolation errors); reasonable values are $b \approx 1.02$--1.10 for well-calibrated optical 
+spectra (see Section \ref{sec:examples} for examples).  If there are $N_{\rm loc}$ local covariance 
+patches (see Section \ref{subsec:MCMC} on how this is determined), then there are $4N_{\rm loc}+2$ 
+elements in the set of covariance (hyper-)parameters, $\cov$.
+
+%\begin{figure*}[!t]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
 The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
 These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
-\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
 The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
 The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
-\label{fig:region}
-\end{center}
-\end{figure*}
+%\label{fig:region}
+%\end{center}
+%\end{figure*}
 
 \subsection{Priors} \label{subsec:priors}
-The final posterior probability distribution of the model parameters is defined by Bayes rule as proportional to the product of the likelihood function (Eqn~\ref{eqn:likelihood}) and the prior probabilities on the model parameters.
-Because a high quality spectrum provides so much information about $\vT$, in general the inference of the stellar parameters is not very sensitive to specific choices of priors.
-We generally assume uniform priors on the stellar parameters, which are flat for any $\vt_{\ast}$ within the range of $\{\vt_{\ast}\}^{\rm grid}$ and otherwise practically $0$.
-For stars with a well-defined continuum, it is generally acceptable to use flat priors on the polynomial parameters $\cheb$.
-However, if there is information about the systematic accuracy of flux calibration, for example from multiple observations of spectrophotometric standards, it is possible to encode this information with a Gaussian prior on the coefficients of $\cheb$, with a mean of $0$ and a width of the typical systematic variation of the flux calibration (e.g., $5\%$).
-For stars with a poorly defined continuum or pseudo-continuum, judicious choice of these priors will ensure that broad spectroscopic features are not destroyed by the polynomial fit.
-For the parameters of local covariance kernels, $\vp_{{\mathsf C},l}$, we assume a prior on the width of the local kernel that is flat below the scale of the instrumental PSF and rapidly Gaussian tapers to zero for widths larger than the PSF.
-This prior is designed to prevent the local kernels from diffusing to large widths and low amplitudes, effectively duplicating the role of the global kernels.
+
+The Bayesian framework of this inference approach permits us to specify prior knowledge about the 
+model parameters, $p(\vM)$.  Since a high quality spectrum provides so much information about 
+$\vT$, the inference of these parameters ends up not being very sensitive to the priors.  To be 
+conservative, we generally recommend assigning uniform priors on $\vT$, such that $p(\vt_{\ast})$
+is flat over the spectral library grid (and zero elsewhere) and $p(\vt_{\rm obs})$ is flat if 
+$\vt_{\rm obs} \ge 0$ (i.e., for physically meaningful values).
+
+For (early type) stars with a clear continuum, it makes sense to assume flat priors on the 
+polynomial parameters $\cheb$.  However, if information about the accuracy of the calibration is 
+available (e.g., from observations of multiple spectrophotometric standards), it is possible to 
+encode this information in a simple prior on the Chebyshev coefficients.  A reasonable example is a 
+set of Gaussian priors with widths that encapsulate the systematic (fractional) variance of the 
+derived calibration functions.  For (late type) stars with a poorly defined pseudo-continuum, 
+some judicious tapering of the priors (such that high-$n$ coefficients are constrained around zero) 
+may be required to ensure that broad spectroscopic features are not absorbed into the polynomial 
+(see Section~\ref{sec:examples}).
+
+We assume uniform priors on the global covariance parameters.  For the local covariance kernels, we 
+adopt flat priors for the amplitudes and means, \{$a_k$, $\mu_k$\}, and construct a specialized 
+prior for the widths, \{$\sigma_k$\}, that is flat when $\sigma_k < \sigma_v$ and has a rapid 
+(Gaussian) taper to zero for larger widths.  This form is designed to prevent the local 
+kernels from diffusing to large $\sigma_k$ and low $a_k$, since such behavior is better treated by 
+the global kernel.
 
@@
 
-The inference problem as framed here has a natural hierarchical structure between the collections 
+The inference problem developed here has a natural hierarchical structure between the collections 
 of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
-$\vP = [\cheb, \cov]$.  To explore the posterior probability distribution of the model conditioned 
-on an observed spectrum in this case, it is convenient to use MCMC simulations with a blocked Gibbs 
-sampler coupled to the Metropolis-Hastings algorithm.  In essence, this procedure works by sampling 
-in a subset of parameters with Metropolis-Hastings conditioned on the current (fixed) values of the 
-other parameters; after each iteration, the Gibbs sampler then updates the values of that parameter 
-subset and then cycles through all the different parameter subsets in the same way (for a more 
-nuanced mathematical description of the process, see Chapter 11 of \citealt{gelman13}).  The 
-step-by-step prescription is as follows, where we denote the $i^{\rm th}$ iteration of the Gibbs 
-sampler as a superscript index (not to be confused with the $i^{\rm th}$ pixel as in our previous 
-notation):
-\begin{enumerate}
-\item Initialize the parameters.  One might set $\vT^0$ based on estimates in the literature or 
-scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the Chebyshev 
-polynomial coefficients ($\cheb^0$) to be flat ($c_0 = 1$,  $c_{>0} = 0$ in all spectral orders) 
-and assume only the trivial noise spectrum contributes to the covariance matrix (all elements of 
-$\cov^0$ are zero, $\vC$ is diagonal).
-
-\item Start a new ($i^{\rm th}$; where $i \in [1,N_G]$) iteration of the Gibbs sampler.  For 
-one iteration of the Metropolis-Hastings algorithm, sample in $\vT$ to evaluate the posterior 
-probability following the framework laid out in Sections \ref{subsec:likelihood} and \ref{subsec:priors}, 
-representing a slice through the posterior space conditioned on the hyperparameters being fixed to 
-their $(i-1)^{\rm th}$ iteration values.  Now update the parameters $\vT^{i-1} \rightarrow \vT^i$.
-
-\item Cycle through and update the hyperparameters, $\vP^{i-1} \rightarrow \vP^i$:
-
-\begin{enumerate}
-\item For each spectral order, sample in the polynomial hyperparameters $\cheb$ using one 
-iteration of the Metropolis-Hastings algorithm, computing posterior probabilities along a subset of the posterior space where $\vT^i$ and $\cov^{i-1}$ are fixed.  Now update the 
-polynomial hyperparameters $\cheb^{i-1} \rightarrow \cheb^i$.
-
-\item Sample in the global (stationary) covariance kernel hyperparameters, $\vp_{{\mathsf C},g} = 
-[a_g, \ell]$, using one iteration of the Metropolis-Hastings algorithm to compute posteriors
-with $\vT^i$ and the local covariance kernel parameters, $\vp_{{\mathsf C}, k}^{i-1}$, fixed.  
-Update the global covariance hyperparameters $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf 
+$\vP = [\cheb, \cov]$.  To explore the posterior distribution of the model conditioned on the data
+\begin{equation} \label{eqn:post}
+p(\vT, \vP | \vD) \propto p(\vD | \vT, \vP) \, p(\vT, \vP)
+\end{equation}
+for this type of structure, it is convenient to use MCMC simulations with a blocked Gibbs 
+sampler coupled to the Metropolis-Hastings algorithm.  To summarize, this procedure works by 
+sampling in a subset of parameters (with Metropolis-Hastings) conditioned on the current (fixed) 
+values of the other parameters; after each iteration, the Gibbs sampler then updates the values of 
+the sampled parameter subset and then cycles through all the (previously fixed) different parameter 
+subsets in the same way (for a more nuanced mathematical description of the process, see Chapter 11 
+of \citealt{gelman13}).  A step-by-step prescription follows, where the $i^{\rm th}$ 
+iteration of the Gibbs sampler is indexed with a superscript: \\
+
+\noindent (1) Initialize the parameters.  One might set $\vT^0$ based on estimates in the 
+literature or scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the 
+Chebyshev coefficients ($\cheb^0$) so that the polynomials are constant ($c_0 = 1$,  $c_{>0} = 0$ 
+in all spectral orders) and assume only the trivial noise spectrum contributes to the covariance 
+matrix ($\cov^0 = 0$, so $\vC$ is diagonal).  \\
+
+\noindent (2) Start the $i^{\rm th}$ (where $i \in [1,N_G]$) iteration of the Gibbs sampler.  For 
+each iteration of the Metropolis-Hastings algorithm, sample in the interesting parameters $\vT$ to 
+evaluate the posterior (Eq.~\ref{eqn:post}) following the framework laid out in Sections 
+\ref{subsec:likelihood} and \ref{subsec:priors}.  This represents a ``slice" through the posterior 
+space conditioned on the nuisance parameters being held fixed ($\vP = \vP^{i-1}$).  Then update the 
+interesting parameters $\vT^{i-1} \rightarrow \vT^i$. \\
+
+\noindent (3) Cycle through and update the nuisance parameters, $\vP^{i-1} \rightarrow \vP^i$.  For 
+each spectral order,
+
+\begin{enumerate}[(a)]
+\item Sample in the polynomial coefficients $\cheb$ along a subset of the posterior space where 
+$\vT^i$ and $\cov^{i-1}$ are fixed.  Update the coefficients, $\cheb^{i-1} \rightarrow \cheb^i$.
+
+\item Sample in the global (stationary) covariance kernel parameters, $\vp_{{\mathsf C},g} = 
+[a_g, \ell]$, with $\vT^i$, $\cheb^i$, and the local covariance kernel parameters, 
+$\vp_{{\mathsf C}, l}^{i-1}$, fixed.  Update $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf 
 C}, g}^i$.
 
-\item Decide whether to instantiate any additional local (non-stationary) covariance kernels.  Here 
-we adopt an iterative thresholding criterion: if a local patch of the residual signal is at least 
-$3\times$ the amplitude of the global covariance kernel (i.e., where $\vR \ge 3
-\, a_g^i$), this will trigger the addition of a local kernel\footnote{For
-practical reasons, we do not delete local kernels from the parameter list once they have been
-instantiated. However, if the stellar parameters have moved to a region of
-parameter space where a given local kernel is no longer required, then the
-amplitude of that kernel will be driven towards $0$, the kernel would have a
-negligible contribution to the covariance matrix $\vC$, and the overall behavior of the model would
-be as if the kernel were deleted.}.
-
-\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},k}$ using one 
-iteration of the Metropolis-Hastings algorithm to compute the posterior probabilities
-with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local covariance 
-hyperparameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
+\item Decide whether to instantiate any local (non-stationary) covariance kernels.  We adopt an 
+iterative thresholding criterion: if a local patch of the residual signal is at least $3\times$ the 
+amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this triggers the 
+addition of a local kernel.\footnote{There is no practical reason to delete local kernels.  If the 
+parameters have changed such that a given local kernel is no longer required, that kernel amplitude 
+will be driven towards zero and represent a negligible contribution to $\vC$; in effect, the model 
+will act as if the kernel were deleted automatically.}
+
+\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
+$\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
+$\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
 \end{enumerate}
 
-\item Return to Step 2 and iterate for $N_G$ iterations.
-
-\item Repeat this procedure with different initializations (Step 1) and compute convergence diagnostics of the multiple chains, to ensure that all of the chains have converged to the same global 
-maximum in the posterior probability distribution and that each chain has converged.
-\end{enumerate}
-How many iterations to run the Gibbs sampler, $N_G$, is empirically determined through visual inspection of the Markov chains and with the Gelman-Rubin convergence diagnostic $\hat{R}$ \citep[Eqn 11.4]{gelman13}, which compares intra-chain and inter-chain variance to assess convergence.
-After a running a series of multiple chains, we compute the diagnostic, which summarizes how the inference of the posterior probability function would be improved if we were to run the Markov chain for an infinite number of iterations.
-If $\hat{R}$$>$1.1, then there is evidence that the chains have not fully converged and so we increase $N_G$ and repeat the simulations until we are confident of convergence.
-
-Gibbs sampling from the posterior with this procedure can be a computational
-challenge.  A typical spectrum might have $N_{\rm pix}$$>$1000, and therefore
-the many repeated evaluations of the matrix product $\vR^{\trans} \vC^{-1} \vR$
-in the likelihood function calculation can be numerically expensive;
-prohibitively so in the brute force sense, since $\vC$ is non-trivial by
-design.  However, there are two aspects of these calculations that enable us to
-take advantage of clever, efficient algorithms designed for specifically these
-kinds of issues.  First, since we are interested in the matrix product
-$\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can directly solve for the product using efficient algorithms and avoid the computational expense of
-matrix inversion. And second, because
-$\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky factorization
-to optimize the evaluation of the matrix product.  The makes Step 2 and Steps
-3a and 3b in the procedure outlined above extremely rapid.  Although this
-factorization must be re-done for each update of the covariance
-hyperparameters, such operations are still modestly efficient because the
-covariance kernels were specifically designed to deliver a sparse matrix (via
+\noindent (4) Return to Step~(2) and iterate to convergence. \\
+
+\noindent (5) Repeat the procedure in Steps~(1)--(4) with different initializations and compute 
+convergence diagnostics to ensure that all of the chains have reached the same global maximum in 
+the posterior probability distribution. \\
+
+After burn-in, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
+diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
+inter-chain variances to assess how the inference of the posterior would be improved if the chains 
+were to be run for an infinite number of iterations.  The Gibbs sampler is eventually stopped if 
+$\hat{R} \, < \, 1.1$, and the chains appear converged in a visual inspection.
+
+Gibbs sampling from the posterior with this procedure can be a computational challenge.  A typical 
+spectrum might have $N_{\rm pix}$$>$1000, and therefore the many repeated evaluations of the matrix 
+product $\vR^{\trans} \vC^{-1} \vR$ in the likelihood function calculation can be numerically 
+expensive; prohibitively so in the brute force sense, since $\vC$ is non-trivial by design.  
+However, there are two aspects of these calculations that enable us to take advantage of some 
+clever algorithms designed for these kinds of issues.  First, since we are only interested in the 
+product $\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can employ some efficient sparse 
+matrix approximations that avoid the computational expense of matrix inversion.  And second, 
+because $\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky factorization
+to optimize the evaluation of the matrix product.  In practice, this means Steps~(2), (3a), and 
+(3b) in the procedure outlined above can be conducted with minimal computation cost.  Although the
+Cholesky factorization must be re-done for each update of $\vp_{\mathsf C}$ (Step~3d), this does 
+not incur a high cost because the covariance kernels were designed to deliver a sparse matrix (via 
 the Hann windows).  We use the high-performance 
 \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} 
-library to implement the sparse matrix and Cholesky factorization operations
-\citep{chen08,davis09}, and extend the Metropolis-Hastings sampler included in
-the {\tt emcee} package \citep{foreman-mackey13} to function within a Gibbs sampler\footnote{\url{https://github.com/iancze/emcee}}.
-
-\section{Examples}
-\label{sec:examples}
-
-\subsection{Datasets and historical analysis}
-% Paragraphs as follows:
-% Introduction to the two types of stars, in order to demonstrate the flexibility.
-To demonstrate the flexibility of our probabilistic framework, we model spectra of two different stars acquired with different instruments.
-The first data set is a high resolution optical spectrum of WASP-14, an F5 star \citep{joshi09}.
-The second data set is a moderate resolution mid-infrared spectrum of Gl~51, an M5 star \citep{rojas-ayala12}.
-
-% Data description of WASP-14
-Our spectrum of WASP-14 was obtained with the Tillinghast Reflector Echelle Spectrograph (\emph{TRES}, \citealt{mink11}) on the Fred Laurence Whipple Observatory 1.5m telescope on June 14th, 2009.
-\emph{TRES} delivers a 51 echelle spectrum spanning the full optical range (3860--9100\AA) at a resolution of $R$$\sim$44,000.
-\citet{torres12} originally acquired the spectrum of WASP-14 to study the stellar properties of transiting exoplanet hosts.
-The raw spectrum was reduced using the well-tested \emph{TRES} pipeline written by L.~Buchhave and in the \ion{Mg}{0}b echelle order ($\sim$5100\AA), the spectrum has a signal to noise of 150 per resolution element.
-
-% What other people get when they fit this star. How does our inference compare?
-\citet{torres12} determined the stellar properties of WASP-14 using several different methods including {\tt MOOG}, {\tt SME}, and {\tt SPC}.
-While the {\tt MOOG} and {\tt SME} methods are not directly comparable to our technique because they measure individual line strengths, {\tt SPC} utilizes a moderately-sized region of spectrum and a pre-computed synthetic library that enables a direct comparison with our technique.
-When we make the same assumptions in \citet{torres12}, such as using the same region of spectrum (5050--5360\AA), the same spectral library, and fixing $\log g$ to the value inferred from the exoplanet transit analysis by \citet{joshi09} (amounting to a $\delta$-function prior on $\log g$$=$4.29), our inference of stellar parameters agrees to within 1$\sigma$ (see Table~\ref{table:tests}, WKX).
-Interestingly, when we allow $\log g$ to float, we infer a lower $T_{\rm eff}$ and $\Z$ (WKX2).
-These results demonstrate the sensitivity to assumptions about $\log g$ as well as the fact that there may exist biases in surface gravity estimates from either spectroscopy, exoplanetary transits, or both.
-
-% Data description of Gl 51
-The spectrum of Gl~51, an M5 star, was downloaded from \emph{IRTF} spectral library \citep{cushing05, rayner09}.
-The data was obtained with the \emph{SPEX} instrument \citep{rayner03} on the NASA \emph{IRTF}.
-The spectrum is a moderate resolution (R$\sim$2,000), mid-infrared spectrum from 0.7--5.5$\mu$m at signal to noise $>$400 at 2.1$\mu$m.
-
-% What other people get when they fit this star. How does our inference compare?
-To determine the stellar parameters of Gl~51, \citep{rojas-ayala12} use the equivalent widths of the 2.205$\mu$m \ion{Na}{1} doublet and 2.263$\mu$m \ion{Ca}{1} triplet along with their \HK\ index, which is defined using three passbands: 2.070--2.090$\mu$m, 2.235--2.255$\mu$m, and 2.360--2.380$\mu$m (see their Figure 2 for a useful visualization).
-This index is designed to probe the curvature of the spectrum pseudo-continuum, which varies to due absorption from water molecules.
-In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
-They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
-\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
-When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{The results of our analysis will be in Table~\ref{table:tests}.}
+library to implement the sparse matrix and Cholesky factorization operations \citep{chen08,
+davis09}, and extend the Metropolis-Hastings sampler included in the {\tt emcee} package 
+\citep{foreman-mackey13} to function within a blocked Gibbs 
+sampler.\footnote{\url{https://github.com/iancze/emcee}} \\
+
+\section{Demonstrations} \label{sec:examples}
+
+In this section, we aim to demonstrate the functionality and highlight the flexibility and utility 
+of the probabilistic inference methodology described above.  To do that, we illustrate how the 
+modeling framework operates for two representative examples.  The first is an elaboration of the 
+example shown throughout Section \ref{sec:method}, using a high resolution optical 
+($\sim$5000-5400\,\AA) spectrum of the F5 dwarf main sequence star (and transiting exoplanet host) 
+WASP-14 \citep{joshi09,torres12}.  This spectrum is interpreted in the context of two different 
+synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
+empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
+$\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
+and the most recent incarnation of the {\sc Phoenix} library \citep{husser13}.  The second 
+example uses a medium resolution near-infrared ($\sim$2.0-2.4\,$\mu$m) spectrum of the M5 dwarf 
+star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
+\citep{cushing05,rayner09}.  In this case we adopt the standard {\sc BT-Settl} modification of the 
+{\sc Phoenix} library \citep{allard12}.
 
@@
 
-We use two synthetic libraries to fit WASP-14.
-The first is the recent publicly available PHOENIX grid (PHOENIX version 16, \citealt{husser13}), which spans wavelengths from 500\AA\ to 5.5$\mu$m and $T_\textrm{eff}$ from 2300-12,000K.
-The second is a custom ``CfA grid'' synthesized using Kurucz models by Laird and Morse and is the same grid used by {\tt SPC}.
-This grid is very high quality because it has been hand-calibrated against the Sun and Vega, therefore it performs very well for solar like stars, however it only spans the comparatively narrow wavelength range 5050--5360\AA.
-
-For demonstrative purposes, the first fits we perform use only the Poisson variances from the original data set (WK1 and WP1) and ignore any correlations that might arise in the residual spectrum.
-This is equivalent to setting $b=1$ and all other amplitudes of $\cov$ to 0, and simplifies the likelihood to the familiar form of $\chi^2$.
-While the parameter uncertainty returned from these fits appears remarkably good, it is also unrealistic.
-This small uncertainty results because any likelihood function assumes that the data set is generated from the model\footnote{For example, other ``robust'' likelihood functions like the Student-t distribution assume that there is some likelihood of non-Gaussian ``outlier'' points that will occur.
-The Gaussian process likelihood that we assume captures the type of ``outliers'' that we might expect from our data.}, meaning that in this paradigm the only way the data set can differ from the model is through the addition of white Poisson noise.
-While both the PHOENIX and CfA stellar grids are \emph{incredibly good}, this assumption is invalid because the three 3 $\vt_\ast$ parameters are not enough to account for the behavior of thousands of spectral lines that might be in a reasonably large swath of spectrum.
-Some line strengths might be different from the synthetic model due to slight inaccuracies in the atomic constants or metallicity compositions that are different from solar in a few elements.
-Thus, in practice, a data spectrum is different from a synthetic model not just by added Poisson noise, but is also \emph{systematically} different in many spectral lines because the model cannot be a perfect representation of reality.
-Thus, assuming only white Poisson noise for a spectroscopic fit is equivalent to fitting a straight line to a series of data points whose variance is drastically underestimated.
-A linear least squares fit $\chi^2$ routine might return a small uncertainty estimate, however the slope and intercept of the line might be drastically biased.
-Therefore, it is clear that they model requires a way to account for this additional source of systematic ``noise.''
-
-While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
+A high resolution ($R\approx44,000$) optical spectrum of WASP-14 was obtained on 2009 June 14 
+using the Tillinghast Reflector Echelle Spectrograph \citep[TRES;][]{furesz08} on the Fred Lawrence 
+Whipple Observatory 1.5\,m telescope.  TRES delivers a 51-order echelle spectrum covering the 
+full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
+standard techniques in the TRES pipeline (cf.,~\citealt{mink11}; see \citealt{torres12} for more 
+specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
+
+\citet{torres12} originally acquired these data in an effort to measure the WASP-14 stellar 
+properties, and thereby better constrain the parameters of the transiting exoplanet it hosts.  To 
+that end, they employed the {\tt MOOG} and {\tt SPC} approaches outlined in Section 
+\ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
+free parameter or assumed to be fixed (to $\log g = 4.29$), based on an inference of the mean 
+stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
+evolution models in the H-R diagram \citep{joshi09}.  Only parameters inferences from the latter 
+assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
+synthesize a suite of individual \ion{Fe}{1} and \ion{Fe}{2} lines in the $\sim$4300--6750\AA\ 
+region, whose equivalent widths were compared with the TRES measurements in a simple least-squares 
+sense.  This approach suggests that $T_{\rm eff} = 6150\pm75$\,K and $[{\rm Fe/H}] = 
+-0.34\pm0.10$\,dex.  The {\tt SPC} analysis relied on a direct comparison between three TRES 
+spectral orders covering $\sim$5100-5400\,\AA\ and synthetic spectra from the {\sc CfA/Kurucz} 
+library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
+i = 3.9\pm0.5$\,km s$^{-1}$.  \hili{It is worth noting that these parameter inferences represent an 
+average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
+the systematic spread between the best-fit parameter values.} \comm{This is something we need to 
+discuss more with Dave and Willie.}
+
+Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
+it is most relevant to compare the \citeauthor{torres12}~{\tt SPC} result with our framework.  
+Using the same TRES spectral orders (for only a single WASP-14 observation, as noted above), the 
+{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer 
+parameters that are entirely consistent with the {\tt SPC} approach: $T_{\rm eff} = 6517\pm13$\,K, 
+$[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and $v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~X displays 
+some relevant marginal projections of the posterior-space, and Table~X compiles our estimates of 
+the peaks of the marginal posteriors and associated confidence intervals for all parameters 
+employed in the modeling.  Figure~Y demonstrates the overall quality of this approach in 
+reproducing the WASP-14 spectrum and, most significantly, the structure of the associated 
+residuals.  Figure~Z compares the inferred parameters with the various determinations made by 
+\citet{torres12}.
+
+\comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
+want to see Table 2 in this paper, or any reference to specific model or test identifications.  It 
+is too confusing, and it obscures the key points we want to make.  I appreciate your desire to 
+demonstrate the complexity of the model sequentially, but it does not add anything useful/tangible 
+here: it is already done very well in Section 2.  I feel like up to here we have 
+established ``reproducibility" well.  The remainder of this section should be solely about 
+systematics.  First, I want a paragraph breaking down the origins of the uncertainties (see the 
+paragraph I commented out below...I don't like it, but it should give you an idea about what I 
+mean.  If I interpret your Google tables correctly, you're getting about half the uncertainty from 
+Poisson noise, and the other half from the global covariance kernel.  This is where you need to be 
+able to defend the seemingly small uncertainties you're finding, even with a non-trivial covariance 
+matrix.  Is it really true that a model with 0.03\,dex lower metallicity cannot describe the data 
+as well?  Why?  [This is not to say I don't believe it, but it {\it will} raise questions...]}
+
+%Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
+%into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
+%(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
+%that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
+%range that no local covariance kernels need to be instantiated; in this specific case we use only a 
+%global covariance kernel.}  To break down their respective contributions, we performed the 
+%same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
+%likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
+%best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
+%$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
+%uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
+%are generated from the model; in this paradigm, the only way to introduce deviation is through 
+%white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
+%($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
+%spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
+%have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
+%noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
+%systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
+%posteriors.
+
+\comm{Next, you want to address the observational systematics issue.  Willie (and basically 
+everyone else) add empirically-estimated systematic terms in quadrature that make some pretty 
+unjustified assumptions: namely that they are Gaussian and independent in each parameter.  If I 
+understant Willie's paper correctly, there are 4 TRES observations of WASP-14.  If that is true, I 
+think we need to fit each of them independently.  I'd like to know what ``systematic" scatter you 
+would derive if you wanted to employ the standard approach.  My gut tells me that it's important to 
+simultaneously model all 4 spectra in orders 22-24, but if you think that's not 
+computationally tractable, we can discuss more.}
+%It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
+Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
+Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
+
+\comm{Finally, you can transition into the model systematics issue.  Here you want to do the 
+modeling for Phoenix in orders 22-24 and demonstrate that you get much different parameters.  You 
+will have to show the results, and then can take that opportunity to discuss the local covariance 
+kernels.  Just show the results and make these two points, don't belabor the details: I don't 
+know what to make of the weird Phoenix parameters.}
+
+\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
+orders with Phoenix.}
+
+%While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
 For demonstrative purposes, for the next test (WP2) we use the common technique of ``sigma-clipping,'' whereby we manually mask any region of the spectrum which has a more than a 3 sigma residual near the best fit parameters, refit, and then proceed until there are no more outliers.
@@
 
-These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+%These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
 For the next tests, (WK2 and WP3), we introduce the global covariance kernel (Eqn.~\ref{eqn:global}) to the fit.
@@
 Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
-
-A major advantage of using the MCMC algorithm to explore the multidimensional probability space is that it provides numerical samples in each dimension.
-Therefore, marginalizing out a parameter (i.e., numerically integrating over a dimension in probability space) is as simple as combining all of the samples in this dimension.
-This enables us to present a posterior of the stellar parameters $\vT$ (Figure~\ref{fig:stellar_posterior}) which has been marginalized over all of the nuisance hyperparameters.
-This posterior is the final estimate of the stellar parameters which incorporates any inherent uncertainty due to flux-calibration (via $\cheb$) and model mismatch (via $\cov$).
 
@@
 
-For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
+%For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
 These kernels can map out the regions of the spectrum which are outliers.
@@
 
-In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
+%In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
 Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.
@@
 \subsection{Gl~51}
-\todo{Tests run using no covariance, covariance, just the lines, just same region of spectrum. Adding local covariance kernels to Na and Ca lines that are underpredicted.}
-
-Because the spectrum of Gl~51 is the mid-infrared, we are unable to use the CfA grid for these comparisons. All Gl~51 tests were made using the PHOENIX grid.
+
+% Data description of Gl 51
+A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
+Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
+(IRTF).  SPEX is a cross-dispersed echelle that covers the red-optical to thermal-infrared spectrum 
+(0.7--5.5\,$\mu$m) in two spectrograph settings.  These data were obtained as part of the IRTF 
+spectral standard library project \citep{cushing05,rayner09}, and were processed through the 
+well-vetted {\tt Spextool} pipeline \citep{cushing04,vacca03} to deliver a fully-calibrated 
+spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.
+
+Modeling late-type stellar atmosphere structures and their spectra is considerably more complex 
+than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
+opacities, as well as the presence of complex condensates (clouds) at the coolest temperatures.  
+\comm{should probably add some citations to that statement.  Make a comment recognizing that 
+various approaches have been taken, in context of Mann.}  \citet{rojas-ayala12} used a $K$-band 
+spectrum at similar resolution to the SPEX data to estimate the key Gl~51 stellar parameters based 
+on a line index technique.  They measured equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} 
+doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, along with a pseudo-continuum index defined over 
+three narrow passbands, 2.070--2.090, 2.235--2.255, and 2.360--2.380\,$\mu$m (see their Figure 2 
+for a useful visualization), designed to probe the spectral curvature induced by H$_2$O absorption 
+bands.  With reference to a custom grid of {\sc BT-Settl} spectra, \citet{rojas-ayala12} find that 
+these indices are consistent with $T_{\rm eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$.
+
+\comm{Now, present your results in a format that parallels the WASP-14 subsection.  You need a new 
+Table and a sequence of figures; introduce them all here.}
+
+\comm{Commentary on the necessity and utility of local covariance kernels here, with discussion of why Na and Ca 
+lines are a pain.  It is also worth commenting on the potential for flexibly extending the shape of 
+the kernel to mimic molecular bandheads, if one wanted to try and fit say the TiO bands.}
+%In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
+They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
+\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
+When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that
+
+\comm{A simple two-sentence (or so) wrap up on the future promise of this approach for late-type 
+stars and their complex spectra.}
 

### 2014-08-19 tex/paper.tex
gl51 fix logg

@@
 
-%Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
-%into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
-%(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
-%that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
-%range that no local covariance kernels need to be instantiated; in this specific case we use only a 
-%global covariance kernel.}  To break down their respective contributions, we performed the 
-%same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
-%likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
-%best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
-%$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
-%uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
-%are generated from the model; in this paradigm, the only way to introduce deviation is through 
-%white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
-%($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
-%spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
-%have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
-%noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
-%systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
-%posteriors.
+Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
+into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
+(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
+that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
+range that no local covariance kernels need to be instantiated; in this specific case we use only a 
+global covariance kernel.}  To break down their respective contributions, we performed the 
+same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
+likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
+best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
+$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
+uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
+are generated from the model; in this paradigm, the only way to introduce deviation is through 
+white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
+($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
+spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
+have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
+noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
+systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
+posteriors.
 

### 2014-08-19 StellarSpectra/model.py
gl51 fix logg

@@ -684,6 +684,8 @@ class StellarSampler(Sampler):
             params = self.model_list[0].zip_stellar_p(p[:4])
             others = p[4:]
         else:
+            #Coming in as temp, Z, vsini, vz, logOmega...
+
             params = self.model_list[0].zip_stellar_p(p[:3])
             others = p[3:]
             params.update({"logg": self.fix_logg})
@@ -932,9 +934,13 @@ class RegionsSampler(GibbsSubController):
             if w in wl[covered]:
                 continue
             else:
-                #instantiate region and update coverage
-                self.create_region_sampler(w)
-                covered = self.order_model.CovarianceMatrix.get_region_coverage()
+                #check to make sure region is not *right* at the edge of the spectrum
+                if w <= np.min(wl) or w >= np.max(wl):
+                    continue
+                else:
+                    #instantiate region and update coverage
+                    self.create_region_sampler(w)
+                    covered = self.order_model.CovarianceMatrix.get_region_coverage()
 
     # def run_mcmc(self, *args, **kwargs):
     #

### 2014-08-19 StellarSpectra/utils.py

@@ -1,5 +1,6 @@
 import numpy as np
 import multiprocessing as mp
+import StellarSpectra.constants as C
 
 def random_draws(cov, num, nprocesses=mp.cpu_count()):
     '''
@@ -18,8 +19,43 @@ def random_draws(cov, num, nprocesses=mp.cpu_count()):
     return np.array(result.get())
 
 
-def k_3_2(r, l):
-    return (1 + np.sqrt(3) * r/l) * np.exp(-np.sqrt(3) * r/l)
+
+#Set of kernels *exactly* as defined in extern/cov.h
+@np.vectorize
+def k_gloabl(r, a, l):
+    r0=6.*l
+    taper = (0.5 + 0.5 * np.cos(np.pi * r/r0))
+    if r >= taper:
+        return 0.
+    else:
+        return taper * a**2 * (1 + np.sqrt(3) * r/l) * np.exp(-np.sqrt(3) * r/l)
+
+@np.vectorize
+def k_local(x0, x1, a, mu, sigma):
+    r0 = 4.0 * sigma #spot where kernel goes to 0
+    rx0 = C.c_kms / mu * np.abs(x0 - mu)
+    rx1 = C.c_kms / mu * np.abs(x1 - mu)
+    r_tap = rx0 if rx0 > rx1 else rx1 #choose the larger distance
+
+    if r_tap >= r0:
+        return 0.
+    else:
+        taper = (0.5 + 0.5 * np.cos(np.pi * r_tap/r0))
+        return taper * a**2 * np.exp(-0.5 * C.c_kms**2/mu**2 * ((x0 - mu)**2 + (x1 - mu)**2)/sigma**2)
+
+
+def k_global_func(x0i, x1i, x0v=None, x1v=None, a=None, l=None):
+    x0 = x0v[x0i]
+    x1 = x1v[x1i]
+    r = np.abs(x1 - x0)
+    pass
+
+def k_local_func(x0i, x1i, x0v=None, x1v=None, a=None, mu=None, sigma=None):
+    x0 = x0v[x0i]
+    x1 = x1v[x1i]
+    pass
+
+
 
 @np.vectorize
 def hann(r, r0):
@@ -91,49 +127,26 @@ def gauss_hann_func(x0i, x1i, x0v=None, x1v=None, amp=None, mu=None, sigma=None)
[…]

### 2014-08-19 hdfutils.py

@@ -1,0 +1,53 @@
+#Base package for all HDF tools
+
+#Basically, import this package and all of the argparsing, file commenting, should be done already.
+
+#Scripts like hdfmultiple, hdfcat, hdfplot, and hdfregion should all use this common core.
+
+
+class FlatchainTree:
+    '''
+    Object defined to wrap a Flatchain structure in order to facilitate combining, burning, etc.
+
+    The Tree will always follow the same structure.
+
+    flatchains.hdf5:
+
+    stellar samples:    stellar
+
+    folder for model:   0
+
+        folder for order: 22
+
+                        cheb
+                        cov
+                        cov_region00
+                        cov_region01
+                        cov_region02
+                        ....
+
+
+        folder for order: 23
+
+                        cheb
+                        cov
+                        cov_region00
+                        cov_region01
+                        cov_region02
+                        ....
+
+    folder for model:   1
+
+
+    '''
+    pass
+
+    def __add__(self, other):
+        pass
+
+
+class OldFlatchainTree:
+    '''
+    The old structure which assumed only 1 DataSpectrum. For legacy's sake.
+    '''
+    pass

### 2014-08-19 scripts/stars/base_lnprob_Kurucz.py

@@ -1,180 +1,0 @@
-from StellarSpectra.model import Model, StellarSampler, ChebSampler, MegaSampler, CovGlobalSampler, RegionsSampler
-from StellarSpectra.spectrum import DataSpectrum
-from StellarSpectra.grid_tools import TRES, HDF5Interface
-import StellarSpectra.constants as C
-import numpy as np
-import yaml
-import os
-import shutil
-import logging
-
-import argparse
-parser = argparse.ArgumentParser(prog="base_lnprob.py", description="Run StellarSpectra fitting model.")
-parser.add_argument("-i", "--input", help="*.yaml file specifying parameters.")
-parser.add_argument("-r", "--run_index", help="Which run (of those running concurrently) is this? All data will "
-                                        "be written into this directory, overwriting any that exists.")
-parser.add_argument("-p", "--perturb", type=float, help="Randomly perturb the starting position of the "
-                                                                 "chain, as a multiple of the jump parameters.")
-args = parser.parse_args()
-
-if args.input: #
-    #assert that we actually specified a *.yaml file
-    if ".yaml" in args.input:
-        yaml_file = args.input
-        f = open(args.input)
-        config = yaml.load(f)
-        f.close()
-
-    else:
-        import sys
-        sys.exit("Must specify a *.yaml file.")
-
-else:
-    #load the default config file
-    yaml_file = "scripts/stars/input.yaml"
-    f = open(yaml_file)
-    config = yaml.load(f)
-    f.close()
-
-outdir = config['outdir']
-name = config['name']
-base = outdir + name + "run{:0>2}/"
-
-#This code is necessary for multiple simultaneous runs on odyssey, so that different runs do not write into the same
-#output directory
-if args.run_index == None:
-    run_index = 0
-    while os.path.exists(base.format(run_index)) and (run_index < 20):
-        print(base.format(run_index), "exists")
-        run_index += 1
-    outdir = base.format(run_index)
-
-else:
-    run_index = args.run_index
-    outdir = base.format(run_index)
-    #Delete this outdir, if it exists
-    if os.path.exists(outdir):
-        print("Deleting", outdir)
[…]

### 2014-08-21 hdfmultiple.py
plotting available

@@ -109,24 +109,24 @@ def get_flatchains(key):
     return [hdf5.get(key)[:] for hdf5 in hdf5list]
 
 #This was good
-# #Order list will always be a 2D list, with the items being flatchains
-# ordersList = []
-# for order in orders:
-#
-#     print("Adding cheb for order {}".format(order))
-#     temp = [get_flatchains("{}/cheb".format(order))]
-#     if yes_cov:
-#         print("Adding cov for order {}".format(order))
-#         temp += [get_flatchains("{}/cov".format(order))]
-#
-#     #do not read in anything about regions here, that is for hdfregions.py, since there needs to be matching logic
-#
-#     #accumulate all of the orders
-#     ordersList += [temp]
+#Order list will always be a 2D list, with the items being flatchains
+ordersList = []
+for order in orders:
 
-# order22list = [order22cheblist, order22covlist]
-# order23list = [order23cheblist, order23covlist]
-# orderlist = [order22list, order23list]
+    print("Adding cheb for order {}".format(order))
+    temp = [get_flatchains("{}/cheb".format(order))]
+    if yes_cov:
+        print("Adding cov for order {}".format(order))
+        temp += [get_flatchains("{}/cov".format(order))]
+
+    #do not read in anything about regions here, that is for hdfregions.py, since there needs to be matching logic
+
+    #accumulate all of the orders
+    ordersList += [temp]
+
+order22list = [order22cheblist, order22covlist]
+order23list = [order23cheblist, order23covlist]
+orderlist = [order22list, order23list]
 # That way, we can do something like gelman_rubin(order22cheblist)
 
 print("Thinning by ", args.thin)

### 2014-08-21 tex/paper.tex
plotting available

@@
 \newcommand{\HK}{$\textrm{H}_2$O-K2}
+
+%Chebyshev coefficients RULE!
+\newcommand{\cc}[2]{c_{#2}^{(#1)}} %echelle order first, then degree coefficient
 
@@
 employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
-a spectral synthesis front-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
+a spectral synthesis back-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
 \citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
@@
 [{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
-{\it any} ``front-end" that generates a model spectrum.  In our adopted approach, the model 
+{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model 
 spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
@@
 a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
-tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
-be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
-could be avoided if the spectral library was computed over a refined grid (with a substantial 
-up-front computational investment); but for the time being, we can empirically propagate these 
-interpolation uncertainties into the likelihood calculations (as will be described in 
+tri-linear interpolation over that refined grid.  Overall, this interpolation
+technique is found to be accurate within a few percent per full-resolution
+model pixel and $\lesssim 1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
+this pre-interpolation could be avoided if the spectral library was computed
+over a refined grid (with a substantial up-front computational investment); but
+for the time being, we can empirically propagate these interpolation
+uncertainties into the likelihood calculations (as will be described in 
 Section~\ref{subsec:likelihood}).
-%% \todo{tests regarding interpolation accuracy}.
 
@@
 they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
-the physical parameters.  This concern is treated externally to any modeling procedure, most often 
+the physical parameters.  This concern is usually treated externally to any modeling procedure, most often 
 by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
@@
 broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
-these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$4 
-pixel scale, corresponding to the 3\,km s$^{-1}$ width of $\mathcal{F}_v^{\rm inst}$.
+these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
+pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
 
@@
   \includegraphics{correlation_panel.pdf}
-  \caption{\textbf{Top}: the big spectrum.
-\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
+  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
+  parameters drawn from the posterior distribution, and the residuals that
+  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
+  to show the mildly covariant structure produced by slight mismatch between
+  the data and model spectra.   \textbf{Right} The autocorrelation of the
+residual sequence shown at left. Notice that there is significant correlation
+for offsets of $\lesssim 5$ pixels.}
 \label{fig:class0}
@@
 to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
-scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matern} demonstrates that random draws from such a 
+scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matrix} demonstrates that random draws from such a 
 kernel readily produce correlated structure similar to those seen in a typical residual spectrum.
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
-\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
-There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
-This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
-\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
-The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
-The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
-\label{fig:matern}
+\includegraphics{figs/matrix_compilation.pdf}
+\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals.
+\textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region.
+Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix.
+\textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
+At the bottom of each panel is the envelope formed by 50 random draws (blue) overlaid with the actual residuals from the gray region.
+\textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum.
+\textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
+\textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
+Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines.
+The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit.
+Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not.
+All parameters of the covariance kernels are determined self-consistently along with the stellar parameters.
+}
+\label{fig:matrix}
 \end{center}
@@
 \begin{equation}
-\vR = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
-\end{equation}
-with amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
+\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
+\end{equation}
+with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
 Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
@@
 \begin{equation} \label{eqn:kregion}
-  \mathcal{K}^{l,k}_{ij} = \frac{a_k^2}{2\pi\sigma_k^2} \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
+  \mathcal{K}^{l,k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
 \end{equation}
@@
 with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
-again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure that a sparse covariance 
+again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
 matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
-width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:region} shows how this non-stationary Gaussian 
+width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:matrix} shows how this non-stationary Gaussian 
 kernel generates a localized region of enhanced variance that successfully mimics the kind of 
@@
 the probabilistic framework (as opposed to manual or threshold-based clipping or masking).
-
-\begin{figure*}[!t]
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
-The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
-These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
-\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
-The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
-The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
-\label{fig:region}
-\end{center}
-\end{figure*}
 
@@
 elements in the set of covariance (hyper-)parameters, $\cov$.
-
-%\begin{figure*}[!t]
-%\begin{center}
-%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
-%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
-%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
-The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
-These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
-%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
-The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
-The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
-%\label{fig:region}
-%\end{center}
-%\end{figure*}
 
@@
 
-We assume uniform priors on the global covariance parameters.  For the local covariance kernels, we 
-adopt flat priors for the amplitudes and means, \{$a_k$, $\mu_k$\}, and construct a specialized 
-prior for the widths, \{$\sigma_k$\}, that is flat when $\sigma_k < \sigma_v$ and has a rapid 
-(Gaussian) taper to zero for larger widths.  This form is designed to prevent the local 
-kernels from diffusing to large $\sigma_k$ and low $a_k$, since such behavior is better treated by 
-the global kernel.
+We assume uniform priors on the global covariance parameters.  For the local
+covariance kernels, we adopt flat priors for the amplitudes and means, \{$a_k$,
+$\mu_k$\}, and construct a specialized prior for the widths, \{$\sigma_k$\},
+that is flat below the instrumental width (when $\sigma_k < \sigma_v$) and has
+a rapid (logistic) taper to zero for larger widths.  This form is designed to
+prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
+since such behavior is better treated by the global kernel.
 
@@
 
-\item Decide whether to instantiate any local (non-stationary) covariance kernels.  We adopt an 
-iterative thresholding criterion: if a local patch of the residual signal is at least $3\times$ the 
-amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this triggers the 
-addition of a local kernel.\footnote{There is no practical reason to delete local kernels.  If the 
-parameters have changed such that a given local kernel is no longer required, that kernel amplitude 
-will be driven towards zero and represent a negligible contribution to $\vC$; in effect, the model 
-will act as if the kernel were deleted automatically.}
-
-\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
+\item If any local (non-stationary) covariance kernels have been instantiated (see below), sample in their hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
 $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
@@
 
-After burn-in, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
+Local kernels are instantiated by the following procedure:  An ``average''
+residual spectrum is created by combining ~500 spectra residuals stored from a
+burn-in period using only global covariance kernels (thinned to account for
+autocorrelation of the Markov chain). Then, using an iterative thresholding
+criterion similar in procedure to
+``sigma clipping,''\footnote{Although the procedure of identifying outliers is
+  similar to sigma clipping, the inference on the stellar parameters is
+  emphatically \emph{not} the same as sigma clipping, since we now determine
+the weights of these spectral outliers inside of a probabilistic framework, rather than
+directly setting their weights to zero as in sigma clipping.} we identify the largest outlier
+pixel not already covered by a local kernel, instantiate a local kernel at its location, and repeat until there are no pixels above a certain
+threshold left.\footnote{There is no practical
+  reason to delete local kernels once instantiated.  If the parameters have
+  changed such that a given local kernel is no longer required, that kernel
+  amplitude will be driven towards zero and represent a negligible contribution
+  to $\vC$; in effect, the model will act as if the kernel were deleted
+automatically.} Alternative local kernel instantiation schemes,
+such as re-evaluating the location of kernels with each Gibbs sampler
+iteration, have yielded similar results to the thresholding approach.
+However, this approach most consistently converges and only has a modest computational overhead.
+We choose four times the standard deviation of the average residual spectrum as our threshold.
+Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which in turn slightly reduces the amplitude of the global kernel.
+Taken to the logical conclusion, if the threshold was set low enough then there would be a local kernel for each spectral line and no global kernel would be required.
+We find that the final stellar posterior is relatively insensitive to the choice of threshold level as long as it is set low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
+Once all of the local kernels have been instantiated, the Gibbs sampler is run for another period of burn-in.
+
+After burn-in is complete, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
 diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_Kurucz_logg.pdf}
+\caption{Kurucz WASP-14.}
+\label{fig:residuals_Kurucz}
+\end{center}
+\end{figure*}
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 6516.8 & 12.6 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.27 & 0.01 dex \\
+$v \sin i$ & 4.29 & 0.05 $\kms$\\
+$v_z$ & -4.62 & 0.02 $\kms$ \\
+$\log_{10} \Omega$ & -12.72 & 0.0004 \\
+\cutinhead{Order 22 nuisance parameters}
+$\cc{22}{0}$ & & \\
+$\cc{22}{1}$ & & \\
+$\cc{22}{2}$ & & \\
+$\cc{22}{3}$ & & \\
+$b_{22}$ & & \\
+$a_{{\rm g}, 22}$ & & \\
+$l_{22}$ & & \\
+\cutinhead{Order 23 nuisance parameters}
+$\cc{23}{0}$ & & \\
+$\cc{23}{1}$ & & \\
+$\cc{23}{2}$ & & \\
+$\cc{23}{3}$ & & \\
+$b_{23}$ & & \\
+$a_{{\rm g}, 23}$ & & \\
+$l_{23}$ & & \\
+\cutinhead{Order 24 nuisance parameters}
+$\cc{24}{0}$ & & \\
+$\cc{24}{1}$ & & \\
+$\cc{24}{2}$ & & \\
+$\cc{24}{3}$ & & \\
+$b_{24}$ & & \\
+$a_{{\rm g}, 24}$ & & \\
+$l_{24}$ & & \\
+\enddata
+\end{deluxetable}
+
 \comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
@@
 know what to make of the weird Phoenix parameters.}
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\caption{Kurucz WASP-14.}
+\label{fig:residuals_PHOENIX}
+\end{center}
+\end{figure*}
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 6516.8 & 12.6 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.27 & 0.01 dex \\
+$v \sin i$ & 4.29 & 0.05 $\kms$\\
+$v_z$ & -4.62 & 0.02 $\kms$ \\
+$\log_{10} \Omega$ & -12.72 & 0.0004 \\
+\cutinhead{Order 22 nuisance parameters}
+$\cc{22}{0}$ & & \\
+$\cc{22}{1}$ & & \\
+$\cc{22}{2}$ & & \\
+$\cc{22}{3}$ & & \\
+$b_{22}$ & & \\
+$a_{{\rm g}, 22}$ & & \\
+$l_{22}$ & & \\
+\cutinhead{Order 23 nuisance parameters}
+$\cc{23}{0}$ & & \\
+$\cc{23}{1}$ & & \\
+$\cc{23}{2}$ & & \\
+$\cc{23}{3}$ & & \\
+$b_{23}$ & & \\
+$a_{{\rm g}, 23}$ & & \\
+$l_{23}$ & & \\
+\cutinhead{Order 24 nuisance parameters}
+$\cc{24}{0}$ & & \\
+$\cc{24}{1}$ & & \\
+$\cc{24}{2}$ & & \\
+$\cc{24}{3}$ & & \\
+$b_{24}$ & & \\
+$a_{{\rm g}, 24}$ & & \\
+$l_{24}$ & & \\
+\enddata
+\end{deluxetable}
 
@@
 
-\begin{deluxetable}{lllrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
-\tablecaption{\label{table:tests} Results}
-\tablehead{
-  \colhead{Test ID} & \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
-  \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
-  \multicolumn{2}{c}{$\Z \pm \sigma$} & \colhead{comments}
-}
-\startdata
-WK1 & WASP-14 & 23 & 5135-5235 & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
-WK2 & WASP-14 & 23 & 5135-5235 & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
-WK3 & WASP-14 & 23 & 5135-5235 & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
-WP1 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
-WP2 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
-WP3 & WASP-14 & 23 & 5135-5235 & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
-WP4 & WASP-14 & 23 & 5135-5235 & PHOENIX & & & & & & & Global and local \\
-&WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
-&WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
-&WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
-&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
-&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
-&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
-&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
-GP1 &Gl~51 & \nodata & IR & PHOENIX & & & & & & & Poisson\\
-\enddata
-\tablenotetext{a}{Using the same assumptions in the \citet{torres12} {\tt SPC} analysis: same region of spectrum, same spectral library, and fixing $\log g$=$4.29$, the value determined by \citet{joshi09} from the exoplanet transit.
-For the WASP-14 spectrum acquired by \emph{TRES}, \citet{torres12} use {\tt SPC} to determine $T_{\rm eff}$=6507$\pm$50K and $\Z$=-0.15$\pm$0.08.
-With {\tt MOOG}, they determine $T_{\rm eff}$=6150$\pm$75K and $\Z$=-0.34$\pm$0.10.
-Because {\tt SPC} uses a cross-correlation technique which only accounts for the white noise and we model residual spectrum as correlated noise, there may be differences in the uncertainty estimate.}
-\end{deluxetable}
-
 \appendix

### 2014-08-21 hdfplot.py

@@ -1,305 +1,14 @@
 #!/usr/bin/env python
-import os
-import numpy as np
+
+#Defines a bunch of global variables and also parses args.
+import hdfutils
+
 
 
 '''
 Script designed to plot the HDF5 output from MCMC runs.
 '''
 
-#Plot kw
-label_dict = {"temp":r"$T_{\rm eff}$", "logg":r"$\log_{10} g$", "Z":r"$[{\rm Fe}/{\rm H}]$", "alpha":r"$[\alpha/{\rm Fe}]$",
-    "vsini":r"$v \sin i$", "vz":r"$v_z$",  "vz1":r"$v_z1$", "logOmega1":r"$\log{10} \Omega 1$", "logOmega":r"$\log_{10} \Omega$", "logc0":r"$\log_{10} c_0$",
-    "c1":r"$c_0$", "c2":r"$c_1$", "c3":r"$c_3$",
-    "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}$", "l":r"$l$",
-    "h":r"$h$", "loga":r"$\log_{10} a$", "mu":r"$\mu$", "sigma":r"$\sigma$"}
+for ftree in hdfutils.flatchainTreeList:
+    hdfutils.plot(ftree, clip_stellar=hdfutils.args.stellar_params)
 
-
-import argparse
-parser = argparse.ArgumentParser(description="Plot the MCMC output. Default is to plot everything possible.")
-parser.add_argument("HDF5file", help="The HDF5 file containing the MCMC samples.")
-parser.add_argument("--lnprob", help="The HDF5 file containing the lnprobability chains.")
-parser.add_argument("-o", "--outdir", default="mcmcplot", help="Output directory to contain all plots.")
-parser.add_argument("--clobber", action="store_true", help="Overwrite existing output directory?")
-
-parser.add_argument("-t", "--triangle", action="store_true", help="Make a triangle (staircase) plot of the parameters.")
-parser.add_argument("--chain", action="store_true", help="Make a plot of the position of the chains.")
-parser.add_argument("--range", nargs=2, help="start and end ranges for chain plot, separated by WHITESPACE")
-parser.add_argument("--acor", action="store_true", help="Calculate the autocorrelation of the chain")
-parser.add_argument("--acor-window", type=int, default=50, help="window to compute acor with")
-
-parser.add_argument("--cov", action="store_true", help="Estimate the covariance between two parameters.")
-
[…]

### 2014-08-21 hdfutils.py
plotting available

@@ -1,9 +1,161 @@
+#!/usr/bin/env python
+import os
+import numpy as np
+import matplotlib
+#matplotlib.rc("axes", labelsize="large")
+import triangle
+
 #Base package for all HDF tools
 
 #Basically, import this package and all of the argparsing, file commenting, should be done already.
 
 #Scripts like hdfmultiple, hdfcat, hdfplot, and hdfregion should all use this common core.
 
+import argparse
+parser = argparse.ArgumentParser(description="Measure statistics across multiple chains.")
+parser.add_argument("--dir", action="store_true", help="Concatenate all of the flatchains stored within run* "
+                                                       "folders in the current directory. Designed to collate runs from a JobArray.")
+parser.add_argument("--outdir", default="mcmcplot", help="Output directory to contain all plots.")
+parser.add_argument("--output", default="combined.hdf5", help="Output HDF5 file.")
+parser.add_argument("--clobber", action="store_true", help="Overwrite existing files?")
+
+parser.add_argument("--lnprob", help="The HDF5 file containing the lnprobability chains.")
+parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
+
+parser.add_argument("--burn", type=int, default=0, help="How many samples to discard from the beginning of the chain "
+                                                        "for burn in.")
+parser.add_argument("--thin", type=int, default=1, help="Thin the chain by this factor. E.g., --thin 100 will take "
+                                                        "every 100th sample.")
+parser.add_argument("--keep", type=int, default=0, help="How many samples to keep from the end of the chain, "
+                                                        "the beginning of the chain will be for burn in.")
+parser.add_argument("--range", nargs=2, help="start and end ranges for chain plot, separated by WHITESPACE")
+
+
+parser.add_argument("--chain", action="store_true", help="Make a plot of the position of the chains.")
+parser.add_argument("-t", "--triangle", action="store_true", help="Make a triangle (staircase) plot of the parameters.")
+parser.add_argument("--stellar_params", nargs="*", default="all", help="A list of which stellar parameters to plot, "
+                                                                       "separated by WHITESPACE. Default is to plot all.")
[…]

### 2014-08-21 hdfplot.py
more hfd tools

@@ -4,11 +4,8 @@
 import hdfutils
 
 
-
 '''
-Script designed to plot the HDF5 output from MCMC runs.
+Script designed to plot the HDF5 output from MCMC runs. It will only plot the first chain, to prevent overwriting.
 '''
 
-for ftree in hdfutils.flatchainTreeList:
-    hdfutils.plot(ftree, clip_stellar=hdfutils.args.stellar_params)
-
+hdfutils.plot(hdfutils.flatchainTreeList[0], clip_stellar=hdfutils.args.stellar_params)

### 2014-08-21 tex/paper.tex
more hfd tools

@@
 corresponding library spectrum.  After an extensive exploration of such calculations 
-\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
-achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
-a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
-tri-linear interpolation over that refined grid.  Overall, this interpolation
+\citep[see also][]{husser12}, we concluded that this interpolation
 technique is found to be accurate within a few percent per full-resolution
@@
 \begin{center}
-  \includegraphics[draft, width=2.5in]{figs/stellar_triangle_Kurucz.png}
+  \includegraphics[width=0.5\textwidth]{figs/kurucz_triangle.pdf}
   \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.

### 2014-08-21 StellarSpectra/utils.py

@@ -33,6 +33,13 @@ def envelope(spectra):
     return np.min(spectra, axis=0), np.max(spectra, axis=0)
 
 
+def std_envelope(spectra):
+    '''
+    Given a 2D array of spectra, shape (Nspectra, Npix), return the std envelope of these as two spectra.
+    '''
+    std = np.std(spectra, axis=0)
+    return -std, std
+
 def visualize_draws(spectra, num=20):
     '''
     Given a 2D array of spectra, shape (Nspectra, Npix), visualize them to choose the most illustrative "random"
@@ -50,14 +57,12 @@ def visualize_draws(spectra, num=20):
     plt.show()
 
 
-
 def saveall(fig, fname, formats=[".png", ".pdf", ".svg"]):
     '''
     Save a matplotlib figure instance to many different formats
     '''
     for format in formats:
         fig.savefig(fname + format)
-
 
 
 #Set of kernels *exactly* as defined in extern/cov.h

### 2014-08-21 hdfmultiple.py

@@ -1,220 +1,5 @@
 #!/usr/bin/env python
-import os
-import numpy as np
 
-'''
-hdfmultiple.py
+import hdfutils
 
-Tools for measuring the output of many runs.
-
-Main functionality:
-
-1) Visualize the chains from all of the runs on the same plot. This way we can check convergence.
-   1.1) Ability to plot the starting positions of each chain, stellar[0], as well.
-2) Measure the between-chain correlation. Compute the Gelman-Rubin statistics.
-'''
-
-#Plot kw
-label_dict = {"temp":r"$T_{\rm eff}$", "logg":r"$\log_{10} g$", "Z":r"$[{\rm Fe}/{\rm H}]$", "alpha":r"$[\alpha/{\rm Fe}]$",
-              "vsini":r"$v \sin i$", "vz":r"$v_z$", "vz1":r"$v_z1$", "logOmega":r"$\log_{10} \Omega$",
-              "logOmega1":r"$\log{10} \Omega 1$", "logc0":r"$\log_{10} c_0$",
-              "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}", "l":r"$l$",
-              "h":r"$h$", "loga":r"$\log_{10} a$", "mu":r"$\mu$", "sigma":r"$\sigma$"}
-
-import argparse
-parser = argparse.ArgumentParser(description="Measure statistics across multiple chains.")
-parser.add_argument("--dir", action="store_true", help="Concatenate all of the flatchains stored within run* "
-                                                       "folders in the current directory. Designed to collate runs from a JobArray.")
-parser.add_argument("-o", "--outdir", default="mcmcplot", help="Output directory to contain all plots.")
-parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
-parser.add_argument("--clobber", action="store_true", help="Overwrite existing files?")
-parser.add_argument("--chain", action="store_true", help="Make a plot of the position of the chains.")
-parser.add_argument("--burn", type=int, default=0, help="How many samples to discard from the beginning of the chain "
-                                                        "for burn in.")
-parser.add_argument("--thin", type=int, default=1, help="Thin the chain by this factor. E.g., --thin 100 will take "
-                                                        "every 100th sample.")
-parser.add_argument("--stellar_params", nargs="*", default="all", help="A list of which stellar parameters to plot, "
[…]

### 2014-08-21 hdfutils.py

@@ -18,6 +18,7 @@ parser.add_argument("--dir", action="store_true", help="Concatenate all of the f
 parser.add_argument("--outdir", default="mcmcplot", help="Output directory to contain all plots.")
 parser.add_argument("--output", default="combined.hdf5", help="Output HDF5 file.")
 parser.add_argument("--clobber", action="store_true", help="Overwrite existing files?")
+parser.add_argument("--old", action="store_true", help="Old format flatchains.hdf5 files?")
 
 parser.add_argument("--lnprob", help="The HDF5 file containing the lnprobability chains.")
 parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
@@ -43,6 +44,17 @@ parser.add_argument("--acor-window", type=int, default=50, help="window to compu
 parser.add_argument("--cov", action="store_true", help="Estimate the covariance between two parameters.")
 
 args = parser.parse_args()
+
+#Check to see if outdir exists. If --clobber, overwrite, otherwise exit.
+if os.path.exists(args.outdir):
+    if not args.clobber:
+        import sys
+        sys.exit("Error: --outdir already exists and --clobber is not set. Exiting.")
+else:
+    os.makedirs(args.outdir)
+
+args.outdir += "/"
+
 
 import h5py
 
@@ -192,7 +204,7 @@ class FlatchainTree:
 
 
     '''
-    def __init__(self, file):
+    def __init__(self, file, old=False):
         #Load everything from the HDF5File into a bunch of Flatchain objects
         #We will always have the stellar samples
 
@@ -201,9 +213,13 @@ class FlatchainTree:
         self.stellar = Flatchain.from_dset("stellar", hdf5.get("stellar"))
         self.shape = self.stellar.shape
 
-        #For each key that is a number, make a model.
-        self.modelTreeDict = {id:ModelTree.from_dset(id, dset) for (id, dset) in ((key, dset) for (key, dset) in hdf5
-                                .items() if key.isdigit())}
+        if old:
+            self.modelTreeDict = {"0":ModelTree.from_dset("0", hdf5)}
+
+        else:
+            #For each key that is a number, make a model.
+            self.modelTreeDict = {id:ModelTree.from_dset(id, dset) for (id, dset) in ((key, dset) for (key, dset) in hdf5
+                                    .items() if key.isdigit())}
 
[…]

### 2014-08-21 hdfutils.py
ignor regions in GR

@@ -68,7 +68,7 @@ label_dict = {"temp":r"$T_{\rm eff}$", "logg":r"$\log_{10} g$", "Z":r"$[{\rm Fe}
               "logOmega2":r"$\log{10} \Omega 2$",
               "logOmega3":r"$\log{10} \Omega 3$",
               "logc0":r"$\log_{10} c_0$", "c1":r"$c_0$", "c2":r"$c_1$", "c3":r"$c_3$",
-              "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}", "l":r"$l$",
+              "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}$", "l":r"$l$",
               "h":r"$h$", "loga":r"$\log_{10} a$", "mu":r"$\mu$", "sigma":r"$\sigma$"}
 
 #Additionally, there should be an object for each flatchain, that stores param_tuple and samples
@@ -350,8 +350,10 @@ def GR_list(flatchainTreeList):
 
     for key in keys:
         print("\n", key)
-        #Accumulate one set of samples (from each FlatchainTree) corresponding to the key
-        gelman_rubin([fchain_dict[key].samples for fchain_dict in fchain_dicts])
+        #We can't do this for the region chains in this format, because they are not guaranteed to match.
+        if "region" not in key:
+            #Accumulate one set of samples (from each FlatchainTree) corresponding to the key
+            gelman_rubin([fchain_dict[key].samples for fchain_dict in fchain_dicts])
 
 
 def cat_list(file, flatchainTreeList):
@@ -386,6 +388,10 @@ def plot(flatchainTree, base=args.outdir, triangle_plot=args.triangle, chain_plo
     '''
     Make a bunch of plots to diagnose how the run went.
     '''
+
+    import matplotlib
+    matplotlib.rc("font", size=16)
+
     #Navigate the flatchain tree, and each time we encounter a flatchain, plot it.
     for flatchain in flatchainTree.flatchains:
 
@@ -397,7 +403,7 @@ def plot(flatchainTree, base=args.outdir, triangle_plot=args.triangle, chain_plo
         labels = [label_dict.get(key, "unknown") for key in params]
 
         figure = triangle.corner(samples, labels=labels, quantiles=[0.16, 0.5, 0.84],
-                                 show_titles=True, title_args={"fontsize": 10})
+                                 show_titles=True, title_args={"fontsize": 16})
         figure.savefig(base + flatchain.id + format)
 
 

### 2014-08-21 tex/paper.tex
ignor regions in GR

@@
 \textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
-At the bottom of each panel is the envelope formed by 50 random draws (blue) overlaid with the actual residuals from the gray region.
+At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region.
 \textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum.
@@
 \includegraphics{figs/residuals_Kurucz_logg.pdf}
-\caption{Kurucz WASP-14.}
+\caption{Kurucz WASP-14. One, two, and three sigma contours formed from 200 random draws from the covariance matrix.}
 \label{fig:residuals_Kurucz}
@@
 \end{figure*}
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
+Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+}
+\label{fig:stellar_posterior}
+\end{center}
+\end{figure}
 
@@
 \cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6516.8 & 12.6 K \\
+$T_{\rm eff}$ & 6517.1 & 12.5 K \\
 $\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
@@
 $v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.72 & 0.0004 \\
+$v_z$ & -4.620 & 0.016 $\kms$ \\
+$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
 \cutinhead{Order 22 nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b_{22}$ & & \\
-$a_{{\rm g}, 22}$ & & \\
-$l_{22}$ & & \\
+$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
+$\cc{22}{1}$ & -0.0142 & 0.0015\\
+$\cc{22}{2}$ & -0.0145 & 0.0017\\
+$\cc{22}{3}$ & -0.0031 & 0.0011\\
+\\
+$b_{22}$ & 1.04 & 0.02 \\
+$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.020 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{22}$ & 7.54 & 0.77 $\kms$ \\
+
 \cutinhead{Order 23 nuisance parameters}
-$\cc{23}{0}$ & & \\
-$\cc{23}{1}$ & & \\
-$\cc{23}{2}$ & & \\
-$\cc{23}{3}$ & & \\
-$b_{23}$ & & \\
-$a_{{\rm g}, 23}$ & & \\
-$l_{23}$ & & \\
+$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
+$\cc{23}{1}$ & -0.0133 & 0.0016 \\
+$\cc{23}{2}$ & -0.0247 & 0.0014 \\
+$\cc{23}{3}$ &  0.0012 &  0.0013\\
+\\
+$b_{23}$ & 1.01 & 0.02 \\
+$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.015 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{23}$ & 6.52 & 0.50 \\
+
 \cutinhead{Order 24 nuisance parameters}
-$\cc{24}{0}$ & & \\
-$\cc{24}{1}$ & & \\
-$\cc{24}{2}$ & & \\
-$\cc{24}{3}$ & & \\
-$b_{24}$ & & \\
-$a_{{\rm g}, 24}$ & & \\
-$l_{24}$ & & \\
+$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00} \\
+$\cc{24}{1}$ & -0.0103 & 0.0016 \\
+$\cc{24}{2}$ & -0.0112 & 0.0013\\
+$\cc{24}{3}$ &  0.0002 & 0.0012\\
+\\
+$b_{24}$ & 1.02 & 0.02 \\
+$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{24}$ & 7.33 & 0.59 $\kms$ \\
 \enddata
 \end{deluxetable}
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics{figs/metacomparison.pdf}
+\caption{Comparing all the posteriors.}
+\label{fig:metacomparison}
+\end{center}
+\end{figure}
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/kurucz_triangle.pdf}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
-Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:stellar_posterior}
-\end{center}
-\end{figure*}
-
 %For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.

### 2014-08-22 tex/paper.tex
ready for PH run

@@
 
-\begin{figure*}[!htb]
+\begin{figure}[!htb]
 \begin{center}
-  \includegraphics{correlation_panel.pdf}
+  \includegraphics{residuals_23.pdf}
+  \includegraphics{class0_residuals.pdf}
+  \includegraphics{class0_autocorrelation.pdf}
   \caption{\textbf{Top}: A typical model fit to a data spectrum, with
@@
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
-  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
-Plotted with the increased variance envelope over the residuals.
-}
-\label{fig:regions}
-\end{center}
-\end{figure*}
-
 %In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+  \includegraphics[draft,width=0.9\textwidth,height=3in]{figs/Gl51_residuals.pdf}
+  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
+Plotted with the increased variance envelope over the residuals.}
+\label{fig:regions}
+\end{center}
+\end{figure*}
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[draft,width=0.45\textwidth,height=3in]{figs/Gl51_posterior.pdf}
+  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
+Plotted with the increased variance envelope over the residuals.}
+\label{fig:gl_posterior}
+\end{center}
+\end{figure}
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:PHOENIX}Gl51 {\sc Phoenix} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ &  &  K \\
+$\log_{10} g$ & \multicolumn{2}{c}{5.0 dex} \\
+$\Z$ &  &  dex \\
+$v \sin i$ &  &  $\kms$\\
+$v_z$ &  &  $\kms$ \\
+$\log_{10} \Omega$ &  & \\
+\cutinhead{Nuisance parameters}
+$\cc{22}{0}$ & & \\
+$\cc{22}{1}$ & & \\
+$\cc{22}{2}$ & & \\
+$\cc{22}{3}$ & & \\
+$b$ & & \\
+$a_{\rm g} $ & & \\
+$l$ & & \\
+\cutinhead{Local kernel parameters}
+& & \\
+& & \\
+& & \\
+\hline\\
+& & \\
+& & \\
+& & \\
+\hline\\
+& & \\
+& & \\
+& & \\
+\enddata
+\end{deluxetable}
+
 \comm{A simple two-sentence (or so) wrap up on the future promise of this approach for late-type 
@@
 
-\begin{figure*}[!htb]
+\begin{figure}[!htb]
 \begin{center}
-  \includegraphics{figs/badlines.pdf}
+  \includegraphics{figs/badlines0.pdf}
+  \includegraphics{figs/badlines1.pdf}
+  \includegraphics{figs/badlines2.pdf}
+  \includegraphics{figs/badlines3.pdf}
   \caption{A collection of spectral lines which have imperfect model fits.
-    From left to right: \textbf{Class 0} The majority of spectral lines
-    ($\gtrsim 60$\%) will have minor differences in strength between data and
-    model spectrum, which produce low-amplitude correlations in the residuals
-    on the length scale of the width of a typical spectral line.  \textbf{Class
-    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-    of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-    correlated residuals of moderate amplitude.  The difficulty
-    with class III lines is that for any specific line, there might exist a
-    $\vT$ that will fit the line, but there does not exist a $\vT$ that
+    \textbf{First panel} The majority of spectral lines ($\gtrsim 60$\%) will
+    have minor differences in strength between data and model spectrum, which
+    produce low-amplitude correlations in the residuals on the length scale of
+    the width of a typical spectral line.  \textbf{Second panel}: Sometimes
+    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
+    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Third panel}: Sometimes ($\lesssim
+    5$\% of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals. \textbf{Fourth panel}: If the line
+    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
+    there will be many correlated residuals of moderate amplitude.  The
+    difficulty with class III lines is that for any specific line, there might
+    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
     will properly fit \emph{all} the lines.}

### 2014-08-22 tex/paper.tex
Sect 4 outline notes

--- 2014-08-21 tex/paper.tex
@@
 \newcommand{\HK}{$\textrm{H}_2$O-K2}
-
-%Chebyshev coefficients RULE!
-\newcommand{\cc}[2]{c_{#2}^{(#1)}} %echelle order first, then degree coefficient
 
@@
 employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
-a spectral synthesis back-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
+a spectral synthesis front-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
 \citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
@@
 [{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
-{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model 
+{\it any} ``front-end" that generates a model spectrum.  In our adopted approach, the model 
 spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
@@
 corresponding library spectrum.  After an extensive exploration of such calculations 
-\citep[see also][]{husser12}, we concluded that this interpolation
-technique is found to be accurate within a few percent per full-resolution
-model pixel and $\lesssim 1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
-this pre-interpolation could be avoided if the spectral library was computed
-over a refined grid (with a substantial up-front computational investment); but
-for the time being, we can empirically propagate these interpolation
-uncertainties into the likelihood calculations (as will be described in 
+\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
+achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
+a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
+tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
+be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
+could be avoided if the spectral library was computed over a refined grid (with a substantial 
+up-front computational investment); but for the time being, we can empirically propagate these 
+interpolation uncertainties into the likelihood calculations (as will be described in 
 Section~\ref{subsec:likelihood}).  
+%% \todo{tests regarding interpolation accuracy}.
 
@@
 they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
-the physical parameters.  This concern is usually treated externally to any modeling procedure, most often 
+the physical parameters.  This concern is treated externally to any modeling procedure, most often 
 by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
@@
 broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
-these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
-pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
+these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$4 
+pixel scale, corresponding to the 3\,km s$^{-1}$ width of $\mathcal{F}_v^{\rm inst}$.
 
@@
   \includegraphics{correlation_panel.pdf}
-  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
-  parameters drawn from the posterior distribution, and the residuals that
-  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
-  to show the mildly covariant structure produced by slight mismatch between
-  the data and model spectra.   \textbf{Right} The autocorrelation of the
-residual sequence shown at left. Notice that there is significant correlation
-for offsets of $\lesssim 5$ pixels.}
+  \caption{\textbf{Top}: the big spectrum.
+\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
+  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
 \label{fig:class0}
@@
 to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
-scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matrix} demonstrates that random draws from such a 
+scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matern} demonstrates that random draws from such a 
 kernel readily produce correlated structure similar to those seen in a typical residual spectrum.
@@
 \begin{center}
-\includegraphics{figs/matrix_compilation.pdf}
-\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals.
-\textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region.
-Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix.
-\textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
-At the bottom of each panel is the envelope formed by 50 random draws (blue) overlaid with the actual residuals from the gray region.
-\textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum.
-\textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
-\textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
-Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines.
-The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit.
-Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not.
-All parameters of the covariance kernels are determined self-consistently along with the stellar parameters.
-}
-\label{fig:matrix}
+\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
+\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
+There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
+This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
+\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
+The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
+The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
+\label{fig:matern}
 \end{center}
@@
 \begin{equation}
-\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
-\end{equation}
-with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
+\vR = \frac{a_k}{\sqrt{2\pi}{\sigma_k}} \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
+\end{equation}
+with amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
 Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
@@
 \begin{equation} \label{eqn:kregion}
-  \mathcal{K}^{l,k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
+  \mathcal{K}^{l,k}_{ij} = \frac{a_k^2}{2\pi\sigma_k^2} \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
 \end{equation}
@@
 with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
-again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
+again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure that a sparse covariance 
 matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
-width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:matrix} shows how this non-stationary Gaussian 
+width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:region} shows how this non-stationary Gaussian 
 kernel generates a localized region of enhanced variance that successfully mimics the kind of 
@@
 the probabilistic framework (as opposed to manual or threshold-based clipping or masking).
+
+\begin{figure*}[!t]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
+The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
+These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
+\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
+The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
+\label{fig:region}
+\end{center}
+\end{figure*}
 
@@
 elements in the set of covariance (hyper-)parameters, $\cov$.
+
+%\begin{figure*}[!t]
+%\begin{center}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
+%\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
+%\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}).
+The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals.
+These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
+%\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix.
+The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix.
+The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
+%\label{fig:region}
+%\end{center}
+%\end{figure*}
 
@@
 
-We assume uniform priors on the global covariance parameters.  For the local
-covariance kernels, we adopt flat priors for the amplitudes and means, \{$a_k$,
-$\mu_k$\}, and construct a specialized prior for the widths, \{$\sigma_k$\},
-that is flat below the instrumental width (when $\sigma_k < \sigma_v$) and has
-a rapid (logistic) taper to zero for larger widths.  This form is designed to
-prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
-since such behavior is better treated by the global kernel.
+We assume uniform priors on the global covariance parameters.  For the local covariance kernels, we 
+adopt flat priors for the amplitudes and means, \{$a_k$, $\mu_k$\}, and construct a specialized 
+prior for the widths, \{$\sigma_k$\}, that is flat when $\sigma_k < \sigma_v$ and has a rapid 
+(Gaussian) taper to zero for larger widths.  This form is designed to prevent the local 
+kernels from diffusing to large $\sigma_k$ and low $a_k$, since such behavior is better treated by 
+the global kernel.
 
@@
 
-\item If any local (non-stationary) covariance kernels have been instantiated (see below), sample in their hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
+\item Decide whether to instantiate any local (non-stationary) covariance kernels.  We adopt an 
+iterative thresholding criterion: if a local patch of the residual signal is at least $3\times$ the 
+amplitude of the global covariance kernel (i.e., where $\vR \ge 3 \, a_g^i$), this triggers the 
+addition of a local kernel.\footnote{There is no practical reason to delete local kernels.  If the 
+parameters have changed such that a given local kernel is no longer required, that kernel amplitude 
+will be driven towards zero and represent a negligible contribution to $\vC$; in effect, the model 
+will act as if the kernel were deleted automatically.}
+
+\item Now sample in the local covariance hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
 $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
@@
 
-Local kernels are instantiated by the following procedure:  An ``average''
-residual spectrum is created by combining ~500 spectra residuals stored from a
-burn-in period using only global covariance kernels (thinned to account for
-autocorrelation of the Markov chain). Then, using an iterative thresholding
-criterion similar in procedure to
-``sigma clipping,''\footnote{Although the procedure of identifying outliers is
-  similar to sigma clipping, the inference on the stellar parameters is
-  emphatically \emph{not} the same as sigma clipping, since we now determine
-the weights of these spectral outliers inside of a probabilistic framework, rather than
-directly setting their weights to zero as in sigma clipping.} we identify the largest outlier
-pixel not already covered by a local kernel, instantiate a local kernel at its location, and repeat until there are no pixels above a certain
-threshold left.\footnote{There is no practical
-  reason to delete local kernels once instantiated.  If the parameters have
-  changed such that a given local kernel is no longer required, that kernel
-  amplitude will be driven towards zero and represent a negligible contribution
-  to $\vC$; in effect, the model will act as if the kernel were deleted
-automatically.} Alternative local kernel instantiation schemes,
-such as re-evaluating the location of kernels with each Gibbs sampler
-iteration, have yielded similar results to the thresholding approach.
-However, this approach most consistently converges and only has a modest computational overhead.
-We choose four times the standard deviation of the average residual spectrum as our threshold.
-Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which in turn slightly reduces the amplitude of the global kernel.
-Taken to the logical conclusion, if the threshold was set low enough then there would be a local kernel for each spectral line and no global kernel would be required.
-We find that the final stellar posterior is relatively insensitive to the choice of threshold level as long as it is set low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
-Once all of the local kernels have been instantiated, the Gibbs sampler is run for another period of burn-in.
-
-After burn-in is complete, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
+After burn-in, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
 diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/residuals_Kurucz_logg.pdf}
-\caption{Kurucz WASP-14.}
-\label{fig:residuals_Kurucz}
-\end{center}
-\end{figure*}
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6516.8 & 12.6 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.72 & 0.0004 \\
-\cutinhead{Order 22 nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b_{22}$ & & \\
-$a_{{\rm g}, 22}$ & & \\
-$l_{22}$ & & \\
-\cutinhead{Order 23 nuisance parameters}
-$\cc{23}{0}$ & & \\
-$\cc{23}{1}$ & & \\
-$\cc{23}{2}$ & & \\
-$\cc{23}{3}$ & & \\
-$b_{23}$ & & \\
-$a_{{\rm g}, 23}$ & & \\
-$l_{23}$ & & \\
-\cutinhead{Order 24 nuisance parameters}
-$\cc{24}{0}$ & & \\
-$\cc{24}{1}$ & & \\
-$\cc{24}{2}$ & & \\
-$\cc{24}{3}$ & & \\
-$b_{24}$ & & \\
-$a_{{\rm g}, 24}$ & & \\
-$l_{24}$ & & \\
-\enddata
-\end{deluxetable}
-
 \comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
@@
 
-Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
-into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
-(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
-that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
-range that no local covariance kernels need to be instantiated; in this specific case we use only a 
-global covariance kernel.}  To break down their respective contributions, we performed the 
-same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
-likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
-best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
-$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
-uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
-are generated from the model; in this paradigm, the only way to introduce deviation is through 
-white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
-($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
-spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
-have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
-noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
-systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
-posteriors.
+%Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
+%into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
+%(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
+%that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
+%range that no local covariance kernels need to be instantiated; in this specific case we use only a 
+%global covariance kernel.}  To break down their respective contributions, we performed the 
+%same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
+%likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
+%best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
+%$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
+%uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
+%are generated from the model; in this paradigm, the only way to introduce deviation is through 
+%white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
+%($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
+%spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
+%have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
+%noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
+%systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
+%posteriors.
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/residuals_PHOENIX_logg.pdf}
-\caption{Kurucz WASP-14.}
-\label{fig:residuals_PHOENIX}
-\end{center}
-\end{figure*}
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6516.8 & 12.6 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.72 & 0.0004 \\
-\cutinhead{Order 22 nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b_{22}$ & & \\
-$a_{{\rm g}, 22}$ & & \\
-$l_{22}$ & & \\
-\cutinhead{Order 23 nuisance parameters}
-$\cc{23}{0}$ & & \\
-$\cc{23}{1}$ & & \\
-$\cc{23}{2}$ & & \\
-$\cc{23}{3}$ & & \\
-$b_{23}$ & & \\
-$a_{{\rm g}, 23}$ & & \\
-$l_{23}$ & & \\
-\cutinhead{Order 24 nuisance parameters}
-$\cc{24}{0}$ & & \\
-$\cc{24}{1}$ & & \\
-$\cc{24}{2}$ & & \\
-$\cc{24}{3}$ & & \\
-$b_{24}$ & & \\
-$a_{{\rm g}, 24}$ & & \\
-$l_{24}$ & & \\
-\enddata
-\end{deluxetable}
-
 \comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/kurucz_triangle.pdf}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[draft, width=2.5in]{figs/stellar_triangle_Kurucz.png}
+%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
 These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
 Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:stellar_posterior}
-\end{center}
-\end{figure*}
+%}
+%\label{fig:stellar_posterior}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
-  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
+%  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
 Plotted with the increased variance envelope over the residuals.
-}
-\label{fig:regions}
-\end{center}
-\end{figure*}
+%}
+%\label{fig:regions}
+%\end{center}
+%\end{figure*}
 
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    From left to right: \textbf{Class 0} The majority of spectral lines
-    ($\gtrsim 60$\%) will have minor differences in strength between data and
-    model spectrum, which produce low-amplitude correlations in the residuals
-    on the length scale of the width of a typical spectral line.  \textbf{Class
-    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-    of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-    correlated residuals of moderate amplitude.  The difficulty
-    with class III lines is that for any specific line, there might exist a
-    $\vT$ that will fit the line, but there does not exist a $\vT$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure*}
-
-\section{Discussion}
-\label{sec:discussion}
-%\begin{rant about why this method rocks}
-There are many advantages to placing spectroscopic inference within a Bayesian framework.
+%\begin{figure*}[!htb]
+%\begin{center}
+%  \includegraphics{figs/badlines.pdf}
+%  \caption{A collection of spectral lines which have imperfect model fits.
+%    From left to right: \textbf{Class 0} The majority of spectral lines
+%    ($\gtrsim 60$\%) will have minor differences in strength between data and
+%    model spectrum, which produce low-amplitude correlations in the residuals
+%    on the length scale of the width of a typical spectral line.  \textbf{Class
+%    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
+%    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+%    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
+%    of all lines), an extraneous line in the model leaves a large, highly
+%    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
+%    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
+%    correlated residuals of moderate amplitude.  The difficulty
+%    with class III lines is that for any specific line, there might exist a
+%    $\vT$ that will fit the line, but there does not exist a $\vT$ that
+%    will properly fit \emph{all} the lines.}
+%\label{fig:badlines}
+%\end{center}
+%\end{figure*}
+
+\clearpage
+
+\section{Discussion} \label{sec:discussion}
+
+\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
+be tough to do concisely, but here's what I think the framework should be: (1) a statement of the 
+problem (i.e., that we want to extract physical information from a spectrum without bias or 
+systematic error, but the models we have are not perfect); (2) a [brief] reminder of the standard 
+approaches to skirt this issue; (3) a summary of our approach; (4) a statement that we demonstrated 
+the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
+
+% SA: The following too paragraphs are too preachy, and address a different issue than we ended up 
+% focusing on in this paper.  You can remove them entirely.
+%
+%There are many advantages to placing spectroscopic inference within a Bayesian framework.
 Given the large costs of acquiring a high signal to noise spectrum, it would be advantageous to utilize as much as possible of the spectrum for performing inference.
@@
 These ``fit metrics'' (line indices) are similar to distance functions or a figure of merit.
-
-These figures of merit do make a lot of sense, and one can see why they are used, but given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, instead of going through the intermediary step of a figure of merit, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
+%
+%These figures of merit do make a lot of sense, and one can see why they are used, but given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, instead of going through the intermediary step of a figure of merit, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
 In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
@@
 
-The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
+\comm{Next, more specifically highlight why we developed this framework, and what its 
+role should be in general spectroscopic inference.  I think its worth highlighting several key 
+advantages: (1) you are casting the standard data selection/weighting/masking procedure in an 
+internally consistent, Bayesian format that appropriately preserves and propagates information 
+content even in the face of imperfect models (along the lines of what you've written in the 
+commented-out paragraphs below, but more concise/direct); (2) by doing that, you allow the user to 
+actually exploit their full datasets, which should then reflect a more appropriate inference on the 
+parameters of interest without concern that some bad or less-informative features will screw them 
+up (sort of like what I commented out above, but less preachy) -- its in here 
+that we should suggest that this framework could be incorporated as a backend to standard methods, 
+where it essentially is employed to define an appropriate likelihood; and (3) it creates an 
+opportunity to develop quantitative feedback to improve the models, or ultimately to 
+pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
+since you will go into depth on it below). (est.~2 paragraphs)} \\
+
+%The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
 When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
@@
 However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
-%\end{rant}
-
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+%
+%The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
+see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
+local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
+what you've already written (commented out below) is in good shape, but you might decide to change 
+the format/organization given the suggested update in structure for this section.  Go out on a 
+positive note: the mechanics already exist, and this tool can be used to build better models!}
+
+%The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
@@
 
-\todo{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
+%\todo{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
 This will tell us about the frequency and distribution of spectral modelling errors.
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
-
 \bibliographystyle{yahapj}
@@
 
-\section{Tables}
-
-\begin{deluxetable}{ll}
-\tablecaption{Nomenclature used in this document}
-\tablehead{
-  \colhead{Symbol} & \colhead{Description}
-}
-\startdata
-$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
-$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-  & are applied during ``post processing'' of the synthetic spectrum\\
-$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
-
-$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
- & a single spectral order $i$ \\
-$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
-  & a multi-order echelle spectrum with $N$ spectral orders  \\
- $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
- & except those for order $2$ \\
-$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
-$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
- & a multi-order echelle spectrum with $N$ spectral orders  \\
- $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
- & except those for order $2$ \\
-
-$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
-$\vT$ & the collection of all stellar parameters \\
-$\vD$ & data spectrum \\
-$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
-$\vR$ & $\vD - \vM $, residuals  \\
-$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
-$\vC_{ij}$ & an individual element in the covariance matrix\\
-$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-
-$\Kglobal$ & global covariance kernel\\
-$\Klocal$ & local covariance kernel\\
-$w$ & Hanning taper for kernels \\
-\enddata
-\end{deluxetable}
-
-\appendix
-
-This version of the paper was generated
- from a git repository available at \url{http://github.com/iancze/StellarSpectra/}
- with git hash \texttt{\githash\,(\gitdate)}.
+%\section{Tables}
+%
+%\begin{deluxetable}{ll}
+%\tablecaption{Nomenclature used in this document}
+%\tablehead{
+%  \colhead{Symbol} & \colhead{Description}
+%}
+%\startdata
+%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+%$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+%$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
+%$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+%  & are applied during ``post processing'' of the synthetic spectrum\\
+%$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
+%
+%
+%$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
+% & a single spectral order $i$ \\
+%$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
+%  & a multi-order echelle spectrum with $N$ spectral orders  \\
+% $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
+% & except those for order $2$ \\
+%$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
+%$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
+% & a multi-order echelle spectrum with $N$ spectral orders  \\
+% $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
+% & except those for order $2$ \\
+%
+%$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
+%$\vT$ & the collection of all stellar parameters \\
+%$\vD$ & data spectrum \\
+%$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
+%$\vR$ & $\vD - \vM $, residuals  \\
+%$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
+%$\vC_{ij}$ & an individual element in the covariance matrix\\
+%$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
+%
+%$\Kglobal$ & global covariance kernel\\
+%$\Klocal$ & local covariance kernel\\
+%$w$ & Hanning taper for kernels \\
+%\enddata
+%\end{deluxetable}
+
+%\begin{deluxetable}{lllrlr@{$\pm$}lr@{$\pm$}lr@{$\pm$}ll} 
+%\tablecaption{\label{table:tests} Results}
+%\tablehead{
+%  \colhead{Test ID} & \colhead{Object} & \colhead{Orders} & \colhead{$\lambda$ range [\AA]} & \colhead{library} &
+%  \multicolumn{2}{c}{$T_{\rm eff} \pm \sigma$} & \multicolumn{2}{c}{$\log g \pm \sigma$} &
+%  \multicolumn{2}{c}{$\Z \pm \sigma$} & \colhead{comments}
+%}
+%\startdata
+%WK1 & WASP-14 & 23 & 5135-5235 & Kurucz & 6241 & 15 & 3.88 & 0.02 & -0.43 & 0.01 & Poisson\\
+%WK2 & WASP-14 & 23 & 5135-5235 & Kurucz & 6264 & 31 & 3.92 & 0.04 & -0.42 & 0.02 & Global\\
+%WK3 & WASP-14 & 23 & 5135-5235 & Kurucz & & & \multicolumn{2}{c}{\nodata} & & & Matern; fixed $\log g = 4.29$\\
+%WP1 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6023.3 & 4.3 & 3.84 & 0.01 & -0.50 & 7.7e-4 & Poisson\\
+%WP2 & WASP-14 & 23 & 5135-5235 & PHOENIX & 6000.72 & 3 & 3.94 & 0.01 & -0.50 & 2.6e-3 & Poisson, masked \\
+%WP3 & WASP-14 & 23 & 5135-5235 & PHOENIX & 5984.9 & 74 & 3.80 & 0.09 & -0.66 & 0.07 & Global, no mask \\
+%WP4 & WASP-14 & 23 & 5135-5235 & PHOENIX & & & & & & & Global and local \\
+%&WASP-14 & 22-24 & 5060-5315 & Kurucz & & & & & & & Poisson\\
+%&WASP-14 & 22-24 & 5060-5315& Kurucz & & & & & & & Global\\
+%&WASP-14 & 22-24 & 5060-5315& Kurucz & & & \multicolumn{2}{c}{\nodata} & & & \tablenotemark{a}Global; fixed $\log g = 4.29$\\
+%&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Poisson\\
+%&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global; low because of 24\\
+%&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & \multicolumn{2}{c}{\nodata} & & & Global; fixed $\log g = 4.29$ \\
+%&WASP-14 & 22-24 & 5060-5315 & PHOENIX & & & & & & & Global and local\\
+%GP1 &Gl~51 & \nodata & IR & PHOENIX & & & & & & & Poisson\\
+%\enddata
+%\tablenotetext{a}{Using the same assumptions in the \citet{torres12} {\tt SPC} analysis: same region of spectrum, same spectral library, and fixing $\log g$=$4.29$, the value determined by \citet{joshi09} from the exoplanet transit.
+For the WASP-14 spectrum acquired by \emph{TRES}, \citet{torres12} use {\tt SPC} to determine $T_{\rm eff}$=6507$\pm$50K and $\Z$=-0.15$\pm$0.08.
+With {\tt MOOG}, they determine $T_{\rm eff}$=6150$\pm$75K and $\Z$=-0.34$\pm$0.10.
+Because {\tt SPC} uses a cross-correlation technique which only accounts for the white noise and we model residual spectrum as correlated noise, there may be differences in the uncertainty estimate.}
+%\end{deluxetable}
+
+%\appendix
+
+%This version of the paper was generated
+% from a git repository available at \url{http://github.com/iancze/StellarSpectra/}
+% with git hash \texttt{\githash\,(\gitdate)}.
 \end{document}

### 2014-08-22 tex/paper.tex

@@
 corresponding library spectrum.  After an extensive exploration of such calculations 
-\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
-achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
-a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
-tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
-be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
-could be avoided if the spectral library was computed over a refined grid (with a substantial 
-up-front computational investment); but for the time being, we can empirically propagate these 
-interpolation uncertainties into the likelihood calculations (as will be described in 
+\citep[see also][]{husser12}, we concluded that this interpolation
+technique is found to be accurate within a few percent per full-resolution
+model pixel and $\lesssim 1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
+this pre-interpolation could be avoided if the spectral library was computed
+over a refined grid (with a substantial up-front computational investment); but
+for the time being, we can empirically propagate these interpolation
+uncertainties into the likelihood calculations (as will be described in 
 Section~\ref{subsec:likelihood}).  
@@
 broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
-these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$4 
-pixel scale, corresponding to the 3\,km s$^{-1}$ width of $\mathcal{F}_v^{\rm inst}$.
-
-\begin{figure*}[!htb]
+these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
+pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
+
+\begin{figure}[!htb]
 \begin{center}
-  \includegraphics{correlation_panel.pdf}
-  \caption{\textbf{Top}: the big spectrum.
-\textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.
-  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
+  \includegraphics{residuals_23.pdf}
+  \includegraphics{class0_residuals.pdf}
+  \includegraphics{class0_autocorrelation.pdf}
+  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
+  parameters drawn from the posterior distribution, and the residuals that
+  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
+  to show the mildly covariant structure produced by slight mismatch between
+  the data and model spectra.   \textbf{Right} The autocorrelation of the
+residual sequence shown at left. Notice that there is significant correlation
+for offsets of $\lesssim 5$ pixels.}
 \label{fig:class0}
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figs/matern_matrix.pdf}
-\includegraphics[width=0.4\textwidth]{figs/matern_draw.pdf}
-\caption{\textbf{Left} Inset zoomed to show a small region of a typical covariance matrix, generated using the kernel in Equation~\ref{eqn:global} and common values for the hyperparameters.
-There is a small degree of covariance in elements within a few pixels off the diagonal, which quickly tapers off so that the majority of the matrix remains sparse ($\sigma_{ij} = 0$).
-This matrix is used to model a spectrum which is correlated on a $\sim$few pixel scale.
-\textbf{Right} To demonstrate that this matrix properly models the correlated structure of the residuals, we compare the residuals to random residuals generated from a multivariate normal distribution with this covariance matrix.
-The top panel shows the same residuals shown in Figure~\ref{fig:class0}, and below are plotted three sets of simulated residuals.
-The amplitude and correlation length of the simulated residuals closely approximates the structure of the actual pixel residuals.}
-\label{fig:matern}
+\includegraphics{figs/matrix_compilation.pdf}
+\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals.
+\textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region.
+Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix.
+\textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
+At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region.
+\textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum.
+\textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
+\textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
+Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines.
+The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit.
+Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not.
+All parameters of the covariance kernels are determined self-consistently along with the stellar parameters.
+}
+\label{fig:matrix}
 \end{center}
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_Kurucz_logg.pdf}
+\caption{Kurucz WASP-14. One, two, and three sigma contours formed from 200 random draws from the covariance matrix.}
+\label{fig:residuals_Kurucz}
+\end{center}
+\end{figure*}
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
+These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
+Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+}
+\label{fig:stellar_posterior}
+\end{center}
+\end{figure}
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 6517.1 & 12.5 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.27 & 0.01 dex \\
+$v \sin i$ & 4.29 & 0.05 $\kms$\\
+$v_z$ & -4.620 & 0.016 $\kms$ \\
+$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
+\cutinhead{Order 22 nuisance parameters}
+$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
+$\cc{22}{1}$ & -0.0142 & 0.0015\\
+$\cc{22}{2}$ & -0.0145 & 0.0017\\
+$\cc{22}{3}$ & -0.0031 & 0.0011\\
+\\
+$b_{22}$ & 1.04 & 0.02 \\
+$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.020 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{22}$ & 7.54 & 0.77 $\kms$ \\
+
+\cutinhead{Order 23 nuisance parameters}
+$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
+$\cc{23}{1}$ & -0.0133 & 0.0016 \\
+$\cc{23}{2}$ & -0.0247 & 0.0014 \\
+$\cc{23}{3}$ &  0.0012 &  0.0013\\
+\\
+$b_{23}$ & 1.01 & 0.02 \\
+$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.015 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{23}$ & 6.52 & 0.50 \\
+
+\cutinhead{Order 24 nuisance parameters}
+$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00} \\
+$\cc{24}{1}$ & -0.0103 & 0.0016 \\
+$\cc{24}{2}$ & -0.0112 & 0.0013\\
+$\cc{24}{3}$ &  0.0002 & 0.0012\\
+\\
+$b_{24}$ & 1.02 & 0.02 \\
+$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
+$l_{24}$ & 7.33 & 0.59 $\kms$ \\
+\enddata
+\end{deluxetable}
+
+\begin{figure}[!htb]
+\begin{center}
+\includegraphics{figs/metacomparison.pdf}
+\caption{Comparing all the posteriors.}
+\label{fig:metacomparison}
+\end{center}
+\end{figure}
+
 \comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
@@
 
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\caption{Kurucz WASP-14.}
+\label{fig:residuals_PHOENIX}
+\end{center}
+\end{figure*}
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 6516.8 & 12.6 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.27 & 0.01 dex \\
+$v \sin i$ & 4.29 & 0.05 $\kms$\\
+$v_z$ & -4.62 & 0.02 $\kms$ \\
+$\log_{10} \Omega$ & -12.72 & 0.0004 \\
+\cutinhead{Order 22 nuisance parameters}
+$\cc{22}{0}$ & & \\
+$\cc{22}{1}$ & & \\
+$\cc{22}{2}$ & & \\
+$\cc{22}{3}$ & & \\
+$b_{22}$ & & \\
+$a_{{\rm g}, 22}$ & & \\
+$l_{22}$ & & \\
+\cutinhead{Order 23 nuisance parameters}
+$\cc{23}{0}$ & & \\
+$\cc{23}{1}$ & & \\
+$\cc{23}{2}$ & & \\
+$\cc{23}{3}$ & & \\
+$b_{23}$ & & \\
+$a_{{\rm g}, 23}$ & & \\
+$l_{23}$ & & \\
+\cutinhead{Order 24 nuisance parameters}
+$\cc{24}{0}$ & & \\
+$\cc{24}{1}$ & & \\
+$\cc{24}{2}$ & & \\
+$\cc{24}{3}$ & & \\
+$b_{24}$ & & \\
+$a_{{\rm g}, 24}$ & & \\
+$l_{24}$ & & \\
+\enddata
+\end{deluxetable}
+
 \comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
@@
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics[draft, width=2.5in]{figs/stellar_triangle_Kurucz.png}
-%  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
-Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-%}
-%\label{fig:stellar_posterior}
-%\end{center}
-%\end{figure*}
-
 %For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
@@
 This can feed into a discussion about tweaking models.}
-
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics[draft,width=0.9\textwidth,height=4in]{figs/spectrum_region.pdf}
-%  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
-Plotted with the increased variance envelope over the residuals.
-%}
-%\label{fig:regions}
-%\end{center}
-%\end{figure*}
 
@@
 
-%\begin{figure*}[!htb]
-%\begin{center}
-%  \includegraphics{figs/badlines.pdf}
-%  \caption{A collection of spectral lines which have imperfect model fits.
-%    From left to right: \textbf{Class 0} The majority of spectral lines
-%    ($\gtrsim 60$\%) will have minor differences in strength between data and
-%    model spectrum, which produce low-amplitude correlations in the residuals
-%    on the length scale of the width of a typical spectral line.  \textbf{Class
-%    I}: Sometimes ($\lesssim 5$\% of all lines), a missing opacity source in
-%    the model (in this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-%    patch of negative residuals.  \textbf{Class II}: Sometimes ($\lesssim 5$\%
-%    of all lines), an extraneous line in the model leaves a large, highly
-%    correlated patch of positive residuals.  \textbf{Class III}: If the line strengths are
-%    substantially discrepant ($\lesssim 10$\% of all lines), there will be many
-%    correlated residuals of moderate amplitude.  The difficulty
-%    with class III lines is that for any specific line, there might exist a
-%    $\vT$ that will fit the line, but there does not exist a $\vT$ that
-%    will properly fit \emph{all} the lines.}
-%\label{fig:badlines}
-%\end{center}
-%\end{figure*}
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines0.pdf}
+  \includegraphics{figs/badlines1.pdf}
+  \includegraphics{figs/badlines2.pdf}
+  \includegraphics{figs/badlines3.pdf}
+  \caption{A collection of spectral lines which have imperfect model fits.
+    \textbf{First panel} The majority of spectral lines ($\gtrsim 60$\%) will
+    have minor differences in strength between data and model spectrum, which
+    produce low-amplitude correlations in the residuals on the length scale of
+    the width of a typical spectral line.  \textbf{Second panel}: Sometimes
+    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
+    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  \textbf{Third panel}: Sometimes ($\lesssim
+    5$\% of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals. \textbf{Fourth panel}: If the line
+    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
+    there will be many correlated residuals of moderate amplitude.  The
+    difficulty with class III lines is that for any specific line, there might
+    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure}
 
@@
 Or we could just speculate about it.}
+
+\section{Tables}
+
+\begin{deluxetable}{ll}
+\tablecaption{Nomenclature used in this document}
+\tablehead{
+  \colhead{Symbol} & \colhead{Description}
+}
+\startdata
+$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
+$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+  & are applied during ``post processing'' of the synthetic spectrum\\
+$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
 
@@
 
+\appendix
+
 %This version of the paper was generated

### 2014-08-23 hdfregion.py
trying regions

@@ -56,7 +56,7 @@ hdf5list = []
 filelist = []
 for file in files:
     try:
-        hdf5list += [h5py.File(file, "r")]
+        hdf5list += [h5py.File(file, "r")["0"]]
         filelist += [file]
     except OSError:
         print("{} does not exist, skipping.".format(file))
@@ -66,6 +66,7 @@ for file in files:
 #Load all of the samples from the a given order into one giant deque
 allRegions = deque()
 
+#Maybe we can just select on model 0 for now.
 hdf5 = hdf5list[0]
 orders = [int(key) for key in hdf5.keys() if key != "stellar"]
 orders.sort()

### 2014-08-23 tex/paper.tex
trying regions

--- 2014-08-22 tex/paper.tex
@@
 \newcommand{\hili}[1]{ \textcolor{green}{#1}}
+\newcommand{\ctext}[1]{ \textcolor{blue}{\% #1}}
 
@@
 \end{equation}
-where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
-interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
-is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
-undesirable level of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation quality 
-can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
-$\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the interpolated spectrum with the 
-corresponding library spectrum.  After an extensive exploration of such calculations 
-\citep[see also][]{husser12}, we concluded that this interpolation
-technique is found to be accurate within a few percent per full-resolution
-model pixel and $\lesssim 1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
+where we assign the symbol $\leadsto$ as an interpolation operator.  The
+multi-dimensional interpolation in Eq.~\ref{eqn:interp} needs to be performed
+many times, so computational efficiency is critical.  In practice, a simple
+tri-linear interpolation is suitably fast, but introduces an undesirable level
+of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation
+quality can be empirically estimated by performing the operation in
+Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ grid point, and then
+comparing the interpolated spectrum with the corresponding library spectrum.
+After an extensive exploration of such calculations \citep[see
+also][]{husser12}, we concluded that this interpolation technique is found to
+be accurate within a few percent per full-resolution model pixel and $\lesssim
+1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
 this pre-interpolation could be avoided if the spectral library was computed
@@
 autocorrelation of the Markov chain). Then, using an iterative thresholding
-criterion similar in procedure to
-``sigma clipping,''\footnote{Although the procedure of identifying outliers is
-  similar to sigma clipping, the inference on the stellar parameters is
-  emphatically \emph{not} the same as sigma clipping, since we now determine
-the weights of these spectral outliers inside of a probabilistic framework, rather than
-directly setting their weights to zero as in sigma clipping.} we identify the largest outlier
-pixel not already covered by a local kernel, instantiate a local kernel at its location, and repeat until there are no pixels above a certain
-threshold left.\footnote{There is no practical
-  reason to delete local kernels once instantiated.  If the parameters have
-  changed such that a given local kernel is no longer required, that kernel
-  amplitude will be driven towards zero and represent a negligible contribution
-  to $\vC$; in effect, the model will act as if the kernel were deleted
-automatically.} Alternative local kernel instantiation schemes,
-such as re-evaluating the location of kernels with each Gibbs sampler
-iteration, have yielded similar results to the thresholding approach.
-However, this approach most consistently converges and only has a modest computational overhead.
-We choose four times the standard deviation of the average residual spectrum as our threshold.
-Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which in turn slightly reduces the amplitude of the global kernel.
-Taken to the logical conclusion, if the threshold was set low enough then there would be a local kernel for each spectral line and no global kernel would be required.
-We find that the final stellar posterior is relatively insensitive to the choice of threshold level as long as it is set low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
-Once all of the local kernels have been instantiated, the Gibbs sampler is run for another period of burn-in.
+criterion similar in procedure to ``sigma clipping,''\footnote{Although the
+  procedure of identifying outliers is similar to sigma clipping, the inference
+  on the stellar parameters is emphatically \emph{not} the same as sigma
+clipping, since we now determine the weights of these spectral outliers inside
+of a probabilistic framework, rather than directly setting their weights to
+zero as in sigma clipping.} we identify the largest outlier pixel not already
+covered by a local kernel, instantiate a local kernel at its location, and
+repeat until there are no pixels above a certain threshold left.\footnote{There
+  is no practical reason to delete local kernels once instantiated.  If the
+  parameters have changed such that a given local kernel is no longer required,
+  that kernel amplitude will be driven towards zero and represent a negligible
+  contribution to $\vC$; in effect, the model will act as if the kernel were
+deleted automatically.} Alternative local kernel instantiation schemes, such as
+re-evaluating the location of kernels with each Gibbs sampler iteration, have
+yielded similar results to the thresholding approach. However, this approach
+most consistently converges and only has a modest computational overhead. We
+choose four times the standard deviation of the average residual spectrum as
+our threshold. Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the
+instantiation of more kernels, which in turn slightly reduces the amplitude of
+the global kernel. Taken to the logical conclusion, if the threshold was set
+low enough then there would be a local kernel for each spectral line and no
+global kernel would be required. We find that the final stellar posterior is
+relatively insensitive to the choice of threshold level as long as it is set
+low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
+Once all of the local kernels have been instantiated, the Gibbs sampler is run
+for another period of burn-in.
 
@@
 computationally tractable, we can discuss more.}
-%It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
+\ctext{It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
 Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
+Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.}
 
@@
 
-%While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
+\ctext{While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
 For demonstrative purposes, for the next test (WP2) we use the common technique of ``sigma-clipping,'' whereby we manually mask any region of the spectrum which has a more than a 3 sigma residual near the best fit parameters, refit, and then proceed until there are no more outliers.
@@
 For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
-Clipping these lines would discard important information and lead to an incorrect inference of stellar parameters.
-
-%These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+Clipping these lines would discard important information and lead to an incorrect inference of stellar parameters.}
+
+\ctext{These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
 For the next tests, (WK2 and WP3), we introduce the global covariance kernel (Eqn.~\ref{eqn:global}) to the fit.
@@
 Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
-
-%For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
+Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.}
+
+\ctext{For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
 These kernels can map out the regions of the spectrum which are outliers.
@@
 The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
-See Figure~\ref{fig:regions} for a cool visualization of what it looks like.
-\todo{Runs are almost converged.
-This can feed into a discussion about tweaking models.}
-
-%In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
-Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.
+See Figure~\ref{fig:regions} for a cool visualization of what it looks like.}
+
+\ctext{In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
+Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.}
 
@@
 
-% Data description of Gl 51
 A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
@@
 the kernel to mimic molecular bandheads, if one wanted to try and fit say the TiO bands.}
-%In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
+\ctext{In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
 They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
 \citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
-When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that
+When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that}
 
@@
 
-\section{Discussion}
-\label{sec:discussion}
-%\begin{rant about why this method rocks}
-There are many advantages to placing spectroscopic inference within a Bayesian framework.
-Given the large costs of acquiring a high signal to noise spectrum, it would be advantageous to utilize as much as possible of the spectrum for performing inference.
-You go to a telescope to take a spectrum, you should use the whole spectrum, not just subsets of the spectrum that are certain well chosen lines.
-True, these lines have been identified by the accumulated knowledge of experts to show high discriminatory power against the stellar parameters of interest, but there is a gap in the chain of logic between these indices and the core fundamental physics knowledge, since calibrations must still be derived against synthetic models.
-These ``fit metrics'' (line indices) are similar to distance functions or a figure of merit.
-
-These figures of merit do make a lot of sense, and one can see why they are used, but given that the raw data product from synthetic models from stars is \emph{actually a spectrum} that has been synthesized from a complicated but \emph{physical} model of a star, instead of going through the intermediary step of a figure of merit, wouldn't it be better to relate the two through an actual \emph{likelihood function}, a physical description of the process that generates the data?
-In this case we use a Gaussian process to absorb the computational artifacts/inaccuracy that is a result from a \emph{imperfect} (but still extraordinarily good) stellar model.
-This allows us to couch the question of what parameters does the star have in a truly \emph{probabilistic} framework, meaning, we can calculate a posterior probability function of the parameters that is grounded in our physical understanding of what a star looks like.
-This is one of the great things about a \emph{forward-model}.
-
-The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
+\clearpage
+
+\section{Discussion} \label{sec:discussion}
+
+\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
+be tough to do concisely, but here's what I think the framework should be: (1) a statement of the 
+problem (i.e., that we want to extract physical information from a spectrum without bias or 
+systematic error, but the models we have are not perfect); (2) a [brief] reminder of the standard 
+approaches to skirt this issue; (3) a summary of our approach; (4) a statement that we demonstrated 
+the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
+
+\comm{Next, more specifically highlight why we developed this framework, and what its 
+role should be in general spectroscopic inference.  I think its worth highlighting several key 
+advantages: (1) you are casting the standard data selection/weighting/masking procedure in an 
+internally consistent, Bayesian format that appropriately preserves and propagates information 
+content even in the face of imperfect models (along the lines of what you've written in the 
+commented-out paragraphs below, but more concise/direct); (2) by doing that, you allow the user to 
+actually exploit their full datasets, which should then reflect a more appropriate inference on the 
+parameters of interest without concern that some bad or less-informative features will screw them 
+up (sort of like what I commented out above, but less preachy) -- its in here 
+that we should suggest that this framework could be incorporated as a backend to standard methods, 
+where it essentially is employed to define an appropriate likelihood; and (3) it creates an 
+opportunity to develop quantitative feedback to improve the models, or ultimately to 
+pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
+since you will go into depth on it below). (est.~2 paragraphs)} \\
+
+\ctext{The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
 When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
@@
 If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
-However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
-%\end{rant}
-
-The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.}
+
+\ctext{The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
 Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
@@
 The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.}
+
+\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
+see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
+local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
+what you've already written (commented out below) is in good shape, but you might decide to change 
+the format/organization given the suggested update in structure for this section.  Go out on a 
+positive note: the mechanics already exist, and this tool can be used to build better models!}
+
+\ctext{The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
@@
 The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
-Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
-
-Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
+Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.}
+
+\ctext{Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
 Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
@@
 In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
-This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
-
-Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
+This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.}
+
+\ctext{Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
 Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
 This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
-
-\todo{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.}
+
+\ctext{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
 This will tell us about the frequency and distribution of spectral modelling errors.
@@
 
-\section{Tables}
-
-\begin{deluxetable}{ll}
-\tablecaption{Nomenclature used in this document}
-\tablehead{
-  \colhead{Symbol} & \colhead{Description}
-}
-\startdata
-$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
-$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-  & are applied during ``post processing'' of the synthetic spectrum\\
-$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
-
-$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
- & a single spectral order $i$ \\
-$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
-  & a multi-order echelle spectrum with $N$ spectral orders  \\
- $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
- & except those for order $2$ \\
-$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
-$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
- & a multi-order echelle spectrum with $N$ spectral orders  \\
- $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
- & except those for order $2$ \\
-
-$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
-$\vT$ & the collection of all stellar parameters \\
-$\vD$ & data spectrum \\
-$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
-$\vR$ & $\vD - \vM $, residuals  \\
-$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
-$\vC_{ij}$ & an individual element in the covariance matrix\\
-$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-
-$\Kglobal$ & global covariance kernel\\
-$\Klocal$ & local covariance kernel\\
-$w$ & Hanning taper for kernels \\
-\enddata
-\end{deluxetable}
+%\section{Tables}
+%
+%\begin{deluxetable}{ll}
+%\tablecaption{Nomenclature used in this document}
+%\tablehead{
+%  \colhead{Symbol} & \colhead{Description}
+%}
+%\startdata
+%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
+%$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
+%$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
+%$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
+%  & are applied during ``post processing'' of the synthetic spectrum\\
+%$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
+%
+%
+%$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
+% & a single spectral order $i$ \\
+%$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
+%  & a multi-order echelle spectrum with $N$ spectral orders  \\
+% $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
+% & except those for order $2$ \\
+%$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
+%$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
+% & a multi-order echelle spectrum with $N$ spectral orders  \\
+% $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
+% & except those for order $2$ \\
+%
+%$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
+%$\vT$ & the collection of all stellar parameters \\
+%$\vD$ & data spectrum \\
+%$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
+%$\vR$ & $\vD - \vM $, residuals  \\
+%$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
+%$\vC_{ij}$ & an individual element in the covariance matrix\\
+%$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
+%
+%$\Kglobal$ & global covariance kernel\\
+%$\Klocal$ & local covariance kernel\\
+%$w$ & Hanning taper for kernels \\
+%\enddata
+%\end{deluxetable}
 
@@
 
-This version of the paper was generated
- from a git repository available at \url{http://github.com/iancze/StellarSpectra/}
- with git hash \texttt{\githash\,(\gitdate)}.
 \end{document}

### 2014-08-23 notebooks/paper/generate_Gl51_residual_draws.py
new PHOENIX run

@@ -1,0 +1,38 @@
+import numpy as np
+from StellarSpectra.model import Model
+from StellarSpectra.spectrum import DataSpectrum
+from StellarSpectra.grid_tools import SPEX, HDF5Interface
+from StellarSpectra import utils
+
+myDataSpectrum = DataSpectrum.open("../../data/Gl51/Gl51RA.hdf5", orders=np.array([0]))
+myInstrument = SPEX()
+myHDF5Interface = HDF5Interface("../../libraries/PHOENIX_SPEX_M.hdf5")
+
+#Load a model using the JSON file
+#Taken from:
+# /home/ian/Grad/Research/Disks/StellarSpectra/output/Gl51/PHOENIX/RA/region/logg/4_8sig/
+myModel = Model.from_json("Gl51_model0_final.json", myDataSpectrum, myInstrument, myHDF5Interface)
+
+myOrderModel = myModel.OrderModels[0]
+model_flux = myOrderModel.get_spectrum()
+
+spec = myModel.get_data()
+wl = spec.wls[0]
+fl = spec.fls[0]
+
+model_fl = myOrderModel.get_spectrum()
+residuals = fl - model_fl
+
+mask = spec.masks[0]
+cov = myModel.OrderModels[0].get_Cov().todense()
+
+np.save("Gl51_covariance_matrix.npy", cov)
+import sys
+sys.exit()
+
+
+cov = np.load("Gl51_covariance_matrix.npy")
+
+draws = utils.random_draws(cov, num=200)
+
+np.save("Gl51_residual_draws.npy", draws)

### 2014-08-24 tex/paper.tex
new Gl run
new PHOENIX run
Working parallel prototype

@@
 the radial velocity.  In this specific case, $\sigma_v = 2.9$\,km s$^{-1}$, $v\sin{i} = 5$\,km 
-s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{sec:examples}.x.
+s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{subsec:wasp}.
 ({\it bottom}) A segment of a raw, full-resolution model spectrum and its post-processed equivalent 
@@
     The spectrum at the overlap of two echelle orders (23 \& 24).
-\textbf{Panel 1}: The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders.
-\textbf{Panel 2}: To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model.
-\textbf{Panels 3 \& 4}: In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be (panel 3), but instead we multiply the model by the polynomials to recover the original discrepancy between orders (panel 4).
+({\it panel 1}) The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders.
+({\it panel 2}) To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model.
+In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be ({\it panel 3}), but instead we multiply the model by the polynomials to recover the original discrepancy between orders ( {\it panel 4}).
 This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
@@
   \includegraphics{class0_autocorrelation.pdf}
-  \caption{\textbf{Top}: A typical model fit to a data spectrum, with
-  parameters drawn from the posterior distribution, and the residuals that
-  result. \textbf{Left} The low-amplitude, mildly covariant residuals enlarged
+  \caption{ ({\it top}) A typical model fit using 
+  parameters drawn from the posterior distribution and the residuals that
+  result. ({\it middle}) The low-amplitude, mildly covariant residuals in the gray band enlarged
   to show the mildly covariant structure produced by slight mismatch between
-  the data and model spectra.   \textbf{Right} The autocorrelation of the
+  the data and model spectra. ({\it bottom}) The autocorrelation of the
 residual sequence shown at left. Notice that there is significant correlation
-for offsets of $\lesssim 5$ pixels.}
+for offsets of $\lesssim 8$ pixels.}
 \label{fig:class0}
@@
 \includegraphics{figs/matrix_compilation.pdf}
-\caption{\textbf{Top}: A typical model fit and residual spectrum with a demonstrative region shaded in grey, showing the scale and structure of the residuals.
-\textbf{Left column}: A small inset of the covariance matrix corresponding to the gray region.
+\caption{ ({\it top}) A typical model fit and residual spectrum (zoomed).
+An illustrative region is shaded in grey, showing the scale and structure of the residuals as an example for the following panels.
+({ \it left column}) A small inset of the covariance matrix corresponding to the gray region.
 Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix.
-\textbf{Right column}: Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
+({\it right column}) Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
 At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region.
-\textbf{First row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) does a poor job of reproducing both the scale and structure of a typical residual spectrum.
-\textbf{Second row}: The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
-\textbf{Third row}: Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
+{\it first row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) poorly reproduces both the scale and structure of a typical residual spectrum.
+({\it second row}) The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
+({\it third row}) Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
 Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines.
 The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit.
-Stated differently, random draws from the likelihood function can now produce behavior that looks like the residuals, whereas before with only white noise it could not.
+Stated differently, random draws from the likelihood function now behave like the residuals, whereas white noise cannot.
 All parameters of the covariance kernels are determined self-consistently along with the stellar parameters.
@@
 
-Local kernels are instantiated by the following procedure:  An ``average''
+Local kernels are instantiated by the following procedure:  an ``average''
 residual spectrum is created by combining ~500 spectra residuals stored from a
-burn-in period using only global covariance kernels (thinned to account for
-autocorrelation of the Markov chain). Then, using an iterative thresholding
-criterion similar in procedure to ``sigma clipping,''\footnote{Although the
-  procedure of identifying outliers is similar to sigma clipping, the inference
-  on the stellar parameters is emphatically \emph{not} the same as sigma
-clipping, since we now determine the weights of these spectral outliers inside
-of a probabilistic framework, rather than directly setting their weights to
-zero as in sigma clipping.} we identify the largest outlier pixel not already
-covered by a local kernel, instantiate a local kernel at its location, and
-repeat until there are no pixels above a certain threshold left.\footnote{There
-  is no practical reason to delete local kernels once instantiated.  If the
-  parameters have changed such that a given local kernel is no longer required,
-  that kernel amplitude will be driven towards zero and represent a negligible
-  contribution to $\vC$; in effect, the model will act as if the kernel were
-deleted automatically.} Alternative local kernel instantiation schemes, such as
-re-evaluating the location of kernels with each Gibbs sampler iteration, have
-yielded similar results to the thresholding approach. However, this approach
-most consistently converges and only has a modest computational overhead. We
-choose four times the standard deviation of the average residual spectrum as
-our threshold. Lower thresholds (e.g., 3$\sigma$ and 3.5$\sigma$) result in the
-instantiation of more kernels, which in turn slightly reduces the amplitude of
-the global kernel. Taken to the logical conclusion, if the threshold was set
-low enough then there would be a local kernel for each spectral line and no
-global kernel would be required. We find that the final stellar posterior is
-relatively insensitive to the choice of threshold level as long as it is set
-low enough to capture the most egregious spectral outliers (e.g., 4$\sigma$).
-Once all of the local kernels have been instantiated, the Gibbs sampler is run
-for another period of burn-in.
+burn-in period using only global covariance kernels (prior to storage the
+Markov chain is thinned to account for autocorrelation of the posterior
+samples). Then, using an iterative thresholding criterion similar in procedure
+to ``sigma clipping,''\footnote{Although the procedure of identifying outliers
+  is similar to sigma clipping, the inference on the stellar parameters is
+  emphatically \emph{not} the same as sigma clipping, since we now determine
+  the weights of these spectral outliers inside of a probabilistic framework,
+  rather than directly setting their weights to zero as in sigma clipping.} we
+  identify the largest outlier pixel not already covered by a local kernel,
+  instantiate a local kernel at its location, and repeat until no residual
+  pixels above a certain threshold remain. Alternative local kernel
+  instantiation schemes, such as re-evaluating the location of kernels with
+  each Gibbs sampler iteration, yield similar results to the thresholding
+  approach. However, we found that this thresholding approach most consistently
+  converged with the least computational overhead. We explored the behavior of
+  setting different threshold levels, and decided to use four times the
+  standard deviation of the average residual spectrum. Lower thresholds (e.g.,
+  3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which
+  in turn slightly reduces the amplitude of the global kernel. Taken to the
+  logical conclusion, if the threshold was set low enough then there would be a
+  local kernel for each spectral line and no global kernel would be required.
+  We find that the final stellar posterior is relatively insensitive to the
+  choice of threshold level as long as it is set low enough to capture the most
+  egregious spectral outliers (e.g., 4$\sigma$).  Once all of the local kernels
+  have been instantiated, the Gibbs sampler is run for another period of
+  burn-in.\footnote{There is no practical reason to delete local kernels once
+    instantiated.  If the parameters have changed such that a given local
+    kernel is no longer required, that kernel amplitude will be driven towards
+    zero and represent a negligible contribution to $\vC$; in effect, the model
+    will act as if the kernel were deleted automatically.}
 
@@
 star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
-\citep{cushing05,rayner09}.  In this case we adopt the standard {\sc BT-Settl} modification of the 
-{\sc Phoenix} library \citep{allard12}.
+\citep{cushing05,rayner09}.  In this case we also use the {\sc Phoenix} library, since it forms a complete grid for late spectral types.
 
@@
 full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
-standard techniques in the TRES pipeline (cf.,~\citealt{mink11}; see \citealt{torres12} for more 
+standard techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for more 
 specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
@@
 stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
-evolution models in the H-R diagram \citep{joshi09}.  Only parameters inferences from the latter 
+evolution models in the H-R diagram \citep{joshi09}.  Only parameter inferences from the latter 
 assumption were reported.  The {\tt MOOG} analysis employed \citet{kurucz93} atmospheres to 
@@
 library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
-i = 3.9\pm0.5$\,km s$^{-1}$.  \hili{It is worth noting that these parameter inferences represent an 
+i = 3.9\pm0.5$\,km s$^{-1}$.  It is worth noting that these parameter inferences represent an 
 average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
-the systematic spread between the best-fit parameter values.} \comm{This is something we need to 
+the systematic spread between the best-fit parameter values. \comm{This is something we need to 
 discuss more with Dave and Willie.}
 
-Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
-it is most relevant to compare the \citeauthor{torres12}~{\tt SPC} result with our framework.  
-Using the same TRES spectral orders (for only a single WASP-14 observation, as noted above), the 
-{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer 
-parameters that are entirely consistent with the {\tt SPC} approach: $T_{\rm eff} = 6517\pm13$\,K, 
-$[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and $v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~X displays 
-some relevant marginal projections of the posterior-space, and Table~X compiles our estimates of 
-the peaks of the marginal posteriors and associated confidence intervals for all parameters 
-employed in the modeling.  Figure~Y demonstrates the overall quality of this approach in 
-reproducing the WASP-14 spectrum and, most significantly, the structure of the associated 
-residuals.  Figure~Z compares the inferred parameters with the various determinations made by 
-\citet{torres12}.
+Given the decision to employ a spectral library to generate models
+(Sect.~\ref{subsec:synthetic}), it is most relevant to compare the
+\citeauthor{torres12}~{\tt SPC} result with our framework.  Using the same TRES
+spectral orders (for only a single WASP-14 observation, as noted above), the
+{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log
+g$, we infer parameters that are entirely consistent with the {\tt SPC}
+approach: $T_{\rm eff} = 6517\pm13$\,K, $[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and
+$v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~\ref{fig:Kurucz_posterior}
+displays some relevant marginal projections of the posterior-space, and
+Table~\ref{table:Kurucz} compiles our estimates of the peaks of the marginal
+posteriors and associated confidence intervals for all parameters employed in
+the modeling.  Figure~\ref{fig:Kurucz_residuals} demonstrates the overall
+quality of this approach in reproducing the WASP-14 spectrum and, most
+significantly, the structure of the associated residuals.
+Figure~\ref{fig:metacomparison} compares the inferred parameters with the
+various determinations made by \citet{torres12}.
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+  \caption{The posterior probability function of the stellar parameters for WASP-14 as explored by the MCMC Gibbs sampler.
+These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
+Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+}
+\label{fig:Kurucz_posterior}
+\end{center}
+\end{figure}
 
@@
 \includegraphics{figs/residuals_Kurucz_logg.pdf}
-\caption{Kurucz WASP-14. One, two, and three sigma contours formed from 200 random draws from the covariance matrix.}
-\label{fig:residuals_Kurucz}
+\caption{A sample of the WASP-14 data set overlaid with a model spectrum produced by parameters drawn from the posterior probability distribution (fixing $\log_{10} g = 4.29$ dex) and using the {\sc CfA/Kurucz} synthetic library.
+The residuals are shown zoomed along with the three-sigma contours of random draws from the covariance matrix, after Figure~\ref{fig:matrix}.}
+\label{fig:Kurucz_residuals}
 \end{center}
 \end{figure*}
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
-  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
-Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:stellar_posterior}
-\end{center}
-\end{figure}
 
@@
 \cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6517.1 & 12.5 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
+$T_{\rm eff}$ & 6517 & 13 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex\tablenotemark{a}} \\
 $\Z$ & -0.27 & 0.01 dex \\
 $v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.620 & 0.016 $\kms$ \\
+$v_z$ & -4.62 & 0.02 $\kms$ \\
 $\log_{10} \Omega$ & -12.7181 & 0.0004 \\
@@
 $\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
-$\cc{22}{1}$ & -0.0142 & 0.0015\\
-$\cc{22}{2}$ & -0.0145 & 0.0017\\
-$\cc{22}{3}$ & -0.0031 & 0.0011\\
+$\cc{22}{1}$ & -0.014 & 0.002\\
+$\cc{22}{2}$ & -0.014 & 0.002\\
+$\cc{22}{3}$ & -0.003 & 0.001\\
 \\
 $b_{22}$ & 1.04 & 0.02 \\
-$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.020 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
-$l_{22}$ & 7.54 & 0.77 $\kms$ \\
+$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
+$l_{22}$ & 7.5 & 0.8 $\kms$ \\
 
@@
 $\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
-$\cc{23}{1}$ & -0.0133 & 0.0016 \\
-$\cc{23}{2}$ & -0.0247 & 0.0014 \\
-$\cc{23}{3}$ &  0.0012 &  0.0013\\
+$\cc{23}{1}$ & -0.013 & 0.002 \\
+$\cc{23}{2}$ & -0.025 & 0.001 \\
+$\cc{23}{3}$ &  0.001 &  0.001\\
 \\
 $b_{23}$ & 1.01 & 0.02 \\
-$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.015 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
-$l_{23}$ & 6.52 & 0.50 \\
+$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{23}$ & 6.5 & 0.5 $\kms$\\
 
 \cutinhead{Order 24 nuisance parameters}
-$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00} \\
-$\cc{24}{1}$ & -0.0103 & 0.0016 \\
-$\cc{24}{2}$ & -0.0112 & 0.0013\\
-$\cc{24}{3}$ &  0.0002 & 0.0012\\
+$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00\tablenotemark{a}} \\
+$\cc{24}{1}$ & -0.010 & 0.002 \\
+$\cc{24}{2}$ & -0.011 & 0.001\\
+$\cc{24}{3}$ &  0.000 & 0.001\\
 \\
 $b_{24}$ & 1.02 & 0.02 \\
-$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}$\\
-$l_{24}$ & 7.33 & 0.59 $\kms$ \\
+$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{24}$ & 7.3 & 0.6 $\kms$ \\
 \enddata
+\tablenotetext{a}{parameter value fixed}
 \end{deluxetable}
@@
 
-\comm{Ok, this is where you went off the rails a little bit in the core-dump.  I frankly do not 
-want to see Table 2 in this paper, or any reference to specific model or test identifications.  It 
-is too confusing, and it obscures the key points we want to make.  I appreciate your desire to 
-demonstrate the complexity of the model sequentially, but it does not add anything useful/tangible 
-here: it is already done very well in Section 2.  I feel like up to here we have 
-established ``reproducibility" well.  The remainder of this section should be solely about 
-systematics.  First, I want a paragraph breaking down the origins of the uncertainties (see the 
-paragraph I commented out below...I don't like it, but it should give you an idea about what I 
-mean.  If I interpret your Google tables correctly, you're getting about half the uncertainty from 
-Poisson noise, and the other half from the global covariance kernel.  This is where you need to be 
-able to defend the seemingly small uncertainties you're finding, even with a non-trivial covariance 
-matrix.  Is it really true that a model with 0.03\,dex lower metallicity cannot describe the data 
-as well?  Why?  [This is not to say I don't believe it, but it {\it will} raise questions...]}
-
-Given the apparent differences in the quoted precisions, it is beneficial to review what factors 
+To understand the apparent differences in the quoted precisions, it is beneficial to review what factors 
 into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
@@
 are generated from the model; in this paradigm, the only way to introduce deviation is through 
-white noise.  Of course, that implicit assumption is invalid, since just a few simple parameters 
+white noise (see Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is invalid, since just a few simple parameters 
 ($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
@@
 systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
-posteriors.
-
-\comm{Next, you want to address the observational systematics issue.  Willie (and basically 
-everyone else) add empirically-estimated systematic terms in quadrature that make some pretty 
-unjustified assumptions: namely that they are Gaussian and independent in each parameter.  If I 
-understant Willie's paper correctly, there are 4 TRES observations of WASP-14.  If that is true, I 
-think we need to fit each of them independently.  I'd like to know what ``systematic" scatter you 
-would derive if you wanted to employ the standard approach.  My gut tells me that it's important to 
-simultaneously model all 4 spectra in orders 22-24, but if you think that's not 
-computationally tractable, we can discuss more.}
-\ctext{It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
-Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.}
+posteriors. For synthetic models which are less accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend to dominate the error budget.
+
+In addition to statistical and systematic error, one might expect that systematic instrumental error might also contribute to the parameter uncertainties.
+Commonly, one might estimate this error by fitting spectra taken by different instruments or at different times and then averaging the results.
+Additionally, one may add in an additional uncertainty ``floor'' that maps to the empirical scatter of typical spectroscopic fits.
+Such procedures generally make the assumption that this additional error contribution is Gaussian and independent in each parameter of interest.
+Our hierarchical Bayesian model avoids such a heuristic by
+parameterizing the quality of fit self-consistently: because the parameter uncertainty is derived through the model itself, it preserves correlations between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
+Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
+However, for well-calibrated data sets from the same instrument any ill-effects are generally small.
+To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim10$ K and $\lesssim 0.001$ dex in $\Z$.
+When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.
 
@@
 
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[draft, width=0.5\textwidth,height=2in]{figs/PHOENIX_triangle.pdf}
+\caption{ {\sc Phoenix} WASP-14 marginal posteriors.}
+\label{fig:PHOENIX_posterior}
+\end{center}
+\end{figure}
+
 \begin{figure*}[!htb]
@@
 \includegraphics{figs/residuals_PHOENIX_logg.pdf}
-\caption{Kurucz WASP-14.}
+\caption{ {\sc Phoenix} WASP-14 model and residuals.}
 \label{fig:residuals_PHOENIX}
@@
 
-\ctext{While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
-For demonstrative purposes, for the next test (WP2) we use the common technique of ``sigma-clipping,'' whereby we manually mask any region of the spectrum which has a more than a 3 sigma residual near the best fit parameters, refit, and then proceed until there are no more outliers.
-Masking out ``outlier'' spectral lines has the effect of actually changing the final stellar estimate significantly, while also reducing the uncertainty in the stellar parameters.
-However, several questions arise about the quality of the inference on the stellar parameters.
+While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
 Because sigma-clipping is a procedure, it is impossible to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
 For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
-Clipping these lines would discard important information and lead to an incorrect inference of stellar parameters.}
-
-\ctext{These results motivate the development of a more sophisticated likelihood function which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
-For the next tests, (WK2 and WP3), we introduce the global covariance kernel (Eqn.~\ref{eqn:global}) to the fit.
-We can see that the amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum.
-Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the tyranny of too many lines, which cause the unreasonably low uncertainty of the previous tests.
-Therefore, including the global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance, as evidenced by the Kurucz ($10^{-14.6}$) v.
-PHOENIX ($10^{-14.2}$) global covariance amplitudes.
-\todo{Figure out a reasonable way to present nuisance parameter values}.
-It is important to note that this ability to parameterize the quality of fit \emph{self-consistently} is unique to a hierarchical Bayesian model, contrary to systematic error estimates that are added in post-facto (I probably don't want to name names, but e.g., \citealt{torres12, buchhave12}).
-Because the parameter uncertainty is derived through the model itself, it \emph{preserves correlations} between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.}
-
-\ctext{For the next tests, (WPX), we introduce the local covariance kernels to track outlier spectral lines.
+\todo{Might be unnecessary from all that we wrote in section 3?}
+
+This behaviour motivates the global and local kernels in a more sophisticated likelihood function, which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum.
+Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the outlier lines.
+The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance, as evidenced by the global covariance amplitudes between the {\sc CfA/Kurucz} and {\sc Phoenix} libraries (Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX}).
+
+Local covariance kernels track outlier spectral lines.
 These kernels can map out the regions of the spectrum which are outliers.
@@
 The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
-See Figure~\ref{fig:regions} for a cool visualization of what it looks like.}
-
-\ctext{In addition to the stark contrast of stellar posteriors derived using only white Poisson noise versus those with the covariance kernels, the results in Table~\ref{table:tests} also reflect the systematic bias that is a function of which spectral library is used as a backend.
-Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.}
+See Figure~\ref{fig:residuals_PHOENIX} for a demonstration of how the local kernels identify and track spectral line outliers.
+
+Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveal that there is a systematic offset depending on which spectral library is used as a backend.
+Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.
 
@@
 for a useful visualization), designed to probe the spectral curvature induced by H$_2$O absorption 
-bands.  With reference to a custom grid of {\sc BT-Settl} spectra, \citet{rojas-ayala12} find that 
+bands.  With reference to a custom grid of {\sc BT-Settl} spectra\footnote{From the web simulator at \url{http://phoenix.ens-lyon.fr/simulator/index.faces}}, \citet{rojas-ayala12} find that 
 these indices are consistent with $T_{\rm eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$.
+
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:Gl51}Gl51 {\sc Phoenix} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 2998  & 17 K \\
+$\log_{10} g$ & \multicolumn{2}{c}{5.0 dex} \\
+$\Z$ & -0.01 & 0.04 dex \\
+$v \sin i$ & 5.68 & 4.15 $\kms$\\
+$v_z$ & 12.1 & 1.3 $\kms$ \\
+$\log_{10} \Omega$ & -19.628  & 0.003 \\
+\cutinhead{Nuisance parameters}
+$\log_{10} c_0$ & \multicolumn{2}{c}{0.0} \\
+$c_1$ & -0.011 & 0.002 \\
+$c_2$ & 0.005 & 0.002 \\
+$c_3$ & -0.002 & 0.001 \\
+$b$ & 1.56 & 0.14 \\
+$\log_{10} a_{\rm g} $ & -15.76 &  0.04 \\
+$l$ & 116 & 20 $\kms$ \\
+\cutinhead{Local kernel parameters}
+$\log_{10} a$ & -14.75 & 0.18 \\
+$\mu$ & 22055.9 & 0.8 \AA\\
+$\sigma$ & 122 & 16 $\kms$ \\
+\hline\\
+$\log_{10}a$ & -14.83 & 0.19 \\
+$\mu$ & 22083.1 & 0.9 \AA\\
+$\sigma$ & 120 & 17 $\kms$  \\
+\hline\\
+$\log_{10}a$ & -14.81 & 0.27 \\
+$\mu$ & 22609.2 & 1.3 \AA\\
+$\sigma$ & 85 & 26 $\kms$  \\
+\hline\\
+$\log_{10}a$ & -14.73 & 0.24 \\
+$\mu$ & 22625.2 & 1.1 \AA\\
+$\sigma$ & 94 & 24 $\kms$ \\
+\hline\\
+$\log_{10}a$ & -14.74 & 0.20 \\
+$\mu$ & 22652.2 & 0.8 \AA\\
+$\sigma$ & 99 & 19 $\kms$ \\
+\hline\\
+$\log_{10}a$ & -14.81 & 0.31 \\
+$\mu$ & 23380.1 & 1.0 \AA\\
+$\sigma$ & 73 & 27 $\kms$ \\
+\enddata
+\end{deluxetable}
 
@@
 
-\comm{Commentary on the necessity and utility of local covariance kernels here, with discussion of why Na and Ca 
-lines are a pain.  It is also worth commenting on the potential for flexibly extending the shape of 
+\comm{Commentary on the necessity and utility of local covariance kernels here, with discussion of why Na and Ca lines are a pain.
+It is also worth commenting on the potential for flexibly extending the shape of
 the kernel to mimic molecular bandheads, if one wanted to try and fit say the TiO bands.}
-\ctext{In order to calibrate these indices to stellar parameters, \citet{rojas-ayala12} determine these indices on a custom grid of PHOENIX BT-Settl models obtained from the PHOENIX web simulator\footnote{\url{http://phoenix.ens-lyon.fr/simulator/index.faces}}.
+
 They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
 \citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
-When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that}
+When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{report parameters}.
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/Gl51_triangle.pdf}
+  \caption{Gl51 stellar posterior.}
+\label{fig:gl_posterior}
+\end{center}
+\end{figure}
 
@@
 \begin{center}
-  \includegraphics[draft,width=0.9\textwidth,height=3in]{figs/Gl51_residuals.pdf}
-  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
-Plotted with the increased variance envelope over the residuals.}
+  \includegraphics{figs/residuals_Gl51_logg.pdf}
+  \caption{The Gl51 spectrum and residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.}
 \label{fig:regions}
@@
 
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics[draft,width=0.45\textwidth,height=3in]{figs/Gl51_posterior.pdf}
-  \caption{The spectrum residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.
-Plotted with the increased variance envelope over the residuals.}
-\label{fig:gl_posterior}
-\end{center}
-\end{figure}
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:PHOENIX}Gl51 {\sc Phoenix} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ &  &  K \\
-$\log_{10} g$ & \multicolumn{2}{c}{5.0 dex} \\
-$\Z$ &  &  dex \\
-$v \sin i$ &  &  $\kms$\\
-$v_z$ &  &  $\kms$ \\
-$\log_{10} \Omega$ &  & \\
-\cutinhead{Nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b$ & & \\
-$a_{\rm g} $ & & \\
-$l$ & & \\
-\cutinhead{Local kernel parameters}
-& & \\
-& & \\
-& & \\
-\hline\\
-& & \\
-& & \\
-& & \\
-\hline\\
-& & \\
-& & \\
-& & \\
-\enddata
-\end{deluxetable}
-
 \comm{A simple two-sentence (or so) wrap up on the future promise of this approach for late-type 
 stars and their complex spectra.}
+
+Complicated molecular structure is poorly known from models.
+
+\clearpage
+
+\section{Discussion} \label{sec:discussion}
+
+\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
+be tough to do concisely, but here's what I think the framework should be: (A1) a statement of the 
+problem (i.e., that we want to extract physical information from a spectrum without bias or 
+systematic error, but the models we have are not perfect); (A2) a [brief] reminder of the standard 
+approaches to skirt this issue; (A3) a summary of our approach; (A4) a statement that we demonstrated 
+the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
+
+\comm{Next, more specifically highlight why we developed this framework, and what its 
+role should be in general spectroscopic inference.  I think its worth highlighting several key 
+advantages: (B1) you are casting the standard data selection/weighting/masking procedure in an 
+internally consistent, Bayesian format that appropriately preserves and propagates information 
+content even in the face of imperfect models (along the lines of what you've written in the 
+commented-out paragraphs below, but more concise/direct); (B2) (some deleted text could be useful here) by doing that, you allow the user to 
+actually exploit their full datasets, which should then reflect a more appropriate inference on the 
+parameters of interest without concern that some bad or less-informative features will screw them 
+up (sort of like what I commented out above, but less preachy) -- its in here 
+that we should suggest that this framework could be incorporated as a backend to standard methods, 
+where it essentially is employed to define an appropriate likelihood; and (B3) it creates an 
+opportunity to develop quantitative feedback to improve the models, or ultimately to 
+pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
+since you will go into depth on it below). (est.~2 paragraphs)} \\
+
+A2: The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
+Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
+The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+
+B1: The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
+When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
+When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters.
+This hierarchical model also works within a single spectrum when fitting more than one spectral order.
+If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
+However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
+
+\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
+see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
+local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
+what you've already written (commented out below) is in good shape, but you might decide to change 
+the format/organization given the suggested update in structure for this section.  Go out on a 
+positive note: the mechanics already exist, and this tool can be used to build better models!}
+
+The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
+By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
+After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
+The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
+Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
+
+B3: Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
+Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
+Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
+In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
+This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
+
+B3: Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
+Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
+This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
+
+Infer the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
+This will tell us about the frequency and distribution of spectral modelling errors.
+Or we could just speculate about it.
+Could also cross correlate with a catalogue of lines to determine which.
 
@@
   \caption{A collection of spectral lines which have imperfect model fits.
-    \textbf{First panel} The majority of spectral lines ($\gtrsim 60$\%) will
+    ( {\it first panel}) The majority of spectral lines ($\gtrsim 60$\%) will
     have minor differences in strength between data and model spectrum, which
     produce low-amplitude correlations in the residuals on the length scale of
-    the width of a typical spectral line.  \textbf{Second panel}: Sometimes
+    the width of a typical spectral line.  ( {\it second panel}): Sometimes
     ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
     this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  \textbf{Third panel}: Sometimes ($\lesssim
+    patch of negative residuals.  ( {\it third panel}): Sometimes ($\lesssim
     5$\% of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals. \textbf{Fourth panel}: If the line
+    correlated patch of positive residuals. ({\it fourth panel}): If the line
     strengths are substantially discrepant ($\lesssim 10$\% of all lines),
-    there will be many correlated residuals of moderate amplitude.  The
-    difficulty with class III lines is that for any specific line, there might
+    there will be many correlated residuals of moderate amplitude. For any specific line, there might
     exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
@@
 \end{figure}
-
-\clearpage
-
-\section{Discussion} \label{sec:discussion}
-
-\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
-be tough to do concisely, but here's what I think the framework should be: (1) a statement of the 
-problem (i.e., that we want to extract physical information from a spectrum without bias or 
-systematic error, but the models we have are not perfect); (2) a [brief] reminder of the standard 
-approaches to skirt this issue; (3) a summary of our approach; (4) a statement that we demonstrated 
-the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
-
-\comm{Next, more specifically highlight why we developed this framework, and what its 
-role should be in general spectroscopic inference.  I think its worth highlighting several key 
-advantages: (1) you are casting the standard data selection/weighting/masking procedure in an 
-internally consistent, Bayesian format that appropriately preserves and propagates information 
-content even in the face of imperfect models (along the lines of what you've written in the 
-commented-out paragraphs below, but more concise/direct); (2) by doing that, you allow the user to 
-actually exploit their full datasets, which should then reflect a more appropriate inference on the 
-parameters of interest without concern that some bad or less-informative features will screw them 
-up (sort of like what I commented out above, but less preachy) -- its in here 
-that we should suggest that this framework could be incorporated as a backend to standard methods, 
-where it essentially is employed to define an appropriate likelihood; and (3) it creates an 
-opportunity to develop quantitative feedback to improve the models, or ultimately to 
-pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
-since you will go into depth on it below). (est.~2 paragraphs)} \\
-
-\ctext{The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
-When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
-When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters.
-This hierarchical model also works within a single spectrum when fitting more than one spectral order.
-If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
-However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.}
-
-\ctext{The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.}
-
-\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
-see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
-local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
-what you've already written (commented out below) is in good shape, but you might decide to change 
-the format/organization given the suggested update in structure for this section.  Go out on a 
-positive note: the mechanics already exist, and this tool can be used to build better models!}
-
-\ctext{The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
-By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
-After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
-The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
-Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.}
-
-\ctext{Linking the covariance matrices of stars could be done serially, where the aggregate covariance matrix of all previous fits is used as a prior for the current iteration.
-Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
-Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
-Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
-In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
-This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.}
-
-\ctext{Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
-Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
-This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.}
-
-\ctext{We can (should?) add an additional level to the hierarchy of hyperparameters and add parameters to describe the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
-This will tell us about the frequency and distribution of spectral modelling errors.
-Or we could just speculate about it.}
 

### 2014-08-24 parallel.py
Working parallel prototype

@@ -1,0 +1,179 @@
+#simple parallel implementatino
+
+from multiprocessing import Process, Pipe, cpu_count
+import os
+import numpy as np
+
+def info(title):
+    print(title)
+    print('module name:', __name__)
+    if hasattr(os, 'getppid'):  # only available on Unix
+        print('parent process:', os.getppid())
+    print('process id:', os.getpid())
+
+
+#Eventually we will want a separate OrderModel started on each process. The only thing we send back and forth is
+# parameter information.
+
+#after the models have been created, now fork
+#
+# class Model:
+#     def __init__(self):
+#         self.param = 0
+#
+#     def square_param(self):
+#         return self.param**2
+#
+#     def set_param(self, param):
+#         self.param = param
+#
+#     def get_communication(self, conn):
+#         info("Model")
+#         print("Communication was", conn.recv())
+#         conn.send("Hello back")
+
+class Model:
+    def __init__(self, DataSpectrum, npoly=4, debug=False):
+        self.DataSpectrum = DataSpectrum
+        self.lnprob = -np.inf
+
+        self.func_dict = {"INIT": self.initialize,
+                          "UPDATE": self.update_stellar,
+                          "REVERT": self.revert_stellar,
+                          "IND" : self.independent_sample,
+                          }
+
+    def initialize(self, order_index=0):
+        '''
+        Designed to be called after the other processes have been created.
+        '''
+        print("Initializing model on order {}.".format(order_index))
+
+        #Load the spectrum interpolator
+
+
+        pass
+
+    def update_stellar(self, params):
+        '''
+        Update the stellar flux. Evaluate the likelihood, return.
+        '''
+        print("Updating stellar parameters", params)
+        lnprob = 1
+        return lnprob
+
+    def revert_stellar(self, *args):
+        '''
+        Revert the downsampled stellar flux.
+        '''
+        print("Reverting stellar parameters")
+        pass
+
+    def independent_sample(self, *args):
+        '''
+        Now, do all the sampling that is specific to this order.
+        '''
+
+        #Take the lnprob from the previous iteration
+
+
+        print("Doing indepedent sampling.")
+
+        #Don't return anything to the main process.
+
+    def brain(self, conn):
+        self.conn = conn
+        alive = True
+        while alive:
+            #Keep listening for messages put on the Pipe
+            alive = self.interpret()
[…]

### 2014-08-24 tex/paper.tex
WASP PHOENIX work

@@
 
-While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range and there is no tuning for specific stars like the Sun, there are often a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
-Because sigma-clipping is a procedure, it is impossible to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
+While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range with no specific tuning for solar type stars, there are likely to be a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
+Some routines choose to ``sigma-clip'' the outlier lines and procede with fitting, however with such an approach it is difficult to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
 For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
-\todo{Might be unnecessary from all that we wrote in section 3?}
 
@@
 Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the outlier lines.
-The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance, as evidenced by the global covariance amplitudes between the {\sc CfA/Kurucz} and {\sc Phoenix} libraries (Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX}).
-
-Local covariance kernels track outlier spectral lines.
+
+Using the {\sc Phoenix} models on WASP-14, we infer substantially different posteriors than with using the {\sc CfA/Kurucz} library.
+Figure~\ref{fig:PHOENIX_posterior} displays some of the relevant marginal posteriors, and Table~\ref{table:PHOENIX} compiles the estimates of the peaks of the posterior distribution and associated confidence intervals.
+Figure~\ref{fig:residuals_PHOENIX} demonstrates the quality of this approach and the structure of the associated residuals.
+Compared to the Kurucz residuals in Figure~\ref{fig:residuals_Kurucz}, there are two significant differences.
+First, the amplitude of the global covariance structure is larger.
+The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance.
+Second, local covariance kernels track outlier spectral lines.
 These kernels can map out the regions of the spectrum which are outliers.
-The glory of these local kernels are that the parameters allow to play a role in data driven spectra.
-\emph{A priori} we do not know which regions of the spectrum are improperly modeled.
 The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
-See Figure~\ref{fig:residuals_PHOENIX} for a demonstration of how the local kernels identify and track spectral line outliers.
-
-Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveal that there is a systematic offset depending on which spectral library is used as a backend.
-Depending on your assumptions, there may be systematic differences of 200K or more in $T_{\rm eff}$.
+The locations and amplitudes of these kernels output from the code would be useful information to consider when refining the models of stellar sythenis.
+
+Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveals that there is a systematic offset depending on which spectral library is used as a backend.
+Depending on the choice of spectral library, there may be systematic differences of 200 K or more in $T_{\rm eff}$ and 0.2 dex or more in $\Z$.
 

### 2014-08-25 tex/paper.tex
Seems like it's done

@@
 
-We opt for the computationally expedient approach that employs a library of model spectra, 
-$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
-[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
-{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model 
-spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
-\begin{equation} \label{eqn:interp}
-\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
-\end{equation}
-where we assign the symbol $\leadsto$ as an interpolation operator.  The
-multi-dimensional interpolation in Eq.~\ref{eqn:interp} needs to be performed
-many times, so computational efficiency is critical.  In practice, a simple
-tri-linear interpolation is suitably fast, but introduces an undesirable level
-of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation
-quality can be empirically estimated by performing the operation in
-Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ grid point, and then
-comparing the interpolated spectrum with the corresponding library spectrum.
-After an extensive exploration of such calculations \citep[see
-also][]{husser12}, we concluded that this interpolation technique is found to
-be accurate within a few percent per full-resolution model pixel and $\lesssim
-1$\% once the spectrum is downsampled to the pixels of the detector. Ideally,
-this pre-interpolation could be avoided if the spectral library was computed
-over a refined grid (with a substantial up-front computational investment); but
-for the time being, we can empirically propagate these interpolation
+We opt for the computationally expedient approach that employs a library of
+model spectra, $\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm
+eff}, \,\, \log{g}, \,\, [{\rm Fe/H}]]$.  However, it is worth noting that the
+techniques developed here are applicable to {\it any} ``back-end" that
+generates a model spectrum.  In our adopted approach, the model spectrum for an
+arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
+\begin{equation} \label{eqn:interp} \flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto
+\flam(\vt_{\ast}), \end{equation} where we assign the symbol $\leadsto$ as an
+interpolation operator.  The multi-dimensional interpolation in
+Eq.~\ref{eqn:interp} needs to be performed many times, so computational
+efficiency is critical.  In practice, a simple tri-linear interpolation is
+suitably fast, but introduces an undesirable level of inaccuracy
+(particularly in the [Fe/H] dimension).  The interpolation quality can be
+empirically estimated by performing the operation in Eq.~\ref{eqn:interp}
+across a $\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the
+interpolated spectrum with the corresponding library spectrum. After an
+extensive exploration of such calculations \citep[see also][]{husser12}, we
+concluded that the best combination of speed and accuracy can be achieved by
+pre-computing a {\it refined} spectral library using a cubic spline
+interpolation with a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex,
+0.1\,dex], and then performing tri-linear interpolation over that refined
+grid.  Overall, this interpolation technique is found to be accurate within
+a few percent per full-resolution model pixel and $\lesssim 1$\% once the
+spectrum is downsampled to the pixels of the detector.  Ideally, this
+pre-interpolation could be avoided if the spectral library was computed
+over a refined grid (with a substantial up-front computational investment);
+but for the time being, we can empirically propagate these interpolation
 uncertainties into the likelihood calculations (as will be described in
@@
 
+\begin{deluxetable}{lr@{$\pm$}l} 
+\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+\startdata
+\cutinhead{Stellar parameters}
+$T_{\rm eff}$ & 6517 & 13 K \\
+$\log_{10} g$\tablenotemark{a} & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.27 & 0.01 dex \\
+$v \sin i$ & 4.29 & 0.05 $\kms$\\
+$v_z$ & -4.62 & 0.02 $\kms$ \\
+$\log_{10} \Omega$\tablenotemark{b} & -12.7181 & 0.0004\\
+\cutinhead{Order 22 nuisance parameters}
+$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
+$\cc{22}{1}$ & -0.014 & 0.002\\
+$\cc{22}{2}$ & -0.014 & 0.002\\
+$\cc{22}{3}$ & -0.003 & 0.001\\
+\\
+$b_{22}$ & 1.04 & 0.02 \\
+$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
+$l_{22}$ & 7.5 & 0.8 $\kms$ \\
+
+\cutinhead{Order 23 nuisance parameters}
+$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
+$\cc{23}{1}$ & -0.013 & 0.002 \\
+$\cc{23}{2}$ & -0.025 & 0.001 \\
+$\cc{23}{3}$ &  0.001 &  0.001\\
+\\
+$b_{23}$ & 1.01 & 0.02 \\
+$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{23}$ & 6.5 & 0.5 $\kms$\\
+
+\cutinhead{Order 24 nuisance parameters}
+$\log_{10} \cc{24}{0}$\tablenotemark{a} & \multicolumn{2}{c}{0.00} \\
+$\cc{24}{1}$ & -0.010 & 0.002 \\
+$\cc{24}{2}$ & -0.011 & 0.001\\
+$\cc{24}{3}$ &  0.000 & 0.001\\
+\\
+$b_{24}$ & 1.02 & 0.02 \\
+$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{24}$ & 7.3 & 0.6 $\kms$ \\
+\enddata
+\tablenotetext{a}{parameter value fixed}
+\tablenotetext{b}{The {\sc CfA/Kurucz} library is scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.}
+\end{deluxetable}
+
 \begin{figure}[!htb]
@@
 
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6517 & 13 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex\tablenotemark{a}} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
-\cutinhead{Order 22 nuisance parameters}
-$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
-$\cc{22}{1}$ & -0.014 & 0.002\\
-$\cc{22}{2}$ & -0.014 & 0.002\\
-$\cc{22}{3}$ & -0.003 & 0.001\\
-\\
-$b_{22}$ & 1.04 & 0.02 \\
-$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
-$l_{22}$ & 7.5 & 0.8 $\kms$ \\
-
-\cutinhead{Order 23 nuisance parameters}
-$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
-$\cc{23}{1}$ & -0.013 & 0.002 \\
-$\cc{23}{2}$ & -0.025 & 0.001 \\
-$\cc{23}{3}$ &  0.001 &  0.001\\
-\\
-$b_{23}$ & 1.01 & 0.02 \\
-$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
-$l_{23}$ & 6.5 & 0.5 $\kms$\\
-
-\cutinhead{Order 24 nuisance parameters}
-$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00\tablenotemark{a}} \\
-$\cc{24}{1}$ & -0.010 & 0.002 \\
-$\cc{24}{2}$ & -0.011 & 0.001\\
-$\cc{24}{3}$ &  0.000 & 0.001\\
-\\
-$b_{24}$ & 1.02 & 0.02 \\
-$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
-$l_{24}$ & 7.3 & 0.6 $\kms$ \\
-\enddata
-\tablenotetext{a}{parameter value fixed}
-\end{deluxetable}
-
 \begin{figure}[!htb]
@@
 \includegraphics{figs/metacomparison.pdf}
-\caption{Comparing all the posteriors.}
+\caption{The posterior peaks and widths for various determination of the stellar parameters of WASP-14.
+The determinations in \citet{torres12} are shown along with the estimates using the {\sc CfA/Kurucz} synthetic library and the {\sc PHOENIX} synthetic library.
+\todo{What else do we want to show on here?}}
 \label{fig:metacomparison}
@@
 
-To understand the apparent differences in the quoted precisions, it is beneficial to review what factors 
-into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
-(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
-that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
-range that no local covariance kernels need to be instantiated; in this specific case we use only a 
-global covariance kernel.}  To break down their respective contributions, we performed the 
-same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
-likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
-best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
-$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
-uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
-are generated from the model; in this paradigm, the only way to introduce deviation is through 
-white noise (see Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is invalid, since just a few simple parameters 
-($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
-spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
-have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
-noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
-systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
-posteriors. For synthetic models which are less accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend to dominate the error budget.
+To understand the apparent differences in the quoted precisions, it is
+beneficial to review what factors into the parameter uncertainties.  In our
+approach, we consider contributions from both trivial (Poisson) noise and a
+non-trivial Gaussian process covariance matrix.\footnote{It is worth noting
+  that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this
+  limited spectral range that no local covariance kernels need to be
+  instantiated; in this specific case we use only a global covariance kernel.}
+  To break down their respective contributions, we performed the same analysis
+  with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies
+  the likelihood function to the familiar $\chi^2$ form.  Doing this, we
+  identify little bias to the best-fit parameters, but find unrealistic
+  precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, $[{\rm Fe/H}]$ to
+  $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small
+  uncertainties are to be expected, since the prescribed likelihood function
+  presumes that the data are generated from the model; in this paradigm, the
+  only way to introduce deviation is through white noise (see
+  Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is
+  invalid, since just a few simple parameters ($\vT$) are generally
+  insufficient to perfectly account for the thousands of pixels and hundreds of
+  spectral features in a typical dataset.  In essence, the data and an
+  imperfect model spectrum also have {\it systematic} differences that
+  contribute to the uncertainties (in addition to the Poisson noise), which we
+  forward-model using a non-trivial covariance matrix.  In this specific case,
+  these systematic uncertainties and the Poisson uncertainties contribute
+  roughly equally to the $\vT$ posteriors. For synthetic models which are less
+  accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend
+  to dominate the error budget.
 
@@
 However, for well-calibrated data sets from the same instrument any ill-effects are generally small.
-To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim10$ K and $\lesssim 0.001$ dex in $\Z$.
+To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim20$ K and $\lesssim 0.02$ dex in $\Z$.
 When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.
-
-\comm{Finally, you can transition into the model systematics issue.  Here you want to do the 
-modeling for Phoenix in orders 22-24 and demonstrate that you get much different parameters.  You 
-will have to show the results, and then can take that opportunity to discuss the local covariance 
-kernels.  Just show the results and make these two points, don't belabor the details: I don't 
-know what to make of the weird Phoenix parameters.}
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics[draft, width=0.5\textwidth,height=2in]{figs/PHOENIX_triangle.pdf}
-\caption{ {\sc Phoenix} WASP-14 marginal posteriors.}
-\label{fig:PHOENIX_posterior}
-\end{center}
-\end{figure}
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/residuals_PHOENIX_logg.pdf}
-\caption{ {\sc Phoenix} WASP-14 model and residuals.}
-\label{fig:residuals_PHOENIX}
-\end{center}
-\end{figure*}
 
@@
 \cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6516.8 & 12.6 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.72 & 0.0004 \\
+$T_{\rm eff}$ & 6273 & 15 K \\
+$\log_{10} g$\tablenotemark{a} & \multicolumn{2}{c}{4.29 dex} \\
+$\Z$ & -0.498 & 0.006 dex \\
+$v \sin i$ & 4.89 & 0.05  $\kms$\\
+$v_z$ & -4.86 & 0.02 $\kms$ \\
+$\log_{10} \Omega$ & -19.678  & 0.001 \\
 \cutinhead{Order 22 nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b_{22}$ & & \\
-$a_{{\rm g}, 22}$ & & \\
-$l_{22}$ & & \\
+$\cc{22}{0}$ & 0.012 & 0.001 \\
+$\cc{22}{1}$ & -0.021 & 0.003 \\
+$\cc{22}{2}$ & -0.018 & 0.002 \\
+$\cc{22}{3}$ & -0.004 & 0.002 \\
+$b_{22}$ & 1.02 & 0.02 \\
+$\log_{10}a_{{\rm g}, 22}$ & -14.33 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{22}$ & 9.0 & 0.5 $\kms$\\
 \cutinhead{Order 23 nuisance parameters}
-$\cc{23}{0}$ & & \\
-$\cc{23}{1}$ & & \\
-$\cc{23}{2}$ & & \\
-$\cc{23}{3}$ & & \\
-$b_{23}$ & & \\
-$a_{{\rm g}, 23}$ & & \\
-$l_{23}$ & & \\
+$\cc{23}{0}$ & 0.004 & 0.001  \\
+$\cc{23}{1}$ & -0.015 & 0.003 \\
+$\cc{23}{2}$ & -0.024 & 0.003 \\
+$\cc{23}{3}$ & -0.002 & 0.003 \\
+$b_{23}$ & 1.00 & 0.02 \\
+$\log_{10}a_{{\rm g}, 23}$ & -14.30 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{23}$ & 9.8 & 0.5 $\kms$\\
 \cutinhead{Order 24 nuisance parameters}
-$\cc{24}{0}$ & & \\
-$\cc{24}{1}$ & & \\
-$\cc{24}{2}$ & & \\
-$\cc{24}{3}$ & & \\
-$b_{24}$ & & \\
-$a_{{\rm g}, 24}$ & & \\
-$l_{24}$ & & \\
+$\cc{24}{0}$\tablenotemark{a} & \multicolumn{2}{c}{0.00} \\
+$\cc{24}{1}$ & -0.009 & 0.003 \\
+$\cc{24}{2}$ & -0.012 & 0.003 \\
+$\cc{24}{3}$ & -0.001 & 0.002 \\
+$b_{24}$ & 1.00 & 0.02 \\
+$\log_{10} a_{{\rm g}, 24}$ & -14.32 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
+$l_{24}$ & 8.5 & 0.5 $\kms$ \\
 \enddata
+\tablenotetext{a}{parameter value fixed}
+\tablecomments{ \todo{There are literally 100 other parameters for the $\sim30$ local kernels. How should I present these in the Table?}}
 \end{deluxetable}
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/PHOENIX_triangle.pdf}
+\caption{ {\sc Phoenix} WASP-14 marginal posteriors. Note $v \sin i$ may differ because the libraries assume different micro and macroturbulence.}
+\label{fig:PHOENIX_posterior}
+\end{center}
+\end{figure}
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\caption{ {\sc Phoenix} WASP-14 model and residuals.}
+\label{fig:PHOENIX_residuals}
+\end{center}
+\end{figure*}
 
@@
 Some routines choose to ``sigma-clip'' the outlier lines and procede with fitting, however with such an approach it is difficult to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
-For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
-
-This behaviour motivates the global and local kernels in a more sophisticated likelihood function, which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
-The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum.
-Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the outlier lines.
-
-Using the {\sc Phoenix} models on WASP-14, we infer substantially different posteriors than with using the {\sc CfA/Kurucz} library.
+For example, if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters, then this information would be lost in a sigma-clipping procedure.
+
+This behavior of the {\sc PHOENIX} library motivates the use of global and local covariance kernels in a sophisticated likelihood function.
+The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum, and acts as a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+Including this covariance broadens the posteriors on the stellar parameters, while the local covariance kernels provide cushion against outlier lines.
+
+Using the {\sc Phoenix} models on WASP-14, we infer substantially different posteriors than with the {\sc CfA/Kurucz} library.
 Figure~\ref{fig:PHOENIX_posterior} displays some of the relevant marginal posteriors, and Table~\ref{table:PHOENIX} compiles the estimates of the peaks of the posterior distribution and associated confidence intervals.
-Figure~\ref{fig:residuals_PHOENIX} demonstrates the quality of this approach and the structure of the associated residuals.
-Compared to the Kurucz residuals in Figure~\ref{fig:residuals_Kurucz}, there are two significant differences.
+Figure~\ref{fig:PHOENIX_residuals} demonstrates the quality of this approach and the structure of the associated residuals.
+Compared to the Kurucz residuals in Figure~\ref{fig:Kurucz_residuals}, there are two significant differences.
 First, the amplitude of the global covariance structure is larger.
-The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance.
-Second, local covariance kernels track outlier spectral lines.
-These kernels can map out the regions of the spectrum which are outliers.
-The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
-The locations and amplitudes of these kernels output from the code would be useful information to consider when refining the models of stellar sythenis.
-
-Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveals that there is a systematic offset depending on which spectral library is used as a backend.
-Depending on the choice of spectral library, there may be systematic differences of 200 K or more in $T_{\rm eff}$ and 0.2 dex or more in $\Z$.
+Because the amplitude of the covariance (both global and local) is an indirect parameterization of the quality of a model fit, better spectral models will have a lower covariance.
+Second, local covariance kernels are instantiated to track outlier spectral lines.
+These kernels map out the regions of the spectrum where there are large outlier spectral lines.
+The parameters of these local kernels provide a framework to identify incorrectly modelled spectral lines iteratively and in a self-consistent manner, and would be a useful data product to consider when refining the synthetic stellar models.
+The most striking difference between the two libraries is the $\sim$200 K systematic offset in temperature.
+Any systematic bias that results is most likely a direct consequence of the various physics and atomic constants included in each synthetic library.
 
@@
 
-Modeling late-type stellar atmosphere structures and their spectra is considerably more complex 
-than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
-opacities, as well as the presence of complex condensates (clouds) at the coolest temperatures.  
-\comm{should probably add some citations to that statement.  Make a comment recognizing that 
-various approaches have been taken, in context of Mann.}  \citet{rojas-ayala12} used a $K$-band 
-spectrum at similar resolution to the SPEX data to estimate the key Gl~51 stellar parameters based 
-on a line index technique.  They measured equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} 
-doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, along with a pseudo-continuum index defined over 
-three narrow passbands, 2.070--2.090, 2.235--2.255, and 2.360--2.380\,$\mu$m (see their Figure 2 
-for a useful visualization), designed to probe the spectral curvature induced by H$_2$O absorption 
-bands.  With reference to a custom grid of {\sc BT-Settl} spectra\footnote{From the web simulator at \url{http://phoenix.ens-lyon.fr/simulator/index.faces}}, \citet{rojas-ayala12} find that 
-these indices are consistent with $T_{\rm eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$.
+Modeling late-type stellar atmosphere structures and their spectra is
+considerably more complex than for Sun-like stars, due to lingering
+uncertainties in the atmosphere physics and molecular opacities, as well as the
+presence of complex condensates (clouds) at the coolest temperatures
+\citep{allard13}, making it considerably more challenging to determine stellar
+properties \citep{rajpurohit14}. Various approaches have been taken to
+determine stellar parameters in the face of these difficulties, including
+iteratively masking the regions with poor spectral agreement \citep{mann13}.
+However, \citet{mann13} note that such a scheme may exclude slightly discrepant
+regions of the spectrum that actually contain a significant amount of
+information useful to discriminate stellar properties, and that a more
+sophisticated approach would weight each spectral region based upon its
+consistency with the data. \citet{rojas-ayala12} used a $K$-band spectrum at
+similar resolution to the SPEX data to estimate the key Gl~51 stellar
+parameters based on a line index technique.  They measured equivalent widths
+for the 2.205\,$\mu$m \ion{Na}{1} doublet and 2.263\,$\mu$m \ion{Ca}{1}
+triplet, along with a pseudo-continuum index defined over three narrow
+passbands, 2.070--2.090, 2.235--2.255, and 2.360--2.380\,$\mu$m (see their
+Figure 2 for a useful visualization), designed to probe the spectral curvature
+induced by H$_2$O absorption bands.  With reference to a custom grid of {\sc
+BT-Settl} spectra\footnote{Created using the web simulator at
+   \url{http://phoenix.ens-lyon.fr/simulator/index.faces}},
+\citet{rojas-ayala12} find that these indices are consistent with $T_{\rm
+eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$.
 
@@
 $T_{\rm eff}$ & 2998  & 17 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{5.0 dex} \\
+$\log_{10} g$\tablenotemark{a} & \multicolumn{2}{c}{5.0 dex} \\
 $\Z$ & -0.01 & 0.04 dex \\
@@
 \cutinhead{Nuisance parameters}
-$\log_{10} c_0$ & \multicolumn{2}{c}{0.0} \\
+$\log_{10} c_0$\tablenotemark{a} & \multicolumn{2}{c}{0.0} \\
 $c_1$ & -0.011 & 0.002 \\
@@
 \enddata
+\tablenotetext{a}{parameter value fixed}
 \end{deluxetable}
 
-\comm{Now, present your results in a format that parallels the WASP-14 subsection.  You need a new 
-Table and a sequence of figures; introduce them all here.}
-
-\comm{Commentary on the necessity and utility of local covariance kernels here, with discussion of why Na and Ca lines are a pain.
-It is also worth commenting on the potential for flexibly extending the shape of
-the kernel to mimic molecular bandheads, if one wanted to try and fit say the TiO bands.}
-
-They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
-\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
-When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{report parameters}.
+Using the same region of the spectrum as \citet{rojas-ayala12} and the {\sc PHOENIX} library \citep{husser13}, and an appropriate $\delta$-function prior on $\log
+g = 5.0$, we infer parameters that are consistent with the \citet{rojas-ayala}
+approach: $T_{\rm eff} = 2998 \pm17$\,K, $[{\rm Fe/H}] = -0.01\pm0.04$\,dex, and
+$v \sin i \lesssim 10$\,km s$^{-1}$.  Figure~\ref{fig:Gl51_posterior}
+displays some relevant marginal projections of the posterior-space, and
+Table~\ref{table:Gl51} compiles our estimates of the peaks of the marginal
+posteriors and associated confidence intervals for all parameters employed in
+the modeling.  Figure~\ref{fig:Gl51_residuals} demonstrates the overall
+quality of this approach in reproducing the Gl51 spectrum and, most
+significantly, the structure of the associated residuals.
 
@@
   \includegraphics[width=0.5\textwidth]{figs/Gl51_triangle.pdf}
-  \caption{Gl51 stellar posterior.}
-\label{fig:gl_posterior}
+  \caption{Gl51 stellar posterior marginalized over the other nuisance parameters, including the local covariance kernels which account for the \ion{Na}{1} and \ion{Ca}{1} lines.}
+\label{fig:Gl51_posterior}
 \end{center}
@@
   \caption{The Gl51 spectrum and residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.}
-\label{fig:regions}
+\label{fig:Gl51_residuals}
 \end{center}
@@
 
-\comm{A simple two-sentence (or so) wrap up on the future promise of this approach for late-type 
-stars and their complex spectra.}
-
-Complicated molecular structure is poorly known from models.
-
-\clearpage
+We find, like \citet{rojas-ayala12} do, that the {\sc PHOENIX} models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using high metallicity models.
+\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
+Fortunately, the local kernels provide a mechanism for self-consistently determining the weights of discrepant spectral lines, like that speculated about by \citet{mann13}, such that they bring their full information content to bear on the stellar posterior distribution without bias.
+Figure~\ref{fig:Gl51_residuals} shows that these kernels have identified these two line groups as problematic and appropriately inflated the covariance for this region of spectrum, reducing the ability of these lines to bias the final stellar parameters.
+
+In other regions of the spectrum, late type stars also present other types of spectral outliers.
+In particular, the TiO bandheads can produce a complicated saw-like residual pattern.
+Like a Gaussian function was used as the starting point for residuals produced by single spectral lines, the local kernel framework could be extended to include an additional kernel which would mimic the appearance of the residuals generated from poorly-modelled molecular bandheads.
+For example, the product of a truncated exponential and a \matern\ kernel or other similar hybrid kernels might capture this residual behavior well.
+Our forward modelling technique has great potential for fitting the spectra of late-type stars to derive stellar parameters.
+Because large regions of the spectrum may be in modest but not perfect agreement with the data, the usage of covariance kernels allows us to utilize these mildly discrepant regions by self-consistently determining their weighted contribution, maximizing the potential inference on the stellar parameters of interest.
 
@@
 
-\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
-be tough to do concisely, but here's what I think the framework should be: (A1) a statement of the 
-problem (i.e., that we want to extract physical information from a spectrum without bias or 
-systematic error, but the models we have are not perfect); (A2) a [brief] reminder of the standard 
-approaches to skirt this issue; (A3) a summary of our approach; (A4) a statement that we demonstrated 
-the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
-
-\comm{Next, more specifically highlight why we developed this framework, and what its 
-role should be in general spectroscopic inference.  I think its worth highlighting several key 
-advantages: (B1) you are casting the standard data selection/weighting/masking procedure in an 
-internally consistent, Bayesian format that appropriately preserves and propagates information 
-content even in the face of imperfect models (along the lines of what you've written in the 
-commented-out paragraphs below, but more concise/direct); (B2) (some deleted text could be useful here) by doing that, you allow the user to 
-actually exploit their full datasets, which should then reflect a more appropriate inference on the 
-parameters of interest without concern that some bad or less-informative features will screw them 
-up (sort of like what I commented out above, but less preachy) -- its in here 
-that we should suggest that this framework could be incorporated as a backend to standard methods, 
-where it essentially is employed to define an appropriate likelihood; and (B3) it creates an 
-opportunity to develop quantitative feedback to improve the models, or ultimately to 
-pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
-since you will go into depth on it below). (est.~2 paragraphs)} \\
-
-A2: The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
+Astronomical spectroscopists generally seek to extract the maximally precise physical information (i.e., the fundamental stellar parameters) from a spectrum without suffering from bias.
+Ideally, any likelihood computation of the most probable stellar model parameters would be directly coupled to a sophisticated synthesis code which incorporates the multitude of atomic constants and atmospheric structure choices required for stellar synthesis as free parameters in the forward model.
+However, the demanding realities of stellar synthesis and the current computational reality necessitate that astronomers compromise by either using a limited spectral range or a simplified forward model.
+For many applications that require large spectral range, it is most practical to use spectra from a pre-computed library.
+Although these models are generally of exceptional quality, the inability to tweak individual line strengths necessitates a technique to account for the complicated structure of the residuals of a model fit.
+The traditional method of dealing with this structure is first to simply mask the regions of the spectrum which do not agree to within a certain tolerance, assume the residuals are due to white Poisson noise, and then proceed with a $\chi^2$ likelihood function.
+Rather than assuming white noise, we advocate that one should include a parameterization that accounts for the mild, globally covariant nature of the residuals, which results from the inability of the forward model to adjust individual line strengths.
+Rather than arbitrarily excluding discrepant spectral lines from the fit, one should include these regions with the appropriate weight.
 A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-B1: The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
+Our fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
+In fact, for some types of stars with difficult-to-model spectra, it may be possible that the mildly discrepant lines actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
+We have demonstrated the flexibility of this approach to model spectra \emph{in general} by determining the stellar parameters of two very different stars, whose spectra were acquired with very different instruments.
+
+One major advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$ is that the model-data mismatch is treated internally to the model framework.
 When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
-When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters.
-This hierarchical model also works within a single spectrum when fitting more than one spectral order.
-If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
-However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
-
-\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
-see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
-local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
-what you've already written (commented out below) is in good shape, but you might decide to change 
-the format/organization given the suggested update in structure for this section.  Go out on a 
-positive note: the mechanics already exist, and this tool can be used to build better models!}
-
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
+When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters and local kernels may be instantiated to cover outlier lines.
+This enables users to use a larger swath of the data spectrum, including regions which may be more poorly modelled (e.g., molecular bands), without undue concern that their stellar posteriors might be biased from outlier lines.
+This self-consistent property of a hierarchical model also works within a single echelle spectrum: if each order has a similar prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
+However, if certain orders lead to the inference of different stellar parameters, then global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
+The local kernels will track spectral outliers and provide useful feedback on how the spectral synthesis could be improved.
+
+The application of this likelihood function we developed should not necessarily be limited to forward models which rely upon interpolation from a synthetic library.
+In practice, any technique which synthesizes a spectrum which has more spectral lines than free parameters in the model will suffer at some level from the globally covariant residual structure.
+In order to avoid overly-precise uncertainty estimates, this more sophisticated likelihood function should be used as a best-fit metric instead of the more common $\chi^2$.
+
+The benefits of using covariance kernels extend beyond the case of fitting a single stellar spectrum.
 If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
-By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
-After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
-The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
-Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
-
-B3: Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
-Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
-Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
+By cataloguing the covariance structure of the residuals, especially the parameters of the local kernels, we collect valuable information about the quality of the synthetic spectra and the regions where they deviate from physical reality.
+The modular approach of the local covariance kernels also enables this framework to be extended to accommodate new types of stellar behavior.
+For example, the complicated structure of molecular bandheads in M stars could be parameterized by a custom saw-shaped covariance kernel, which would alleviate some of the tension that is present in these spectroscopic fits (e.g., the TiO band, \citealt{mann13}, Fig.
+9).
+
+After fitting several stars of similar spectral type, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
+Linking the parameters of the covariance kernels in a hierarchical Bayesian model allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted and enables us to learn about the population and frequency of spectral modelling errors.
 In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
@@
 
-B3: Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
-Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
-This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
-
-Infer the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
-This will tell us about the frequency and distribution of spectral modelling errors.
-Or we could just speculate about it.
-Could also cross correlate with a catalogue of lines to determine which.
+Once determined, this average covariance matrix could be delivered to the spectral synthesis communities, which would enable them to rapidly pinpoint and correct non-physical lines in the model spectra.
+Alternatively, users could make corrections to the models themselves by using the chain of logic and mathematical post-processing (\S\ref{subsec:postprocess}) that were used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
+This application of machine learning would enable users to create their own library of data driven, semi-empirical stellar models.
+Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach naturally provides a physical anchor point for the fundamental stellar parameters, since they are tied down by the laws of stellar physics that were used to create the synthetic spectra.
+Such a technique holds great promise in the era of large high-resolution spectral datasets like SLOAN/\emph{APOGEE} and \emph{HERMES}, where large homogeneous collections of stars will provide enormous leverage for first identifying which parts of a stellar spectrum may be discrepant due to systematic error in the models themselves and then inferring what the most accurate representation of the star in this region should be.
 
@@
     produce low-amplitude correlations in the residuals on the length scale of
-    the width of a typical spectral line.  ( {\it second panel}): Sometimes
+    the width of a typical spectral line.  ( {\it second panel}) Sometimes
     ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
     this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  ( {\it third panel}): Sometimes ($\lesssim
+    patch of negative residuals.  ( {\it third panel}) Sometimes ($\lesssim
     5$\% of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals. ({\it fourth panel}): If the line
+    correlated patch of positive residuals. ({\it fourth panel}) If the line
     strengths are substantially discrepant ($\lesssim 10$\% of all lines),

### 2014-08-25 notebooks/paper/generate_PHOENIX_residual_draws.py
Seems like it's done

@@ -4,27 +4,31 @@ from StellarSpectra.spectrum import DataSpectrum
 from StellarSpectra.grid_tools import TRES, HDF5Interface
 from StellarSpectra import utils
 
-myDataSpectrum = DataSpectrum.open("../../data/WASP14/WASP14-2009-06-14.hdf5", orders=np.array([22]))
+myDataSpectrum = DataSpectrum.open("../../data/WASP14/WASP14-2009-06-14.hdf5", orders=np.array([21,22,23]))
 
 myInstrument = TRES()
 
 myHDF5Interface = HDF5Interface("../../libraries/PHOENIX_TRES_F.hdf5")
 
 #Load a model using the JSON file
-myModel = Model.from_json("../WASP14_22_model_final_region.json", myDataSpectrum, myInstrument, myHDF5Interface)
+myModel = Model.from_json("WASP14_PHOENIX_model0_final.json", myDataSpectrum, myInstrument, myHDF5Interface)
 
-myOrderModel = myModel.OrderModels[0]
+myOrderModel = myModel.OrderModels[1]
 model_flux = myOrderModel.get_spectrum()
 
 spec = myModel.get_data()
-wl = spec.wls[0]
-fl = spec.fls[0]
+wl = spec.wls[1]
+fl = spec.fls[1]
 
 model_fl = myOrderModel.get_spectrum()
 residuals = fl - model_fl
 
-mask = spec.masks[0]
-cov = myModel.OrderModels[0].get_Cov().todense()
+mask = spec.masks[1]
+cov = myModel.OrderModels[1].get_Cov().todense()
+
+np.save("PHOENIX_covariance_matrix.npy", cov)
+
+import sys; sys.exit()
 
 draws = utils.random_draws(cov, num=50)
 

### 2014-08-25 hdfplot.py
Seems like it's done

@@ -8,4 +8,8 @@ import hdfutils
 Script designed to plot the HDF5 output from MCMC runs. It will only plot the first chain, to prevent overwriting.
 '''
 
+if hdfutils.args.cov:
+    hdfutils.estimate_covariance(hdfutils.flatchainTreeList[0])
+
 hdfutils.plot(hdfutils.flatchainTreeList[0], clip_stellar=hdfutils.args.stellar_params)
+

### 2014-08-25 tex/paper.tex
Paper draft ready for SA review

@@
 
-In addition to statistical and systematic error, one might expect that systematic instrumental error might also contribute to the parameter uncertainties.
-Commonly, one might estimate this error by fitting spectra taken by different instruments or at different times and then averaging the results.
-Additionally, one may add in an additional uncertainty ``floor'' that maps to the empirical scatter of typical spectroscopic fits.
-Such procedures generally make the assumption that this additional error contribution is Gaussian and independent in each parameter of interest.
-Our hierarchical Bayesian model avoids such a heuristic by
-parameterizing the quality of fit self-consistently: because the parameter uncertainty is derived through the model itself, it preserves correlations between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
-However, for well-calibrated data sets from the same instrument any ill-effects are generally small.
-To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim20$ K and $\lesssim 0.02$ dex in $\Z$.
-When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.
+In addition to statistical and systematic error, one might expect that
+systematic instrumental error might also contribute to the parameter
+uncertainties. Commonly, one might estimate this error by fitting spectra taken
+by different instruments or at different times and then averaging the results.
+Additionally, one may add in an additional uncertainty ``floor'' that maps to
+the empirical scatter of typical spectroscopic fits. Such procedures generally
+make the assumption that this additional error contribution is Gaussian and
+independent in each parameter of interest. Our hierarchical Bayesian model
+avoids such a heuristic by parameterizing the quality of fit self-consistently:
+because the parameter uncertainty is derived through the model itself, it
+preserves correlations between parameters, like the well known degeneracies
+between $T_{\rm eff}$, $\log g$, and $\Z$. Arbitrarily inflating the posterior
+estimates to reflect intuition would artificially destroy any natural
+degeneracy of parameters. However, for well-calibrated data sets from the same
+instrument any ill-effects are generally small. To quantify this error
+contribution for the WASP-14 spectrum in the traditional manner, we
+independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of
+$\sim20$ K and $\lesssim 0.02$ dex in $\Z$. When fitting these 4 spectra
+simultaneously in a hierarchical manner, we find that the posteriors shrink by
+a similar amount as one would achieve by combining the posterior estimates of
+the individual spectra in a weighted average. However, modelling the different
+spectra in a simultaneous manner allows us to preserve the intrinsic
+correlations between stellar parameters, something that would be destroyed in a
+naive weighted average.
 

### 2014-08-27 tex/paper.tex
edits on revised Sect 3, 4

--- 2014-08-24 tex/paper.tex
@@
 
-\title{A Method for the Spectroscopic Inference of Stellar Parameters}
-\author{Ian Czekala, Sean M.~Andrews, et al.}
+%\title{A Method for the Spectroscopic Inference of Stellar Parameters}
+\title{Robust Spectroscopic Inference with Imperfect Models}
+\author{Ian Czekala, Sean M.~Andrews, Guillermo Torres, \& David W.~Latham}
 \affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 (on $\lambda_i$ and $\lambda_j$), and not simply the distance between them ($r_{ij}$).
+
+\begin{figure}[!htb]
+\begin{center}
+  \includegraphics{figs/badlines0.pdf}
+  \includegraphics{figs/badlines1.pdf}
+  \includegraphics{figs/badlines2.pdf}
+  \includegraphics{figs/badlines3.pdf}
+  \caption{A collection of spectral lines which have imperfect model fits.
+    ( {\it first panel}) The majority of spectral lines ($\gtrsim 60$\%) will
+    have minor differences in strength between data and model spectrum, which
+    produce low-amplitude correlations in the residuals on the length scale of
+    the width of a typical spectral line.  ( {\it second panel}) Sometimes
+    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
+    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
+    patch of negative residuals.  ( {\it third panel}) Sometimes ($\lesssim
+    5$\% of all lines), an extraneous line in the model leaves a large, highly
+    correlated patch of positive residuals. ({\it fourth panel}) If the line
+    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
+    there will be many correlated residuals of moderate amplitude. For any specific line, there might
+    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
+    will properly fit \emph{all} the lines.}
+\label{fig:badlines}
+\end{center}
+\end{figure}
 
@@
 star Gliese\,51 (hereafter Gl\,51), observed as part of the NASA/IRTF library of spectral standards 
-\citep{cushing05,rayner09}.  In this case we also use the {\sc Phoenix} library, since it forms a complete grid for late spectral types.
+\citep{cushing05,rayner09}.  In this case we use the {\sc Phoenix} library to generate models, 
+since it is well-sampled at cooler temperatures.
 
@@
 full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
-standard techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for more 
-specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
+standard techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for 
+more specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
 
@@
 \ref{sec:intro}.  Their analysis was performed in two ways, with the surface gravity treated as a 
-free parameter or assumed to be fixed (to $\log g = 4.29$), based on an inference of the mean 
+free parameter or assumed to be fixed (to $\log g = 4.29$) based on an inference of the mean 
 stellar density from the exoplanet transit depth and a comparison of WASP-14 photometry to stellar 
@@
 
-Given the decision to employ a spectral library to generate models
-(Sect.~\ref{subsec:synthetic}), it is most relevant to compare the
-\citeauthor{torres12}~{\tt SPC} result with our framework.  Using the same TRES
-spectral orders (for only a single WASP-14 observation, as noted above), the
-{\sc CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log
-g$, we infer parameters that are entirely consistent with the {\tt SPC}
-approach: $T_{\rm eff} = 6517\pm13$\,K, $[{\rm Fe/H}] = -0.27\pm0.01$\,dex, and
-$v \sin i = 4.3\pm0.1$\,km s$^{-1}$.  Figure~\ref{fig:Kurucz_posterior}
-displays some relevant marginal projections of the posterior-space, and
-Table~\ref{table:Kurucz} compiles our estimates of the peaks of the marginal
-posteriors and associated confidence intervals for all parameters employed in
-the modeling.  Figure~\ref{fig:Kurucz_residuals} demonstrates the overall
-quality of this approach in reproducing the WASP-14 spectrum and, most
-significantly, the structure of the associated residuals.
-Figure~\ref{fig:metacomparison} compares the inferred parameters with the
-various determinations made by \citet{torres12}.
-
-\begin{figure}[!htb]
+Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
+a direct comparison of the \citeauthor{torres12}~{\tt SPC} result with our framework is relevant.  
+Using the same TRES spectral orders (for a single WASP-14 observation, as noted above), the {\sc 
+CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer parameters 
+that are entirely consistent with the {\tt SPC} approach; see Table~\ref{table:Kurucz}.  
+Figure~\ref{fig:Kurucz_residuals} demonstrates the quality of our approach in reproducing the 
+WASP-14 spectrum and, especially significant, the structure of the residuals.  
+Figure~\ref{fig:Kurucz_posterior} shows some marginal projections of the posterior-space, 
+highlighting key parameter degeneracies.  
+%Figure~\ref{fig:metacomparison} compares the inferred 
+%parameters with the various determinations made by \citet{torres12}.
+
+\begin{figure*}[!t]
+\includegraphics{figs/residuals_Kurucz_logg.pdf}
+\figcaption{({\it top}) A representative segment of the TRES spectrum of WASP-14 ({\it blue}), 
+overlaid with a {\sc CfA/Kurucz} model ({\it red}) generated by drawing parameters from the 
+posterior distribution (for a fixed $\log g = 4.29$).  ({\it bottom}) The residual spectrum along 
+with contours representing the distributions of a large number of random draws from the covariance 
+matrix (the shading is representative of the 1, 2, and 3\,$\sigma$ spreads of that distribution of 
+draws), as in Fig.~\ref{fig:matrix}.  With such a finely tuned model library, local covariance 
+kernels are not required.  \label{fig:Kurucz_residuals}}
+\end{figure*}
+
+\begin{deluxetable}{lr@{ $\pm$ }l|r@{ $\pm$ }l}[!b]
+\tablecaption{\label{table:Kurucz}Inferred Parameters for WASP-14}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc CfA/Kurucz}} & \multicolumn{2}{c}{{\sc Phoenix}}}
+\startdata
+$T_{\rm eff}$ (K) & 6517     & 13                      &  6273   & 15        \\
+$\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
+$\Z$          & -0.27    & 0.01                    & -0.498  & 0.006     \\
+$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.29     & 0.05                    &  4.89   & 0.05      \\
+$v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
+$\log \Omega$\tablenotemark{b} & -12.7181 & 0.0004 & -19.678 & 0.001
+\enddata
+\tablenotetext{a}{Small differences in the rotational broadening between the two libraries are
+expected, since they assume different micro- and macro-turbulence behavior in the stellar
+atmospheres.}
+\tablenotetext{b}{The {\sc CfA/Kurucz} spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.
+The inference on $\Omega$ from the {\sc Phoenix} library is in steradian units.}
+\tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
+distributions for each parameter.  The quoted uncertainties correspond to the 68.3\%\ 
+($\sim$1\,$\sigma$) confidence intervals.  Expanded lists of the inferred values and uncertainties 
+for all nuisance parameters are available electronically.}
+\end{deluxetable}
+
+\begin{figure}[!b]
+  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+  \figcaption{The posterior probability distributions for the interesting stellar parameters 
+(marginalized over all other model parameters) of WASP-14 based on the {\sc CfA/Kurucz} model 
+library, as explored by the MCMC Gibbs sampler.  For the two-dimensional posterior projections, 
+contours are drawn at the standard 1, 2, and 3\,$\sigma$ levels (of a normal distribution) for 
+reference.  \label{fig:Kurucz_posterior} }
+\end{figure}
+
+%\begin{figure}[!htb]
+%\begin{center}
+%\includegraphics{figs/metacomparison.pdf}
+%\caption{The posterior peaks and widths for various determination of the stellar parameters of WASP-14.
+The determinations in \citet{torres12} are shown along with the estimates using the {\sc CfA/Kurucz} synthetic library and the {\sc PHOENIX} synthetic library.
+\todo{What else do we want to show on here?}}
+%\label{fig:metacomparison}
+%\end{center}
+%\end{figure}
+
+It is beneficial to review what factors into the parameter uncertainties.  In our approach, we 
+consider contributions from both trivial (Poisson) noise and a non-trivial Gaussian process 
+covariance matrix.\footnote{Note that the {\sc CfA/Kurucz} library is so well-tuned for stars like 
+WASP-14 over this limited spectral range that no local covariance kernels need to be instantiated; 
+in this specific case, we employ only global covariance kernels to reproduce the residual 
+structure.} To decompose their respective contributions, we performed the same analysis with only a 
+trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the likelihood function to the 
+familiar $\chi^2$ form.  Doing this, we identify little bias to the best-fit parameters, but find 
+unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, $[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v 
+\sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small uncertainties are to be expected, since the 
+prescribed likelihood function presumes that the data are generated from the model; in this 
+paradigm, the only way to introduce deviation is through white noise (see Fig.~\ref{fig:matrix}).  
+Of course, that implicit assumption is invalid, since just a few simple parameters ($\vT$) are 
+generally insufficient to account for the thousands of pixels and hundreds of spectral features in 
+a typical dataset.  In essence, the data and an imperfect model spectrum also have {\it systematic} 
+differences that also contribute to the uncertainties, which we forward-model using a non-trivial 
+covariance matrix.  In this specific case, these systematic and the Poisson uncertainties 
+contribute roughly equally to the $\vT$ posteriors.  For synthetic models that are less finely 
+tuned than the {\sc CfA/Kurucz} grid, the systematic uncertainties will dominate (see below).
+
+In addition to noise in the data (statistical uncertainties) and systematic uncertainties in the 
+models, we might also expect that systematics associated with the instrument and/or calibration 
+contribute to the parameter uncertainties.  Common practice is to estimate these latter systematics 
+by comparing parameter inferences made using different spectra (e.g., from different instruments or 
+observing nights): the results are usually presented as an average, with uncertainties inflated by 
+adding in a ``floor" term that accounts for the dispersion between the inferences from individual 
+spectra.  That procedure implicitly assumes that these systematics are Gaussian (add in quadrature) 
+and independent in each parameter.  But the approach we are advocating here avoids such a heuristic 
+by self-consistently parameterizing the fit quality.  Because the parameter uncertainties are 
+derived through a forward model, this method (crucially) preserves the morphologies of intrinsic 
+parameter degeneracies in a way that cannot be reproduced by artificially (i.e., manually) 
+inflating the posteriors.
+
+To quantify these instrumental systematics for WASP-14, we modeled the three other TRES spectra 
+obtained by \citet{torres12} independently (under the same assumptions).  We found a minimal 
+scatter of $\sim$20\,K in $T_{\rm eff}$ and $\sim$0.02\,dex in $\Z$ between the individual 
+inferences, a testament to quality calibration and an exceptionally stable spectrograph.  The more 
+appropriate way of combining the inferences is to model these four spectra simultaneously in our 
+hierarchical framework: in that case, we found that the parameter inferences become slightly more 
+precise (as one would expect for a weighted average and well-tuned models), while preserving the 
+intrinsic correlations between parameters (something expressly destroyed in a weighted average).
+
+%\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
+%orders with Phoenix.}
+
+While comparisons with the {\sc CfA/Kurucz} models confirm that our methodology can faithfully 
+reproduce the inferences made with standard spectral fitting routines, they do little to illustrate 
+a key feature -- the treatment of local patches of large covariance (see 
+Sect.~\ref{subsec:local_covariance}) in a more sophisticated likelihood function.  For that we turn 
+to the {\sc Phoenix} library, comprised of model spectra with a much broader spectral range that 
+have not been fine-tuned for solar type stars.  As a result, users should expect that there will 
+likely be a population of ``outlier" residual lines present when comparing any observed spectrum 
+with the {\sc Phoenix} models \citep[e.g.,][their Fig.~8]{husser13}.  The standard approach is to 
+outright ignore, or perhaps ``$\sigma$-clip", the outlier lines and proceed with fitting the 
+remainder of the spectrum.  However, it is difficult to quantify how such decisions ultimately 
+affect the parameter inferences \citep{hogg10}.  By parameterizing these residual features with 
+local covariance kernels, we are self-consistently preserving any discriminatory power against the 
+stellar parameters that might still be contained in their imperfect representations.
+
+\begin{figure*}[!htb]
+\includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\figcaption{Same as Fig.~\ref{fig:Kurucz_residuals} for the {\sc Phoenix} models.  Note the 
+increased vertical scale and local covariance structure for the residuals.
+ \label{fig:PHOENIX_residuals}}
+\end{figure*}
+
+\begin{figure}[!b]
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
-  \caption{The posterior probability function of the stellar parameters for WASP-14 as explored by the MCMC Gibbs sampler.
-These stellar parameters are marginalized over the Chebyshev ($\cheb$) and noise parameters ($\cov$).
-Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-}
-\label{fig:Kurucz_posterior}
+  \includegraphics[width=0.5\textwidth]{figs/PHOENIX_triangle.pdf}
+\figcaption{Same as Fig.~\ref{fig:Kurucz_posterior}, but for the {\sc Phoenix} models.  
+\label{fig:PHOENIX_posterior}}
 \end{center}
@@
 
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/residuals_Kurucz_logg.pdf}
-\caption{A sample of the WASP-14 data set overlaid with a model spectrum produced by parameters drawn from the posterior probability distribution (fixing $\log_{10} g = 4.29$ dex) and using the {\sc CfA/Kurucz} synthetic library.
-The residuals are shown zoomed along with the three-sigma contours of random draws from the covariance matrix, after Figure~\ref{fig:matrix}.}
-\label{fig:Kurucz_residuals}
-\end{center}
-\end{figure*}
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:Kurucz}WASP-14 {\sc CfA/Kurucz} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6517 & 13 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex\tablenotemark{a}} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.7181 & 0.0004 \\
-\cutinhead{Order 22 nuisance parameters}
-$\log_{10} \cc{22}{0}$ &  0.0085 & 0.0005\\
-$\cc{22}{1}$ & -0.014 & 0.002\\
-$\cc{22}{2}$ & -0.014 & 0.002\\
-$\cc{22}{3}$ & -0.003 & 0.001\\
-\\
-$b_{22}$ & 1.04 & 0.02 \\
-$\log_{10} a_{{\rm g}, 22}$ & -14.63 & 0.02 $\log_{10}\left[ \textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1} \right ]$\\
-$l_{22}$ & 7.5 & 0.8 $\kms$ \\
-
-\cutinhead{Order 23 nuisance parameters}
-$\log_{10} \cc{23}{0}$ & 0.0030 & 0.0006 \\
-$\cc{23}{1}$ & -0.013 & 0.002 \\
-$\cc{23}{2}$ & -0.025 & 0.001 \\
-$\cc{23}{3}$ &  0.001 &  0.001\\
-\\
-$b_{23}$ & 1.01 & 0.02 \\
-$\log_{10} a_{{\rm g}, 23}$ & -14.55 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
-$l_{23}$ & 6.5 & 0.5 $\kms$\\
-
-\cutinhead{Order 24 nuisance parameters}
-$\log_{10} \cc{24}{0}$ & \multicolumn{2}{c}{0.00\tablenotemark{a}} \\
-$\cc{24}{1}$ & -0.010 & 0.002 \\
-$\cc{24}{2}$ & -0.011 & 0.001\\
-$\cc{24}{3}$ &  0.000 & 0.001\\
-\\
-$b_{24}$ & 1.02 & 0.02 \\
-$\log_{10} a_{{\rm g}, 24}$ & -14.61 & 0.02 $\log_{10}\left[\textrm{erg}\;\textrm{cm}^{-2}\;\textrm{s}^{-1}\;\textrm{\AA}^{-1}\right]$\\
-$l_{24}$ & 7.3 & 0.6 $\kms$ \\
-\enddata
-\tablenotetext{a}{parameter value fixed}
-\end{deluxetable}
-
-\begin{figure}[!htb]
-\begin{center}
-\includegraphics{figs/metacomparison.pdf}
-\caption{Comparing all the posteriors.}
-\label{fig:metacomparison}
-\end{center}
-\end{figure}
-
-To understand the apparent differences in the quoted precisions, it is beneficial to review what factors 
-into the parameter uncertainties.  In our approach, we consider contributions from both trivial 
-(Poisson) noise and a non-trivial Gaussian process covariance matrix.\footnote{It is worth noting 
-that the {\sc CfA/Kurucz} library is so good for stars like WASP-14 over this limited spectral 
-range that no local covariance kernels need to be instantiated; in this specific case we use only a 
-global covariance kernel.}  To break down their respective contributions, we performed the 
-same analysis with only a trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the 
-likelihood function to the familiar $\chi^2$ form.  Doing this, we identify little bias to the 
-best-fit parameters, but find unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, 
-$[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v \sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small 
-uncertainties are to be expected, since the prescribed likelihood function presumes that the data 
-are generated from the model; in this paradigm, the only way to introduce deviation is through 
-white noise (see Figure~\ref{fig:matrix}, panel 1).  Of course, that implicit assumption is invalid, since just a few simple parameters 
-($\vT$) are generally insufficient to perfectly account for the thousands of pixels and hundreds of 
-spectral features in a typical dataset.  In essence, the data and an imperfect model spectrum also 
-have {\it systematic} differences that contribute to the uncertainties (in addition to the Poisson 
-noise), which we forward-model using a non-trivial covariance matrix.  In this specific case, these 
-systematic uncertainties and the Poisson uncertainties contribute roughly equally to the $\vT$
-posteriors. For synthetic models which are less accurate than the {\sc CfA/Kurucz} grid, systematic uncertainties will tend to dominate the error budget.
-
-In addition to statistical and systematic error, one might expect that systematic instrumental error might also contribute to the parameter uncertainties.
-Commonly, one might estimate this error by fitting spectra taken by different instruments or at different times and then averaging the results.
-Additionally, one may add in an additional uncertainty ``floor'' that maps to the empirical scatter of typical spectroscopic fits.
-Such procedures generally make the assumption that this additional error contribution is Gaussian and independent in each parameter of interest.
-Our hierarchical Bayesian model avoids such a heuristic by
-parameterizing the quality of fit self-consistently: because the parameter uncertainty is derived through the model itself, it preserves correlations between parameters, like the well known degeneracies between $T_{\rm eff}$, $\log g$, and $\Z$.
-Arbitrarily inflating the posterior estimates to reflect intuition would artificially destroy any natural degeneracy of parameters.
-However, for well-calibrated data sets from the same instrument any ill-effects are generally small.
-To quantify this error contribution for the WASP-14 spectrum in the traditional manner, we independently fit 4 separate \emph{TRES} spectra, and find a minimal scatter of $\sim10$ K and $\lesssim 0.001$ dex in $\Z$.
-When fitting these 4 spectra simultaneously in a hierarchical manner, we find that \todo{numbers}.
-
-\comm{Finally, you can transition into the model systematics issue.  Here you want to do the 
-modeling for Phoenix in orders 22-24 and demonstrate that you get much different parameters.  You 
-will have to show the results, and then can take that opportunity to discuss the local covariance 
-kernels.  Just show the results and make these two points, don't belabor the details: I don't 
-know what to make of the weird Phoenix parameters.}
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics[draft, width=0.5\textwidth,height=2in]{figs/PHOENIX_triangle.pdf}
-\caption{ {\sc Phoenix} WASP-14 marginal posteriors.}
-\label{fig:PHOENIX_posterior}
-\end{center}
-\end{figure}
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/residuals_PHOENIX_logg.pdf}
-\caption{ {\sc Phoenix} WASP-14 model and residuals.}
-\label{fig:residuals_PHOENIX}
-\end{center}
-\end{figure*}
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:PHOENIX}WASP-14 {\sc Phoenix} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
-\startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 6516.8 & 12.6 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{4.29 dex} \\
-$\Z$ & -0.27 & 0.01 dex \\
-$v \sin i$ & 4.29 & 0.05 $\kms$\\
-$v_z$ & -4.62 & 0.02 $\kms$ \\
-$\log_{10} \Omega$ & -12.72 & 0.0004 \\
-\cutinhead{Order 22 nuisance parameters}
-$\cc{22}{0}$ & & \\
-$\cc{22}{1}$ & & \\
-$\cc{22}{2}$ & & \\
-$\cc{22}{3}$ & & \\
-$b_{22}$ & & \\
-$a_{{\rm g}, 22}$ & & \\
-$l_{22}$ & & \\
-\cutinhead{Order 23 nuisance parameters}
-$\cc{23}{0}$ & & \\
-$\cc{23}{1}$ & & \\
-$\cc{23}{2}$ & & \\
-$\cc{23}{3}$ & & \\
-$b_{23}$ & & \\
-$a_{{\rm g}, 23}$ & & \\
-$l_{23}$ & & \\
-\cutinhead{Order 24 nuisance parameters}
-$\cc{24}{0}$ & & \\
-$\cc{24}{1}$ & & \\
-$\cc{24}{2}$ & & \\
-$\cc{24}{3}$ & & \\
-$b_{24}$ & & \\
-$a_{{\rm g}, 24}$ & & \\
-$l_{24}$ & & \\
-\enddata
-\end{deluxetable}
-
-\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
-orders with Phoenix.}
-
-While the CfA grid spectra fit well with no obvious features in the residual spectrum, because the PHOENIX spectrum is computed over a very broad spectral range with no specific tuning for solar type stars, there are likely to be a small amount of  ``outlier'' lines that appear in the residual spectrum (e.g., \citealt{husser13}, Figure 8).
-Some routines choose to ``sigma-clip'' the outlier lines and procede with fitting, however with such an approach it is difficult to quantify the effect that clipping a given line has on the final parameter estimates \citep{hogg10}.
-For example, what if a series of lines with $\sim$3 sigma disagreement actually has substantial discriminatory power against the stellar parameters?
-
-This behaviour motivates the global and local kernels in a more sophisticated likelihood function, which includes a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
-The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum.
-Including this global covariance broadens the posteriors on the stellar parameters, providing cushion against the outlier lines.
-
-Using the {\sc Phoenix} models on WASP-14, we infer substantially different posteriors than with using the {\sc CfA/Kurucz} library.
-Figure~\ref{fig:PHOENIX_posterior} displays some of the relevant marginal posteriors, and Table~\ref{table:PHOENIX} compiles the estimates of the peaks of the posterior distribution and associated confidence intervals.
-Figure~\ref{fig:residuals_PHOENIX} demonstrates the quality of this approach and the structure of the associated residuals.
-Compared to the Kurucz residuals in Figure~\ref{fig:residuals_Kurucz}, there are two significant differences.
-First, the amplitude of the global covariance structure is larger.
-The global covariance acts as a way to parameterize the goodness or badness of a fit: better spectral models will have a lower covariance.
-Second, local covariance kernels track outlier spectral lines.
-These kernels can map out the regions of the spectrum which are outliers.
-The covariance hyperparameters provide a framework to identify these regions iteratively and in a self-consistent manner.
-The locations and amplitudes of these kernels output from the code would be useful information to consider when refining the models of stellar sythenis.
-
-Comparing the results in Tables~\ref{table:Kurucz}~\&~\ref{table:PHOENIX} reveals that there is a systematic offset depending on which spectral library is used as a backend.
-Depending on the choice of spectral library, there may be systematic differences of 200 K or more in $T_{\rm eff}$ and 0.2 dex or more in $\Z$.
+%This behavior of the {\sc PHOENIX} library motivates the use of global and local covariance kernels in a sophisticated likelihood function.
+The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum, and acts as a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
+Including this covariance broadens the posteriors on the stellar parameters, while the local covariance kernels provide cushion against outlier lines.
+
+Figure~\ref{fig:PHOENIX_residuals} illustrates the approach in practice through a comparison of the 
+WASP-14 data and the {\sc Phoenix} models.  There are two notable differences in the residual 
+spectra presented there and in the corresponding Figure~\ref{fig:Kurucz_residuals} for the {\sc 
+CfA/Kurucz} models.  First, the amplitude of the global covariance structure is $\sim$2$\times$ 
+larger for the {\sc Phoenix} models: since that amplitude is an indirect parameterization of the 
+fit quality, this indicates that the {\sc Phoenix} models present an overall worse match to the 
+data.  Second, there are several local regions of strong residuals (outlier lines), which are 
+enveloped in a parametric treatment of the local covariance structure.  In effect, we have modeled 
+the large residuals with a suite of individual covariance kernels, with parameters that identify 
+and appropriately down-weight imperfect model spectral lines in a self-consistent manner.
+
+Aside from the structure of the covariance matrix, there is also a striking difference between the 
+posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
+CfA/Kurucz} model libraries: see Table~\ref{table:Kurucz} and compare 
+Figures~\ref{fig:Kurucz_posterior} and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find 
+a systematically cooler temperature (by $\sim$240\,K) and lower metallicity (by a factor of 
+$\sim$2) than inferred from the {\sc CfA/Kurucz} models.  Because our framework explicitly 
+mitigates the potential for posterior bias from imperfect matches between data and models, we can 
+safely conclude that these apparent discrepancies are intrinsic to the different input physics of 
+the model libraries.\footnote{It is worth noting that these differences are not a manifestation of 
+the stringent prior on surface gravity.  When the spectra are modeled with $\log g$ unconstrained, 
+the posteriors for both libraries shift, but their respective discrepancies actually increase 
+slightly.}  Such a comparison serves as a stark reminder that incomplete and/or over-simplified 
+physical prescriptions remain the most significant source of systematic uncertainty in estimates 
+of fundamental stellar parameters.
 
@@
 Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
-(IRTF).  SPEX is a cross-dispersed echelle that covers the red-optical to thermal-infrared spectrum 
-(0.7--5.5\,$\mu$m) in two spectrograph settings.  These data were obtained as part of the IRTF 
-spectral standard library project \citep{cushing05,rayner09}, and were processed through the 
-well-vetted {\tt Spextool} pipeline \citep{cushing04,vacca03} to deliver a fully-calibrated 
-spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.
+(IRTF).  SPEX is a cross-dispersed echelle spectrograph that covers the red-optical to 
+thermal-infrared spectrum (0.7--5.5\,$\mu$m) in two settings.  These data were obtained as part of 
+the IRTF spectral standard library project \citep{cushing05,rayner09}, and were processed through 
+the well-vetted {\tt Spextool} reduction pipeline \citep{cushing04,vacca03} to deliver a fully 
+calibrated spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.
 
@@
 than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
-opacities, as well as the presence of complex condensates (clouds) at the coolest temperatures.  
-\comm{should probably add some citations to that statement.  Make a comment recognizing that 
-various approaches have been taken, in context of Mann.}  \citet{rojas-ayala12} used a $K$-band 
-spectrum at similar resolution to the SPEX data to estimate the key Gl~51 stellar parameters based 
-on a line index technique.  They measured equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} 
-doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, along with a pseudo-continuum index defined over 
-three narrow passbands, 2.070--2.090, 2.235--2.255, and 2.360--2.380\,$\mu$m (see their Figure 2 
-for a useful visualization), designed to probe the spectral curvature induced by H$_2$O absorption 
-bands.  With reference to a custom grid of {\sc BT-Settl} spectra\footnote{From the web simulator at \url{http://phoenix.ens-lyon.fr/simulator/index.faces}}, \citet{rojas-ayala12} find that 
-these indices are consistent with $T_{\rm eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$.
-
-\begin{deluxetable}{lr@{$\pm$}l} 
-\tablecaption{\label{table:Gl51}Gl51 {\sc Phoenix} results}
-\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{Value}}
+opacities.  Especially confounding is the presence of complex condensates (clouds) at the coolest 
+temperatures \citep{allard13}, making it considerably more challenging to determine (sub-)stellar 
+properties \citep{rajpurohit14}.  Various approaches have been taken to infer stellar parameters in 
+the face of these difficulties, including iteratively masking regions with poor spectral agreement 
+\citep[e.g.,][]{mann13}.  Astutely, \citeauthor{mann13}~note that such a scheme may exclude 
+regions of the spectrum that contain intrinsically useful information for discriminating between 
+physical properties, and that a more sophisticated approach would weight each spectral region based 
+on its consistency with the data.  The modeling framework that we have constructed here does 
+exactly that.
+
+\begin{figure*}[!htb]
+  \includegraphics{figs/residuals_Gl51_logg.pdf}
+  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a {\sc Phoenix} model 
+({\it red}) generated by drawing parameters from the posterior distribution (for a fixed $\log g = 
+5.0$).  ({\it bottom}) The residual spectrum along with contours representing the distributions of 
+a large number of random draws from the covariance matrix (the shading is representative of the 1, 
+2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:}.  Note how the 
+``outlier" features are clearly identified and treated by the local covariance kernels.
+\label{fig:Gl51_residuals}}
+\end{figure*}
+
+\begin{figure}[!b]
+  \includegraphics[width=0.5\textwidth]{figs/Gl51_triangle.pdf}
+  \figcaption{The posterior probability distributions for the interesting stellar parameters 
+(marginalized over all other model parameters) of Gl~51 based on the {\sc Phoenix} model library, 
+as explored by the MCMC Gibbs sampler.  For the two-dimensional posterior projections, contours are 
+drawn at the standard 1, 2, and 3\,$\sigma$ levels (of a normal distribution) for reference.  
+\label{fig:Gl51_posterior}}
+\end{figure}
+
+As a point of reference, \citet{rojas-ayala12} used a $K$-band spectrum at similar resolution to 
+the SPEX data to estimate the Gl~51 stellar parameters using an index technique.  They measured 
+equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, 
+along with a pseudo-continuum index defined over three narrow passbands, 2.070--2.090, 
+2.235--2.255, and 2.360--2.380\,$\mu$m (see their Fig.~2 for a useful visualization), designed to 
+probe the spectral curvature induced by H$_2$O absorption bands.  With reference to a custom grid 
+of the {\sc BT-Settl} modifications of the {\sc Phoenix} library \citep{allard12}, 
+\citeauthor{rojas-ayala12}~infer $T_{\rm eff} = 3039\pm56$\,K and $[{\rm Fe/H}] = 0.28\pm0.17$ for 
+a fixed $\log g = 5.0$.
+
+By modeling the $K$-band region of the SPEX data with the {\sc Phoenix} library\footnote{The 
+differences between the \citet{husser13} {\sc Phoenix} library and the {\sc BT-Settl} modification 
+of the {\sc Phoenix} models at these (relatively) high effective temperatures are minor.  Since the 
+former is defined over a well-sampled, more regular grid of stellar parameters, we prefer it here 
+for computational simplicity.} and a $\delta$-function prior on $\log g$ (at 5.0), we infer 
+parameters for Gl~51 that are consistent with \citet{rojas-ayala12}.  
+Figure~\ref{fig:Gl51_residuals} displays the spectral modeling results, and 
+Figure~\ref{fig:Gl51_posterior} shows the marginal posterior distributions for the relevant stellar 
+parameters.  The best-fit parameter values and their associated uncertainties are compiled in 
+Table~\ref{table:Gl51}.
+
+We concur with \citet{rojas-ayala12} that the {\sc Phoenix} models manage to match the overall 
+spectral morphology of mid-M type stars well, but consistently under-predict the strengths of the 
+\ion{Na}{1} and \ion{Ca}{1} resonance lines, even for high metallicities.  \citet{rajpurohit10} 
+speculated that this discrepancy may be a consequence of inaccurate atomic data (oscillator 
+strengths and/or opacities).  \hili{The indexing technique adopted by \citeauthor{rojas-ayala12} 
+presumes that these discrepancies do not affect the line {\it ratios}}.  Our approach offers an 
+alternative that makes no such assumption, instead using local covariance kernels to 
+self-consistently determine the ``weights" of discrepant spectral lines in the overall fit.  In 
+that sense, these outlier features still bring their full information content to bear on the 
+posterior distributions of the stellar parameters without imposing a systematic bias.  
+Figure~\ref{fig:Gl51_residuals} demonstrates that these features have been identified as
+problematic, resulting in localized patches of appropriately inflated covariance.
+
+\begin{deluxetable}{lr@{ $\pm$ }l}[h]
+\tablecaption{\label{table:Gl51}Inferred Parameters for Gl 51}
+\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc Phoenix}}}
 \startdata
-\cutinhead{Stellar parameters}
-$T_{\rm eff}$ & 2998  & 17 K \\
-$\log_{10} g$ & \multicolumn{2}{c}{5.0 dex} \\
-$\Z$ & -0.01 & 0.04 dex \\
-$v \sin i$ & 5.68 & 4.15 $\kms$\\
-$v_z$ & 12.1 & 1.3 $\kms$ \\
-$\log_{10} \Omega$ & -19.628  & 0.003 \\
-\cutinhead{Nuisance parameters}
-$\log_{10} c_0$ & \multicolumn{2}{c}{0.0} \\
-$c_1$ & -0.011 & 0.002 \\
-$c_2$ & 0.005 & 0.002 \\
-$c_3$ & -0.002 & 0.001 \\
-$b$ & 1.56 & 0.14 \\
-$\log_{10} a_{\rm g} $ & -15.76 &  0.04 \\
-$l$ & 116 & 20 $\kms$ \\
-\cutinhead{Local kernel parameters}
-$\log_{10} a$ & -14.75 & 0.18 \\
-$\mu$ & 22055.9 & 0.8 \AA\\
-$\sigma$ & 122 & 16 $\kms$ \\
-\hline\\
-$\log_{10}a$ & -14.83 & 0.19 \\
-$\mu$ & 22083.1 & 0.9 \AA\\
-$\sigma$ & 120 & 17 $\kms$  \\
-\hline\\
-$\log_{10}a$ & -14.81 & 0.27 \\
-$\mu$ & 22609.2 & 1.3 \AA\\
-$\sigma$ & 85 & 26 $\kms$  \\
-\hline\\
-$\log_{10}a$ & -14.73 & 0.24 \\
-$\mu$ & 22625.2 & 1.1 \AA\\
-$\sigma$ & 94 & 24 $\kms$ \\
-\hline\\
-$\log_{10}a$ & -14.74 & 0.20 \\
-$\mu$ & 22652.2 & 0.8 \AA\\
-$\sigma$ & 99 & 19 $\kms$ \\
-\hline\\
-$\log_{10}a$ & -14.81 & 0.31 \\
-$\mu$ & 23380.1 & 1.0 \AA\\
-$\sigma$ & 73 & 27 $\kms$ \\
+$T_{\rm eff}$ (K) & 2998  & 17 \\
+$\log g$ & 5.0 & 0 (fixed) \\
+$\Z$ & -0.01 & 0.04 \\
+$v \sin i$ ($\kms$) & \hili{5.68} & \hili{4.15} \\
+$v_z$ ($\kms$) & 12.1 & 1.3 \\
+$\log \Omega$ (sr) & -19.628  & 0.003
 \enddata
+\tablecomments{Notation as in Table~\ref{table:Kurucz}.}
 \end{deluxetable}
 
-\comm{Now, present your results in a format that parallels the WASP-14 subsection.  You need a new 
-Table and a sequence of figures; introduce them all here.}
-
-\comm{Commentary on the necessity and utility of local covariance kernels here, with discussion of why Na and Ca lines are a pain.
-It is also worth commenting on the potential for flexibly extending the shape of
-the kernel to mimic molecular bandheads, if one wanted to try and fit say the TiO bands.}
-
-They note that the BT-Settl models manage to fit the overall spectral shape well, but consistently under-predict the strength of the \ion{Na}{1} and \ion{Ca}{1} lines, even when using the highest metallicity models.
-\citet{rajpurohit10} speculate that this discrepancy may result from inaccuracies in the oscillator strengths and opacities for the \ion{Na}{1} and \ion{Ca}{1} lines.
-When modelling the same region of spectrum as \citet{rojas-ayala12} (2.070--2.380$\mu$m) using an updated set of PHOENIX models \citep{husser13}, we find that \todo{report parameters}.
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/Gl51_triangle.pdf}
-  \caption{Gl51 stellar posterior.}
-\label{fig:gl_posterior}
-\end{center}
-\end{figure}
-
-\begin{figure*}[!htb]
-\begin{center}
-  \includegraphics{figs/residuals_Gl51_logg.pdf}
-  \caption{The Gl51 spectrum and residuals and many random draws from the final covariance matrix showing how the uncertainty in the ``outlier'' lines is identified by the local covariance kernels.}
-\label{fig:regions}
-\end{center}
-\end{figure*}
-
-\comm{A simple two-sentence (or so) wrap up on the future promise of this approach for late-type 
-stars and their complex spectra.}
-
-Complicated molecular structure is poorly known from models.
-
-\clearpage
+The fitting methodology adopted here could prove especially useful in spectroscopic inference for
+cool stars like Gl~51, where substantial uncertainties in their more complex atmospheres will 
+naturally produce systematic deviations between models and data.  However, many of 
+those discrepancies will be manifested in molecular features, which likely result in considerably 
+more complex residual structures than noted here \citep[e.g., the TiO bands in the red-optical; 
+see][their Fig.~9]{mann13}.  The overall framework we have employed should still function, although 
+more appropriate local covariance kernels may need to be developed to capture the different nature 
+of these outliers.  For example, one might employ hybrid kernels (like the product of a truncated 
+exponential and a \matern\ kernel) or empirically-motivated parametric shapes (e.g., a saw-tooth 
+pattern) to provide a better representation than a simple Gaussian feature.
 
@@
 
-\comm{It might seem silly, but every good discussion section needs to start with a recap.  This can 
-be tough to do concisely, but here's what I think the framework should be: (A1) a statement of the 
-problem (i.e., that we want to extract physical information from a spectrum without bias or 
-systematic error, but the models we have are not perfect); (A2) a [brief] reminder of the standard 
-approaches to skirt this issue; (A3) a summary of our approach; (A4) a statement that we demonstrated 
-the functionality of that approach in some different scenarios.  (est.~1 paragraph)} \\
-
-\comm{Next, more specifically highlight why we developed this framework, and what its 
-role should be in general spectroscopic inference.  I think its worth highlighting several key 
-advantages: (B1) you are casting the standard data selection/weighting/masking procedure in an 
-internally consistent, Bayesian format that appropriately preserves and propagates information 
-content even in the face of imperfect models (along the lines of what you've written in the 
-commented-out paragraphs below, but more concise/direct); (B2) (some deleted text could be useful here) by doing that, you allow the user to 
-actually exploit their full datasets, which should then reflect a more appropriate inference on the 
-parameters of interest without concern that some bad or less-informative features will screw them 
-up (sort of like what I commented out above, but less preachy) -- its in here 
-that we should suggest that this framework could be incorporated as a backend to standard methods, 
-where it essentially is employed to define an appropriate likelihood; and (B3) it creates an 
-opportunity to develop quantitative feedback to improve the models, or ultimately to 
-pursue data-driven modifications of the models (this last bit should be short, a basic summary, 
-since you will go into depth on it below). (est.~2 paragraphs)} \\
-
-A2: The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.
-Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight.
-A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit.
-The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters.
-In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are incorrect by a small amount actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.
-
-B1: The main advantage of using a more sophisticated hierarchical likelihood function (Eqn~\ref{eqn:likelihood}) than $\chi^2$, is that the model-data mismatch becomes a parameter in the model.
-When the model is of higher quality, the confidence in the stellar posteriors grows and is reflected by a smaller uncertainty.
-When the model is of lesser quality, the stellar posteriors inflate to reflect the increased uncertainty in the stellar parameters.
-This hierarchical model also works within a single spectrum when fitting more than one spectral order.
-If each order has a disparate prediction about the stellar parameters, then the global kernels will have a low amplitude and the final posterior will be an improved estimate.
-However, if certain orders predict different stellar parameters, the global kernels will inflate to absorb the model-data mismatch and consequently inflate the posteriors on the stellar parameters.
-
-\comm{Now you really want to launch into the data-driven model stuff.  As we discussed, I'd like to 
-see a short paragraph in here that speaks to the M dwarf folks about developing a bandhead-shaped 
-local covariance kernel to deal with particularly problematic molecular features (TiO).  Most of 
-what you've already written (commented out below) is in good shape, but you might decide to change 
-the format/organization given the suggested update in structure for this section.  Go out on a 
-positive note: the mechanics already exist, and this tool can be used to build better models!}
-
-The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum.
-If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves.
-By cataloguing the covariance structure of the residuals, especially those generated from strong spectral line mismatch, we collect valuable information about the quality of the synthetic spectra.
-After fitting many stars, the accumulated knowledge of the data-model mismatch can be used when fitting a new star.
-The previous structure of the covariance matrix allows us to set priors on what the covariance in certain regions of the spectrum should be, which will speed convergence for this new star.
-Additionally, after fitting several stars, the average value of the covariance matrix will inform us about the quality of specific spectral lines in the synthetic models.
-
-B3: Each stellar spectrum will have a slightly different realization of a spectral line, which will have some scatter about the average residual height.
-Linking the covariance matrix between spectra of similar stars allows us to grow more confident in our assessment that certain synthetic spectral lines are indeed outliers and should be appropriately down-weighted.
-Or, the covariance matrices could instead be linked hierarchically, in that the parameters describing the depth and width of a line residual for a specific star $a_1$ and $\sigma_1$, are modeled as coming from of a population of possible depths and widths for a given spectral type.
-In turn, as we become more certain of the weights, the stellar posteriors will become narrower and make our estimates of the stellar parameters more precise.
-This ability of the model to mutually inform sets of parameters is one of the major advantages of hierarchical Bayesian analysis \citep{kruschke10}.
-
-B3: Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models.
-Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution.
-This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models.
-Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.
-
-Infer the \emph{population characteristics} of poorly modeled spectral lines (mostly the typical width, amplitude of these lines).
-This will tell us about the frequency and distribution of spectral modelling errors.
-Or we could just speculate about it.
-Could also cross correlate with a catalogue of lines to determine which.
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines0.pdf}
-  \includegraphics{figs/badlines1.pdf}
-  \includegraphics{figs/badlines2.pdf}
-  \includegraphics{figs/badlines3.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    ( {\it first panel}) The majority of spectral lines ($\gtrsim 60$\%) will
-    have minor differences in strength between data and model spectrum, which
-    produce low-amplitude correlations in the residuals on the length scale of
-    the width of a typical spectral line.  ( {\it second panel}): Sometimes
-    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
-    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  ( {\it third panel}): Sometimes ($\lesssim
-    5$\% of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals. ({\it fourth panel}): If the line
-    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
-    there will be many correlated residuals of moderate amplitude. For any specific line, there might
-    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure}
+Astronomers exploit spectroscopy to retrieve physical information about their targets.  Ideally, 
+such inferences are made with the maximal precision afforded by the measurement noise, and do not 
+suffer from any systematic bias.  But in practice, the spectral models used as references are never 
+perfect representations.  Even modest mismatches between data and model can propagate substantial 
+systematic uncertainty into the inference problem.  In high-sensitivity applications (e.g., stellar 
+and exoplanetary astrophysics), ignoring these systematics can give a false sense of both precision 
+and accuracy in the inferences of key parameters.  Typically, the more egregious of these 
+imperfections are ``mitigated" by dismissal (explicitly not considering a subset of the data; e.g., 
+masking, clipping).  Rarely, they are confronted directly with painstaking, computationally 
+expensive fine-tuning of more general (nuisance) parameters in the model (e.g., oscillator 
+strengths, opacities), albeit only over a very limited spectral range and region of physical 
+parameter-space (e.g., the {\sc CfA/Kurucz} library; Sect.~\ref{subsec:wasp}).
+
+We have presented an alternative approach to dealing with this fundamental issue, grounded in a 
+generative, hierarchical Bayesian framework.  The method advocated here constructs a sophisticated 
+likelihood function, employing a non-trivial covariance matrix to treat the correlated 
+pixel-to-pixel residuals generated from intrinsically imperfect models.  That matrix is composed of 
+a linear combination of {\it global} (stationary) and {\it local} (non-stationary) Gaussian 
+Process kernels, which parameterize an overall mild covariance structure as well as small patches 
+of highly discrepant outlier features.  The approach we describe is generally applicable to any 
+spectroscopic inference problem (e.g., population synthesis in unresolved star clusters or 
+galaxies, physical/chemical models of emission line spectra in star-forming regions, etc.).  
+Moreover, it has the flexibility to incorporate additional information (as priors) or parametric 
+complexity (if desired), and could be deployed as a substitute for a simplistic $\chi^2$ metric in 
+already-established fitting tools (e.g., {\tt SME}, {\tt MOOG}).  To demonstrate how it is used, we 
+determined the surface parameters of main-sequence stars with mid-F and mid-M spectral types from 
+high-S/N optical and near-infrared data, with reference to pre-computed model libraries 
+(Sect.~\ref{sec:examples}).
+
+The novelty of employing this kind of likelihood function in the spectroscopic inference problem is 
+that the treatment of data--model mismatches (in essence, the fit quality) is explicitly built into 
+the forward-modeling framework.  This offers the unique advantage that discrepant spectral features 
+(outliers), which may contain substantial (even crucial) information about the parameters of 
+interest, can still effectively propagate their useful information content into the posteriors with 
+a weighting that is determined self-consistently.  From a practical standpoint, this means that 
+{\it all of the data} can be employed in the inference problem without undue concern that the 
+posteriors will be biased by model imperfections, even in the case of models with limited 
+parametric flexibility (e.g., pre-computed spectral libraries).
+
+Ultimately, the benefits of employing covariance kernels to accommodate imperfect models could 
+be extended well beyond modeling the spectra of individual targets.  In principle, the approach we 
+have described here can be used to systematically discover and quantify imperfections in spectral 
+models and eventually to build data-driven improvements of those models that are more appropriate 
+for spectroscopic inference.  Consider the specific application to stellar spectroscopy, which has 
+been the focus here.  By fitting many stellar spectra with the same family of models, we can 
+catalog the covariant structure of the fit residuals -- especially the parameters of the local 
+covariance kernels -- to collate quantitative information about where and how the models tend to 
+deviate from observational reality.  That information can be passed to the spectral synthesis 
+community, in some cases enabling modifications that will improve the quality of the spectral 
+models.  On a large enough scale, this feedback between observers and modelers could be used to 
+refine inputs like atomic and molecular data (oscillator strengths, opacities), elemental abundance 
+patterns, and perhaps the stellar atmosphere structures.
+
+Alternatively, this kind of feedback could be used to make data-driven modifications to the already 
+existing models, creating a new library of semi-empirical spectral models.  This could be 
+accomplished by linking the parameters of the covariance kernels derived from fitting many spectra 
+in a hierarchical Bayesian model, which would add confidence to the assessment that certain 
+spectral features are {\it systematic} outliers and offer general quantitative guidance on how to 
+weight them in the likelihood calculation.  Rather than simply assembling an empirical spectral 
+library using only observations, this combined machine-learning approach would naturally provide a 
+physical anchoring for the key physical parameters, since they are reflected in the spectra based 
+on the physical assumptions in the original models.  This kind of large-scale analysis holds great 
+promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
+like those being collected in programs like the APOGEE and HERMES surveys), since they provide 
+enormous leverage for identifying and improving the underlying model systematics.
+
+\comm{Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
+include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
+linked to that, since there is a popular Herschel HERMES program that will confuse people...}
+
+\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
+Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
+acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
+for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
 
@@
 
-%\section{Tables}
-%
-%\begin{deluxetable}{ll}
-%\tablecaption{Nomenclature used in this document}
-%\tablehead{
-%  \colhead{Symbol} & \colhead{Description}
-%}
-%\startdata
-%$\lambda_i$ & wavelength corresponding to a given pixel $i$\\
-%$\vt_{\ast}$ & fundamental stellar parameters, $T_{\rm eff}, \log(g), \Z, \A$\\
-%$\{\vt_{\ast}\}^{\rm grid}$ & stellar parameters that specify a spectrum from a synthetic library\\
-%$\vt_{\rm obs}$ & stellar parameters $v \sin i$, $v_z$, $A_V$, and $R^2/d^2$ that\\
-%  & are applied during ``post processing'' of the synthetic spectrum\\
-%$\vt$ & $\{\vt_{\ast},\vt_{\rm obs}\}$, all stellar parameters\\
-%
-%
-%$\chebi{i}$ & $\{c_0, c_1, \ldots, c_{n-1}\}$ the set of Chebyshev polynomial coefficients  for \\
-% & a single spectral order $i$ \\
-%$\Cheb$ & $\{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$, the collection of all Chebyshev parameters for \\
-%  & a multi-order echelle spectrum with $N$ spectral orders  \\
-% $\Chebi{2}$ & $\{\chebi{1}, \chebi{3}, \ldots, \chebi{N}\}$, the collection of all Chebyshev parameters \\
-% & except those for order $2$ \\
-%$\covi{i}$ & $[a_g, \ell, \{a_k, \mu_k, \sigma_k, h_k\}^{N_{\rm reg}}]$, the set of covariance hyperparameters for a single spectral order $i$ \\
-%$\Cov$ & $\{ \covi{1}, \covi{2}, \ldots, \covi{N} \}$, the collection of all covariance hyper-parameters for \\
-% & a multi-order echelle spectrum with $N$ spectral orders  \\
-% $\Covi{2}$ & $\{\covi{1}, \covi{3}, \ldots, \covi{N}\}$, the collection of all covariance hyperparameters \\
-% & except those for order $2$ \\
-%
-%$\vP$ & $\{\Cheb, \Cov \}$, the collection of all nuisance parameters \\
-%$\vT$ & the collection of all stellar parameters \\
-%$\vD$ & data spectrum \\
-%$\vM$ & model, a function of the above stellar, Chebyshev, and covariance (hyper)parameters \\
-%$\vR$ & $\vD - \vM $, residuals  \\
-%$\vC$ & covariance matrix, a function of the covariance hyper-parameters\\
-%$\vC_{ij}$ & an individual element in the covariance matrix\\
-%$r(\lambda_i, \lambda_j)$ & radial distance in wavelength space corresponding to $\Delta v$\\
-%
-%$\Kglobal$ & global covariance kernel\\
-%$\Klocal$ & local covariance kernel\\
-%$w$ & Hanning taper for kernels \\
-%\enddata
-%\end{deluxetable}
-
-\appendix
-
 \end{document}

### 2014-08-27 tex/paper.tex
Figure updates

@@
 
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics{figs/residuals_Kurucz_logg.pdf}
+\caption{A sample of the WASP-14 data set overlaid with a model spectrum produced by parameters drawn from the posterior probability distribution (fixing $\log_{10} g = 4.29$ dex) and using the {\sc CfA/Kurucz} synthetic library.
+The residuals are shown zoomed along with the three-sigma contours of random draws from the covariance matrix, after Figure~\ref{fig:matrix}.}
+\label{fig:Kurucz_residuals}
+\end{center}
+\end{figure*}
+
 %\begin{figure}[!htb]

### 2014-08-29 StellarSpectra/grid_tools.py
About to start error spec interpolation

@@ -634,6 +634,8 @@ class HDF5Interface:
 
         :raises KeyError: if spectrum is not found in the HDF5 file.
         '''
+        print("loading {}".format(parameters))
+
         key = self.flux_name.format(**parameters)
         with h5py.File(self.filename, "r") as hdf5:
             if self.ind is not None:
@@ -660,6 +662,54 @@ class HDF5Interface:
         #Note: will raise a KeyError if the file is not found.
 
         return (fl, hdr)
+
+    def get_error_spectrum(self, parameters, axis="temp"):
+        '''
+        Given a grid point specified by parameters, and a direction on the axis, return the error spectrum.
+        '''
+        #load the file specified by the parameters
+        fl = self.load_flux(parameters)
+
+        #identify the axis
+        points = self.points[axis]
+        print(points)
+        #Find which index this point is
+        idx = (np.abs(points - parameters[axis])).argmin()
+
+        #Get the two nearest spectra in the grid, as long as we aren't already at a grid edge
+        #Better checking here, because it overlaps
+        try:
+            low = points[idx - 1]
+            high = points[idx + 1]
+        except KeyError:
+            return None
+
+        #load the spectra that bracket either side of these
+        parameters.update({axis:low})
+        fl_low = self.load_flux(parameters)
+        parameters.update({axis:high})
+        fl_high = self.load_flux(parameters)
+
+        #compare the average of the two side spectra to the central spectrum and return this as the error spectrum
+        errspec = fl - (fl_low + fl_high)/2
+        return errspec
+
+
+    def get_error_spectra(self, parameters):
+        '''
+        Do get_error_spectrum but for each one of the parameters in the grid
+        '''
+        return [self.get_error_spectrum(parameters, axis) for axis in ("temp", "logg", "Z")]
+
+
+#We're going to need a couple methods:
+
+#1 Given a grid point, and an axis (one of temp, logg, Z), return an error spectrum
+# addressed as a function of HDF5Interface?
+
+#2 Create an HDF5 file with all three error spectra for each grid point
+
+#3 Use the error spectra along with the degree of interpolation to return an estimate of the interpolation
 
 class HDF5InstGridCreator:
     '''
@@ -999,6 +1049,7 @@ class Interpolator:
         :param parameters: stellar parameters
         :type parameters: dict
[…]

### 2014-09-07 StellarSpectra/spectrum.py
Last commit before rewriting the C/cython routines

@@ -793,7 +793,9 @@ class ModelSpectrum:
         self.downsampled_fls = np.empty(self.DataSpectrum.shape)
         self.downsampled_fls_last = self.downsampled_fls
 
-        self.downsampled_errors = np.empty((24,) + self.DataSpectrum.shape)
+        #self.downsampled_errors = np.zeros((24,) + self.DataSpectrum.shape)
+        self.downsampled_errors = None
+        self.other_downsampled_errors = None
 
         self.grid_params = {}
         self.vz = 0
@@ -1006,6 +1008,7 @@ class ModelSpectrum:
         :returns fls: the downsampled fluxes that has the same shape as DataSpectrum.fls
         '''
 
+        #print("inside downsample")
         #Check to make sure that the grid is within self.wl range, because a spline will not naturally raise an error!
         wls = self.DataSpectrum.wls.flatten()
         if min(wls) < min(self.wl) or max(wls) > max(self.wl):
@@ -1018,13 +1021,34 @@ class ModelSpectrum:
         self.downsampled_fls = np.reshape(interp(wls), self.DataSpectrum.shape)
         del interp
 
+        #if self.downsampled_errors is not None:
+        #    self.downsampled_errors_last = np.copy(self.downsampled_errors)
+
+        # if self.other_downsampled_errors is None:
+        #     print("setting other downsampled errors")
+        #     self.other_downsampled_errors = np.zeros((24,) + self.DataSpectrum.shape)
+        #     for i, errspec in enumerate(self.errors):
+        #         interp = InterpolatedUnivariateSpline(self.wl, errspec, k=5)
+        #
+        #         self.other_downsampled_errors[i, :, :] = np.reshape(interp(wls), self.DataSpectrum.shape)
+        #         del interp
+
         #Interpolate each of the 24 error spectra to the grid points
-        self.downsampled_errors_last = self.downsampled_errors.copy()
+        #if self.downsampled_errors is None:
+        self.downsampled_errors = np.zeros((24,) + self.DataSpectrum.shape)
+        print("setting downsampled errors")
         for i, errspec in enumerate(self.errors):
             interp = InterpolatedUnivariateSpline(self.wl, errspec, k=5)
 
             self.downsampled_errors[i, :, :] = np.reshape(interp(wls), self.DataSpectrum.shape)
             del interp
+
+        #assert np.allclose(self.downsampled_errors, self.other_downsampled_errors), "No longer downsampling the same " \
[…]

### 2014-09-07 scripts/stars/base_lnprob.py

@@ -157,11 +157,11 @@ region_priors = config['region_priors']
 model_list = [
     Model(myDataSpectrum, myInstrument, HDF5Interface(config['HDF5_path']), stellar_tuple=stellar_tuple,
           cheb_tuple=cheb_tuple, cov_tuple=cov_tuple, region_tuple=region_tuple, outdir=outdir,
-        debug=False)
+          max_v=config["interp_cut"], debug=False)
     for myDataSpectrum in myDataSpectra]
 
 myStellarSampler = StellarSampler(model_list=model_list, cov=stellar_MH_cov, starting_param_dict=stellar_Starting,
-                                  fix_logg=fix_logg, outdir=outdir, debug=False)
+                                  fix_logg=fix_logg, outdir=outdir, debug=True)
 
 samplerList = []
 #Create the samplers for each DataSpectrum
@@ -169,14 +169,14 @@ for model_index in range(len(model_list)):
     #Create the three subsamplers for each order
     for order_index in range(len(config['orders'])):
         samplerList.append(ChebSampler(model_list=model_list, model_index=model_index, cov=cheb_MH_cov,
-            starting_param_dict=cheb_Starting, order_index=order_index, outdir=outdir, debug=False))
+            starting_param_dict=cheb_Starting, order_index=order_index, outdir=outdir, debug=True))
         if not config['no_cov']:
             samplerList.append(CovGlobalSampler(model_list=model_list, model_index=model_index, cov=cov_MH_cov,
-            starting_param_dict=cov_Starting, order_index=order_index, outdir=outdir, debug=False))
+            starting_param_dict=cov_Starting, order_index=order_index, outdir=outdir, debug=True))
             if not config['no_region']:
                 samplerList.append(RegionsSampler(model_list=model_list, model_index=model_index, cov=region_MH_cov,
                 default_param_dict=region_Starting, priors=region_priors, order_index=order_index, outdir=outdir,
-                                                                                                       debug=False))
+                                                                                                       debug=True))
 
 mySampler = MegaSampler(samplers=[myStellarSampler] + samplerList, debug=False)
 
@@ -200,8 +200,14 @@ def main():
     print(mySampler.acor)
     for i, model in enumerate(model_list):
         model.to_json("model{}_final.json".format(i))
-        np.save("residuals.npy", model.OrderModels[0].get_residual_array())
[…]

### 2014-09-07 StellarSpectra/grid_tools.py
Last commit before rewriting the C/cython routines

@@ -1182,7 +1182,8 @@ class Interpolator:
             self.fluxes[i,:] = self.cache[key]*weight_list[i]
             spec_tuple = self.cache_errors[key]
             for j, spec in enumerate(spec_tuple):
-                self.errors[i*3 + j,:] = spec*weight_sq[i] * Weight_list[j] #Do not square these weights here, since
+                self.errors[i*3 + j,:] = spec * weight_sq[i] * Weight_list[j] #Do not square these weights here, since
+                #self.errors[i*3 + j,:] = spec*0.204124 #Do not square these weights here, since
                 #the routine for adding them to the sparse matrix will automatically square them.
 
         return np.sum(self.fluxes, axis=0), self.errors

### 2014-09-07 tex/paper.tex
Last commit before rewriting the C/cython routines

--- 2014-08-27 tex/paper.tex
@@
 average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
-the systematic spread between the best-fit parameter values. \comm{This is something we need to 
-discuss more with Dave and Willie.}
+the systematic spread between the best-fit parameter values. 
+%\comm{This is something we need to discuss more with Dave and Willie.}
 
@@
 
-\comm{Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
-include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
-linked to that, since there is a popular Herschel HERMES program that will confuse people...}
-
-\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
-Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
-acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
-for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+%\comm{Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
+%include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
+%linked to that, since there is a popular Herschel HERMES program that will confuse people...}
+
+%\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
+%Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
+%acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
+%for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+%
 

### 2014-09-09 tex/paper.tex
Work on text, added abstract and appendix

@@
 \begin{abstract}
-An abstract needs to go here. \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd \\
-tbd
+We present a modular and extensible framework for forward-modelling spectra with synthetic models.
+Due to the nature of spectra, that adjacent pixels are not independent samples of the intrinsic flux distribution, it is common to result in residual spectra that exhibit some degree of correlation.
+Neglecting this covariance in the residual spectrum can result in a biased estimates of the parameters which describe the model spectrum.
+In the high signal-to noise, large spectral range limit common to stellar parameter estimation, the covariant structure of the residuals can bias the resulting parameter determinations.
+We present a likelihood function designed to account for this covariant structure, utilizing the machinery of Gaussian process covariance kernels to self-consistently identify and model the covariance.
+Specific to stellar parameter estimation, a common problem is the presence of spectral lines with an incorrect strength, often due to incorrect atomic constants or radiative treatment.
+We also present the development of a local covariance kernel that can track and self-consistently down-weight pathologically wrong spectra lines.
+Local kernels that track line weights provide a mechanism to learn corrections to spectral models, by fitting in a hierarchical manner.
+We demonstrate the application of this method, implemented as an open source code, to fit an echelle optical spectrum WASP-14, an F star, and a mid-infra spectrum of Gl51, an M5 star.
 \end{abstract}
@@
 
+\appendix
+
+\section{Interpolation Errors}
+Interpolating a model spectrum from a synthetic library of spectra on a grid is conceptually straightforward; however, in practice there are several subtle effects to consider.
+Spectra are stored in a synthetic library as samples of the intensity at specific wavelengths, or pixels, at high resolution (typically R$\sim$500,000).
+Typically, to interpolate a spectrum for a specific.
+
+Moreover, often the flux value of the pixel will not vary monontonically.
+Visually inspecting values of pixels as a function of parameter values makes it apparent that the spectra have not be Nyquist sampled in parameter space.
+That the synthetic grid does not contain the full amount of information about the spectral behavior.
+However, this is not surprising giving the computational complexity of stellar synthesis.
+Therefore, we are left to make the most conservative assumption about the behavior of the pixels between grid points and decide on a linear interpolation scheme.
+Going beyond this behavior, for example using a spline or other band-limited interpolation scheme, ignores the fact that the grid itself is not Nyquist sampled.
+
+We choose to do linear interpolation, but empirically determine the interpolation error by a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, for example T=6000, logg=4.5, and Z=-0.5, and then create three sets of interpolated spectra, over the various dimensions.
+Therefore, for each spectrum in the synthetic grid, there are also three error spectra Generally, the interpolation error is small, typically no larger than 1\% across many lines.
+This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself.
+However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}.
+When the signal to noise of the spectrum is particularly large, this coherent trend can manifest itself as a behavior that the spectral lines in an interpolated spectrum are less ``physical'' looking than those directly near a gridpoint, with minor or no interpolation.
+Therefore, it is necessary to quantify the degree of interpolation error, in order to give the proper likelihood to interpolated spectra.
+
+[Figure here of actuall interpolation error, as alluded to in text].
+
+[Possible second figure showing a 3D cube, with vertices labelled].
+
+Because the interpolation error is covariant, i.e., the interpolation error spectrum is highly dependent on whether the pixel is on the continuum, or on a spectral line (and specifically where on the spectral line), we must model the covariance of the errors.
+We must multiply and then add in order to preserve the covariance.
+A simple sum would destroy the covariance and the effect.
+
+For computational expediency and numerical stability, we smoothly taper the covariance to zero after a similar length scale as the global covariance kernel.
+In reality, this interpolation error matrix is dense: pixels separated by a wide spectral range may in fact exhibit a strong degree of correlation because they happen to fall on the same relative location of a spectral line.
+For example, all pixels at the continuum will tend to act similarly, regardless of where on the spectrum the pixels fall.
+Any pixel that falls on the blue side of a spectral line will tend to exhibit a similar error behaviour regardless of which spectral line, as long as that line was changing in a similar manner.
+
+The elements of the interpolation covariance matrix $\mathsf{C}_{ij}^\textrm{int}$ are determined empirically from the drop-out spectra, and are a function of the degree of the interpolation.
+
+\begin{equation}
+  \mathsf{C}_{ij}^\textrm{int} = 
+\end{equation}
+
+[Figure showing what the truncated covariance matrix looks like, with high values about lines].
+
 \end{document}

### 2014-09-10 tex/paper.tex
more orders

@@
 Therefore, for each spectrum in the synthetic grid, there are also three error spectra Generally, the interpolation error is small, typically no larger than 1\% across many lines.
-This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself.
+This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit.
 However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}.
@@
 
-[Figure here of actuall interpolation error, as alluded to in text].
-
-[Possible second figure showing a 3D cube, with vertices labelled].
+\begin{figure}[!t]
+\begin{center}
+  \includegraphics{figs/interp_errors.pdf}
+  \figcaption{Also showing the envelope for a two gridpoint interpolation. All errors measured relative to the value of the true spectrum.}
+  \label{fig:interp_errors}
+\end{center}
+\end{figure}
+
+Nearly all interpolation errors are below the spread in change for an interpolated value.
+Those few pixels for which the change in value is \emph{less} than the bracketed change are due to values for which the pixel value changes non-monotonically with parameters.
+In such a situation, because the synthetic grid is not Nyquist sampled, it is difficult to accurately predict what the true pixel value should be.
+In this case, the most conservative approach is to incorporate the large interpolation error budget, as we do.
+
+\begin{figure}[!t]
+\begin{center}
+  \includegraphics{figs/interp_grid.pdf}
+  \figcaption{Tri-linear interpolation on a grid cell the shape of a rectangular prism.}
+  \label{fig:interp_grid}
+\end{center}
+\end{figure}
 

### 2014-09-10 tex/paper.tex
Removed reset after burn in

@@
 The elements of the interpolation covariance matrix $\mathsf{C}_{ij}^\textrm{int}$ are determined empirically from the drop-out spectra, and are a function of the degree of the interpolation.
+The variables are notated such that the last index changes most rapidly.
+
+We could represent the average covariance matrix, for a specific axis $k$, completed from all of the grid points as
 
 \begin{equation}
-  \mathsf{C}_{ij}^\textrm{int} = 
-\end{equation}
+  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  b_{kl} \lambda_{ikl} \lambda_{jkl}
+\end{equation}
+Then, the final combined matrix requires some weighting based upon interpolation distance in each of the three axes, $k$
+\begin{equation}
+  \mathsf{C}_{ij}^\textrm{int} = w_{ij}(r_0) \sum_{k=1}^3 a_k \mathsf{C}_{ijk}^\textrm{int}
+\end{equation}
+We also add in a Hann taper to ensure sparsity of the covariance matrix for computational efficiency, however note that the interpolation error covariance matrix is itself dense.
+
+Here, $b_{kl}$ corresponds to the same weights that would be applied to each spectra in the interpolation case, for a specific axis $k$.
+The estimate of covariance is determined as a weighted average between the covariance matrices at all of the grid points.
+
+Then, the final covariance
 

### 2014-09-10 tex/paper.tex
Appendix edits

@@
 \section{Interpolation Errors}
-Interpolating a model spectrum from a synthetic library of spectra on a grid is conceptually straightforward; however, in practice there are several subtle effects to consider.
-Spectra are stored in a synthetic library as samples of the intensity at specific wavelengths, or pixels, at high resolution (typically R$\sim$500,000).
-Typically, to interpolate a spectrum for a specific.
-
-Moreover, often the flux value of the pixel will not vary monontonically.
-Visually inspecting values of pixels as a function of parameter values makes it apparent that the spectra have not be Nyquist sampled in parameter space.
-That the synthetic grid does not contain the full amount of information about the spectral behavior.
-However, this is not surprising giving the computational complexity of stellar synthesis.
-Therefore, we are left to make the most conservative assumption about the behavior of the pixels between grid points and decide on a linear interpolation scheme.
-Going beyond this behavior, for example using a spline or other band-limited interpolation scheme, ignores the fact that the grid itself is not Nyquist sampled.
-
-We choose to do linear interpolation, but empirically determine the interpolation error by a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, for example T=6000, logg=4.5, and Z=-0.5, and then create three sets of interpolated spectra, over the various dimensions.
-Therefore, for each spectrum in the synthetic grid, there are also three error spectra Generally, the interpolation error is small, typically no larger than 1\% across many lines.
+The interpolation of a model spectrum from a synthetic library of spectra is a conceptually straightforward process; however, in practice there may be several subtle effects that require consideration, depending on the application.
+Spectra are generally stored in a synthetic library as 1-D arrays of flux values, sampled at specific wavelengths at high resolution (typically R$\gtrsim$100,000).
+Each spectrum has several fundamental parameters which describe the model that generated it.
+To use a concrete example, the stellar libraries used in this paper have three parameters, $T_\textrm{eff}$, $\log_{10} g$, and $\Z$, each of which constitutes an axis in the synthetic library.
+Therefore, the full synthetic grid can be stored as a 4-D array, where the first three axes are the stellar parameters and the 4th axis describes the wavelength.
+The value stored in the array is the flux of the spectrum described by current value of the four axes.
+
+Spectral line formation is a non-linear process, with the shape of the line determined by complicated radiative synthesis, which is a function of the stellar parameters.
+Interpolating a \emph{physical} spectrum short-cuts this non-linear process and attempts to predict the behavior of the spectral line based upon the behavior of nearby spectra in the grid.
+While this approximation is generally accurate enough for our applications, for high signal-to-noise applications it becomes necessary to account for any inaccuracies introduced by the interpolation.
+In practice, there are two factors which influence the choice of interpolation method.
+First, for a small percentage of pixels, the flux value will not vary monotonically with change in fundamental parameters.
+Second, visual inspection of the pixel values as a function of fundamental parameters makes it apparent that the spectra are not Nyquist sampled in parameter space.
+This means that synthetic grid does is not dense enough to losslessly reconstruct any synthetic spectrum.
+However, this fact is not surprising given the computational complexity of stellar synthesis.
+Given this reality, we make the most conservative assumption about the behavior of the pixels between grid points and decide to use a linear interpolation scheme.
+Using a more advanced interpolation method, such as a spline or other band-limited interpolation scheme, would not remedy the fact that the grid itself is not Nyquist sampled.
+To account for the loss of spectral information during the interpolation process, we build a model of interpolation error and incorporate this into the likelihood function (Eqn.~\ref{eqn:lnlikelihood}).
+
+We empirically determine the interpolation error by performing a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, generate an interpolated spectrum from the nearest two grid points, and compare the difference between the two.
+For example, consider the {\sc Phoenix} grid used for the WASP-14 parameter estimation.
+In this parameter space, the grid is sampled every 100K in $T_\textrm{eff}$ and every 0.5 dex in $\log_{10} g$ and $\Z$.
+Therefore, for each spectrum in the synthetic grid, there are also three error spectra associated with it, one for each axis of stellar parameters.
+For example, consider a spectrum with stellar parameters $T_\textrm{eff} = $ 6000K, $\log_{10} g=$4.0 dex, and $\Z = $ -0.5 dex.
+A temperature ``error spectrum'' would be formed by subtracting this true spectrum from the spectrum formed by the average of the $T_\textrm{eff} = $ 6100K spectrum and the $T_\textrm{eff} = $ 5900K spectrum, with $\log_{10} g$ and $\Z$ held constant.
+Likewise, the $\log_{10} g$ and $\Z$ error spectra are created by performing the same operation over the $\log_{10} g$ and $\Z$ axes of the grid (see Figure~\ref{fig:interp_errors}).
+
+These are done over a 200K range, and the actual range will only be 100K.
+Tests with other interpolation show that the interpolation error increases approximately linearly with interpolation distance.
+Therefore, when using the spectrum, we divide the total interpolation error spectrum by 2.
+
+Generally, the interpolation error is small, typically no larger than 1\% across many lines.
 This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit.

### 2014-09-11 StellarSpectra/model.py
draft on interpolation errors roughly complete

@@ -448,6 +448,9 @@ class OrderModel:
     def get_Cov(self):
         return self.CovarianceMatrix.cholmod_to_scipy_sparse()
 
+    def get_Cov_interp(self):
+        return self.CovarianceMatrix.cholmod_to_scipy_sparse_interp()
+
     def get_regions_dict(self):
         return self.CovarianceMatrix.get_regions_dict()
 

### 2014-09-11 tex/paper.tex
draft on interpolation errors roughly complete

@@
 Likewise, the $\log_{10} g$ and $\Z$ error spectra are created by performing the same operation over the $\log_{10} g$ and $\Z$ axes of the grid (see Figure~\ref{fig:interp_errors}).
-
-These are done over a 200K range, and the actual range will only be 100K.
-Tests with other interpolation show that the interpolation error increases approximately linearly with interpolation distance.
-Therefore, when using the spectrum, we divide the total interpolation error spectrum by 2.
-
-Generally, the interpolation error is small, typically no larger than 1\% across many lines.
-This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit.
-However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}.
-When the signal to noise of the spectrum is particularly large, this coherent trend can manifest itself as a behavior that the spectral lines in an interpolated spectrum are less ``physical'' looking than those directly near a gridpoint, with minor or no interpolation.
-Therefore, it is necessary to quantify the degree of interpolation error, in order to give the proper likelihood to interpolated spectra.
+Because the error spectra are determined over an interpolation distance twice that (e.g., 100K) of the actual distance used (e.g., 50K), we reduce the amplitude of the error spectra by two.
+Drop-out tests at larger interpolation distances (e.g., 200K and 300K) show that the amplitude of the error spectrum grows approximately linearly with interpolation distance.
 
@@
   \includegraphics{figs/interp_errors.pdf}
-  \figcaption{Also showing the envelope for a two gridpoint interpolation. All errors measured relative to the value of the true spectrum.}
+  \figcaption{(\emph{top}) A small range of the ``true'' {\sc Phoenix} spectrum at $T_\textrm{eff} = $ 6000K, $\log_{10} g=$4.0 dex, and $\Z = $ -0.5 dex, shown with the drop-out interpolation spectra along each of the three parameter axes.
+These interpolation error spectra are determined over an interpolation distance twice that (e.g., 100K) of the actual distance used in practice.
+Typically, the interpolated spectra closely match the ``true'' spectrum.
+(\emph{bottom}) The three ``error spectra'', or difference between the interpolated spectrum and the true spectrum, relative to the true spectrum value.
+The shaded region shows the total change in flux between the two spectra used for interpolation, relative to the true spectrum.
+Note that the error spectrum is generally much smaller than the typical change in flux.
+For the few points where this is not true (e.g., $\log_{10} g$ near 5161.5\AA) this implies a non-monotonic change in flux between grid points.
+Therefore, it is best to conservatively use the error spectrum, which protects against assumptions of monotonicity.}
   \label{fig:interp_errors}
@@
 
-Nearly all interpolation errors are below the spread in change for an interpolated value.
-Those few pixels for which the change in value is \emph{less} than the bracketed change are due to values for which the pixel value changes non-monotonically with parameters.
-In such a situation, because the synthetic grid is not Nyquist sampled, it is difficult to accurately predict what the true pixel value should be.
-In this case, the most conservative approach is to incorporate the large interpolation error budget, as we do.
+Generally, the interpolation error is small: typically no larger than 2\% of the flux and usually--but not always--much smaller than the total flux difference between spectra at adjacent grid points (see Figure~\ref{fig:interp_errors}).
+This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the large number of possible lines to fit.
+However, while the interpolation error is small, because all spectral lines have roughly the same shape, it is often true that the error adds \emph{coherently}.
+When the signal to noise of the spectrum is particularly large, this coherent trend can contribute substantially by diminishing the likelihood of an interpolated spectrum, since the lines in the spectrum are less ``physical'' in shape than that of a spectrum directly on a gridpoint, which requires minor or no interpolation.
+Therefore in order to deliver the proper likelihood for interpolated spectra it is necessary to quantify the degree of interpolation error and add this into the covariance matrix.
 
@@
 
-Because the interpolation error is covariant, i.e., the interpolation error spectrum is highly dependent on whether the pixel is on the continuum, or on a spectral line (and specifically where on the spectral line), we must model the covariance of the errors.
-We must multiply and then add in order to preserve the covariance.
-A simple sum would destroy the covariance and the effect.
-
-For computational expediency and numerical stability, we smoothly taper the covariance to zero after a similar length scale as the global covariance kernel.
-In reality, this interpolation error matrix is dense: pixels separated by a wide spectral range may in fact exhibit a strong degree of correlation because they happen to fall on the same relative location of a spectral line.
+Because the interpolation error is highly dependent on whether the wavelength is on the continuum or on a spectral line, we must model the covariance of the errors.
 For example, all pixels at the continuum will tend to act similarly, regardless of where on the spectrum the pixels fall.
 Any pixel that falls on the blue side of a spectral line will tend to exhibit a similar error behaviour regardless of which spectral line, as long as that line was changing in a similar manner.
-
 The elements of the interpolation covariance matrix $\mathsf{C}_{ij}^\textrm{int}$ are determined empirically from the drop-out spectra, and are a function of the degree of the interpolation.
-The variables are notated such that the last index changes most rapidly.
-
-We could represent the average covariance matrix, for a specific axis $k$, completed from all of the grid points as
-
+Consider the manner in which a typical spectrum is tri-linearly interpolated from the synthetic grid (Figure~\ref{fig:interp_grid}).
+The interpolated spectrum is a weighted average of the neighboring 8 spectra, with weights $d_l$ based upon the distance from the grid points.
+To generate a composite covariance matrix which reflects the interpolation error present for the interpolated spectra, we must also combine the error spectra.
+Because the error is covariant, we must first multiply and then add the error spectra, in order to preserve the covariance.
+A simple sum of error spectra would destroy any information about the covariance.
+For a specific axis $k$ (e.g., $T_\textrm{eff}$, $\log_{10} g$, or $\Z$), we represent the average covariance matrix as a composite of all of the error spectra of this species
 \begin{equation}
-  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  b_{kl} \lambda_{ikl} \lambda_{jkl}
-\end{equation}
-Then, the final combined matrix requires some weighting based upon interpolation distance in each of the three axes, $k$
+  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  d_{kl} \lambda_{ikl} \lambda_{jkl}
+\end{equation}
+
+We combine the covariance matrices of each axis into a total interpolation covariance matrix by adding them together, with weights $g_k$ based upon the total interpolation distance in that axis.
+For example, if we were to interpolate a spectrum with $T_\textrm{eff} = 6120$K, then $g_{T_\textrm{eff}} = 0.4$.
+$g_{T_\textrm{eff}} = 0.0$ would signify no interpolation distance (0K, $T_\textrm{eff} = 6100$K) while $g_{T_\textrm{eff}} = 1.0$ would signify maximal interpolation distance (50K, $T_\textrm{eff} = 6150$K).
+Therefore, the final interpolation covariance matrix is
 \begin{equation}
-  \mathsf{C}_{ij}^\textrm{int} = w_{ij}(r_0) \sum_{k=1}^3 a_k \mathsf{C}_{ijk}^\textrm{int}
-\end{equation}
-We also add in a Hann taper to ensure sparsity of the covariance matrix for computational efficiency, however note that the interpolation error covariance matrix is itself dense.
-
-Here, $b_{kl}$ corresponds to the same weights that would be applied to each spectra in the interpolation case, for a specific axis $k$.
-The estimate of covariance is determined as a weighted average between the covariance matrices at all of the grid points.
-
-Then, the final covariance
-
-[Figure showing what the truncated covariance matrix looks like, with high values about lines].
+  \mathsf{C}_{ij}^\textrm{int} = w_{ij}(r_0) \sum_{k=1}^3 g_k \mathsf{C}_{ijk}^\textrm{int}
+\end{equation}
+
+For computational expediency and numerical stability, we smoothly taper the covariance to zero after a similar length scale as the global covariance kernel.
+In reality, the interpolation error matrix is dense: pixels separated by a wide spectral range may in fact exhibit a strong degree of correlation because they happen to fall on the same relative location of a spectral line.
+A typical covariance matrix is shown in Figure~\ref{fig:interp_matrix}.
+
+\begin{figure}[!t]
+\begin{center}
+  \includegraphics{figs/interp_matrix.pdf}
+  \figcaption{A typical interpolation error matrix for $T_\textrm{eff} = $ 6220K, $\log_{10} g=$4.2 dex, and $\Z = $ -0.4 dex.}
+  \label{fig:interp_matrix}
+\end{center}
+\end{figure}
 

### 2014-09-13 tex/paper.tex
Draft for Sean

@@
 \begin{abstract}
-We present a modular and extensible framework for forward-modelling spectra with synthetic models.
-Due to the nature of spectra, that adjacent pixels are not independent samples of the intrinsic flux distribution, it is common to result in residual spectra that exhibit some degree of correlation.
-Neglecting this covariance in the residual spectrum can result in a biased estimates of the parameters which describe the model spectrum.
+We present a modular and extensible framework for inference of model parameters using synthetic spectra.
+By the nature of continuously sampled spectra, when subtracting a synthetic model, adjacent pixels in the residual spectrum exhibit some degree of correlation.
 In the high signal-to noise, large spectral range limit common to stellar parameter estimation, the covariant structure of the residuals can bias the resulting parameter determinations.
 We present a likelihood function designed to account for this covariant structure, utilizing the machinery of Gaussian process covariance kernels to self-consistently identify and model the covariance.
-Specific to stellar parameter estimation, a common problem is the presence of spectral lines with an incorrect strength, often due to incorrect atomic constants or radiative treatment.
-We also present the development of a local covariance kernel that can track and self-consistently down-weight pathologically wrong spectra lines.
-Local kernels that track line weights provide a mechanism to learn corrections to spectral models, by fitting in a hierarchical manner.
-We demonstrate the application of this method, implemented as an open source code, to fit an echelle optical spectrum WASP-14, an F star, and a mid-infra spectrum of Gl51, an M5 star.
+Additionally, a common problem specific to stellar parameter estimation is the presence of spectral lines with mismatched strength, often due to incorrect atomic constants or radiative treatment.
+To address this issue, we also present the development of a novel local covariance kernel that can identify and self-consistently down-weight pathologically wrong spectral line ``outliers.'' By fitting multiple spectra in a hierarchical manner, these local kernels provide a mechanism to learn empirical corrections to spectral models.
+We demonstrate the application of this method, implemented as a freely available open source code, by fitting an high resolution optical spectrum WASP-14, an F star, and a moderate resolution mid-infrared spectrum of Gl51, an M5 star.
 \end{abstract}
@@
 \startdata
-$T_{\rm eff}$ (K) & 6517     & 13                      &  6273   & 15        \\
+$T_{\rm eff}$ (K) & 6523     & 14                      &  6234   & 17        \\
 $\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
-$\Z$          & -0.27    & 0.01                    & -0.498  & 0.006     \\
-$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.29     & 0.05                    &  4.89   & 0.05      \\
+$\Z$          & -0.28    & 0.01                    & -0.52  & 0.02     \\
+$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.28     & 0.06                    &  4.89   & 0.06      \\
 $v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
-$\log \Omega$\tablenotemark{b} & -12.7181 & 0.0004 & -19.678 & 0.001
+$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -19.676 & 0.001
 \enddata
@@
 \section{Interpolation Errors}
-The interpolation of a model spectrum from a synthetic library of spectra is a conceptually straightforward process; however, in practice there may be several subtle effects that require consideration, depending on the application.
+The interpolation of a model spectrum from a synthetic library of spectra is a conceptually straightforward process; however, depending on the application, in practice there may be several subtle effects that require consideration.
 Spectra are generally stored in a synthetic library as 1-D arrays of flux values, sampled at specific wavelengths at high resolution (typically R$\gtrsim$100,000).
 Each spectrum has several fundamental parameters which describe the model that generated it.
-To use a concrete example, the stellar libraries used in this paper have three parameters, $T_\textrm{eff}$, $\log_{10} g$, and $\Z$, each of which constitutes an axis in the synthetic library.
-Therefore, the full synthetic grid can be stored as a 4-D array, where the first three axes are the stellar parameters and the 4th axis describes the wavelength.
-The value stored in the array is the flux of the spectrum described by current value of the four axes.
-
-Spectral line formation is a non-linear process, with the shape of the line determined by complicated radiative synthesis, which is a function of the stellar parameters.
-Interpolating a \emph{physical} spectrum short-cuts this non-linear process and attempts to predict the behavior of the spectral line based upon the behavior of nearby spectra in the grid.
-While this approximation is generally accurate enough for our applications, for high signal-to-noise applications it becomes necessary to account for any inaccuracies introduced by the interpolation.
-In practice, there are two factors which influence the choice of interpolation method.
+For example, the stellar libraries used in this paper have three parameters, $T_\textrm{eff}$, $\log_{10} g$, and $\Z$, each of which constitutes an axis in the synthetic library.
+Therefore, the full synthetic grid is stored as a 4-D array, where the first three axes are the stellar parameters and the 4th axis describes the wavelength.
+The value stored in the array is the flux of the spectrum described by the current value of the four axes.
+
+Spectral line formation is a non-linear process: line shapes are determined by the complicated radiative transfer through the stellar atmosphere, the structure of which is iterated to convergence using a specific set of stellar parameters.
+Interpolating a spectrum short-cuts this non-linear process and attempts to predict the behavior of the spectral line based upon the behavior of nearby spectra in the grid.
+While this approximation is generally accurate enough for our applications, in the high signal-to-noise regime it becomes necessary to account for any inaccuracies introduced by the interpolation.
+In practice, there are two factors to consider when interpolating spectra.
 First, for a small percentage of pixels, the flux value will not vary monotonically with change in fundamental parameters.
-Second, visual inspection of the pixel values as a function of fundamental parameters makes it apparent that the spectra are not Nyquist sampled in parameter space.
-This means that synthetic grid does is not dense enough to losslessly reconstruct any synthetic spectrum.
-However, this fact is not surprising given the computational complexity of stellar synthesis.
-Given this reality, we make the most conservative assumption about the behavior of the pixels between grid points and decide to use a linear interpolation scheme.
+This informs the decision about which functional form of interpolation to use.
+Second, inspecting the flux values of a sample of pixels as a function of fundamental parameters makes it apparent that not all pixels of the spectrum are Nyquist sampled with regards to the stellar parameters.
+This unfortunately means that the synthetic grid is not sampled densely enough to losslessly reconstruct a synthetic spectrum.
+However, given the computational complexity of stellar synthesis, this fact is not surprising.
+In order to proceed, we make the most conservative assumption about the behavior of the pixels between grid points and decide to use a linear interpolation scheme.
 Using a more advanced interpolation method, such as a spline or other band-limited interpolation scheme, would not remedy the fact that the grid itself is not Nyquist sampled.
@@
 We empirically determine the interpolation error by performing a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, generate an interpolated spectrum from the nearest two grid points, and compare the difference between the two.
-For example, consider the {\sc Phoenix} grid used for the WASP-14 parameter estimation.
-In this parameter space, the grid is sampled every 100K in $T_\textrm{eff}$ and every 0.5 dex in $\log_{10} g$ and $\Z$.
+As a concrete example, the {\sc Phoenix} grid used for the WASP-14 parameter estimation is sampled every 100K in $T_\textrm{eff}$ and every 0.5 dex in $\log_{10} g$ and $\Z$.
 Therefore, for each spectrum in the synthetic grid, there are also three error spectra associated with it, one for each axis of stellar parameters.
@@
   \figcaption{(\emph{top}) A small range of the ``true'' {\sc Phoenix} spectrum at $T_\textrm{eff} = $ 6000K, $\log_{10} g=$4.0 dex, and $\Z = $ -0.5 dex, shown with the drop-out interpolation spectra along each of the three parameter axes.
-These interpolation error spectra are determined over an interpolation distance twice that (e.g., 100K) of the actual distance used in practice.
-Typically, the interpolated spectra closely match the ``true'' spectrum.
+These interpolation error spectra are determined over an interpolation distance twice that (e.g., 100K) of the actual distance used while fitting.
+The interpolated spectra closely match the ``true'' spectrum.
 (\emph{bottom}) The three ``error spectra'', or difference between the interpolated spectrum and the true spectrum, relative to the true spectrum value.
@@
 For the few points where this is not true (e.g., $\log_{10} g$ near 5161.5\AA) this implies a non-monotonic change in flux between grid points.
-Therefore, it is best to conservatively use the error spectrum, which protects against assumptions of monotonicity.}
+Therefore, it is most conservative to use the error spectrum in an interpolation error model, which accounts for this behavior and protects against assumptions of monotonicity.}
   \label{fig:interp_errors}
@@
 
-Generally, the interpolation error is small: typically no larger than 2\% of the flux and usually--but not always--much smaller than the total flux difference between spectra at adjacent grid points (see Figure~\ref{fig:interp_errors}).
-This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the large number of possible lines to fit.
+Generally, the interpolation error is small: typically no larger than 2\% of the flux (smaller than the Poisson noise for moderate to high signal-to-noise spectra) and usually--but not always--much smaller than the total flux difference between spectra at adjacent grid points (see Figure~\ref{fig:interp_errors}).
+This test gives us confidence that it is possible to generate a sufficiently accurate interpolated spectrum, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the large number of possible lines to fit.
 However, while the interpolation error is small, because all spectral lines have roughly the same shape, it is often true that the error adds \emph{coherently}.
-When the signal to noise of the spectrum is particularly large, this coherent trend can contribute substantially by diminishing the likelihood of an interpolated spectrum, since the lines in the spectrum are less ``physical'' in shape than that of a spectrum directly on a gridpoint, which requires minor or no interpolation.
+When the signal to noise of the spectrum is particularly high, this coherent trend can contribute substantially by diminishing the likelihood of an interpolated spectrum, since the lines in the spectrum are less ``physical'' in shape than that of a spectrum directly on a gridpoint, which requires minor or no interpolation.
 Therefore in order to deliver the proper likelihood for interpolated spectra it is necessary to quantify the degree of interpolation error and add this into the covariance matrix.
@@
   \includegraphics{figs/interp_grid.pdf}
-  \figcaption{Tri-linear interpolation on a grid cell the shape of a rectangular prism.}
+  \figcaption{Tri-linear interpolation on a grid cell the shape of a rectangular prism.
+I'm not sure if this figure is necessary, but I think it could help guide the reader.
+Perhaps by labelling the axis with temperature and so on.
+It is here as a prototype.}
   \label{fig:interp_grid}
@@
 For example, all pixels at the continuum will tend to act similarly, regardless of where on the spectrum the pixels fall.
-Any pixel that falls on the blue side of a spectral line will tend to exhibit a similar error behaviour regardless of which spectral line, as long as that line was changing in a similar manner.
+All pixels that fall on the blue side of a spectral line will tend to exhibit a similar error behaviour regardless of which spectral line they belong to, as long as the general behavior of the spectral lines was similar.
 The elements of the interpolation covariance matrix $\mathsf{C}_{ij}^\textrm{int}$ are determined empirically from the drop-out spectra, and are a function of the degree of the interpolation.
 Consider the manner in which a typical spectrum is tri-linearly interpolated from the synthetic grid (Figure~\ref{fig:interp_grid}).
-The interpolated spectrum is a weighted average of the neighboring 8 spectra, with weights $d_l$ based upon the distance from the grid points.
+The interpolated spectrum is a weighted average of the neighboring 8 spectra, with weights $d_l$ based upon the distance from the nearest grid points $P$.
 To generate a composite covariance matrix which reflects the interpolation error present for the interpolated spectra, we must also combine the error spectra.
-Because the error is covariant, we must first multiply and then add the error spectra, in order to preserve the covariance.
+Because the error is covariant, we must first determine the empirical covariance matrix for each error spectrum and then add the covariance matrices.
 A simple sum of error spectra would destroy any information about the covariance.
-For a specific axis $k$ (e.g., $T_\textrm{eff}$, $\log_{10} g$, or $\Z$), we represent the average covariance matrix as a composite of all of the error spectra of this species
+The empirical covariance matrix generated by a single error spectrum $l$, for a specific axis $k$ (e.g., $T_\textrm{eff}$, $\log_{10} g$, or $\Z$), is
 \begin{equation}
-  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  d_{kl} \lambda_{ikl} \lambda_{jkl}
+  \mathsf{C}_{ijkl}^\textrm{int} = \lambda_{ikl} \lambda_{jkl}
+\end{equation}
+We represent the average covariance matrix as a composite of all of the error covariance matrices of this species
+\begin{equation}
+  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  d_{l} \mathsf{C}_{ijl}^\textrm{int}
 \end{equation}
@@
   \includegraphics{figs/interp_matrix.pdf}
-  \figcaption{A typical interpolation error matrix for $T_\textrm{eff} = $ 6220K, $\log_{10} g=$4.2 dex, and $\Z = $ -0.4 dex.}
+  \figcaption{A typical interpolation error matrix for $T_\textrm{eff} = $ 6220K, $\log_{10} g=$4.2 dex, and $\Z = $ -0.4 dex. Some elements of the covariance matrix may be negative.}
   \label{fig:interp_matrix}

### 2014-09-19 tex/paper.tex
last call edits on full draft; ready to circulate after your changes (if any)

--- 2014-09-10 tex/paper.tex
@@
 %\documentclass[preprint]{aastex} 
-\documentclass[iop,floatfix]{emulateapj}
+\documentclass[iop,floatfix,twocolappendix]{emulateapj}
 
@@
 \newcommand{\HK}{$\textrm{H}_2$O-K2}
-
-%Chebyshev coefficients RULE!
-\newcommand{\cc}[2]{c_{#2}^{(#1)}} %echelle order first, then degree coefficient
-
-%flux
+\newcommand{\cc}[2]{c_{#2}^{(#1)}}
+
 \newcommand{\flam}{f_\lambda}
-
-%\vt stands for ``vector theta''
 \newcommand{\vt}{ {\bm \theta}}
 \newcommand{\vT}{ {\bm \Theta}}
-
-%\vp stands for ``vector phi''
 \newcommand{\vp}{ {\bm \phi}}
 \newcommand{\vP}{ {\bm \Phi}}
-
-%\phi and \Phi are used for nuisance parameters
 \newcommand{\cheb}{ \vp_{\mathsf{P}}}
-\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}} %for signifying a specific order
+\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}}
 \newcommand{\Cheb}{ \vP_{\textrm{Cheb}}}
-\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} %for specifying everything but a specific order
+\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} 
 \newcommand{\cov}{ \vp_{\mathsf{C}}}
-\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} %for signifying a specific order
+\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} 
 \newcommand{\Cov}{ \vP_{\textrm{cov}}}
-\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}} %for specifying everything but a specific order
-
-\newcommand{\allParameters}{\vT} %all parameters
-\newcommand{\nuisanceParameters}{\vP} %all nuisance parameters
-
-%kernel notation
+\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}}
+
+\newcommand{\allParameters}{\vT} 
+\newcommand{\nuisanceParameters}{\vP}
+
 \newcommand{\KK}{\mathcal{K}}
@@
 
-\shorttitle{Spectroscopic inference}
+\shorttitle{Robust Spectroscopic Inference with Imperfect Models}
 \shortauthors{Czekala et al.}
@@
 
-%\title{A Method for the Spectroscopic Inference of Stellar Parameters}
 \title{Robust Spectroscopic Inference with Imperfect Models}
-\author{Ian Czekala, Sean M.~Andrews, Guillermo Torres, \& David W.~Latham}
+
+\author{Ian Czekala, Sean M.~Andrews, et al.}
 \affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
@@
 \begin{abstract}
-We present a modular and extensible framework for forward-modelling spectra with synthetic models.
-Due to the nature of spectra, that adjacent pixels are not independent samples of the intrinsic flux distribution, it is common to result in residual spectra that exhibit some degree of correlation.
-Neglecting this covariance in the residual spectrum can result in a biased estimates of the parameters which describe the model spectrum.
-In the high signal-to noise, large spectral range limit common to stellar parameter estimation, the covariant structure of the residuals can bias the resulting parameter determinations.
-We present a likelihood function designed to account for this covariant structure, utilizing the machinery of Gaussian process covariance kernels to self-consistently identify and model the covariance.
-Specific to stellar parameter estimation, a common problem is the presence of spectral lines with an incorrect strength, often due to incorrect atomic constants or radiative treatment.
-We also present the development of a local covariance kernel that can track and self-consistently down-weight pathologically wrong spectra lines.
-Local kernels that track line weights provide a mechanism to learn corrections to spectral models, by fitting in a hierarchical manner.
-We demonstrate the application of this method, implemented as an open source code, to fit an echelle optical spectrum WASP-14, an F star, and a mid-infra spectrum of Gl51, an M5 star.
+We present a modular, extensible framework for the spectroscopic inference of 
+physical parameters based on synthetic model spectra.  The subtraction of an 
+imperfect model from a continuously sampled spectrum introduces covariance 
+between adjacent datapoints (pixels) into the residual spectrum.  In the limit 
+of high signal-to noise data with large spectral range that is common for 
+stellar parameter estimation, that covariant structure can bias the parameter 
+determinations.  We have designed a likelihood function formalism to account 
+for the structure of the covariance matrix, utilizing the machinery of Gaussian 
+process kernels.  We specifically address the common problem of mismatches in 
+model spectral line strengths (with respect to data) due to intrinsic model 
+imperfections (e.g., in the atomic or molecular data, or radiative transfer 
+treatment) by developing a novel local covariance kernel framework that 
+identifies and self-consistently downweights pathological spectral line 
+``outliers."  By fitting multiple spectra in a hierarchical manner, these local 
+kernels provide a mechanism to learn about and build data-driven corrections to 
+synthetic model spectral libraries.  The application of this method, 
+implemented as a freely available open source code, is demonstrated by fitting 
+the high resolution optical ($V$-band) spectrum of WASP-14, an F5 dwarf with a 
+transiting exoplanet, and the moderate resolution near-infrared ($K$-band) 
+spectrum of Gliese~51, an M5 dwarf. 
 \end{abstract}
@@
 
-All astronomers recognize that spectroscopy offers a wealth of information that can help
-characterize the properties of the observing target.  In the context of stellar astrophysics, 
-spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption 
-lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface 
-gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the 
-masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and 
-molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]), 
-facilitating study of the chemical hallmarks of different stellar populations.  With sufficient 
-spectral resolution, the velocity content of a spectrum can convey crucial information about 
-stellar rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through 
-the radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of 
-these stellar properties, there is acute interest in the exoplanet community.  There, all estimates 
-of the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host 
-mass or radius {\it ratio} is constrained with the radial velocity or transit techniques, 
-respectively).  Moreover, essential clues to the planet formation process are encapsulated in the 
-dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
+All astronomers recognize that spectroscopy offers a wealth of information that 
+can help characterize the properties of the observing target.  In the context 
+of stellar astrophysics, spectroscopy plays many fundamental roles.  The 
+relative strengths and widths of stellar absorption lines provide access to key 
+parameters like effective temperature ($T_{\rm eff}$) and surface gravity 
+($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to 
+estimate the masses and ages so crucial to understanding stellar evolution, as 
+well as individual elemental (and molecular) abundances or the collective 
+``metallicity" (typically parameterized as [Fe/H]), facilitating study of the 
+chemical hallmarks of different stellar populations.  With sufficient spectral 
+resolution, the velocity content of a spectrum can convey crucial information 
+about stellar rotation ($v \sin i$) and kinematics (e.g., association with a 
+cluster or companion through the radial velocity, $v_z$).  While many fields 
+benefit from the spectroscopic measurements of these stellar properties, there 
+is acute interest in the exoplanet community.  There, all estimates of the 
+planet properties are made {\it relative} to the host properties (e.g., the 
+planet-to-host mass or radius {\it ratio} is constrained with the radial 
+velocity or transit techniques, respectively).  Moreover, essential clues to 
+the planet formation process are encapsulated in the dependences of planet 
+frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
 \citep[e.g.,][]{fischer05,buchhave14}.
-% SA: we ought to check with exoplanet people on these references. Can do later...
 
@@
 shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
-to one key parameter in the models, and therefore can be used to quantify it 
-\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
-the advantage of being trivially fast, but each index relationship is only informative over a 
-limited region of parameter-space.  A second technique exploits the cross-correlation of an 
-observed spectrum with a suite of model templates to identify an optimized set of parameters, 
-usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
-SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
-used (particularly in the spectral dimension), thereby achieving higher precision even for data 
-with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
-inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
-employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
-a spectral synthesis back-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
-\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
-for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
-least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
-it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
-practice, this often requires an effort focused on a pre-selected subsample of the data.
+to one key parameter in the models, and therefore can be used to quantify it \citep[e.g., {\tt 
+MOOG};][]{sneden73,gray94,reid95,rojas-ayala10,rojas-ayala12}.  This ``indexing" approach has the 
+advantage of being trivially fast, but each condensed relationship is only informative over a 
+limited swath of parameter-space.  A second technique exploits the cross-correlation of a spectrum 
+with a suite of model templates to optimize a set of parameters, usually with some weighting 
+applied to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the 
+speed advantage is maintained (perhaps enhanced) and more data content is used (particularly in the 
+spectral dimension), thereby achieving higher precision even for data with comparatively low 
+sensitivity; the disadvantage is that the model quality and parameter inferences are assessed in an 
+empirical, rather than probabilistic, framework.  A third approach employs a direct, pixel-by-pixel 
+comparison between model and data, and is usually associated with a spectral synthesis back-end 
+(e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric flexibility 
+(e.g., one can fit for arbitrary abundances or structures) and a proper inference framework 
+\citep[usually a least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,
+schoenrich13}, but it is computationally expensive and can be biased by systematics 
+\citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a pre-selected 
+subsample of the data.
 
@@
 systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
-observations to guide substantially improvements in the models themselves.  An overview of the 
+observations to guide substantial improvements in the models themselves.  An overview of the 
 methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
@@
 
-We opt for the computationally expedient approach that employs a library of
-model spectra, $\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm
-eff}, \,\, \log{g}, \,\, [{\rm Fe/H}]]$.  However, it is worth noting that the
-techniques developed here are applicable to {\it any} ``back-end" that
-generates a model spectrum.  In our adopted approach, the model spectrum for an
-arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
-\begin{equation} \label{eqn:interp} \flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto
-\flam(\vt_{\ast}), \end{equation} where we assign the symbol $\leadsto$ as an
-interpolation operator.  The multi-dimensional interpolation in
-Eq.~\ref{eqn:interp} needs to be performed many times, so computational
-efficiency is critical.  In practice, a simple tri-linear interpolation is
-suitably fast, but introduces an undesirable level of inaccuracy
-(particularly in the [Fe/H] dimension).  The interpolation quality can be
-empirically estimated by performing the operation in Eq.~\ref{eqn:interp}
-across a $\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the
-interpolated spectrum with the corresponding library spectrum. After an
-extensive exploration of such calculations \citep[see also][]{husser12}, we
-concluded that the best combination of speed and accuracy can be achieved by
-pre-computing a {\it refined} spectral library using a cubic spline
-interpolation with a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex,
-0.1\,dex], and then performing tri-linear interpolation over that refined
-grid.  Overall, this interpolation technique is found to be accurate within
-a few percent per full-resolution model pixel and $\lesssim 1$\% once the
-spectrum is downsampled to the pixels of the detector.  Ideally, this
-pre-interpolation could be avoided if the spectral library was computed
-over a refined grid (with a substantial up-front computational investment);
-but for the time being, we can empirically propagate these interpolation
-uncertainties into the likelihood calculations (as will be described in
-Section~\ref{subsec:likelihood}).
+We opt for the computationally expedient approach that employs a library of model spectra, 
+$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
+[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
+{\it any} ``back-end" that generates a model spectrum.  In our adopted approach, the model spectrum 
+for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library,
+\begin{equation} \label{eqn:interp} 
+\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}), 
+\end{equation} 
+where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
+interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
+is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces a 
+non-negligible level of uncertainty.  These interpolation uncertainties can be empirically 
+estimated by performing the operation in Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ 
+grid point, and then comparing the interpolated spectrum with the corresponding library spectrum 
+\citep[see also][]{husser12}.  We have developed this approach to propagate the interpolation 
+uncertainties into the likelihood calculations, as described briefly in 
+Section~\ref{subsec:likelihood} and elaborated in the Appendix.
 
@@
 Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
-observed spectrum, and does not account for several additional observational and 
-instrumental effects that become important in comparisons with real data.  Therefore, a certain 
-amount of post-processing is required before assessing the model quality.  We treat that 
-post-processing in two stages: the first deals with an additional set of ``observational" 
-parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
-location of the target, while the second employs a suite of nuisance (hyper-)parameters ($\vp$) 
-designed to mitigate an imperfect data calibration.
+observed spectrum, and does not account for several additional observational and instrumental 
+effects that become important in comparisons with real data.  Therefore, a certain amount of 
+post-processing is required before assessing the model quality.  We treat that post-processing in 
+two stages: the first deals with an additional set of ``observational" parameters, $\vt_{\rm obs}$, 
+that incorporate dynamical effects, geometry, and the relative location of the target, while the 
+second employs a suite of nuisance (hyper-)parameters ($\vp$) designed to mitigate an imperfect 
+data calibration.
 
@@
 where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
-$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
-spectrum).  Figure \ref{fig:broadening} shows a (condensed) graphical representation 
-of these post-processing steps.
+$N_{\rm pix}$-element model vector $\vM$ ($N_{\rm pix}$ is the number of pixels in the spectrum).  
+Figure \ref{fig:broadening} shows a (condensed) graphical representation of these post-processing 
+steps.
 
@@
 
+% NEEDS UPDATED FIGURE
+\begin{figure}[b!]
+\includegraphics{figs/chebyshev.pdf}
+\vspace{-0.75cm}
+\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
+spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
+shapes.  Chebyshev polynomials ({\it middle, top}) are designed to account for that mismatch ({\it 
+middle, bottom}); rather, in practice, the model is scaled (multiplied) by those polynomials ({\it 
+bottom}).  \label{fig:chebyshev}}
+\end{figure}
+
 The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
@@
 they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
-the physical parameters.  This concern is usually treated externally to any modeling procedure, most often 
-by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
-``normalization" implicitly assumes that there is no relevant information content in the spectral 
-shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
-the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
-an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
-molecular bands for cool stars).
+the physical parameters.  This issue is usually treated externally to any modeling procedure, most 
+often by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But 
+that ``normalization" implicitly assumes that there is no relevant information content in the 
+spectral shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach 
+will corrupt the inferences of these parameters.  Moreover, in practice this approach is limited, 
+since defining an appropriate polynomial becomes difficult in cases where the spectral line density 
+is high (e.g., molecular bands for cool stars).
 
@@
 \end{eqnarray}
-and $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
+where $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
 coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
@@
 multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
-$c^{(0)}$ in other orders to be different.  It is worth noting that this formalism can, in 
-principle, be extended to develop models for completely uncalibrated spectra, rather than the 
-residual calibration mismatches described here (with a suitable relaxation of the priors on 
-$\cheb$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how these nuisance 
-parameters are applied.
-
-\begin{figure}[!b]
-\begin{center}
-  \includegraphics{figs/chebyshev.pdf}
-  \figcaption{
-    The spectrum at the overlap of two echelle orders (23 \& 24).
-({\it panel 1}) The flux-calibrated dataset shows a slight discrepancy of $\lesssim 3$\% between orders.
-({\it panel 2}) To account for this residual error in the flux calibration, we multiply the model spectrum by a Chebyshev polynomial, whose coefficients are parameters that are part of the model.
-In principle, we could divide the data by these polynomials to recover what the true flux-calibrated data should be ({\it panel 3}), but instead we multiply the model by the polynomials to recover the original discrepancy between orders ( {\it panel 4}).
-This has the advantage of preserving the original dataset as a fixed quantity and makes the model (Equation~\ref{eqn:chebyshev}) linear in the Chebyshev polynomial coefficients.
-\label{fig:chebyshev}}
-\end{center}
-\end{figure}
+$c^{(0)}$ in other orders to be different.  Figure \ref{fig:chebyshev} offers a practical 
+demonstration of how these nuisance parameters are applied to the model.
 
@@
 line-spread function with at least a few pixels.  Therefore, adjacent pixels never record 
-completely independent samples of the true spectrum.  In that case, a difference between an 
-observed and modeled spectral feature will create a correlated residual that spans multiple 
-pixels.  This can be demonstrated practically through the autocorrelation of $\vR$: a slight model 
-mismatch will produce correlated residuals over a characteristic scale similar to the instrumental
-broadening kernel width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of 
-these correlated residuals in real data; a significant autocorrelation signal is found on a $\sim$5 
-pixel scale, corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
+independent samples of the true spectrum.  In that case, a difference between an observed and 
+modeled spectral feature will create a correlated residual that spans multiple pixels.  This can be 
+demonstrated practically through the autocorrelation of $\vR$: a slight model mismatch will produce 
+correlated residuals over a characteristic scale similar to the instrumental broadening kernel 
+width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of these correlated 
+residuals in real data; a significant autocorrelation signal is found on a $\sim$5 pixel scale, 
+corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
 
@@
   \includegraphics{class0_autocorrelation.pdf}
-  \caption{ ({\it top}) A typical model fit using 
-  parameters drawn from the posterior distribution and the residuals that
-  result. ({\it middle}) The low-amplitude, mildly covariant residuals in the gray band enlarged
-  to show the mildly covariant structure produced by slight mismatch between
-  the data and model spectra. ({\it bottom}) The autocorrelation of the
-residual sequence shown at left. Notice that there is significant correlation
-for offsets of $\lesssim 8$ pixels.}
-\label{fig:class0}
+  \figcaption{({\it top}) A comparison of the data and a typical model with parameters drawn from 
+the posterior distribution, along with the corresponding residual spectrum.  ({\it middle}) A 
+zoomed view of the gray band in the top panels, highlighting the mildly covariant residual 
+structure that is produced by slight mismatches between the data and model spectra.  ({\it bottom}) 
+The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
+offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not described well by 
+white (Poisson) noise alone. \label{fig:class0}}
 \end{center}
@@
 \end{equation}
-to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
-scale (we set $r_0 = 4\ell$).  Figure \ref{fig:matrix} demonstrates that random draws from such a 
-kernel readily produce correlated structure similar to those seen in a typical residual spectrum.
-
-\begin{figure*}[!htb]
-\begin{center}
-\includegraphics{figs/matrix_compilation.pdf}
-\caption{ ({\it top}) A typical model fit and residual spectrum (zoomed).
-An illustrative region is shaded in grey, showing the scale and structure of the residuals as an example for the following panels.
-({ \it left column}) A small inset of the covariance matrix corresponding to the gray region.
-Note that the colorbar scale changes with each inset in order to best illustrate the structure of the matrix.
-({\it right column}) Each panel shows three random draws (gray) from a multivariate normal distribution using the covariance matrix in the left column.
-At the bottom of each panel are the one, two, and three sigma contours (orange) of 200 random draws overlaid with the actual residuals from the gray region.
-{\it first row}: White, Poisson-only noise ($\delta_{ij}\sigma_i$) poorly reproduces both the scale and structure of a typical residual spectrum.
-({\it second row}) The addition of a global kernel ($\mathcal{K}^{\rm g}$) yields random draws that closely approximate the structure and amplitude of the residuals, but miss the strong outlier line at 5202.2 \AA.
-({\it third row}) Finally, adding in a local kernel ($\mathcal{K}^l$) at the location of the outlier produces random draws that completely match the structure of the residuals.
-Interestingly, note that the covariance matrix can produce random draws that have negative, positive, and even flat residual lines.
-The point is that with a non-trivial covariance matrix, the likelihood function can now accommodate the residuals generated from a typical spectroscopic fit.
-Stated differently, random draws from the likelihood function now behave like the residuals, whereas white noise cannot.
-All parameters of the covariance kernels are determined self-consistently along with the stellar parameters.
-}
-\label{fig:matrix}
-\end{center}
-\end{figure*}
+as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
+scale (we set $r_0 = 4\ell$).
 
@@
 
-Aside from the global covariance structure, there can also be local regions of strong (highly 
+In addition to the global covariance structure, there can be local regions of strong (highly 
 correlated) residuals that need to be treated in the modeling framework.  These patches of large 
-$\vR$ are usually produced by imperfect spectral lines in the models (missing opacity sources, 
-uncertain oscillator strengths, etc.); some representative examples are highlighted in Figure 
-\ref{fig:badlines}.  To parameterize such regions in $\vC$, we introduce a sequence of 
+$\vR$ are usually produced by imperfect model spectral lines (due to missing opacity sources, 
+uncertain oscillator strengths, etc.); some representative examples are shown in 
+Figure~\ref{fig:badlines}.  To parameterize such regions in $\vC$, we introduce a sequence of 
 non-stationary kernels that explicitly depend on the actual wavelength values of a pair of pixels 
 (on $\lambda_i$ and $\lambda_j$), and not simply the distance between them ($r_{ij}$).
-
-\begin{figure}[!htb]
-\begin{center}
-  \includegraphics{figs/badlines0.pdf}
-  \includegraphics{figs/badlines1.pdf}
-  \includegraphics{figs/badlines2.pdf}
-  \includegraphics{figs/badlines3.pdf}
-  \caption{A collection of spectral lines which have imperfect model fits.
-    ( {\it first panel}) The majority of spectral lines ($\gtrsim 60$\%) will
-    have minor differences in strength between data and model spectrum, which
-    produce low-amplitude correlations in the residuals on the length scale of
-    the width of a typical spectral line.  ( {\it second panel}) Sometimes
-    ($\lesssim 5$\% of all lines), a missing opacity source in the model (in
-    this case a line-blended \ion{Ca}{2}) leaves a large, highly correlated
-    patch of negative residuals.  ( {\it third panel}) Sometimes ($\lesssim
-    5$\% of all lines), an extraneous line in the model leaves a large, highly
-    correlated patch of positive residuals. ({\it fourth panel}) If the line
-    strengths are substantially discrepant ($\lesssim 10$\% of all lines),
-    there will be many correlated residuals of moderate amplitude. For any specific line, there might
-    exist a $\vT$ that will fit the line, but there does not exist a $\vT$ that
-    will properly fit \emph{all} the lines.}
-\label{fig:badlines}
-\end{center}
-\end{figure}
 
@@
 again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
-matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the Gaussian 
-width (we use $r_0 = 4\sigma_k$).  Figure \ref{fig:matrix} shows how this non-stationary Gaussian 
-kernel generates a localized region of enhanced variance that successfully mimics the kind of 
-residuals produced by an inaccurate spectral line model.  In effect, these kernels down-weight the 
-influence of such strong residuals in the likelihood calculation, mitigating any potential bias 
-they might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
+matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the width 
+parameter (e.g., $r_0 = 4\sigma_k$).  In effect, these kernels systematically down-weight the 
+influence of strong residuals in the likelihood calculation, mitigating any potential bias they 
+might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
 flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
-the probabilistic framework (as opposed to manual or threshold-based clipping or masking).
-
-These local kernels can be further modified to account for more complex residual structures.  For 
-example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
-pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
-spectral scale.  This phenomenologically different local covariance behavior can still be treated 
-in this framework if we design an appropriate kernel.
+the probabilistic framework (as opposed to the common manual or threshold-based techniques of 
+clipping or masking).
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics{figs/badlines.pdf}
+\figcaption{A particularly illustrative spectral region with substantial localized structure in 
+the residuals due to ``outlier" spectral lines in the model library.  For any specific line, there 
+might exist a set of model parameters, $\vT$, that will improve its match with the data, but a 
+$\vT$ that will properly fit \emph{all} the outlier lines does not exist in a pre-computed library 
+with (necessarily) limited parametric flexibility.  Out of concern that such intrinsic mismatches 
+can bias the inference on $\vT$, the methodology advocated here introduces local kernels to inflate 
+the covariance around these outliers, self-consistently down-weighting their influence on the fit.  
+\label{fig:badlines}}
+\end{center}
+\end{figure}
+
+These local kernels can be modified to account for more complex residual structures.  For example, 
+late-type stars with imperfectly modeled molecular bandheads may produce a complicated pattern of 
+positive and negative residuals or a pronounced mismatch over a relatively large spectral scale.  
+This phenomenologically different local covariance behavior can still be treated in this framework 
+if an appropriate kernel morphology is adopted.
+
+\begin{figure*}[!htb]
+\begin{center}
+\includegraphics[scale=1.05]{figs/matrix_compilation.pdf}
+\figcaption{A decomposition of the modeling procedure, explicitly highlighting the roles of the 
+various contributions to the covariance matrix.  The top panels show a typical comparison between 
+the data and model spectrum, along with the associated residual spectrum.  The subsequent panels 
+focus on the illustrative region shaded in grey.  The left column of panels show the corresponding 
+region of the covariance matrix $\vC$, decomposed into its primary contributions: from top to 
+bottom, the trivial noise matrix, then combined with the global covariance kernel, and finally 
+including an appropriate local covariance kernel.  In the right column, we show the zoomed-in 
+residual spectrum ({\it black}) along with example random draws from the subsets of $\vC$ exhibited 
+to the left.  The shaded contours ({\it orange}) represent the 1, 2, and 3\,$\sigma$ dispersions of 
+an ensemble of 200 random draws from $\vC$.  Note that the trivial noise matrix ($\delta_{ij} 
+\sigma_i$) poorly reproduces both the scale and structure of the residual spectrum.  The addition 
+of a global kernel ($\mathcal{K}^{\rm g}$) more closely approximates the structure and amplitude of 
+the residuals, but misses the outlier line at 5202.2\,\AA.  Including a local kernel 
+($\mathcal{K}^l$) at that location results in a covariance structure that does an excellent job 
+reproducing all the key residual features.  \label{fig:matrix}}
+\end{center}
+\end{figure*}
 
@@
 We can now compute the covariance matrix employed in the likelihood calculation 
-(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
-and these global and local kernels, 
+(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix, 
+an interpolation error kernel ($\mathcal{K}^{\rm int}$; see the Appendix), and the global and local 
+kernels discussed above, 
 \begin{equation}
-\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \,\, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
+\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + \mathcal{K}_{ij}^{\rm int} + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
                   \Klocal_{ij}(\vp_{{\mathsf C}, l}), 
@@
 with (hyper-)parameters $\cov = [\vp_{{\mathsf C}, g}, \vp_{{\mathsf C}, l}]$.  The factor $b$ is 
-a (constant) parameter that scales up the pixel noise values to account for additional detector or 
-data reduction uncertainties (e.g., read noise, uncertainties in the spectral extraction procedure, 
-and interpolation errors); reasonable values are $b \approx 1.02$--1.10 for well-calibrated optical 
-spectra (see Section \ref{sec:examples} for examples).  If there are $N_{\rm loc}$ local covariance 
-patches (see Section \ref{subsec:MCMC} on how this is determined), then there are $4N_{\rm loc}+2$ 
-elements in the set of covariance (hyper-)parameters, $\cov$.
+a parameter that scales up the Poisson noise in each pixel by a constant factor to account for 
+additional detector or data reduction uncertainties (e.g., read noise, uncertainties in the 
+spectral extraction procedure, etc.); typically $b < 1.1$ for well-calibrated optical spectra.  If 
+there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
+determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance (hyper-)parameters, 
+$\cov$.
+
+Figure \ref{fig:matrix} provides a graphical illustration of how the kernels that comprise the 
+covariance matrix are able to reproduce the structure present in a typical residual spectrum (for 
+the sake of clarity, the interpolation error kernel is discussed separately, in the Appendix).
 
@@
 
-We assume uniform priors on the global covariance parameters.  For the local
-covariance kernels, we adopt flat priors for the amplitudes and means, \{$a_k$,
-$\mu_k$\}, and construct a specialized prior for the widths, \{$\sigma_k$\},
-that is flat below the instrumental width (when $\sigma_k < \sigma_v$) and has
-a rapid (logistic) taper to zero for larger widths.  This form is designed to
-prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
+We suggest assuming uniform priors on the global kernel parameters.  For the local kernels, we 
+typically adopt flat priors for the amplitudes and means \{$a_k$, $\mu_k$\}, but construct a 
+specialized prior for the widths \{$\sigma_k$\} that is flat below the instrumental broadening 
+width (when $\sigma_k < \sigma_v$) and has a rapid (logistic) taper to zero at larger values.  This 
+form is designed to prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
 since such behavior is better treated by the global kernel.
@@
 literature or scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the 
-Chebyshev coefficients ($\cheb^0$) so that the polynomials are constant ($c_0 = 1$,  $c_{>0} = 0$ 
-in all spectral orders) and assume only the trivial noise spectrum contributes to the covariance 
-matrix ($\cov^0 = 0$, so $\vC$ is diagonal).  \\
+Chebyshev coefficients ($\cheb^0$) so that the polynomials are constant ($c_o^{(0)} = 1$,  
+$c_o^{(>0)} = 0$, $\forall$ $o$) and assume only the trivial noise spectrum and interpolation error 
+kernel contribute to the covariance matrix ($\cov^0 = 0$).  \\
 
@@
 
-\item Sample in the global (stationary) covariance kernel parameters, $\vp_{{\mathsf C},g} = 
-[a_g, \ell]$, with $\vT^i$, $\cheb^i$, and the local covariance kernel parameters, 
-$\vp_{{\mathsf C}, l}^{i-1}$, fixed.  Update $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf 
-C}, g}^i$.
-
-\item If any local (non-stationary) covariance kernels have been instantiated (see below), sample in their hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, 
-$\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  Update the local kernel parameters 
-$\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.  
+\item Sample in the global (stationary) kernel parameters, $\vp_{{\mathsf C},g} = [a_g, \ell]$, 
+with $\vT^i$, $\cheb^i$, and the local kernel parameters, $\vp_{{\mathsf C}, l}^{i-1}$, fixed.  
+Update $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf C}, g}^i$.
+
+\item If any local (non-stationary) kernels have been instantiated (see below), sample in their 
+hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  
+Update the local kernel parameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.
+
 \end{enumerate}
 
-\noindent (4) Return to Step~(2) and iterate to convergence. \\
+\noindent (4) Return to Step~(2) and iterate.  After a burn-in period, periodically pause the Gibbs 
+sampler and compute the Gelman-Rubin convergence diagnostic, $\hat{R}$ \citep[][their 
+Eq.~11.4]{gelman13}.  Break the loop if $\hat{R} \, < \, 1.1$.  \\
 
@@
 
-Local kernels are instantiated by the following procedure:  an ``average''
-residual spectrum is created by combining ~500 spectra residuals stored from a
-burn-in period using only global covariance kernels (prior to storage the
-Markov chain is thinned to account for autocorrelation of the posterior
-samples). Then, using an iterative thresholding criterion similar in procedure
-to ``sigma clipping,''\footnote{Although the procedure of identifying outliers
-  is similar to sigma clipping, the inference on the stellar parameters is
-  emphatically \emph{not} the same as sigma clipping, since we now determine
-  the weights of these spectral outliers inside of a probabilistic framework,
-  rather than directly setting their weights to zero as in sigma clipping.} we
-  identify the largest outlier pixel not already covered by a local kernel,
-  instantiate a local kernel at its location, and repeat until no residual
-  pixels above a certain threshold remain. Alternative local kernel
-  instantiation schemes, such as re-evaluating the location of kernels with
-  each Gibbs sampler iteration, yield similar results to the thresholding
-  approach. However, we found that this thresholding approach most consistently
-  converged with the least computational overhead. We explored the behavior of
-  setting different threshold levels, and decided to use four times the
-  standard deviation of the average residual spectrum. Lower thresholds (e.g.,
-  3$\sigma$ and 3.5$\sigma$) result in the instantiation of more kernels, which
-  in turn slightly reduces the amplitude of the global kernel. Taken to the
-  logical conclusion, if the threshold was set low enough then there would be a
-  local kernel for each spectral line and no global kernel would be required.
-  We find that the final stellar posterior is relatively insensitive to the
-  choice of threshold level as long as it is set low enough to capture the most
-  egregious spectral outliers (e.g., 4$\sigma$).  Once all of the local kernels
-  have been instantiated, the Gibbs sampler is run for another period of
-  burn-in.\footnote{There is no practical reason to delete local kernels once
-    instantiated.  If the parameters have changed such that a given local
-    kernel is no longer required, that kernel amplitude will be driven towards
-    zero and represent a negligible contribution to $\vC$; in effect, the model
-    will act as if the kernel were deleted automatically.}
-
-After burn-in is complete, we periodically pause the Gibbs sampler and compute the Gelman-Rubin convergence 
-diagnostic  $\hat{R}$ \citep[][their Eq.~11.4]{gelman13}, which compares the intra-chain and 
-inter-chain variances to assess how the inference of the posterior would be improved if the chains 
-were to be run for an infinite number of iterations.  The Gibbs sampler is eventually stopped if 
-$\hat{R} \, < \, 1.1$, and the chains appear converged in a visual inspection.
-
-Gibbs sampling from the posterior with this procedure can be a computational challenge.  A typical 
-spectrum might have $N_{\rm pix}$$>$1000, and therefore the many repeated evaluations of the matrix 
-product $\vR^{\trans} \vC^{-1} \vR$ in the likelihood function calculation can be numerically 
-expensive; prohibitively so in the brute force sense, since $\vC$ is non-trivial by design.  
-However, there are two aspects of these calculations that enable us to take advantage of some 
-clever algorithms designed for these kinds of issues.  First, since we are only interested in the 
-product $\vR^{\trans} \vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can employ some efficient sparse 
-matrix approximations that avoid the computational expense of matrix inversion.  And second, 
-because $\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky factorization
-to optimize the evaluation of the matrix product.  In practice, this means Steps~(2), (3a), and 
-(3b) in the procedure outlined above can be conducted with minimal computation cost.  Although the
-Cholesky factorization must be re-done for each update of $\vp_{\mathsf C}$ (Step~3d), this does 
-not incur a high cost because the covariance kernels were designed to deliver a sparse matrix (via 
-the Hann windows).  We use the high-performance 
-\texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} 
-library to implement the sparse matrix and Cholesky factorization operations \citep{chen08,
-davis09}, and extend the Metropolis-Hastings sampler included in the {\tt emcee} package 
-\citep{foreman-mackey13} to function within a blocked Gibbs 
+In Step~(3c), local covariance kernels are instantiated into the model according to the following 
+procedure.  First, an ``average" residual spectrum is generated by combining $\sim$500 residual 
+spectra that were stored during a burn-in period using only the global kernels (prior to this 
+storage, the Markov chain is thinned to account for autocorrelation of the posterior samples).  
+This average spectrum is then iteratively examined for deviations outside a critical threshold; 
+when a large residual is identified, a local kernel is introduced with a mean ($\mu_k$) at its 
+location.\footnote{Although this may seem similar to the procedure of ``sigma-clipping", there is a 
+crucial difference.  Rather than rejecting outlier data once it is found (i.e., setting its weight 
+in the inference problem to zero), this procedure will actually self-consistently determine how to 
+weight the outliers inside the probabilistic framework.}  After some experimentation with different 
+threshold levels, we have adopted a criterion whereby a kernel is instantiated where the local 
+residual is at least 4$\times$ the standard deviation in the average residual 
+spectrum.\footnote{Lower thresholds result in more local kernels, thereby reducing the amplitude of 
+the global kernel.  In the extreme case of a very low threshold, a local kernel would be 
+instantiated for every spectral line (and no global kernel would be required).  We found that 
+ultimately the posterior inferences on the parameters of interest are relatively insensitive to the 
+choice of a threshold level, so long as it is set low enough to capture the egregious outliers.}  
+Alternative instantiation schemes, such as re-evaluating the locations of kernels with each 
+iteration of the Gibbs sampler, yield similar results; however, the adopted approach was found to 
+consistently converge with minimal computational overhead.  Once all of the necessary local kernels 
+have been instantiated, the Gibbs sampler is run for another period of burn-in.\footnote{There is 
+no practical reason to delete local kernels once instantiated.  If the parameters have changed such 
+that a given local kernel is no longer required, that kernel amplitude will be driven towards zero 
+and represent a negligible contribution to $\vC$; in effect, the model will act as if the kernel 
+were deleted automatically.}
+
+The procedure outlined above can be a significant computational challenge.  A typical data spectrum 
+has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the many repeated evaluations of the matrix 
+product $\vR^{\trans} \vC^{-1} \vR$ in the likelihood calculation can be numerically expensive; 
+prohibitively so in the brute force sense, since $\vC$ is non-trivial by design.  However, there 
+are two aspects of these calculations that enable us to take advantage of some clever algorithms 
+designed for such issues.  First, since we are only interested in the product $\vR^{\trans} 
+\vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can employ some efficient sparse matrix approximations 
+that avoid the computational expense of matrix inversion.  And second, because $\vC$ is a 
+symmetric, positive semi-definite matrix, we can employ Cholesky factorization to optimize the 
+evaluation of the matrix product.  In practice, this means Steps~(2), (3a), and (3b) in the 
+procedure outlined above can be conducted with minimal computation cost.  Although the Cholesky 
+factorization must be re-done for each update of $\vp_{\mathsf C}$ (Step~3d), this does not incur a 
+high cost because the covariance kernels were designed to deliver a sparse matrix (via the Hann 
+windows).  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization 
+operations \citep{chen08,davis09}, and extend the Metropolis-Hastings sampler included in the {\tt 
+emcee} package \citep{foreman-mackey13} to function within a blocked Gibbs 
 sampler.\footnote{\url{https://github.com/iancze/emcee}} \\
@@
 using the Tillinghast Reflector Echelle Spectrograph \citep[TRES;][]{furesz08} on the Fred Lawrence 
-Whipple Observatory 1.5\,m telescope.  TRES delivers a 51-order echelle spectrum covering the 
-full optical wavelength range (3860--9100\,\AA).  The data were reduced and calibrated using 
-standard techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for 
-more specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
+Whipple Observatory 1.5\,m telescope.  TRES delivers an echelle spectrum with 51 orders that cover 
+the optical range from 3860--9100\,\AA.  The data were reduced and calibrated using standard 
+techniques in the TRES pipeline (cf.,~\citealt{buchhave10}; see \citealt{torres12} for more 
+specific details).  At 5100\,\AA, the S/N is $\sim$150 per resolution element.
 
@@
 library; it indicates $T_{\rm eff} = 6507\pm50$\,K, $[{\rm Fe/H}] = -0.15\pm0.08$\,dex, and $v \sin 
-i = 3.9\pm0.5$\,km s$^{-1}$.  It is worth noting that these parameter inferences represent an 
-average of those determined from four distinct TRES datasets, such that the uncertainties reflect 
-the systematic spread between the best-fit parameter values. 
-%\comm{This is something we need to discuss more with Dave and Willie.}
+i = 3.9\pm0.5$\,km s$^{-1}$.  The uncertainties are dominated by a ``floor" contribution, imposed 
+independently on each parameter and meant to be representative of the systematics (a standard 
+procedure in this field).
 
 Given the decision to employ a spectral library to generate models (Sect.~\ref{subsec:synthetic}), 
-a direct comparison of the \citeauthor{torres12}~{\tt SPC} result with our framework is relevant.  
-Using the same TRES spectral orders (for a single WASP-14 observation, as noted above), the {\sc 
-CfA/Kurucz} library, and an appropriate $\delta$-function prior on $\log g$, we infer parameters 
-that are entirely consistent with the {\tt SPC} approach; see Table~\ref{table:Kurucz}.  
-Figure~\ref{fig:Kurucz_residuals} demonstrates the quality of our approach in reproducing the 
-WASP-14 spectrum and, especially significant, the structure of the residuals.  
-Figure~\ref{fig:Kurucz_posterior} shows some marginal projections of the posterior-space, 
-highlighting key parameter degeneracies.  
-%Figure~\ref{fig:metacomparison} compares the inferred 
-%parameters with the various determinations made by \citet{torres12}.
+a direct comparison of the \citeauthor{torres12}~{\tt SPC} result with our framework is most 
+relevant.  Using the same TRES spectral orders, the {\sc CfA/Kurucz} library, and a 
+$\delta$-function prior on $\log g$ (at 4.29), we infer parameters that are consistent with the 
+{\tt SPC} approach; see Table~\ref{table:Kurucz}.  Figure~\ref{fig:Kurucz_residuals} demonstrates 
+the quality of our approach in reproducing the WASP-14 spectrum and, especially significant, the 
+structure of the residuals.  Figure~\ref{fig:Kurucz_posterior} shows some marginal projections of 
+the posterior-space, highlighting key parameter degeneracies.
 
@@
 \includegraphics{figs/residuals_Kurucz_logg.pdf}
+\vspace{-0.3cm}
 \figcaption{({\it top}) A representative segment of the TRES spectrum of WASP-14 ({\it blue}), 
@@
 kernels are not required.  \label{fig:Kurucz_residuals}}
+\vspace{0.2cm}
 \end{figure*}
@@
 \startdata
-$T_{\rm eff}$ (K) & 6517     & 13                      &  6273   & 15        \\
+$T_{\rm eff}$ (K) & 6523     & 14                      &  6234   & 17        \\
 $\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
-$\Z$          & -0.27    & 0.01                    & -0.498  & 0.006     \\
-$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.29     & 0.05                    &  4.89   & 0.05      \\
+$\Z$          & -0.28    & 0.01                    & -0.52  & 0.02     \\
+$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.28     & 0.06                    &  4.89   & 0.06      \\
 $v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
-$\log \Omega$\tablenotemark{b} & -12.7181 & 0.0004 & -19.678 & 0.001
+$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -19.676 & 0.001 \\
+$A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
 \enddata
@@
   \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+\vspace{-0.5cm}
   \figcaption{The posterior probability distributions for the interesting stellar parameters 
@@
 
-%\begin{figure}[!htb]
-%\begin{center}
-%\includegraphics{figs/metacomparison.pdf}
-%\caption{The posterior peaks and widths for various determination of the stellar parameters of WASP-14.
-The determinations in \citet{torres12} are shown along with the estimates using the {\sc CfA/Kurucz} synthetic library and the {\sc PHOENIX} synthetic library.
-\todo{What else do we want to show on here?}}
-%\label{fig:metacomparison}
-%\end{center}
-%\end{figure}
-
-It is beneficial to review what factors into the parameter uncertainties.  In our approach, we 
-consider contributions from both trivial (Poisson) noise and a non-trivial Gaussian process 
-covariance matrix.\footnote{Note that the {\sc CfA/Kurucz} library is so well-tuned for stars like 
-WASP-14 over this limited spectral range that no local covariance kernels need to be instantiated; 
-in this specific case, we employ only global covariance kernels to reproduce the residual 
-structure.} To decompose their respective contributions, we performed the same analysis with only a 
-trivial covariance matrix ($b = 1$; $\cov = 0$), which simplifies the likelihood function to the 
-familiar $\chi^2$ form.  Doing this, we identify little bias to the best-fit parameters, but find 
-unrealistic precision (i.e., $T_{\rm eff}$ to $\pm$5\,K, $[{\rm Fe/H}]$ to $\pm0.004$\,dex, and $v 
-\sin i$ to $\pm0.01$\,km s$^{-1}$).  Such small uncertainties are to be expected, since the 
-prescribed likelihood function presumes that the data are generated from the model; in this 
-paradigm, the only way to introduce deviation is through white noise (see Fig.~\ref{fig:matrix}).  
-Of course, that implicit assumption is invalid, since just a few simple parameters ($\vT$) are 
-generally insufficient to account for the thousands of pixels and hundreds of spectral features in 
-a typical dataset.  In essence, the data and an imperfect model spectrum also have {\it systematic} 
-differences that also contribute to the uncertainties, which we forward-model using a non-trivial 
-covariance matrix.  In this specific case, these systematic and the Poisson uncertainties 
-contribute roughly equally to the $\vT$ posteriors.  For synthetic models that are less finely 
-tuned than the {\sc CfA/Kurucz} grid, the systematic uncertainties will dominate (see below).
-
-In addition to noise in the data (statistical uncertainties) and systematic uncertainties in the 
-models, we might also expect that systematics associated with the instrument and/or calibration 
-contribute to the parameter uncertainties.  Common practice is to estimate these latter systematics 
-by comparing parameter inferences made using different spectra (e.g., from different instruments or 
-observing nights): the results are usually presented as an average, with uncertainties inflated by 
-adding in a ``floor" term that accounts for the dispersion between the inferences from individual 
-spectra.  That procedure implicitly assumes that these systematics are Gaussian (add in quadrature) 
-and independent in each parameter.  But the approach we are advocating here avoids such a heuristic 
-by self-consistently parameterizing the fit quality.  Because the parameter uncertainties are 
-derived through a forward model, this method (crucially) preserves the morphologies of intrinsic 
-parameter degeneracies in a way that cannot be reproduced by artificially (i.e., manually) 
-inflating the posteriors.
-
-To quantify these instrumental systematics for WASP-14, we modeled the three other TRES spectra 
-obtained by \citet{torres12} independently (under the same assumptions).  We found a minimal 
-scatter of $\sim$20\,K in $T_{\rm eff}$ and $\sim$0.02\,dex in $\Z$ between the individual 
-inferences, a testament to quality calibration and an exceptionally stable spectrograph.  The more 
-appropriate way of combining the inferences is to model these four spectra simultaneously in our 
-hierarchical framework: in that case, we found that the parameter inferences become slightly more 
-precise (as one would expect for a weighted average and well-tuned models), while preserving the 
-intrinsic correlations between parameters (something expressly destroyed in a weighted average).
-
-%\comm{I am not sure how/if it will fit in, but I want to see the results of fitting many more 
-%orders with Phoenix.}
-
-While comparisons with the {\sc CfA/Kurucz} models confirm that our methodology can faithfully 
-reproduce the inferences made with standard spectral fitting routines, they do little to illustrate 
-a key feature -- the treatment of local patches of large covariance (see 
-Sect.~\ref{subsec:local_covariance}) in a more sophisticated likelihood function.  For that we turn 
-to the {\sc Phoenix} library, comprised of model spectra with a much broader spectral range that 
-have not been fine-tuned for solar type stars.  As a result, users should expect that there will 
-likely be a population of ``outlier" residual lines present when comparing any observed spectrum 
-with the {\sc Phoenix} models \citep[e.g.,][their Fig.~8]{husser13}.  The standard approach is to 
-outright ignore, or perhaps ``$\sigma$-clip", the outlier lines and proceed with fitting the 
-remainder of the spectrum.  However, it is difficult to quantify how such decisions ultimately 
-affect the parameter inferences \citep{hogg10}.  By parameterizing these residual features with 
-local covariance kernels, we are self-consistently preserving any discriminatory power against the 
-stellar parameters that might still be contained in their imperfect representations.
+It is worthwhile to comment on the inferred parameter uncertainties.  In this specific example, 
+some experimentation with restricted forms of the covariance matrix confirm that the uncertainties 
+of the interesting parameters have roughly equal contributions from trivial (Poisson) noise, grid 
+interpolation errors, and a non-trivial global covariance term.\footnote{Note that the {\sc 
+CfA/Kurucz} library has been specifically well-tuned for stars like WASP-14, by careful variation 
+of atomic data (e.g., line lists, oscillator strengths) with reference to observations of the Sun 
+and Vega, by J.~Morse and J.~B.~Laird.  Over this limited spectral range, this means that no local 
+covariance kernels need to be instantiated to reproduce the residual structure.}  For synthetic 
+models that are less well-tuned than the {\sc CfA/Kurucz} grid, the latter two terms will certainly 
+dominate.  In any case, these uncertainties are still considerably smaller than the ``floor" values 
+imposed by \citet{torres12}, but they are representative of our expectations for comparing data of 
+high sensitivity and volume with models of limited parametric flexibility (i.e., the uncertainties 
+would inflate considerably if the many fixed or restricted parameters that went into building the 
+model libraries were available for variation).
 
@@
 \includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\vspace{-0.3cm}
 \figcaption{Same as Fig.~\ref{fig:Kurucz_residuals} for the {\sc Phoenix} models.  Note the
@@
 \label{fig:PHOENIX_residuals}}
+\vspace{0.2cm}
 \end{figure*}
@@
   \includegraphics[width=0.5\textwidth]{figs/PHOENIX_triangle.pdf}
+\vspace{-0.5cm}
 \figcaption{Same as Fig.~\ref{fig:Kurucz_posterior}, but for the {\sc Phoenix} models.
@@
 
-%This behavior of the {\sc PHOENIX} library motivates the use of global and local covariance kernels in a sophisticated likelihood function.
-The amplitude and scale of the global covariance grows to capture the discrepancy between data spectrum and synthetic spectrum, and acts as a mechanism to account for the systematic mismatch of data spectrum and model spectrum.
-Including this covariance broadens the posteriors on the stellar parameters, while the local covariance kernels provide cushion against outlier lines.
+There are two additional, {\it systematic} sources of uncertainty that merit attention.  First are 
+those associated with the instrument and/or calibration procedure.  Common practice is to quantify 
+these contributions by comparing parameter inferences made using different spectra (e.g., from 
+different instruments or observing nights): the results are usually presented as an average, with 
+uncertainties inflated by adding in quadrature parameter-independent terms that account for the 
+dispersion between the inferences from individual spectra.  To quantify these instrumental 
+systematics for WASP-14, we modeled three other TRES spectra obtained by \citet{torres12} 
+independently (under the same assumptions).  We found a scatter of $\sim$20\,K in $T_{\rm eff}$ and 
+$\sim$0.02\,dex in $\Z$ between the individual inferences, a testament to quality calibration and 
+an exceptionally stable spectrograph.  The more appropriate way of combining the inferences is to 
+model these four spectra simultaneously in our hierarchical framework: we found that the parameter 
+inferences become slightly more precise (as would be expected for a weighted average and well-tuned 
+models), while preserving the intrinsic parameter degeneracies (which is not possible in the 
+weighted average approach).
+
+A second source of systematic uncertainty lies with the models themselves, including the treatment 
+of structure, assumptions about boundary conditions (convection, etc.), and the employment of 
+fundamental atomic and molecular data.  As a benchmark for estimating the scope of this kind of 
+uncertainty, we repeated the analysis described above using instead the {\sc Phoenix} model 
+library.  Because the {\sc Phoenix} model spectra cover a much broader spectral range that has not 
+been fine-tuned for solar type stars (unlike the {\sc CfA/Kurucz} library), this has the added 
+benefit of demonstrating a key feature of the modeling framework -- the treatment of spectral line 
+``outliers" with local patches of high covariance (Sect.~\ref{subsec:local_covariance}) in a 
+sophisticated likelihood function.
 
@@
 spectra presented there and in the corresponding Figure~\ref{fig:Kurucz_residuals} for the {\sc 
-CfA/Kurucz} models.  First, the amplitude of the global covariance structure is $\sim$2$\times$ 
-larger for the {\sc Phoenix} models: since that amplitude is an indirect parameterization of the 
-fit quality, this indicates that the {\sc Phoenix} models present an overall worse match to the 
+CfA/Kurucz} models.  First, the amplitude of the global covariance structure is twice as large for 
+the {\sc Phoenix} models: since that amplitude is an indirect parameterization of the {\it fit 
+quality}, this indicates that the {\sc Phoenix} models present an overall worse match to the 
 data.  Second, there are several local regions of strong residuals (outlier lines), which are 
@@
 posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
-CfA/Kurucz} model libraries: see Table~\ref{table:Kurucz} and compare 
-Figures~\ref{fig:Kurucz_posterior} and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find 
-a systematically cooler temperature (by $\sim$240\,K) and lower metallicity (by a factor of 
-$\sim$2) than inferred from the {\sc CfA/Kurucz} models.  Because our framework explicitly 
-mitigates the potential for posterior bias from imperfect matches between data and models, we can 
-safely conclude that these apparent discrepancies are intrinsic to the different input physics of 
-the model libraries.\footnote{It is worth noting that these differences are not a manifestation of 
-the stringent prior on surface gravity.  When the spectra are modeled with $\log g$ unconstrained, 
-the posteriors for both libraries shift, but their respective discrepancies actually increase 
-slightly.}  Such a comparison serves as a stark reminder that incomplete and/or over-simplified 
-physical prescriptions remain the most significant source of systematic uncertainty in estimates 
-of fundamental stellar parameters.
-
-\subsection{Gl~51}
-
-A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
-Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
-(IRTF).  SPEX is a cross-dispersed echelle spectrograph that covers the red-optical to 
-thermal-infrared spectrum (0.7--5.5\,$\mu$m) in two settings.  These data were obtained as part of 
-the IRTF spectral standard library project \citep{cushing05,rayner09}, and were processed through 
-the well-vetted {\tt Spextool} reduction pipeline \citep{cushing04,vacca03} to deliver a fully 
-calibrated spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.
-
-Modeling late-type stellar atmosphere structures and their spectra is considerably more complex 
-than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
-opacities.  Especially confounding is the presence of complex condensates (clouds) at the coolest 
-temperatures \citep{allard13}, making it considerably more challenging to determine (sub-)stellar 
-properties \citep{rajpurohit14}.  Various approaches have been taken to infer stellar parameters in 
-the face of these difficulties, including iteratively masking regions with poor spectral agreement 
-\citep[e.g.,][]{mann13}.  Astutely, \citeauthor{mann13}~note that such a scheme may exclude 
-regions of the spectrum that contain intrinsically useful information for discriminating between 
-physical properties, and that a more sophisticated approach would weight each spectral region based 
-on its consistency with the data.  The modeling framework that we have constructed here does 
-exactly that.
+CfA/Kurucz} libraries: see Table~\ref{table:Kurucz} and compare Figures~\ref{fig:Kurucz_posterior} 
+and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find a systematically cooler temperature 
+(by $\sim$240\,K) and lower metallicity (by a factor of $\sim$2) than inferred from the {\sc 
+CfA/Kurucz} models.  Because our framework explicitly mitigates the potential for posterior bias 
+from imperfect matches between data and models, we can safely conclude that these apparent 
+discrepancies are intrinsic to the different input physics of the model libraries.\footnote{It is 
+worth noting that these differences are not a manifestation of the stringent prior on surface 
+gravity.  When the spectra are modeled with $\log g$ unconstrained, the posteriors for both 
+libraries shift, but their respective discrepancies actually increase slightly.}  Such a comparison 
+serves as a stark reminder that incomplete and/or over-simplified physical prescriptions remain the 
+most significant source of systematic uncertainty in estimates of fundamental stellar parameters.
 
@@
   \includegraphics{figs/residuals_Gl51_logg.pdf}
+\vspace{-0.3cm}
   \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a {\sc Phoenix} model
@@
 a large number of random draws from the covariance matrix (the shading is representative of the 1,
-2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:}.  Note how the 
-``outlier" features are clearly identified and treated by the local covariance kernels.
+2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:Kurucz_residuals}.  
+Note how the ``outlier" features are clearly identified and treated by the local covariance kernels.
 \label{fig:Gl51_residuals}}
+\vspace{0.2cm}
 \end{figure*}
@@
 
+\subsection{Gl~51}
+
+A moderate resolution ($R\approx2,000$) near-infrared spectrum of Gl~51 was obtained on 2000 
+Nov 6 using the SPEX instrument \citep{rayner03} on the 2.3\,m NASA Infrared Telescope Facility 
+(IRTF).  SPEX is a cross-dispersed echelle spectrograph that covers the red-optical to 
+thermal-infrared spectrum (0.7--5.5\,$\mu$m) in two settings.  These data were obtained as part of 
+the IRTF spectral standard library project \citep{cushing05,rayner09}, and were processed through 
+the well-vetted {\tt Spextool} reduction pipeline \citep{cushing04,vacca03} to deliver a fully 
+calibrated spectrum.  At 2.1\,$\mu$m, the S/N is $\sim$400 per resolution element.
+
+Modeling late-type stellar atmosphere structures and their spectra is considerably more complex 
+than for Sun-like stars, due to lingering uncertainties in the atmosphere physics and molecular 
+opacities.  Especially confounding is the presence of complex condensates (clouds) at the coolest 
+temperatures \citep{allard13}, making it considerably more challenging to determine (sub-)stellar 
+properties \citep{rajpurohit14}.  Various approaches have been taken to infer the key parameters in 
+the face of these difficulties, including iteratively masking regions with poor spectral agreement 
+\citep[e.g.,][]{mann13}.  Astutely, \citeauthor{mann13}~note that such a scheme may exclude 
+regions of the spectrum that contain intrinsically useful information for discriminating between 
+physical properties, and that a more sophisticated approach would weight each spectral region based 
+on its consistency with the data.  The modeling framework that we have constructed here does 
+exactly that.
+
 As a point of reference, \citet{rojas-ayala12} used a $K$-band spectrum at similar resolution to 
-the SPEX data to estimate the Gl~51 stellar parameters using an index technique.  They measured 
+the SPEX data to estimate the basic Gl~51 parameters using an index technique.  They measured 
 equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, 
@@
 speculated that this discrepancy may be a consequence of inaccurate atomic data (oscillator 
-strengths and/or opacities).  \hili{The indexing technique adopted by \citeauthor{rojas-ayala12} 
-presumes that these discrepancies do not affect the line {\it ratios}}.  Our approach offers an 
-alternative that makes no such assumption, instead using local covariance kernels to 
-self-consistently determine the ``weights" of discrepant spectral lines in the overall fit.  In 
-that sense, these outlier features still bring their full information content to bear on the 
-posterior distributions of the stellar parameters without imposing a systematic bias.  
-Figure~\ref{fig:Gl51_residuals} demonstrates that these features have been identified as
-problematic, resulting in localized patches of appropriately inflated covariance.
+strengths and/or opacities).  The indexing technique adopted by \citeauthor{rojas-ayala12} presumes 
+that these discrepancies do not affect the line {\it ratios}.  Our approach offers an alternative 
+that makes no such assumption, instead using local covariance kernels to self-consistently 
+determine the ``weights" of discrepant spectral lines in the overall fit.  In that sense, these 
+outlier features still bring their full information content to bear on the posterior distributions 
+of the stellar parameters without imposing a systematic bias.  Figure~\ref{fig:Gl51_residuals} 
+demonstrates that these features have been identified as problematic, resulting in localized 
+patches of appropriately inflated covariance.
 
@@
 $v_z$ ($\kms$) & 12.1 & 1.3 \\
-$\log \Omega$ (sr) & -19.628  & 0.003
+$\log \Omega$ (sr) & -19.628  & 0.003 \\
+$A_V$ & 0 & 0 (fixed)
 \enddata
@@
 exponential and a \matern\ kernel) or empirically-motivated parametric shapes (e.g., a saw-tooth 
-pattern) to provide a better representation than a simple Gaussian feature.
+pattern) to provide a better representation than a simple Gaussian feature.  \\
 
@@
 a linear combination of {\it global} (stationary) and {\it local} (non-stationary) Gaussian 
-Process kernels, which parameterize an overall mild covariance structure as well as small patches 
+process kernels, which parameterize an overall mild covariance structure as well as small patches 
 of highly discrepant outlier features.  The approach we describe is generally applicable to any 
@@
 complexity (if desired), and could be deployed as a substitute for a simplistic $\chi^2$ metric in 
-already-established fitting tools (e.g., {\tt SME}, {\tt MOOG}).  To demonstrate how it is used, we 
-determined the surface parameters of main-sequence stars with mid-F and mid-M spectral types from 
-high-S/N optical and near-infrared data, with reference to pre-computed model libraries 
-(Sect.~\ref{sec:examples}).
+already-established fitting tools (e.g., {\tt SME}).  To demonstrate how it is used, we determined 
+the surface parameters of main-sequence stars with mid-F and mid-M spectral types from high-S/N 
+optical and near-infrared data, with reference to pre-computed model libraries 
+(Sect.~\ref{sec:examples}).  The source code developed here is open and freely available for use: 
+see \url{http://github.com/iancze/StellarSpectra}.
 
@@
 like those being collected in programs like the APOGEE and HERMES surveys), since they provide 
-enormous leverage for identifying and improving the underlying model systematics.
-
-%\comm{Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
+enormous leverage for identifying and improving the underlying model systematics. \\
+%Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
 %include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
-%linked to that, since there is a popular Herschel HERMES program that will confuse people...}
-
-%\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
-%Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
-%acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
-%for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-%
+%linked to that, since there is a popular Herschel HERMES program that will confuse people...
+
+\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
+Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
+acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
+for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}. \\
+
+\appendix
+
+\section{Interpolation Errors} \label{sec:Appendix}
+
+Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
+application discussed here there are some subtleties that require detailed consideration.  
+Spectral line formation is a non-linear process, where line shapes are determined by the complex 
+radiative transfer through a stellar atmosphere.  We aim to bypass this non-linear 
+process and predict (through interpolation) the behavior of spectral lines based on pre-computed 
+library spectra with similar physical parameters.  The goal in this Appendix is to describe a 
+procedure by which we propagate an empirical approximation of the {\it inaccuracy} in that 
+prediction through the probabilistic framework we have developed.
+
+There are two important factors to consider when interpolating in libraries of model stellar 
+spectra.  First, the flux values in some non-negligible fraction of the spectra do not vary 
+monotonically with changes in the physical parameters that define the library: this helps inform 
+the decision about which functional form of interpolation should be used.  And second, the spectra 
+are usually not Nyquist sampled with respect to the physical parameters that define the library.  
+This unfortunately (but not surprisingly) means that the library is not sampled densely enough in 
+parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
+information loss.  Given these factors, we make the conservative (and computationally tractable) 
+assumption and adopt a linear interpolation scheme.  More advanced interpolation methods, like a 
+spline or other band-limited approach, do not change the fact that the library grid is not Nyquist 
+sampled, and so do not improve performance.  To account for the loss of spectral information during 
+the interpolation process, we have to build a model of the interpolation errors and incorporate it 
+into the likelihood function (via the covariance matrix).
+
+Model library spectra, $f_{\lambda}(\{\vt_{\ast}\})$, are stored as (1-dimensional) arrays of flux 
+values, sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).  In the case 
+of interest here, the collections of model parameters $\{\vt_{\ast}\} = [\{T_{\rm eff}, \log g,  \Z 
+\}]$ define the dimensions of the library grid; the full library is therefore encapsulated in a 
+4-dimensional array.  Both the libraries used here have grid spacings of 0.5 dex in $\log g$ and 
+0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but the {\sc 
+Phoenix} library has finer coverage in 100\,K increments.  We empirically determine a model grid of 
+interpolation errors through a series of ``drop-out" interpolation tests.  The idea is to remove a 
+spectrum from the library, $f_{\lambda}(\vt_{\ast})$, generate an interpolated spectrum, 
+$\tilde{f}_{\lambda}(\vt_{\ast})$, at the parameters corresponding to that removed spectrum from 
+the neighboring spectra, and then compare the two.  The result is a set of three ``error" (or, 
+rather ``difference") spectra, each representing an individual dimension in parameter-space, where
+\begin{eqnarray} \label{eqn:interr}
+e_{\lambda}(\vt_{\ast}^{k}) &=& f_{\lambda}(\vt_{\ast}^k) - \tilde{f}_{\lambda}(\vt_{\ast}^k) \\
+                            &=& f_{\lambda}(\vt_{\ast}^k) - \frac{1}{2}\left[f_{\lambda}(\vt_{\ast}^{k+1})+f_{\lambda}(\vt_{\ast}^{k-1}))\right], \nonumber
+\end{eqnarray}
+for $k = 1, 2, 3$ indexing the library parameters ($\vt_{\ast}$).
+
+As a concrete example, consider the stellar parameters $\vt_{\ast}$ = [6000\,K, 4.0 dex, -0.5 
+dex].  An ``error spectrum" in the $T_{\rm eff}$ dimension, $e_{\lambda}(\vt_{\ast}^1)$, would be 
+calculated from the difference between the grid spectrum at that $\vt_{\ast}$ and the average of 
+$f_{\lambda}$([5900\,K, 4.0 dex, -0.5 dex]) and $f_{\lambda}$([6100\,K, 4.0 dex, -0.5 dex]), as 
+defined in Eq.~\ref{eqn:interr}.  Likewise for the error spectra in the $\log g$ and $\Z$ 
+dimensions, $e_{\lambda}(\vt_{\ast}^2)$ and $e_{\lambda}(\vt_{\ast}^3)$.  A graphical demonstration 
+is shown in Figure~\ref{fig:interp_errors}.  These kinds of drop-out tests over larger spans of the 
+grid show that the amplitude of $e_{\lambda}$ grows approximately linearly with the interpolation 
+distance.
+
+\begin{figure*}[!t]
+\begin{center}
+  \includegraphics{figs/interp_errors.pdf}
+\vspace{-0.1cm}
+  \figcaption{(\emph{top}) A small subset of a {\sc Phoenix} model spectrum at library gridpoint 
+defined by $\vt_{\ast} = [T_\textrm{eff} = 6000$\,K, $\log g=4.0$ dex, and $\Z = $-0.5 dex$]$, 
+shown with the drop-out interpolation spectra ($\tilde{f}_{\lambda}$) along each of the three 
+parameter axes.  The interpolated spectra closely match the ``true'' spectrum.  (\emph{bottom}) The 
+three corresponding ``error" spectra plotted {\it relative to the true spectrum} (i.e., 
+$e_{\lambda}/f_{\lambda}$).  The shaded regions show the total change in $f_{\lambda}$ between the 
+two neighboring spectra in the library used for the interpolation (again relative to the true 
+spectrum).  Note that the error spectrum is generally much smaller than the typical change in flux 
+across gridpoints.  For the few points where this is not true (e.g., $\log g$ near 5161.5\AA), this 
+indicates non-monotonic changes between gridpoints.  Therefore, it is most conservative to build an 
+interpolation error model that accounts for this behavior and protects against assumptions of 
+monotonicity. \label{fig:interp_errors}}
+\end{center}
+\end{figure*}
+
+The interpolation error is generally small, $\lesssim$2\%, much less than the Poisson noise for 
+spectra with moderate sensitivity and usually (although not always) less than the difference 
+between spectra at adjacent grid points.  This gives us confidence that it is possible to generate 
+accurate interpolated spectra, and that any inference on the stellar parameters can be made at a 
+level substantially more precise than dictated by the coarse spacing of the library itself 
+(especially for such a large data volume).  However, although $e_{\lambda}$ is small in magnitude, 
+the fact that spectral lines have similar shapes means that the interpolation error often compounds 
+\emph{coherently}.  That coherence can substantially diminish the likelihood of an interpolated 
+spectrum, since its spectral line shapes are less ``physical" than for a spectrum located directly 
+on a computed gridpoint.  In essence, this highlights the underlying {\it necessity} of propagating 
+an interpolation error to deliver a proper likelihood at interpolated parameter values.
+
+Because the magnitude of $e_{\lambda}$ is strongly dependent on its association with the continuum 
+or a spectral line, the interpolation error model must be based on a (spectral) covariance kernel.  
+The elements of that kernel, $\mathcal{K}_{ij}^{\rm int}$, are determined from the empirical model 
+grid of $e_{\lambda}$ calculated from the drop-out technique.  The interpolated model spectrum and 
+is a weighted average of the neighboring 8 gridpoints (standard tri-linear interpolation).  To 
+construct $\mathcal{K}_{ij}^{\rm int}$ from the associated 24 interpolation error spectra, we first 
+need to generate sub-kernels that represent each dimension in parameter-space, 
+\begin{equation} \label{eq:subsubkernel}
+^k_l\mathcal{K}_{ij}^{\rm int} = \langle e_{\lambda_i}^k \; e_{\lambda_j}^k \rangle_l,
+\end{equation}
+where again $k$ indexes each element of $\vt_{\ast}$, and $l$ indexes each of the neighboring 
+gridpoints (and $i, j$ denote a pixel pairing, as in Sect.~\ref{sec:method}).  The sub-kernel for 
+dimension $k$ corresponding to the desired interpolate location is a simple weighted sum of the 
+matrices in Eq.~\ref{eq:subsubkernel},
+\begin{equation} \label{eq:subkernel}
+^k\mathcal{K}_{ij}^{\rm int} = \sum_{l=1}^8 \Delta_l \,\, ^k_l\mathcal{K}_{ij}^{\rm int},
+\end{equation}
+where the weights $\Delta_l$ depend on the distance from each grid point to the interpolate 
+location.
+
+\begin{figure}[!b]
+\begin{center}
+  \includegraphics{figs/interp_matrix.pdf}
+  \figcaption{A sub-region of a typical interpolation error kernel, $\mathcal{K}_{\ij}^{\rm int}$, 
+for an arbitrary set of parameters; $\vt_{\ast} = T_\textrm{eff} = 6220$, K, $\log g=4.2$ dex, and 
+$\Z = $-0.4 dex$]$.  The color scale is the same as in Fig.~\ref{fig:matrix}; since elements of 
+this covariance kernel can be negative, an analogous scale (in {\it red}) is duplicated to 
+highlight such regions with the same logarithmic stretch.  \label{fig:interp_matrix}}
+\end{center}
+\end{figure}
+
+Finally, we average the 3 sub-kernels in Eq.~\ref{eq:subkernel}, representing the covariance 
+introduced in each parameter dimension at the interpolate location, as 
+\begin{equation}
+\mathcal{K}_{ij}^{\rm int} = w_{ij}(r_0) \sum_{k=1}^3 \varrho_k \, ^k\mathcal{K}_{ij}^{\rm int}.
+\end{equation}
+Here, the weights $\varrho_k$ are based on the interpolation distance in parameter-space.  For 
+example, if the target interpolate distance were $T_{\rm eff} = 6120$\,K then $\varrho_1 = 0.4$ 
+(as a reminder, the library gridpoints are spaced by 100\,K in this dimension); for $T_{\rm eff} = 
+6100$\,K, $\varrho_1 = 0.0$; and for $T_{\rm eff} = 6150$\,K, the maximal interpolation distance, 
+$\varrho_1 = 1.0$.  
+For computational expediency and numerical stability, we smoothly taper the covariance kernel to 
+zero at the same length scale ($r_0 = 4 \ell$) as the global covariance kernel, again employing a 
+Hann window.  In reality, $\mathcal{K}^{\rm int}$ is a dense matrix: pixels separated by a wide 
+spectral range may in fact exhibit a strong degree of correlation because they happen to fall on 
+the same relative location of a spectral line.  A representative interpolation error kernel is 
+shown in Figure~\ref{fig:interp_matrix}. \\
 
@@
 
-\appendix
-
-\section{Interpolation Errors}
-The interpolation of a model spectrum from a synthetic library of spectra is a conceptually straightforward process; however, in practice there may be several subtle effects that require consideration, depending on the application.
-Spectra are generally stored in a synthetic library as 1-D arrays of flux values, sampled at specific wavelengths at high resolution (typically R$\gtrsim$100,000).
-Each spectrum has several fundamental parameters which describe the model that generated it.
-To use a concrete example, the stellar libraries used in this paper have three parameters, $T_\textrm{eff}$, $\log_{10} g$, and $\Z$, each of which constitutes an axis in the synthetic library.
-Therefore, the full synthetic grid can be stored as a 4-D array, where the first three axes are the stellar parameters and the 4th axis describes the wavelength.
-The value stored in the array is the flux of the spectrum described by current value of the four axes.
-
-Spectral line formation is a non-linear process, with the shape of the line determined by complicated radiative synthesis, which is a function of the stellar parameters.
-Interpolating a \emph{physical} spectrum short-cuts this non-linear process and attempts to predict the behavior of the spectral line based upon the behavior of nearby spectra in the grid.
-While this approximation is generally accurate enough for our applications, for high signal-to-noise applications it becomes necessary to account for any inaccuracies introduced by the interpolation.
-In practice, there are two factors which influence the choice of interpolation method.
-First, for a small percentage of pixels, the flux value will not vary monotonically with change in fundamental parameters.
-Second, visual inspection of the pixel values as a function of fundamental parameters makes it apparent that the spectra are not Nyquist sampled in parameter space.
-This means that synthetic grid does is not dense enough to losslessly reconstruct any synthetic spectrum.
-However, this fact is not surprising given the computational complexity of stellar synthesis.
-Given this reality, we make the most conservative assumption about the behavior of the pixels between grid points and decide to use a linear interpolation scheme.
-Using a more advanced interpolation method, such as a spline or other band-limited interpolation scheme, would not remedy the fact that the grid itself is not Nyquist sampled.
-To account for the loss of spectral information during the interpolation process, we build a model of interpolation error and incorporate this into the likelihood function (Eqn.~\ref{eqn:lnlikelihood}).
-
-We empirically determine the interpolation error by performing a series of ``drop-out'' interpolation tests, whereby we remove a spectrum from the grid, generate an interpolated spectrum from the nearest two grid points, and compare the difference between the two.
-For example, consider the {\sc Phoenix} grid used for the WASP-14 parameter estimation.
-In this parameter space, the grid is sampled every 100K in $T_\textrm{eff}$ and every 0.5 dex in $\log_{10} g$ and $\Z$.
-Therefore, for each spectrum in the synthetic grid, there are also three error spectra associated with it, one for each axis of stellar parameters.
-For example, consider a spectrum with stellar parameters $T_\textrm{eff} = $ 6000K, $\log_{10} g=$4.0 dex, and $\Z = $ -0.5 dex.
-A temperature ``error spectrum'' would be formed by subtracting this true spectrum from the spectrum formed by the average of the $T_\textrm{eff} = $ 6100K spectrum and the $T_\textrm{eff} = $ 5900K spectrum, with $\log_{10} g$ and $\Z$ held constant.
-Likewise, the $\log_{10} g$ and $\Z$ error spectra are created by performing the same operation over the $\log_{10} g$ and $\Z$ axes of the grid (see Figure~\ref{fig:interp_errors}).
-
-These are done over a 200K range, and the actual range will only be 100K.
-Tests with other interpolation show that the interpolation error increases approximately linearly with interpolation distance.
-Therefore, when using the spectrum, we divide the total interpolation error spectrum by 2.
-
-Generally, the interpolation error is small, typically no larger than 1\% across many lines.
-This test gives us confidence that it is possible to generate an interpolated spectra, and that inference on the stellar parameters may be drawn at a substantially finer level that the coarsity of the grid spacing itself, especially given the multitude of possible lines to fit.
-However, while the interpolation error is small, it is often the case that the error adds \emph{coherently}.
-When the signal to noise of the spectrum is particularly large, this coherent trend can manifest itself as a behavior that the spectral lines in an interpolated spectrum are less ``physical'' looking than those directly near a gridpoint, with minor or no interpolation.
-Therefore, it is necessary to quantify the degree of interpolation error, in order to give the proper likelihood to interpolated spectra.
-
-\begin{figure}[!t]
-\begin{center}
-  \includegraphics{figs/interp_errors.pdf}
-  \figcaption{Also showing the envelope for a two gridpoint interpolation. All errors measured relative to the value of the true spectrum.}
-  \label{fig:interp_errors}
-\end{center}
-\end{figure}
-
-Nearly all interpolation errors are below the spread in change for an interpolated value.
-Those few pixels for which the change in value is \emph{less} than the bracketed change are due to values for which the pixel value changes non-monotonically with parameters.
-In such a situation, because the synthetic grid is not Nyquist sampled, it is difficult to accurately predict what the true pixel value should be.
-In this case, the most conservative approach is to incorporate the large interpolation error budget, as we do.
-
-\begin{figure}[!t]
-\begin{center}
-  \includegraphics{figs/interp_grid.pdf}
-  \figcaption{Tri-linear interpolation on a grid cell the shape of a rectangular prism.}
-  \label{fig:interp_grid}
-\end{center}
-\end{figure}
-
-Because the interpolation error is covariant, i.e., the interpolation error spectrum is highly dependent on whether the pixel is on the continuum, or on a spectral line (and specifically where on the spectral line), we must model the covariance of the errors.
-We must multiply and then add in order to preserve the covariance.
-A simple sum would destroy the covariance and the effect.
-
-For computational expediency and numerical stability, we smoothly taper the covariance to zero after a similar length scale as the global covariance kernel.
-In reality, this interpolation error matrix is dense: pixels separated by a wide spectral range may in fact exhibit a strong degree of correlation because they happen to fall on the same relative location of a spectral line.
-For example, all pixels at the continuum will tend to act similarly, regardless of where on the spectrum the pixels fall.
-Any pixel that falls on the blue side of a spectral line will tend to exhibit a similar error behaviour regardless of which spectral line, as long as that line was changing in a similar manner.
-
-The elements of the interpolation covariance matrix $\mathsf{C}_{ij}^\textrm{int}$ are determined empirically from the drop-out spectra, and are a function of the degree of the interpolation.
-The variables are notated such that the last index changes most rapidly.
-
-We could represent the average covariance matrix, for a specific axis $k$, completed from all of the grid points as
-
-\begin{equation}
-  \mathsf{C}_{ijk}^\textrm{int} = \sum_{l=1}^8  b_{kl} \lambda_{ikl} \lambda_{jkl}
-\end{equation}
-Then, the final combined matrix requires some weighting based upon interpolation distance in each of the three axes, $k$
-\begin{equation}
-  \mathsf{C}_{ij}^\textrm{int} = w_{ij}(r_0) \sum_{k=1}^3 a_k \mathsf{C}_{ijk}^\textrm{int}
-\end{equation}
-We also add in a Hann taper to ensure sparsity of the covariance matrix for computational efficiency, however note that the interpolation error covariance matrix is itself dense.
-
-Here, $b_{kl}$ corresponds to the same weights that would be applied to each spectra in the interpolation case, for a specific axis $k$.
-The estimate of covariance is determined as a weighted average between the covariance matrices at all of the grid points.
-
-Then, the final covariance
-
-[Figure showing what the truncated covariance matrix looks like, with high values about lines].
-
 \end{document}

### 2014-09-22 tex/paper.tex
PCA work begun

@@
 
-\acknowledgments  List your acknowledgments (don't forget anyone!).  I can think of: 
-Foreman-Mackey, Hogg, people on Johnson's group?, Lars?, TRES archive people?  You might want to 
-acknowledge your NSF Fellowship here, and the Smithsonian Institution (how you're being paid now), 
-for funding.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}. \\
+\acknowledgments  The authors would like to graciously acknowledge the following people for many helpful discussions and key insights: Gregory M.
+Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Allyson Bieryla, and Maxwell Moe.
+IC was supported by the NSF Graduate Fellowship and the Smithsonian Institution.
+Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
+Astropy.
+JuliaLang.\\
 
@@
 for an arbitrary set of parameters; $\vt_{\ast} = T_\textrm{eff} = 6220$, K, $\log g=4.2$ dex, and 
-$\Z = $-0.4 dex$]$.  The color scale is the same as in Fig.~\ref{fig:matrix}; since elements of 
+$\Z = $-0.4 dex.  The color scale is the same as in Fig.~\ref{fig:matrix}; since elements of 
 this covariance kernel can be negative, an analogous scale (in {\it red}) is duplicated to 

### 2014-10-03 Starfish/model.py
changed around jumiping

@@ -249,9 +249,10 @@ class Model:
         '''
         Define the prior here
         '''
-        logg = params["logg"]
+        #logg = params["logg"]
         #if (logg >= 4.8) and (logg <= 5.2):
-        return -0.5 * (logg - 5.0)**2/(0.15)**2
+        #return -0.5 * (logg - 5.0)**2/(0.15)**2
+        return 0.0
         #else:
         #    return -np.inf
 
@@ -697,7 +698,7 @@ class StellarSampler(Sampler):
         self.param_tuple = C.dictkeys_to_tuple(starting_param_dict)
         print("param_tuple is {}".format(self.param_tuple))
 
-        kwargs.update({"revertfn":self.revertfn, "lnprobfn":self.lnprob})
+        kwargs.update({"revertfn":self.revertfn, "lnprobfn":self.lnprob, "resample":True})
         super(StellarSampler, self).__init__(**kwargs)
         self.fname = "stellar"
 

### 2014-10-08 hdfplot.py
Updated for a longer Gl51 run

@@ -11,5 +11,9 @@ Script designed to plot the HDF5 output from MCMC runs. It will only plot the fi
 if hdfutils.args.cov:
     hdfutils.estimate_covariance(hdfutils.flatchainTreeList[0])
 
-hdfutils.plot(hdfutils.flatchainTreeList[0], clip_stellar=hdfutils.args.stellar_params)
+if hdfutils.args.paper:
+    hdfutils.plot_paper(hdfutils.flatchainTreeList[0], clip_stellar=hdfutils.args.stellar_params)
+else:
+    hdfutils.plot(hdfutils.flatchainTreeList[0], clip_stellar=hdfutils.args.stellar_params)
+
 

### 2014-10-08 hdfutils.py

@@ -42,6 +42,8 @@ parser.add_argument("--acor", action="store_true", help="Calculate the autocorre
 parser.add_argument("--acor-window", type=int, default=50, help="window to compute acor with")
 
 parser.add_argument("--cov", action="store_true", help="Estimate the covariance between two parameters.")
+parser.add_argument("--paper", action="store_true", help="Change the figure plotting options appropriate for the "
+                                                         "paper.")
 
 args = parser.parse_args()
 
@@ -59,11 +61,11 @@ args.outdir += "/"
 import h5py
 
 #Plot kw
-label_dict = {"temp":r"$T_{\rm eff}$", "logg":r"$\log_{10} g$", "Z":r"$[{\rm Fe}/{\rm H}]$",
+label_dict = {"temp":r"$T_{\rm eff}\;[{\rm K}]$", "logg":r"$\log g$", "Z":r"$[{\rm Fe}/{\rm H}]$",
               "alpha":r"$[\alpha/{\rm Fe}]$",
-              "vsini":r"$v \sin i$",
+              "vsini":r"$v \sin i\;[{\rm km\;s}^{-1}]$",
               "vz":r"$v_z$", "vz1":r"$v_z1$", "vz2":r"$v_z2$", "vz3":r"$v_z3$",
-              "logOmega":r"$\log_{10} \Omega$",
+              "logOmega":r"$\log \Omega$",
               "logOmega1":r"$\log{10} \Omega 1$",
               "logOmega2":r"$\log{10} \Omega 2$",
               "logOmega3":r"$\log{10} \Omega 3$",
@@ -404,9 +406,60 @@ def plot(flatchainTree, base=args.outdir, triangle_plot=args.triangle, chain_plo
         labels = [label_dict.get(key, "unknown") for key in params]
 
         figure = triangle.corner(samples, labels=labels, quantiles=[0.16, 0.5, 0.84],
-                                 show_titles=True, title_args={"fontsize": 16})
+                                 show_titles=True, title_args={"fontsize": 16}, plot_contours=True,
+                                 plot_datapoints=False)
         figure.savefig(base + flatchain.id + format)
 
+
+def plot_paper(flatchainTree, base=args.outdir, triangle_plot=args.triangle, chain_plot=args.chain,
+         lnprob=args.lnprob, clip_stellar='all', format=".pdf"):
+    '''
+    Make a bunch of plots to diagnose how the run went.
+    '''
+
[…]

### 2014-10-13 tex/paper.tex
changed around jumiping
Updated for a longer Gl51 run
Some progress towards marginalizing over weights

@@
 
-\usepackage[breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[backref,breaklinks,colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage[all]{hypcap}
+\renewcommand*{\backref}[1]{[#1]}
 \usepackage{graphicx}
@@
 
+\capstartfalse
 \begin{deluxetable}{lr@{ $\pm$ }l|r@{ $\pm$ }l}[!b]
@@
 \end{deluxetable}
+\capstarttrue
 
 \begin{figure}[!b]
-  \includegraphics[width=0.5\textwidth]{figs/Kurucz_triangle.pdf}
+  \includegraphics{figs/Kurucz_triangle.pdf}
 \vspace{-0.5cm}
@@
 
+\capstartfalse
 \begin{deluxetable}{lr@{ $\pm$ }l}[h]
@@
 \end{deluxetable}
+\capstarttrue
 
@@
 Astropy.
-JuliaLang.\\
+JuliaLang.
+Python/Scikit-learn.\\
 
@@
 
-\section{Interpolation Errors} \label{sec:Appendix}
+\section{Spectral Emulator} \label{sec:Appendix}
 
@@
 library spectra with similar physical parameters.  The goal in this Appendix is to describe a 
-procedure by which we propagate an empirical approximation of the {\it inaccuracy} in that 
+procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
 prediction through the probabilistic framework we have developed.
@@
 parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
-information loss.  Given these factors, we make the conservative (and computationally tractable) 
-assumption and adopt a linear interpolation scheme.  More advanced interpolation methods, like a 
-spline or other band-limited approach, do not change the fact that the library grid is not Nyquist 
-sampled, and so do not improve performance.  To account for the loss of spectral information during 
-the interpolation process, we have to build a model of the interpolation errors and incorporate it 
-into the likelihood function (via the covariance matrix).
+information loss.
+Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra, seeks to non-deterministically predict the behavior of a spectrum for stellar parameters not sampled by the library.
 
@@
 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but the {\sc 
-Phoenix} library has finer coverage in 100\,K increments.  We empirically determine a model grid of 
-interpolation errors through a series of ``drop-out" interpolation tests.  The idea is to remove a 
-spectrum from the library, $f_{\lambda}(\vt_{\ast})$, generate an interpolated spectrum, 
-$\tilde{f}_{\lambda}(\vt_{\ast})$, at the parameters corresponding to that removed spectrum from 
-the neighboring spectra, and then compare the two.  The result is a set of three ``error" (or, 
-rather ``difference") spectra, each representing an individual dimension in parameter-space, where
-\begin{eqnarray} \label{eqn:interr}
-e_{\lambda}(\vt_{\ast}^{k}) &=& f_{\lambda}(\vt_{\ast}^k) - \tilde{f}_{\lambda}(\vt_{\ast}^k) \\
-                            &=& f_{\lambda}(\vt_{\ast}^k) - \frac{1}{2}\left[f_{\lambda}(\vt_{\ast}^{k+1})+f_{\lambda}(\vt_{\ast}^{k-1}))\right], \nonumber
-\end{eqnarray}
-for $k = 1, 2, 3$ indexing the library parameters ($\vt_{\ast}$).
-
-As a concrete example, consider the stellar parameters $\vt_{\ast}$ = [6000\,K, 4.0 dex, -0.5 
-dex].  An ``error spectrum" in the $T_{\rm eff}$ dimension, $e_{\lambda}(\vt_{\ast}^1)$, would be 
-calculated from the difference between the grid spectrum at that $\vt_{\ast}$ and the average of 
-$f_{\lambda}$([5900\,K, 4.0 dex, -0.5 dex]) and $f_{\lambda}$([6100\,K, 4.0 dex, -0.5 dex]), as 
-defined in Eq.~\ref{eqn:interr}.  Likewise for the error spectra in the $\log g$ and $\Z$ 
-dimensions, $e_{\lambda}(\vt_{\ast}^2)$ and $e_{\lambda}(\vt_{\ast}^3)$.  A graphical demonstration 
-is shown in Figure~\ref{fig:interp_errors}.  These kinds of drop-out tests over larger spans of the 
-grid show that the amplitude of $e_{\lambda}$ grows approximately linearly with the interpolation 
-distance.
-
-\begin{figure*}[!t]
-\begin{center}
-  \includegraphics{figs/interp_errors.pdf}
-\vspace{-0.1cm}
-  \figcaption{(\emph{top}) A small subset of a {\sc Phoenix} model spectrum at library gridpoint 
-defined by $\vt_{\ast} = [T_\textrm{eff} = 6000$\,K, $\log g=4.0$ dex, and $\Z = $-0.5 dex$]$, 
-shown with the drop-out interpolation spectra ($\tilde{f}_{\lambda}$) along each of the three 
-parameter axes.  The interpolated spectra closely match the ``true'' spectrum.  (\emph{bottom}) The 
-three corresponding ``error" spectra plotted {\it relative to the true spectrum} (i.e., 
-$e_{\lambda}/f_{\lambda}$).  The shaded regions show the total change in $f_{\lambda}$ between the 
-two neighboring spectra in the library used for the interpolation (again relative to the true 
-spectrum).  Note that the error spectrum is generally much smaller than the typical change in flux 
-across gridpoints.  For the few points where this is not true (e.g., $\log g$ near 5161.5\AA), this 
-indicates non-monotonic changes between gridpoints.  Therefore, it is most conservative to build an 
-interpolation error model that accounts for this behavior and protects against assumptions of 
-monotonicity. \label{fig:interp_errors}}
-\end{center}
-\end{figure*}
-
-The interpolation error is generally small, $\lesssim$2\%, much less than the Poisson noise for 
-spectra with moderate sensitivity and usually (although not always) less than the difference 
-between spectra at adjacent grid points.  This gives us confidence that it is possible to generate 
-accurate interpolated spectra, and that any inference on the stellar parameters can be made at a 
-level substantially more precise than dictated by the coarse spacing of the library itself 
-(especially for such a large data volume).  However, although $e_{\lambda}$ is small in magnitude, 
-the fact that spectral lines have similar shapes means that the interpolation error often compounds 
-\emph{coherently}.  That coherence can substantially diminish the likelihood of an interpolated 
-spectrum, since its spectral line shapes are less ``physical" than for a spectrum located directly 
-on a computed gridpoint.  In essence, this highlights the underlying {\it necessity} of propagating 
-an interpolation error to deliver a proper likelihood at interpolated parameter values.
-
-Because the magnitude of $e_{\lambda}$ is strongly dependent on its association with the continuum 
-or a spectral line, the interpolation error model must be based on a (spectral) covariance kernel.  
-The elements of that kernel, $\mathcal{K}_{ij}^{\rm int}$, are determined from the empirical model 
-grid of $e_{\lambda}$ calculated from the drop-out technique.  The interpolated model spectrum and 
-is a weighted average of the neighboring 8 gridpoints (standard tri-linear interpolation).  To 
-construct $\mathcal{K}_{ij}^{\rm int}$ from the associated 24 interpolation error spectra, we first 
-need to generate sub-kernels that represent each dimension in parameter-space, 
-\begin{equation} \label{eq:subsubkernel}
-^k_l\mathcal{K}_{ij}^{\rm int} = \langle e_{\lambda_i}^k \; e_{\lambda_j}^k \rangle_l,
-\end{equation}
-where again $k$ indexes each element of $\vt_{\ast}$, and $l$ indexes each of the neighboring 
-gridpoints (and $i, j$ denote a pixel pairing, as in Sect.~\ref{sec:method}).  The sub-kernel for 
-dimension $k$ corresponding to the desired interpolate location is a simple weighted sum of the 
-matrices in Eq.~\ref{eq:subsubkernel},
-\begin{equation} \label{eq:subkernel}
-^k\mathcal{K}_{ij}^{\rm int} = \sum_{l=1}^8 \Delta_l \,\, ^k_l\mathcal{K}_{ij}^{\rm int},
-\end{equation}
-where the weights $\Delta_l$ depend on the distance from each grid point to the interpolate 
-location.
-
-\begin{figure}[!b]
-\begin{center}
-  \includegraphics{figs/interp_matrix.pdf}
-  \figcaption{A sub-region of a typical interpolation error kernel, $\mathcal{K}_{\ij}^{\rm int}$, 
-for an arbitrary set of parameters; $\vt_{\ast} = T_\textrm{eff} = 6220$, K, $\log g=4.2$ dex, and 
-$\Z = $-0.4 dex.  The color scale is the same as in Fig.~\ref{fig:matrix}; since elements of 
-this covariance kernel can be negative, an analogous scale (in {\it red}) is duplicated to 
-highlight such regions with the same logarithmic stretch.  \label{fig:interp_matrix}}
-\end{center}
-\end{figure}
-
-Finally, we average the 3 sub-kernels in Eq.~\ref{eq:subkernel}, representing the covariance 
-introduced in each parameter dimension at the interpolate location, as 
-\begin{equation}
-\mathcal{K}_{ij}^{\rm int} = w_{ij}(r_0) \sum_{k=1}^3 \varrho_k \, ^k\mathcal{K}_{ij}^{\rm int}.
-\end{equation}
-Here, the weights $\varrho_k$ are based on the interpolation distance in parameter-space.  For 
-example, if the target interpolate distance were $T_{\rm eff} = 6120$\,K then $\varrho_1 = 0.4$ 
-(as a reminder, the library gridpoints are spaced by 100\,K in this dimension); for $T_{\rm eff} = 
-6100$\,K, $\varrho_1 = 0.0$; and for $T_{\rm eff} = 6150$\,K, the maximal interpolation distance, 
-$\varrho_1 = 1.0$.  
-For computational expediency and numerical stability, we smoothly taper the covariance kernel to 
-zero at the same length scale ($r_0 = 4 \ell$) as the global covariance kernel, again employing a 
-Hann window.  In reality, $\mathcal{K}^{\rm int}$ is a dense matrix: pixels separated by a wide 
-spectral range may in fact exhibit a strong degree of correlation because they happen to fall on 
-the same relative location of a spectral line.  A representative interpolation error kernel is 
-shown in Figure~\ref{fig:interp_matrix}. \\
+Phoenix} library has finer coverage in 100\,K increments.
+
+The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation.
+We use principal component analysis (PCA) to decompose the synthetic library into a set of ``eigenspectra'' $\varphi$.
+Following techniques of SLOAN (cite Ivecik 2014).
+After examining how well the spectra reconstruct the grid, we keep the first $m$ of the eigenspectra, and discard the latter eigenspectra.
+We keep enough eigenspectra to reconstruct any pixel flux to better than $2\%$, however most pixels are more like $XX\%$.
+$\varphi_k$, $k = \{1, 2, \ldots, m\}$.
+Before we do the PCA operation, however, we first standardize the synthetic library by subtracting the mean spectrum and ``whitening'' the spectra by dividing by the standard deviation in each pixel.
+First, we select only a subset of the spectra synthetic library which will be relevant to our  The mean spectrum is
+\begin{equation}
+  f_{\lambda, \mu} = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
+\end{equation}
+and the standard deviation spectrum is
+\begin{equation}
+  f_{\lambda, \sigma} = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - f_{\lambda, \mu} \bigr]^2 } 
+\end{equation}
+So a spectrum from the standardized library is given by
+\begin{equation}
+  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - f_{\lambda, \mu}}{f_{\lambda, \sigma}}
+\end{equation}
+Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik). To speed computation, we first select a
+
+Using the PCA basis, we can lossily reconstruct any spectrum from the library by a linear combination of the eigenspactra
+\begin{equation}
+  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}) \varphi_k
+\end{equation}
+where $w_k$ is the weight of a given eigenspectrum, $k$.
+This weight is a three dimensional scalar function, which depends on the stellar parameters $\vt_\ast$.
+By projecting the spectra from the synthetic library onto the eigenspectra, we can determine the weights at the grid points in the library, $\{\vt_\ast \}^\textrm{grid}$.
+For a given eigenspectrum $\varphi_k$, the weight is
+\begin{equation}
+  w_k(\{\vt_\ast \}^\textrm{grid}) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \varphi_k
+\end{equation}
+These weights are generally a smooth function of stellar parameters.
 

### 2014-10-13 scripts/stars/base_lnprob.py

@@ -108,8 +108,22 @@ myInstrument = TRES()
 stellar_Starting = config['stellar_params']
 stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
 #go for each item in stellar_tuple, and assign the appropriate covariance to it
-stellar_MH_cov = np.array([float(config["stellar_jump"][key]) for key in stellar_tuple])**2 \
-                 * np.identity(len(stellar_Starting))
+#stellar_MH_cov = np.array([float(config["stellar_jump"][key]) for key in stellar_tuple])**2 \
+#                 * np.identity(len(stellar_Starting))
+stellar_MH_cov = np.array([float(config["stellar_jump"][key]) for key in stellar_tuple])**2
+
+temulator = Emulator.open(config['PCA_path'])
+#Call the emulator at the starting stellar parameters
+pp = np.array([stellar_Starting["temp"], stellar_Starting["logg"], stellar_Starting["Z"]])
+starting_Weights = temulator.draw_weights(pp)
+stellar_Starting["weights"] = starting_Weights
+
+weight_mu, weight_cov = temulator(pp)
+weight_cov = weight_cov * config["frac_weight"]
+
+stellar_MH_cov = np.concatenate((stellar_MH_cov, weight_cov))
+stellar_MH_cov = stellar_MH_cov * np.identity(len(stellar_MH_cov))
+print(len(stellar_MH_cov))
 
 fix_logg = config.get("fix_logg", None)
 

### 2014-10-13 Starfish/emulator.py
Some progress towards marginalizing over weights

@@ -316,14 +316,10 @@ class WeightEmulator:
         self.mu.shape = (-1)
         self.sig = self.V22 - self.V12.T.dot(np.linalg.solve(self.V11, self.V12))
 
-    def draw_weights(self):
+    def draw_weights(self, *args):
         '''
         Using the current settings, draw a sample of PCA weights
-        '''
-        return np.random.multivariate_normal(self.mu, self.sig)
 
-    def __call__(self, *args):
-        '''
         If you call this with an arg that is an np.array, it will set the emulator to these parameters first and then
         draw weights.
 
@@ -347,8 +343,34 @@ class WeightEmulator:
             print("No parameters are set, yet. Must set parameters first.")
             return
 
-        weights = self.draw_weights()
-        return weights
+
+        return np.random.multivariate_normal(self.mu, self.sig)
+
+    def __call__(self, *args):
+        '''
+        If you call this with an arg that is an np.array, it will set the emulator to these parameters first and then
+        draw weights.
+
+        If no args are provided, then the emulator uses the previous parameters.
+
+        If there are samples defined, it will also reseed the emulator parameters by randomly draw a parameter
+        combination from MCMC samples.
+        '''
+        if self.samples is not None:
+            self.emulator_params = self.samples[np.random.choice(self.indexes)]
+            if not args:
+                #Reset V12, V22, since we also changed emulator paramaters.
+                self.params = self._params
+
+        if args:
+            params, *junk = args
+            self.params = params
+
+        if self.V12 is None:
+            print("No parameters are set, yet. Must set parameters first.")
+            return
+
+        return (self.mu, self.sig)
 
 class Emulator:
     '''
@@ -426,15 +448,18 @@ class Emulator:
 
         self._params = pars
 
-    def draw_weights(self):
+    def draw_weights(self, *args):
         '''
         Draw a weight from each WeightEmulator and return as an array.
         '''
-        return np.array([WE(self._params) for WE in self.WEs])
+        if args:
+            params, *junk = args
+            self.params = params
+        return np.array([WE.draw_weights(self._params) for WE in self.WEs])
 
-    def __call__(self, *args):
+    def reconstruct_draw(self, *args):
         '''
-        Same behavior as with WeightEmulator.
+        Reconstructing a spectrum. Taking the place of the old __call__ behavior.
         '''
[…]

### 2014-10-13 Starfish/model.py

@@ -246,12 +246,17 @@ class Model:
         '''
         Define the prior here
         '''
+
+        #Evaluate prior values of weights.
+        w = params["weights"]
+        #Get the mean and sigma of p(w | theta_stellar)
+        mu, sigma = self.ModelSpectrum.Emulator(w)
+        wprior = -np.log(sigma) - 0.5 * (w - mu)**2 - 0.5 * np.log(2. * np.pi)
+
         logg = params["logg"]
-        #if (logg >= 4.8) and (logg <= 5.2):
-        return -0.5 * (logg - 5.0)**2/(0.05)**2
-        #return 0.0
-        #else:
-        #    return -np.inf
+        logg_prior = -0.5 * (logg - 5.0)**2/(0.05)**2
+
+        return wprior + logg_prior
 
     def to_json(self, fname="model.json"):
         '''
@@ -692,6 +697,10 @@ class StellarSampler(Sampler):
         self.fix_logg = kwargs.get("fix_logg", False)
 
         starting_param_dict = kwargs.get("starting_param_dict")
+        #Expand weights component
+        weights = starting_param_dict["weights"]
+
+
         self.param_tuple = C.dictkeys_to_tuple(starting_param_dict)
         print("param_tuple is {}".format(self.param_tuple))
 
@@ -734,13 +743,19 @@ class StellarSampler(Sampler):
 
         self.logger.debug("params in lnprob are {}".format(params))
 
+
+        #Now we need to have weights as the last part of the array.
+        ncomp = self.model_list[0].ModelSpectrum.Emulator.PCAGrid.ncomp
+        weights = others[-ncomp:]
+        others = others[:-ncomp]
+
         #others should now be either [vz, logOmega] or [vz0, logOmega0, vz1, logOmega1, ...] etc. Always div by 2.
         #split p up into [vz, logOmega], [vz, logOmega] pairs that update the other parameters.
         #mparams is now a list of parameter dictionaries
         mparams = deque()
         for vz, logOmega in grouper(others, 2):
             p = params.copy()
-            p.update({"vz":vz, "logOmega":logOmega})
+            p.update({"vz":vz, "logOmega":logOmega, "weights":weights})
             mparams.append(p)
 
         self.logger.debug("updated lnprob params: {}".format([params for params in mparams]))

### 2014-10-29 Starfish/model.py
Discarding regions in new format

@@ -1,20 +1,11 @@
 import numpy as np
-import sys
-import emcee
-from emcee import GibbsSampler, GibbsSubController, GibbsController, ParallelSampler
+from emcee import GibbsSampler, ParallelSampler
 from . import constants as C
-from .grid_tools import Interpolator, ErrorInterpolator
-from .spectrum import ModelSpectrum, ChebyshevSpectrum, ModelSpectrumHA
-from .covariance import CovarianceMatrix #from Starfish.spectrum import CovarianceMatrix #pure Python
-from .emulator import Emulator
-import time
+from .grid_tools import Interpolator
+from .spectrum import ModelSpectrum, ModelSpectrumHA
 import json
 import h5py
-from astropy.stats.funcs import sigma_clip
 import logging
-import os
-from collections import deque
-from operator import itemgetter
 import matplotlib.pyplot as plt
 from itertools import zip_longest
 
@@ -805,7 +796,8 @@ class NuisanceSampler(Sampler):
 
         self.model = kwargs.get("OrderModel")
         spectrum_id, order_id = self.model.id
-        self.fname = "{}/{}/{}".format(spectrum_id, order_id, "nuisance")
+        order = kwargs.get("order", order_id)
+        self.fname = "{}/{}/{}".format(spectrum_id, order, "nuisance")
         self.params = None
         self.prior_params = kwargs.get("prior_params", None)
         if self.prior_params:

### 2014-10-29 hdfutils.py

@@ -73,6 +73,19 @@ label_dict = {"temp":r"$T_{\rm eff}\;[{\rm K}]$", "logg":r"$\log g$", "Z":r"$[{\
               "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}$", "l":r"$l$",
               "h":r"$h$", "loga":r"$\log_{10} a$", "mu":r"$\mu$", "sigma":r"$\sigma$"}
 
+import re
+p = re.compile("r\d\d*", re.IGNORECASE)
+
+def not_region(str):
+    '''
+    Return true if this is not a region
+    '''
+    m = p.match(str)
+    if m:
+        return False
+    else:
+        return True
+
 #Additionally, there should be an object for each flatchain, that stores param_tuple and samples
 class Flatchain:
     '''
@@ -106,7 +119,13 @@ class Flatchain:
         param_tuple = tuple([param.strip("'() ") for param in param_tuple.split(",")])
         samples = dset[:]
 
-        return cls(id, param_tuple, samples)
+        #If we have regions, then separate them and discard for now.
+        ind = np.array([not_region(param) for param in param_tuple], dtype='bool')
+        print("Non regions", ind)
+
+        param_tuple = tuple([param_tuple[i] for i in range(len(param_tuple)) if ind[i]])
+
+        return cls(id, param_tuple, samples[:,ind])
 
     @property
     def shape(self):
@@ -571,17 +590,12 @@ def estimate_covariance(flatchainTree):
         print(cor)
 
 
-# plot_walkers(self.outdir + self.fname + "_chain_pos.png", samples, labels=self.param_tuple)
-# plt.close(figure)
-
 if args.acor:
     from emcee import autocorr
     print("Stellar Autocorrelation")
     acors = autocorr.integrated_time(stellar, axis=0, window=args.acor_window)
     for param,acor in zip(stellar_tuple, acors):
         print("{} : {}".format(param, acor))
-
-
 
 #Now that all of the structures have been declared, do the initialization stuff.
 

### 2014-10-29 scripts/stars/region_lnprob.py

@@ -1,191 +1,0 @@
-from StellarSpectra.model import Model, StellarSampler, ChebSampler, CovGlobalSampler, RegionsSampler, MegaSampler
-from StellarSpectra.spectrum import DataSpectrum
-from StellarSpectra.grid_tools import TRES, HDF5Interface
-import StellarSpectra.constants as C
-import numpy as np
-import yaml
-import os
-import shutil
-
-
-import argparse
-parser = argparse.ArgumentParser(prog="base_lnprob.py", description="Run StellarSpectra fitting model.")
-parser.add_argument("-i", "--input", help="*.yaml file specifying parameters.")
-parser.add_argument("-r", "--run_index", help="Which run (of those running concurrently) is this? All data will "
-                                        "be written into this directory, overwriting any that exists.")
-parser.add_argument("-p", "--perturb", type=float, help="Randomly perturb the starting position of the "
-                                                                 "chain, as a multiple of the jump parameters.")
-args = parser.parse_args()
-
-if args.input: #
-    #assert that we actually specified a *.yaml file
-    if ".yaml" in args.input:
-        yaml_file = args.input
-        f = open(args.input)
-        config = yaml.load(f)
-        f.close()
-
-    else:
-        import sys
-        sys.exit("Must specify a *.yaml file.")
-
-else:
-    #load the default config file
-    yaml_file = "scripts/stars/input.yaml"
-    f = open(yaml_file)
-    config = yaml.load(f)
-    f.close()
-
-def perturb(startingDict, jumpDict, factor=3.):
-    '''
-    Given a starting parameter dictionary loaded from a config file, perturb the values as a multiple of the jump
-    distribution. This is designed so that chains do not all start in the same place, when run in parallel.
-
-    Modifies the startingDict
-    '''
-    for key in startingDict.keys():
-        startingDict[key] += factor * np.random.normal(loc=0, scale=jumpDict[key])
-
-
-myDataSpectrum = DataSpectrum.open(config['data'], orders=config['orders'])
-#Load mask and add it to DataSpectrum
-mask = np.load("notebooks/WASP14_24.mask.npy")
[…]

### 2014-10-29 tex/paper.tex
Discarding regions in new format

@@
 emcee} package \citep{foreman-mackey13} to function within a blocked Gibbs 
-sampler.\footnote{\url{https://github.com/iancze/emcee}} \\
+sampler.\footnote{\url{https://github.com/iancze/emcee}} For echelle spectra with multiple orders or multiple spectra of the same object, taken with different instruments, the nuisance parameters for each chunk of spectrum are independent.
+Additionally, the computationally intensive steps of generating a model spectrum for a specific wavelength range and evaluating the likelihood on this spectral chunk can also be parallelized.
+Therefore, the only part of the code that needs to be synchronized is the MCMC proposal of stellar parameters.
+Because all of the computationally expensive steps can be parallelize, this algorithim has the ability to be massively parallelized on a cluster, enabling simultaneous inference for wide spectral ranges or multiple spectral datasets acquired with different instruments.\\
 
@@
 We keep enough eigenspectra to reconstruct any pixel flux to better than $2\%$, however most pixels are more like $XX\%$.
-$\varphi_k$, $k = \{1, 2, \ldots, m\}$.
+An $N_{\rm pix}$ eigenspectrum $\xi_k$, $k = \{1, 2, \ldots, m\}$.
 Before we do the PCA operation, however, we first standardize the synthetic library by subtracting the mean spectrum and ``whitening'' the spectra by dividing by the standard deviation in each pixel.
@@
 \begin{equation}
-  f_{\lambda, \mu} = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
+  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
 \end{equation}
@@
 \begin{equation}
-  f_{\lambda, \sigma} = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - f_{\lambda, \mu} \bigr]^2 } 
-\end{equation}
-So a spectrum from the standardized library is given by
-\begin{equation}
-  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - f_{\lambda, \mu}}{f_{\lambda, \sigma}}
-\end{equation}
-Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik). To speed computation, we first select a
-
-Using the PCA basis, we can lossily reconstruct any spectrum from the library by a linear combination of the eigenspactra
-\begin{equation}
-  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}) \varphi_k
+  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu \bigr]^2 } 
+\end{equation}
+We standardize all spectra in the library by subtracting the mean spectrum and dividing by the standard deviation spectrum.
+\begin{equation}
+  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
+\end{equation}
+Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik, scikit-learn). To speed computation, we first select $M$ subset spectra from the grid.
+
+[For an example of the eigenspectra for Gl51, see figure A1 (left) and reconstruction of a synthetic spectrum at stellar parameters XX (right).]
+
+Using the PCA basis, we can lossily reconstruct any spectrum from the library by a linear combination of the eigenspectra
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}) \xi_k
+  \label{eqn:reconstruct}
 \end{equation}
@@
 This weight is a three dimensional scalar function, which depends on the stellar parameters $\vt_\ast$.
-By projecting the spectra from the synthetic library onto the eigenspectra, we can determine the weights at the grid points in the library, $\{\vt_\ast \}^\textrm{grid}$.
-For a given eigenspectrum $\varphi_k$, the weight is
-\begin{equation}
-  w_k(\{\vt_\ast \}^\textrm{grid}) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \varphi_k
-\end{equation}
-These weights are generally a smooth function of stellar parameters.
+We can determine the weights at the grid points in the library by taking the dot product of the standardized synthetic spectrum with each of the eigenspectra.
+For a given eigenspectrum $\xi_k$, the weight is
+\begin{equation}
+  w_k(\{\vt_\ast \}^\textrm{grid}) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \xi_k
+\end{equation}
+These weights are generally a smooth function of stellar parameters. To simplify the notation, we can write the weights in a length-$m$ column vector
+\begin{equation}
+  \mathbf{w} = 
+  \begin{bmatrix}
+    w_1\\
+    w_2\\
+    \vdots\\
+    w_m
+  \end{bmatrix}
+\end{equation}
+and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
+\begin{equation}
+  \mathbf{\Xi} = 
+  \begin{bmatrix}
+    \xi_1 & \xi_2 & \cdots & \xi_m \\
+  \end{bmatrix}
+\end{equation}
+Therefore we can rewrite Equation~\ref{eqn:reconstruct} as
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}
+\end{equation}
+Where $I_{N_{\rm pix}}$ is an $N_{\rm pix}$ by $N_{\rm pix}$ identity matrix.
+To summarize, thus far we have decomposed the spectra in the synthetic library into a principal component basis, which allows us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
+The weight corresponding to each eigenspectrum is a relatively-smooth three-dimensional scalar function of the stellar parameters.
+Therefore, to interpolate a spectrum from the synthetic library for a value of stellar parameters which is not in the spectral library, we must interpolate the value of the weights to an off-grid value.
+In practice, it might be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly.
+However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum could constitute a significant portion of the uncertainty budget.
+Therefore, we decide to develop a Bayesian spectral emulator (cite Habib, Heitmann) which allows us to propagate the uncertainty in interpolation forward to the likelihood calculation.
+First, we describe how we use a Gaussian process to model the interpolation of the weights then we marginalize over all possible interpolations to arrive at a modified likelihood function.
+
+In Section XX, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral regions of the residuals due to pathological differences in the data spectrum and synthetic spectrum.
+Now, we use a separate Gaussian process in a slightly-different manner, in order to serve as an interpolating function for the weights.
+This is using a Gaussian process for prediction, given the values of the weights at several known locations (the values of stellar parameters at the grid points) and some assumed properties of the unknown functional form of the weights, what are the most probable values of the weights at stellar parameters in between the grid points?
+This is similar to drawing random values from the covariance matrix in Section XX given covariance matrices filled in by covariance kernels.
+Drawing from one giant Gaussian.
+Only now, we are drawing from a giant Gaussian conditioned on several points, which are the weights at the grid point locations.
+
+[Fig (three panels?) showing the values of the weight at different 1-d locations, and then many possible interpolations of that weight.
+Also one showing the probability distribution of the weight at a fixed stellar parameters (Gaussian in shape) and mentioning that we can marginalize over.]
+
+The smooth the distribution of weights (see Fig A.2, left panel) with changing stellar parameters motivated the use of a squared-exponential covariance kernel for the Gaussian process (cite Rasmussen + Williams for it's smooth properties).
+This is in contrast to the \matern\ kernel used in Section XX, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
+
+Consider the modelling of a single weight. We model this using a squared exponential covariance kernel
+
+Eqn of Squared exponential.
+
+We create a different covariance matrix $C_\textrm{w}$ which models the covariance between the weights of all the $M$ points in the grid. Here show V11, V12, V22, etc.
+
+Given the values of the weights at the grid point locations, we can define a posterior probability distribution for the parameters of the Gaussian process.
+Following Habib, Heitmann, we adopt priors on the parameters of the GP in order to regularize it [TODO: better exp].
+
+Eqn of full posterior distribution for a given weight.
+
+We use a downhill simplex algorithm (Scipy/fmin) to find the maximum of the posterior with respect to the GP parameters.
+From here on, we fix the GP hyperparamaters to their optimum values and consider the GP specified.
+The benefit of using a Gaussian process for prediction/regression is that we now specified a probability distribution over possible interpolated values of the weights.
+We can draw multivariate samples from this probability distribution, like we did for the covariance matrices in Section XX to demonstrate the residuals (add note there to see here), only now we can also condition on the observed weight values.
+Notice how the GP directly interpolates the values at the grid points.
+For example, Fig A.2 (middle) shows draws from vectors of densely packed stellar values in order to show that simultaneous multivariate draws along a vector exhibit a certain structure of correlations (similar to the residuals in Section XX.
+However, the spread of subsequent multivariate draws at a specific single stellar value is given by a Gaussian distribution (Fig A.2 right).
+
+Initially, we presented the reconstruction as a linear reconstruction of the eigenspectra resulting from the decomposition of the synthetic grid.
+In practice, however, we must still post-process the synthetic spectra to match the data spectrum (Section XX), applying convolutions for instrumental broadening, rotational broadening, and Doppler shift, and then lastly re-sampling the spectra to the exact pixels of the dataspectrum.
+Fortunately, because the convolution operation is a linear operator, we can first apply these operations to the eigenspectra themselves, without loss of information.
+Unfortunately, the Doppler shift and resampling operation is a non-linear operation and there will be some loss of information.
+However, we find that in practice when using a reasonable oversampling ratio of the sythetic to the dataspectra, applying the Doppler shift/resampling operation to the eigenspectra results in smaller than a 0.2\% error across all pixels, and thus the effect is negligable.
+
+Therefore, we can rewrite the reconstruction of the model spectrum as a linear combination of eigenspectra.
+
+[Equation rewriting the blended/resampled eigenspectra as a function of thetaOBS]
+[Equation rewriting reconstruction but using the operators on the eigenspectra]
+
+Since the Gaussian process delivers a probability distribution (Gaussian in form) for each weight, given a value of stellar parameters, we now have a probability distribution of a given reconstruction of a spectrum, which is just a linear combination of the eigenspectra.
+Because both the data likelihood and the probability distribution for the weights are multidimensional Gaussians, we can write down the joint probability distribution for the data and the interpolations, and then \emph{analytically} marginalize over the weights, thus marginalizing over the space of all possible interpolations given the principal component basis.
+
+Cite Mandel, Bishop, Gelman.
+
+The final modified covariance matrix looks like Fig A.3.
 

### 2014-10-29 Starfish/parallel.py

@@ -81,6 +81,7 @@ if type(data) != list:
     data = [data]
 print("loading data spectra {}".format(data))
 orders = config["orders"] #list of which orders to fit
+order_ids = np.arange(len(orders))
 DataSpectra = [DataSpectrum.open(data_file, orders=orders) for data_file in data]
 
 spectra = np.arange(len(DataSpectra)) #Number of different data sets we are fitting. Used for indexing purposes.
@@ -185,7 +186,7 @@ class OrderModel:
         self.id = key
         self.spectrum_id, self.order_id = self.id
 
-        print("Initializing model on order {}, DataSpectrum {}.".format(self.order_id, self.spectrum_id))
+        print("Initializing model on Spectrum {}, order {}.".format(self.spectrum_id, self.order_id))
 
         self.instrument = Instruments[self.spectrum_id]
         self.DataSpectrum = DataSpectra[self.spectrum_id]
@@ -256,7 +257,7 @@ class OrderModel:
 
         #Because this initialization is happening on the subprocess, I think the random state should be fine.
         self.sampler = NuisanceSampler(OrderModel=self, starting_param_dict=nuisance_starting, cov=nuisance_MH_cov,
-                                       debug=True, outdir=outdir)
+                                       debug=True, outdir=outdir, order=self.order)
         self.p0 = self.sampler.p0
 
         # Udpate the nuisance parameters to the starting values so that we at least have a self.data_mat
@@ -341,7 +342,7 @@ class OrderModel:
 
         # Initialize a new sampler, replacing the old one
         self.sampler = NuisanceSampler(OrderModel=self, starting_param_dict=starting_dict, cov=nuisance_MH_cov,
-                                       debug=True, outdir=outdir, prior_params=prior_params)
+                                       debug=True, outdir=outdir, prior_params=prior_params, order=self.order)
 
         self.p0 = self.sampler.p0
 
@@ -653,9 +654,9 @@ pconns = {} #Parent connections
 cconns = {} #Child connections
 ps = {}
 for spectrum in spectra:
-    for order in orders:
+    for order_id in order_ids:
         pconn, cconn = Pipe()
-        key = (spectrum, order)
+        key = (spectrum, order_id)
         pconns[key], cconns[key] = pconn, cconn
         p = Process(target=model.brain, args=(cconn,))
         p.start()
@@ -670,7 +671,7 @@ if args.perturb:
[…]

### 2014-10-30 test_profile.py
Removed a lot of code. Now for kurucz multi-order run

@@ -2,10 +2,9 @@ import cProfile
 import pstats
 
 #Your module here
-from scripts.stars.base_lnprob import main
-cProfile.run("main()", "prof")
+#from Starfish.parallel import profile_code
 
-
+#cProfile.run("profile_code()", "prof")
 
 def display_stats(pfile):
     p = pstats.Stats(pfile)

### 2014-10-30 Starfish/parallel.py
Removed a lot of code. Now for kurucz multi-order run

@@ -162,9 +162,7 @@ class OrderModel:
         self.lnprob_last = -np.inf
 
         self.func_dict = {"INIT": self.initialize,
-                          #"UPDATE": self.update_stellar,
                           "DECIDE": self.decide_stellar,
-                          #"IND": self.independent_sample,
                           "INST": self.instantiate,
                           "LNPROB": self.stellar_lnprob,
                           "GET_LNPROB": self.get_lnprob,
@@ -267,7 +265,7 @@ class OrderModel:
 
     def instantiate(self, *args):
         '''
-        Clear the old NuisanceSamplre, Instantiate the regions, and create a new NuisanceSampler.
+        Clear the old NuisanceSampler, Instantiate the regions, and create a new NuisanceSampler.
         '''
 
         # Using the stored residual spectra, find the largest outliers
@@ -365,7 +363,7 @@ class OrderModel:
         try:
             #lnp = None
             self.update_stellar(params)
-            lnp = self.evaluate() #Also sets self.lnprob to new value
+            lnp = self.evaluate() # Also sets self.lnprob to new value
             return lnp
         except C.ModelError:
             self.logger.debug("ModelError in stellar parameters, sending back -np.inf {}".format(params))
@@ -383,23 +381,23 @@ class OrderModel:
 
         R = self.fl - self.ChebyshevSpectrum.k * self.flux_mean - X.dot(self.mus)
 
-        sign, logdet = slogdet(CC)
-        if sign <= 0:
-            self.logger.debug("The determinant is negative {}".format(logdet))
-            #np.save("CC.npy", CC)
-            self.logger.debug("self.sampler.params are {}".format(self.sampler.params))
-            self.exceptions.append(self.sampler.params)
-
-            raise C.ModelError("The determinant is negative.")
+        # sign, logdet = slogdet(CC)
+        # if sign <= 0:
+        #     self.logger.debug("The determinant is negative {}".format(logdet))
+        #     #np.save("CC.npy", CC)
+        #     self.logger.debug("self.sampler.params are {}".format(self.sampler.params))
+        #     self.exceptions.append(self.sampler.params)
+        #
+        #     raise C.ModelError("The determinant is negative.")
 
         try:
             factor, flag = cho_factor(CC)
         except np.linalg.LinAlgError as e:
[…]

### 2014-10-30 attic/bokeh_plot_model.py

@@ -1,0 +1,170 @@
+#!/usr/bin/env python
+from StellarSpectra.model import Model
+from StellarSpectra.spectrum import DataSpectrum
+from StellarSpectra.grid_tools import TRES, HDF5Interface
+import StellarSpectra.constants as C
+import numpy as np
+import yaml
+import json
+
+import bokeh
+from bokeh.plotting import *
+from bokeh.objects import Range1d
+from bokeh.resources import Resources
+from bokeh.embed import autoload_static
+
+#Use argparse to determine if we've specified a config file
+import argparse
+parser = argparse.ArgumentParser(prog="bokeh_plot_model.py", description="Plot the model.")
+parser.add_argument("json", help="*.json file describing the model.")
+parser.add_argument("params", help="*.yaml file specifying run parameters.")
+parser.add_argument("-o", "--output", help="*.html file for output")
+args = parser.parse_args()
+
+
+if args.json: #
+    #assert that we actually specified a *.json file
+    if ".json" not in args.json:
+        import sys
+        sys.exit("Must specify a *.json file.")
+
+if args.params: #
+    #assert that we actually specified a *.yaml file
+    if ".yaml" in args.params:
+        yaml_file = args.params
+        f = open(args.params)
+        config = yaml.load(f)
+        f.close()
+
+    else:
+        import sys
+        sys.exit("Must specify a *.yaml file.")
+        yaml_file = args.params
+
+
+myDataSpectrum = DataSpectrum.open(config['data_dir'], orders=config['orders'])
+myInstrument = TRES()
+myHDF5Interface = HDF5Interface(config['HDF5_path'])
+
+stellar_Starting = config['stellar_params']
+stellar_tuple = C.dictkeys_to_tuple(stellar_Starting)
+
+cheb_Starting = config['cheb_params']
+cheb_tuple = ("logc0", "c1", "c2", "c3")
+
+cov_Starting = config['cov_params']
+cov_tuple = C.dictkeys_to_cov_global_tuple(cov_Starting)
+
+region_tuple = ("h", "loga", "mu", "sigma")
+region_MH_cov = np.array([0.05, 0.04, 0.02, 0.02])**2 * np.identity(len(region_tuple))
+
+
+myModel = Model.from_json(args.json, myDataSpectrum, myInstrument, myHDF5Interface)
+
[…]

### 2014-10-30 attic/cheb.py

@@ -1,0 +1,102 @@
+import matplotlib.pyplot as plt
+from numpy.polynomial import Chebyshev as Ch
+import numpy as np
+import model as m
+
+
+def test_chebyshev():
+    '''Domain controls the x-range, while window controls the y-range.'''
+    coef = np.array([0., 1.])
+    #coef2 = np.array([0.,0.,1,-1,0])
+    myCh = Ch(coef, window=[-10, 10])
+    #Domain c
+    #myCh2 = Ch(coef2)
+    #xs = np.linspace(0,3.)
+    x0 = np.linspace(-1, 1)
+    plt.plot(x0, myCh(x0))
+    #plt.plot(x0, myCh2(x0))
+    #plt.plot(xs, myCh2(xs))
+
+    plt.show()
+
+
+#test_chebyshev()
+
+
+xs = np.arange(2299)
+T0 = np.ones_like(xs)
+
+Ch1 = Ch([0,1], domain=[0,2298])
+T1 = Ch1(xs)
+
+Ch2 = Ch([0,0,1],domain=[0,2298])
+T2 = Ch2(xs)
+
+Ch3 = Ch([0,0,0,1],domain=[0,2298])
+T3 = Ch3(xs)
+
+T = np.array([T0,T1,T2,T3]) #multiply this by the flux and sigma vector for each order
+TT = np.einsum("in,jn->ijn",T,T)
+
+c = np.array([1,0,0,0])
+
+
+orders = [21,22]
+
+wls = m.wls
+fls = m.fls
+
+sigmas = m.sigmas #has shape (51, 2299), a sigma array for each order
+
+fmods = m.model(wls, 6001,3.5,0.0, 3, 0.0, 1e-10)
+#fls = m.model(wls,6020,3.6,40,2e-27)
+
+TT = np.einsum("in,jn->ijn",T,T)
+
+mu = np.array([1,0,0,0])
+sigmac = 0.2
+D = sigmac**(-2) * np.eye(4)
+Dmu = np.einsum("ij,j->j",D,mu)
+muDmu = np.einsum("j,j->",mu,Dmu)
+
+a= fmods**2/sigmas**2
+A = np.einsum("in,jkn->ijk",a,TT)
+#add in prior
+A = A + D
+detA = np.array(list(map(np.linalg.det, A)))
+invA = np.array(list(map(np.linalg.inv, A)))
+
[…]

### 2014-10-30 attic/correlated.py

@@ -1,0 +1,125 @@
+import scipy.sparse as sp
+import numpy as np
+from numpy.linalg import slogdet
+from scipy.sparse.linalg import spsolve
+import matplotlib.pyplot as plt
+
+def money(x, y):
+    pass
+
+def gauss(x, amp, mu, sigma):
+    return amp/np.sqrt(2 * np.pi * sigma**2) * np.exp(-0.5 * (x - mu)**2 / sigma**2)
+
+def line(x, b, m):
+    return b + m * x
+
+xs = np.linspace(-20,20)
+npoints = len(xs)
+
+#Create a continuum with a gaussian absorption line superimposed. Add Gaussian noise.
+ys = line(xs, 10, 0.05) - gauss(xs, amp=15, mu=0, sigma=1) + np.random.normal(size=npoints)
+
+def gauss_func(x0i, x1i, x0v=None, x1v=None, amp=None, mu=None, sigma=None):
+    x0 = x0v[x0i]
+    x1 = x1v[x1i]
+    return amp**2/(2 * np.pi * sigma**2) * np.exp(-((x0 - mu)**2 + (x1 - mu)**2)/(2 * sigma**2))
+
+def Cregion(xs, amp, mu, sigma, var=1):
+    '''Create a sparse covariance matrix using identity and block_diagonal'''
+    #In the region of the Gaussian, the matrix will be dense, so just create it as `fromfunction`
+    #and then later turn it into a sparse matrix with size xs x xs
+
+    #Given mu, and the extent of sigma, estimate the data points that are above, in Gaussian, and below
+    n_above = np.sum(xs < (mu - 4 * sigma))
+    n_below = np.sum(xs > (mu + 4 * sigma))
+
+    #Create dense matrix and indexes, then convert to lists so that you can pack things in as:
+
+    #csc_matrix((data, ij), [shape=(M, N)])
+    #where data and ij satisfy the relationship a[ij[0, k], ij[1, k]] = data[k]
+
+    len_x = len(xs)
+    ind_in = (xs >= (mu - 4 * sigma)) & (xs <= (mu + 4 * sigma)) #indices to grab the x values
+    len_in = np.sum(ind_in)
+    #print(n_above, n_below, len_in)
+    #that will be needed to evaluate the Gaussian
+
+
+    #Create Gaussian matrix fromfunction
+    x_gauss = xs[ind_in]
[…]

### 2014-10-30 attic/flot_model.py

@@ -1,0 +1,127 @@
+#!/usr/bin/env python
+
+#Use flot JavaScript plotting library to visualize a single order plot
+
+import numpy as np
+import json
+import jinja2
+
+def np_to_json(arr0, arr1):
+    '''
+    Take two numpy arrays, as in a plot, and return the JSON serialization for flot.
+    '''
+    data = np.array([arr0, arr1]).T #Transpose the arrays
+    listdata = data.tolist() #Convert numpy array to a list
+    return json.dumps(listdata) #Serialize to JSON
+
+def order_json(wl, fl, sigma, mask, flm, cheb):
+    '''
+    Given the quantities from a fit, create the JSON necessary for flot.
+    '''
+
+    residuals = fl - flm
+
+    # create the three lines necessary for the plot, in JSON
+    # data = [[wl0, fl0], [wl1, fl1], ...]
+    # model = [wl0, flm0], [wl1, flm1], ...]
+    # residuals = [[wl0, residuals0], [wl1, residuals1], ...]
+    plot_data = {"data":np_to_json(wl[mask], fl[mask]), "model":np_to_json(wl, flm), "residuals": np_to_json(wl[mask],
+                residuals[mask]), "sigma": np_to_json(wl[mask], sigma[mask]), "cheb": np_to_json(wl, cheb) }
+    return plot_data
+
+
+def render_template(base, plot_data):
+
+    templateLoader = jinja2.FileSystemLoader(searchpath=base + "templates")
+    # An environment provides the data necessary to read and
+    # parse our templates.  We pass in the loader object here.
+    templateEnv = jinja2.Environment(loader=templateLoader)
+    template = templateEnv.get_template('flot_plot.jinja')
+
+    templateVars = {"base": base}
+    templateVars.update(plot_data)
+
+    #Render plot using plot_data
+    outputText = template.render(templateVars)
+    f = open('index_flot.html', 'w')
+    f.write(outputText)
+    f.close()
+
+
+def main():
+    #Use argparse to determine if we've specified a config file
+    import argparse
+    parser = argparse.ArgumentParser(prog="flot_model.py", description="Plot the model and residuals using flot.")
+    parser.add_argument("json", help="*.json file describing the model.")
+    parser.add_argument("params", help="*.yaml file specifying run parameters.")
+    # parser.add_argument("-o", "--output", help="*.html file for output")
+    args = parser.parse_args()
+
+    import json
+    import yaml
+
+    if args.json: #
+        #assert that we actually specified a *.json file
[…]

### 2014-10-30 attic/kernel.py

@@ -1,0 +1,128 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.ndimage.filters import convolve
+from matplotlib.ticker import FormatStrFormatter as FSF
+#from scipy.optimize import fsolve
+
+c = 2.99792458e18 #A s^-1
+
+
+def karray(center, width, res):
+    '''Creates a kernel array with an odd number of elements, the central element centered at `center` and spanning
+    out to +/- width in steps of resolution. Works similar to arange in that it may or may not get all the way to the
+    edge.'''
+    neg = np.arange(center - res, center - width, -res)[::-1]
+    pos = np.arange(center, center + width, res)
+    kar = np.concatenate([neg, pos])
+    return kar
+
+#Convolution
+@np.vectorize
+def gauss_kernel(dlam, V=6.8, lam0=6500.):
+    '''V is the FWHM in km/s. lam0 is the central wavelength in A'''
+    sigma = V / 2.355 * 1e13 #A/s
+    return c_ang / lam0 * 1 / (sigma * np.sqrt(2 * np.pi)) * np.exp(- (c_ang * dlam / lam0) ** 2 / (2. * sigma ** 2))
+
+
+def plot_gauss():
+    lams = np.linspace(-1, 1, num=200)
+    plt.plot(lams, gauss_kernel(lams))
+    plt.show()
+
+
+def gauss_series(dlam, V=6.5, lam0=6500.):
+    '''sampled from +/- 3sigma at dlam. V is the FWHM in km/s'''
+    sigma_l = V / (2.355 * 3e5) * 6500. #A
+    wl = karray(0., 4 * sigma_l, dlam)
+    gk = gauss_kernel(wl)
+    return gk / np.sum(gk)
+
+
+def plot_gauss_kernel():
+    sigma_l = 6.5 / (2.355 * 3e5) * 6500.
+    wl = karray(0, 4 * sigma_l, 0.01)
+    ys = gauss_series(0.01)
+    plt.plot(wl, ys, "o")
+    plt.show()
+
+#plot_gauss_kernel()
+
+@np.vectorize
+def vsini_kernel(v, vsini, epsilon=0.6):
[…]

### 2014-10-30 attic/plotting.py

@@ -1,0 +1,476 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import FormatStrFormatter as FSF
+from matplotlib.ticker import MultipleLocator
+from astropy.io import ascii
+from scipy.optimize import fmin
+from StellarSpectra.spectrum import DataSpectrum
+
+c_ang = 2.99792458e18 #A s^-1
+
+#k.xaxis.set_major_formatter(FSF("%.0f"))
+#k.locator_params(axis='x', nbins=5)
+
+
+def compare_kurucz():
+    wg_full = np.load('wave_grids/PHOENIX_2kms_air.npy')
+    wg = np.load("wave_grids/kurucz_2kms_air.npy")
+
+    ind = (wg_full >= wg[0]) & (wg_full <= wg[-1])
+
+    kurucz = m.load_hdf5_spectrum(6000, 4.0, 0.0, "kurucz", "LIB_kurucz_2kms_air.hdf5")
+    BTSettl = m.load_hdf5_spectrum(6000, 4.0, 0.0, "BTSettl", "LIB_BTSettl_2kms_air.hdf5")[ind]
+    PHOENIX = m.load_hdf5_spectrum(6000, 4.0, 0.0, "PHOENIX", "LIB_PHOENIX_2kms_air.hdf5")[ind]
+
+    #normalize avg value to 1
+    kurucz *= c_ang/wg**2
+    kurucz = kurucz/np.average(kurucz)
+    BTSettl = BTSettl/np.average(BTSettl)
+    PHOENIX = PHOENIX/np.average(PHOENIX)
+
+    fig, ax = plt.subplots(nrows=2, sharex=True, figsize=(9,6))
+
+    ax[0].plot(wg, kurucz, label="Kurucz")
+    ax[0].plot(wg, PHOENIX, label="Husser")
+    ax[0].legend()
+
+    residuals = (kurucz - PHOENIX)/kurucz
+    ax[1].plot(wg, residuals, label="(Kurucz-Husser)/Kurucz")
+    ax[1].legend()
+    line_list = pltMC.return_lines(wg, residuals, sigma=0.3, tol=0.2)
+
+    offsets = np.linspace(-0.4, 0.4, num=10)
+    off_counter = 0
+    for line, label in line_list:
+        ax[1].axvline(line, color='k', lw=0.1)
[…]

### 2014-10-30 attic/synthphot.py

@@ -1,0 +1,84 @@
+#!/usr/bin/env python2
+
+import pysynphot as S
+import pyfits as pf
+from collections import OrderedDict
+import numpy as np
+import matplotlib.pyplot as plt
+from PHOENIX_tools import load_flux_full, w_full
+from deredden import deredden
+
+c_ang = 2.99792458e18 #A s^-1
+
+#data for GWOri
+#wl [microns], f_nu [Jy], sys_err (as fraction of f_nu)
+data = np.array([[0.36, 6.482e-02, 0.12],
+                 [0.44, 2.451e-01, 0.30],
+                 [0.55, 4.103e-01, 0.08],
+                 [0.64, 7.844e-01, 0.11],
+                 [0.79, 9.174e-01, 0.07]])
+
+wl = data[:, 0] * 1e4 #angstroms
+f_nu = data[:, 1] * 1e-23 #ergs/cm^2/s/Hz
+f_nu_err = f_nu * data[:, 2] #ergs/cm^2/s/Hz
+
+#Convert to f_lambda
+f_lam = f_nu * c_ang / wl ** 2
+f_lam_err = f_nu_err * c_ang / wl ** 2
+
+filters = ["U", "B", "V", "R", "I"]
+
+ind = (w_full > 2000) & (w_full < 40000)
+ww = w_full[ind]
+
+ff = load_flux_full(5900, 3.5)[ind] * 2e-28
+#redden spectrum
+#red = ff/deredden(ww,1.5,mags=False)
+
+sp = S.ArraySpectrum(wave=ww, flux=ff, waveunits='angstrom', fluxunits='flam', name='T5900K')
+
+#sdss_u=S.FileBandpass("/home/ian/.builds/stsci_python-2.12/pysynphot/cdbs/comp/nonhst/sdss_u_005_syn.fits")
+#sdss_g=S.FileBandpass("/home/ian/.builds/stsci_python-2.12/pysynphot/cdbs/comp/nonhst/sdss_g_005_syn.fits")
+#sdss_r=S.FileBandpass("/home/ian/.builds/stsci_python-2.12/pysynphot/cdbs/comp/nonhst/sdss_r_005_syn.fits")
+#sdss_i=S.FileBandpass("/home/ian/.builds/stsci_python-2.12/pysynphot/cdbs/comp/nonhst/sdss_i_005_syn.fits")
[…]

### 2014-10-31 tex/paper.tex
new hdf5 format, rerunning gl, wasp

@@
 \acknowledgments  The authors would like to graciously acknowledge the following people for many helpful discussions and key insights: Gregory M.
-Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Allyson Bieryla, and Maxwell Moe.
+Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Rebecca Dawson, Tom Loredo and the ExoStat group, Allyson Bieryla, and Maxwell Moe.
 IC was supported by the NSF Graduate Fellowship and the Smithsonian Institution.
 Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-Astropy.
-JuliaLang.
-Python/Scikit-learn.\\
+This research made use of Astropy \citep{astropy13} and the Julia programming language \citep{julia12}.
 
@@
 Spectral line formation is a non-linear process, where line shapes are determined by the complex 
-radiative transfer through a stellar atmosphere.  We aim to bypass this non-linear 
-process and predict (through interpolation) the behavior of spectral lines based on pre-computed 
-library spectra with similar physical parameters.  The goal in this Appendix is to describe a 
+radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
+process and predict the behavior of spectral lines based on pre-computed 
+library spectra with similar physical parameters.  This appendix describes the 
 procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
-prediction through the probabilistic framework we have developed.
+prediction--which can be substantial--through the probabilistic framework we have developed.
 
 There are two important factors to consider when interpolating in libraries of model stellar 
-spectra.  First, the flux values in some non-negligible fraction of the spectra do not vary 
+spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
 monotonically with changes in the physical parameters that define the library: this helps inform 
@@
 are usually not Nyquist sampled with respect to the physical parameters that define the library.  
-This unfortunately (but not surprisingly) means that the library is not sampled densely enough in 
+This unfortunate fact means that the library is not sampled densely enough in 
 parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
 information loss.
-Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra, seeks to non-deterministically predict the behavior of a spectrum for stellar parameters not sampled by the library.
+Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra and a chosen interpolation point, delivers a probability distribution over the range of possible interpolated spectra.
 
@@
 
-The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation.
-We use principal component analysis (PCA) to decompose the synthetic library into a set of ``eigenspectra'' $\varphi$.
-Following techniques of SLOAN (cite Ivecik 2014).
-After examining how well the spectra reconstruct the grid, we keep the first $m$ of the eigenspectra, and discard the latter eigenspectra.
-We keep enough eigenspectra to reconstruct any pixel flux to better than $2\%$, however most pixels are more like $XX\%$.
+The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}.
+We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}.
+We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels, however the mean pixel error is generally much smaller ($\lesssim 0.5\%$).
 An $N_{\rm pix}$ eigenspectrum $\xi_k$, $k = \{1, 2, \ldots, m\}$.
@@
 First, we select only a subset of the spectra synthetic library which will be relevant to our  The mean spectrum is
+
+\begin{figure*}[!t]
+  \begin{center}
+  \includegraphics{figs/pca_reconstruct.pdf}
+  \figcaption{A bunch of reconstrcuted.
+\label{fig:pca_reconstruct}}
+  \end{center}
+\end{figure*}
+
 \begin{equation}

### 2014-10-31 hdfregion.py

@@ -8,11 +8,14 @@ Works by matching regions within some degree of variance.
 import os
 import numpy as np
 from collections import deque
+from astropy.table import Table
+from astropy.io import ascii
+import sys
 
 import argparse
 parser = argparse.ArgumentParser(description="Measure statistics across multiple chains.")
-parser.add_argument("--dir", action="store_true", help="Concatenate all of the flatchains stored within run* "
-                                                       "folders in the current directory. Designed to collate runs from a JobArray.")
+parser.add_argument("--glob", help="Do something on this glob. Must be given as a quoted expression to avoid shell "
+                                   "expansion.")
 parser.add_argument("-o", "--outdir", default="mcmcplot", help="Output directory to contain all plots.")
 parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
 parser.add_argument("--clobber", action="store_true", help="Overwrite existing files?")
@@ -37,58 +40,84 @@ else:
 
 args.outdir += "/"
 
-if args.dir:
-    #assemble all of the flatchains.hdf5 files from the run* subdirectories.
-    import glob
-    folders = glob.glob("run*")
-    files = [folder + "/flatchains.hdf5" for folder in folders]
+if args.glob:
+    from glob import glob
+    files = glob(args.glob)
 elif args.files:
-    assert len(args.files) >= 2, "Must provide 2 or more HDF5 files to combine."
     files = args.files
 else:
     import sys
-    sys.exit("Must specify either --dir or --files")
+    sys.exit("Must specify either --glob or --files")
 
-import h5py
+#At this point, files is a list of flatchains.hdf5 files, which contain both cheb, cov, and region parameters.
+
+from hdfutils import Flatchain
+
 #Because we are impatient and want to compute statistics before all the jobs are finished, there may be some
 # directories that do not have a flatchains.hdf5 file
-hdf5list = []
-filelist = []
+flatchainList = []
 for file in files:
     try:
-        hdf5list += [h5py.File(file, "r")["0"]]
-        filelist += [file]
-    except OSError:
-        print("{} does not exist, skipping.".format(file))
+        flatchainList.append(Flatchain.open(file, type="region"))
+    except OSError as e:
+        print("{} does not exist, skipping. Or error {}".format(file, e))
 
-#I think we should separate this by order
[…]

### 2014-10-31 hdfutils.py

@@ -1,8 +1,10 @@
 #!/usr/bin/env python
 import os
 import numpy as np
-import matplotlib
-#matplotlib.rc("axes", labelsize="large")
+import Starfish.constants as C
+from astropy.table import Table
+from astropy.io import ascii
+
 import triangle
 
 #Base package for all HDF tools
@@ -15,13 +17,13 @@ import argparse
 parser = argparse.ArgumentParser(description="Measure statistics across multiple chains.")
 parser.add_argument("--glob", help="Do something on this glob. Must be given as a quoted expression to avoid shell "
                                   "expansion.")
+parser.add_argument("--ftype", default="stellar", help="What type of flatchains do we have?")
 #parser.add_argument("--dir", action="store_true", help="Concatenate all of the flatchains stored within run* "
 #                                                       "folders in the current directory. Designed to collate runs
 # from a JobArray.")
 parser.add_argument("--outdir", default="mcmcplot", help="Output directory to contain all plots.")
 parser.add_argument("--output", default="combined.hdf5", help="Output HDF5 file.")
 parser.add_argument("--clobber", action="store_true", help="Overwrite existing files?")
-parser.add_argument("--old", action="store_true", help="Old format flatchains.hdf5 files?")
 
 parser.add_argument("--lnprob", help="The HDF5 file containing the lnprobability chains.")
 parser.add_argument("--files", nargs="+", help="The HDF5 files containing the MCMC samples, separated by whitespace.")
@@ -45,6 +47,7 @@ parser.add_argument("--acor", action="store_true", help="Calculate the autocorre
 parser.add_argument("--acor-window", type=int, default=50, help="window to compute acor with")
 
 parser.add_argument("--cov", action="store_true", help="Estimate the covariance between two parameters.")
+parser.add_argument("--ndim", type=int, help="How many dimensions to use for estimating the 'optimal jump'.")
 parser.add_argument("--paper", action="store_true", help="Change the figure plotting options appropriate for the "
                                                          "paper.")
 
@@ -59,7 +62,6 @@ else:
     os.makedirs(args.outdir)
 
 args.outdir += "/"
-
 
 import h5py
 
@@ -76,18 +78,45 @@ label_dict = {"temp":r"$T_{\rm eff}\;[{\rm K}]$", "logg":r"$\log g$", "Z":r"$[{\
               "sigAmp":r"$b$", "logAmp":r"$\log_{10} a_{\rm g}$", "l":r"$l$",
[…]

### 2014-11-01 tex/paper.tex
Sys import error on hdf

@@
 (Sect.~\ref{sec:examples}).  The source code developed here is open and freely available for use: 
-see \url{http://github.com/iancze/StellarSpectra}.
+see \url{http://github.com/iancze/Starfish}.
 
@@
 
-The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}.
-We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}.
-We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels, however the mean pixel error is generally much smaller ($\lesssim 0.5\%$).
-An $N_{\rm pix}$ eigenspectrum $\xi_k$, $k = \{1, 2, \ldots, m\}$.
-Before we do the PCA operation, however, we first standardize the synthetic library by subtracting the mean spectrum and ``whitening'' the spectra by dividing by the standard deviation in each pixel.
-First, we select only a subset of the spectra synthetic library which will be relevant to our  The mean spectrum is
-
 \begin{figure*}[!t]
@@
   \includegraphics{figs/pca_reconstruct.pdf}
-  \figcaption{A bunch of reconstrcuted.
+  \figcaption{\emph{top} : The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51.
+\emph{bottom}: The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
 \label{fig:pca_reconstruct}}
@@
 
+The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}.
+We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}.
+Prior to the PCA decomposition, we first identify a subset of $M$ synthetic spectra which will be relevant to the star at hand.
+We then standardize the synthetic library by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid.
+The mean spectrum is
 \begin{equation}
@@
 \end{equation}
-We standardize all spectra in the library by subtracting the mean spectrum and dividing by the standard deviation spectrum.
+All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
 \begin{equation}
@@
 \end{equation}
-Then, the eigenspectra are determined from the standardize grid using PCA analysis (see Ivecik, scikit-learn). To speed computation, we first select $M$ subset spectra from the grid.
-
-[For an example of the eigenspectra for Gl51, see figure A1 (left) and reconstruction of a synthetic spectrum at stellar parameters XX (right).]
-
-Using the PCA basis, we can lossily reconstruct any spectrum from the library by a linear combination of the eigenspectra
+Then, the eigenspectra are computed from the standardize grid using PCA analysis (\cite{ivezic13}).
+We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$).
+We denote a single $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$.
+As an example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.
+
+Using the PCA basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
 \begin{equation}
@@
 where $w_k$ is the weight of a given eigenspectrum, $k$.
-This weight is a three dimensional scalar function, which depends on the stellar parameters $\vt_\ast$.
+This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$.
 We can determine the weights at the grid points in the library by taking the dot product of the standardized synthetic spectrum with each of the eigenspectra.
-For a given eigenspectrum $\xi_k$, the weight is
+The weight for a given eigenspectrum $\xi_k$ is
 \begin{equation}
@@
 \end{equation}
-These weights are generally a smooth function of stellar parameters. To simplify the notation, we can write the weights in a length-$m$ column vector
+which is generally a smooth function of stellar parameters. To simplify the notation, we can write the weights in a length-$m$ column vector
 \begin{equation}
@@
 \end{equation}
-Therefore we can rewrite Equation~\ref{eqn:reconstruct} as
+Then we can rewrite Equation~\ref{eqn:reconstruct} as
 \begin{equation}
@@
 Where $I_{N_{\rm pix}}$ is an $N_{\rm pix}$ by $N_{\rm pix}$ identity matrix.
-To summarize, thus far we have decomposed the spectra in the synthetic library into a principal component basis, which allows us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
-The weight corresponding to each eigenspectrum is a relatively-smooth three-dimensional scalar function of the stellar parameters.
-Therefore, to interpolate a spectrum from the synthetic library for a value of stellar parameters which is not in the spectral library, we must interpolate the value of the weights to an off-grid value.
+To recapitulate thus far, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
+The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters.
+Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value.
 In practice, it might be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly.
 However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum could constitute a significant portion of the uncertainty budget.
+This, combined with the Nyquist undersampling of the synthetic grid caused behavior where stellar parameters corresponding to a grid point in the synthetic library might be preferentially favored and bias the results, even after empirically accounting for interpolation error by doing ``drop-out'' interpolation tests.
+[This is more of a reason to use PCA vs.
+regular spectra, right?]
+
 Therefore, we decide to develop a Bayesian spectral emulator (cite Habib, Heitmann) which allows us to propagate the uncertainty in interpolation forward to the likelihood calculation.

### 2014-11-01 tex/paper.tex
more GP text

@@
 $\vP = [\cheb, \cov]$.  To explore the posterior distribution of the model conditioned on the data
-\begin{equation} \label{eqn:post}
+\begin{equation} 
 p(\vT, \vP | \vD) \propto p(\vD | \vT, \vP) \, p(\vT, \vP)
+\label{eqn:post}
 \end{equation}
@@
 
-Model library spectra, $f_{\lambda}(\{\vt_{\ast}\})$, are stored as (1-dimensional) arrays of flux 
+Model library spectra, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, are stored as (1-dimensional) arrays of flux 
 values, sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).  In the case 
-of interest here, the collections of model parameters $\{\vt_{\ast}\} = [\{T_{\rm eff}, \log g,  \Z 
+of interest here, the collections of model parameters $\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z 
 \}]$ define the dimensions of the library grid; the full library is therefore encapsulated in a 
@@
   \includegraphics{figs/pca_reconstruct.pdf}
-  \figcaption{\emph{top} : The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51.
-\emph{bottom}: The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
+  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51.
+(\emph{bottom}) The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
 \label{fig:pca_reconstruct}}
@@
 \begin{equation}
-  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid})
+  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
 \end{equation}
@@
 \begin{equation}
-  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu \bigr]^2 } 
-\end{equation}
+  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
+\end{equation}
+$\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of stellar parameters sampled in the spectral library, while $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection.
 All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
@@
 \begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}) \xi_k
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
   \label{eqn:reconstruct}
@@
 \begin{equation}
-  w_k(\{\vt_\ast \}^\textrm{grid}) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) \xi_k
-\end{equation}
-which is generally a smooth function of stellar parameters. To simplify the notation, we can write the weights in a length-$m$ column vector
-\begin{equation}
-  \mathbf{w} = 
+  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
+\end{equation}
+which is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel). To simplify the notation, we can write the weights in a length-$m$ column vector
+\begin{equation}
+  \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i) = 
   \begin{bmatrix}
-    w_1\\
-    w_2\\
+    w_1(\{\vt_\ast \}^\textrm{grid}_i)\\
+    w_2(\{\vt_\ast \}^\textrm{grid}_i)\\
     \vdots\\
-    w_m
+    w_m(\{\vt_\ast \}^\textrm{grid}_i)
   \end{bmatrix}
@@
 \begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
+  \label{eqn:reconstruct2}
 \end{equation}
@@
 This, combined with the Nyquist undersampling of the synthetic grid caused behavior where stellar parameters corresponding to a grid point in the synthetic library might be preferentially favored and bias the results, even after empirically accounting for interpolation error by doing ``drop-out'' interpolation tests.
-[This is more of a reason to use PCA vs.
-regular spectra, right?]
-
-Therefore, we decide to develop a Bayesian spectral emulator (cite Habib, Heitmann) which allows us to propagate the uncertainty in interpolation forward to the likelihood calculation.
-First, we describe how we use a Gaussian process to model the interpolation of the weights then we marginalize over all possible interpolations to arrive at a modified likelihood function.
-
-In Section XX, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral regions of the residuals due to pathological differences in the data spectrum and synthetic spectrum.
-Now, we use a separate Gaussian process in a slightly-different manner, in order to serve as an interpolating function for the weights.
-This is using a Gaussian process for prediction, given the values of the weights at several known locations (the values of stellar parameters at the grid points) and some assumed properties of the unknown functional form of the weights, what are the most probable values of the weights at stellar parameters in between the grid points?
+
+Now, we describe how we design the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters.
+In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum.
+Alternatively, we now use a different Gaussian process in a separate application to serve as an interpolating function for the weights.
+
+We model each eigenspectrum weight independently; therefore there is a separate Gaussian process ($\mathcal{GP}$) for each eigenspectrum.
+For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$.
+The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
+\begin{equation}
+  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( 0, \mathcal{K}_\textrm{SE} (\{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})\right )
+\end{equation}
+The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$.
+Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
+\begin{widetext}
+\begin{equation}
+  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
+  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2 + (\log g_i - \log g_j)^2 + (\Z_i - \Z_j)^2 }{2 l_\textrm{SE}^2} \right )
+\end{equation}
+\end{widetext}
+This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
+Then the $M$ by $M$ covariance matrix is
+\begin{equation}
+\mathsf{C}^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | a_\textrm{SE}, l_\textrm{SE})
+\end{equation}
+which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters.
+
+The optimal parameters for the hyperparameters of the squared exponential are given by the multi-variate Gaussian probability distribution, which is called the marginal likelihood function or ``evidence'' for the weights \citep{rasmussen05},
+\begin{equation}
+  p(\mathbf{w}_k^\textrm{grid} | C^\textrm{grid}) = \mathcal{N}\left (\mathbf{w}_k^\textrm{grid} \left | \; \mathbf{0}, \mathsf{C}^\textrm{grid} \right . \right )
+\end{equation}
+Because $\mathsf{C}^\textrm{grid}$ is defined by $\mathcal{K}_\textrm{SE}$, this likelihood function is equivalent to $p(\mathbf{w}_k^\textrm{grid} |\; a_\textrm{SE}, l_\textrm{SE})$.
+By multiplying this with a suitable prior over $a_\textrm{SE}$ and $l_\textrm{SE}$, we arrive at a posterior distribution for the hyperparameters of the Gaussian process.
+For computational efficiency, we optimize the hyperparameters by finding the peak of the posterior distribution with a downhill gradient algorithm and then consider the hyperparameters fixed.
+
+Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
+\begin{equation}
+  \begin{bmatrix}
+    \mathbf{w}_k^\textrm{grid}\\
+    w_k
+  \end{bmatrix} \sim
+  \mathcal{N} \left (
+  \mathbf{0},
+  \begin{bmatrix}
+    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
+\mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
+  \end{bmatrix}
+  \right )
+\end{equation}
+where the dependence of the covariance kernel on the optimized hyperparameters is implied.
+Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$.
+And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$.
+The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$).
+The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
+\begin{multline}
+  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) = \\
+  \mathcal{N} \Bigl ( w_k \Bigl |\; \Bigr . \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}, \Bigr . \\
+ \Bigl .
+\mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \Bigr )
+ \label{eqn:conditional_gp}
+\end{multline}
+Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
+
+\begin{figure*}[!t]
+\begin{center}
+  \includegraphics{figs/GP_left.pdf}
+  \includegraphics{figs/GP_right.pdf}
+  \figcaption{ (\emph{left}) the blue dots mark the weights $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ for $\vt_\ast$ for various values of $T_\textrm{eff}$ with $\log g = 5.0$ and $\Z = 0.0$.
+The thin blue lines show 50 random draws of possible functional forms described by the Gaussian process, constrained to pass through the grid weights.
+(\emph{inset}) a zoomed portion showing the scatter in the possible functional forms.
+The black vertical line represents a slice through the scatter of the predicted weight value at $\vt_\ast = [T_\textrm{eff} = 3150\,\textrm{K}, \log g = 5.0\,\textrm{dex}, \Z = 0.0\, \textrm{dex}]$.
+(\emph{right}) The posterior probability of $w_k$ at this value of $\vt_\ast$ is completely described by Eqn~\ref{eqn:conditional_gp}, allowing us to analytically marginalize over all probable values of the weights, and thus marginalize over all probable spectral interpolations.
+\label{fig:GP_interp}}
+\end{center}
+\end{figure*}
+
+If we desired an actual value of the interpolated weight, for example to reconstruct a model spectrum, we could simply draw a Gaussian random variable from the probability distribution in Eqn~\ref{eqn:conditional_gp}.
+However, because we now know the probability distribution of the weight as a function of $\vt_\ast$, we can rewrite our data likelihood function (Eqn~\ref{eqn:lnlikelihood}) in such a way that it is possible to analytically marginalize over all possible values of $w_k$, and thus all probable spectral interpolations.
+
+EDIT As an aside, this process of prediction is completely analogous to the process by which we generated the residual draws in the the covariance diagram.
+These were draws from the prior distribution (specified by Cdata).
 This is similar to drawing random values from the covariance matrix in Section XX given covariance matrices filled in by covariance kernels.
@@
 Only now, we are drawing from a giant Gaussian conditioned on several points, which are the weights at the grid point locations.
-
-[Fig (three panels?) showing the values of the weight at different 1-d locations, and then many possible interpolations of that weight.
-Also one showing the probability distribution of the weight at a fixed stellar parameters (Gaussian in shape) and mentioning that we can marginalize over.]
-
-The smooth the distribution of weights (see Fig A.2, left panel) with changing stellar parameters motivated the use of a squared-exponential covariance kernel for the Gaussian process (cite Rasmussen + Williams for it's smooth properties).
-This is in contrast to the \matern\ kernel used in Section XX, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
-
-Consider the modelling of a single weight. We model this using a squared exponential covariance kernel
-
-Eqn of Squared exponential.
-
-We create a different covariance matrix $C_\textrm{w}$ which models the covariance between the weights of all the $M$ points in the grid. Here show V11, V12, V22, etc.
-
-Given the values of the weights at the grid point locations, we can define a posterior probability distribution for the parameters of the Gaussian process.
-Following Habib, Heitmann, we adopt priors on the parameters of the GP in order to regularize it [TODO: better exp].
-
-Eqn of full posterior distribution for a given weight.
-
-We use a downhill simplex algorithm (Scipy/fmin) to find the maximum of the posterior with respect to the GP parameters.
-From here on, we fix the GP hyperparamaters to their optimum values and consider the GP specified.
 The benefit of using a Gaussian process for prediction/regression is that we now specified a probability distribution over possible interpolated values of the weights.
@@
 
-Initially, we presented the reconstruction as a linear reconstruction of the eigenspectra resulting from the decomposition of the synthetic grid.
-In practice, however, we must still post-process the synthetic spectra to match the data spectrum (Section XX), applying convolutions for instrumental broadening, rotational broadening, and Doppler shift, and then lastly re-sampling the spectra to the exact pixels of the dataspectrum.
-Fortunately, because the convolution operation is a linear operator, we can first apply these operations to the eigenspectra themselves, without loss of information.
-Unfortunately, the Doppler shift and resampling operation is a non-linear operation and there will be some loss of information.
-However, we find that in practice when using a reasonable oversampling ratio of the sythetic to the dataspectra, applying the Doppler shift/resampling operation to the eigenspectra results in smaller than a 0.2\% error across all pixels, and thus the effect is negligable.
-
-Therefore, we can rewrite the reconstruction of the model spectrum as a linear combination of eigenspectra.
-
-[Equation rewriting the blended/resampled eigenspectra as a function of thetaOBS]
-[Equation rewriting reconstruction but using the operators on the eigenspectra]
-
-Since the Gaussian process delivers a probability distribution (Gaussian in form) for each weight, given a value of stellar parameters, we now have a probability distribution of a given reconstruction of a spectrum, which is just a linear combination of the eigenspectra.
-Because both the data likelihood and the probability distribution for the weights are multidimensional Gaussians, we can write down the joint probability distribution for the data and the interpolations, and then \emph{analytically} marginalize over the weights, thus marginalizing over the space of all possible interpolations given the principal component basis.
-
-Cite Mandel, Bishop, Gelman.
+So far, we have presented the reconstruction of a spectrum as a linear combination of the eigenspectra of the synthetic library (Figure~\ref{fig:pca_reconstruct}).
+In practice, however, the reconstructed spectrum must be further post-processed to match the observational parameters ($\vt_{\rm obs}$) of the data spectrum by applying convolutions for instrumental broadening, stellar rotation broadening, and Doppler shift, and then lastly re-sampling the spectrum to the exact pixels of the CCD (Section~\ref{subsec:postprocess}).
+Fortunately, because convolution is a linear operation, we can first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the reconstructed spectrum as a linear combination of these modified eigenspectra without loss of information.
+Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them as linear.
+However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the error across all pixels is smaller than 0.2\%, and thus any effect is negligible.
+We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra.
+This is the model.
+Then, we can rewrite Eqn~\ref{eqn:reconstruct2} as
+\begin{equation}
+  f_\lambda(\vt_\textrm{obs}) \approx \widetilde{\xi}_\mu + \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}} \mathbf{w}
+\end{equation}
+accompanied by the conditional probability on $\mathbf{w}$ (Eqn~\ref{eqn:conditional_gp}).
+
+The final task of designing the spectral emulator is to combine the posterior distribution of the data (Eqn~\ref{eqn:post}) with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:conditional_gp}) to form the joint distribution and then marginalize out the weights, such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.
+
+Cite Bishop, Gelman.
 

### 2014-11-01 hdfcov.py

@@ -3,5 +3,4 @@
 #Defines a bunch of global variables and also parses args.
 import hdfutils
 
-if hdfutils.args.cov:
-    hdfutils.estimate_covariance(hdfutils.flatchainList[0])
+hdfutils.estimate_covariance(hdfutils.flatchainList[0])

### 2014-11-02 tex/paper.tex
rough draft of appendix complete

@@
 
-The procedure outlined above can be a significant computational challenge.  A typical data spectrum 
-has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the many repeated evaluations of the matrix 
-product $\vR^{\trans} \vC^{-1} \vR$ in the likelihood calculation can be numerically expensive; 
-prohibitively so in the brute force sense, since $\vC$ is non-trivial by design.  However, there 
-are two aspects of these calculations that enable us to take advantage of some clever algorithms 
-designed for such issues.  First, since we are only interested in the product $\vR^{\trans} 
-\vC^{-1} \vR$ and not $\vC^{-1}$ itself, we can employ some efficient sparse matrix approximations 
-that avoid the computational expense of matrix inversion.  And second, because $\vC$ is a 
-symmetric, positive semi-definite matrix, we can employ Cholesky factorization to optimize the 
-evaluation of the matrix product.  In practice, this means Steps~(2), (3a), and (3b) in the 
-procedure outlined above can be conducted with minimal computation cost.  Although the Cholesky 
-factorization must be re-done for each update of $\vp_{\mathsf C}$ (Step~3d), this does not incur a 
-high cost because the covariance kernels were designed to deliver a sparse matrix (via the Hann 
-windows).  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization 
-operations \citep{chen08,davis09}, and extend the Metropolis-Hastings sampler included in the {\tt 
-emcee} package \citep{foreman-mackey13} to function within a blocked Gibbs 
-sampler.\footnote{\url{https://github.com/iancze/emcee}} For echelle spectra with multiple orders or multiple spectra of the same object, taken with different instruments, the nuisance parameters for each chunk of spectrum are independent.
-Additionally, the computationally intensive steps of generating a model spectrum for a specific wavelength range and evaluating the likelihood on this spectral chunk can also be parallelized.
-Therefore, the only part of the code that needs to be synchronized is the MCMC proposal of stellar parameters.
-Because all of the computationally expensive steps can be parallelize, this algorithim has the ability to be massively parallelized on a cluster, enabling simultaneous inference for wide spectral ranges or multiple spectral datasets acquired with different instruments.\\
+The procedure outlined above can be a significant computational challenge.  A
+typical data spectrum has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the
+many repeated evaluations of the matrix product $\vR^{\trans} \vC^{-1} \vR$ in
+the likelihood calculation can be numerically expensive.  However, because
+$\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky
+factorization to optimize the evaluation of the matrix product ($\vR^{\trans}
+\vC^{-1} \vR$) and avoid the direct calculation of the matrix inverse
+($\vC^{-1}$) itself.  For echelle spectra with multiple orders or multiple
+spectra of the same object, taken with different instruments, the nuisance
+parameters for each chunk of spectrum are independent from the other chunks. This
+means that the computationally intensive steps of generating a model spectrum
+for a specific wavelength range and evaluating the likelihood can be
+parallelized. The only segment of the code that needs to be synchronized is the
+MCMC proposal of stellar parameters, which are shared between all chunks of the
+spectrum. Because all of the computationally expensive steps can be
+parallelized, this algorithm can be massively parallelized on a cluster,
+enabling simultaneous inference for wide spectral ranges at high resolution or
+multiple spectral datasets acquired with different instruments. To sample this
+model, we extend the Metropolis-Hastings sampler included in the {\tt emcee}
+package \citep{foreman-mackey13} to function within a parallelized blocked
+Gibbs sampler.\footnote{\url{https://github.com/iancze/emcee}}\\
 
@@
 promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
-like those being collected in programs like the APOGEE and HERMES surveys), since they provide 
-enormous leverage for identifying and improving the underlying model systematics. \\
-%Should make note of surveys like the ESA/Gaia program, SEGUE, RAVE, too.  Probably need to 
-%include citations to each.  Also, I think HERMES is really GALAH, right?  If not, it should be 
-%linked to that, since there is a popular Herschel HERMES program that will confuse people...
+like those being collected in programs like the APOGEE and HERMES surveys \citep{nidever12, zucker12}), since they provide enormous leverage for identifying and improving the underlying model systematics.
+\\
 
@@
 \end{equation}
-Then, the eigenspectra are computed from the standardize grid using PCA analysis (\cite{ivezic13}).
+Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}).
 We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$).
@@
 Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
-\begin{widetext}
-\begin{equation}
+\begin{multline}
   \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
-  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2 + (\log g_i - \log g_j)^2 + (\Z_i - \Z_j)^2 }{2 l_\textrm{SE}^2} \right )
-\end{equation}
-\end{widetext}
+  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right ) \times \\
+  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
+\end{multline}
 This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
@@
 For computational efficiency, we optimize the hyperparameters by finding the peak of the posterior distribution with a downhill gradient algorithm and then consider the hyperparameters fixed.
-
-Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
-\begin{equation}
-  \begin{bmatrix}
-    \mathbf{w}_k^\textrm{grid}\\
-    w_k
-  \end{bmatrix} \sim
-  \mathcal{N} \left (
-  \mathbf{0},
-  \begin{bmatrix}
-    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
-\mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
-  \end{bmatrix}
-  \right )
-\end{equation}
-where the dependence of the covariance kernel on the optimized hyperparameters is implied.
-Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$.
-And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$.
-The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$).
-The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
-\begin{multline}
-  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) = \\
-  \mathcal{N} \Bigl ( w_k \Bigl |\; \Bigr . \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}, \Bigr . \\
- \Bigl .
-\mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \Bigr )
- \label{eqn:conditional_gp}
-\end{multline}
-Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
 
@@
 
+Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
+\begin{equation}
+  \begin{bmatrix}
+    \mathbf{w}_k^\textrm{grid}\\
+    w_k
+  \end{bmatrix} \sim
+  \mathcal{N} \left (
+  \mathbf{0},
+  \begin{bmatrix}
+    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
+\mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
+  \end{bmatrix}
+  \right )
+\end{equation}
+where the dependence of the covariance kernel on the optimized hyperparameters is implied.
+Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$.
+And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$.
+The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$).
+The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
+\begin{equation}
+  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) =  \mathcal{N} \left ( w_k \Bigl | \Bigr . \,  \mu_k , \sigma_k \right )
+ \label{eqn:conditional_gp}
+\end{equation}
+where
+\begin{equation}
+ \mu_k =   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}
+\end{equation}
+\begin{multline}
+  \sigma_k = \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \\
+  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)
+\end{multline}
+
+Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
+
 If we desired an actual value of the interpolated weight, for example to reconstruct a model spectrum, we could simply draw a Gaussian random variable from the probability distribution in Eqn~\ref{eqn:conditional_gp}.
@@
 
-EDIT As an aside, this process of prediction is completely analogous to the process by which we generated the residual draws in the the covariance diagram.
-These were draws from the prior distribution (specified by Cdata).
-This is similar to drawing random values from the covariance matrix in Section XX given covariance matrices filled in by covariance kernels.
-Drawing from one giant Gaussian.
-Only now, we are drawing from a giant Gaussian conditioned on several points, which are the weights at the grid point locations.
-The benefit of using a Gaussian process for prediction/regression is that we now specified a probability distribution over possible interpolated values of the weights.
-We can draw multivariate samples from this probability distribution, like we did for the covariance matrices in Section XX to demonstrate the residuals (add note there to see here), only now we can also condition on the observed weight values.
-Notice how the GP directly interpolates the values at the grid points.
-For example, Fig A.2 (middle) shows draws from vectors of densely packed stellar values in order to show that simultaneous multivariate draws along a vector exhibit a certain structure of correlations (similar to the residuals in Section XX.
-However, the spread of subsequent multivariate draws at a specific single stellar value is given by a Gaussian distribution (Fig A.2 right).
+Next, we consider all $m$ of the Gaussian processes describing the weight for each eigenspectra collectively.
+For a given value of $\vt_\ast$, these Gaussian processes deliver a collection of $m$ one-dimensional Gaussian probability distributions.
+Because we have modelled each $w_k$ independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a multidimensional Gaussian with only diagonal elements filled
+\begin{equation}
+  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | \right . \mathbf{\mu}_\mathbf{w}, \mathbf{\Sigma}_\mathbf{w} \right )
+\end{equation}
+\begin{equation}
+  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | 
+  \begin{bmatrix}
+    \mu_1 \\
+    \vdots \\
+    \mu_m \\
+  \end{bmatrix},
+  \begin{bmatrix}
+    \sigma_1^2 & \ldots & 0 \\
+    \vdots & \ddots & \vdots \\
+    0 & \ldots & \sigma_m^2 \\
+  \end{bmatrix}
+  \right .
+  \right)
+  \label{eqn:weight_conditional}
+\end{equation}
 
@@
 However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the error across all pixels is smaller than 0.2\%, and thus any effect is negligible.
-We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra.
-This is the model.
-Then, we can rewrite Eqn~\ref{eqn:reconstruct2} as
-\begin{equation}
-  f_\lambda(\vt_\textrm{obs}) \approx \widetilde{\xi}_\mu + \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}} \mathbf{w}
-\end{equation}
-accompanied by the conditional probability on $\mathbf{w}$ (Eqn~\ref{eqn:conditional_gp}).
-
-The final task of designing the spectral emulator is to combine the posterior distribution of the data (Eqn~\ref{eqn:post}) with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:conditional_gp}) to form the joint distribution and then marginalize out the weights, such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.
-
-Cite Bishop, Gelman.
-
-The final modified covariance matrix looks like Fig A.3.
+We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra according to the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance parameters $(\phi_\mathsf{P})$.
+Now, the model spectrum is specified conditional on the vector of eigenspectra weights as
+\begin{equation}
+  \mathsf{M} \bigl | \bigr . \,\mathbf{w} = \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w}
+\end{equation}
+where 
+\begin{equation}
+  \mathbf{X} = \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}}
+\end{equation}
+Because the Gaussian process describes a probability distribution of the weights, we really have a probability distribution of possible (interpolated) models
+\begin{equation}
+  p(\mathsf{M} |\, \mathbf{w}) = p( \mathbf{w} |\, \vt_\ast)
+\end{equation}
+and the likelihood function (Eqn~\ref{eqn:likelihood}) is now also specified conditional on the weights
+\begin{equation}
+  p( \mathsf{D} |\, \mathsf{M}) = p( \mathsf{D} |\, \mathbf{w}) = 
+  \mathcal{N} \left ( \mathsf{D} |\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w} , \mathsf{C} \right )
+  \label{eqn:likelihood_conditional}
+\end{equation}
+
+The final task of designing the spectral emulator is to combine this data likelihood function with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:weight_conditional}) and then marginalize over the weights
+\begin{equation}
+  p(\mathsf{D} | \vt_\ast) = \int p(\mathsf{D} | \mathbf{w}) p( \mathbf{w} | \vt_\ast) d \mathbf{w}
+  \label{eqn:marginal}
+\end{equation}
+such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.
+To perform this multidimensional integral, we use a convenient lemma found in \citet[Appendix A]{gelman13}.
+
+If the probability distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in Eqns~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the marginal distribution (Eqn~\ref{eqn:marginal}) is
+\begin{equation}
+  p(\mathsf{D} | \vt_\ast, \vt_\textrm{obs}, \mathbf{\Phi}) = \mathcal{N} \left ( \mathsf{D} \bigl | \bigr .\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}, \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C} \right)
+\end{equation}
+where the dependence on the model parameters is now made explicit.
+We can couch this in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
+\begin{equation}
+  \mathsf{R}^\prime = \mathsf{D} - (\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w})
+\end{equation}
+ and
+\begin{equation}
+  \mathsf{C}^\prime = \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C}
+\end{equation}
+where $\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}$ can be thought of as the ``mean model spectrum'' given the model parameters, and the covariance matrix has been modified to account for the various possible interpolations.
 

### 2014-11-03 tex/paper.tex
Ready for SA edits

@@
 \startdata
-$T_{\rm eff}$ (K) & 6523     & 14                      &  6234   & 17        \\
+$T_{\rm eff}$ (K) & 6424     & 19                      &  6354   & 39        \\
 $\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
-$\Z$          & -0.28    & 0.01                    & -0.52  & 0.02     \\
-$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.28     & 0.06                    &  4.89   & 0.06      \\
-$v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
-$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -19.676 & 0.001 \\
+$\Z$          & -0.26    & 0.01                    & -0.30  & 0.02     \\
+$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.47 & 0.06       &  5.09   & 0.06      \\
+$v_z$ ($\kms$)        & -4.60    & 0.02                    & -4.86   & 0.02      \\
+$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -12.717 & 0.0007 \\
 $A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
@@
 atmospheres.}
-\tablenotetext{b}{The {\sc CfA/Kurucz} spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.
-The inference on $\Omega$ from the {\sc Phoenix} library is in steradian units.}
+\tablenotetext{b}{For the spectral emulator (Appendix~\ref{sec:Appendix}) the spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.}
 \tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
@@
 
-Aside from the structure of the covariance matrix, there is also a striking difference between the 
+Aside from the structure of the covariance matrix, there is also a moderate difference between the 
 posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
@@
 and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find a systematically cooler temperature 
-(by $\sim$240\,K) and lower metallicity (by a factor of $\sim$2) than inferred from the {\sc 
+(by $\sim$75\,K) and lower metallicity than inferred from the {\sc 
 CfA/Kurucz} models.  Because our framework explicitly mitigates the potential for posterior bias 
@@
 \vspace{-0.3cm}
-  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a {\sc Phoenix} model
-({\it red}) generated by drawing parameters from the posterior distribution (for a fixed $\log g = 
-5.0$).  ({\it bottom}) The residual spectrum along with contours representing the distributions of
-a large number of random draws from the covariance matrix (the shading is representative of the 1,
+  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a
+  {\sc Phoenix} model ({\it red}) generated by drawing parameters from the
+  posterior distribution. ({\it bottom}) The residual spectrum along with
+  contours representing the distributions of a large number of random draws
+  from the covariance matrix (the shading is representative of the 1,
 2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:Kurucz_residuals}.  
@@
 former is defined over a well-sampled, more regular grid of stellar parameters, we prefer it here 
-for computational simplicity.} and a $\delta$-function prior on $\log g$ (at 5.0), we infer 
+for computational simplicity.} and a Gaussian prior on $\log g$ (at $5.0 \pm 0.05$ dex), we infer 
 parameters for Gl~51 that are consistent with \citet{rojas-ayala12}.  
@@
 speculated that this discrepancy may be a consequence of inaccurate atomic data (oscillator 
-strengths and/or opacities).  The indexing technique adopted by \citeauthor{rojas-ayala12} presumes 
-that these discrepancies do not affect the line {\it ratios}.  Our approach offers an alternative 
-that makes no such assumption, instead using local covariance kernels to self-consistently 
+strengths and/or opacities).  Our approach uses local covariance kernels to self-consistently 
 determine the ``weights" of discrepant spectral lines in the overall fit.  In that sense, these 
@@
 \startdata
-$T_{\rm eff}$ (K) & 2998  & 17 \\
-$\log g$ & 5.0 & 0 (fixed) \\
-$\Z$ & -0.01 & 0.04 \\
-$v \sin i$ ($\kms$) & \hili{5.68} & \hili{4.15} \\
-$v_z$ ($\kms$) & 12.1 & 1.3 \\
-$\log \Omega$ (sr) & -19.628  & 0.003 \\
+$T_{\rm eff}$ (K) & 3071  & 28 \\
+$\log g$ & 5.19 & 0.04 \\
+$\Z$ & 0.14 & 0.07 \\
+$v \sin i$ ($\kms$) & \multicolumn{2}{c}{$\leq 15.0$} \\
+$v_z$ ($\kms$) & 11.6 & 1.3 \\
+$\log \Omega$ (sr) & -13.481  & 0.001 \\
 $A_V$ & 0 & 0 (fixed)
@@
 
-Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
+Interpolation from a library of synthetic spectra is a conceptually straightforward process, but in the 
 application discussed here there are some subtleties that require detailed consideration.  
-Spectral line formation is a non-linear process, where line shapes are determined by the complex 
+Spectral line formation is a non-linear process, where line shapes are determined by complex 
 radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
-process and predict the behavior of spectral lines based on pre-computed 
+process and predict the behavior of spectral lines at intermediate points based on pre-computed 
 library spectra with similar physical parameters.  This appendix describes the 
@@
 
-There are two important factors to consider when interpolating in libraries of model stellar 
+There are two important factors to consider when interpolating libraries of model stellar 
 spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
@@
 
-Model library spectra, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, are stored as (1-dimensional) arrays of flux 
-values, sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).  In the case 
-of interest here, the collections of model parameters $\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z 
-\}]$ define the dimensions of the library grid; the full library is therefore encapsulated in a 
-4-dimensional array.  Both the libraries used here have grid spacings of 0.5 dex in $\log g$ and 
-0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but the {\sc 
-Phoenix} library has finer coverage in 100\,K increments.
+Model library spectra are stored as (1-dimensional) arrays of flux values,
+sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).
+In the case of interest here, the collections of model parameters
+$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the
+dimensions of the library grid; the full library
+$f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$ is therefore encapsulated in a
+4-dimensional array.  Both the libraries used here have grid spacings of 0.5
+dex in $\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by
+250\,K in $T_{\rm eff}$, but the {\sc Phoenix} library has finer coverage in
+100\,K increments.
 
@@
   \includegraphics{figs/pca_reconstruct.pdf}
-  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51.
+  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51, generated using a subset of the \textsc{Phoenix} grid parameters space most relevant for M dwarfs.
 (\emph{bottom}) The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
@@
 We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}.
-Prior to the PCA decomposition, we first identify a subset of $M$ synthetic spectra which will be relevant to the star at hand.
-We then standardize the synthetic library by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid.
+Prior to the PCA decomposition, we first isolate a subset of the library (containing $M$ spectra) with parameter values that will be most relevant to the star at hand.
+We then standardize these spectra by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid.
 The mean spectrum is
 \begin{equation}
-  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
+  \xi_\mu = \frac{1}{M} \sum_{i = 1}^M f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
 \end{equation}
@@
 \begin{equation}
-  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
-\end{equation}
-$\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of stellar parameters sampled in the spectral library, while $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection.
+  \xi_\sigma = \sqrt{\frac{1}{M} \sum_{i=1}^M \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
+\end{equation}
+where $\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of the $M$ pairings of stellar parameters sampled in the spectral library and $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection.
 All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
 \begin{equation}
-  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
+  \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
 \end{equation}
 Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}).
-We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$).
-We denote a single $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$.
-As an example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.
-
-Using the PCA basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
-\begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
-  \label{eqn:reconstruct}
-\end{equation}
-where $w_k$ is the weight of a given eigenspectrum, $k$.
-This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$.
-We can determine the weights at the grid points in the library by taking the dot product of the standardized synthetic spectrum with each of the eigenspectra.
-The weight for a given eigenspectrum $\xi_k$ is
-\begin{equation}
-  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
-\end{equation}
-which is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel). To simplify the notation, we can write the weights in a length-$m$ column vector
-\begin{equation}
-  \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i) = 
-  \begin{bmatrix}
-    w_1(\{\vt_\ast \}^\textrm{grid}_i)\\
-    w_2(\{\vt_\ast \}^\textrm{grid}_i)\\
-    \vdots\\
-    w_m(\{\vt_\ast \}^\textrm{grid}_i)
-  \end{bmatrix}
-\end{equation}
-and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
-\begin{equation}
-  \mathbf{\Xi} = 
-  \begin{bmatrix}
-    \xi_1 & \xi_2 & \cdots & \xi_m \\
-  \end{bmatrix}
-\end{equation}
-Then we can rewrite Equation~\ref{eqn:reconstruct} as
-\begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
-  \label{eqn:reconstruct2}
-\end{equation}
-Where $I_{N_{\rm pix}}$ is an $N_{\rm pix}$ by $N_{\rm pix}$ identity matrix.
-To recapitulate thus far, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
-The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters.
-Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value.
-In practice, it might be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly.
-However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum could constitute a significant portion of the uncertainty budget.
-This, combined with the Nyquist undersampling of the synthetic grid caused behavior where stellar parameters corresponding to a grid point in the synthetic library might be preferentially favored and bias the results, even after empirically accounting for interpolation error by doing ``drop-out'' interpolation tests.
-
-Now, we describe how we design the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters.
-In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum.
-Alternatively, we now use a different Gaussian process in a separate application to serve as an interpolating function for the weights.
-
-We model each eigenspectrum weight independently; therefore there is a separate Gaussian process ($\mathcal{GP}$) for each eigenspectrum.
-For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$.
-The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
-\begin{equation}
-  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( 0, \mathcal{K}_\textrm{SE} (\{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})\right )
-\end{equation}
-The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$.
-Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
-\begin{multline}
-  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
-  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right ) \times \\
-  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
-\end{multline}
-This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
-Then the $M$ by $M$ covariance matrix is
-\begin{equation}
-\mathsf{C}^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | a_\textrm{SE}, l_\textrm{SE})
-\end{equation}
-which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters.
-
-The optimal parameters for the hyperparameters of the squared exponential are given by the multi-variate Gaussian probability distribution, which is called the marginal likelihood function or ``evidence'' for the weights \citep{rasmussen05},
-\begin{equation}
-  p(\mathbf{w}_k^\textrm{grid} | C^\textrm{grid}) = \mathcal{N}\left (\mathbf{w}_k^\textrm{grid} \left | \; \mathbf{0}, \mathsf{C}^\textrm{grid} \right . \right )
-\end{equation}
-Because $\mathsf{C}^\textrm{grid}$ is defined by $\mathcal{K}_\textrm{SE}$, this likelihood function is equivalent to $p(\mathbf{w}_k^\textrm{grid} |\; a_\textrm{SE}, l_\textrm{SE})$.
-By multiplying this with a suitable prior over $a_\textrm{SE}$ and $l_\textrm{SE}$, we arrive at a posterior distribution for the hyperparameters of the Gaussian process.
-For computational efficiency, we optimize the hyperparameters by finding the peak of the posterior distribution with a downhill gradient algorithm and then consider the hyperparameters fixed.
+We decide to truncate our basis to the first $m$ eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$).
+We denote a single length $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$.
+For example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.
 
@@
   \includegraphics{figs/GP_right.pdf}
-  \figcaption{ (\emph{left}) the blue dots mark the weights $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ for $\vt_\ast$ for various values of $T_\textrm{eff}$ with $\log g = 5.0$ and $\Z = 0.0$.
+  \figcaption{The Gaussian process modelling of the principal component weights for the Gl51 \textsc{Phoenix} spectral library.
+(\emph{left}) the blue dots mark the weights $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ computed at a one-dimensional slice of the spectral library for grid points with $\log g = 5.0$, $\Z = 0.0$ and various values of $T_\textrm{eff}$.
+In reality, the weights are a three-dimensional function of $\vt_\ast$.
 The thin blue lines show 50 random draws of possible functional forms described by the Gaussian process, constrained to pass through the grid weights.
@@
 
-Now, with the Gaussian process specified, we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
+Using the principal component basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
+  \label{eqn:reconstruct}
+\end{equation}
+where $w_k$ is the weight of a given eigenspectrum, $k$.
+This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$.
+We can determine the weight at any grid point in the library by taking the dot product of the standardized synthetic spectrum with the eigenspectrum
+\begin{equation}
+  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
+\end{equation}
+and is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel).
+To simplify notation, we can write the collection of eigenspectra weights in a length-$m$ column vector
+\begin{equation}
+  \mathbf{w}(\vt_\ast) = 
+  \begin{bmatrix}
+    w_1(\vt_\ast )\\
+    w_2(\vt_\ast )\\
+    \vdots\\
+    w_m(\vt_\ast )
+  \end{bmatrix}
+\end{equation}
+and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
+\begin{equation}
+  \mathbf{\Xi} = 
+  \begin{bmatrix}
+    \xi_1 & \xi_2 & \cdots & \xi_m \\
+  \end{bmatrix} .
+\end{equation}
+Then, we can rewrite Equation~\ref{eqn:reconstruct} as
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
+  \label{eqn:reconstruct2}
+\end{equation}
+where $I_{N_{\rm pix}}$ is an ($N_{\rm pix}, N_{\rm pix}$) identity matrix.
+
+To recapitulate, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
+The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters.
+Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value.
+In practice, it may be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly.
+However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum can constitute a significant portion of the uncertainty budget.
+This, combined with the Nyquist undersampling of the synthetic grid can cause artificial ``peaking'' of the posterior near grid points in the synthetic library, because the interpolated spectrum is not as good as the raw spectrum at the grid point.
+Even explicitly accounting for interpolation error by doing ``drop-out'' interpolation tests and empricially propagating it forward did not relieve the peaking problem.
+We address this problem by using a Gaussian process to encapsulate the range of possible interpolated spectra.
+
+Now, we describe the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters.
+In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum.
+We now use a separate Gaussian process to serve as an interpolating function for the weights.
+Because each eigenspectrum weight is modelled independently, there is a new Gaussian process ($\mathcal{GP}$) for each eigenspectrum.
+For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$.
+The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
+\begin{equation}
+  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right )
+\end{equation}
+The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$.
+Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
+\begin{multline}
+  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_{T_\textrm{eff}},  l_{\log g}, l_{\Z}) = \\
+  a^2 \exp \left ( - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right )
+  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \times \\
+  \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
+\end{multline}
+This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
+The squared exponential kernel produces a smooth functional form which closely matches the behavior of the eigenspectra weights (Fig~\ref{fig:GP_interp}, left panel).
+Then the ($M,M$) covariance matrix is
+\begin{equation}
+\mathbf{\Sigma}_k^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})
+\end{equation}
+which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters, and the conditional dependence on the kernel hyperparameters is implied.
+
+The optimal parameters for the hyperparameters of the squared exponential are given by the multi-variate Gaussian probability distribution, which is called the marginal likelihood function or ``evidence'' for the weights \citep{rasmussen05},
+\begin{equation}
+  p(\mathbf{w}_k^\textrm{grid} | \mathbf{\Sigma}_k^\textrm{grid} ) = \mathcal{N}\left (\mathbf{w}_k^\textrm{grid} \left | \; \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right . \right )
+\end{equation}
+Because $\mathbf{\Sigma}_k^\textrm{grid}$ is defined by $\mathcal{K}_\textrm{SE}$ and its hyperparameters, this likelihood function is equivalent to $p(\mathbf{w}_k^\textrm{grid} |\; a_\textrm{SE}, l_{T_\textrm{eff}},  l_{\log g}, l_{\Z})$.
+By multiplying this with a suitable prior over $a_\textrm{SE}$ and $l$, we arrive at a posterior distribution for the hyperparameters of the Gaussian process.
+For computational efficiency, we optimize the hyperparameters by finding the peak of the posterior distribution with a downhill gradient algorithm and then consider the hyperparameters fixed.
+
+Now the Gaussian process is specified and we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
 \begin{equation}
@@
   \begin{bmatrix}
-    \mathsf{C}^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
+  \mathbf{\Sigma}_k^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
@@
 \end{equation}
-where the dependence of the covariance kernel on the optimized hyperparameters is implied.
-Notation means a multivariate Gaussian with $\mu$ vector = 0, length $M$.
-And then $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ vector which is horizontally stacked to  $\mathsf{C}^\textrm{grid}$.
-The combination of all of these products results in a total covariance matrix that is size ($M + 1$ by $M + 1$).
+where $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ column vector where $\mathcal{K}_\textrm{SE}$ is evaluated pairwise with all values in $\{\vt_\ast \}^\textrm{grid}$ and the proposed value of $\vt_\ast$.
+This vector is stacked horizontally next to  $\mathbf{\Sigma}^\textrm{grid}$, and its transpose is stacked vertically underneath $\mathbf{\Sigma}^\textrm{grid}$.
+The combination of all of these products results in an augmented covariance matrix of size ($M + 1, M + 1$).
 The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
@@
 \begin{equation}
- \mu_k =   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}
+ \mu_k =   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathbf{\Sigma}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}
 \end{equation}
@@
   \sigma_k = \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \\
-  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathsf{C}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)
+  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathbf{\Sigma}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)
 \end{multline}
 
-Although this equation is complex, the result is simpler to interpret: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
+Although this equation is complex, the interpretation is straightforward: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
 
@@
 For a given value of $\vt_\ast$, these Gaussian processes deliver a collection of $m$ one-dimensional Gaussian probability distributions.
-Because we have modelled each $w_k$ independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a multidimensional Gaussian with only diagonal elements filled
+Because we have modelled each $w_k$ independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a multidimensional Gaussian with only the diagonal elements filled
 \begin{equation}
@@
 Fortunately, because convolution is a linear operation, we can first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the reconstructed spectrum as a linear combination of these modified eigenspectra without loss of information.
-Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them as linear.
-However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the error across all pixels is smaller than 0.2\%, and thus any effect is negligible.
-We let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra according to the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance parameters $(\phi_\mathsf{P})$.
+Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them in this manner.
+However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the flux error due to resampling is smaller than 0.2\% across all pixels , and thus any effect is negligible.
+For notational compactness, we let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra, with an implied dependence on the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance parameters $(\phi_\mathsf{P})$.
 Now, the model spectrum is specified conditional on the vector of eigenspectra weights as
@@
 \end{equation}
-Because the Gaussian process describes a probability distribution of the weights, we really have a probability distribution of possible (interpolated) models
+Because the Gaussian process describes a probability distribution of the weights, we now have a probability distribution of possible (interpolated) models
 \begin{equation}
@@
 \end{equation}
-and the likelihood function (Eqn~\ref{eqn:likelihood}) is now also specified conditional on the weights
+and the likelihood function (Eqn~\ref{eqn:likelihood}) is specified conditional on the weights
 \begin{equation}
@@
 such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.
-To perform this multidimensional integral, we use a convenient lemma found in \citet[Appendix A]{gelman13}.
-
-If the probability distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in Eqns~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the marginal distribution (Eqn~\ref{eqn:marginal}) is
+To perform this multidimensional integral, we use a convenient lemma found in \citet[Appendix A]{gelman13}: if the probability distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in Eqns~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the marginal distribution (Eqn~\ref{eqn:marginal}) is
 \begin{equation}
@@
 where the dependence on the model parameters is now made explicit.
-We can couch this in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
-\begin{equation}
-  \mathsf{R}^\prime = \mathsf{D} - (\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w})
-\end{equation}
- and
+We can couch this modified likelihood function in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
+\begin{equation}
+  \mathsf{M}^\prime = \widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}
+\end{equation}
+\begin{equation}
+  \mathsf{R}^\prime = \mathsf{D} - \mathsf{M}^\prime 
+\end{equation}
 \begin{equation}
@@
 \end{equation}
-where $\widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}$ can be thought of as the ``mean model spectrum'' given the model parameters, and the covariance matrix has been modified to account for the various possible interpolations.
+where $\mathsf{M}^\prime$ can be thought of as the ``mean model spectrum'' given the model parameters, and the covariance matrix has been modified to account for the various probable manifestations of the model spectrum about the mean spectrum.
 

### 2014-11-03 tex/paper.tex
full editing by SA (last time!)

@@
 %\documentclass[preprint]{aastex} 
-\documentclass[iop,floatfix,twocolappendix]{emulateapj}
+\documentclass[iop,floatfix,numberedappendix,twocolappendix]{emulateapj}
 
@@
 imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
-systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
-observations to guide substantial improvements in the models themselves.  An overview of the 
+systematic uncertainties into the parameter inferences, and ultimately can serve as a tool for 
+using observations to guide substantial improvements in the models themselves.  An overview of the 
 methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
@@
 The remainder of this Section describes the mechanics of this modeling framework.  First, a model 
-spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
-then post-processed to mimic reality using a set of observational and practical nuisance parameters 
-(Section \ref{subsec:postprocess}).  Next, a direct, pixel-by-pixel comparison between the data and 
-model spectra is made with a prescribed likelihood function and a parametric treatment of the 
-covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
-under the guise of hierarchical Markov Chain Monte Carlo (MCMC) simulations to numerically explore 
-the posterior probability density of the model conditioned on the data, and thereby to determine 
-constraints on the parameters of interest (Section \ref{subsec:MCMC}).  Along the way, these 
-procedures are illustrated with real observations of the high resolution optical spectrum from a 
-nearby F star.  That specific application, along with some alternative demonstrations of the 
-method, are discussed in more detail in Section \ref{sec:examples}.
+spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}; 
+Appendix~\ref{sec:Appendix}), and then post-processed to mimic reality using a set of observational 
+and practical nuisance parameters (Section \ref{subsec:postprocess}).  Next, a direct, 
+pixel-by-pixel comparison between the data and model spectra is made with a prescribed likelihood 
+function and a parametric treatment of the covariances between pixel residuals (Section 
+\ref{subsec:likelihood}).  That process is iterated under the guise of hierarchical Markov Chain 
+Monte Carlo (MCMC) simulations to numerically explore the posterior probability density of the 
+model conditioned on the data, and thereby to determine constraints on the parameters of interest 
+(Section \ref{subsec:MCMC}).  Along the way, these procedures are illustrated with observations of 
+the high resolution optical spectrum from a nearby F star.  That specific application, along with 
+some alternative demonstrations of the method, are discussed in more detail in Section 
+\ref{sec:examples}.
 
@@
 grid point, and then comparing the interpolated spectrum with the corresponding library spectrum 
-\citep[see also][]{husser12}.  We have developed this approach to propagate the interpolation 
-uncertainties into the likelihood calculations, as described briefly in 
-Section~\ref{subsec:likelihood} and elaborated in the Appendix.
+\citep[see also][]{husser12}.  To suitably propagate the interpolation uncertainties into the 
+likelihood calculations, we employ a Bayesian emulation technique that is described in detail in 
+Appendix \ref{sec:Appendix}.
 
@@
 \begin{eqnarray} \label{eqn:broadening}
-\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \otimes \varphi_v \\
-                                           &=& \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop}, \nonumber
+\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \ast \varphi_v \\
+                                           &=& \flam(\vt_{\ast}) \ast \mathcal{F}_v^{\rm inst} \ast \mathcal{F}_v^{\rm rot} \ast \mathcal{F}_v^{\rm dop}, \nonumber
 \end{eqnarray} 
@@
 \end{eqnarray}
-where $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
+where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
 coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
@@
 \begin{equation}
-  \Kglobal_{ij}(\vp_{{\mathsf C}, g}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm g}}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
@@
 \end{equation}
-where the (hyper-)parameters $\vp_{{\mathsf C}, g} = [a_g, \ell]$ include an amplitude ($a_g$) and 
-scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
+where the (hyper-)parameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
+($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
 enables computational expendiency, we employ a Hann window function
@@
 \end{equation}
-as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the 
-scale (we set $r_0 = 4\ell$).
+as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
+set $r_0 = 4\ell$).
 
@@
 
-The procedure outlined above can be a significant computational challenge.  A
-typical data spectrum has $N_{\rm pix} > \mathcal{O}(10^3)$, and therefore the
-many repeated evaluations of the matrix product $\vR^{\trans} \vC^{-1} \vR$ in
-the likelihood calculation can be numerically expensive.  However, because
-$\vC$ is a symmetric, positive semi-definite matrix, we can employ Cholesky
-factorization to optimize the evaluation of the matrix product ($\vR^{\trans}
-\vC^{-1} \vR$) and avoid the direct calculation of the matrix inverse
-($\vC^{-1}$) itself.  For echelle spectra with multiple orders or multiple
-spectra of the same object, taken with different instruments, the nuisance
-parameters for each chunk of spectrum are independent from the other chunks. This
-means that the computationally intensive steps of generating a model spectrum
-for a specific wavelength range and evaluating the likelihood can be
-parallelized. The only segment of the code that needs to be synchronized is the
-MCMC proposal of stellar parameters, which are shared between all chunks of the
-spectrum. Because all of the computationally expensive steps can be
-parallelized, this algorithm can be massively parallelized on a cluster,
-enabling simultaneous inference for wide spectral ranges at high resolution or
-multiple spectral datasets acquired with different instruments. To sample this
-model, we extend the Metropolis-Hastings sampler included in the {\tt emcee}
-package \citep{foreman-mackey13} to function within a parallelized blocked
-Gibbs sampler.\footnote{\url{https://github.com/iancze/emcee}}\\
+This procedure can be a significant computational challenge.  A typical data spectrum has $N_{\rm 
+pix} > \mathcal{O}(10^3)$, and therefore the many evaluations of the matrix product $\vR^{\trans} 
+\vC^{-1} \vR$ in the likelihood calculation can be numerically expensive.  However, because $\vC$ 
+is a symmetric, positive semi-definite matrix, we can employ Cholesky factorization to optimize the 
+evaluation of the matrix product and avoid the direct calculation of the matrix inversion 
+($\vC^{-1}$).  For multi-order echelle spectra or multiple spectra of the same target (perhaps 
+taken with different instruments), the nuisance parameters for each segment of the spectrum are 
+independent. This means that the computationally intensive steps of generating a model spectrum
+for a specific wavelength range and evaluating the likelihood can be parallelized. The only segment 
+of the code that needs to be synchronized is the MCMC proposal of stellar parameters, which are 
+shared between all chunks of the spectrum.  The massive parallelization of this algorithm on a 
+computer cluster therefore enables the simultaneous inference of interesting parameters over wide 
+spectral ranges at high resolution, or from multiple datasets.  To sample the posteriors in this 
+mode, we extend the Metropolis-Hastings sampler included in the {\tt emcee} package 
+\citep{foreman-mackey13} to function within a parallelized blocked Gibbs 
+sampler.\footnote{See \url{https://github.com/iancze/emcee}.}\\
 
@@
 relevant.  Using the same TRES spectral orders, the {\sc CfA/Kurucz} library, and a 
-$\delta$-function prior on $\log g$ (at 4.29), we infer parameters that are consistent with the 
-{\tt SPC} approach; see Table~\ref{table:Kurucz}.  Figure~\ref{fig:Kurucz_residuals} demonstrates 
-the quality of our approach in reproducing the WASP-14 spectrum and, especially significant, the 
-structure of the residuals.  Figure~\ref{fig:Kurucz_posterior} shows some marginal projections of 
-the posterior-space, highlighting key parameter degeneracies.
+$\delta$-function prior on $\log g$ (at 4.29), we infer parameters that are roughly consistent with 
+the {\tt SPC} approach; see Table~\ref{table:Kurucz}.  Figure~\ref{fig:Kurucz_residuals} 
+demonstrates the quality of our approach in reproducing the WASP-14 spectrum and, especially 
+significant, the structure of the residuals.  Figure~\ref{fig:Kurucz_posterior} shows marginal 
+projections of the posterior-space, highlighting key parameter degeneracies.
 
@@
 atmospheres.}
-\tablenotetext{b}{For the spectral emulator (Appendix~\ref{sec:Appendix}) the spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.}
+\tablenotetext{b}{For the spectral emulator (Appendix~\ref{sec:Appendix}), the models are scaled to 
+an average flux of unity, rather than flux at the stellar surface; the quoted values are unknown 
+multiples of $R^2/d^2$.}
 \tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
@@
 It is worthwhile to comment on the inferred parameter uncertainties.  In this specific example, 
-some experimentation with restricted forms of the covariance matrix confirm that the uncertainties 
+some experimentation with restricted forms of the covariance matrix confirms that the uncertainties 
 of the interesting parameters have roughly equal contributions from trivial (Poisson) noise, grid 
@@
 
-Aside from the structure of the covariance matrix, there is also a moderate difference between the 
+Aside from the structure of the covariance matrix, there is also a modest difference between the 
 posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
@@
 and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find a systematically cooler temperature 
-(by $\sim$75\,K) and lower metallicity than inferred from the {\sc 
-CfA/Kurucz} models.  Because our framework explicitly mitigates the potential for posterior bias 
-from imperfect matches between data and models, we can safely conclude that these apparent 
-discrepancies are intrinsic to the different input physics of the model libraries.\footnote{It is 
-worth noting that these differences are not a manifestation of the stringent prior on surface 
-gravity.  When the spectra are modeled with $\log g$ unconstrained, the posteriors for both 
-libraries shift, but their respective discrepancies actually increase slightly.}  Such a comparison 
-serves as a stark reminder that incomplete and/or over-simplified physical prescriptions remain the 
-most significant source of systematic uncertainty in estimates of fundamental stellar parameters.
+(by $\sim$75\,K) and lower metallicity than the {\sc CfA/Kurucz} models.  Because our framework 
+explicitly mitigates the potential for posterior bias from imperfect matches between data and 
+models, these apparent discrepancies must be intrinsic to the different input physics of the model 
+libraries.\footnote{It is worth noting that these differences are not a manifestation of the 
+stringent prior on surface gravity.  When the spectra are modeled with $\log g$ unconstrained, the 
+posteriors for both libraries shift, but their respective discrepancies actually increase 
+slightly.}  Such a comparison serves as a stark reminder that incomplete and/or over-simplified 
+physical prescriptions remain the most significant source of systematic uncertainty in estimates of 
+fundamental stellar parameters.
 
@@
 As a point of reference, \citet{rojas-ayala12} used a $K$-band spectrum at similar resolution to 
-the SPEX data to estimate the basic Gl~51 parameters using an index technique.  They measured 
+the SPEX data to estimate the basic Gl~51 parameters using an indexing technique.  They measured 
 equivalent widths for the 2.205\,$\mu$m \ion{Na}{1} doublet and 2.263\,$\mu$m \ion{Ca}{1} triplet, 
@@
 promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
-like those being collected in programs like the APOGEE and HERMES surveys \citep{nidever12, zucker12}), since they provide enormous leverage for identifying and improving the underlying model systematics.
-\\
-
-\acknowledgments  The authors would like to graciously acknowledge the following people for many helpful discussions and key insights: Gregory M.
-Green, Kaisey Mandel, Daniel Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Rebecca Dawson, Tom Loredo and the ExoStat group, Allyson Bieryla, and Maxwell Moe.
-IC was supported by the NSF Graduate Fellowship and the Smithsonian Institution.
-Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.
-This research made use of Astropy \citep{astropy13} and the Julia programming language \citep{julia12}.
+like those being collected in programs like the APOGEE and HERMES surveys; \citealt{nidever12,
+zucker12}), since they provide enormous leverage for identifying and improving the underlying model 
+systematics. \\
+
+\acknowledgments  The authors would like to acknowledge the following people for many 
+extraordinarily helpful discussions and key insights: Gregory M.~Green, Kaisey Mandel, Daniel 
+Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Rebekah Dawson, Tom Loredo and the 
+ExoStat group, Allyson Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship 
+and the Smithsonian Institution.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, 
+and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.  
+This research made extensive use of Astropy \citep{astropy13} and the Julia programming language 
+\citep{julia12}.
 
@@
 
-\section{Spectral Emulator} \label{sec:Appendix}
-
-Interpolation from a library of synthetic spectra is a conceptually straightforward process, but in the 
-application discussed here there are some subtleties that require detailed consideration.  
-Spectral line formation is a non-linear process, where line shapes are determined by complex 
-radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
-process and predict the behavior of spectral lines at intermediate points based on pre-computed 
-library spectra with similar physical parameters.  This appendix describes the 
-procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
-prediction--which can be substantial--through the probabilistic framework we have developed.
-
-There are two important factors to consider when interpolating libraries of model stellar 
-spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
-monotonically with changes in the physical parameters that define the library: this helps inform 
-the decision about which functional form of interpolation should be used.  And second, the spectra 
-are usually not Nyquist sampled with respect to the physical parameters that define the library.  
-This unfortunate fact means that the library is not sampled densely enough in 
-parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
-information loss.
-Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra and a chosen interpolation point, delivers a probability distribution over the range of possible interpolated spectra.
-
-Model library spectra are stored as (1-dimensional) arrays of flux values,
-sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).
-In the case of interest here, the collections of model parameters
-$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the
-dimensions of the library grid; the full library
-$f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$ is therefore encapsulated in a
-4-dimensional array.  Both the libraries used here have grid spacings of 0.5
-dex in $\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by
-250\,K in $T_{\rm eff}$, but the {\sc Phoenix} library has finer coverage in
-100\,K increments.
+\section{Spectral Emulator for Interpolation} \label{sec:Appendix}
+
+Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
+application being considered here there are some subtleties that necessitate a more sophisticated 
+approach.  Spectral line formation is a non-linear process, where line shapes are determined by 
+complex radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass 
+this non-linear process and predict the behavior of many different spectral lines for an arbitrary 
+set of parameters based on a pre-computed set of library spectra with similar (neighboring) 
+parameters.  In this Appendix, we describe the procedure through which we propagate an empirical 
+approximation of the {\it uncertainty} in that prediction -- which can be substantial -- into and 
+through the probabilistic inference framework we have developed.
+
+There are two important factors to consider when interpolating libraries of model stellar spectra.  
+First, the fluxes in a non-negligible fraction of the spectral pixels do not vary monotonically 
+with changes in the physical parameters that define the library, $\vt_{\ast}$: this helps inform 
+the decision about which functional form of interpolation should be used.  Second, the spectra 
+are usually not Nyquist sampled with respect to $\vt_{\ast}$.  Unfortunately, this fact means that 
+the library is not sampled densely enough in parameter-space to reconstruct a synthetic spectrum 
+for arbitrary parameters without some information loss.  Given these factors, we choose to develop 
+a spectral ``emulator'' to aid in the interpolation.  Given the information contained in 
+the spectral library and an arbitrary desired $\vt_{\ast}$, this emulator delivers a probability 
+distribution representing the range of possible interpolated spectra.
+
+Model library spectra are stored as (1-dimensional) arrays of fluxes, sampled on high resolution 
+wavelength grids.  In the case of interest here, the sets of model parameters 
+$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the dimensions of the 
+library grid.  The full spectral library, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, is therefore 
+encapsulated in a 4-dimensional array.  The libraries used here have grid spacings of 0.5\,dex in 
+$\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but 
+the {\sc Phoenix} library has finer coverage in 100\,K increments.
 
@@
   \includegraphics{figs/pca_reconstruct.pdf}
-  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51, generated using a subset of the \textsc{Phoenix} grid parameters space most relevant for M dwarfs.
-(\emph{bottom}) The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
-\label{fig:pca_reconstruct}}
+  \vspace{-0.6cm}
+  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra 
+that form the basis of the \textsc{Phoenix} synthetic library used to model Gl~51, generated using 
+a subset of the parameter-space most relevant for M dwarfs.  (\emph{bottom}) The original synthetic 
+spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear 
+combination of the derived eigenspectra using Eqn~\ref{eqn:reconstruct} (with the weights $w_k$ 
+listed in the top panel figure).  \label{fig:pca_reconstruct}}
   \end{center}
@@
 
-The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}.
-We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}.
-Prior to the PCA decomposition, we first isolate a subset of the library (containing $M$ spectra) with parameter values that will be most relevant to the star at hand.
-We then standardize these spectra by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid.
-The mean spectrum is
+The first step in designing a spectral emulator is to break down the library into an appropriate 
+basis \citep{habib07, heitmann09}.  We chose the principal component basis to decompose the library 
+into a set of ``eigenspectra", following the techniques of \citet{ivezic13}.  Prior to this 
+decomposition, we isolate a subset of the library (containing $M$ spectra) with parameter values 
+that will be most relevant to the target being considered (e.g., for Gl~51, this means considering 
+only effective temperatures below $\sim$3800\,K).  We then standardize these spectra by subtracting 
+off their mean spectrum and then ``whitening" them by dividing off the standard deviation spectrum 
+measured in each pixel across the grid.  The mean spectrum is 
 \begin{equation}
@@
 \begin{equation}
-  \xi_\sigma = \sqrt{\frac{1}{M} \sum_{i=1}^M \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
-\end{equation}
-where $\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of the $M$ pairings of stellar parameters sampled in the spectral library and $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection.
-All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
-\begin{equation}
-  \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
-\end{equation}
-Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}).
-We decide to truncate our basis to the first $m$ eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$).
-We denote a single length $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$.
-For example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.
+  \xi_\sigma = \sqrt{\frac{1}{M} \sum_{i=1}^M \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 }, 
+\end{equation}
+where $\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of the $M$ sets of stellar 
+parameters under consideration in the library and $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single 
+set of those parameters drawn from this collection.  In effect, all library spectra are 
+standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
+\begin{equation}
+  \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}.
+\end{equation}
+
+The eigenspectra are computed from this standardized grid using principal component analysis 
+\citep[PCA;][]{ivezic13}.  We decided to truncate our basis to the first $m$ eigenspectra, where 
+$m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid 
+to better than 2\% accuracy for all pixels (the mean pixel error is generally much smaller than 
+this, $\lesssim 0.5\%$).  We denote a single $N_{\rm pix}$-dimensional eigenspectrum as $\xi_k$, 
+with principal component index $k = \{1, 2, \ldots, m\}$.  As an example, the eigenspectra basis 
+computed for Gl~51 using the \textsc{Phoenix} library is shown in the top panel of 
+Figure~\ref{fig:pca_reconstruct}.
 
@@
 
-Using the principal component basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
-\begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
+Using the principal component basis, we can lossily reconstruct any spectrum from the library with 
+a linear combination of the eigenspectra
+% Is "lossily" really a word?  If so, wtf does it mean?
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \, \xi_k
   \label{eqn:reconstruct}
 \end{equation}
-where $w_k$ is the weight of a given eigenspectrum, $k$.
-This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$.
-We can determine the weight at any grid point in the library by taking the dot product of the standardized synthetic spectrum with the eigenspectrum
-\begin{equation}
-  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
-\end{equation}
-and is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel).
-To simplify notation, we can write the collection of eigenspectra weights in a length-$m$ column vector
+where $w_k$ is the weight of the $k^{\rm th}$ eigenspectrum.  These weights are 3-dimensional 
+scalar functions that depend on the stellar parameters $\vt_\ast$.  These weights, which are 
+generally smooth functions of the stellar parameters (see the left panel of 
+Figure~\ref{fig:GP_interp}), can be determined at any grid point in the library by taking the dot 
+product of the standardized synthetic spectrum with the eigenspectrum  
+\begin{equation}
+  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \, \xi_k.
+\end{equation}
+To simplify notation, we can write the collection of eigenspectra weights in a length-$m$ column 
+vector
 \begin{equation}
@@
 and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
+% Christ, you know your paper's a mess when you start throwing around capital xi's.
 \begin{equation}
@@
 \end{equation}
-Then, we can rewrite Equation~\ref{eqn:reconstruct} as
-\begin{equation}
-  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
+Then, we can rewrite Eq.~(\ref{eqn:reconstruct}) as
+\begin{equation}
+  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \, \mathsf{I}_{N_{\rm pix}} \, \mathbf{\Xi} \, \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
   \label{eqn:reconstruct2}
 \end{equation}
-where $I_{N_{\rm pix}}$ is an ($N_{\rm pix}, N_{\rm pix}$) identity matrix.
-
-To recapitulate, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra.
-The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters.
-Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value.
-In practice, it may be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly.
-However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum can constitute a significant portion of the uncertainty budget.
-This, combined with the Nyquist undersampling of the synthetic grid can cause artificial ``peaking'' of the posterior near grid points in the synthetic library, because the interpolated spectrum is not as good as the raw spectrum at the grid point.
-Even explicitly accounting for interpolation error by doing ``drop-out'' interpolation tests and empricially propagating it forward did not relieve the peaking problem.
-We address this problem by using a Gaussian process to encapsulate the range of possible interpolated spectra.
-
-Now, we describe the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters.
-In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum.
-We now use a separate Gaussian process to serve as an interpolating function for the weights.
-Because each eigenspectrum weight is modelled independently, there is a new Gaussian process ($\mathcal{GP}$) for each eigenspectrum.
-For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$.
-The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
-\begin{equation}
-  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right )
-\end{equation}
-The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$.
-Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
-\begin{multline}
-  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_{T_\textrm{eff}},  l_{\log g}, l_{\Z}) = \\
-  a^2 \exp \left ( - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right )
-  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \times \\
-  \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
-\end{multline}
-This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals.
-The squared exponential kernel produces a smooth functional form which closely matches the behavior of the eigenspectra weights (Fig~\ref{fig:GP_interp}, left panel).
-Then the ($M,M$) covariance matrix is
-\begin{equation}
-\mathbf{\Sigma}_k^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})
-\end{equation}
-which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters, and the conditional dependence on the kernel hyperparameters is implied.
-
-The optimal parameters for the hyperparameters of the squared exponential are given by the multi-variate Gaussian probability distribution, which is called the marginal likelihood function or ``evidence'' for the weights \citep{rasmussen05},
-\begin{equation}
-  p(\mathbf{w}_k^\textrm{grid} | \mathbf{\Sigma}_k^\textrm{grid} ) = \mathcal{N}\left (\mathbf{w}_k^\textrm{grid} \left | \; \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right . \right )
-\end{equation}
-Because $\mathbf{\Sigma}_k^\textrm{grid}$ is defined by $\mathcal{K}_\textrm{SE}$ and its hyperparameters, this likelihood function is equivalent to $p(\mathbf{w}_k^\textrm{grid} |\; a_\textrm{SE}, l_{T_\textrm{eff}},  l_{\log g}, l_{\Z})$.
-By multiplying this with a suitable prior over $a_\textrm{SE}$ and $l$, we arrive at a posterior distribution for the hyperparameters of the Gaussian process.
-For computational efficiency, we optimize the hyperparameters by finding the peak of the posterior distribution with a downhill gradient algorithm and then consider the hyperparameters fixed.
-
-Now the Gaussian process is specified and we can use it to predict the values of the weights at a new set of stellar parameters $\vt_\ast$ by considering them drawn from joint distribution 
+where $\mathsf{I}_{N_{\rm pix}}$ is an $N_{\rm pix} \times N_{\rm pix}$ identity matrix.
+
+To recapitulate, the framework described above can be used to decompose the synthetic spectra in a 
+model library into a principal component basis, allowing us to lossily reconstruct any spectrum in 
+the library as a (weighted) linear combination of $m$ eigenspectra.  The weights corresponding to 
+each eigenspectrum are moderately-smooth scalar functions of the three stellar parameters, 
+$\vt_{\ast}$.  Therefore, to create a spectrum corresponding to an arbitrary set of these 
+parameters that is not in the spectral library, we must interpolate the weights to this new set.  
+In practice, it may be possible to use a traditional scheme like spline interpolation to do this 
+directly.  However, we found that with sensitive spectra (e.g., for Gl~51 the S/N is $>$400), the 
+uncertainty in the interpolated representation of the spectrum can constitute a significant portion 
+of the total uncertainty budget.  This, combined with the under-sampling of the synthetic grid can 
+cause artificial ``peaking" of the posterior near grid points in the synthetic library, because the 
+interpolated spectrum is not as good as the raw spectrum at the grid point.  Even explicitly 
+accounting for interpolation error by doing ``drop-out" interpolation tests and empricially 
+propagating it forward does not relieve this ``peaking" issue.  So instead, we address this problem 
+by employing a Gaussian process to model the interpolation of the eigenspectra weights over 
+$\vt_{\ast}$, thereby encapsulating the range of possible interpolated spectra.
+
+Each weight is modeled independently, so there is a new Gaussian process for each eigenspectrum.  
+For a single eigenspectrum with index $k$, we denote the collection of 
+$w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated for all the spectra in the library as 
+$\mathbf{w}_k^\textrm{grid}$.  The Gaussian process treats $\mathbf{w}_k$ as a collection of random 
+variables drawn from a joint multi-variate Gaussian distribution 
+\citep{rasmussen05},
+\begin{equation}
+  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right ),
+\end{equation}
+with $\mathbf{\Sigma}_k^{\rm grid}$ denoting the covariances.  The kernel that describes the 
+covariance matrix for this distribution is assumed to be a 3-dimensional squared exponential,
+\begin{eqnarray}
+  \mathcal{K}(\vt_i, \vt_j | \phi_{\rm int} ) &=& a_{\rm int}^2 \exp \left [ - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 \, \ell_{T_\textrm{eff}}^2 } \right] \nonumber \\
+  & & \times \exp \left [-\frac{(\log g_i - \log g_j)^2}{2 \, \ell_{\log g}^2} \right ] \\
+  & & \times \exp \left [ -\frac{ (\Z_i - \Z_j)^2}{2 \, \ell_{\Z}^2} \right ] \nonumber,
+\end{eqnarray}
+with hyper-parameters $\phi_{\rm int} = \{a_{\rm int}$, $\ell_{T_{\rm eff}}$, $\ell_{\log g}$, 
+$\ell_{\Z}$\} representing an amplitude and length scale for each dimension of $\vt_{\ast}$.  
+Unlike the \matern\ kernel used in Section~\ref{sec:method} (which produces a more structured 
+behavior reminiscent of the spectral residuals), this squared exponential kernel has a smooth 
+functional form that better represents the behavior of the eigenspectra weights across the library 
+grid, as demonstrated in Figure~\ref{fig:GP_interp}.  The $M\times M$-dimensional covariance matrix 
+is 
+\begin{equation}
+\mathbf{\Sigma}_k^\textrm{grid} = \mathcal{K}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | \, \phi_{\rm int}),
+\end{equation}
+the evaluation of the covariance kernel for all pairings of stellar parameters at library 
+gridpoints.
+
+The optimal values for the hyperparameters $\phi_{\rm int}$ are given by the multi-variate Gaussian 
+probability distribution, called the marginal likelihood function or ``evidence" for the weights 
+\citep{rasmussen05},
+\begin{equation}
+  p(\mathbf{w}_k^\textrm{grid} | \mathbf{\Sigma}_k^\textrm{grid} ) = \mathcal{N}\left (\mathbf{w}_k^\textrm{grid} \left | \; \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right . \right ).
+\end{equation}
+Because $\mathbf{\Sigma}_k^\textrm{grid}$ is defined by the kernel $\mathcal{K}$ and its associated 
+hyperparameters, this likelihood function is equivalent to $p(\mathbf{w}_k^\textrm{grid} |\; 
+a_\textrm{int}, \ell_{T_\textrm{eff}},  \ell_{\log g}, \ell_{\Z})$.  By multiplying this with a 
+suitable prior over $a_\textrm{int}$ and $\ell$, we can construct a posterior distribution for 
+$\phi_{\rm int}$.  For computational efficiency, we optimize the hyperparameters by finding the 
+peak of the posterior distribution with a downhill gradient algorithm and then consider 
+$\phi_{\rm int}$ fixed.
+
+Now the Gaussian process is fully specified and can be used to predict the values of the weights at 
+any arbitrary set of stellar parameters $\vt_\ast$ by considering them drawn from the joint 
+distribution 
 \begin{equation}
@@
   \begin{bmatrix}
-  \mathbf{\Sigma}_k^\textrm{grid} & \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
-  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast)
+  \mathbf{\Sigma}_k^\textrm{grid} & \mathcal{K}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \\
+  \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) & \mathcal{K}(\vt_\ast,\vt_\ast)
   \end{bmatrix}
@@
 \end{equation}
-where $\mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length $M$ column vector where $\mathcal{K}_\textrm{SE}$ is evaluated pairwise with all values in $\{\vt_\ast \}^\textrm{grid}$ and the proposed value of $\vt_\ast$.
-This vector is stacked horizontally next to  $\mathbf{\Sigma}^\textrm{grid}$, and its transpose is stacked vertically underneath $\mathbf{\Sigma}^\textrm{grid}$.
-The combination of all of these products results in an augmented covariance matrix of size ($M + 1, M + 1$).
-The probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution \citep{rasmussen05}
-\begin{equation}
-  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) =  \mathcal{N} \left ( w_k \Bigl | \Bigr . \,  \mu_k , \sigma_k \right )
+where $\mathcal{K}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)$ represents a length-$M$ column vector 
+evaluated pairwise with all values in $\{\vt_\ast \}^\textrm{grid}$ and the proposed value of 
+$\vt_\ast$.  This vector is stacked horizontally next to $\mathbf{\Sigma}^\textrm{grid}$, and its 
+transpose is stacked vertically underneath $\mathbf{\Sigma}^\textrm{grid}$.  The combination of all 
+these products results in an augmented covariance matrix of size ($M + 1, M + 1$).  The probability 
+distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution 
+\citep{rasmussen05}
+\begin{equation}
+  p(w_k |\; \{\vt_\ast \}^\textrm{grid}, \mathbf{w}_k^\textrm{grid}, \vt_\ast) =  \mathcal{N} \left ( w_k \Bigl | \Bigr . \,  \mu_k , \sigma_k \right ),
  \label{eqn:conditional_gp}
@@
 \begin{equation}
- \mu_k =   \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathbf{\Sigma}^{\textrm{grid}, -1} \mathbf{w}_k^\textrm{grid}
-\end{equation}
-\begin{multline}
-  \sigma_k = \mathcal{K}_\textrm{SE}(\vt_\ast,\vt_\ast) - \\
-  \mathcal{K}_\textrm{SE}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) \mathbf{\Sigma}^{\textrm{grid}, -1} \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast)
-\end{multline}
-
-Although this equation is complex, the interpretation is straightforward: the probability distribution of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of $\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared exponential hyperparameters (see Figure~\ref{fig:GP_interp}, right panel).
-
-If we desired an actual value of the interpolated weight, for example to reconstruct a model spectrum, we could simply draw a Gaussian random variable from the probability distribution in Eqn~\ref{eqn:conditional_gp}.
-However, because we now know the probability distribution of the weight as a function of $\vt_\ast$, we can rewrite our data likelihood function (Eqn~\ref{eqn:lnlikelihood}) in such a way that it is possible to analytically marginalize over all possible values of $w_k$, and thus all probable spectral interpolations.
-
-Next, we consider all $m$ of the Gaussian processes describing the weight for each eigenspectra collectively.
-For a given value of $\vt_\ast$, these Gaussian processes deliver a collection of $m$ one-dimensional Gaussian probability distributions.
-Because we have modelled each $w_k$ independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a multidimensional Gaussian with only the diagonal elements filled
-\begin{equation}
-  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | \right . \mathbf{\mu}_\mathbf{w}, \mathbf{\Sigma}_\mathbf{w} \right )
-\end{equation}
-\begin{equation}
-  p(\mathbf{w} |\, \vt_\ast) = \mathcal{N} \left ( \mathbf{w} \left | 
+ \mu_k =   \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) (\mathbf{\Sigma}^{\textrm{grid}})^{-1} \mathbf{w}_k^\textrm{grid}
+\end{equation}
+and 
+\begin{eqnarray}
+  \sigma_k &=& \mathcal{K}(\vt_\ast,\vt_\ast) - \\
+  & & \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) (\mathbf{\Sigma}^{\textrm{grid}})^{-1} \mathcal{K}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \nonumber.
+\end{eqnarray}
+% Should one of these be transposed?
+
+Though the notation is complex, the interpretation is straightforward: the probability distribution 
+of $w_k$ is a univariate Gaussian distribution whose mean and variance are a function of 
+$\vt_\ast$, conditional upon the (fixed) values of $\mathbf{w}_k^\textrm{grid}$ and the squared 
+exponential hyperparameters (an example is shown in Figure~\ref{fig:GP_interp}, right panel).
+
+If we desired an actual value of the interpolated weight, for example to reconstruct a model spectrum, we could simply draw a Gaussian random variable from the probability distribution in 
+Eq.~(\ref{eqn:conditional_gp}).  However, because we now know the probability distribution of the 
+weight as a function of $\vt_\ast$, we can rewrite our data likelihood function 
+(Eq.~\ref{eqn:lnlikelihood}) in such a way that it is possible to analytically marginalize over all 
+possible values of $w_k$, and thus all probable spectral interpolations.
+
+Next, we consider the combined $m$ Gaussian processes that describe the weights for each 
+eigenspectrum collectively.  For a given $\vt_\ast$, these Gaussian processes deliver a collection 
+of $m$ one-dimensional Gaussian probability distributions.  Because we have modeled each $w_k$ 
+independently, the joint probability of the vector of weights $\mathbf{w}$ is simply a 
+multidimensional Gaussian matrix with only the diagonal elements populated,
+\begin{eqnarray} \label{eqn:weight_conditional}
+  p(\mathbf{w} |\, \vt_\ast) &=& \mathcal{N} \left ( \mathbf{w} \left | \right . \mathbf{\mu}_\mathbf{w}, \mathbf{\Sigma}_\mathbf{w} \right ) \\ \nonumber
+  & = & \mathcal{N} \left ( \mathbf{w} \left | 
   \begin{bmatrix}
@@
   \right .
-  \right)
-  \label{eqn:weight_conditional}
-\end{equation}
-
-So far, we have presented the reconstruction of a spectrum as a linear combination of the eigenspectra of the synthetic library (Figure~\ref{fig:pca_reconstruct}).
-In practice, however, the reconstructed spectrum must be further post-processed to match the observational parameters ($\vt_{\rm obs}$) of the data spectrum by applying convolutions for instrumental broadening, stellar rotation broadening, and Doppler shift, and then lastly re-sampling the spectrum to the exact pixels of the CCD (Section~\ref{subsec:postprocess}).
-Fortunately, because convolution is a linear operation, we can first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the reconstructed spectrum as a linear combination of these modified eigenspectra without loss of information.
-Unfortunately, the Doppler shift and resampling operations are not linear operations, and there will be some loss of information when trying to approximate them in this manner.
-However, we find that in practice when the synthetic spectra are oversampled relative to the instrument resolution by a reasonable factor, the flux error due to resampling is smaller than 0.2\% across all pixels , and thus any effect is negligible.
-For notational compactness, we let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and $\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra, with an implied dependence on the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance parameters $(\phi_\mathsf{P})$.
-Now, the model spectrum is specified conditional on the vector of eigenspectra weights as
+  \right).
+\end{eqnarray}
+
+Up until this point, we have described the reconstruction of a spectrum as a linear combination of 
+the eigenspectra that characterize the synthetic library (Figure~\ref{fig:pca_reconstruct}).  But 
+in practice, that reconstructed spectrum must be further post-processed as detailed in 
+Section~\ref{subsec:postprocess}.  Fortunately, because convolution is a linear operation, we can 
+first post-process the raw eigenspectra according to $\vt_{\rm obs}$, and then represent the 
+reconstructed spectrum as a linear combination of these modified eigenspectra without loss of 
+information.  Unfortunately, the Doppler shift and resampling operations are not linear operations, 
+and there will be some loss of information when trying to approximate them in this manner.  
+However, we find that in practice when the synthetic spectra are oversampled relative to the 
+instrument resolution by a reasonable factor, the flux error due to resampling is smaller than 
+0.2\%\ across all pixels, and thus any effect of that information loss is negligible.  For 
+notational compactness, we let $\widetilde{\xi}_\mu$, $\widetilde{\xi}_\sigma$, and 
+$\widetilde{\mathbf{\Xi}}$ represent the post-processed eigenspectra, with an implied dependence on 
+the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance 
+parameters $(\phi_\mathsf{P})$.  Now, the model spectrum is specified conditional on the vector of 
+eigenspectra weights as 
 \begin{equation}
@@
 \begin{equation}
-  \mathbf{X} = \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}}
+  \mathbf{X} = \widetilde{\xi}_\sigma I_{N_{\rm pix}} \widetilde{\mathbf{\Xi}}.
 \end{equation}
@@
 \end{equation}
-and the likelihood function (Eqn~\ref{eqn:likelihood}) is specified conditional on the weights
+and the likelihood function (Eq.~\ref{eqn:likelihood}) is specified conditional on the weights,
 \begin{equation}
   p( \mathsf{D} |\, \mathsf{M}) = p( \mathsf{D} |\, \mathbf{w}) = 
-  \mathcal{N} \left ( \mathsf{D} |\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w} , \mathsf{C} \right )
+  \mathcal{N} \left ( \mathsf{D} |\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w} , \mathsf{C} \right ).
   \label{eqn:likelihood_conditional}
@@
 
-The final task of designing the spectral emulator is to combine this data likelihood function with the posterior distribution of the eigenspectra weights (Eqn~\ref{eqn:weight_conditional}) and then marginalize over the weights
+The final task of designing the spectral emulator is to combine this data likelihood function with the posterior distribution of the eigenspectra weights (Eq.~\ref{eqn:weight_conditional}) and then marginalize over the weights
 \begin{equation}
@@
 \end{equation}
-such that we are left with a modified posterior distribution of the data that incorporates the range of probable interpolation values for the model.
-To perform this multidimensional integral, we use a convenient lemma found in \citet[Appendix A]{gelman13}: if the probability distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in Eqns~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the marginal distribution (Eqn~\ref{eqn:marginal}) is
-\begin{equation}
-  p(\mathsf{D} | \vt_\ast, \vt_\textrm{obs}, \mathbf{\Phi}) = \mathcal{N} \left ( \mathsf{D} \bigl | \bigr .\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}, \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C} \right)
-\end{equation}
-where the dependence on the model parameters is now made explicit.
-We can couch this modified likelihood function in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
+such that we are left with a modified posterior distribution of the data that incorporates the 
+range of probable interpolation values for the model.  To perform this multidimensional integral, 
+we use a convenient lemma found in \citet[their Appendix A]{gelman13}: if the probability 
+distributions of $\mathbf{w}$ and $\mathsf{D} | \mathbf{w}$ are specified conditionally as in 
+Eq.~\ref{eqn:weight_conditional} and \ref{eqn:likelihood_conditional}, respectively, then the 
+marginal distribution (Eq.~\ref{eqn:marginal}) is
+\begin{equation}
+  p(\mathsf{D} | \vt_\ast, \vt_\textrm{obs}, \mathbf{\Phi}) = \mathcal{N} \left ( \mathsf{D} \bigl | \bigr .\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{\mu}_\mathbf{w}, \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C} \right),
+\end{equation}
+where the dependence on the model parameters is now made explicit.  We can couch this modified 
+likelihood function in the form of Eqn~\ref{eqn:lnlikelihood} by rewriting
 \begin{equation}
@@
 \end{equation}
-where $\mathsf{M}^\prime$ can be thought of as the ``mean model spectrum'' given the model parameters, and the covariance matrix has been modified to account for the various probable manifestations of the model spectrum about the mean spectrum.
+where $\mathsf{M}^\prime$ can be thought of as the ``mean model spectrum" given the model 
+parameters, and the covariance matrix has been modified to account for the various probable 
+manifestations of the model spectrum about that mean spectrum.
 

### 2014-11-03 tex/paper.tex
just kidding: now its the last time!

@@
 We can now compute the covariance matrix employed in the likelihood calculation 
-(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix, 
-an interpolation error kernel ($\mathcal{K}^{\rm int}$; see the Appendix), and the global and local 
-kernels discussed above, 
-\begin{equation}
-\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + \mathcal{K}_{ij}^{\rm int} + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
+(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
+and the global and local kernels discussed above,\footnote{See the Appendix for a modification of 
+the likelihood function that incorporates interpolation uncertainty.} 
+\begin{equation}
+\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
                   \Klocal_{ij}(\vp_{{\mathsf C}, l}), 

### 2014-11-04 tex/paper.tex
More figure tweaks

@@
 \startdata
-$T_{\rm eff}$ (K) & 6424     & 19                      &  6354   & 39        \\
+$T_{\rm eff}$ (K) & 6424     & 19                      &  6349   & 20        \\
 $\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
-$\Z$          & -0.26    & 0.01                    & -0.30  & 0.02     \\
+$\Z$          & -0.26    & 0.01                    & -0.31  & 0.02     \\
 $v \sin i$ ($\kms$)\tablenotemark{a}   & 4.47 & 0.06       &  5.09   & 0.06      \\
 $v_z$ ($\kms$)        & -4.60    & 0.02                    & -4.86   & 0.02      \\
-$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -12.717 & 0.0007 \\
+$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -12.717 & 0.001 \\
 $A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
@@
 a subset of the parameter-space most relevant for M dwarfs.  (\emph{bottom}) The original synthetic 
-spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear 
+spectrum from the \textsc{Phoenix} library ($\vt_\ast = [T_\textrm{eff} = 3000\,\textrm{K}, \log g =
+  5.0\,\textrm{dex}, \Z = 0.0\, \textrm{dex}]$) compared with a spectrum reconstructed from a linear 
 combination of the derived eigenspectra using Eqn~\ref{eqn:reconstruct} (with the weights $w_k$ 
@@
 \end{eqnarray}
-% Should one of these be transposed?
 

### 2014-11-04 Starfish/parallel.py
More figure tweaks

@@ -704,7 +704,7 @@ def main():
     # import sys; sys.exit()
 
     mySampler = StellarSampler(pconns=pconns, starting_param_dict=stellar_Starting, cov=stellar_MH_cov,
-                               outdir=outdir, debug=True, fix_logg=config["fix_logg"])
+                               outdir=outdir, debug=True, fix_logg=fix_logg)
 
     mySampler.run_mcmc(mySampler.p0, config['burn_in'])
     #mySampler.reset()

### 2014-11-04 hdfutils.py

@@ -50,6 +50,8 @@ parser.add_argument("--cov", action="store_true", help="Estimate the covariance 
 parser.add_argument("--ndim", type=int, help="How many dimensions to use for estimating the 'optimal jump'.")
 parser.add_argument("--paper", action="store_true", help="Change the figure plotting options appropriate for the "
                                                          "paper.")
+parser.add_argument("--replicate", action="store_true", help="Repeat the (converged) chains for the paper plot to "
+                                                             "fill in the appearance of the histogram.")
 
 args = parser.parse_args()
 
@@ -383,6 +385,11 @@ def plot_paper(flatchain, base=args.outdir, triangle_plot=args.triangle, chain_p
 
     params = flatchain.param_tuple
     samples = flatchain.samples
+
+    if args.replicate:
+        #Concatenate the samples a few times
+        samples = np.vstack([samples for i in range(30)])
+
     labels = [label_dict.get(key, "unknown") for key in params]
 
     K = len(labels)
@@ -392,7 +399,7 @@ def plot_paper(flatchain, base=args.outdir, triangle_plot=args.triangle, chain_p
                              show_titles=False, title_args={"fontsize": 8}, plot_contours=True,
                              plot_datapoints=False, fig=fig)
     if K == 3:
-        figure.subplots_adjust(left=0.13, right=0.87, top=0.95, bottom=0.15)
+        figure.subplots_adjust(left=0.13, right=0.87, top=0.95, bottom=0.16)
     if K == 4:
         figure.subplots_adjust(left=0.13, right=0.87, top=0.95, bottom=0.15)
 
@@ -410,9 +417,15 @@ def plot_paper(flatchain, base=args.outdir, triangle_plot=args.triangle, chain_p
     if K == 3:
         #Yaxis
         for ax in axes[:, 0]:
-            ax.yaxis.set_label_coords(-0.38, 0.5)
+            ax.yaxis.set_label_coords(-0.40, 0.5)
         for ax in axes[-1, :]:
-            ax.xaxis.set_label_coords(0.5, -0.34)
+            ax.xaxis.set_label_coords(0.5, -0.39)
+
+        for ax in axes[:, 1]:
+            ax.xaxis.set_major_locator(MaxNLocator(nbins=4, prune='lower'))
+
+        axes[1, 0].yaxis.set_major_locator(MaxNLocator(nbins=5))
+
 
     figure.savefig(base + flatchain.id + format)
 

### 2014-11-08 Starfish/__init__.py
trying requirements file

@@ -1,2 +1,2 @@
-__version__ = '0.1'
+__version__ = '1.0'
 __all__ = ["spectrum", "model", "grid_tools", "constants", "covariance", "utils", "emulator", "em_cov"]

### 2014-12-05 tex/paper.tex
Open sourcing
Travis
trying requirements file
Kurucz fig updated
Partial comments addressed

@@
 \newcommand{\KK}{\mathcal{K}}
-\newcommand{\Kglobal}{\KK^{\textrm{g}}}
-\newcommand{\Klocal}{\KK^l}
+\newcommand{\Kglobal}{\KK^{\textrm{G}}}
+\newcommand{\Klocal}{\KK^{\textrm{L}}}
 
@@
 
-\shorttitle{Robust Spectroscopic Inference with Imperfect Models}
+\shorttitle{Robust Spectroscopic Inference}
 \shortauthors{Czekala et al.}
@@
 
-\slugcomment{draft: \today{}}
-
 \title{Robust Spectroscopic Inference with Imperfect Models}
 
-\author{Ian Czekala, Sean M.~Andrews, et al.}
-\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
+\newcommand{\harvard}{1}
+\newcommand{\nyu}{2}
+\newcommand{\mpia}{3}
+\newcommand{\cds}{4}
+
+\author{%
+  Ian~Czekala\altaffilmark{\harvard }, 
+  Sean~M.~Andrews\altaffilmark{\harvard },
+  Kaisey~S.~Mandel\altaffilmark{\harvard },
+  David~W.~Hogg\altaffilmark{\nyu },
+  and Gregory~M.~Green\altaffilmark{\harvard}
+}
+
 \email{iczekala@cfa.harvard.edu}
+\altaffiltext{\harvard}  {Harvard-Smithsonian Center for Astrophysics, 
+			 60 Garden Street, Cambridge, MA 02138}
+\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
+                          Department of Physics, New York University,
+                          4 Washington Place, New York, NY, 10003, USA}
+%\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
+%			  K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+%\altaffiltext{\cds}      {Center for Data Science,
+%                          New York University,
+%                          4 Washington Place, New York, NY, 10003, USA}
 
@@
 between adjacent datapoints (pixels) into the residual spectrum.  In the limit 
-of high signal-to noise data with large spectral range that is common for 
+of high signal-to-noise data with large spectral range that is common for 
 stellar parameter estimation, that covariant structure can bias the parameter 
@@
 \end{abstract}
-\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}
+\keywords{stars: fundamental parameters --- techniques: spectroscopic --- stars: late-type --- 
+  stars: statistics --- methods: data analysis --- methods: statistical}
 
@@
 
-All astronomers recognize that spectroscopy offers a wealth of information that 
-can help characterize the properties of the observing target.  In the context 
-of stellar astrophysics, spectroscopy plays many fundamental roles.  The 
-relative strengths and widths of stellar absorption lines provide access to key 
-parameters like effective temperature ($T_{\rm eff}$) and surface gravity 
-($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to 
-estimate the masses and ages so crucial to understanding stellar evolution, as 
-well as individual elemental (and molecular) abundances or the collective 
-``metallicity" (typically parameterized as [Fe/H]), facilitating study of the 
-chemical hallmarks of different stellar populations.  With sufficient spectral 
-resolution, the velocity content of a spectrum can convey crucial information 
-about stellar rotation ($v \sin i$) and kinematics (e.g., association with a 
-cluster or companion through the radial velocity, $v_z$).  While many fields 
-benefit from the spectroscopic measurements of these stellar properties, there 
-is acute interest in the exoplanet community.  There, all estimates of the 
-planet properties are made {\it relative} to the host properties (e.g., the 
-planet-to-host mass or radius {\it ratio} is constrained with the radial 
-velocity or transit techniques, respectively).  Moreover, essential clues to 
-the planet formation process are encapsulated in the dependences of planet 
-frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity 
+All astronomers recognize that spectroscopy offers a wealth of information that can help
+characterize the properties of the observing target.  In the context of stellar astrophysics,
+spectroscopy plays many fundamental roles.  The relative strengths and widths of stellar absorption
+lines provide access to key parameters like effective temperature ($T_{\rm eff}$) and surface
+gravity ($\log g$), enabling model comparisons in the Hertzsprung-Russell diagram to estimate the
+masses and ages so crucial to understanding stellar evolution, as well as individual elemental (and
+molecular) abundances or the collective ``metallicity" (typically parameterized as [Fe/H]),
+facilitating study of the chemical hallmarks of different stellar populations.  With sufficient
+spectral resolution, the velocity content of a spectrum can convey crucial information about stellar
+rotation ($v \sin i$) and kinematics (e.g., association with a cluster or companion through the
+radial velocity, $v_z$).  While many fields benefit from the spectroscopic measurements of these
+stellar properties, they are of acute interest to the exoplanet community.  There, all estimates of
+the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host mass
+function or radius {\it ratio} is constrained with the radial velocity or transit techniques,
+respectively).  Moreover, essential clues to the planet formation process are encapsulated in the
+dependences of planet frequency on host mass \citep[e.g.,][]{johnson07,howard10} and metallicity
 \citep[e.g.,][]{fischer05,buchhave14}.
 
-That said, the robust and quantitative extraction of physical (or empirical) parameters from an 
+The robust and quantitative extraction of physical (or empirical) parameters from an 
 observed spectrum can be an extraordinary challenge.  Stellar models are ultimately required as a 
@@
 limited swath of parameter-space.  A second technique exploits the cross-correlation of a spectrum
-with a suite of model templates to optimize a set of parameters, usually with some weighting 
-applied to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the 
-speed advantage is maintained (perhaps enhanced) and more data content is used (particularly in the 
+with a suite of model templates to optimize a set of parameters, usually with some weighting applied
+to specific spectral regions \citep[e.g., {\tt SPC};][]{buchhave12}.  In this case, the speed
+advantage is maintained (perhaps enhanced) and more data content is used (particularly in the
 spectral dimension), thereby achieving higher precision even for data with comparatively low
-sensitivity; the disadvantage is that the model quality and parameter inferences are assessed in an 
-empirical, rather than probabilistic, framework.  A third approach employs a direct, pixel-by-pixel 
-comparison between model and data, and is usually associated with a spectral synthesis back-end 
-(e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric flexibility 
-(e.g., one can fit for arbitrary abundances or structures) and a proper inference framework 
-\citep[usually a least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,
-schoenrich13}, but it is computationally expensive and can be biased by systematics 
-\citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a pre-selected 
-subsample of the data.
+signal-to-noise; the disadvantage is that the model quality and parameter inferences are assessed in
+an heuristic, rather than probabilistic, framework, making it difficult to quantify uncertainty in
+the stellar parameters without a well-defined model of how the data is generated.  A third approach
+employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with a
+spectral synthesis back-end (e.g., {\tt SME}; \citealt{valenti96}).  This technique has the benefits
+of parametric flexibility (e.g., one can fit for arbitrary abundances or structures) and a proper
+inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian
+format;][]{shkedy07, schoenrich13}, but it is computationally expensive and can be biased by
+systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a
+pre-selected subsample of the data.
 
@@
 function and a parametric treatment of the covariances between pixel residuals (Section
-\ref{subsec:likelihood}).  That process is iterated under the guise of hierarchical Markov Chain 
-Monte Carlo (MCMC) simulations to numerically explore the posterior probability density of the 
-model conditioned on the data, and thereby to determine constraints on the parameters of interest 
-(Section \ref{subsec:MCMC}).  Along the way, these procedures are illustrated with observations of 
-the high resolution optical spectrum from a nearby F star.  That specific application, along with 
-some alternative demonstrations of the method, are discussed in more detail in Section 
-\ref{sec:examples}.
+\ref{subsec:likelihood}).  That process is iterated using Markov Chain Monte Carlo (MCMC)
+simulations in a multi-stage Gibbs sampler to numerically explore the posterior probability density
+of the model conditioned on the data, and thereby to determine constraints on the parameters of
+interest (Section \ref{subsec:MCMC}).  Along the way, these procedures are illustrated with
+observations of the high resolution optical spectrum from a nearby F star.  That specific
+application, along with some alternative demonstrations of the method, are discussed in more detail
+in Section \ref{sec:examples}.
 
@@
 interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency
-is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces a 
-non-negligible level of uncertainty.  These interpolation uncertainties can be empirically 
-estimated by performing the operation in Eq.~\ref{eqn:interp} across a $\{\vt_{\ast}\}^{\rm grid}$ 
-grid point, and then comparing the interpolated spectrum with the corresponding library spectrum 
-\citep[see also][]{husser12}.  To suitably propagate the interpolation uncertainties into the 
-likelihood calculations, we employ a Bayesian emulation technique that is described in detail in 
-Appendix \ref{sec:Appendix}.
+is critical.  In practice, a simple linear interpolation from nearby grid points in the spectral
+library is suitably fast, but introduces a non-negligible level of interpolation error. Although
+generally small, this error can significantly bias the inference of stellar parameters
+when the signal-to-noise of the data is sufficiently high. To account for the uncertainty in a given
+interpolated spectrum, we design a Bayesian spectral emulator, which we describe in detail in
+Appendix \ref{sec:Appendix}. Rather than delivering a single interpolated spectrum for a given
+$\vt_\ast$, the emulator delivers a probability distribution over a set of possible interpolated
+spectra. By marginalizing over all of the possible interpolated models, we can modify the likelihood
+function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation uncertainty. For
+simplicity, however, we now proceed as if we used linear interpolation from a synthetic library to
+generate model spectra and that the error introduced by this process is negligible, since the
+following concepts are applicable regardless of the interpolation method.
 
@@
 that incorporate dynamical effects, geometry, and the relative location of the target, while the 
-second employs a suite of nuisance (hyper-)parameters ($\vp$) designed to mitigate an imperfect 
+second employs a suite of nuisance parameters ($\vp$) designed to mitigate an imperfect 
 data calibration.
@@
 
-% NEEDS UPDATED FIGURE
 \begin{figure}[b!]
@@
 spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
-shapes.  Chebyshev polynomials ({\it middle, top}) are designed to account for that mismatch ({\it 
-middle, bottom}); rather, in practice, the model is scaled (multiplied) by those polynomials ({\it 
-bottom}).  \label{fig:chebyshev}}
+shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
+adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
+by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
+(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
 \end{figure}
@@
 
-We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
-the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
+We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internally}
+to the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
 in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as
@@
 possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance
-parameters and then marginalized.  This way, the inferences for the stellar parameters will 
-properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
-practice, this is achieved by distorting segments of the model spectrum with polynomials, 
-$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
+parameters and then marginalized.  This way, the inferences for the stellar parameters will properly
+account for the underlying uncertainty in the calibration of the spectral shape.  In practice, this
+is achieved by distorting segments of the model spectrum with polynomials, $\mathsf{P}$
+\citep[e.g.,][]{eisenstein06,koleva09}. Figure \ref{fig:chebyshev} offers a practical demonstration
+of how these nuisance parameters are applied to the model. For data that has $N_{\rm ord}$ spectral
 orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
@@
 where $T^{(n)}$ is an $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
-coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
+coefficients are considered a set of nuisance parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
 \ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
@@
 multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
-$c^{(0)}$ in other orders to be different.  Figure \ref{fig:chebyshev} offers a practical 
-demonstration of how these nuisance parameters are applied to the model.
+$c^{(0)}$ in other orders to be different.
 
@@
 \delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
-the Kronecker delta function, and Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
-\begin{equation}
-\ln{p(\vD | \vM)} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
+the Kronecker delta function. Under the assumption that the covariance matrix is held constant, Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
+\begin{equation}
+  \ln{p(\vD | \vM)} - \textrm{constant} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
 \label{eqn:chisq}
@@
 The autocorrelation of the residual spectrum.  Notice the substantial autocorrelation signal for 
-offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not described well by 
+offsets of $\lesssim 8$ pixels, demonstrating clearly that the residuals are not well described by 
 white (Poisson) noise alone. \label{fig:class0}}
@@
 For a well-designed spectrograph and sufficiently accurate model, this {\it global} (i.e., present 
-throughout the spectrum) covariant structure should have a relatively low amplitude and small 
+throughout the spectrum) covariance should have a relatively low amplitude and small 
 correlation length.  To describe that structure, we use a stationary covariance kernel (or radial 
@@
 \end{equation}
-where the (hyper-)parameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
+where the hyperparameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
 ($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
-enables computational expendiency, we employ a Hann window function
-\begin{equation}
-  w_{ij}^{\rm g} \,(r_0) = \left \{ 
+enables computational expediency, we employ a Hann window function
+\begin{equation}
+  w_{ij}^{\rm G} \,(r_0) = \left \{ 
     \begin{array}{cc}
@@
 \begin{equation} \label{eqn:kregion}
-  \mathcal{K}^{l,k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
+  \mathcal{K}^{\textrm{L},k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
 \end{equation}
@@
 \begin{equation} \label{eqn:klocal}
-\Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{l,k}_{ij},
-\end{equation}
-with a set of (hyper-)parameters $\vp_{{\mathsf C},l} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
+  \Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{\textrm{L},k}_{ij},
+\end{equation}
+with a set of hyperparameters $\vp_{{\mathsf C},\textrm{L}} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
 again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
@@
 \sigma_i$) poorly reproduces both the scale and structure of the residual spectrum.  The addition 
-of a global kernel ($\mathcal{K}^{\rm g}$) more closely approximates the structure and amplitude of 
+of a global kernel ($\mathcal{K}^\textrm{G}$) more closely approximates the structure and amplitude of 
 the residuals, but misses the outlier line at 5202.2\,\AA.  Including a local kernel 
-($\mathcal{K}^l$) at that location results in a covariance structure that does an excellent job 
+($\mathcal{K}^\textrm{L}$) at that location results in a covariance structure that does an excellent job 
 reproducing all the key residual features.  \label{fig:matrix}}
@@
 (Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel noise matrix 
-and the global and local kernels discussed above,\footnote{See the Appendix for a modification of 
-the likelihood function that incorporates interpolation uncertainty.} 
+and the global and local kernels discussed above, 
 \begin{equation}
@@
 \end{equation}
-with (hyper-)parameters $\cov = [\vp_{{\mathsf C}, g}, \vp_{{\mathsf C}, l}]$.  The factor $b$ is 
+with hyperparameters $\cov = [\vp_{{\mathsf C}, \textrm{G}}, \vp_{{\mathsf C}, \textrm{L}}]$.  The factor $b$ is 
 a parameter that scales up the Poisson noise in each pixel by a constant factor to account for 
@@
 there are $N_{\rm loc}$ local covariance patches (see Section \ref{subsec:MCMC} on how this is 
-determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance (hyper-)parameters, 
+determined), then there are $4N_{\rm loc}+2$ elements in the set of covariance hyperparameters, 
 $\cov$.
@@
 since such behavior is better treated by the global kernel.
+
+\begin{figure}[!htb]
+  \includegraphics{figs/PGM.pdf}
+  \figcaption{ Probabilistic graphical model.
+\label{fig:PGM}}
+\end{figure}
 
@@
 \startdata
-$T_{\rm eff}$ (K) & 6424     & 19                      &  6349   & 20        \\
+$T_{\rm eff}$ (K) & 6426     & 21                      &  6349   & 20        \\
 $\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
@@
 $v_z$ ($\kms$)        & -4.60    & 0.02                    & -4.86   & 0.02      \\
-$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -12.717 & 0.001 \\
+$\log \Omega$\tablenotemark{b} & -12.7182 & 0.0004 & -12.717 & 0.001 \\
 $A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
@@
 \startdata
-$T_{\rm eff}$ (K) & 3071  & 28 \\
+$T_{\rm eff}$ (K) & 3066  & 23 \\
 $\log g$ & 5.19 & 0.04 \\
-$\Z$ & 0.14 & 0.07 \\
+$\Z$ & 0.13 & 0.06 \\
 $v \sin i$ ($\kms$) & \multicolumn{2}{c}{$\leq 15.0$} \\
-$v_z$ ($\kms$) & 11.6 & 1.3 \\
-$\log \Omega$ (sr) & -13.481  & 0.001 \\
+$v_z$ ($\kms$) & 11.7 & 1.3 \\
+$\log \Omega$ (sr) & -13.481  & 0.0003 \\
 $A_V$ & 0 & 0 (fixed)
@@
 We have presented an alternative approach to dealing with this fundamental issue, grounded in a 
-generative, hierarchical Bayesian framework.  The method advocated here constructs a sophisticated 
+generative Bayesian framework.  The method advocated here constructs a sophisticated 
 likelihood function, employing a non-trivial covariance matrix to treat the correlated 
@@
 (Sect.~\ref{sec:examples}).  The source code developed here is open and freely available for use: 
-see \url{http://github.com/iancze/Starfish}.
+see \url{http://iancze.github.io/Starfish}.
 
@@
 existing models, creating a new library of semi-empirical spectral models.  This could be
-accomplished by linking the parameters of the covariance kernels derived from fitting many spectra 
-in a hierarchical Bayesian model, which would add confidence to the assessment that certain 
-spectral features are {\it systematic} outliers and offer general quantitative guidance on how to 
-weight them in the likelihood calculation.  Rather than simply assembling an empirical spectral 
-library using only observations, this combined machine-learning approach would naturally provide a 
-physical anchoring for the key physical parameters, since they are reflected in the spectra based 
-on the physical assumptions in the original models.  This kind of large-scale analysis holds great 
-promise in the (ongoing) era of large, homogeneous high resolution spectroscopic datasets (e.g., 
-like those being collected in programs like the APOGEE and HERMES surveys; \citealt{nidever12,
-zucker12}), since they provide enormous leverage for identifying and improving the underlying model 
-systematics. \\
-
-\acknowledgments  The authors would like to acknowledge the following people for many 
-extraordinarily helpful discussions and key insights: Gregory M.~Green, Kaisey Mandel, Daniel 
-Foreman-Mackey, John Johnson and the ExoLab, Daniel Eisenstein, Rebekah Dawson, Tom Loredo and the 
-ExoStat group, Allyson Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship 
-and the Smithsonian Institution.  Figures \ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, 
-and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py} \citep{foreman-mackey14}.  
-This research made extensive use of Astropy \citep{astropy13} and the Julia programming language 
-\citep{julia12}.
+accomplished by linking the parameters of the covariance kernels while fitting many 
+stars of similar spectral type in a hierarchical Bayesian model, which would add confidence to
+the assessment that certain spectral features are {\it systematic} outliers and offer general
+quantitative guidance on how to weight them in the likelihood calculation.  Rather than simply
+assembling an empirical spectral library using only observations, this combined machine-learning
+approach would naturally provide a physical anchoring for the key physical parameters, since they
+are reflected in the spectra based on the physical assumptions in the original models.  This kind of
+large-scale analysis holds great promise in the (ongoing) era of large, homogeneous high resolution
+spectroscopic datasets (e.g., like those being collected in programs like the APOGEE and HERMES
+surveys; \citealt{nidever12, zucker12}), since they provide enormous leverage for identifying and
+improving the underlying model systematics. \\
+
+\acknowledgments  The authors would like to acknowledge the following people
+for many extraordinarily helpful discussions and key insights: Daniel
+Foreman-Mackey, Guillermo Torres, David Latham, John Johnson and the ExoLab,
+Daniel Eisenstein, Rebekah Dawson, Tom Loredo and the ExoStat group, Allyson
+Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship and
+the Smithsonian Institution.  Figure~\ref{fig:PGM} was generated with
+\texttt{Daft}\footnote{\url{http://daft-pgm.org/}} and Figures
+\ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and
+\ref{fig:Gl51_posterior} were generated with \texttt{triangle.py}
+\citep{foreman-mackey14}.  This research made extensive use of Astropy
+\citep{astropy13} and the Julia programming language \citep{julia12}.
 
@@
 parameters under consideration in the library and $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single 
-set of those parameters drawn from this collection.  In effect, all library spectra are 
+set of those parameters drawn from this collection. Both $\xi_\mu$ and $\xi_\sigma$ are
+vectors with length $N_\textrm{pix}$, the same size as any
+raw synthetic spectrum ($f_\lambda$). In effect, all library spectra are 
 standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
@@
 The eigenspectra are computed from this standardized grid using principal component analysis
-\citep[PCA;][]{ivezic13}.  We decided to truncate our basis to the first $m$ eigenspectra, where 
-$m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid 
-to better than 2\% accuracy for all pixels (the mean pixel error is generally much smaller than 
-this, $\lesssim 0.5\%$).  We denote a single $N_{\rm pix}$-dimensional eigenspectrum as $\xi_k$, 
-with principal component index $k = \{1, 2, \ldots, m\}$.  As an example, the eigenspectra basis 
-computed for Gl~51 using the \textsc{Phoenix} library is shown in the top panel of 
-Figure~\ref{fig:pca_reconstruct}.
+\citep[PCA;][]{ivezic13}. Each eigenspectrum is a vector with length $N_\textrm{pix}$, denoted as
+$\xi_k$, where $k$ is the principal component index $k = \{1, 2, \ldots, m\}$. We decided to
+truncate our basis to the first $m$ eigenspectra, where $m$ is decided by the minimum number of
+eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all
+pixels (the typical error for any given pixel is generally much smaller than this, $\lesssim
+0.5\%$). As an example, the eigenspectra basis computed for Gl~51 using the \textsc{Phoenix} library
+is shown in the top panel of Figure~\ref{fig:pca_reconstruct}.
 
@@
   \includegraphics{figs/GP_right.pdf}
-  \figcaption{The Gaussian process modelling of the principal component weights for the Gl51 \textsc{Phoenix} spectral library.
-(\emph{left}) the blue dots mark the weights $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ computed at a one-dimensional slice of the spectral library for grid points with $\log g = 5.0$, $\Z = 0.0$ and various values of $T_\textrm{eff}$.
-In reality, the weights are a three-dimensional function of $\vt_\ast$.
-The thin blue lines show 50 random draws of possible functional forms described by the Gaussian process, constrained to pass through the grid weights.
-(\emph{inset}) a zoomed portion showing the scatter in the possible functional forms.
-The black vertical line represents a slice through the scatter of the predicted weight value at $\vt_\ast = [T_\textrm{eff} = 3150\,\textrm{K}, \log g = 5.0\,\textrm{dex}, \Z = 0.0\, \textrm{dex}]$.
-(\emph{right}) The posterior probability of $w_k$ at this value of $\vt_\ast$ is completely described by Eqn~\ref{eqn:conditional_gp}, allowing us to analytically marginalize over all probable values of the weights, and thus marginalize over all probable spectral interpolations.
+  \figcaption{The Gaussian process modelling of the principal component weights for the Gl51
+    \textsc{Phoenix} spectral library. (\emph{left}) the blue dots mark the weights
+    $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ computed at a one-dimensional
+    slice of the spectral library for grid points with $\log g = 5.0$, $\Z = 0.0$ and various values
+    of $T_\textrm{eff}$. In reality, the weights are a three-dimensional function of $\vt_\ast$. The
+    thin blue lines show 50 random draws of possible functional forms described by the Gaussian
+    process, constrained to pass through the grid weights. (\emph{inset}) a zoomed portion showing
+    the scatter in the possible functional forms. The black vertical line represents a slice through
+    the scatter of the predicted weight value at $\vt_\ast = [T_\textrm{eff} = 3150\,\textrm{K},
+    \log g = 5.0\,\textrm{dex}, \Z = 0.0\, \textrm{dex}]$. (\emph{right}) The posterior predictive
+    probability of $w_k$ at this value of $\vt_\ast$ is completely described by
+    Eqn~\ref{eqn:conditional_gp}, allowing us to analytically marginalize over all probable values
+    of the weights, and thus marginalize over all probable spectral interpolations.
 \label{fig:GP_interp}}
@@
 interpolated spectrum is not as good as the raw spectrum at the grid point.  Even explicitly 
-accounting for interpolation error by doing ``drop-out" interpolation tests and empricially 
+accounting for interpolation error by doing ``drop-out" interpolation tests and empirically 
 propagating it forward does not relieve this ``peaking" issue.  So instead, we address this problem 
@@
 \end{eqnarray}
-with hyper-parameters $\phi_{\rm int} = \{a_{\rm int}$, $\ell_{T_{\rm eff}}$, $\ell_{\log g}$, 
+with hyperparameters $\phi_{\rm int} = \{a_{\rm int}$, $\ell_{T_{\rm eff}}$, $\ell_{\log g}$, 
 $\ell_{\Z}$\} representing an amplitude and length scale for each dimension of $\vt_{\ast}$.  
@@
 transpose is stacked vertically underneath $\mathbf{\Sigma}^\textrm{grid}$.  The combination of all
-these products results in an augmented covariance matrix of size ($M + 1, M + 1$).  The probability 
-distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian distribution 
+these products results in an augmented covariance matrix of size ($M + 1, M + 1$).  The posterior
+predictive probability distribution of $w_k(\vt_\ast)$ is given by the conditional Gaussian
+distribution 
 \citep{rasmussen05}
@@
 \begin{equation}
- \mu_k =   \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) (\mathbf{\Sigma}^{\textrm{grid}})^{-1} \mathbf{w}_k^\textrm{grid}
+ \mu_k(\vt_\ast) =   \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) (\mathbf{\Sigma}^{\textrm{grid}})^{-1} \mathbf{w}_k^\textrm{grid}
 \end{equation}
@@
 \begin{eqnarray}
-  \sigma_k &=& \mathcal{K}(\vt_\ast,\vt_\ast) - \\
+  \sigma_k(\vt_\ast) &=& \mathcal{K}(\vt_\ast,\vt_\ast) - \\
   & & \mathcal{K}(\vt_\ast, \{\vt_\ast \}^\textrm{grid}) (\mathbf{\Sigma}^{\textrm{grid}})^{-1} \mathcal{K}( \{\vt_\ast \}^\textrm{grid}, \vt_\ast) \nonumber.
@@
 \begin{eqnarray} \label{eqn:weight_conditional}
-  p(\mathbf{w} |\, \vt_\ast) &=& \mathcal{N} \left ( \mathbf{w} \left | \right . \mathbf{\mu}_\mathbf{w}, \mathbf{\Sigma}_\mathbf{w} \right ) \\ \nonumber
+  p(\mathbf{w} |\, \vt_\ast) &=& \mathcal{N} \left ( \mathbf{w} \left | \right .
+  \mathbf{\mu}_\mathbf{w}(\vt_\ast), \mathbf{\Sigma}_\mathbf{w}(\vt_\ast) \right ) \\ \nonumber
   & = & \mathcal{N} \left ( \mathbf{w} \left | 
@@
 the current values of the observational parameters ($\vt_\textrm{obs}$) and the polynomial nuisance 
-parameters $(\phi_\mathsf{P})$.  Now, the model spectrum is specified conditional on the vector of 
-eigenspectra weights as 
-\begin{equation}
-  \mathsf{M} \bigl | \bigr . \,\mathbf{w} = \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w}
+parameters $(\phi_\mathsf{P})$.  Now, the model spectrum is a function of the vector of 
+eigenspectra weights  
+\begin{equation}
+  \mathsf{M}(\mathbf{w}) = \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w}
 \end{equation}
@@
 \end{equation}
-Because the Gaussian process describes a probability distribution of the weights, we now have a probability distribution of possible (interpolated) models
-\begin{equation}
-  p(\mathsf{M} |\, \mathbf{w}) = p( \mathbf{w} |\, \vt_\ast)
-\end{equation}
-and the likelihood function (Eq.~\ref{eqn:likelihood}) is specified conditional on the weights,
-\begin{equation}
-  p( \mathsf{D} |\, \mathsf{M}) = p( \mathsf{D} |\, \mathbf{w}) = 
+Because the Gaussian process describes a probability distribution of the weights, we now have a
+distribution of possible (interpolated) models and the likelihood function
+(Eq.~\ref{eqn:likelihood}) is specified conditionally on the weights,
+\begin{equation}
+  p \bigl( \mathsf{D} |\, \mathsf{M}(\mathbf{w}) \bigr) = p( \mathsf{D} |\, \mathbf{w}) = 
   \mathcal{N} \left ( \mathsf{D} |\, \widetilde{\xi}_\mu + \mathbf{X} \mathbf{w} , \mathsf{C} \right ).

### 2014-12-06 tex/paper.tex
Coauthor comments addressed.

@@
 \newcommand{\ctext}[1]{ \textcolor{blue}{\% #1}}
+
+\hyphenation{ExoLab}
+\hyphenation{ExoStat}
 
@@
 stellar properties, they are of acute interest to the exoplanet community.  There, all estimates of
-the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host mass
-function or radius {\it ratio} is constrained with the radial velocity or transit techniques,
+the planet properties are made {\it relative} to the host properties (e.g., the planet-to-host
+radius {\it ratio} or mass function is constrained with the radial velocity or transit techniques,
 respectively).  Moreover, essential clues to the planet formation process are encapsulated in the
@@
 spectra. By marginalizing over all of the possible interpolated models, we can modify the likelihood
-function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation uncertainty. For
-simplicity, however, we now proceed as if we used linear interpolation from a synthetic library to
-generate model spectra and that the error introduced by this process is negligible, since the
-following concepts are applicable regardless of the interpolation method.
+function (introduced in Section~\ref{subsec:likelihood}) to account for the interpolation
+uncertainty. However, since the following concepts are applicable regardless of the interpolation
+method, for simplicity we now proceed as if we were using linear interpolation to generate model spectra
+and assume that the error introduced by this process is negligible.
 
@@
 that incorporate dynamical effects, geometry, and the relative location of the target, while the 
-second employs a suite of nuisance parameters ($\vp$) designed to mitigate an imperfect 
+second employs a suite of nuisance parameters $\vp$ designed to mitigate an imperfect 
 data calibration.
@@
 
-\begin{figure}[b!]
-\includegraphics{figs/chebyshev.pdf}
-\vspace{-0.75cm}
-\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
-spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
-shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
-adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
-by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
-(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
-\end{figure}
-
 The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
@@
 
+\begin{figure}[t!]
+\includegraphics{figs/chebyshev.pdf}
+\vspace{-0.75cm}
+\figcaption{A demonstration of our treatment for residual calibration mismatches.  The observed 
+spectra at the overlap of two echelle orders ({\it top}) have slightly ($\sim$1--3\%) discrepant 
+shapes.  By using Chebyshev polynomials ({\it middle, top}) one can correct for that mismatch by
+adjusting the data ({\it middle, bottom}); instead, in practice we equivalently distort the model
+by these polynomials ({\it bottom}) such that the model remains linear in the Chebyshev coefficients
+(Eq.~\ref{eqn:chebyshev}).  \label{fig:chebyshev}}
+\end{figure}
+
 We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internally}
-to the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
+within the modeling framework to propagate the uncertainties introduced by additional degrees of freedom
 in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as
@@
 
+\vspace{1cm}
+
 \subsection{Model Evaluation} \label{subsec:likelihood}
@@
 that penalizes models which yield larger residuals and explicitly allows for covariances in the
-residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix $\vC$.  For practical 
-reasons, the log-likelihood is used as the quality metric, where
+residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix $\vC$\footnote{When using a
+spectral emulator to interpolate model spectra, this matrix will be further modified to account
+for the propagated interpolation uncertainty (Appendix \ref{sec:Appendix}, Eq.~\ref{eqn:modC})}.
+For practical reasons, the log-likelihood is used as the quality metric, where
 \begin{equation}
@@
 width ($\sigma_v$).  Figure \ref{fig:class0} highlights a specific example of these correlated 
-residuals in real data; a significant autocorrelation signal is found on a $\sim$5 pixel scale, 
+residuals in real data; a significant autocorrelation signal is found on an $\sim$8 pixel scale, 
 corresponding to the 6.8\,km s$^{-1}$ FWHM of $\mathcal{F}_v^{\rm inst}$.
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
 \begin{center}
@@
 \begin{equation}
-  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm g}}) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
+  \Kglobal_{ij}(\vp_{{\mathsf C}, {\rm G}}) = a_{\rm G} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
    \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
@@
 \end{equation}
-where the hyperparameters $\vp_{{\mathsf C}, {\rm g}} = [a_{\rm g}, \ell]$ include an amplitude 
-($a_{\rm g}$) and scale ($\ell$).  To ensure that $\vC$ remains a relatively sparse matrix that 
-enables computational expediency, we employ a Hann window function
-\begin{equation}
-  w_{ij}^{\rm G} \,(r_0) = \left \{ 
+where the hyperparameters $\vp_{{\mathsf C}, {\rm G}} = [a_{\rm G}, \ell]$ include an amplitude
+($a_{\rm G}$) and scale ($\ell$). We distinguish the kernel parameters from other parameters in the
+model by labelling them hyperparameters. Because a Gaussian process describes a population of
+functions that can be generated by random draws from the covariance matrix, the covariance kernel
+parameters are truly part of a hierarchical model. In our case, the functions described by these
+hyperparameters are actually the many different realizations of covariant residuals that result from
+a spectral fit. For an example of the Gaussian process kernel and the population of covariant
+residual ``noise'' that could be generated from it, see Figure~\ref{fig:matrix}.
+To ensure that $\vC$ remains a relatively sparse matrix that enables computational expediency, we
+employ a Hann window function
+\begin{equation}
+  w_{ij} \,(r_0) = \left \{ 
     \begin{array}{cc}
@@
 \end{equation}
-as a taper.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
+to taper the kernel.  The truncation distance $r_0$ can be fixed to a reasonable multiple of the scale (we 
 set $r_0 = 4\ell$).
-
-\subsubsection{Local Covariance Structure} \label{subsec:local_covariance}
-
-In addition to the global covariance structure, there can be local regions of strong (highly 
-correlated) residuals that need to be treated in the modeling framework.  These patches of large 
-$\vR$ are usually produced by imperfect model spectral lines (due to missing opacity sources, 
-uncertain oscillator strengths, etc.); some representative examples are shown in 
-Figure~\ref{fig:badlines}.  To parameterize such regions in $\vC$, we introduce a sequence of 
-non-stationary kernels that explicitly depend on the actual wavelength values of a pair of pixels 
-(on $\lambda_i$ and $\lambda_j$), and not simply the distance between them ($r_{ij}$).
-
-Assuming that these local residuals are produced primarily by pathological differences in the 
-spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
-model.  In that case, the $k^{\rm th}$ such local residual could be described as
-\begin{equation}
-\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
-\end{equation}
-with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
-Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
-residual regions can be written
-\begin{equation} \label{eqn:kregion}
-  \mathcal{K}^{\textrm{L},k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
-\end{equation}
-such that the full local covariance kernel composed of a sequence of residual regions is the linear 
-combination,
-\begin{equation} \label{eqn:klocal}
-  \Klocal_{ij}(\vp_{{\mathsf C},l}) = \sum_k w^k_{ij} \, \mathcal{K}^{\textrm{L},k}_{ij},
-\end{equation}
-with a set of hyperparameters $\vp_{{\mathsf C},\textrm{L}} = [\{a_k, \mu_k, \sigma_k\}]$.  Note that we 
-again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure a sparse covariance 
-matrix; in this case, the truncation distance $r_0$ can be set to some multiple of the width 
-parameter (e.g., $r_0 = 4\sigma_k$).  In effect, these kernels systematically down-weight the 
-influence of strong residuals in the likelihood calculation, mitigating any potential bias they 
-might induce on inferences of the interesting parameters ($\vT$).  This acts like a robust, 
-flexible, and unbiased method for (correlated) outlier rejection that preserves the integrity of 
-the probabilistic framework (as opposed to the common manual or threshold-based techniques of 
-clipping or masking).
-
-\begin{figure}[!t]
-\begin{center}
-\includegraphics{figs/badlines.pdf}
-\figcaption{A particularly illustrative spectral region with substantial localized structure in 
-the residuals due to ``outlier" spectral lines in the model library.  For any specific line, there 
-might exist a set of model parameters, $\vT$, that will improve its match with the data, but a 
-$\vT$ that will properly fit \emph{all} the outlier lines does not exist in a pre-computed library 
-with (necessarily) limited parametric flexibility.  Out of concern that such intrinsic mismatches 
-can bias the inference on $\vT$, the methodology advocated here introduces local kernels to inflate 
-the covariance around these outliers, self-consistently down-weighting their influence on the fit.  
-\label{fig:badlines}}
-\end{center}
-\end{figure}
-
-These local kernels can be modified to account for more complex residual structures.  For example, 
-late-type stars with imperfectly modeled molecular bandheads may produce a complicated pattern of 
-positive and negative residuals or a pronounced mismatch over a relatively large spectral scale.  
-This phenomenologically different local covariance behavior can still be treated in this framework 
-if an appropriate kernel morphology is adopted.
 
@@
 
+\subsubsection{Local Covariance Structure} \label{subsec:local_covariance}
+
+In addition to the global covariance structure, there can be local regions of strong (highly 
+correlated) residuals that need to be treated in the modeling framework.  These patches of large 
+$\vR$ are usually produced by imperfect model spectral lines (due to missing opacity sources, 
+uncertain oscillator strengths, etc.); some representative examples are shown in 
+Figure~\ref{fig:badlines}.  To parameterize such regions in $\vC$, we introduce a sequence of 
+non-stationary kernels that explicitly depend on the actual wavelength values of a pair of pixels 
+(on $\lambda_i$ and $\lambda_j$), and not simply the distance between them ($r_{ij}$).
+
+Assuming that these local residuals are produced primarily by pathological differences in the 
+spectral line strength (rather than shape or center), a simple Gaussian is a reasonable residual 
+model.  In that case, the $k^{\rm th}$ such local residual could be described as
+\begin{equation}
+\vR = a_k \exp \left[ - \frac{r^2(\lambda,\mu_k)}{2\sigma_k^2} \right],
+\end{equation}
+with peak amplitude $a_k$, mean wavelength $\mu_k$, and width $\sigma_k$.  Following 
+Eq.~\ref{eqn:expectation}, the covariance of any two pixels related to the $k^{\rm th}$ local 
+residual regions can be written
+\begin{equation} \label{eqn:kregion}
+  \mathcal{K}^{\textrm{L},k}_{ij} = a_k^2 \exp \left [ - \, \frac{r^2(\lambda_i, \mu_k) + r^2(\lambda_j, \mu_k)}{2 \sigma_k^2}\right ],
+\end{equation}
+such that the full local covariance kernel composed of a sequence of residual regions is the linear 
+combination,
+\begin{equation} \label{eqn:klocal}
+  \Klocal_{ij}(\vp_{{\mathsf C},L}) = \sum_k^N w^k_{ij} \, \mathcal{K}^{\textrm{L},k}_{ij},
+\end{equation}
+with a set of hyperparameters $\vp_{{\mathsf C},\textrm{L}} = [\{a_1, \mu_1, \sigma_1, \ldots, a_N,
+\mu_N, \sigma_N\}]$.  Note that we again taper the kernels with Hann windows (Eq.~\ref{eqn:Hann}) to ensure
+a sparse covariance matrix; in this case, the truncation distance $r_0$ can be set to some multiple
+of the width parameter (e.g., $r_0 = 4\sigma_k$).  In effect, these kernels systematically
+down-weight the influence of strong residuals in the likelihood calculation, mitigating any
+potential bias they might induce on inferences of the interesting parameters ($\vT$).  This acts
+like a robust, flexible, and unbiased method for (correlated) outlier rejection that preserves the
+integrity of the probabilistic framework (as opposed to the common manual or threshold-based
+techniques of masking or clipping).
+
+Alternatively, one can also think of adding in this Gaussian component to the covariance matrix as
+being equivalent to marginalizing over the existence of an outlier spectral line with a prior on the
+line strength that is strictly Gaussian, where adjusting the hyperparameters of the local kernel
+changes the prior on the strength and shape of the outlier residual. We find that once such a local
+kernel is instantiated at the location of an outlier spectral line, the inference on the stellar
+parameters ($\vT$) is moderately robust to the precise values of the kernel parameters. This means
+that the marginalization over outlier lines can be performed at minimal computational cost, since
+MCMC iterations which would otherwise be spent exploring the parameter space of the local kernel
+hyperparameters are largely unnecessary if one is only interested in the stellar parameters. For
+completeness, however, the local kernel hyperparameters of the spectral fits presented in
+Section~\ref{sec:examples} are allowed to float and explored simultaneously with the other stellar
+and nuisance parameters.
+
+\begin{figure}[!t]
+\begin{center}
+\includegraphics{figs/badlines.pdf}
+\figcaption{A particularly illustrative spectral region with substantial localized structure in 
+the residuals due to ``outlier" spectral lines in the model library.  For any specific line, there 
+might exist a set of model parameters, $\vT$, that will improve its match with the data, but a 
+$\vT$ that will properly fit \emph{all} the outlier lines does not exist in a pre-computed library 
+with (necessarily) limited parametric flexibility.  Out of concern that such intrinsic mismatches 
+can bias the inference on $\vT$, the methodology advocated here introduces local kernels to inflate 
+the covariance around these outliers, self-consistently down-weighting their influence on the fit.  
+\label{fig:badlines}}
+\end{center}
+\end{figure}
+
+These local kernels can be modified to account for more complex residual structures.  For example, 
+late-type stars with imperfectly modeled molecular bandheads may produce a complicated pattern of 
+positive and negative residuals or a pronounced mismatch over a relatively large spectral scale.  
+This phenomenologically different local covariance behavior can still be treated in this framework 
+if an appropriate kernel morphology is adopted.
+
 \subsubsection{Composite Covariance Matrix}
@@
 \begin{equation}
-\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, g}) + 
-                  \Klocal_{ij}(\vp_{{\mathsf C}, l}), 
+\vC_{ij}(\cov)  = b \, \delta_{ij} \, \sigma_i + w_{ij} \, \Kglobal_{ij}(\vp_{{\mathsf C}, G}) + 
+                  \Klocal_{ij}(\vp_{{\mathsf C}, L}), 
 \end{equation}
@@
 Figure \ref{fig:matrix} provides a graphical illustration of how the kernels that comprise the 
-covariance matrix are able to reproduce the structure present in a typical residual spectrum (for 
-the sake of clarity, the interpolation error kernel is discussed separately, in the Appendix).
+covariance matrix are able to reproduce the structure present in a typical residual spectrum.
 
@@
 $\vt_{\rm obs} \ge 0$ (i.e., for physically meaningful values).
+
+\begin{figure}[!b]
+  \includegraphics{figs/PGM.pdf}
+  \figcaption{A probabilistic graphical model representing how the parameters of the model are
+    combined to forward model a spectrum.  When interpolating models using a spectral emulator
+    (Appendix \ref{sec:Appendix}), the stellar parameters ($\vt_\ast$) determine the weights
+    ($\mathbf{w}$) of the eigenspectra ($\mathbf{\Xi}$), which are modified according to the
+    observational parameters ($\vt_\textrm{obs}$) and polynomial parameters ($\cheb$). Together,
+    these parameters specify the model spectrum ($\vM$). If one uses linear interpolation instead of
+    a spectral emulator, then there would be no intermediate nodes for $\mathbf{w}$ and $\mathbf{\Xi}$. The structure
+    of the covariance matrix, which is included in the likelihood function, is determined by the
+    covariance hyperparameters ($\cov$). Together, the model spectrum and the covariance matrix
+    predict the resulting dataset ($\vD$).
+\label{fig:PGM}}
+\end{figure}
 
@@
 typically adopt flat priors for the amplitudes and means \{$a_k$, $\mu_k$\}, but construct a
-specialized prior for the widths \{$\sigma_k$\} that is flat below the instrumental broadening 
-width (when $\sigma_k < \sigma_v$) and has a rapid (logistic) taper to zero at larger values.  This 
-form is designed to prevent the local kernels from diffusing to large $\sigma_k$ and low $a_k$,
-since such behavior is better treated by the global kernel.
-
-\begin{figure}[!htb]
-  \includegraphics{figs/PGM.pdf}
-  \figcaption{ Probabilistic graphical model.
-\label{fig:PGM}}
-\end{figure}
+specialized prior for the widths \{$\sigma_k$\} that is flat below the combined broadening width
+(when $\sigma_k$ is less than the FWHM of $\varphi$) and has a rapid (logistic) taper to zero at
+larger values.  This form is designed to prevent the local kernels from diffusing to large
+$\sigma_k$ and low $a_k$, since such behavior is better treated by the global kernel.
 
@@
 
-The inference problem developed here has a natural hierarchical structure between the collections 
-of ``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, and ``nuisance" hyperparameters, 
-$\vP = [\cheb, \cov]$.  To explore the posterior distribution of the model conditioned on the data
-\begin{equation} 
-p(\vT, \vP | \vD) \propto p(\vD | \vT, \vP) \, p(\vT, \vP)
+The inference problem developed here has a natural structure between the collections of
+``interesting" parameters, $\vT = [\vt_{\ast}, \vt_{\rm obs}]$, the nuisance calibration parameters
+$\cheb$, and the covariance hyperparameters $\cov$. The conditional dependencies of these parameters
+are shown graphically in Figure~\ref{fig:PGM}. To explore the posterior distribution of the model
+conditioned on the data
+\begin{equation} 
+p(\vT, \cheb, \cov | \vD) \propto p(\vD | \vT, \cheb, \cov) \, p(\vT, \cheb, \cov)
 \label{eqn:post}
@@
 sampler coupled to the Metropolis-Hastings algorithm.  To summarize, this procedure works by 
-sampling in a subset of parameters (with Metropolis-Hastings) conditioned on the current (fixed) 
+sampling in a subset of parameters (with Metropolis-Hastings parameter proposals) conditioned on the current (fixed) 
 values of the other parameters; after each iteration, the Gibbs sampler then updates the values of 
@@
 \noindent (1) Initialize the parameters.  One might set $\vT^0$ based on estimates in the 
-literature or scaling behaviors, and make simple assumptions about $\vP^0$.  Here, we set the 
+literature or scaling behaviors, and make simple assumptions about $\cheb^0$.  Here, we set the 
 Chebyshev coefficients ($\cheb^0$) so that the polynomials are constant ($c_o^{(0)} = 1$,  
@@
 
-\noindent (2) Start the $i^{\rm th}$ (where $i \in [1,N_G]$) iteration of the Gibbs sampler.  For 
+\noindent (2) Start the $i^{\rm th}$ iteration of the Gibbs sampler.  For
 each iteration of the Metropolis-Hastings algorithm, sample in the interesting parameters $\vT$ to
@@
 \ref{subsec:likelihood} and \ref{subsec:priors}.  This represents a ``slice" through the posterior
-space conditioned on the nuisance parameters being held fixed ($\vP = \vP^{i-1}$).  Then update the 
-interesting parameters $\vT^{i-1} \rightarrow \vT^i$. \\
-
-\noindent (3) Cycle through and update the nuisance parameters, $\vP^{i-1} \rightarrow \vP^i$.  For 
-each spectral order,
-
-\begin{enumerate}[(a)]
-\item Sample in the polynomial coefficients $\cheb$ along a subset of the posterior space where 
-$\vT^i$ and $\cov^{i-1}$ are fixed.  Update the coefficients, $\cheb^{i-1} \rightarrow \cheb^i$.
-
-\item Sample in the global (stationary) kernel parameters, $\vp_{{\mathsf C},g} = [a_g, \ell]$, 
-with $\vT^i$, $\cheb^i$, and the local kernel parameters, $\vp_{{\mathsf C}, l}^{i-1}$, fixed.  
-Update $\vp_{{\mathsf C}, g}^{i-1} \rightarrow \vp_{{\mathsf C}, g}^i$.
-
-\item If any local (non-stationary) kernels have been instantiated (see below), sample in their 
-hyperparameters $\vp_{{\mathsf C},l}$ with $\vT^i$, $\cheb^i$, and $\vp_{{\mathsf C}, g}^i$ fixed.  
-Update the local kernel parameters $\vp_{{\mathsf C}, k}^{i-1} \rightarrow \vp_{{\mathsf C}, k}^i$.
-
-\end{enumerate}
+space conditioned on the other parameters being held fixed ($\cheb = \cheb^{i-1}$ and $\cov =
+\cov^{i-1}$).  Then update $\vT^{i-1} \rightarrow \vT^i$. \\
+
+\noindent (3) For each spectral order, sample in the polynomial parameters $\cheb$ and covariance
+hyperparameters $\cov$, conditioned on the other parameters being held fixed $\vT^i = \vT^{i-1}$.
+Then update $\cheb^{i-1} \rightarrow \cheb^i$ and $\cov^{i-1} \rightarrow \cov^i$. \\
 
@@
 
-In Step~(3c), local covariance kernels are instantiated into the model according to the following 
+In Step~(3), local covariance kernels are instantiated into the model according to the following 
 procedure.  First, an ``average" residual spectrum is generated by combining $\sim$500 residual 
@@
 synthetic model libraries: a custom modification of the \citet{castelli04} models designed and 
-empirically calibrated to reproduce well the optical spectrum around the \ion{Mg}{1}\,b triplet at 
+empirically calibrated to accurately reproduce the optical spectrum around the \ion{Mg}{1}\,b triplet at 
 $\sim$5100\,\AA\ for Sun-like stars (used in {\tt SPC}; hereafter the {\sc CfA/Kurucz} library), 
@@
 opacities.  Especially confounding is the presence of complex condensates (clouds) at the coolest 
-temperatures \citep{allard13}, making it considerably more challenging to determine (sub-)stellar 
+temperatures \citep{allard13}, making it considerably more challenging to determine (sub-) stellar 
 properties \citep{rajpurohit14}.  Various approaches have been taken to infer the key parameters in 
@@
 
-Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
-application being considered here there are some subtleties that necessitate a more sophisticated 
-approach.  Spectral line formation is a non-linear process, where line shapes are determined by 
-complex radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass 
-this non-linear process and predict the behavior of many different spectral lines for an arbitrary 
-set of parameters based on a pre-computed set of library spectra with similar (neighboring) 
-parameters.  In this Appendix, we describe the procedure through which we propagate an empirical 
-approximation of the {\it uncertainty} in that prediction -- which can be substantial -- into and 
-through the probabilistic inference framework we have developed.
-
-There are two important factors to consider when interpolating libraries of model stellar spectra.  
-First, the fluxes in a non-negligible fraction of the spectral pixels do not vary monotonically 
-with changes in the physical parameters that define the library, $\vt_{\ast}$: this helps inform 
-the decision about which functional form of interpolation should be used.  Second, the spectra 
-are usually not Nyquist sampled with respect to $\vt_{\ast}$.  Unfortunately, this fact means that 
-the library is not sampled densely enough in parameter-space to reconstruct a synthetic spectrum 
-for arbitrary parameters without some information loss.  Given these factors, we choose to develop 
-a spectral ``emulator'' to aid in the interpolation.  Given the information contained in 
-the spectral library and an arbitrary desired $\vt_{\ast}$, this emulator delivers a probability 
-distribution representing the range of possible interpolated spectra.
-
-Model library spectra are stored as (1-dimensional) arrays of fluxes, sampled on high resolution 
-wavelength grids.  In the case of interest here, the sets of model parameters 
-$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the dimensions of the 
-library grid.  The full spectral library, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, is therefore 
-encapsulated in a 4-dimensional array.  The libraries used here have grid spacings of 0.5\,dex in 
-$\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but 
-the {\sc Phoenix} library has finer coverage in 100\,K increments.
-
-\begin{figure*}[!t]
+\begin{figure*}[!htb]
   \begin{center}
@@
 
+Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
+application being considered here there are some subtleties that necessitate a more sophisticated 
+approach.  Spectral line formation is a non-linear process, where line shapes are determined by 
+complex radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass 
+this non-linear process and predict the behavior of many different spectral lines for an arbitrary 
+set of parameters based on a pre-computed set of library spectra with similar (neighboring) 
+parameters.  In this Appendix, we describe the procedure through which we propagate an empirical 
+approximation of the {\it uncertainty} in that prediction -- which can be substantial -- into and 
+through the probabilistic inference framework we have developed.
+
+There are two important factors to consider when interpolating libraries of model stellar spectra.  
+First, the fluxes in a non-negligible fraction of the spectral pixels do not vary monotonically 
+with changes in the physical parameters that define the library, $\vt_{\ast}$: this helps inform 
+the decision about which functional form of interpolation should be used.  Second, the spectra 
+are usually not Nyquist sampled with respect to $\vt_{\ast}$.  Unfortunately, this fact means that 
+the library is not sampled densely enough in parameter-space to reconstruct a synthetic spectrum 
+for arbitrary parameters without some information loss.  Given these factors, we choose to develop 
+a spectral ``emulator'' to aid in the interpolation.  Given the information contained in 
+the spectral library and an arbitrary desired $\vt_{\ast}$, this emulator delivers a probability 
+distribution representing the range of possible interpolated spectra.
+
+Model library spectra are stored as (1-dimensional) arrays of fluxes, sampled on high resolution 
+wavelength grids.  In the case of interest here, the sets of model parameters 
+$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the dimensions of the 
+library grid.  The full spectral library, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, is therefore 
+encapsulated in a 4-dimensional array.  The libraries used here have grid spacings of 0.5\,dex in 
+$\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but 
+the {\sc Phoenix} library has finer coverage in 100\,K increments.
+
 The first step in designing a spectral emulator is to break down the library into an appropriate 
@@
 set of those parameters drawn from this collection. Both $\xi_\mu$ and $\xi_\sigma$ are
-vectors with length $N_\textrm{pix}$, the same size as any
+vectors with length $N_\textrm{pix}$, the same size as a
 raw synthetic spectrum ($f_\lambda$). In effect, all library spectra are 
@@
 where $w_k$ is the weight of the $k^{\rm th}$ eigenspectrum.  These weights are 3-dimensional 
-scalar functions that depend on the stellar parameters $\vt_\ast$.  These weights, which are 
+scalar functions that depend on the stellar parameters $\vt_\ast$.  The weights, which are 
 generally smooth functions of the stellar parameters (see the left panel of 
@@
 To recapitulate, the framework described above can be used to decompose the synthetic spectra in a 
-model library into a principal component basis, allowing us to lossily reconstruct any spectrum in 
+model library into a principal component basis, allowing us to reconstruct any spectrum in 
 the library as a (weighted) linear combination of $m$ eigenspectra.  The weights corresponding to 
@@
 \begin{eqnarray}
-  \mathcal{K}(\vt_i, \vt_j | \phi_{\rm int} ) &=& a_{\rm int}^2 \exp \left [ - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 \, \ell_{T_\textrm{eff}}^2 } \right] \nonumber \\
+  \mathcal{K}(\vt_{\ast, i}, \vt_{\ast, j} | \phi_{\rm int} ) &=& a_{\rm int}^2 \exp \left [ - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 \, \ell_{T_\textrm{eff}}^2 } \right] \nonumber \\
   & & \times \exp \left [-\frac{(\log g_i - \log g_j)^2}{2 \, \ell_{\log g}^2} \right ] \\
@@
   \mathsf{C}^\prime = \mathbf{X} \mathbf{\Sigma}_\mathbf{w} \mathbf{X}^T + \mathsf{C}
+  \label{eqn:modC}
 \end{equation}

### 2014-12-08 tex/paper.tex
Submitted to ApJ

@@
 \usepackage[utf8]{inputenc}
-
-%version-control tagging based off of github.com/bd-j/speccal
-\input{vc}
 
@@
 
-\graphicspath{{figs/}}
+% it was great while it lasted
+%\graphicspath{{figs/}}
 
@@
 \begin{center}
-  \includegraphics{figs/kernels.pdf}
-  \includegraphics{figs/high2low.pdf}
+  \includegraphics{kernels.pdf}
+  \includegraphics{high2low.pdf}
   \figcaption{({\it top}) The line-of-sight velocity distribution function, $\varphi_v$, and its 
@@
 \begin{figure}[t!]
-\includegraphics{figs/chebyshev.pdf}
+\includegraphics{chebyshev.pdf}
 \vspace{-0.75cm}
@@
 \begin{center}
-\includegraphics[scale=1.05]{figs/matrix_compilation.pdf}
+\includegraphics[scale=1.05]{matrix_compilation.pdf}
 \figcaption{A decomposition of the modeling procedure, explicitly highlighting the roles of the 
@@
 \begin{center}
-\includegraphics{figs/badlines.pdf}
+\includegraphics{badlines.pdf}
 \figcaption{A particularly illustrative spectral region with substantial localized structure in 
@@
 \begin{figure}[!b]
-  \includegraphics{figs/PGM.pdf}
+  \includegraphics{PGM.pdf}
   \figcaption{A probabilistic graphical model representing how the parameters of the model are
@@
 \begin{figure*}[!t]
-\includegraphics{figs/residuals_Kurucz_logg.pdf}
+\includegraphics{residuals_Kurucz_logg.pdf}
 \vspace{-0.3cm}
@@
 \begin{figure}[!b]
-  \includegraphics{figs/Kurucz_triangle.pdf}
+  \includegraphics{Kurucz_triangle.pdf}
 \vspace{-0.5cm}
@@
 \begin{figure*}[!htb]
-\includegraphics{figs/residuals_PHOENIX_logg.pdf}
+\includegraphics{residuals_PHOENIX_logg.pdf}
 \vspace{-0.3cm}
@@
 \begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/PHOENIX_triangle.pdf}
+  \includegraphics[width=0.5\textwidth]{PHOENIX_triangle.pdf}
 \vspace{-0.5cm}
@@
 \begin{figure*}[!htb]
-  \includegraphics{figs/residuals_Gl51_logg.pdf}
+  \includegraphics{residuals_Gl51_logg.pdf}
 \vspace{-0.3cm}
@@
 \begin{figure}[!b]
-  \includegraphics[width=0.5\textwidth]{figs/Gl51_triangle.pdf}
+  \includegraphics[width=0.5\textwidth]{Gl51_triangle.pdf}
   \figcaption{The posterior probability distributions for the interesting stellar parameters
@@
 
-\acknowledgments  The authors would like to acknowledge the following people
-for many extraordinarily helpful discussions and key insights: Daniel
-Foreman-Mackey, Guillermo Torres, David Latham, John Johnson and the ExoLab,
-Daniel Eisenstein, Rebekah Dawson, Tom Loredo and the ExoStat group, Allyson
-Bieryla, and Maxwell Moe.  IC was supported by the NSF Graduate Fellowship and
-the Smithsonian Institution.  Figure~\ref{fig:PGM} was generated with
-\texttt{Daft}\footnote{\url{http://daft-pgm.org/}} and Figures
-\ref{fig:Kurucz_posterior}, \ref{fig:PHOENIX_posterior}, and
-\ref{fig:Gl51_posterior} were generated with \texttt{triangle.py}
-\citep{foreman-mackey14}.  This research made extensive use of Astropy
-\citep{astropy13} and the Julia programming language \citep{julia12}.
+\acknowledgments  The authors would like to acknowledge the following people for many
+extraordinarily helpful discussions and key insights: Daniel Foreman-Mackey, Guillermo Torres, David
+Latham, John Johnson and the ExoLab, Daniel Eisenstein, Rebekah Dawson, Tom Loredo and the ExoStat
+group, Allyson Bieryla, and Maxwell Moe.  IC is supported by the NSF Graduate Fellowship and the
+Smithsonian Institution. K.M. is supported at Harvard by NSF grant AST-1211196. Figure~\ref{fig:PGM} was generated with
+\texttt{Daft}\footnote{\url{http://daft-pgm.org/}} and Figures \ref{fig:Kurucz_posterior},
+\ref{fig:PHOENIX_posterior}, and \ref{fig:Gl51_posterior} were generated with \texttt{triangle.py}
+\citep{foreman-mackey14}.  This research made extensive use of Astropy \citep{astropy13} and the
+Julia programming language \citep{julia12}.
 
@@
   \begin{center}
-  \includegraphics{figs/pca_reconstruct.pdf}
+  \includegraphics{pca_reconstruct.pdf}
   \vspace{-0.6cm}
@@
 \begin{center}
-  \includegraphics{figs/GP_left.pdf}
-  \includegraphics{figs/GP_right.pdf}
+  \includegraphics{GP_left.pdf}
+  \includegraphics{GP_right.pdf}
   \figcaption{The Gaussian process modelling of the principal component weights for the Gl51

### 2014-12-17 tex/paper.tex
Moved files

@@
 \bibliographystyle{yahapj}
-\bibliography{stellarspectra}
+%\bibliography{stellarspectra}
+
+\begin{thebibliography}{48}
+\providecommand\natexlab[1]{#1}
+\providecommand\JournalTitle[1]{#1}
+
+\bibitem[{{Allard} {et~al.}(2012){Allard}, {Homeier}, \& {Freytag}}]{allard12}
+{Allard}, F., {Homeier}, D., \& {Freytag}, B. 2012,
+  \href{http://dx.doi.org/10.1098/rsta.2011.0269}{\JournalTitle{Royal Society
+  of London Philosophical Transactions Series A}, 370, 2765}
+
+\bibitem[{{Allard} {et~al.}(2013){Allard}, {Homeier}, {Freytag},
+  {Schaffenberger}, \& {Rajpurohit}}]{allard13}
+{Allard}, F., {Homeier}, D., {Freytag}, B., {Schaffenberger}, \& {Rajpurohit},
+  A.~S. 2013, \JournalTitle{Memorie della Societa Astronomica Italiana
+  Supplementi}, 24, 128
+
+\bibitem[{{Astropy Collaboration} {et~al.}(2013){Astropy Collaboration},
+  {Robitaille}, {Tollerud}, {Greenfield}, {Droettboom}, {Bray}, {Aldcroft},
+  {Davis}, {Ginsburg}, {Price-Whelan}, {Kerzendorf}, {Conley}, {Crighton},
+  {Barbary}, {Muna}, {Ferguson}, {Grollier}, {Parikh}, {Nair}, {Unther},
+  {Deil}, {Woillez}, {Conseil}, {Kramer}, {Turner}, {Singer}, {Fox}, {Weaver},
+  {Zabalza}, {Edwards}, {Azalee Bostroem}, {Burke}, {Casey}, {Crawford},
+  {Dencheva}, {Ely}, {Jenness}, {Labrie}, {Lian Lim}, {Pierfederici},
+  {Pontzen}, {Ptak}, {Refsdal}, {Servillat}, \& {Streicher}}]{astropy13}
+{Astropy Collaboration}, {Robitaille}, T.~P., {Tollerud}, E.~J., {et~al.} 2013,
+  \href{http://dx.doi.org/10.1051/0004-6361/201322068}{\JournalTitle{\aap},
+  558, A33}
+
+\bibitem[{{Bezanson} {et~al.}(2012){Bezanson}, {Karpinski}, {Shah}, \&
+  {Edelman}}]{julia12}
+{Bezanson}, J., {Karpinski}, S., {Shah}, V.~B., \& {Edelman}, A. 2012,
+  \JournalTitle{ArXiv e-prints},
+  \href{http://arxiv.org/abs/1209.5145}{{\sffamily arXiv:1209.5145 [cs.PL]}}
+
+\bibitem[{{Buchhave} {et~al.}(2010){Buchhave}, {Bakos}, {Hartman}, {Torres},
+  {Kov\'{a}cs}, {Latham}, {Noyes}, {Esquerdo}, {Everett}, {Howard}, {Marcy},
+  {Fischer}, {Johnson}, {Andersen}, {F{\H u}r\'{e}sz}, {Perumpilly},
+  {Sasselov}, {Stefanik}, {B\'{e}ky}, {L\'{a}z\'{a}r}, {Papp}, \&
+  {S\'{a}ri}}]{buchhave10}
+{Buchhave}, L.~A., {Bakos}, G.~A., {Hartman}, J.~D., {et~al.} 2010,
+  \href{http://dx.doi.org/10.1088/0004-637X/720/2/1118}{\JournalTitle{\apj},
+  720, 1118}
+
+\bibitem[{{Buchhave} {et~al.}(2012){Buchhave}, {Latham}, {Johansen},
+  {Bizzarro}, {Torres}, {Rowe}, {Batalha}, {Borucki}, {Brugamyer}, {Caldwell},
+  {Bryson}, {Ciardi}, {Cochran}, {Endl}, {Esquerdo}, {Ford}, {Geary},
+  {Gilliland}, {Hansen}, {Isaacson}, {Laird}, {Lucas}, {Marcy}, {Morse},
+  {Robertson}, {Shporer}, {Stefanik}, {Still}, \& {Quinn}}]{buchhave12}
+{Buchhave}, L.~A., {Latham}, D.~W., {Johansen}, A., {et~al.} 2012,
+  \href{http://dx.doi.org/10.1038/nature11121}{\JournalTitle{\nat}, 486, 375}
+
+\bibitem[{{Buchhave} {et~al.}(2014){Buchhave}, {Bizzarro}, {Latham},
+  {Sasselov}, {Cochran}, {Endl}, {Isaacson}, {Juncher}, \&
+  {Marcy}}]{buchhave14}
+{Buchhave}, L.~A., {Bizzarro}, M., {Latham}, D.~W., {et~al.} 2014,
+  \href{http://dx.doi.org/10.1038/nature13254}{\JournalTitle{\nat}, 509, 593}
+
+\bibitem[{{Cardelli} {et~al.}(1989){Cardelli}, {Clayton}, \&
+  {Mathis}}]{cardelli89}
+{Cardelli}, J.~A., {Clayton}, G.~C., \& {Mathis}, J.~S. 1989,
+  \href{http://dx.doi.org/10.1086/167900}{\JournalTitle{\apj}, 345, 245}
+
+\bibitem[{{Castelli} \& {Kurucz}(2004)}]{castelli04}
+{Castelli}, F., \& {Kurucz}, R.~L. 2004, \JournalTitle{ArXiv Astrophysics
+  e-prints}, \href{http://arxiv.org/abs/astro-ph/0405087}{{\sffamily
+  astro-ph/0405087}}
+
+\bibitem[{{Cushing} {et~al.}(2005){Cushing}, {Rayner}, \& {Vacca}}]{cushing05}
+{Cushing}, M.~C., {Rayner}, J.~T., \& {Vacca}, W.~D. 2005,
+  \href{http://dx.doi.org/10.1086/428040}{\JournalTitle{\apj}, 623, 1115}
+
+\bibitem[{{Cushing} {et~al.}(2004){Cushing}, {Vacca}, \& {Rayner}}]{cushing04}
+{Cushing}, M.~C., {Vacca}, W.~D., \& {Rayner}, J.~T. 2004,
+  \href{http://dx.doi.org/10.1086/382907}{\JournalTitle{\pasp}, 116, 362}
+
+\bibitem[{{Eisenstein} {et~al.}(2006){Eisenstein}, {Liebert}, {Harris},
+  {Kleinman}, {Nitta}, {Silvestri}, {Anderson}, {Barentine}, {Brewington},
+  {Brinkmann}, {Harvanek}, {Krzesi\'{n}ski}, {Neilsen}, {Long}, {Schneider}, \&
+  {Snedden}}]{eisenstein06}
+{Eisenstein}, D.~J., {Liebert}, J., {Harris}, H.~C., {et~al.} 2006,
+  \href{http://dx.doi.org/10.1086/507110}{\JournalTitle{\apjs}, 167, 40}
+
+\bibitem[{{F{\H{u}}r\'{e}sz}(2008)}]{furesz08}
+{F{\H{u}}r\'{e}sz}, G. 2008
+
+\bibitem[{{Fischer} \& {Valenti}(2005)}]{fischer05}
+{Fischer}, D.~A., \& {Valenti}, J. 2005,
+  \href{http://dx.doi.org/10.1086/428383}{\JournalTitle{\apj}, 622, 1102}
+
+\bibitem[{{Foreman-Mackey} {et~al.}(2013){Foreman-Mackey}, {Hogg}, {Lang}, \&
+  {Goodman}}]{foreman-mackey13}
+{Foreman-Mackey}, D., {Hogg}, D.~W., {Lang}, D., \& {Goodman}, J. 2013,
+  \href{http://dx.doi.org/10.1086/670067}{\JournalTitle{\pasp}, 125, 306}
+
+\bibitem[{Foreman-Mackey {et~al.}(2014)Foreman-Mackey, Price-Whelan, Ryan,
+  Rice, Smith, Barbary, Hogg, \& Brewer}]{foreman-mackey14}
+Foreman-Mackey, D., Price-Whelan, A., Ryan, G., {et~al.}
+  \href{http://dx.doi.org/10.5281/zenodo.10598}{2014}
+
+\bibitem[{Gelman {et~al.}(2013)Gelman, Carlin, Stern, Dunson, Vehtari, \&
+  Rubin}]{gelman13}
+Gelman, A., Carlin, J., Stern, H., {et~al.} 2013, {Bayesian Data Analysis,
+  Third Edition}, {Chapman \& Hall/CRC Texts in Statistical Science} (Taylor \&
+  Francis)
+
+\bibitem[{{Gray}(1994)}]{gray94}
+{Gray}, D.~F. 1994,
+  \href{http://dx.doi.org/10.1086/133502}{\JournalTitle{\pasp}, 106, 1248}
+
+\bibitem[{{Gray}(2008)}]{gray08}
+---. 2008, {The Observation and Analysis of Stellar Photospheres}
+
+\bibitem[{{Habib} {et~al.}(2007){Habib}, {Heitmann}, {Higdon}, {Nakhleh}, \&
+  {Williams}}]{habib07}
+{Habib}, S., {Heitmann}, K., {Higdon}, D., {Nakhleh}, C., \& {Williams}, B.
+  2007,
+  \href{http://dx.doi.org/10.1103/PhysRevD.76.083503}{\JournalTitle{\prd}, 76,
+  083503}
+
+\bibitem[{{Hauschildt} {et~al.}(1999){Hauschildt}, {Allard}, \&
+  {Baron}}]{hauschildt99}
+{Hauschildt}, P.~H., {Allard}, F., \& {Baron}, E. 1999,
+  \href{http://dx.doi.org/10.1086/306745}{\JournalTitle{\apj}, 512, 377}
+
+\bibitem[{{Heitmann} {et~al.}(2009){Heitmann}, {Higdon}, {White}, {Habib},
+  {Williams}, {Lawrence}, \& {Wagner}}]{heitmann09}
+{Heitmann}, K., {Higdon}, D., {White}, M., {et~al.} 2009,
+  \href{http://dx.doi.org/10.1088/0004-637X/705/1/156}{\JournalTitle{\apj},
+  705, 156}
+
+\bibitem[{{Howard} {et~al.}(2010){Howard}, {Marcy}, {Johnson}, {Fischer},
+  {Wright}, {Isaacson}, {Valenti}, {Anderson}, {Lin}, \& {Ida}}]{howard10}
+{Howard}, A.~W., {Marcy}, G.~W., {Johnson}, J.~A., {et~al.} 2010,
+  \href{http://dx.doi.org/10.1126/science.1194854}{\JournalTitle{Science}, 330,
+  653}
+
+\bibitem[{{Husser} {et~al.}(2013){Husser}, {Wende-von Berg}, {Dreizler},
+  {Homeier}, {Reiners}, {Barman}, \& {Hauschildt}}]{husser13}
+{Husser}, T.-O., {Wende-von Berg}, S., {Dreizler}, S., {et~al.} 2013,
+  \href{http://dx.doi.org/10.1051/0004-6361/201219058}{\JournalTitle{\aap},
+  553, A6}
+
+\bibitem[{{Ivezi\'{c}} {et~al.}(2013){Ivezi\'{c}}, {Connolly}, {VanderPlas}, \&
+  {Gray}}]{ivezic13}
+{Ivezi\'{c}}, {\.Z}., {Connolly}, A., {VanderPlas}, J., \& {Gray}, A. 2013,
+  {Statistics, Data Mining, and Machine Learning in Astronomy}
+
+\bibitem[{{Johnson} {et~al.}(2007){Johnson}, {Butler}, {Marcy}, {Fischer},
+  {Vogt}, {Wright}, \& {Peek}}]{johnson07}
+{Johnson}, J.~A., {Butler}, R.~P., {Marcy}, G.~W., {et~al.} 2007,
+  \href{http://dx.doi.org/10.1086/521720}{\JournalTitle{\apj}, 670, 833}
+
+\bibitem[{{Joshi} {et~al.}(2009){Joshi}, {Pollacco}, {Collier Cameron},
+  {Skillen}, {Simpson}, {Steele}, {Street}, {Stempels}, {Christian}, {Hebb},
+  {Bouchy}, {Gibson}, {H\'{e}brard}, {Keenan}, {Loeillet}, {Meaburn}, {Moutou},
+  {Smalley}, {Todd}, {West}, {Anderson}, {Bentley}, {Enoch}, {Haswell},
+  {Hellier}, {Horne}, {Irwin}, {Lister}, {McDonald}, {Maxted}, {Mayor},
+  {Norton}, {Parley}, {Perrier}, {Pont}, {Queloz}, {Ryans}, {Smith}, {Udry},
+  {Wheatley}, \& {Wilson}}]{joshi09}
+{Joshi}, Y.~C., {Pollacco}, D., {Collier Cameron}, A., {et~al.} 2009,
+  \href{http://dx.doi.org/10.1111/j.1365-2966.2008.14178.x}{\JournalTitle{\mnras},
+  392, 1532}
+
+\bibitem[{{Koleva} {et~al.}(2009){Koleva}, {Prugniel}, {Bouchard}, \&
+  {Wu}}]{koleva09}
+{Koleva}, M., {Prugniel}, P., {Bouchard}, A., \& {Wu}, Y. 2009,
+  \href{http://dx.doi.org/10.1051/0004-6361/200811467}{\JournalTitle{\aap},
+  501, 1269}
+
+\bibitem[{{Kurucz}(1993)}]{kurucz93}
+{Kurucz}, R.~L. 1993, {SYNTHE spectrum synthesis programs and line data}
+
+\bibitem[{{Mann} {et~al.}(2013){Mann}, {Gaidos}, \& {Ansdell}}]{mann13}
+{Mann}, A.~W., {Gaidos}, E., \& {Ansdell}, M. 2013,
+  \href{http://dx.doi.org/10.1088/0004-637X/779/2/188}{\JournalTitle{\apj},
+  779, 188}
+
+\bibitem[{{Nidever} {et~al.}(2012){Nidever}, {Zasowski}, {Majewski}, {Bird},
+  {Robin}, {Martinez-Valpuesta}, {Beaton}, {Sch\"{o}nrich}, {Schultheis},
+  {Wilson}, {Skrutskie}, {O'Connell}, {Shetrone}, {Schiavon}, {Johnson},
+  {Weiner}, {Gerhard}, {Schneider}, {Allende Prieto}, {Sellgren}, {Bizyaev},
+  {Brewington}, {Brinkmann}, {Eisenstein}, {Frinchaboy}, {Elia Garc\'{\i}a
+  P\'{e}rez}, {Holtzman}, {Hearty}, {Malanushenko}, {Malanushenko}, {Muna},
+  {Oravetz}, {Pan}, {Simmons}, {Snedden}, \& {Weaver}}]{nidever12}
+{Nidever}, D.~L., {Zasowski}, G., {Majewski}, S.~R., {et~al.} 2012,
+  \href{http://dx.doi.org/10.1088/2041-8205/755/2/L25}{\JournalTitle{\apjl},
+  755, L25}
+
+\bibitem[{{Paxton} {et~al.}(2011){Paxton}, {Bildsten}, {Dotter}, {Herwig},
+  {Lesaffre}, \& {Timmes}}]{paxton11}
+{Paxton}, B., {Bildsten}, L., {Dotter}, A., {et~al.} 2011,
+  \href{http://dx.doi.org/10.1088/0067-0049/192/1/3}{\JournalTitle{\apjs}, 192,
+  3}
+
+\bibitem[{{Rajpurohit} {et~al.}(2014){Rajpurohit}, {Reyl\'{e}}, {Allard},
+  {Scholz}, {Homeier}, {Schultheis}, \& {Bayo}}]{rajpurohit14}
+{Rajpurohit}, A.~S., {Reyl\'{e}}, C., {Allard}, F., {et~al.} 2014,
+  \href{http://dx.doi.org/10.1051/0004-6361/201322881}{\JournalTitle{\aap},
+  564, A90}
+
+\bibitem[{{Rajpurohit} {et~al.}(2010){Rajpurohit}, {Reyl\'{e}}, {Schultheis},
+  \& {Allard}}]{rajpurohit10}
+{Rajpurohit}, A.~S., {Reyl\'{e}}, C., {Schultheis}, M., \& {Allard}, F. 2010,
+  in {SF2A-2010: Proceedings of the Annual meeting of the French Society of
+  Astronomy and Astrophysics}, ed. S.~{Boissier}, M.~{Heydari-Malayeri},
+  R.~{Samadi}, \& D.~{Valls-Gabaud}, 275
+
+\bibitem[{Rasmussen \& Williams(2005)}]{rasmussen05}
+Rasmussen, C.~E., \& Williams, C. K.~I. 2005, {Gaussian Processes for Machine
+  Learning (Adaptive Computation and Machine Learning)} (The MIT Press)
+
+\bibitem[{{Rayner} {et~al.}(2009){Rayner}, {Cushing}, \& {Vacca}}]{rayner09}
+{Rayner}, J.~T., {Cushing}, M.~C., \& {Vacca}, W.~D. 2009,
+  \href{http://dx.doi.org/10.1088/0067-0049/185/2/289}{\JournalTitle{\apjs},
+  185, 289}
+
+\bibitem[{{Rayner} {et~al.}(2003){Rayner}, {Toomey}, {Onaka}, {Denault},
+  {Stahlberger}, {Vacca}, {Cushing}, \& {Wang}}]{rayner03}
+{Rayner}, J.~T., {Toomey}, D.~W., {Onaka}, P.~M., {et~al.} 2003,
+  \href{http://dx.doi.org/10.1086/367745}{\JournalTitle{\pasp}, 115, 362}
+
+\bibitem[{{Reid} {et~al.}(1995){Reid}, {Hawley}, \& {Gizis}}]{reid95}
+{Reid}, I.~N., {Hawley}, S.~L., \& {Gizis}, J.~E. 1995,
+  \href{http://dx.doi.org/10.1086/117655}{\JournalTitle{\aj}, 110, 1838}
+
+\bibitem[{{Rojas-Ayala} {et~al.}(2010){Rojas-Ayala}, {Covey}, {Muirhead}, \&
+  {Lloyd}}]{rojas-ayala10}
+{Rojas-Ayala}, B., {Covey}, K.~R., {Muirhead}, P.~S., \& {Lloyd}, J.~P. 2010,
+  \href{http://dx.doi.org/10.1088/2041-8205/720/1/L113}{\JournalTitle{\apjl},
+  720, L113}
+
+\bibitem[{{Rojas-Ayala} {et~al.}(2012){Rojas-Ayala}, {Covey}, {Muirhead}, \&
+  {Lloyd}}]{rojas-ayala12}
+---. 2012,
+  \href{http://dx.doi.org/10.1088/0004-637X/748/2/93}{\JournalTitle{\apj}, 748,
+  93}
+
+\bibitem[{{Sch\"{o}nrich} \& {Bergemann}(2013)}]{schoenrich13}
+{Sch\"{o}nrich}, R., \& {Bergemann}, M. 2013, \JournalTitle{ArXiv e-prints},
+  \href{http://arxiv.org/abs/1311.5558}{{\sffamily arXiv:1311.5558
+  [astro-ph.SR]}}
+
+\bibitem[{{Shkedy} {et~al.}(2007){Shkedy}, {Decin}, {Molenberghs}, \&
+  {Aerts}}]{shkedy07}
+{Shkedy}, Z., {Decin}, L., {Molenberghs}, G., \& {Aerts}, C. 2007,
+  \href{http://dx.doi.org/10.1111/j.1365-2966.2007.11508.x}{\JournalTitle{\mnras},
+  377, 120}
+
+\bibitem[{{Sneden}(1973)}]{sneden73}
+{Sneden}, C.~A. 1973, PhD thesis, The University of Texas at Austin.
+
+\bibitem[{{Tonry} \& {Davis}(1979)}]{tonry79}
+{Tonry}, J., \& {Davis}, M. 1979,
+  \href{http://dx.doi.org/10.1086/112569}{\JournalTitle{\aj}, 84, 1511}
+
+\bibitem[{{Torres} {et~al.}(2012){Torres}, {Fischer}, {Sozzetti}, {Buchhave},
+  {Winn}, {Holman}, \& {Carter}}]{torres12}
+{Torres}, G., {Fischer}, D.~A., {Sozzetti}, A., {et~al.} 2012,
+  \href{http://dx.doi.org/10.1088/0004-637X/757/2/161}{\JournalTitle{\apj},
+  757, 161}
+
+\bibitem[{{Vacca} {et~al.}(2003){Vacca}, {Cushing}, \& {Rayner}}]{vacca03}
+{Vacca}, W.~D., {Cushing}, M.~C., \& {Rayner}, J.~T. 2003,
+  \href{http://dx.doi.org/10.1086/346193}{\JournalTitle{\pasp}, 115, 389}
+
+\bibitem[{{Valenti} \& {Piskunov}(1996)}]{valenti96}
+{Valenti}, J.~A., \& {Piskunov}, N. 1996, \JournalTitle{\aaps}, 118, 595
+
+\bibitem[{{Zucker} {et~al.}(2012){Zucker}, {de Silva}, {Freeman},
+  {Bland-Hawthorn}, \& {Hermes Team}}]{zucker12}
+{Zucker}, D.~B., {de Silva}, G., {Freeman}, K., {Bland-Hawthorn}, J., \&
+  {Hermes Team}. 2012, in {Astronomical Society of the Pacific Conference
+  Series}, Vol. 458, {Galactic Archaeology: Near-Field Cosmology and the
+  Formation of the Milky Way}, ed. W.~{Aoki}, M.~{Ishigaki}, T.~{Suda},
+  T.~{Tsujimoto}, \& N.~{Arimoto}, 421
+
+\end{thebibliography}
 
